---
title: "02_modelFitting_globalForestModel"
author: "Alice Stears"
date: "`r lubridate::today()`"
output:
  html_document:
    toc: true
    df_print: paged
    code_folding: hide
    dev: "jpeg"
params:
  run: FALSE      
  save_figs: TRUE
  ecoregion: CONUS
  response: TotalTreeCover
  treeThreshold: 10
  whichSecondBestMod: "halfse"# 1se OR halfse OR auto (chose the model w/ fewest # of predictors)
  thresholdMethod: "PredPrev=Obs" # approach for identifying best threshold
subtitle: "Based on code from Martin Holdrege's Sagebrush-Fire paper"
---
This workflow fits a model across all of CONUS that predicts whether a location does or does not have trees. 

The data consists of vegetation % cover by functional group from across CONUS (from FIA and RAP), 
as well as climate variables from DayMet, which have been aggregated into mean interannual conditions accross multiple temporal windows. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE,
                      message = FALSE, dpi = 72)
```

# Dependencies 

Set user defined parameters

```{r}
print(params)
# set to true if want to run for a limited number of rows (i.e. for code testing)
test_run <- params$test_run
save_figs <- params$save_figs
response <- params$response
fit_sample <- TRUE # fit model to a sample of the data
n_train <- 5e4 # sample size of the training data
n_test <- 1e6 # sample size of the testing data (if this is too big the decile dotplot code throws memory errors)
run <- params$run
whichSecondBestMod <- params$whichSecondBestMod
treeThreshold <- params$treeThreshold
thresholdMethod <- params$thresholdMethod
```

Load packages and source functions
```{r warning=FALSE, message=FALSE}
# set option so resampled dataset created here reproduces earlier runs of this code with dplyr 1.0.10
source("../../../Functions/glmTransformsIterates.R")
source("../../../Functions/transformPreds.R")
source("../../../Functions/betaLASSO.R")

#source("../../../Functions/StepBeta_mine.R")
#source("src/fig_params.R")
#source("src/modeling_functions.R")
 
library(betareg)
library(ggspatial)
library(terra)
library(tidyterra)
library(sf)
library(caret)
library(tidyverse)
library(GGally) # for ggpairs()
library(pdp) # for partial dependence plots
library(gridExtra)
library(knitr)
library(patchwork) # for figure insets etc. 
library(ggtext)
library(StepBeta)
theme_set(theme_classic())
library(here)
library(rsample)
library(kableExtra)
library(glmnet)
library(USAboundaries)
library(cvms)
library(rsvg)
library(ggimage)
library(doMC)
library(parallel)
library(pROC)

## functions
           get_inset <- function(df, densPlotName) {
              p <- ggplot(data = df) +
             #facet_wrap(~lyr, ncol = 1, scales = "free_y") +
                geom_density(aes(densPlotName)) +
                theme_bw(base_size=9) +  ## makes everything smaller
                theme(panel.background = element_rect(fill="transparent"),  ## white plot background
                      panel.border = element_blank(),
          axis.title.y = element_blank(),
          axis.title.x = element_blank(),
          axis.text.x = element_text(size=rel(2)), ## tiny axis text
          axis.text.y = element_text(size=rel(2)), ## tiny axis text
          axis.line.x = element_line(size = .5),
          axis.line.y = element_line(size = .5),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          plot.background = element_blank(),
          strip.background = element_blank(),
          strip.text = element_blank()) 
  return(p)
           }
           # custom function to add inset to main plot
           annotation_custom2 <- function (grob, xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, data)
{
  layer(data = data, stat = StatIdentity, position = PositionIdentity,
        geom = ggplot2:::GeomCustomAnn,
        inherit.aes = TRUE, params = list(grob = grob,
                                          xmin = xmin, xmax = xmax,
                                          ymin = ymin, ymax = ymax))
           }
```

Source the script that calculates the proportion of CONUS by area that is un-forested, which we will use to calibrate our model
```{r message=FALSE}
source("../DataPrep/09_non-forest_area.R")
# object w/ the raster of the forest locations is called notForest
# object w/ the proportion of the unmasked raster area that is not forested (<1% trees in a grid cell) is called area_not_forest_km2
# object w/ the mean proportion non-forest (area-weighted) is called mean_frac
```

# Read in data

Data compiled in the `prepDataForModels.R` script

```{r}
here::i_am("Analysis/VegComposition/ModelFitting/02_ModelFitting_globalForestModel.Rmd")
modDat <- readRDS( here("Data_processed", "CoverData", "DataForModels_spatiallyAveraged_withSoils_noSf_sampledLANDFIRE.rds")) %>% st_drop_geometry()
```

# Prep data

We will fit a binomial model that predicts whether or not there are trees at a location. Because the tree cover data we have is continuous between 0 and 100, we convert it to be binomial be forcing any values ≤ `r print(treeThreshold)` % to be 0, and any values > `r print(treeThreshold)`% to be 1. 

```{r}
modDat <- modDat %>% 
  mutate(TotalTreeCover_binom = replace(TotalTreeCover, TotalTreeCover <=treeThreshold, 0)) %>% 
  mutate(TotalTreeCover_binom = replace(TotalTreeCover_binom, TotalTreeCover_binom > treeThreshold, 1))
```

```{r rename variables}  
set.seed(1234)
# now, rename columns for brevity
modDat_1 <- modDat %>% 
  dplyr::select(-c(prcp_annTotal:annVPD_min)) %>% 
  # mutate(Lon = st_coordinates(.)[,1], 
  #        Lat = st_coordinates(.)[,2])  %>% 
  # st_drop_geometry() %>% 
  # filter(!is.na(newRegion))
  rename("tmin" = tmin_meanAnnAvg_CLIM, 
     "tmax" = tmax_meanAnnAvg_CLIM, #1
     "tmean" = tmean_meanAnnAvg_CLIM, 
     "prcp" = prcp_meanAnnTotal_CLIM, 
     "t_warm" = T_warmestMonth_meanAnnAvg_CLIM,
     "t_cold" = T_coldestMonth_meanAnnAvg_CLIM, 
     "prcp_wet" = precip_wettestMonth_meanAnnAvg_CLIM,
     "prcp_dry" = precip_driestMonth_meanAnnAvg_CLIM, 
     "prcp_seasonality" = precip_Seasonality_meanAnnAvg_CLIM, #2
     "prcpTempCorr" = PrecipTempCorr_meanAnnAvg_CLIM,  #3
     "abvFreezingMonth" = aboveFreezing_month_meanAnnAvg_CLIM, 
     "isothermality" = isothermality_meanAnnAvg_CLIM, #4
     "annWatDef" = annWaterDeficit_meanAnnAvg_CLIM, 
     "annWetDegDays" = annWetDegDays_meanAnnAvg_CLIM,
     "VPD_mean" = annVPD_mean_meanAnnAvg_CLIM, 
     "VPD_max" = annVPD_max_meanAnnAvg_CLIM, #5
     "VPD_min" = annVPD_min_meanAnnAvg_CLIM, #6
     "VPD_max_95" = annVPD_max_95percentile_CLIM, 
     "annWatDef_95" = annWaterDeficit_95percentile_CLIM, 
     "annWetDegDays_5" = annWetDegDays_5percentile_CLIM, 
     "frostFreeDays_5" = durationFrostFreeDays_5percentile_CLIM, 
     "frostFreeDays" = durationFrostFreeDays_meanAnnAvg_CLIM, 
     "soilDepth" = soilDepth, #7
     "clay" = surfaceClay_perc, 
     "sand" = avgSandPerc_acrossDepth, #8
     "coarse" = avgCoarsePerc_acrossDepth, #9
     "carbon" = avgOrganicCarbonPerc_0_3cm, #10
     "AWHC" = totalAvailableWaterHoldingCapacity,
     ## anomaly variables
     tmean_anom = tmean_meanAnnAvg_3yrAnom, #15
     tmin_anom = tmin_meanAnnAvg_3yrAnom, #16
     tmax_anom = tmax_meanAnnAvg_3yrAnom, #17
    prcp_anom = prcp_meanAnnTotal_3yrAnom, #18
      t_warm_anom = T_warmestMonth_meanAnnAvg_3yrAnom,  #19
     t_cold_anom = T_coldestMonth_meanAnnAvg_3yrAnom, #20
      prcp_wet_anom = precip_wettestMonth_meanAnnAvg_3yrAnom, #21
      precp_dry_anom = precip_driestMonth_meanAnnAvg_3yrAnom,  #22
    prcp_seasonality_anom = precip_Seasonality_meanAnnAvg_3yrAnom, #23 
     prcpTempCorr_anom = PrecipTempCorr_meanAnnAvg_3yrAnom, #24
      aboveFreezingMonth_anom = aboveFreezing_month_meanAnnAvg_3yrAnom, #25  
    isothermality_anom = isothermality_meanAnnAvg_3yrAnom, #26
       annWatDef_anom = annWaterDeficit_meanAnnAvg_3yrAnom, #27
     annWetDegDays_anom = annWetDegDays_meanAnnAvg_3yrAnom,  #28
      VPD_mean_anom = annVPD_mean_meanAnnAvg_3yrAnom, #29
      VPD_min_anom = annVPD_min_meanAnnAvg_3yrAnom,  #30
      VPD_max_anom = annVPD_max_meanAnnAvg_3yrAnom,  #31
     VPD_max_95_anom = annVPD_max_95percentile_3yrAnom, #32
      annWatDef_95_anom = annWaterDeficit_95percentile_3yrAnom, #33 
      annWetDegDays_5_anom = annWetDegDays_5percentile_3yrAnom ,  #34
    frostFreeDays_5_anom = durationFrostFreeDays_5percentile_3yrAnom, #35 
      frostFreeDays_anom = durationFrostFreeDays_meanAnnAvg_3yrAnom #36
  ) %>% 
  dplyr::select(-c(tmin_meanAnnAvg_3yr:durationFrostFreeDays_meanAnnAvg_3yr))

# get necessary raster data for background data on maps
test_rastTemp <-  rast("/Users/astears/Documents/Dropbox_static/Work/NAU_USGS_postdoc/cleanPED/PED_vegClimModels/Data_raw/dayMet/rawMonthlyData/orders/70e0da02b9d2d6e8faa8c97d211f3546/Daymet_Monthly_V4R1/data/daymet_v4_prcp_monttl_na_1980.tif") %>%
  terra::crop(ext(-2000000, 2600000, -1900000, 1500000))


## want to use x_vegDat_sf and y_vegDat_sf, so rename to x and y
modDat_1 <- modDat_1 %>% 
  rename(x = x_vegDat_sf, 
         y = y_vegDat_sf)
```


#### Predictors
The following are the candidate predictor variables for this ecoregion: 
```{r get candidate predictor variables}
  prednames <- c(
      "tmean","prcp","prcpTempCorr","isothermality", "annWetDegDays", "clay","coarse", "AWHC"
 
 #"tmean", "prcp", "prcp_seasonality", "prcpTempCorr", "isothermality", "annWetDegDays", "sand", "coarse", "AWHC"
  )

print(prednames)
```

#### Scale the predictor variables for the model-fitting process
```{r scaling predictors}
allPreds <- modDat_1 %>% 
  dplyr::select(tmin:frostFreeDays,tmean_anom:frostFreeDays_anom, soilDepth:AWHC) %>% 
  names()
modDat_1_s_temp <- modDat_1 %>% 
  mutate(across(all_of(allPreds), base::scale, .names = "{.col}_s")) 
saveRDS(modDat_1_s_temp, file = "./models/scaledModelInputData_yesNoTreeMod.rds")

# Remove the rows that have no observations for total tree cover
modDat_1_s <- modDat_1_s_temp[!is.na(modDat_1_s_temp[,"TotalTreeCover_binom"]),]

# subset the data to only include these predictors, and remove any remaining NAs 
modDat_1_s <- modDat_1_s %>% 
  dplyr::select(all_of(prednames), paste0(prednames, "_s"), TotalTreeCover, TotalTreeCover_binom, newRegion, Year, x, y, NA_L1NAME, NA_L2NAME) %>% 
  drop_na()

names(prednames) <- prednames
df_pred <- modDat_1_s[, prednames]

response <- "TotalTreeCover"
```


#### Visualize the response variable 
```{r visualize response variable , fig.height=4, fig.width=5}
ggplot(modDat_1_s) + 
  geom_histogram(aes(TotalTreeCover/100), fill = "forestgreen", col = "forestgreen", alpha = .5) + 
  xlab("Tree Cover") + 
  ggtitle("Untransformed, observed tree cover")

ggplot(modDat_1_s) + 
  geom_histogram(aes(TotalTreeCover_binom), fill = "purple", alpha = .5, col = "purple") +
  xlab("Tree Cover") + 
  ggtitle(paste0("Tree cover, converted to binomial with a ",treeThreshold,"cutoff"))

```

```{r summary_table}
create_summary <- function(df) {
  df %>% 
    pivot_longer(cols = everything(),
                 names_to = 'variable') %>% 
    group_by(variable) %>% 
    summarise(across(value, .fns = list(mean = ~mean(.x, na.rm = TRUE), min = ~min(.x, na.rm = TRUE), 
                                        median = ~median(.x, na.rm = TRUE), max = ~max(.x, na.rm = TRUE)))) %>% 
    mutate(across(where(is.numeric), round, 4))
}

modDat_1_s[prednames] %>% 
  create_summary() %>% 
  knitr::kable(caption = 'summaries of possible predictor variables') %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed")) 

```

#### Histograms of raw and scaled predictors
```{r}
  
scaleFigDat_1 <- modDat_1_s %>% 
  dplyr::select(c(x, y, Year, prednames)) %>% 
  pivot_longer(cols = all_of(names(prednames)), 
               names_to = "predNames", 
               values_to = "predValues_unScaled")
scaleFigDat_2 <- modDat_1_s %>% 
  dplyr::select(c(x, y, Year, paste0(prednames, "_s"))) %>% 
  pivot_longer(cols = all_of(paste0(prednames,"_s"
                                    )), 
               names_to = "predNames", 
               values_to = "predValues_scaled", 
               names_sep = ) %>% 
  mutate(predNames = str_replace(predNames, pattern = "_s$", replacement = ""))

scaleFigDat_3 <- scaleFigDat_1 %>% 
  left_join(scaleFigDat_2)

ggplot(scaleFigDat_3) + 
  facet_wrap(~predNames, scales = "free") +
  geom_histogram(aes(predValues_unScaled), fill = "lightgrey", col = "darkgrey") + 
  geom_histogram(aes(predValues_scaled), fill = "lightblue", col = "blue") +
  xlab ("predictor variable values") + 
  ggtitle("Comparing the distribution of unscaled (grey) to scaled (blue) predictor variables")
```

```{r}
modDat_1_s <- modDat_1_s %>% 
  rename_with(~paste0(.x, "_raw"), 
                any_of(names(prednames))) %>% 
  rename_with(~str_remove(.x, "_s$"), 
              any_of(paste0(names(prednames), "_s")))
  
```

#### Predictor variables compared to binned response variables

```{r climVar_boxplots, fig.height=9, fig.width=8, eval=FALSE}
set.seed(12011993)
# vector of name of response variables
vars_response <- response
# longformat dataframes for making boxplots
df_sample_plots <-  modDat_1_s  %>% 
  slice_sample(n = 5e4) %>% 
   rename(response = all_of("TotalTreeCover_binom")) %>% 
  mutate(response = case_when(
    response == 0 ~ "0", 
    response > 0  ~ "1", 
  )) %>% 
  dplyr::select(c(response, prednames)) %>% 
  tidyr::pivot_longer(cols = unname(prednames), 
               names_to = "predictor", 
               values_to = "value"
               )  
 

  ggplot(df_sample_plots, aes_string(x= "response", y = 'value')) +
  geom_boxplot() +
  facet_wrap(~predictor , scales = 'free_y') + 
  ylab("Predictor Variable Values") + 
    xlab(response)

```

#### If desired, split the data into training and test data (just leave one year out for now--somewhat randomly chose 2019, 2003, and 2012 )
```{r}
# # # do a pca of climate across years 
# pca_yrs <- prcomp(modDat_1_s[,paste0(prednames)])
# # library(factoextra)
# (fviz_pca_ind(pca_yrs, habillage = modDat_1_s$Year, label = "none", addEllipses = TRUE, ellipse.level = .95, ggtheme = theme_minimal(), alpha.ind = .1))
# 
# ggplot(modDat_1_s) + 
#   geom_point(aes(modDat_1_s$Year, modDat_1_s$TotalTreeCover)) + 
#   geom_smooth(aes(modDat_1_s$Year, modDat_1_s$TotalTreeCover))

# save test set
modDat_1_sTest <- modDat_1_s[modDat_1_s$Year %in% c(2003, 2012, 2019),]
modDat_1_s <- modDat_1_s[!modDat_1_s$Year %in% c(2003, 2012, 2019),]
```


# Model Fitting

## Visualize the spatial blocks and how they differ across environmental space
First, if there are observations in Louisiana, sub-sample them so they're not so over-represented in the dataset
```{r}
## make data into spatial format
modDat_1_sf <- modDat_1_s %>% 
  st_as_sf(coords = c("x", "y"), crs = st_crs("EPSG:4326"))


# get state boundaries for plotting
cropped_states <- us_states(resolution = "low")
cropped_states <- suppressMessages(cropped_states %>%
  dplyr::filter(name!="Hawaii") %>%
  dplyr::filter(name!="Alaska") %>%
  dplyr::filter(name!="Puerto Rico") %>%
  dplyr::filter(name!="American Samoa") %>%
  dplyr::filter(name!="Guam") %>%
  dplyr::filter(name!="Commonwealth of the Northern Mariana Islands") %>%
  dplyr::filter(name!="United States Virgin Islands")  %>%
  sf::st_transform(sf::st_crs(modDat_1_sf)))

```

```{r Visualize ecoregion distribution, fig.height=10, fig.width=14, message=FALSE, warning=FALSE}
## do a pca of climate across level 2 ecoregions
pca <- prcomp(modDat_1_s[,paste0(prednames)])
library(factoextra)
(fviz_pca_ind(pca, habillage = modDat_1_s$NA_L2NAME, label = "none", addEllipses = TRUE, ellipse.level = .95, ggtheme = theme_minimal(), alpha.ind = .1))

# make a table of n for each region
modDat_1_s %>% 
  group_by(NA_L2NAME) %>% 
  dplyr::summarize("Number_Of_Observations" = length(NA_L2NAME)) %>% 
  rename("Level_2_Ecoregion" = NA_L2NAME)%>% 
  kable() %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed")) 

```
## Then, look at the spatial distribution and environmental characteristics of the grouped ecoregions 
```{r fig.width = 14, fig.height=10, message=FALSE}

map1 <- ggplot() +
  geom_sf(data=cropped_states,fill='white') +
  geom_sf(data=modDat_1_sf#[modDat_1_sf$NA_L2NAME %in% c("MIXED WOOD PLAINS"),]
          ,
          aes(fill=as.factor(NA_L2NAME)),linewidth=0.5,alpha=0.5) +
  geom_point(data=modDat_1_s#[modDat_1_s$NA_L2NAME %in% c("MIXED WOOD PLAINS"),]
             ,
             alpha=0.5, 
             aes(x = x, y = y, color=as.factor(NA_L2NAME)), alpha = .1) +
  #scale_fill_okabeito() +
  #scale_color_okabeito() +
 # theme_default() +
  theme(legend.position = 'none') +
  labs(title = "Level 2 Ecoregions as spatial blocks")

hull <- modDat_1_sf %>%
  ungroup() %>%
  group_by(NA_L2NAME) %>%
  slice(chull(tmean, prcp))

plot1<-ggplot(data=modDat_1_sf,aes(x=tmean,y=prcp)) +
  geom_polygon(data = hull, alpha = 0.25,aes(fill=NA_L2NAME) )+
  geom_point(aes(group=NA_L2NAME,color=NA_L2NAME),alpha=0.25) +
  theme_minimal() + xlab("Annual Average T_mean - long-term average") +
  ylab("Annual Average Precip - long-term average") #+
  #scale_color_okabeito() +
  #scale_fill_okabeito()

plot2<-ggplot(data=modDat_1_sf %>%
                pivot_longer(cols=tmean:prcp),
              aes(x=value,group=name)) +
  # geom_polygon(data = hull, alpha = 0.25,aes(fill=fold) )+
  geom_density(aes(group=NA_L2NAME,fill=NA_L2NAME),alpha=0.25) +
  theme_minimal() +
  facet_wrap(~name,scales='free')
 
library(patchwork)
(combo <- (map1+plot1)/plot2) 

```


## Fit a global model with all of the data 

### First, fit a LASSO regression model using the glmnet R package
- This regression is a beta glm with a logit link (using custom function from Daniel)
- Use cross validation across level 2 ecoregions to tune the lambda parameter in the LASSO model 
- this model is fit to using the scaled weather/climate/soils variables
- this list of possible predictors includes: 
  1. main effects
  2. interactions between all soils variables
  3. interactions between climate and weather variables
  4. transformed main effects (squared, log-transformed (add a uniform integer -- 20-- to all variables prior to log-transformation), square root-transformed (add a uniform integer -- 20-- to all variables prior to log-transformation))
```{r fit global model, echo = FALSE}
# penalty values
  lambdas <- 10^seq(1,-5, length.out = 200)  # sequence of penalties to test

 ## if the model parameters indicate that we DO want anomalies... 
    # make model matrix (which includes transformations and interactions)
  ## only include interactions within and between climate and anomaly variables, and within and between soils variables

 ## break predictors into climate, weather, and soils 
  prednames_clim <- prednames[which(prednames %in% c("tmin" , "tmax",  "tmean",
     "prcp", "t_warm" , "t_cold" , "prcp_wet" , "prcp_dry" , "prcp_seasonality", "prcpTempCorr", "abvFreezingMonth" , "isothermality" , "annWatDef" , "annWetDegDays",  "VPD_mean",  "VPD_max",  "VPD_min" , "VPD_max_95", "annWatDef_95", "annWetDegDays_5",  "frostFreeDays_5", "frostFreeDays"))]

  prednames_weath <- prednames[which(prednames %in% c("tmean_anom", "tmin_anom",  "tmax_anom", 
    "prcp_anom",      "t_warm_anom" , "t_cold_anom",  "prcp_wet_anom" , "precp_dry_anom" ,  "prcp_seasonality_anom" ,  "prcpTempCorr_anom" , "aboveFreezingMonth_anom",  "isothermality_anom" ,  "annWatDef_anom", "annWetDegDays_anom",  "VPD_mean_anom" , "VPD_min_anom" , "VPD_max_anom", "VPD_max_95_anom", "annWatDef_95_anom" , "annWetDegDays_5_anom",  "frostFreeDays_5_anom" , "frostFreeDays_anom"))]
  
  prednames_soils <- prednames[which(prednames %in% c( "soilDepth" , "clay", 
     "sand" , "coarse" ,  "carbon", "AWHC"))]
  
  
  # get a text string of possible interactions between climate and weather variables
  combos <- data.frame(gtools::permutations(n = length(c(prednames_clim, prednames_weath)),
                                            r = 2,
                                            v = c(prednames_clim, prednames_weath)))
  
  combos$interactions <- paste0(combos$X1, "*", combos$X2)
  
  
  # get a text string of possible interactions between soils variables
  combos_2 <- data.frame(gtools::permutations(n = length(c(prednames_soils)),
                                            r = 2,
                                            v = c(prednames_soils)))
  
  combos_2$interactions <- paste0(combos_2$X1, "*", combos_2$X2)
  
  # make text string of all possible interactions
  int_string <-str_flatten(c(combos$interactions, combos_2$interactions), collapse = " + ")
  
  # make text string of transformations (log, squared)
  transform_string <- str_flatten(c(# add square root term 
                                    paste0("I(log(",prednames_clim,"+30))"), 
                                    paste0("I(log(",prednames_soils,"+30))"),
                                    # 
                                    paste0("I(sqrt(",prednames_clim,"+30))"),
                                    paste0("I(sqrt(",prednames_soils,"+30))"),

                                    paste0("I(",prednames_clim,"^2)"),
                                    paste0("I(",prednames_soils,"^2)")),
                                    collapse = " + ")
              
  # get a text string of the possible untransformed model terms
  mod_string <- paste("TotalTreeCover_binom", "~", paste0(prednames, collapse = "+"))
  
  # make a text string of the model formula w/ interactions and transformations included
  modWithInteractions <- paste(mod_string, "+",int_string , "+", transform_string
                               )

  # now, make a model formula object
  f <- as.formula(modWithInteractions)
  
  # transform dataframe to model matrix
  X <- model.matrix(f, modDat_1_s)
  # get response variable
  y <- as.matrix(modDat_1_s[,"TotalTreeCover_binom"])

```

#### Get rid of transformed predictions and interactions that are correlated 
```{r}
# get first pass of names correlated variables
X_df <- X %>% 
  as.data.frame() %>% 
  dplyr::select(-'(Intercept)')  
corrNames_i <- X_df %>% 
  cor()  %>% 
   caret::findCorrelation(cutoff = .7, verbose = FALSE, names = TRUE, exact = TRUE)
# remove those names that are untransformed main effects 
  # vector of columns to remove 
badNames <- corrNames_i[!(corrNames_i %in% prednames)]
while (sum(!(corrNames_i %in% prednames))>0) {
 corrNames_i <-  X_df %>% 
    dplyr::select(-badNames) %>% 
     cor()  %>% 
   caret::findCorrelation(cutoff = .7, verbose = FALSE, names = TRUE, exact = TRUE)
 # update the vector of names to remove 
 badNames <- unique(c(badNames, corrNames_i[!(corrNames_i %in% prednames)]))
}

## see if there are any correlated variables left (would be all main effects at this point)
# if there are, step through and remove the variable that each is most correlated with 
if (length(corrNames_i)>1) {
  for (i in 1:length(corrNames_i)) {
    X_i <- X_df %>% 
      dplyr::select(-badNames)
    if (corrNames_i[i] %in% names(X_i)) {
    corMat_i <- cor(x = X_i[corrNames_i[i]], y = X_i %>% dplyr::select(-corrNames_i[i])) 
    badNames_i <- colnames(corMat_i)[abs(corMat_i)>=.7]
    # if there are any predictors in the 'badNames_i', remove them from this list
    if (length(badNames_i) > 0 & sum(c(badNames_i %in% prednames))>0) {
        badNames_i <- badNames_i[!(badNames_i %in% prednames)]
    }
    badNames <- unique(c(badNames, badNames_i))
    }
  }
}
## update the X matrix to exclude these correlated variables
X <- X[,!(colnames(X) %in% badNames)]
```

#### Run the global LASSO model
```{r}
if (run == TRUE) {
  # set up custom folds
    # get the ecoregions for training lambda
  train_eco <- modDat_1_s$NA_L2NAME#[train]
  
  # Fit model -----------------------------------------------
  # specify leave-one-year-out cross-validation
  my_folds <- as.numeric(as.factor(train_eco))

    # set up parallel processing
    # this computer has 16 cores (parallel::detectCores())
    registerDoMC(cores = 8)
    
    fit <- cv.glmnet(
      x = X[,2:ncol(X)], 
      y = y, 
      family = "binomial",
      keep = FALSE,
      alpha = 1,  # 0 == ridge regression, 1 == lasso, 0.5 ~~ elastic net
      lambda = lambdas,
      relax = ifelse(response == "ShrubCover", yes = TRUE, no = FALSE),
      #nlambda = 100,
      type.measure="mse",
      #penalty.factor = pen_facts,
      foldid = my_folds,
      #thresh = thresh,
      standardize = FALSE, ## scales variables prior to the model sequence... coefficients are always returned on the original scale
      parallel = TRUE
    )
    base::saveRDS(fit, paste0("../ModelFitting/models/ForestMod_globalLASSOmod_binomial_treeCutoff_",treeThreshold,".rds"))
    
  
     best_lambda <- fit$lambda.min
  # save the lambda for the most regularized model w/ an MSE that is still 1SE w/in the best lambda model
  lambda_1SE <- fit$lambda.1se
  # save the lambda for the most regularized model w/ an MSE that is still .5SE w/in the best lambda model
  lambda_halfSE <- best_lambda + ((lambda_1SE - best_lambda)/2)
 
## Now, we need to do stability selection to ensure the coefficients that are being chosen with each lambda are stable 

## stability selection for best lambda model 
# setup params
p <- ncol(X[,2:ncol(X)]) # of parameters
n <- length(y) # of observations
n_iter <- 100        # number of subsamples
sample_frac <- 0.75  # fraction of data to subsample
lambda_val <- best_lambda    # fixed lambda value (could be chosen via CV)

# Track selection
selection_counts_bestL <- matrix(0, nrow = p, ncol = 1)

for (i in 1:n_iter) {
  # Subsample rows
  sample_idx <- sample(1:n, size = floor(sample_frac * n), replace = FALSE)
  X_sub <- X[sample_idx, ]
  y_sub <- y[sample_idx]

  # Fit Lasso model
  fit_stab_i <- glmnet(x = X_sub[,2:ncol(X_sub)], y = y_sub, 
    family = "binomial",
    alpha = 1, lambda = lambda_val, standardize = FALSE)

  # Get non-zero coefficients (excluding intercept)
  select_bestL <- which(as.vector(coef(fit_stab_i))[-1] != 0)
  selection_counts_bestL[select_bestL] <- selection_counts_bestL[select_bestL] + 1
}

# Convert counts to selection probabilities (the probability that a variable is selected over 100 iterations)
selection_prob_bestL <- selection_counts_bestL / n_iter
selection_prob_bestL_df <- data.frame(
  VariableNumber = paste0("X", 1:p),
  VariableName = rownames(coef(fit_stab_i))[2:(p+1)],
  SelectionProb = as.vector(selection_prob_bestL)
)

# get those variables that are selected in ≥70% of subsets (Meinshausen and Bühlmann, 2010)
bestLambda_coef <- selection_prob_bestL_df[selection_prob_bestL_df$SelectionProb>=.7, c("VariableName", "SelectionProb")]

#//////
# stability selection for 1se lambda model
lambda_val <-  lambda_1SE    # fixed lambda value (could be chosen via CV)

# Track selection
selection_counts_1seL <- matrix(0, nrow = p, ncol = 1)

for (i in 1:n_iter) {
  # Subsample rows
  sample_idx <- sample(1:n, size = floor(sample_frac * n), replace = FALSE)
  X_sub <- X[sample_idx, ]
  y_sub <- y[sample_idx]

  # Fit Lasso model
  fit_stab_i <- glmnet(x = X_sub[,2:ncol(X_sub)], y = y_sub, 
    family = "binomial",
    alpha = 1, lambda = lambda_val, standardize = FALSE)

  # Get non-zero coefficients (excluding intercept)
  selected_1seL <- which(as.vector(coef(fit_stab_i))[-1] != 0)
  selection_counts_1seL[selected_1seL] <- selection_counts_1seL[selected_1seL] + 1
}

# Convert counts to selection probabilities (the probability that a variable is selected over 100 iterations)
selection_prob_1seL <- selection_counts_1seL / n_iter
selection_prob_1seL_df <- data.frame(
  VariableNumber = paste0("X", 1:p),
  VariableName = rownames(coef(fit_stab_i))[2:(p+1)],
  SelectionProb = as.vector(selection_prob_1seL)
)

# get those variables that are selected in ≥70% of subsets (Meinshausen and Bühlmann, 2010)
seLambda_coef <- selection_prob_1seL_df[selection_prob_1seL_df$SelectionProb>=.7, c("VariableName", "SelectionProb")]

# stability selection for half se lambda model
lambda_val <- lambda_halfSE    # fixed lambda value (could be chosen via CV)

# Track selection
selection_counts_halfseL <- matrix(0, nrow = p, ncol = 1)

for (i in 1:n_iter) {
  # Subsample rows
  sample_idx <- sample(1:n, size = floor(sample_frac * n), replace = FALSE)
  X_sub <- X[sample_idx, ]
  y_sub <- y[sample_idx]

  # Fit Lasso model
  fit_stab_i <- glmnet(x = X_sub[,2:ncol(X_sub)], y = y_sub, 
    family = "binomial",
    alpha = 1, lambda = lambda_val, standardize = FALSE)

  # Get non-zero coefficients (excluding intercept)
  selected_halfseL <- which(as.vector(coef(fit_stab_i))[-1] != 0)
  selection_counts_halfseL[selected_halfseL] <- selection_counts_halfseL[selected_halfseL] + 1
}

# Convert counts to selection probabilities (the probability that a variable is selected_halfseL over 100 iterations)
selection_prob_halfseL <- selection_counts_halfseL / n_iter
selection_prob_halfseL_df <- data.frame(
  VariableNumber = paste0("X", 1:p),
  VariableName = rownames(coef(fit_stab_i))[2:(p+1)],
  SelectionProb = as.vector(selection_prob_halfseL)
)

# get those variables that are selected_halfseL_halfseL in ≥70% of subsets (Meinshausen and Bühlmann, 2010)
halfseLambda_coef <- selection_prob_halfseL_df[selection_prob_halfseL_df$SelectionProb>=.7, c("VariableName", "SelectionProb")]

## fit w/ the identified coefficients from the 'best' lambda, but using the glm function
  mat_glmnet_best <- bestLambda_coef$VariableName 

  if (length(mat_glmnet_best) == 0) {
    f_glm_bestLambda <- as.formula(paste0("TotalTreeCover_binom", "~ 1"))
  } else {
  f_glm_bestLambda <- as.formula(paste0("TotalTreeCover_binom", " ~ ", paste0(mat_glmnet_best, collapse = " + ")))
  }
  
## fit using glm
  fit_glm_bestLambda <- fit_glm_bestLambda_binomial <- glm(formula = f_glm_bestLambda, data = modDat_1_s, family = binomial)
    
   ## fit w/ the identified coefficients from the '1se' lambda, but using the glm function
  mat_glmnet_1se <- seLambda_coef$VariableName
    
  if (length(mat_glmnet_1se) == 0) {
    f_glm_1se <- as.formula(paste0("TotalTreeCover_binom", "~ 1"))
  } else {
  f_glm_1se <- as.formula(paste0("TotalTreeCover_binom", " ~ ", paste0(mat_glmnet_1se, collapse = " + ")))
  }


  fit_glm_se <- glm(formula = f_glm_1se, data = modDat_1_s, family = binomial)
  # glm(data = modDat_1_s, formula = f_glm_1se,
  #                   family =  stats::Gamma(link = "log"))
  
     ## fit w/ the identified coefficients from the '.5se' lambda, but using the glm function
  mat_glmnet_halfse <- halfseLambda_coef$VariableName
  
  if (length(mat_glmnet_halfse) == 0) {
    f_glm_halfse <- as.formula(paste0("TotalTreeCover_binom", "~ 1"))
  } else {
  f_glm_halfse <- as.formula(paste0("TotalTreeCover_binom", " ~ ", paste0(mat_glmnet_halfse, collapse = " + ")))
  }

  fit_glm_halfse <- glm(formula = f_glm_halfse, data = modDat_1_s, family = binomial )
  
  ## save models 
  saveRDS(fit_glm_bestLambda, paste0("./models/ForestMod_bestLambdaGLM_",treeThreshold,".rds"))
  saveRDS(fit_glm_halfse, paste0("./models/ForestMod_halfSELambdaGLM_",treeThreshold,".rds"))
  saveRDS(fit_glm_se, paste0("./models/ForestMod_oneSELambdaGLM_",treeThreshold,".rds"))
    
  ## save the R environment after running the models 
  save(f_glm_halfse, mat_glmnet_halfse, halfseLambda_coef,
              f_glm_1se, mat_glmnet_1se, seLambda_coef,
              f_glm_bestLambda, mat_glmnet_best, bestLambda_coef,
              file = paste0("./models/interimModelFittingObjects_ForestMod_binomial_",treeThreshold,".rds"))
  } else {
    # read in LASSO object
    fit <- readRDS(paste0("../ModelFitting/models/ForestMod_globalLASSOmod_binomial_treeCutoff_",treeThreshold,".rds"))
    
    # read in R objects having to do w/ model fitting 
    load(file = paste0("./models/interimModelFittingObjects_ForestMod_binomial_",treeThreshold,".rds"))
      
  fit_glm_bestLambda <- readRDS(paste0("./models/ForestMod_bestLambdaGLM_",treeThreshold,".rds"))
  fit_glm_halfse <- readRDS(paste0("./models/ForestMod_halfSELambdaGLM_",treeThreshold,".rds"))
  fit_glm_se <- readRDS(paste0("./models/ForestMod_oneSELambdaGLM_",treeThreshold,".rds"))
  }


  # assess model fit
  # assess.glmnet(fit$fit.preval, #newx = X[,2:293], 
  #               newy = y, family = stats::Gamma(link = "log"))
  # save the minimum lambda
  best_lambda <- fit$lambda.min
  # save the lambda for the most regularized model w/ an MSE that is still 1SE w/in the best lambda model
  lambda_1SE <- fit$lambda.1se
  # save the lambda for the most regularized model w/ an MSE that is still .5SE w/in the best lambda model
  lambda_halfSE <- best_lambda + ((lambda_1SE - best_lambda)/2)
 
  print(fit)     
  
  plot(fit)

```

#### Then, we predict (on the training set) using both of these models (best lambda and 1se lambda) as an itial test
```{r}
  ## predict on the test data
  # lasso model predictions with the optimal lambda
  optimal_pred <- stats::predict(fit_glm_bestLambda, #newx=X[,2:ncol(X)], 
                                 type = "response")
  optimal_pred_1se <-  stats::predict(fit_glm_se, newx=X[,2:ncol(X)], type = "response")
  optimal_pred_halfse <- stats::predict(fit_glm_halfse, newx = X[,2:ncol(X)], type = "response")
  
    null_fit <- glm(
      formula = y ~ 1, #data = modDat_1_s, 
      family = binomial
      )
  null_pred <- predict(null_fit, newdata = as.data.frame(X), type = "response"
                       )

  # save data
  fullModOut <- list(
    "modelObject" = fit,
    "nullModelObject" = null_fit,
    "modelPredictions" = data.frame(#ecoRegion_holdout = rep(test_eco,length(y)),
      obs=y,
                    pred_opt=optimal_pred, 
                    pred_opt_se = optimal_pred_1se,
                    pred_opt_halfse = optimal_pred_halfse,
                    pred_null=null_pred#,
                    #pred_nopenalty=nopen_pred
                    ))

ggplot() + 
  geom_point(aes(X[,2], fullModOut$modelPredictions$obs), col = "black", alpha = .1) + 
  geom_point(aes(X[,2], fullModOut$modelPredictions$pred_opt), col = "red", alpha = .1) + ## predictions w/ the CV model (best lambda)
  geom_point(aes(X[,2], fullModOut$modelPredictions$pred_opt_halfse), col = "orange", alpha = .1) + ## predictions w/ the CV model (.5se lambda)
  geom_point(aes(X[,2], fullModOut$modelPredictions$pred_opt_se), col = "green", alpha = .1) + ## predictions w/ the CV model (1se lambda)
  geom_point(aes(X[,2], fullModOut$modelPredictions$pred_null), col = "blue", alpha = .1) + 
  labs(title = "A rough comparison of observed and model-predicted values", 
       subtitle = "black = observed values \n red = predictions from 'best lambda' model \n orange = predictions for '1/2se' lambda model \n green = predictions from '1se' lambda model \n blue = predictions from null model") +
  xlab(colnames(X)[2])
  #ylim(c(0,200))
 
```

#### The internal cross-validation process to fit the global LASSO model identified an optimal lambda value (regularization parameter) of `r{print(best_lambda)}`. The lambda value such that the cross validation error is within 1 standard error of the minimum ("1se lambda") was `r{print(fit$lambda.1se)}``. The following coefficients were kept in each model: 
```{r coefficients in the global model}
# the coefficient matrix from the 'best model' -- find and print those coefficients that aren't 0 in a table
coef_glm_bestLambda <- coef(fit_glm_bestLambda) %>% 
  data.frame() 
coef_glm_bestLambda$coefficientName <- rownames(coef_glm_bestLambda)
names(coef_glm_bestLambda)[1] <- "coefficientValue_bestLambda"
# coefficient matrix from the '1se' model 
coef_glm_1se <- coef(fit_glm_se) %>% 
  data.frame() 
coef_glm_1se$coefficientName <- rownames(coef_glm_1se)
names(coef_glm_1se)[1] <- "coefficientValue_1seLambda"
# coefficient matrix from the 'half se' model 
coef_glm_halfse <- coef(fit_glm_halfse) %>% 
  data.frame() 
coef_glm_halfse$coefficientName <- rownames(coef_glm_halfse)
names(coef_glm_halfse)[1] <- "coefficientValue_halfseLambda"
# add together
coefs <- full_join(coef_glm_bestLambda, coef_glm_halfse) %>% 
  full_join(coef_glm_1se) %>% 
  dplyr::select(coefficientName, coefficientValue_bestLambda,
                coefficientValue_halfseLambda, coefficientValue_1seLambda)

globModTerms <- coefs[!is.na(coefs$coefficientValue_bestLambda), "coefficientName"]

## also, get the number of unique variables in each model 
var_prop_pred <- paste0(response, "_pred")
response_vars <- c(response, var_prop_pred)
# for best lambda model
prednames_fig <- paste(str_split(globModTerms, ":", simplify = TRUE)) 
prednames_fig <- str_replace(prednames_fig, "I\\(", "")
prednames_fig <- str_replace(prednames_fig, "\\^2\\)", "")
prednames_fig <- unique(prednames_fig[prednames_fig>0])
prednames_fig <- prednames_fig
prednames_fig_num <- length(prednames_fig)
# for 1SE lambda model
globModTerms_1se <- coefs[!is.na(coefs$coefficientValue_1seLambda), "coefficientName"]
if (length(globModTerms_1se) == 1) {
prednames_fig_1se <- paste(str_split(globModTerms_1se, ":", simplify = TRUE)) 
prednames_fig_1se <- str_replace(prednames_fig_1se, "I\\(", "")
prednames_fig_1se <- str_replace(prednames_fig_1se, "\\^2\\)", "")
prednames_fig_1se <- unique(prednames_fig_1se[prednames_fig_1se>0])
prednames_fig_1se_num <- c(0)
} else {
prednames_fig_1se <- paste(str_split(globModTerms_1se, ":", simplify = TRUE)) 
prednames_fig_1se <- str_replace(prednames_fig_1se, "I\\(", "")
prednames_fig_1se <- str_replace(prednames_fig_1se, "\\^2\\)", "")
prednames_fig_1se <- unique(prednames_fig_1se[prednames_fig_1se>0])
prednames_fig_1se_num <- length(prednames_fig_1se)
}
# for 1/2SE lambda model
globModTerms_halfse <- coefs[!is.na(coefs$coefficientValue_halfseLambda), "coefficientName"]
if (length(globModTerms_halfse) == 1) {
prednames_fig_halfse <- paste(str_split(globModTerms_halfse, ":", simplify = TRUE)) 
prednames_fig_halfse <- str_replace(prednames_fig_halfse, "I\\(", "")
prednames_fig_halfse <- str_replace(prednames_fig_halfse, "\\^2\\)", "")
prednames_fig_halfse <- unique(prednames_fig_halfse[prednames_fig_halfse>0])
prednames_fig_halfse_num <- c(0)
} else {
prednames_fig_halfse <- paste(str_split(globModTerms_halfse, ":", simplify = TRUE)) 
prednames_fig_halfse <- str_replace(prednames_fig_halfse, "I\\(", "")
prednames_fig_halfse <- str_replace(prednames_fig_halfse, "\\^2\\)", "")
prednames_fig_halfse <- unique(prednames_fig_halfse[prednames_fig_halfse>0])
prednames_fig_halfse_num <- length(prednames_fig_halfse)
}
# make a table
kable(coefs, col.names = c("Coefficient Name", "Value from best lambda model", 
                           "Value from 1/2 se lambda", "Value from 1se lambda model")
      ) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed")) 

# calculate RMSE of all models 
RMSE_best <- yardstick::rmse(fullModOut$modelPredictions[,c("obs", "pred_opt")], truth = "obs", estimate = "pred_opt")$.estimate
RMSE_halfse <- yardstick::rmse(fullModOut$modelPredictions[,c("obs", "pred_opt_halfse")], truth = "obs", estimate = "pred_opt_halfse")$.estimate
RMSE_1se <- yardstick::rmse(fullModOut$modelPredictions[,c("obs", "pred_opt_se")], truth = "obs", estimate = "pred_opt_se")$.estimate
# calculate bias of all models
bias_best <-  mean((fullModOut$modelPredictions$obs) - fullModOut$modelPredictions$pred_opt)
bias_halfse <-  mean((fullModOut$modelPredictions$obs) - fullModOut$modelPredictions$pred_opt_halfse)
bias_1se <- mean((fullModOut$modelPredictions$obs) - fullModOut$modelPredictions$pred_opt_se)

uniqueCoeffs <- data.frame("Best lambda model" = c(signif(RMSE_best,3), as.character(signif(bias_best, 3)),
  as.integer(length(globModTerms)-1), as.integer(prednames_fig_num), 
                                                   as.integer(sum(prednames_fig %in% c(prednames_clim))),
                                                   as.integer(sum(prednames_fig %in% c(prednames_weath))),
                                                   as.integer(sum(prednames_fig %in% c(prednames_soils)))
                                                   ), 
                           "1/2 se lambda model" = c(signif(RMSE_halfse,3), as.character(signif(bias_halfse, 3)),
                             length(globModTerms_halfse)-1, prednames_fig_halfse_num,
                                                   sum(prednames_fig_halfse %in% c(prednames_clim)),
                                                   sum(prednames_fig_halfse %in% c(prednames_weath)),
                                                   sum(prednames_fig_halfse %in% c(prednames_soils))), 
                           "1se lambda model" = c(signif(RMSE_1se,3), as.character(signif(bias_1se, 3)),
                             length(globModTerms_1se)-1, prednames_fig_1se_num,
                                                   sum(prednames_fig_1se %in% c(prednames_clim)),
                                                   sum(prednames_fig_1se %in% c(prednames_weath)),
                                                   sum(prednames_fig_1se %in% c(prednames_soils))))
row.names(uniqueCoeffs) <- c("RMSE", "bias: mean(obs-pred.)", "Total number of coefficients", "Number of unique coefficients",
                             "Number of unique climate coefficients", 
                             "Number of unique weather coefficients",  
                             "Number of unique soils coefficients"
                             )

kable(uniqueCoeffs, 
      col.names = c("Best lambda model", "1/2 se lambda model", "1se lambda model"), row.names = TRUE) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed")) 
```

# Calibration and Visualizations of Model Predictions and Residuals 

In the following figures, we show model predictions made using the best lambda model, as well as an alternative "second-best" lambda model. As the alternative to the best lambda model, we use the model (1se or 1/2se of best Lambda) that has the fewest number of unique predictors, but is not a null model. 
```{r}
if (whichSecondBestMod == "auto") {
  # name of model w/ fewest # of predictors (but more than 0)
uniqueCoeff_min <- min(as.numeric(uniqueCoeffs[4,2:3])[which(as.numeric(uniqueCoeffs[4,2:3]) > 0)])
alternativeModel <- names(uniqueCoeffs[4,2:3])[which(uniqueCoeffs[4,2:3] == uniqueCoeff_min)]

if (is.finite(uniqueCoeff_min)) {
  if (length(alternativeModel) == 1) {
  if (alternativeModel == "X1.2.se.lambda.model") {
  mod_secondBest <- fit_glm_halfse
  name_secondBestMod <- "1/2 SE Model"
  prednames_secondBestMod <- prednames_fig_halfse
  # get the name of the column that contains the prediction data for only the training data, for this model
trainDatPredName <- "pred_opt_halfse"
} else if (alternativeModel == "X1se.lambda.model") {
  mod_secondBest <- fit_glm_se
  name_secondBestMod <- "1 SE Model"
  prednames_secondBestMod <- prednames_fig_1se
  # get the name of the column that contains the prediction data for only the training data, for this model
trainDatPredName <- "pred_opt_se"
}
} else {
  # if both alternative models have the same number of unique coefficients, chose the model that has the fewest number of total predictors
  uniqueCoeff_min2 <- min(as.numeric(uniqueCoeffs[3,alternativeModel]))
alternativeModel2 <- names(uniqueCoeffs[3,alternativeModel])[which(uniqueCoeffs[3,alternativeModel] == uniqueCoeff_min2)]
if (length(alternativeModel2) == 1) {
  if (alternativeModel2 == "X1.2.se.lambda.model") {
  mod_secondBest <- fit_glm_halfse
  name_secondBestMod <- "1/2 SE Model"
  prednames_secondBestMod <- prednames_fig_halfse
   # get the name of the column that contains the prediction data for only the training data, for this model
trainDatPredName <- "pred_opt_halfse"
} else if (alternativeModel2 == "X1se.lambda.model") {
  mod_secondBest <- fit_glm_se
  name_secondBestMod <- "1 SE Model"
  prednames_secondBestMod <- prednames_fig_1se
  # get the name of the column that contains the prediction data for only the training data, for this model
trainDatPredName <- "pred_opt_se"
}
} else {
  mod_secondBest <- fit_glm_halfse
  name_secondBestMod <- "1/2 SE Model"
  prednames_secondBestMod <- prednames_fig_halfse
  # get the name of the column that contains the prediction data for only the training data, for this model
trainDatPredName <- "pred_opt_halfse"
}

}
  }else {
    mod_secondBest <- NULL
  name_secondBestMod <- "Intercept_Only"
  prednames_secondBestMod <- NULL
}
} else if (whichSecondBestMod == "1se") {
  mod_secondBest <- fit_glm_se
  name_secondBestMod <- "1 SE Model"
  prednames_secondBestMod <- prednames_fig_1se
# get the name of the column that contains the prediction data for only the training data, for this model
trainDatPredName <- "pred_opt_se"
} else if (whichSecondBestMod == "halfse") {
  mod_secondBest <- fit_glm_halfse
  name_secondBestMod <- "1/2 SE Model"
  prednames_secondBestMod <- prednames_fig_halfse
  # get the name of the column that contains the prediction data for only the training data, for this model
trainDatPredName <- "pred_opt_halfse"
}
```


```{r}
## Get wall-to-wall climate data calculated from dayMet, which we will use for calibration
# read in unscaled climate data
climDat_Temp <- readRDS("../../../Data_processed/WallToWallClimateData/DayMetData_allCONUS_2023ClimateValues_withSoils.rds")
#climDat_Temp <- readRDS("../../../Data_processed/EcoRegion_climSoilData.rds")

# rename
climDat <- climDat_Temp %>% 
  #filter(year == 2016) %>% 
  dplyr::select(year, tmin_meanAnnAvg_CLIM:geometry) %>%
rename("tmin" = tmin_meanAnnAvg_CLIM, 
     "tmax" = tmax_meanAnnAvg_CLIM, #1 
     "tmean" = tmean_meanAnnAvg_CLIM, 
     "prcp" = prcp_meanAnnTotal_CLIM, 
     "t_warm" = T_warmestMonth_meanAnnAvg_CLIM,
     "t_cold" = T_coldestMonth_meanAnnAvg_CLIM, 
     "prcp_wet" = precip_wettestMonth_meanAnnAvg_CLIM,
     "prcp_dry" = precip_driestMonth_meanAnnAvg_CLIM, 
     "prcp_seasonality" = precip_Seasonality_meanAnnAvg_CLIM, #2
     "prcpTempCorr" = PrecipTempCorr_meanAnnAvg_CLIM,  #3
     "abvFreezingMonth" = aboveFreezing_month_meanAnnAvg_CLIM, 
     "isothermality" = isothermality_meanAnnAvg_CLIM, #4
     "annWatDef" = annWaterDeficit_meanAnnAvg_CLIM, 
     "annWetDegDays" = annWetDegDays_meanAnnAvg_CLIM,
     "VPD_mean" = annVPD_mean_meanAnnAvg_CLIM, 
     "VPD_max" = annVPD_max_meanAnnAvg_CLIM, #5
     "VPD_min" = annVPD_min_meanAnnAvg_CLIM, #6
     "VPD_max_95" = annVPD_max_95percentile_CLIM, 
     "annWatDef_95" = annWaterDeficit_95percentile_CLIM, 
     "annWetDegDays_5" = annWetDegDays_5percentile_CLIM, 
     "frostFreeDays_5" = durationFrostFreeDays_5percentile_CLIM, 
     "frostFreeDays" = durationFrostFreeDays_meanAnnAvg_CLIM, 
     "soilDepth" = soilDepth, #7
     "clay" = surfaceClay_perc, 
     "sand" = avgSandPerc_acrossDepth, #8
     "coarse" = avgCoarsePerc_acrossDepth, #9
     "carbon" = avgOrganicCarbonPerc_0_3cm, #10
     "AWHC" = totalAvailableWaterHoldingCapacity,
  )

#. now, scale the model predictions according to the scale parameters used in the fitting data 
scaleParams <- modDat_1_s_temp %>% 
  #filter(Year == 2016) %>% 
  dplyr::select(tmin_s:AWHC_s) %>% 
  reframe(across(all_of(names(.)), attributes)) 

# apply the scaling factors to the contemporary climate data 
namesToScale <- climDat %>% 
  st_drop_geometry() %>% 
  dplyr::select(tmin:frostFreeDays, soilDepth:carbon, AWHC) %>% 
  names()
climDat_noSF <- climDat %>% 
  st_drop_geometry()
climDat_scaled <- map(namesToScale, .f = function(x) {
  x_new <- (climDat_noSF[,x] - scaleParams[,paste0(x, "_s")]$`scaled:center`)/scaleParams[,paste0(x, "_s")]$`scaled:scale`
  return(data.frame(x_new))
}) %>% 
  purrr::list_cbind()
names(climDat_scaled) <- paste0(namesToScale, "_s")

# get the Lon and Lat from the climDat geometry (crs is the same as dayMet)
climDatPred <- climDat %>% 
  mutate("Lon" = st_coordinates(.)[,1],
         "Lat" = st_coordinates(.)[,2]) %>% 
  #st_drop_geometry() %>% 
  dplyr::select(Lon, Lat) %>% 
  cbind(climDat_scaled)
names(climDatPred)[3:30] <- str_remove(names(climDatPred)[3:30], pattern = "_s$")
rm(climDat_Temp, climDat_noSF, climDatPred_extract, climDat_scaled, climDat_noSF, climDat)
gc()
```

<!-- Make sure that the locations we're using for the model-fitting are the same as the locations used to determine the proportion of CONUS that is un-forested (i.e. masks out the same locations due to fire and/or land use) -->
```{r}

# rasterize the model-fitting data on the dayMet grid
# modDat_1_rast <- modDat_1 %>%
#   st_as_sf(coords = c("x", "y"), crs = crs("PROJCRS[\"unnamed\",\n    BASEGEOGCRS[\"unknown\",\n        DATUM[\"unknown\",\n            ELLIPSOID[\"Spheroid\",6378137,298.257223563,\n                LENGTHUNIT[\"metre\",1,\n                    ID[\"EPSG\",9001]]]],\n        PRIMEM[\"Greenwich\",0,\n            ANGLEUNIT[\"degree\",0.0174532925199433,\n                ID[\"EPSG\",9122]]]],\n    CONVERSION[\"Lambert Conic Conformal (2SP)\",\n        METHOD[\"Lambert Conic Conformal (2SP)\",\n            ID[\"EPSG\",9802]],\n        PARAMETER[\"Latitude of false origin\",42.5,\n            ANGLEUNIT[\"degree\",0.0174532925199433],\n            ID[\"EPSG\",8821]],\n        PARAMETER[\"Longitude of false origin\",-100,\n            ANGLEUNIT[\"degree\",0.0174532925199433],\n            ID[\"EPSG\",8822]],\n        PARAMETER[\"Latitude of 1st standard parallel\",25,\n            ANGLEUNIT[\"degree\",0.0174532925199433],\n            ID[\"EPSG\",8823]],\n        PARAMETER[\"Latitude of 2nd standard parallel\",60,\n            ANGLEUNIT[\"degree\",0.0174532925199433],\n            ID[\"EPSG\",8824]],\n        PARAMETER[\"Easting at false origin\",0,\n            LENGTHUNIT[\"metre\",1],\n            ID[\"EPSG\",8826]],\n        PARAMETER[\"Northing at false origin\",0,\n            LENGTHUNIT[\"metre\",1],\n            ID[\"EPSG\",8827]]],\n    CS[Cartesian,2],\n        AXIS[\"easting\",east,\n            ORDER[1],\n            LENGTHUNIT[\"metre\",1,\n                ID[\"EPSG\",9001]]],\n        AXIS[\"northing\",north,\n            ORDER[2],\n            LENGTHUNIT[\"metre\",1,\n                ID[\"EPSG\",9001]]]]")) %>%
#   terra::vect() %>%
#   terra::rasterize(y = test_rastTemp, field = "TotalTreeCover")

## compare to the the RAP forest data from Martin (is the 'notForest' object)
# reproject notForest to match data I'm using
notForest2 <- notForest %>%
  terra::project(crs(climDatPred))

# notForestTrim <- notForest2 %>%
#   terra::crop(terra::ext(climDatPred[1:1000,]))
# make sure crs matches
#crs(notForest2) == crs(climDatPred)

climDatPred_vect <- climDatPred %>%
  terra::vect()

# crop climDatPred to match the extent of the notForest data
climDatPred_extract <- notForest2 %>%
  terra::extract(y = climDatPred_vect) %>%
  cbind(climDatPred) %>%
  st_as_sf()

## remove climate data from the locations that have been masked in Martin's tree data
climDatPred_new <- climDatPred_extract %>%
  filter(!is.na(fracNotForest))
```

## Predict on the test data, the complete dataset, and then the masked wall-to-wall data

```{r predict with best lambda on data used for fitting}
  # create prediction for each each model
# (i.e. for each fire proporation variable)
predict_by_response <- function(mod, df, scale = FALSE) {
  df_out <- df
    if(scale == TRUE) {
  df_out <- df_out %>% 
  mutate(across(all_of(prednames), base::scale,scale = FALSE, center = FALSE)) 
  }
  response_name <- paste0("TotalTreeCover_binom", "_pred")
  preds <- predict(mod, newdata = df_out, #s="lambda.min", 
                                     type = "response")
  #preds[preds<0] <- 0
  #preds[preds>100] <- 100
  df_out <- df_out %>% cbind(preds)
   colnames(df_out)[ncol(df_out)] <- response_name
  return(df_out)
}
## predict on the test dataset
pred_glm1_test <- predict_by_response(fit_glm_bestLambda, modDat_1_sTest)#X[,2:ncol(X)])
pred_glm1_test$resid <- pred_glm1_test[,"TotalTreeCover_binom"] - pred_glm1_test[,paste0("TotalTreeCover_binom", "_pred")]
pred_glm1_test$extremeResid <- NA
pred_glm1_test[pred_glm1_test$resid > .5 | pred_glm1_test$resid < -.5,"extremeResid"] <- 1

## predict on entire dataset
pred_glm1_full <- predict_by_response(fit_glm_bestLambda, rbind(modDat_1_sTest, modDat_1_s))#X[,2:ncol(X)])
pred_glm1_full$resid <- pred_glm1_full[,"TotalTreeCover_binom"] - pred_glm1_full[,paste0("TotalTreeCover_binom", "_pred")]
pred_glm1_full$extremeResid <- NA
pred_glm1_full[pred_glm1_full$resid > .5 | pred_glm1_full$resid < -.5,"extremeResid"] <- 1

## predict on the wall-to-wall dataset
pred_glm1_w2w <- predict_by_response(fit_glm_bestLambda, climDatPred_new %>% st_drop_geometry(), scale = TRUE)

```

```{r predict with second best lambda model on data used for model fitting}
if ( name_secondBestMod == "Intercept_Only") {
  print("The next best lambda model only contains one predictor (an intercept)")
} else {

pred_glm1_1se_test <- predict_by_response(mod_secondBest, modDat_1_sTest)#X[,2:ncol(X)])
pred_glm1_1se_test$resid <- pred_glm1_1se_test[,"TotalTreeCover_binom"] - pred_glm1_1se_test[,paste0("TotalTreeCover_binom", "_pred")]
pred_glm1_1se_test$extremeResid <- NA
pred_glm1_1se_test[pred_glm1_1se_test$resid > .5 | pred_glm1_1se_test$resid < -.5,"extremeResid"] <- 1

## for full dataset
pred_glm1_1se_full <- predict_by_response(mod_secondBest, rbind(modDat_1_sTest, modDat_1_s))#X[,2:ncol(X)])
pred_glm1_1se_full$resid <- pred_glm1_1se_full[,"TotalTreeCover_binom"] - pred_glm1_1se_full[,paste0("TotalTreeCover_binom", "_pred")]
pred_glm1_1se_full$extremeResid <- NA
pred_glm1_1se_full[pred_glm1_1se_full$resid > .5 | pred_glm1_1se_full$resid < -.5,"extremeResid"] <- 1

## predict on the wall-to-wall dataset
pred_glm1_1se_w2w <- predict_by_response(mod_secondBest, climDatPred_new %>% st_drop_geometry(), scale = TRUE)
}
```


## Inital Model Diagnostics 
Identifying the optimal threshold to make the continuous predictions binomial based on the prevalence in the training data. 
ROC Curves are based on classification accuracy of the model when used on the testing set (held-out years 2003, 2012 and 2019)

For the bestLambda model: 
```{r ROC for best lambda model, warning=FALSE}
library(PresenceAbsence)
DATA <- fullModOut$modelPredictions[,c("obs", "pred_opt")] # only training data
DATA$ID <- 1:nrow(DATA)
#pred.prev <- predicted.prevalence(DATA = DATA[,c("ID", "TotalTreeCover_binom", "TotalTreeCover_binom_pred")], threshold = 20)
#accu <- presence.absence.accuracy(DATA[,c("ID", "TotalTreeCover_binom", "TotalTreeCover_binom_pred")], threshold = 20)
optimalThresholds <- optimal.thresholds(DATA = DATA[,c("ID", "obs", "pred_opt")], 
                                        threshold = 200, obs.prev = mean(DATA[,"obs"])) # tests 100 thresholds)
rm(DATA)
best_threshold_bestLambda <- optimalThresholds[optimalThresholds$Method == thresholdMethod, 2]
cat(paste0("Optimal threshold for best lambda model based on ROC curve, based on the ",thresholdMethod, " method: ", best_threshold_bestLambda, "\n"))

# Now make an ROC curve based on the test set
roc_curve <- roc(pred_glm1_test$TotalTreeCover_binom, pred_glm1_test$TotalTreeCover_binom_pred)
# Plot ROC curve
plot(roc_curve, main =  paste0("ROC Curve for best lambda model \n","AUC: ",round(roc_curve$auc,2)))

## save this prevalence for later 
saveRDS(best_threshold_bestLambda, "./models/ForestModelPrevalence.rds")
```

For the second-bestLambda model: 
```{r ROC for next best lambda model}
# best_threshold_secondBest <- coords$threshold
DATA <- fullModOut$modelPredictions[,c("obs", trainDatPredName)]
DATA$ID <- 1:nrow(DATA)
#pred.prev <- predicted.prevalence(DATA = DATA[,c("ID", "TotalTreeCover_binom", "TotalTreeCover_binom_pred")], threshold = 20)
#accu <- presence.absence.accuracy(DATA[,c("ID", "TotalTreeCover_binom", "TotalTreeCover_binom_pred")], threshold = 20)
optimalThresholds <- optimal.thresholds(DATA = DATA[,c("ID", "obs", trainDatPredName)], 
                                        threshold = 200, obs.prev = mean(DATA[,"obs"])) # tests 100 thresholds)
rm(DATA)
best_threshold_secondBest <- optimalThresholds[optimalThresholds$Method == thresholdMethod, 2]
cat(paste0("Optimal threshold for second-best lambda model based on ROC curve, based on the ",thresholdMethod, " method: ", best_threshold_secondBest, "\n"))

# Now make an ROC curve based on the test set
roc_curve_secondBest <- roc(pred_glm1_1se_test$TotalTreeCover_binom, pred_glm1_1se_test$TotalTreeCover_binom_pred, smooth = FALSE)
# Plot ROC curve
plot(roc_curve_secondBest, main =  paste0("ROC Curve for second-best lambda model \n","AUC: ",round(roc_curve_secondBest$auc,2)))

```
## Interpolation
We use the threshold identified above to generate binary predictions.

<!-- , we also add a calibration step to ensure that this model estimates locations with no tree cover at a frequency that matches estimates of the area of CONUS that is tree-less. To do this, we classify all RAP cells in un-developed areas as either treeless (RAP fractional tree cover < 1%) or treed. We then determine the proportion of each 1km x 1km dayMet grid cell that is treeless based on the RAP cells contained within that daymet cell. We can then use this information estimate the proportion of the area within CONUS that is treeless. When we use our forest model to predict forest probability using dayMet data across CONUS, we want to predict an area that is treeless that corresponds to the estimate we made using RAP data.  -->

Below are figures of the empirical cumulative distribution functions for model-predicted forest probability from the best and second-best lambda models *before* this interpolation step. The vertical dashed line shows the threshold identified above to make the predicted values binary. The  dotted lines on each side show the values of P(forest) within which we will use the predicted probabilities to "blend" the subsequent total tree cover model predictions together. We identified these boundaries such that there was an equal percentage (0.05%) of values on each side of the cutoff.  

```{r}
## make an empirical cumulative probability function plot of the model predictions on wall-to-wall data
pred_glm1_w2w_cdf <- stats::ecdf(pred_glm1_w2w$TotalTreeCover_binom_pred)
# ## find the quantile of observations that corresponds to the desired proportion of unforested locations
thresholdPercentile_bestLambda <- pred_glm1_w2w_cdf(best_threshold_bestLambda)
  #quantile(pred_glm1_w2w$TotalTreeCover_binom_pred %>% na.omit(), probs = best_threshold_bestLambda) 
# get P(forest) for percentiles +/- 0.05 from the percentile of the optimal threshold
lowBuffProb <- quantile(pred_glm1_w2w$TotalTreeCover_binom_pred %>% na.omit(), probs = thresholdPercentile_bestLambda - 0.05) 

highBuffProb <- quantile(pred_glm1_w2w$TotalTreeCover_binom_pred %>% na.omit(), probs = thresholdPercentile_bestLambda +  0.05) 

# plot the ecdf
plot(pred_glm1_w2w_cdf, xlab = c("P(Forest)"), ylab = "Percentile",
     main = "Empirical Cumulative Distribution Function for \nUn-transformed model predictions of \nP(Forest) from the best-lambda model")
# lines for optimal threshold
abline(v = best_threshold_bestLambda, lty = 2)
abline(h = thresholdPercentile_bestLambda, lty = 2)
abline(v = lowBuffProb, lty = 3)
abline(h = thresholdPercentile_bestLambda - 0.05, lty = 3)
abline(v = highBuffProb, lty = 3)
abline(h = thresholdPercentile_bestLambda + 0.05, lty = 3)
text(x = best_threshold_bestLambda + 0.05, y = 0.1, labels = round(best_threshold_bestLambda, 4))
text(x = lowBuffProb +.05, y = 0.05, labels = round(lowBuffProb, 4))
text(x = highBuffProb +.05, y = 0.15, labels = round(highBuffProb, 4))

## save buffers for later

saveRDS(lowBuffProb, "./models/ForestModel_interpLowerBuffer.rds")
saveRDS(highBuffProb, "./models/ForestModel_interpHigherBuffer.rds")
```

For the best-lambda model: 
If we want 31% of model predictions to be zero, then any model predictions ≤ 0.136147 should be set to zero. 

```{r}
## make an empirical cumulative probability function plot of the model predictions on wall-to-wall data
pred_glm1_1se_w2w_cdf <- stats::ecdf(pred_glm1_1se_w2w$TotalTreeCover_binom_pred)
## find the quantile of observations that corresponds to the desired proportion of unforested locations
#0 <- quantile(pred_glm1_1se_w2w$TotalTreeCover_binom_pred %>% na.omit(), probs = 0.3133397) 
# plot the ecdf
plot(pred_glm1_w2w_cdf, xlab = c("P(Total Tree Cover)"), ylab = "Percentile",
     main = "Empirical Cumulative Distribution Function for \nUn-transformed model predictions of \nP(Forest) from the second best-lambda model")
abline(v = best_threshold_secondBest, lty = 2)
text(x = best_threshold_secondBest+.05, y = 0.65, labels = round(best_threshold_secondBest, 4))
abline(v = best_threshold_secondBest - 0.1, lty = 3)
text(x = best_threshold_secondBest - 0.1 +.05, y = 0.55, labels = round(best_threshold_secondBest - 0.1, 4))
abline(v = best_threshold_secondBest + 0.1, lty = 3)
text(x = best_threshold_secondBest + 0.1 +.05, y = 0.75, labels = round(best_threshold_secondBest + 0.1, 4))
```

For the 1se-lambda model: 
If we want 31% of model predictions to be zero, then any model predictions ≤ 0.1398974 should be set to zero. 

## Interpolation

 We use the identified thresholds and 0-calibration points identified above to calibrate continuous and binary model predictions of forest probability

First, interpolate the continuous predictions
```{r}
## set up function for linear interpolation - best lambda model
approxfun_bestLambda <- approxfun(x = c(0,  best_threshold_bestLambda, 1), # values we're interpolating *from*
          y =c(0, 0.1, 1), # what we're interpolation to (desired values at the end)
          method = "linear")

## set up function for linear interpolation - second-best lambda model
approxfun_secondBestLambda <- approxfun(x = c(0,  best_threshold_secondBest, 1), # values we're interpolating *from*
          y =c(0, 0.1, 1), # what we're interpolation to (desired values at the end)
          method = "linear") 

## for test dataset
# best lambda model
# "binomialize" the continuous predictions
pred_glm1_test <- pred_glm1_test %>% 
  mutate(TotalTreeCover_interp = approxfun_bestLambda(TotalTreeCover_binom_pred)) %>% 
  mutate(TotalTreeCover_interp = replace(TotalTreeCover_interp, list = TotalTreeCover_binom_pred < 0, 0))

# second-best lambda model
# "binomialize" the continuouse predictions
pred_glm1_1se_test <- pred_glm1_1se_test %>% 
  mutate(TotalTreeCover_interp = approxfun_secondBestLambda(TotalTreeCover_binom_pred))  %>% 
  mutate(TotalTreeCover_interp = replace(TotalTreeCover_interp, list = TotalTreeCover_binom_pred < 0, 0))

## for full dataset
# best lambda model
# "binomialize" the continuous predictions
pred_glm1_full <- pred_glm1_full %>% 
  mutate(TotalTreeCover_interp = approxfun_bestLambda(TotalTreeCover_binom_pred)) %>% 
  mutate(TotalTreeCover_interp = replace(TotalTreeCover_interp, list = TotalTreeCover_binom_pred < 0, 0))
# second-best lambda model
# "binomialize" the continuouse predictions
pred_glm1_1se_full <- pred_glm1_1se_full %>% 
  mutate(TotalTreeCover_interp = approxfun_secondBestLambda(TotalTreeCover_binom_pred))  %>% 
  mutate(TotalTreeCover_interp = replace(TotalTreeCover_interp, list = TotalTreeCover_binom_pred < 0, 0))

## for wall-to-wall data
# best lambda model
# "binomialize" the continuous predictions
pred_glm1_w2w <- pred_glm1_w2w %>% 
  mutate(TotalTreeCover_interp = approxfun_bestLambda(TotalTreeCover_binom_pred)) %>% 
  mutate(TotalTreeCover_interp = replace(TotalTreeCover_interp, list = TotalTreeCover_binom_pred < 0, 0))

# second-best lambda model
# "binomialize" the continuouse predictions
pred_glm1_1se_w2w <- pred_glm1_1se_w2w %>% 
  mutate(TotalTreeCover_interp = approxfun_secondBestLambda(TotalTreeCover_binom_pred))  %>% 
  mutate(TotalTreeCover_interp = replace(TotalTreeCover_interp, list = TotalTreeCover_binom_pred < 0, 0))

```

Now, Make the continuous predictions binary
```{r}
## for test dataset
# best lambda model
# "binomialize" the continuous predictions
pred_glm1_test <- pred_glm1_test %>% 
  mutate(TotalTreeCover_binom_pred_binary = TotalTreeCover_interp,
    TotalTreeCover_binom_pred_binary = replace(TotalTreeCover_binom_pred_binary, TotalTreeCover_binom_pred_binary > 0.1, 1),
         TotalTreeCover_binom_pred_binary = replace(TotalTreeCover_binom_pred_binary, TotalTreeCover_binom_pred_binary <= 0.1, 0)) 

# second-best lambda model
# "binomialize" the continuouse predictions
pred_glm1_1se_test <- pred_glm1_1se_test %>% 
  mutate(TotalTreeCover_binom_pred_binary = TotalTreeCover_interp,
    TotalTreeCover_binom_pred_binary = replace(TotalTreeCover_binom_pred_binary, TotalTreeCover_binom_pred_binary > 0.1, 1),
         TotalTreeCover_binom_pred_binary = replace(TotalTreeCover_binom_pred_binary, TotalTreeCover_binom_pred_binary <= 0.1, 0)) 

## for full dataset
# best lambda model
# "binomialize" the continuous predictions
pred_glm1_full <- pred_glm1_full %>% 
  mutate(TotalTreeCover_binom_pred_binary = TotalTreeCover_interp,
    TotalTreeCover_binom_pred_binary = replace(TotalTreeCover_binom_pred_binary, TotalTreeCover_binom_pred_binary > 0.1, 1),
         TotalTreeCover_binom_pred_binary = replace(TotalTreeCover_binom_pred_binary, TotalTreeCover_binom_pred_binary <= 0.1, 0)) 

# second-best lambda model
# "binomialize" the continuouse predictions
pred_glm1_1se_full <- pred_glm1_1se_full %>% 
  mutate(TotalTreeCover_binom_pred_binary = TotalTreeCover_interp,
    TotalTreeCover_binom_pred_binary = replace(TotalTreeCover_binom_pred_binary, TotalTreeCover_binom_pred_binary > 0.1, 1),
         TotalTreeCover_binom_pred_binary = replace(TotalTreeCover_binom_pred_binary, TotalTreeCover_binom_pred_binary <= 0.1, 0))

## for wall-to-wall data
# best lambda model
# "binomialize" the continuous predictions
pred_glm1_w2w <- pred_glm1_w2w %>% 
  mutate(TotalTreeCover_binom_pred_binary = TotalTreeCover_interp,
    TotalTreeCover_binom_pred_binary = replace(TotalTreeCover_binom_pred_binary, TotalTreeCover_binom_pred_binary > 0.1, 1),
         TotalTreeCover_binom_pred_binary = replace(TotalTreeCover_binom_pred_binary, TotalTreeCover_binom_pred_binary <= 0.1, 0)) 

# second-best lambda model
# "binomialize" the continuouse predictions
pred_glm1_1se_w2w <- pred_glm1_1se_w2w %>% 
  mutate(TotalTreeCover_binom_pred_binary = TotalTreeCover_interp,
    TotalTreeCover_binom_pred_binary = replace(TotalTreeCover_binom_pred_binary, TotalTreeCover_binom_pred_binary > 0.1, 1),
         TotalTreeCover_binom_pred_binary = replace(TotalTreeCover_binom_pred_binary, TotalTreeCover_binom_pred_binary <= 0.1, 0)) 
```

Below, we show empirical cumulative distribution functions for model-predicted forest probability from the best and second-best lambda models *after* interpolation and calibration. 
```{r}
# best lambda model
## make an empirical cumulative probability function plot of the model predictions on wall-to-wall data
pred_glm1_w2w_cdf_postInterp <- stats::ecdf(pred_glm1_w2w$TotalTreeCover_interp)
# plot the ecdf
plot(pred_glm1_w2w_cdf_postInterp, xlab = c("P(Total Tree Cover)"), ylab = "Percentile",
     main = "Empirical Cumulative Distribution Function for \nCalibrated and Interpolated model predictions of \nP(Forest) from the best-lambda model")
```
```{r}
# best lambda model
## make an empirical cumulative probability function plot of the model predictions on wall-to-wall data
pred_glm1_1se_w2w_cdf_postInterp <- stats::ecdf(pred_glm1_1se_w2w$TotalTreeCover_interp)
# plot the ecdf
plot(pred_glm1_1se_w2w_cdf_postInterp, xlab = c("P(Total Tree Cover)"), ylab = "Percentile",
     main = "Empirical Cumulative Distribution Function for \nCalibrated and Interpolated model predictions of \nP(Forest) from the second best-lambda model")
```

### Identify Misclassifications
```{r}
# true positive = tree (1)
# false positive = model says tree, data says no-tree (2) 
# true negative = no tree (3) 
# false negative = model says no tree, data says tree (4)

# best lambda model - test dataset
pred_glm1_test$missClass <- NA
pred_glm1_test[pred_glm1_test$TotalTreeCover_binom == 1 & 
                 pred_glm1_test$TotalTreeCover_binom_pred_binary == 1, "missClass"] <- 1
pred_glm1_test[pred_glm1_test$TotalTreeCover_binom == 0 & 
                 pred_glm1_test$TotalTreeCover_binom_pred_binary == 0, "missClass"] <- 3
pred_glm1_test[pred_glm1_test$TotalTreeCover_binom == 0 & 
                 pred_glm1_test$TotalTreeCover_binom_pred_binary == 1, "missClass"] <- 2
pred_glm1_test[pred_glm1_test$TotalTreeCover_binom == 1 & 
                 pred_glm1_test$TotalTreeCover_binom_pred_binary == 0, "missClass"] <- 4

# best lambda model - full dataset
pred_glm1_full$missClass <- NA
pred_glm1_full[pred_glm1_full$TotalTreeCover_binom == 1 & 
                 pred_glm1_full$TotalTreeCover_binom_pred_binary == 1, "missClass"] <- 1
pred_glm1_full[pred_glm1_full$TotalTreeCover_binom == 0 & 
                 pred_glm1_full$TotalTreeCover_binom_pred_binary == 0, "missClass"] <- 3
pred_glm1_full[pred_glm1_full$TotalTreeCover_binom == 0 & 
                 pred_glm1_full$TotalTreeCover_binom_pred_binary == 1, "missClass"] <- 2
pred_glm1_full[pred_glm1_full$TotalTreeCover_binom == 1 & 
                 pred_glm1_full$TotalTreeCover_binom_pred_binary == 0, "missClass"] <- 4

# second-best lambda model - test dataset
pred_glm1_1se_test$missClass <- NA
pred_glm1_1se_test[pred_glm1_1se_test$TotalTreeCover_binom == 1 & 
                 pred_glm1_1se_test$TotalTreeCover_binom_pred_binary == 1, "missClass"] <- 1
pred_glm1_1se_test[pred_glm1_1se_test$TotalTreeCover_binom == 0 & 
                 pred_glm1_1se_test$TotalTreeCover_binom_pred_binary == 0, "missClass"] <- 3
pred_glm1_1se_test[pred_glm1_1se_test$TotalTreeCover_binom == 0 & 
                 pred_glm1_1se_test$TotalTreeCover_binom_pred_binary == 1, "missClass"] <- 2
pred_glm1_1se_test[pred_glm1_1se_test$TotalTreeCover_binom == 1 & 
                 pred_glm1_1se_test$TotalTreeCover_binom_pred_binary == 0, "missClass"] <- 4

# second-best lambda model - full dataset
pred_glm1_1se_full$missClass <- NA
pred_glm1_1se_full[pred_glm1_1se_full$TotalTreeCover_binom == 1 & 
                 pred_glm1_1se_full$TotalTreeCover_binom_pred_binary == 1, "missClass"] <- 1
pred_glm1_1se_full[pred_glm1_1se_full$TotalTreeCover_binom == 0 & 
                 pred_glm1_1se_full$TotalTreeCover_binom_pred_binary == 0, "missClass"] <- 3
pred_glm1_1se_full[pred_glm1_1se_full$TotalTreeCover_binom == 0 & 
                 pred_glm1_1se_full$TotalTreeCover_binom_pred_binary == 1, "missClass"] <- 2
pred_glm1_1se_full[pred_glm1_1se_full$TotalTreeCover_binom == 1 & 
                 pred_glm1_1se_full$TotalTreeCover_binom_pred_binary == 0, "missClass"] <- 4

```

## Maps of Observations, Predictions, and Residuals

### Observations - continuous and binary total tree cover for the full dataset 
```{r Residual maps, fig.height=14, fig.width=9}
## get information for plotting
# aggregate to a grid of 8km x 8km cells
test_rast <- test_rastTemp %>%
 terra::aggregate(fact = 8, fun = "mean")#%>% 

# function for getting the statistical mode
Mode <- function(x, ...) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

# string w/ crs of veg data
crs_veg <- "PROJCRS[\"unnamed\",\n    BASEGEOGCRS[\"unknown\",\n        DATUM[\"unknown\",\n            ELLIPSOID[\"Spheroid\",6378137,298.257223563,\n                LENGTHUNIT[\"metre\",1,\n                    ID[\"EPSG\",9001]]]],\n        PRIMEM[\"Greenwich\",0,\n            ANGLEUNIT[\"degree\",0.0174532925199433,\n                ID[\"EPSG\",9122]]]],\n    CONVERSION[\"Lambert Conic Conformal (2SP)\",\n        METHOD[\"Lambert Conic Conformal (2SP)\",\n            ID[\"EPSG\",9802]],\n        PARAMETER[\"Latitude of false origin\",42.5,\n            ANGLEUNIT[\"degree\",0.0174532925199433],\n            ID[\"EPSG\",8821]],\n        PARAMETER[\"Longitude of false origin\",-100,\n            ANGLEUNIT[\"degree\",0.0174532925199433],\n            ID[\"EPSG\",8822]],\n        PARAMETER[\"Latitude of 1st standard parallel\",25,\n            ANGLEUNIT[\"degree\",0.0174532925199433],\n            ID[\"EPSG\",8823]],\n        PARAMETER[\"Latitude of 2nd standard parallel\",60,\n            ANGLEUNIT[\"degree\",0.0174532925199433],\n            ID[\"EPSG\",8824]],\n        PARAMETER[\"Easting at false origin\",0,\n            LENGTHUNIT[\"metre\",1],\n            ID[\"EPSG\",8826]],\n        PARAMETER[\"Northing at false origin\",0,\n            LENGTHUNIT[\"metre\",1],\n            ID[\"EPSG\",8827]]],\n    CS[Cartesian,2],\n        AXIS[\"easting\",east,\n            ORDER[1],\n            LENGTHUNIT[\"metre\",1,\n                ID[\"EPSG\",9001]]],\n        AXIS[\"northing\",north,\n            ORDER[2],\n            LENGTHUNIT[\"metre\",1,\n                ID[\"EPSG\",9001]]]]"
# make a multi-layer raster w/ layers for each variable of interest
pred_glm1_full_sf <- pred_glm1_full %>% 
  st_as_sf(coords = c("x", "y"), crs = crs(crs_veg)) %>% 
  st_transform(crs(test_rast)) %>% 
  terra::vect()
testRast2 <- lapply(c("TotalTreeCover", "TotalTreeCover_binom", "TotalTreeCover_interp", "TotalTreeCover_binom_pred_binary", "resid", "missClass"), FUN = function(x) {
  # S4 method for class 'data.frame'
  if (x %in% c("TotalTreeCover", "TotalTreeCover_binom", "TotalTreeCover_interp", "TotalTreeCover_binom_pred_binary", "resid")) {
    temp <- terra::rasterize(x = pred_glm1_full_sf, y = test_rast, 
                         field = x, fun = mean, na.rm = TRUE)
    names(temp) <- x
  } else {
    temp <- terra::rasterize(x = pred_glm1_full_sf, y = test_rast, 
                         field = x, fun = Mode, na.rm = TRUE)
    # get the % aggreement between values in the dayMetx8 cell 
    temp2 <- terra::rasterize(x = pred_glm1_full_sf, y = test_rast$daymet_v4_prcp_monttl_na_1980_1, 
                         field = x, fun = function(y) {
                           yTemp <- y[!is.na(y)]
                           yMode <- Mode(yTemp)
                           (sum(yTemp == yMode)/length(yTemp))*100
                         })
    temp <- c(as.factor(temp$missClass), temp2)
    names(temp) <- c("missClass", "PercAgreement")
  }
    
 return(temp)
  }
)
pred_glm1_full_RAST <- c(testRast2[[1]], testRast2[[2]], testRast2[[3]], testRast2[[4]], testRast2[[5]], testRast2[[6]])

# reproject state boundaries
cropped_states_2 <- cropped_states  %>%
  sf::st_transform(sf::st_crs(test_rast))
# make figure of raw tree cover
map_obs_cont <- ggplot() +
geom_spatraster(data = pred_glm1_full_RAST, aes(fill = TotalTreeCover), maxcell = 50000000) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) +
  coord_sf(datum = st_crs(test_rastTemp)) +
             scale_fill_viridis_c(na.value = "white", "magma", direction = -1, limits = c(0,212)) +
 #geom_sf(data = pred_glm1_full_RAST, aes(col = TotalTreeCover)) + 
  # stat_summary_2d(data = plotObs, aes(x = x, y = y, z = TotalTreeCover), fun = mean, binwidth = .05) + 
labs(title = paste0("Observations of Continuous Total Tree Cover - all data")) +
  scale_fill_gradient2(low = "brown",
                       mid = "wheat" ,
                       high = "forestgreen" , 
                       midpoint = 0,   na.value = "white") +
  theme(axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) 
# add inset
           inset_plot <-  ggplot(pred_glm1_full) + 
             geom_density(aes(TotalTreeCover)) +
                theme_bw(base_size=9) +  ## makes everything smaller
                theme(panel.background = element_rect(fill="transparent"),  ## white plot background
                      panel.border = element_blank(),
          axis.title.y = element_blank(),
          axis.title.x = element_blank(),
          axis.text.x = element_text(size=rel(2)), ## tiny axis text
          axis.text.y = element_text(size=rel(2)), ## tiny axis text
          axis.line.x = element_line(size = .5),
          axis.line.y = element_line(size = .5),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          plot.background = element_blank(),
          strip.background = element_blank(),
          strip.text = element_blank()) 
     
 map_obs_cont <-    map_obs_cont +   annotation_custom(grob = ggplotGrob(inset_plot), ymin = -2025000, ymax=-1205000, xmin=-2280000, xmax=-180000 ) 

 # make figure of binomialized tree cover
map_obs <- ggplot() +
geom_spatraster(data = pred_glm1_full_RAST, aes(fill = TotalTreeCover_binom), maxcell = 50000000) + 
 #geom_sf(data = plotObs, aes(col = TotalTreeCover_binom)) + 
   #stat_summary_2d(data = plotObs, aes(x = x, y = y, z = TotalTreeCover_binom), fun = mean, binwidth = .05) + 
     scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"), 
                          limits = c(0,100))  +
  geom_sf(data=cropped_states_2 ,fill=NA ) +
labs(title = paste0("Observations of Binomial Total Tree Cover - all data")) +
  scale_fill_continuous(breaks =  c(0, 0.5, 1), palette = c("wheat", "forestgreen"),
                    labels = c("No Trees",
                               "",
                      "Trees"), na.value = "white") + 
  coord_sf(datum = st_crs(test_rastTemp)) +
    theme(axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) 


# add inset
           inset_plot <-  ggplot(pred_glm1_full) + 
             geom_density(aes(TotalTreeCover_binom)) +
                theme_bw(base_size=9) +  ## makes everything smaller
                theme(panel.background = element_rect(fill="transparent"),  ## white plot background
                      panel.border = element_blank(),
          axis.title.y = element_blank(),
          axis.title.x = element_blank(),
          axis.text.x = element_text(size=rel(2)), ## tiny axis text
          axis.text.y = element_text(size=rel(2)), ## tiny axis text
          axis.line.x = element_line(size = .5),
          axis.line.y = element_line(size = .5),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          plot.background = element_blank(),
          strip.background = element_blank(),
          strip.text = element_blank()) 
     
 map_obs <-    map_obs +   annotation_custom(grob = ggplotGrob(inset_plot), ymin = -2155000, ymax=-1409000, xmin=-2280000, xmax=-180000 ) 

library(ggpubr)
ggarrange(map_obs_cont,  map_obs, heights = c(1, 1), ncol = 1)

```

### Observations - continuous and binary total tree cover for the test dataset 

```{r observations in the test dataset , fig.height = 14, fig.width = 9}
pred_glm1_test_sf <- pred_glm1_test %>% 
  st_as_sf(coords = c("x", "y"), crs = crs(crs_veg)) %>% 
  st_transform(crs(test_rast)) %>% 
  terra::vect()
# make raster of test set 
# make a multi-layer raster w/ layers for each variable of interest
testRast_test <- lapply(c("TotalTreeCover", "TotalTreeCover_binom", "TotalTreeCover_interp", "TotalTreeCover_binom_pred_binary", "resid", "missClass"), FUN = function(x) {
  # S4 method for class 'data.frame'
  if (x %in% c("TotalTreeCover", "TotalTreeCover_binom", "TotalTreeCover_interp", "TotalTreeCover_binom_pred_binary", "resid")) {
    temp <- terra::rasterize(x = pred_glm1_test_sf, y = test_rast, 
                         field = x, fun = mean, na.rm = TRUE)
    names(temp) <- x
  } else {
    temp <- terra::rasterize(x = pred_glm1_test_sf, y = test_rast, 
                         field = x, fun = Mode, na.rm = TRUE)
    # get the % aggreement between values in the dayMetx8 cell 
    temp2 <- terra::rasterize(x = pred_glm1_test_sf, y = test_rast$daymet_v4_prcp_monttl_na_1980_1, 
                         field = x, fun = function(y) {
                           yTemp <- y[!is.na(y)]
                           yMode <- Mode(yTemp)
                           (sum(yTemp == yMode)/length(yTemp))*100
                         })
    temp <- c(as.factor(temp$missClass), temp2)
    names(temp) <- c("missClass", "PercAgreement")
  }
    
 return(temp)
  }
)
pred_glm1_test_RAST <- c(testRast_test[[1]], testRast_test[[2]], testRast_test[[3]], testRast_test[[4]], testRast_test[[5]], testRast_test[[6]])

# make maps 
map_obs_cont_test <- ggplot() +
geom_spatraster(data = pred_glm1_test_RAST, aes(fill = TotalTreeCover), maxcell = 50000000) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
labs(title = paste0("Observations of Continuous Total Tree Cover - test data")) +
  scale_fill_gradient2(low = "brown",
                       mid = "wheat" ,
                       high = "forestgreen" , 
                       midpoint = 0,   na.value = "white") + 
  coord_sf(datum = st_crs(test_rastTemp)) +
    theme(axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) 
  
 inset_plot <-  ggplot(pred_glm1_test) + 
             geom_density(aes(TotalTreeCover)) +
                theme_bw(base_size=9) +  ## makes everything smaller
                theme(panel.background = element_rect(fill="transparent"),  ## white plot background
                      panel.border = element_blank(),
          axis.title.y = element_blank(),
          axis.title.x = element_blank(),
          axis.text.x = element_text(size=rel(2)), ## tiny axis text
          axis.text.y = element_text(size=rel(2)), ## tiny axis text
          axis.line.x = element_line(size = .5),
          axis.line.y = element_line(size = .5),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          plot.background = element_blank(),
          strip.background = element_blank(),
          strip.text = element_blank()) 
     
 map_obs_cont_test <-    map_obs_cont_test +   annotation_custom(grob = ggplotGrob(inset_plot), ymin = -2155000, ymax=-1409000, xmin=-2280000, xmax=-180000 ) 

# make figure of binomialized tree cover
map_obs_test <- ggplot() +
geom_spatraster(data = pred_glm1_test_RAST, aes(fill = TotalTreeCover_binom), maxcell = 50000000) +  
     scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"), 
                          limits = c(0,100))  +
  geom_sf(data=cropped_states_2 ,fill=NA ) +
labs(title = paste0("Observations of Binomial Total Tree Cover - test data")) +
  scale_fill_continuous(breaks =  c(0, 0.5, 1), palette = c("wheat", "forestgreen"),
                    labels = c("No Trees",
                               "",
                      "Trees"), na.value = "white") + 
  coord_sf(datum = st_crs(test_rastTemp)) +
    theme(axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) 

 inset_plot <-  ggplot(pred_glm1_test) + 
             geom_density(aes(TotalTreeCover_binom)) +
                theme_bw(base_size=9) +  ## makes everything smaller
                theme(panel.background = element_rect(fill="transparent"),  ## white plot background
                      panel.border = element_blank(),
          axis.title.y = element_blank(),
          axis.title.x = element_blank(),
          axis.text.x = element_text(size=rel(2)), ## tiny axis text
          axis.text.y = element_text(size=rel(2)), ## tiny axis text
          axis.line.x = element_line(size = .5),
          axis.line.y = element_line(size = .5),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          plot.background = element_blank(),
          strip.background = element_blank(),
          strip.text = element_blank()) 
     
 map_obs_test <-    map_obs_test +   annotation_custom(grob = ggplotGrob(inset_plot), ymin = -2155000, ymax=-1409000, xmin=-2280000, xmax=-180000 ) 

 
ggarrange(map_obs_cont_test,map_obs_test,heights = c(1,  1), ncol = 1)

```

### Predictions on the full and test dataset using the *best lambda* model
```{r predicted values map best lambda model, fig.height = 14, fig.width = 9 }

# make figures of predictions w/ the full dataset
map_preds1_cont <- ggplot() +
geom_spatraster(data = pred_glm1_full_RAST, aes(fill = TotalTreeCover_interp), maxcell = 20000000) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
labs(title = paste0("Predictions from the bestaLambda model of Yes/No Trees \ncontinuous probabilities \nfull dataset")) +
  scale_fill_gradient2(low = "brown",
                       mid = "wheat" ,
                       high = "forestgreen" , 
                       midpoint = 0,   na.value = "white") + 
  coord_sf(datum = st_crs(test_rastTemp)) +
    theme(axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) 

# add inset
           inset_plot <-  ggplot(pred_glm1_full) + 
             geom_density(aes(TotalTreeCover_interp)) +
                theme_bw(base_size=9) +  ## makes everything smaller
                theme(panel.background = element_rect(fill="transparent"),  ## white plot background
                      panel.border = element_blank(),
          axis.title.y = element_blank(),
          axis.title.x = element_blank(),
          axis.text.x = element_text(size=rel(2)), ## tiny axis text
          axis.text.y = element_text(size=rel(2)), ## tiny axis text
          axis.line.x = element_line(size = .5),
          axis.line.y = element_line(size = .5),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          plot.background = element_blank(),
          strip.background = element_blank(),
          strip.text = element_blank()) 
     
 map_preds1_cont <-    map_preds1_cont +   annotation_custom(grob = ggplotGrob(inset_plot), ymin = -2155000, ymax=-1409000, xmin=-2280000, xmax=-180000 ) 

# make figure of binomialized tree cover
map_preds1 <- ggplot() +
geom_spatraster(data = pred_glm1_full_RAST, aes(fill = TotalTreeCover_binom_pred_binary), maxcell = 20000000) + 
    geom_sf(data=cropped_states_2 ,fill=NA ) +
     scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"), 
                          limits = c(0,100))  +
labs(title = paste0("Predictions from the bestaLambda model of Yes/No Trees \nbinary probabilities \nfull dataset")) +
  scale_fill_continuous(breaks =  c(0, 0.5, 1), palette = c("wheat", "forestgreen"),
                    labels = c("No Trees",
                               "",
                      "Trees"), na.value = "white") + 
  coord_sf(datum = st_crs(test_rastTemp)) +
    theme(axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) 

# add inset
           inset_plot <-  ggplot(pred_glm1_full) + 
             geom_density(aes(TotalTreeCover_binom_pred_binary)) +
                theme_bw(base_size=9) +  ## makes everything smaller
                theme(panel.background = element_rect(fill="transparent"),  ## white plot background
                      panel.border = element_blank(),
          axis.title.y = element_blank(),
          axis.title.x = element_blank(),
          axis.text.x = element_text(size=rel(2)), ## tiny axis text
          axis.text.y = element_text(size=rel(2)), ## tiny axis text
          axis.line.x = element_line(size = .5),
          axis.line.y = element_line(size = .5),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          plot.background = element_blank(),
          strip.background = element_blank(),
          strip.text = element_blank()) 
     
 map_preds1 <-    map_preds1 +   annotation_custom(grob = ggplotGrob(inset_plot), ymin = -2155000, ymax=-1409000, xmin=-2280000, xmax=-180000 ) 

ggarrange(map_preds1_cont, map_preds1, heights = c(1,1), ncol = 1)


# make figures of predictions w/ the test dataset
map_preds1_cont_test <- ggplot() +
geom_spatraster(data = pred_glm1_test_RAST, aes(fill = TotalTreeCover_interp), maxcell = 20000000) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
labs(title = paste0("Predictions from the bestaLambda model of Yes/No Trees \ncontinuous probabilities \ntest dataset")) +
  scale_fill_gradient2(low = "brown",
                       mid = "wheat" ,
                       high = "forestgreen" , 
                       midpoint = 0,   na.value = "white") + 
  coord_sf(datum = st_crs(test_rastTemp)) +
    theme(axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) 

# add inset
           inset_plot <-  ggplot(pred_glm1_test) + 
             geom_density(aes(TotalTreeCover_interp)) +
                theme_bw(base_size=9) +  ## makes everything smaller
                theme(panel.background = element_rect(fill="transparent"),  ## white plot background
                      panel.border = element_blank(),
          axis.title.y = element_blank(),
          axis.title.x = element_blank(),
          axis.text.x = element_text(size=rel(2)), ## tiny axis text
          axis.text.y = element_text(size=rel(2)), ## tiny axis text
          axis.line.x = element_line(size = .5),
          axis.line.y = element_line(size = .5),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          plot.background = element_blank(),
          strip.background = element_blank(),
          strip.text = element_blank()) 
     
 map_preds1_cont_test <-    map_preds1_cont_test +   annotation_custom(grob = ggplotGrob(inset_plot), ymin = -2155000, ymax=-1409000, xmin=-2280000, xmax=-180000 ) 

 
# make figure of binomialized tree cover
map_preds1_test <- ggplot() +
geom_spatraster(data = pred_glm1_test_RAST, aes(fill = TotalTreeCover_binom_pred_binary), maxcell = 20000000) + 
     scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"), 
                          limits = c(0,100))  +
  geom_sf(data=cropped_states_2 ,fill=NA ) +
labs(title = paste0("Predictions from the bestaLambda model of Yes/No Trees \nbinary probabilities \ntest dataset")) +
  scale_fill_continuous(breaks =  c(0, 0.5, 1), palette = c("wheat", "forestgreen"),
                    labels = c("No Trees",
                               "",
                      "Trees"), na.value = "white") + 
  coord_sf(datum = st_crs(test_rastTemp)) +
    theme(axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) 

# add inset
           inset_plot <-  ggplot(pred_glm1_test) + 
             geom_density(aes(TotalTreeCover_binom_pred_binary)) +
                theme_bw(base_size=9) +  ## makes everything smaller
                theme(panel.background = element_rect(fill="transparent"),  ## white plot background
                      panel.border = element_blank(),
          axis.title.y = element_blank(),
          axis.title.x = element_blank(),
          axis.text.x = element_text(size=rel(2)), ## tiny axis text
          axis.text.y = element_text(size=rel(2)), ## tiny axis text
          axis.line.x = element_line(size = .5),
          axis.line.y = element_line(size = .5),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          plot.background = element_blank(),
          strip.background = element_blank(),
          strip.text = element_blank()) 
     
 map_preds1_test <-    map_preds1_test +   annotation_custom(grob = ggplotGrob(inset_plot), ymin = -2155000, ymax=-1409000, xmin=-2280000, xmax=-180000 ) 


ggarrange(map_preds1_cont_test, map_preds1_test, heights = c(1,1), ncol = 1)

```


### Predictions on the full and test datasets using the *second best lambda* model
```{r predicted values map 1se lambda model, fig.height = 28, fig.width = 9, }
if ( name_secondBestMod == "Intercept_Only") {
  print("The next best lambda model only contains one predictor (an intercept)")
} else {
  pred_glm1_1se_full_sf <- pred_glm1_1se_full %>% 
  st_as_sf(coords = c("x", "y"), crs = crs(crs_veg)) %>% 
  st_transform(crs(test_rast)) %>% 
  terra::vect()
  # make a raster of the full dataset
# make a multi-layer raster w/ layers for each variable of interest
testRast3 <- lapply(c("TotalTreeCover", "TotalTreeCover_binom", "TotalTreeCover_interp", "TotalTreeCover_binom_pred_binary", "resid", "missClass"), FUN = function(x) {
  # S4 method for class 'data.frame'
 if (x %in% c("TotalTreeCover", "TotalTreeCover_binom", "TotalTreeCover_interp", "TotalTreeCover_binom_pred_binary", "resid")) {
    temp <- terra::rasterize(x = pred_glm1_1se_full_sf, y = test_rast, 
                         field = x, fun = mean, na.rm = TRUE)
    names(temp) <- x
  } else {
    temp <- terra::rasterize(x = pred_glm1_1se_full_sf, y = test_rast, 
                         field = x, fun = Mode, na.rm = TRUE)
    # get the % aggreement between values in the dayMetx8 cell 
    temp2 <- terra::rasterize(x = pred_glm1_1se_full_sf, y = test_rast$daymet_v4_prcp_monttl_na_1980_1, 
                         field = x, fun = function(y) {
                           yTemp <- y[!is.na(y)]
                           yMode <- Mode(yTemp)
                           (sum(yTemp == yMode)/length(yTemp))*100
                         })
    temp <- c(as.factor(temp$missClass), temp2)
    names(temp) <- c("missClass", "PercAgreement")
  }
    
 return(temp)
  }
)
pred_glm1_1se_full_RAST <- c(testRast3[[1]], testRast3[[2]], testRast3[[3]], testRast3[[4]], testRast3[[5]], testRast3[[6]])

# make raster of test set 
 pred_glm1_1se_test_sf <- pred_glm1_1se_test %>% 
  st_as_sf(coords = c("x", "y"), crs = crs(crs_veg)) %>% 
  st_transform(crs(test_rast)) %>% 
  terra::vect()
# make a multi-layer raster w/ layers for each variable of interest
testRast_test2 <- lapply(c("TotalTreeCover", "TotalTreeCover_binom", "TotalTreeCover_interp", "TotalTreeCover_binom_pred_binary", "resid", "missClass"), FUN = function(x) {
  # S4 method for class 'data.frame'
 if (x %in% c("TotalTreeCover", "TotalTreeCover_binom", "TotalTreeCover_interp", "TotalTreeCover_binom_pred_binary", "resid")) {
    temp <- terra::rasterize(x = pred_glm1_1se_test_sf, y = test_rast, 
                         field = x, fun = mean, na.rm = TRUE)
    names(temp) <- x
  } else {
    temp <- terra::rasterize(x = pred_glm1_1se_test_sf, y = test_rast, 
                         field = x, fun = Mode, na.rm = TRUE)
    # get the % aggreement between values in the dayMetx8 cell 
    temp2 <- terra::rasterize(x = pred_glm1_1se_test_sf, y = test_rast$daymet_v4_prcp_monttl_na_1980_1, 
                         field = x, fun = function(y) {
                           yTemp <- y[!is.na(y)]
                           yMode <- Mode(yTemp)
                           (sum(yTemp == yMode)/length(yTemp))*100
                         })
    temp <- c(as.factor(temp$missClass), temp2)
    names(temp) <- c("missClass", "PercAgreement")
  }
    
 return(temp)
  }
)
pred_glm1_1se_test_RAST <- c(testRast_test2[[1]], testRast_test2[[2]], testRast_test2[[3]], testRast_test2[[4]], testRast_test2[[5]], testRast_test2[[6]])

# make figures of predictions w/ the full dataset
map_preds2_cont <- ggplot() +
geom_spatraster(data = pred_glm1_1se_full_RAST, aes(fill = TotalTreeCover_interp), maxcell = 20000000) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
labs(title = paste0("Predictions from the ", name_secondBestMod, " of Yes/No Trees \ncontinuous probabilities \nfull dataset")) +
  scale_fill_gradient2(low = "brown",
                       mid = "wheat" ,
                       high = "forestgreen" , 
                       midpoint = 0,   na.value = "white") + 
  coord_sf(datum = st_crs(test_rastTemp)) +
    theme(axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) 

# add inset
           inset_plot <-  ggplot(pred_glm1_1se_full) + 
             geom_density(aes(TotalTreeCover_interp)) +
                theme_bw(base_size=9) +  ## makes everything smaller
                theme(panel.background = element_rect(fill="transparent"),  ## white plot background
                      panel.border = element_blank(),
          axis.title.y = element_blank(),
          axis.title.x = element_blank(),
          axis.text.x = element_text(size=rel(2)), ## tiny axis text
          axis.text.y = element_text(size=rel(2)), ## tiny axis text
          axis.line.x = element_line(size = .5),
          axis.line.y = element_line(size = .5),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          plot.background = element_blank(),
          strip.background = element_blank(),
          strip.text = element_blank()) 
     
 map_preds2_cont <-    map_preds2_cont +   annotation_custom(grob = ggplotGrob(inset_plot), ymin = -2155000, ymax=-1409000, xmin=-2280000, xmax=-180000 ) 
 
# make figure of binomialized tree cover
map_preds2 <- ggplot() +
geom_spatraster(data = pred_glm1_1se_full_RAST, aes(fill = TotalTreeCover_binom_pred_binary), maxcell = 20000000) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
labs(title = paste0("Predictions from the ", name_secondBestMod, " of Yes/No Trees \nbinary probabilities \nfull dataset")) +
  scale_fill_continuous(breaks =  c(0, 0.5, 1), palette = c("wheat", "forestgreen"),
                    labels = c("No Trees",
                               "",
                      "Trees"), na.value = "white") + 
  coord_sf(datum = st_crs(test_rastTemp)) +
    theme(axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) 

# add inset
           inset_plot <-  ggplot(pred_glm1_1se_full) + 
             geom_density(aes(TotalTreeCover_binom_pred_binary)) +
                theme_bw(base_size=9) +  ## makes everything smaller
                theme(panel.background = element_rect(fill="transparent"),  ## white plot background
                      panel.border = element_blank(),
          axis.title.y = element_blank(),
          axis.title.x = element_blank(),
          axis.text.x = element_text(size=rel(2)), ## tiny axis text
          axis.text.y = element_text(size=rel(2)), ## tiny axis text
          axis.line.x = element_line(size = .5),
          axis.line.y = element_line(size = .5),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          plot.background = element_blank(),
          strip.background = element_blank(),
          strip.text = element_blank()) 
     
 map_preds2 <-    map_preds2 +   annotation_custom(grob = ggplotGrob(inset_plot), ymin = -2155000, ymax=-1409000, xmin=-2280000, xmax=-180000 ) 


# make figures of predictions w/ the test dataset
map_preds2_cont_test <- ggplot() +
geom_spatraster(data = pred_glm1_1se_test_RAST, aes(fill = TotalTreeCover_interp), maxcell = 20000000) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
labs(title = paste0("Predictions from the ", name_secondBestMod, " of Yes/No Trees \ncontinuous probabilities \ntest dataset")) +
  scale_fill_gradient2(low = "brown",
                       mid = "wheat" ,
                       high = "forestgreen" , 
                       midpoint = 0,   na.value = "white") +
  coord_sf(datum = st_crs(test_rastTemp)) +
    theme(axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) 

# add inset
           inset_plot <-  ggplot(pred_glm1_1se_test) + 
             geom_density(aes(TotalTreeCover_interp)) +
                theme_bw(base_size=9) +  ## makes everything smaller
                theme(panel.background = element_rect(fill="transparent"),  ## white plot background
                      panel.border = element_blank(),
          axis.title.y = element_blank(),
          axis.title.x = element_blank(),
          axis.text.x = element_text(size=rel(2)), ## tiny axis text
          axis.text.y = element_text(size=rel(2)), ## tiny axis text
          axis.line.x = element_line(size = .5),
          axis.line.y = element_line(size = .5),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          plot.background = element_blank(),
          strip.background = element_blank(),
          strip.text = element_blank()) 
     
 map_preds2_cont_test <-    map_preds2_cont_test +   annotation_custom(grob = ggplotGrob(inset_plot), ymin = -2155000, ymax=-1409000, xmin=-2280000, xmax=-180000 ) 

# make figure of binomialized tree cover
map_preds2_test <- ggplot() +
geom_spatraster(data = pred_glm1_1se_test_RAST, aes(fill = TotalTreeCover_binom_pred_binary), maxcell = 20000000) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
labs(title = paste0("Predictions from the ", name_secondBestMod, " of Yes/No Trees \nbinary probabilities \ntest dataset")) +
  scale_fill_continuous(breaks =  c(0, 0.5, 1), palette = c("wheat", "forestgreen"),
                    labels = c("No Trees",
                               "",
                      "Trees"), na.value = "white") + 
  coord_sf(datum = st_crs(test_rastTemp)) +
    theme(axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) 

# add inset
           inset_plot <-  ggplot(pred_glm1_1se_test) + 
             geom_density(aes(TotalTreeCover_binom_pred_binary)) +
                theme_bw(base_size=9) +  ## makes everything smaller
                theme(panel.background = element_rect(fill="transparent"),  ## white plot background
                      panel.border = element_blank(),
          axis.title.y = element_blank(),
          axis.title.x = element_blank(),
          axis.text.x = element_text(size=rel(2)), ## tiny axis text
          axis.text.y = element_text(size=rel(2)), ## tiny axis text
          axis.line.x = element_line(size = .5),
          axis.line.y = element_line(size = .5),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          plot.background = element_blank(),
          strip.background = element_blank(),
          strip.text = element_blank()) 
     
 map_preds2_test <-    map_preds2_test +   annotation_custom(grob = ggplotGrob(inset_plot), ymin = -2155000, ymax=-1409000, xmin=-2280000, xmax=-180000 ) 

ggarrange(ggarrange(map_preds2_cont, map_preds2, heights = c(1, 1), ncol = 1),
ggarrange(map_preds2_cont_test, map_preds2_test, heights = c(1,1), ncol = 1),
ncol = 1
)
}
```

### A map of classification error of predictions on the full and test datasets made using the *best lambda model*
```{r classification error map for best lambda model, fig.height = 10, fig.width = 9, warning = FALSE }
# true positive = tree (1) (forest green)
# true negative = no tree (3) (wheat) 
# false positive = model says tree, data says no-tree (2) (red)
# false negative = model says no tree, data says tree (4) (blue)
# make figures
map_missClass <- ggplot() +
geom_spatraster(data =pred_glm1_full_RAST
                , aes(fill= missClass), maxcell = 20000000) + 
  geom_sf(data=cropped_states_2,fill=NA )  + 
labs(title = paste0("Misclassification and correct classifications \nfrom the bestLambda model of Yes/No Trees \nfull dataset"),
     subtitle = "bestLambda model \nvalues show are the most common classification in each 8km x 8km grid cell") +
    scale_fill_discrete(palette = 
                        c("forestgreen", "red", "wheat", "blue"),
                        na.value = "white",
                        labels = c("forest", "Data = no trees; Model = trees", "non-forest", "Data = trees; Model = no trees", "")) +
  coord_sf(datum = st_crs(test_rastTemp)) +
    theme(axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) 

# make a map of the aggreement in classification w/in each 8x8km cell 
map_missClass_ag <- ggplot() +
  geom_spatraster(data = pred_glm1_full_RAST, aes(fill= PercAgreement), maxcell = 20000000) + 
  geom_sf(data=cropped_states_2,fill=NA )  + 
labs(title = paste0("% Agreement of the classification values within \neach 8x8km^2 gridcell classifications using \nthe bestLambda model of Yes/No Trees \nfull dataset"))+
    scale_fill_viridis_c(option = "B", na.value = "white", name = "% agreement", ) +
  coord_sf(datum = st_crs(test_rastTemp)) +
    theme(axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) 

ggarrange(map_missClass, map_missClass_ag, ncol = 1, heights = c(1, 1), align = "v")
```

```{r fig.width = 7, fig.height=7}
# make a confusion matrix 
# prepare data
matData <- pred_glm1_full %>% 
  mutate(predClass = TotalTreeCover_binom_pred_binary, 
         obsClass = TotalTreeCover_binom) %>% 
  mutate(predClass = replace(predClass, predClass == 0, "no trees"),
         predClass = replace(predClass, predClass == 1, "trees"),
         obsClass = replace(obsClass, obsClass == 0, "no trees"),
         obsClass = replace(obsClass, obsClass == 1, "trees")) %>% 
mutate(predClass = as.factor(predClass) ,
       obsClass = as.factor(obsClass))

# make matrix as a data.frame
# confMat <-  confusionMatrix(data = matData$predClass,
#                  reference = matData$obsClass)
ConfusionTableR::binary_visualiseR(
  train_labels = as.factor(matData$predClass),
  truth_labels = as.factor(matData$obsClass),
   class_label1 = "No trees",
   class_label2 = "Trees",
   quadrant_col1 = "wheat",
   quadrant_col2 = "forestgreen",
  text_col = "white", custom_title = "Confusion Matrix: classification of full dataset \nwith the best lambda model  using the full dataset"
)
```
```{r fig.height = 10, fig.width = 9, warning = FALSE }
### test dataset

# make figures
map_missClass_test <- ggplot() +
geom_spatraster(data =pred_glm1_test_RAST, aes(fill= missClass), maxcell = 20000000) + 
  geom_sf(data=cropped_states_2,fill=NA )  + 
labs(title = paste0("Misclassification and correct classifications \nfrom the bestLambda model of Yes/No Trees \ntest dataset"),
     subtitle = "bestLambda model \nvalues show are the most common classification in each 8km x 8km grid cell") +
    scale_fill_discrete(palette = 
                        c("forestgreen", "red", "wheat", "blue"),
                        na.value = "white",
                        labels = c("forest", "Data = no trees; Model = trees", "non-forest", "Data = trees; Model = no trees", "")) +
  coord_sf(datum = st_crs(test_rastTemp)) +
    theme(axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) 

map_missClass_test_ag <- ggplot() +
  geom_spatraster(data = pred_glm1_test_RAST, aes(fill= PercAgreement), maxcell = 20000000) + 
  geom_sf(data=cropped_states_2,fill=NA )  + 
labs(title = paste0("% Agreement of the classification values within \neach 8x8km^2 gridcell classifications using \nthe bestLambda model of Yes/No Trees \ntest dataset"))+
    scale_fill_viridis_c(option = "B", na.value = "white", name = "% agreement", ) +
  coord_sf(datum = st_crs(test_rastTemp)) +
    theme(axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) 
  

  ggarrange(map_missClass_test, map_missClass_test_ag, ncol = 1, heights = c(1, 1), align = "v")
```

```{r fig.width = 7, fig.height=6.5}
# make a confusion matrix 
# prepare data
matData_test <- pred_glm1_test %>% 
  mutate(predClass = TotalTreeCover_binom_pred_binary, 
         obsClass = TotalTreeCover_binom) %>% 
  mutate(predClass = replace(predClass, predClass == 0, "no trees"),
         predClass = replace(predClass, predClass == 1, "trees"),
         obsClass = replace(obsClass, obsClass == 0, "no trees"),
         obsClass = replace(obsClass, obsClass == 1, "trees")) %>% 
mutate(predClass = as.factor(predClass) ,
       obsClass = as.factor(obsClass))

# make matrix as a data.frame
# confMat <-  confusionMatrix(data = matData$predClass,
#                  reference = matData$obsClass)
ConfusionTableR::binary_visualiseR(
  train_labels = as.factor(matData_test$predClass),
  truth_labels = as.factor(matData_test$obsClass),
   class_label1 = "No trees",
   class_label2 = "Trees",
   quadrant_col1 = "wheat",
   quadrant_col2 = "forestgreen",
  text_col = "white", custom_title = "Confusion Matrix: classification of test dataset \nwith the best lambda model using the test dataset"
)
```

### A map of classification error of predictions on the full and test datasets made using the *second best lambda model*
```{r classification error map for second best lambda model, fig.height = 10, fig.width = 9, warning = FALSE }
### full dataset
# make figures
map_missClass <- ggplot() +
geom_spatraster(data =pred_glm1_1se_full_RAST, aes(fill= missClass), maxcell = 20000000) + 
  geom_sf(data=cropped_states_2,fill=NA )  + 
labs(title = paste0("Misclassification and correct classifications \nfrom the second-best lambda model of Yes/No Trees \nfull dataset"),
     subtitle = "second-best lambda model \nvalues show are the most common classification in each 8km x 8km grid cell") + 
  scale_fill_discrete(palette = 
                        c("forestgreen", "red", "wheat", "blue"),
                        na.value = "white",
                        labels = c("forest", "Data = no trees; Model = trees", "non-forest", "Data = trees; Model = no trees", "")) +
  coord_sf(datum = st_crs(test_rastTemp)) +
    theme(axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) 

map_missClass_ag <- ggplot() +
  geom_spatraster(data = pred_glm1_1se_full_RAST, aes(fill= PercAgreement), maxcell = 20000000) + 
  geom_sf(data=cropped_states_2,fill=NA )  + 
labs(title = paste0("% Agreement of the classification values within \neach 8x8km^2 gridcell classifications using \nthe second-best lambda model of Yes/No Trees \nfull dataset"))+
    scale_fill_viridis_c(option = "B", na.value = "white", name = "% agreement", ) +
  coord_sf(datum = st_crs(test_rastTemp)) +
    theme(axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) 
  
ggarrange(map_missClass, map_missClass_ag, ncol = 1, heights = c(1, 1), align = "v")
```

```{r fig.width = 7, fig.height=6.5}  
# make a confusion matrix 
# prepare data
matData <- pred_glm1_1se_full %>% 
  mutate(predClass = TotalTreeCover_binom_pred_binary, 
         obsClass = TotalTreeCover_binom) %>% 
  mutate(predClass = replace(predClass, predClass == 0, "no trees"),
         predClass = replace(predClass, predClass == 1, "trees"),
         obsClass = replace(obsClass, obsClass == 0, "no trees"),
         obsClass = replace(obsClass, obsClass == 1, "trees")) %>% 
mutate(predClass = as.factor(predClass) ,
       obsClass = as.factor(obsClass))

# make matrix as a data.frame
# confMat <-  confusionMatrix(data = matData$predClass,
#                  reference = matData$obsClass)
ConfusionTableR::binary_visualiseR(
  train_labels = as.factor(matData$predClass),
  truth_labels = as.factor(matData$obsClass),
   class_label1 = "No trees",
   class_label2 = "Trees",
   quadrant_col1 = "wheat",
   quadrant_col2 = "forestgreen",
  text_col = "white", custom_title = "Confusion Matrix: classification of full dataset \nwith the second-best lambda model using the full dataset"
)

```

```{r fig.height = 10, fig.width = 9, warning = FALSE }
### test dataset
# make figures
map_missClass_test <- ggplot() +
geom_spatraster(data =pred_glm1_1se_test_RAST, aes(fill= missClass), maxcell = 20000000) + 
  geom_sf(data=cropped_states_2,fill=NA )  + 
labs(title = paste0("Misclassification and correct classifications \nfrom the second-best lambda model of Yes/No Trees \ntest dataset"),
     subtitle = "second-best lambda model \nvalues show are the most common classification in each 8km x 8km grid cell") +
     scale_fill_discrete(palette = 
                        c("forestgreen", "red", "wheat", "blue"),
                        na.value = "white",
                        labels = c("forest", "Data = no trees; Model = trees", "non-forest", "Data = trees; Model = no trees", "")) +
  coord_sf(datum = st_crs(test_rastTemp)) +
    theme(axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) 

map_missClass_test_ag <- ggplot() +
  geom_spatraster(data = pred_glm1_1se_test_RAST, aes(fill= PercAgreement), maxcell = 20000000) + 
  geom_sf(data=cropped_states_2,fill=NA )  + 
labs(title = paste0("% Agreement of the classification values within \neach 8x8km^2 gridcell classifications using \nthe second-best lambda model of Yes/No Trees \ntest dataset"))+
    scale_fill_viridis_c(option = "B", na.value = "white", name = "% agreement", ) +
  coord_sf(datum = st_crs(test_rastTemp)) +
    theme(axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) 

  ggarrange(map_missClass_test, map_missClass_test_ag, ncol = 1, heights = c(1, 1), align = "v")
```
```{r fig.width = 7, fig.height=6.5}
# make a confusion matrix 
# prepare data
matData_test <- pred_glm1_1se_test %>% 
  mutate(predClass = TotalTreeCover_binom_pred_binary, 
         obsClass = TotalTreeCover_binom) %>% 
  mutate(predClass = replace(predClass, predClass == 0, "no trees"),
         predClass = replace(predClass, predClass == 1, "trees"),
         obsClass = replace(obsClass, obsClass == 0, "no trees"),
         obsClass = replace(obsClass, obsClass == 1, "trees")) %>% 
mutate(predClass = as.factor(predClass) ,
       obsClass = as.factor(obsClass))

# make matrix as a data.frame
# confMat <-  confusionMatrix(data = matData$predClass,
#                  reference = matData$obsClass)
ConfusionTableR::binary_visualiseR(
  train_labels = as.factor(matData_test$predClass),
  truth_labels = as.factor(matData_test$obsClass),
   class_label1 = "No trees",
   class_label2 = "Trees",
   quadrant_col1 = "wheat",
   quadrant_col2 = "forestgreen",
  text_col = "white", custom_title = "Confusion Matrix: classification of test dataset \nwith the second-best lambda model using the test dataset"
)
```

## Maps of wall-to-wall model predictions for locations that are *not* developed, agriculture, or water
```{r , fig.height = 16, fig.width = 9}
### best lambda model
## rasterize 
 pred_glm1_w2w_sf <- pred_glm1_w2w %>% 
  st_as_sf(coords = c("Lon", "Lat"), crs = crs(crs_veg)) %>% 
  st_transform(crs(test_rast)) %>% 
  terra::vect()
  # make a raster of the full dataset
# make a multi-layer raster w/ layers for each variable of interest
testRast6 <- lapply(c( "TotalTreeCover_interp", "TotalTreeCover_binom_pred_binary"), FUN = function(x) {
  # S4 method for class 'data.frame'
    temp <- terra::rasterize(x = pred_glm1_w2w_sf, y = test_rast, 
                         field = x, fun = mean, na.rm = TRUE)
    names(temp) <- x
 return(temp)
  }
)
pred_glm1_w2w_RAST <- c(testRast6[[1]], testRast6[[2]])

## map of continuous predictions
pred_glm1_w2w_cont<- ggplot() +
geom_spatraster(data = pred_glm1_w2w_RAST, aes(fill = TotalTreeCover_interp), maxcell = 20000000) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
labs(title = paste0("Predictions from the best lambda model of Yes/No Trees \ncontinuous probabilities on the wall-to-wall dayMet data")) +
  scale_fill_gradient2(low = "brown",
                       mid = "wheat" ,
                       high = "forestgreen" , 
                       midpoint = 0,   na.value = "white") +
  coord_sf(datum = st_crs(test_rastTemp)) +
    theme(axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) 

# add inset
           inset_plot <-  ggplot(pred_glm1_w2w) + 
             geom_density(aes(TotalTreeCover_interp)) +
                theme_bw(base_size=9) +  ## makes everything smaller
                theme(panel.background = element_rect(fill="transparent"),  ## white plot background
                      panel.border = element_blank(),
          axis.title.y = element_blank(),
          axis.title.x = element_blank(),
          axis.text.x = element_text(size=rel(2)), ## tiny axis text
          axis.text.y = element_text(size=rel(2)), ## tiny axis text
          axis.line.x = element_line(size = .5),
          axis.line.y = element_line(size = .5),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          plot.background = element_blank(),
          strip.background = element_blank(),
          strip.text = element_blank()) 
     
 pred_glm1_w2w_cont <-    pred_glm1_w2w_cont +   annotation_custom(grob = ggplotGrob(inset_plot), ymin = -2155000, ymax=-1409000, xmin=-2280000, xmax=-180000 ) 

## map of binary predictions
 pred_glm1_w2w_bin <- ggplot() +
geom_spatraster(data = pred_glm1_w2w_RAST, aes(fill = TotalTreeCover_binom_pred_binary), maxcell = 20000000) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
labs(title = paste0("Predictions from the best lambda model of Yes/No Trees \n binary probabilities on the wall-to-wall dayMet data")) +
  scale_fill_gradient2(low = "brown",
                       mid = "wheat" ,
                       high = "forestgreen" , 
                       midpoint = 0,   na.value = "white") +
  coord_sf(datum = st_crs(test_rastTemp)) +
    theme(axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) 

# add inset
           inset_plot <-  ggplot(pred_glm1_w2w) + 
             geom_density(aes(TotalTreeCover_binom_pred_binary)) +
                theme_bw(base_size=9) +  ## makes everything smaller
                theme(panel.background = element_rect(fill="transparent"),  ## white plot background
                      panel.border = element_blank(),
          axis.title.y = element_blank(),
          axis.title.x = element_blank(),
          axis.text.x = element_text(size=rel(2)), ## tiny axis text
          axis.text.y = element_text(size=rel(2)), ## tiny axis text
          axis.line.x = element_line(size = .5),
          axis.line.y = element_line(size = .5),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          plot.background = element_blank(),
          strip.background = element_blank(),
          strip.text = element_blank()) 
     
 pred_glm1_w2w_bin <-    pred_glm1_w2w_bin +   annotation_custom(grob = ggplotGrob(inset_plot), ymin = -2155000, ymax=-1409000, xmin=-2280000, xmax=-180000 ) 


### second best lambda model
## rasterize 
 pred_glm1_1se_w2w_sf <- pred_glm1_1se_w2w %>% 
  st_as_sf(coords = c("Lon", "Lat"), crs = crs(crs_veg)) %>% 
  st_transform(crs(test_rast)) %>% 
  terra::vect()
  # make a raster of the full dataset
# make a multi-layer raster w/ layers for each variable of interest
testRast7 <- lapply(c( "TotalTreeCover_interp", "TotalTreeCover_binom_pred_binary"), FUN = function(x) {
  # S4 method for class 'data.frame'
    temp <- terra::rasterize(x = pred_glm1_1se_w2w_sf, y = test_rast, 
                         field = x, fun = mean, na.rm = TRUE)
    names(temp) <- x
 return(temp)
  }
)
pred_glm1_1se_w2w_RAST <- c(testRast7[[1]], testRast7[[2]])

## map of continuous predictions
pred_glm1_1se_w2w_cont<- ggplot() +
geom_spatraster(data = pred_glm1_1se_w2w_RAST, aes(fill = TotalTreeCover_interp), maxcell = 20000000) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
labs(title = paste0("Predictions from the second-best lambda model of Yes/No Trees \ncontinuous probabilities on the wall-to-wall dayMet data")) +
  scale_fill_gradient2(low = "brown",
                       mid = "wheat" ,
                       high = "forestgreen" , 
                       midpoint = 0,   na.value = "white") +
  coord_sf(datum = st_crs(test_rastTemp)) +
    theme(axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) 

# add inset
           inset_plot <-  ggplot(pred_glm1_1se_w2w) + 
             geom_density(aes(TotalTreeCover_interp)) +
                theme_bw(base_size=9) +  ## makes everything smaller
                theme(panel.background = element_rect(fill="transparent"),  ## white plot background
                      panel.border = element_blank(),
          axis.title.y = element_blank(),
          axis.title.x = element_blank(),
          axis.text.x = element_text(size=rel(2)), ## tiny axis text
          axis.text.y = element_text(size=rel(2)), ## tiny axis text
          axis.line.x = element_line(size = .5),
          axis.line.y = element_line(size = .5),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          plot.background = element_blank(),
          strip.background = element_blank(),
          strip.text = element_blank()) 
     
 pred_glm1_1se_w2w_cont <-    pred_glm1_1se_w2w_cont +   annotation_custom(grob = ggplotGrob(inset_plot), ymin = -2155000, ymax=-1409000, xmin=-2280000, xmax=-180000 ) 

## map of binary predictions
 pred_glm1_1se_w2w_bin <- ggplot() +
geom_spatraster(data = pred_glm1_1se_w2w_RAST, aes(fill = TotalTreeCover_binom_pred_binary), maxcell = 20000000) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
labs(title = paste0("Predictions from the second-best lambda model of Yes/No Trees \n binary probabilities on the wall-to-wall dayMet data")) +
  scale_fill_gradient2(low = "brown",
                       mid = "wheat" ,
                       high = "forestgreen" , 
                       midpoint = 0,   na.value = "white") +
  coord_sf(datum = st_crs(test_rastTemp)) +
    theme(axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) 

# add inset
           inset_plot <-  ggplot(pred_glm1_1se_w2w) + 
             geom_density(aes(TotalTreeCover_binom_pred_binary)) +
                theme_bw(base_size=9) +  ## makes everything smaller
                theme(panel.background = element_rect(fill="transparent"),  ## white plot background
                      panel.border = element_blank(),
          axis.title.y = element_blank(),
          axis.title.x = element_blank(),
          axis.text.x = element_text(size=rel(2)), ## tiny axis text
          axis.text.y = element_text(size=rel(2)), ## tiny axis text
          axis.line.x = element_line(size = .5),
          axis.line.y = element_line(size = .5),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          plot.background = element_blank(),
          strip.background = element_blank(),
          strip.text = element_blank()) 
     
 pred_glm1_1se_w2w_bin <-   pred_glm1_1se_w2w_bin +   annotation_custom(grob = ggplotGrob(inset_plot), ymin = -2155000, ymax=-1409000, xmin=-2280000, xmax=-180000 ) 
 
ggarrange(pred_glm1_w2w_cont, pred_glm1_w2w_bin, heights = c(1,1), ncol = 1)
ggarrange(pred_glm1_1se_w2w_cont, pred_glm1_1se_w2w_bin, heights = c(1,1), ncol = 1)
```

### Are the locations that the model predicts to be un-forested (i.e. the continuous P(Forest) == 0) similar to the locations that are unforested based on Martin's data derived from RAP?
```{r, fig.height = 16, fig.width = 9}
## map of model-predicted zeros 
pred_glm1_w2w_ZeroRast <- pred_glm1_w2w_RAST$TotalTreeCover_interp

pred_glm1_w2w_ZeroRast[pred_glm1_w2w_ZeroRast != 0] <- NA

pred_glm1_w2w_Zeros <- ggplot() +
geom_spatraster(data = pred_glm1_w2w_ZeroRast, aes(fill = TotalTreeCover_interp), maxcell = 20000000) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
labs(title = paste0("Predictions of No Forest from the best lambda model of Yes/No Trees \ncontinuous probabilities on the wall-to-wall dayMet data")) +
  scale_fill_continuous(na.value = "white") +
  # scale_fill_gradient2(low = "brown",
  #                      mid = "wheat" ,
  #                      high = "forestgreen" , 
  #                      midpoint = 0,   na.value = "white") +
  coord_sf(datum = st_crs(test_rastTemp)) +
    theme(axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) 


## map of zeros from Martin
notForest_ZeroRast <- notForest2

notForest_ZeroRast[notForest_ZeroRast <=0.5 ] <- NA

notForest_ZeroRast_Zeros <- ggplot() +
geom_spatraster(data = notForest_ZeroRast, aes(fill = fracNotForest), maxcell = 20000000) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
labs(title = paste0("Locations where forested area from RAP indicates that \n >50% of a dayMet grid cell is unforested")) +
  scale_fill_continuous(na.value = "white") +
  # scale_fill_gradient2(low = "brown",
  #                      mid = "wheat" ,
  #                      high = "forestgreen" , 
  #                      midpoint = 0,   na.value = "white") +
  coord_sf(datum = st_crs(test_rastTemp)) +
    theme(axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) 


ggarrange(pred_glm1_w2w_Zeros, notForest_ZeroRast_Zeros, ncol = 1)
```


### Are there biases of the model predictions across year/lat/long? (for the full dataset)
```{r Misclassification figures best lambda model, fig.align=8, fig.height=10}
if ( name_secondBestMod == "Intercept_Only") {
  print("The next best lambda model only contains one predictor (an intercept)")
  
  ## calculate Misclassification in the input pred_glm1 or pred_glm1se datasets
  pred_glm1_full$missClass2 <- pred_glm1_full$TotalTreeCover_binom - pred_glm1_full$TotalTreeCover_binom_pred_binary
  pred_glm1_test$missClass2 <- pred_glm1_test$TotalTreeCover_binom - pred_glm1_test$TotalTreeCover_binom_pred_binary

  # plot misclassifications against Year
yearResidMod_bestLambda <- ggplot(pred_glm1_full) + 
  geom_point(aes(x = jitter(Year), y = jitter(missClass2)), alpha = .1) + 
  geom_hline(aes(yintercept = 0), lty = 2, col = "red") + 
  geom_smooth(aes(x = Year, y = missClass2), method = "lm") + 
  xlab("Year") + 
  ylab("misclassifications") +
  ggtitle("from best lamba model")

# plot misclassifications against Lat
latResidMod_bestLambda <- ggplot(pred_glm1_full) + 
  geom_point(aes(x = y, y = jitter(missClass2)), alpha = .1) + 
  geom_hline(aes(yintercept = 0), lty = 2, col = "red") + 
  geom_smooth(aes(x = y, y = missClass2)) + 
  xlab("Latitude") + 
  ylab("misclassifications") +
  ggtitle("from best lamba model")

# plot misclassifications against Long
longResidMod_bestLambda <- ggplot(pred_glm1_full) + 
  geom_point(aes(x = x, y = jitter(missClass2)), alpha = .1) + 
  geom_hline(aes(yintercept = 0), lty = 2, col = "red") + 
  geom_smooth(aes(x = x, y = missClass2)) + 
  xlab("Longitude") + 
  ylab("misclassifications") +
  ggtitle("from best lamba model")
library(patchwork)
(yearResidMod_bestLambda ) / 
(  latResidMod_bestLambda ) /
(  longResidMod_bestLambda )
} else {
  
  ## calculate missClass2ification in the input pred_glm1 or pred_glm1se datasets
  pred_glm1_full$missClass2 <- pred_glm1_full$TotalTreeCover_binom - pred_glm1_full$TotalTreeCover_binom_pred_binary
  pred_glm1_test$missClass2 <- pred_glm1_test$TotalTreeCover_binom - pred_glm1_test$TotalTreeCover_binom_pred_binary
  pred_glm1_1se_full$missClass2 <- pred_glm1_1se_full$TotalTreeCover_binom - pred_glm1_1se_full$TotalTreeCover_binom_pred_binary
  pred_glm1_1se_test$missClass2 <- pred_glm1_1se_test$TotalTreeCover_binom - pred_glm1_1se_test$TotalTreeCover_binom_pred_binary

# plot misclassifications against Year
yearResidMod_bestLambda <- ggplot(pred_glm1_full) + 
  geom_point(aes(x = jitter(Year), y = jitter(missClass2)), alpha = .1) + 
  geom_hline(aes(yintercept = 0), lty = 2, col = "red") + 
  geom_smooth(aes(x = Year, y = missClass2), method = "lm") + 
  xlab("Year") + 
  ylab("misclassifications") +
  ggtitle("from best lamba model")
yearResidMod_1seLambda <- ggplot(pred_glm1_1se_full) + 
  geom_point(aes(x = jitter(Year), y = jitter(missClass2)), alpha = .1) + 
  geom_hline(aes(yintercept = 0), lty = 2, col = "red") + 
  geom_smooth(aes(x = Year, y = missClass2), method = "lm") + 
  xlab("Year") + 
  ylab("misclassifications") +
  ggtitle(paste0("from ", name_secondBestMod))

# plot misclassifications against Lat
latResidMod_bestLambda <- ggplot(pred_glm1_full) + 
  geom_point(aes(x = y, y = jitter(missClass2)), alpha = .1) + 
  geom_hline(aes(yintercept = 0), lty = 2, col = "red") + 
  geom_smooth(aes(x = y, y = missClass2)) + 
  xlab("Latitude") + 
  ylab("misclassifications") +
  ggtitle("from best lamba model")
latResidMod_1seLambda <- ggplot(pred_glm1_1se_full) + 
  geom_point(aes(x = y, y = jitter(missClass2)), alpha = .1) + 
  geom_hline(aes(yintercept = 0), lty = 2, col = "red") + 
  geom_smooth(aes(x = y, y = missClass2)) + 
  xlab("Latitude") + 
  ylab("misclassifications") +
  ggtitle(paste0("from ", name_secondBestMod))

# plot misclassifications against Long
longResidMod_bestLambda <- ggplot(pred_glm1_full) + 
  geom_point(aes(x = x, y = jitter(missClass2)), alpha = .1) + 
  geom_hline(aes(yintercept = 0), lty = 2, col = "red") + 
  geom_smooth(aes(x = x, y = missClass2)) + 
  xlab("Longitude") + 
  ylab("misclassifications") +
  ggtitle("from best lamba model")
longResidMod_1seLambda <- ggplot(pred_glm1_1se_full) + 
  geom_point(aes(x = x, y = jitter(missClass2)), alpha = .1) + 
  geom_hline(aes(yintercept = 0), lty = 2, col = "red") + 
  geom_smooth(aes(x = x, y = missClass2)) + 
  xlab("Longitude") + 
  ylab("misclassifications") +
  ggtitle(paste0("from ", name_secondBestMod))

library(patchwork)
(yearResidMod_bestLambda + yearResidMod_1seLambda) / 
(  latResidMod_bestLambda + latResidMod_1seLambda) /
(  longResidMod_bestLambda + longResidMod_1seLambda)
}
```

<!-- ### Quantile plots - comparing "binary" observations across the full dataset to continuous, interpolated model-predicted probabilities -->

<!-- Binning predictor variables into "Quantiles"and looking at the mean -->
<!-- predicted probability for each quantile  -->
<!-- ```{r} -->
<!-- response_vars <- c("TotalTreeCover_binom", "TotalTreeCover_interp") -->
<!-- # get deciles for best lambda model  -->
<!-- if (length(prednames_fig) == 0) { -->
<!--   print("The best lambda model only contains one predictor (an intercept), so decile plots aren't possible to generate") -->
<!-- } else { -->
<!--   pred_glm1_deciles <- predvars2deciles(pred_glm1_full, -->
<!--                                       response_vars = response_vars, -->
<!--                                         pred_vars = prednames_fig,  -->
<!--                                        cut_points = seq(0, 1, 0.01)) -->
<!-- } -->
<!-- # get deciles for 1 SE lambda model  -->
<!-- if ( name_secondBestMod == "Intercept_Only") { -->
<!--   print("The next best lambda model only contains one predictor (an intercept)")} else { -->
<!--   pred_glm1_deciles_1se <- predvars2deciles(pred_glm1_1se_full, -->
<!--                                       response_vars = response_vars, -->
<!--                                         pred_vars = prednames_secondBestMod,  -->
<!--                                        cut_points = seq(0, 1, 0.01)) -->
<!--   } -->
<!-- ``` -->

<!-- Below are quantile plots for the *best lambda model*  -->
<!-- (note that the predictor variables are scaled)  -->
<!-- ```{r Quantile best lambda, fig.height= 16, fig.width = 12, warning=FALSE, message=FALSE} -->
<!-- if (length(prednames_fig) == 0) { -->
<!--   print("The model only contains one predictor (an intercept), so decile plots aren't possible to generate") -->
<!-- } else { -->

<!-- # publication quality version -->
<!-- g3 <- decile_dotplot_pq(df = pred_glm1_deciles, response = c("TotalTreeCover_binom", "TotalTreeCover_interp"), IQR = FALSE, -->
<!--                         CI = FALSE -->
<!--                         ) + ggtitle("Decile Plot") -->

<!-- g4 <- add_dotplot_inset(g3, df = pred_glm1_deciles, response = c("TotalTreeCover_binom", "TotalTreeCover_interp"), dfRaw = pred_glm1_full, add_smooth = TRUE, deciles = FALSE) -->


<!-- if(save_figs) { -->
<!--   # png(paste0("../../../Figures/CoverDatFigures/ figures/quantile_plots/quantile_plot_", response,  "_",ecoregion,".png"),  -->
<!--   #    units = "in", res = 600, width = 5.5, height = 3.5 ) -->
<!--   #   print(g4) -->
<!--   # dev.off() -->
<!-- } -->

<!-- g4 -->
<!-- } -->
<!-- ``` -->

<!-- Below are quantile plots from the second best lambda model (`r print(name_secondBestMod)`) -->
<!-- ```{r quantile_plot 1 SE lambda, fig.height= 16, fig.width = 12, warning=FALSE, message=FALSE} -->

<!-- if ( name_secondBestMod == "Intercept_Only") { -->
<!--   print("The next best lambda model only contains one predictor (an intercept)") -->
<!--   } else { -->

<!-- # publication quality version -->
<!-- g3 <- decile_dotplot_pq(pred_glm1_deciles_1se, response = c("TotalTreeCover_binom", "TotalTreeCover_interp"), IQR = FALSE) + ggtitle("Decile Plot") -->

<!-- g4 <- add_dotplot_inset(g3, df = pred_glm1_deciles_1se, response = c("TotalTreeCover_binom", "TotalTreeCover_interp"), dfRaw = pred_glm1_1se_full, add_smooth = TRUE, deciles = FALSE) -->


<!-- if(save_figs) { -->
<!--   # png(paste0("figures/quantile_plots/quantile_plot_", response,  "_",ecoregion,".png"),  -->
<!--   #    units = "in", res = 600, width = 5.5, height = 3.5 ) -->
<!--   #   print(g4) -->
<!--   # dev.off() -->
<!-- } -->

<!-- g4 -->
<!-- } -->

<!-- ``` -->

<!-- ### Filtered Quantiles -  comparing "binary" observations in the testing dataset to continuous model-predicted probabilities -->
<!-- #### For the best lambda model -->

<!-- Filtered quantile plots of data. These plots show each vegetation variable, -->
<!-- but only based on data that falls into the upper and lower 20th percentiles of -->
<!-- each predictor variable.  -->

<!-- ```{r glm_deciles_filtered best lambda model, fig.height = 24, fig.width = 24, message = FALSE} -->
<!-- if (length(prednames_fig) == 0) { -->
<!--   print("The model only contains one predictor (an intercept), so decile plots aren't possible to generate") -->
<!-- } else { -->
<!-- pred_glm1_deciles_filt <- predvars2deciles( pred_glm1_full,  -->
<!--                          response_vars = c("TotalTreeCover_binom", "TotalTreeCover_interp"), -->
<!--                          pred_vars = prednames_fig, -->
<!--                          filter_var = TRUE, -->
<!--                          filter_vars = prednames_fig, -->
<!--                          cut_points = seq(0, 1, 0.01))  -->
<!-- g <- decile_dotplot_filtered_pq(df = pred_glm1_deciles_filt, response = "TotalTreeCover_binom",  -->
<!--                                 xvars = prednames_fig) -->

<!-- if(save_figs) { -->
<!-- # jpeg(paste0("figures/quantile_plots/quantile_plot_filtered_mid_v1", , ".jpeg"), -->
<!-- #      units = "in", res = 600, width = 5.5, height = 6 ) -->
<!-- #   g  -->
<!-- # dev.off() -->
<!-- } -->
<!-- g -->
<!-- } -->
<!-- ``` -->

<!-- Filtered quantile figure with middle 2 deciles also shown -->

<!-- ```{r filtered quantile plot for best lambda model with middle 2 deciles, fig.height = 24, fig.width = 24, eval = TRUE, message = FALSE} -->
<!-- if ( name_secondBestMod == "Intercept_Only") { -->
<!--   print("The next best lambda model only contains one predictor (an intercept)") -->
<!-- } else { -->
<!-- pred_glm1_deciles_filt_mid <- predvars2deciles(pred_glm1_full,  -->
<!--                          response_vars = c("TotalTreeCover_binom", "TotalTreeCover_interp"), -->
<!--                          pred_vars = prednames_fig, -->
<!--                          filter_vars = prednames_fig, -->
<!--                          filter_var = TRUE, -->
<!--                          add_mid = TRUE, -->
<!--                          cut_points = seq(0, 1, 0.01)) -->

<!-- g <- decile_dotplot_filtered_pq(df = pred_glm1_deciles_filt_mid, response = "TotalTreeCover_binom",  -->
<!--                                 xvars = prednames_fig) -->

<!-- if(save_figs) { -->
<!-- # jpeg(paste0("figures/quantile_plots/quantile_plot_filtered_mid_v1", , ".jpeg"), -->
<!-- #      units = "in", res = 600, width = 5.5, height = 6) -->
<!-- #   g  -->
<!-- # dev.off() -->
<!-- } -->
<!-- g -->
<!-- } -->
<!-- ``` -->

<!-- #### For the second best lambda model (`r print(name_secondBestMod)`) -->

<!-- Filtered 'Quantile' plots of data. These plots show each vegetation variable, -->
<!-- but only based on data that falls into the upper and lower 20th percentiles of -->
<!-- each predictor variable.  -->

<!-- ```{r glm_deciles_filtered second best lambda model, fig.height = 24, fig.width = 24, message = FALSE} -->
<!-- if (length(prednames_secondBestMod) == 0) { -->
<!--   print("The model only contains one predictor (an intercept), so decile plots aren't possible to generate") -->
<!-- } else { -->

<!-- pred_glm1_deciles_filt <- predvars2deciles( pred_glm1_1se_full,  -->
<!--                          response_vars = c("TotalTreeCover_binom", "TotalTreeCover_interp"), -->
<!--                          pred_vars = prednames_secondBestMod, -->
<!--                          filter_var = TRUE, -->
<!--                          filter_vars = prednames_secondBestMod, -->
<!--                          cut_points = seq(0, 1, 0.01))  -->
<!-- g <- decile_dotplot_filtered_pq(df = pred_glm1_deciles_filt, response = "TotalTreeCover_binom",  -->
<!--                                 xvars = prednames_fig) -->

<!-- if(save_figs) { -->
<!-- # jpeg(paste0("figures/quantile_plots/quantile_plot_filtered_mid_v1", , ".jpeg"), -->
<!-- #      units = "in", res = 600, width = 5.5, height = 6 ) -->
<!-- #   g  -->
<!-- # dev.off() -->
<!-- } -->
<!-- g -->
<!-- } -->
<!-- ``` -->

<!-- Filtered quantile figure with middle 2 deciles also shown -->

<!-- ```{r filtered quantile plot for second best lambda model with middle 2 deciles, fig.height = 24, fig.width = 24, eval = TRUE, message = FALSE} -->
<!-- if (length(prednames_secondBestMod) == 0) { -->
<!--   print("The model only contains one predictor (an intercept), so decile plots aren't possible to generate") -->
<!-- } else { -->
<!-- pred_glm1_deciles_filt_mid <- predvars2deciles(pred_glm1_1se_full,  -->
<!--                          response_vars = c("TotalTreeCover_binom", "TotalTreeCover_interp"), -->
<!--                          pred_vars = prednames_secondBestMod, -->
<!--                          filter_vars = prednames_secondBestMod, -->
<!--                          filter_var = TRUE, -->
<!--                          add_mid = TRUE, -->
<!--                          cut_points = seq(0, 1, 0.01)) -->

<!-- g <- decile_dotplot_filtered_pq(df = pred_glm1_deciles_filt_mid, response = "TotalTreeCover_binom",  -->
<!--                                 xvars = prednames_fig) -->


<!-- if(save_figs) { -->
<!-- # jpeg(paste0("figures/quantile_plots/quantile_plot_filtered_mid_v1", , ".jpeg"), -->
<!-- #      units = "in", res = 600, width = 5.5, height = 6 ) -->
<!-- #   g  -->
<!-- # dev.off() -->
<!-- } -->
<!-- g -->
<!-- } -->
<!-- ``` -->



### Quantile plots - comparing "binary" testing data to "binary" model-predictions (i.e. converting the continuous model predictions to discrete values (1 or 0) based on the previously-identified optimal threshold)

```{r}
response_vars <- c("TotalTreeCover_binom", "TotalTreeCover_binom_pred_binary")
# get deciles for best lambda model 
if (length(prednames_fig) == 0) {
  print("The best lambda model only contains one predictor (an intercept), so decile plots aren't possible to generate")
} else {
  pred_glm1_deciles <- predvars2deciles(pred_glm1_full,
                                      response_vars = response_vars,
                                        pred_vars = prednames_fig, 
                                       cut_points = seq(0, 1, 0.01))
}
# get deciles for 1 SE lambda model 
if ( name_secondBestMod == "Intercept_Only") {
  print("The next best lambda model only contains one predictor (an intercept)")} else {
  pred_glm1_deciles_1se <- predvars2deciles(pred_glm1_1se_full,
                                      response_vars = response_vars,
                                        pred_vars = prednames_secondBestMod, 
                                       cut_points = seq(0, 1, 0.01))
  }
```

Below are quantile plots for the *best lambda model* 
(note that the predictor variables are scaled) 
```{r quantile_plot best lambda binary response, fig.height= 16, fig.width = 12, warning=FALSE, message=FALSE}
if (length(prednames_fig) == 0) {
  print("The model only contains one predictor (an intercept), so decile plots aren't possible to generate")
} else {
# ####  code to generate plots analagous to "decile plots" by hand, do double check that they're suitable for use w/ binary predictions 
#   pred_glm1 %>% 
#     #slice_sample(n = 10000) %>% 
#     pivot_longer(cols = all_of(prednames_fig)) %>% 
#     select(Year, x, y, TotalTreeCover_binom, TotalTreeCover_binom_pred_binary, name, value) %>% 
#     ggplot() + 
#     facet_wrap(~name, scales = "free") + 
#     geom_point(aes(x = value, y = jitter(TotalTreeCover_binom)), alpha = .5) + 
#     geom_point(aes(x = value, y = jitter(TotalTreeCover_binom_pred_binary)), col = "blue", alpha = .5) + 
#     geom_smooth(aes(x = value, y = TotalTreeCover_binom), col = "black") + 
#     geom_smooth(aes(x = value, y = TotalTreeCover_binom_pred_binary), col = "blue") 
#                  
  
  
  #####
# publication quality version
g3 <- decile_dotplot_pq(df = pred_glm1_deciles, response = c("TotalTreeCover_binom", paste0("TotalTreeCover_binom", "_pred_binary")), IQR = FALSE,
                        CI = FALSE
                        ) + ggtitle("Decile Plot")

g4 <- add_dotplot_inset(g3, df = pred_glm1_deciles, response = c("TotalTreeCover_binom", paste0("TotalTreeCover_binom", "_pred_binary")), dfRaw = pred_glm1_full, add_smooth = TRUE, deciles = FALSE)

  
if(save_figs) {
  # png(paste0("../../../Figures/CoverDatFigures/ figures/quantile_plots/quantile_plot_", response,  "_",ecoregion,".png"), 
  #    units = "in", res = 600, width = 5.5, height = 3.5 )
  #   print(g4)
  # dev.off()
}

g4
}
```

Below are quantile plots from the second best lambda model (not that the predictor variables are scaled)
```{r quantile_plot 1 SE lambda binary response, fig.height= 16, fig.width = 12, warning=FALSE, message=FALSE}

if ( name_secondBestMod == "Intercept_Only") {
  print("The next best lambda model only contains one predictor (an intercept)")
  } else {

# publication quality version
g3 <- decile_dotplot_pq(pred_glm1_deciles_1se, response = c("TotalTreeCover_binom", paste0("TotalTreeCover_binom", "_pred_binary")), IQR = FALSE) + ggtitle("Decile Plot")

g4 <- add_dotplot_inset(g3, df = pred_glm1_deciles_1se, response = c("TotalTreeCover_binom", paste0("TotalTreeCover_binom", "_pred_binary")), dfRaw = pred_glm1_1se_full, add_smooth = TRUE, deciles = FALSE)

  
if(save_figs) {
  # png(paste0("figures/quantile_plots/quantile_plot_", response,  "_",ecoregion,".png"), 
  #    units = "in", res = 600, width = 5.5, height = 3.5 )
  #   print(g4)
  # dev.off()
}

g4
}

```

### Show model RMSE w/in each quantile
```{r}
# get deciles for best lambda model 
if (length(prednames_fig) == 0) {
  print("The best lambda model only contains one predictor (an intercept), so decile plots aren't possible to generate")
} else {
  pred_glm1_deciles %>% 
    ggplot(aes(x = mean_value, y = RMSE)) +
    facet_wrap(~name, scales = "free_x") +
    geom_point(alpha = .2, size = .5) + 
    geom_smooth(lwd = .5) + 
    xlab("Scaled predictor value") + 
    ggtitle("RMSE by decile for bestLambda model")
}
# get deciles for 1 SE lambda model 
if (length(prednames_secondBestMod) == 0) {
  print("The 1SE (or 1/2 SE) lambda model only contains one predictor (an intercept), so decile plots aren't possible to generate")
} else {
  pred_glm1_deciles_1se %>% 
    ggplot(aes(x = mean_value, y = RMSE)) +
    facet_wrap(~name, scales = "free_x") +
    geom_point(alpha = .2, size = .5) + 
    geom_smooth(lwd = .5) + 
    xlab("Scaled predictor value") + 
    ggtitle(paste0("RMSE by decile for ", name_secondBestMod, "model"))
}
```

## Equations for final models 
```{r}
# get the parameters that were used to scale the data
scaleParams <- modDat_1_s_temp %>% 
  filter(Year == 2016) %>% 
  dplyr::select(tmin_s:AWHC_s) %>% 
  reframe(across(all_of(names(.)), attributes)) 

# function for printing model statement
## function to get model statements
getModelStatement <- function(coefficientTable, # name of the d.f that has model coefficients
                  modelName, # name of the column in the coefficient table that has the parameters of interest
                  responseVar # name of the response variable
                  ) {
  ##
  # coefficientTable <- grassShrub_totalHerb_trimAnoms
  # modelName <- "coefficientValue_bestLambda"
  # responseVar <- "TotalHerbaceousCover"
  ##
  temp <- coefficientTable[,c("coefficientName", modelName)] %>% 
    drop_na()
  rownames(temp) <- temp$coefficientName
  temp[,modelName] <- round(temp[,modelName], 8)
  
  # print out coefficients w/ coefficient names
tempNames <- paste0(
  apply(temp, MARGIN = 1, FUN = function(x) {
    if (x["coefficientName"] == "(Intercept)") {
      paste0(x[modelName])
    } else {  
      paste0(x[modelName], "*", x["coefficientName"])
    }
    }
  ),
  collapse = " + "
)

# print the unscaled model statement
  unscaledModelName <- paste0(responseVar, "~ 1/(1 + exp(-(", tempNames, ")))")
  
  # now add in the scale parameters
tempNames <- str_replace_all(tempNames, pattern = "annWetDegDays_anom", 
                         replacement = paste0("((annWetDegDays_anom - ", 
                                              round(scaleParams$annWetDegDays_anom_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$annWetDegDays_anom_s$`scaled:scale`,8), ")"))

tempNames <- str_replace_all(tempNames, pattern = "prcpTempCorr_anom", 
                         replacement = paste0("((prcpTempCorr_anom - ", 
                                              round(scaleParams$prcpTempCorr_anom_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$prcpTempCorr_anom_s$`scaled:scale`,8), ")"))


tempNames <- str_replace_all(tempNames, pattern = "prcp_seasonality_anom", 
                         replacement = paste0("((prcp_seasonality_anom - ", 
                                              round(scaleParams$prcp_seasonality_anom_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$prcp_seasonality_anom_s$`scaled:scale`,8), ") "))

tempNames <- str_replace_all(tempNames, pattern = "annWatDef_anom", 
                         replacement = paste0("((annWatDef_anom - ", 
                                              round(scaleParams$annWatDef_anom_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$annWatDef_anom_s$`scaled:scale`,8), ") "))

tempNames <- str_replace_all(tempNames, pattern = "isothermality_anom", 
                         replacement = paste0("((isothermality_anom - ", 
                                              round(scaleParams$isothermality_anom_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$isothermality_anom_s$`scaled:scale`,8), ") "))

tempNames <- str_replace_all(tempNames, pattern = "prcp_anom", 
                         replacement = paste0("((prcp_anom - ", 
                                              round(scaleParams$prcp_anom_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$prcp_anom_s$`scaled:scale`,8), ") "))

tempNames <- str_replace_all(tempNames, pattern = "prcp ", 
                         replacement = paste0("((prcp - ", 
                                              round(scaleParams$prcp_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$prcp_s$`scaled:scale`,8), ") "))

tempNames <- str_replace_all(tempNames, pattern = "prcp\\^", 
                         replacement = paste0("((prcp - ", 
                                              round(scaleParams$prcp_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$prcp_s$`scaled:scale`,8), ")^"))

tempNames <- str_replace_all(tempNames, pattern = "prcp:", 
                         replacement = paste0("((prcp - ", 
                                              round(scaleParams$prcp_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$prcp_s$`scaled:scale`,8), "):"))

tempNames <- str_replace_all(tempNames, pattern = "prcpTempCorr ", 
                         replacement = paste0("((prcpTempCorr - ", 
                                              round(scaleParams$prcpTempCorr_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$prcpTempCorr_s$`scaled:scale`,8), ") "))

tempNames <- str_replace_all(tempNames, pattern = "prcpTempCorr\\^", 
                         replacement = paste0("((prcpTempCorr - ", 
                                              round(scaleParams$prcpTempCorr_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$prcpTempCorr_s$`scaled:scale`,8), ")^"))

tempNames <- str_replace_all(tempNames, pattern = "prcpTempCorr:", 
                         replacement = paste0("((prcpTempCorr - ", 
                                              round(scaleParams$prcpTempCorr_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$prcpTempCorr_s$`scaled:scale`,8), "):"))

tempNames <- str_replace_all(tempNames, pattern = "prcpTempCorr\\)", 
                         replacement = paste0("((prcpTempCorr - ", 
                                              round(scaleParams$prcpTempCorr_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$prcpTempCorr_s$`scaled:scale`,8), "))"))

tempNames <- str_replace_all(tempNames, pattern = "prcpTempCorr$", 
                         replacement = paste0("((prcpTempCorr - ", 
                                              round(scaleParams$prcpTempCorr_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$prcpTempCorr_s$`scaled:scale`,8), ")"))

tempNames <- str_replace_all(tempNames, pattern = "isothermality ", 
                         replacement = paste0("((isothermality - ", 
                                              round(scaleParams$isothermality_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$isothermality_s$`scaled:scale`,8), ") "))

tempNames <- str_replace_all(tempNames, pattern = "isothermality\\^", 
                         replacement = paste0("((isothermality - ", 
                                              round(scaleParams$isothermality_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$isothermality_s$`scaled:scale`,8), ")^"))

tempNames <- str_replace_all(tempNames, pattern = "isothermality:", 
                         replacement = paste0("((isothermality - ", 
                                              round(scaleParams$isothermality_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$isothermality_s$`scaled:scale`,8), "):"))

tempNames <- str_replace_all(tempNames, pattern = "sand", 
                         replacement = paste0("((sand - ", 
                                              round(scaleParams$sand_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$sand_s$`scaled:scale`,8), ")"))

tempNames <- str_replace_all(tempNames, pattern = "coarse", 
                         replacement = paste0("((coarse - ", 
                                              round(scaleParams$coarse_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$coarse_s$`scaled:scale`,8), ")"))

tempNames <- str_replace_all(tempNames, pattern = "AWHC", 
                         replacement = paste0("((AWHC - ", 
                                              round(scaleParams$AWHC_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$AWHC_s$`scaled:scale`,8), ")"))

tempNames <- str_replace_all(tempNames, pattern = "carbon", 
                         replacement = paste0("((carbon - ", 
                                              round(scaleParams$carbon_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$carbon_s$`scaled:scale`,8), ")"))

tempNames <- str_replace_all(tempNames, pattern = "clay", 
                         replacement = paste0("((clay - ", 
                                              round(scaleParams$clay_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$clay_s$`scaled:scale`,8), ")"))

tempNames <- str_replace_all(tempNames, pattern = "annWatDef ", 
                         replacement = paste0("((annWatDef - ", 
                                              round(scaleParams$annWatDef_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$annWatDef_s$`scaled:scale`,8), ") "))

tempNames <- str_replace_all(tempNames, pattern = "annWatDef:", 
                         replacement = paste0("((annWatDef - ", 
                                              round(scaleParams$annWatDef_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$annWatDef_s$`scaled:scale`,8), "):"))

tempNames <- str_replace_all(tempNames, pattern = "prcp_seasonality ", 
                         replacement = paste0("((prcp_seasonality - ", 
                                              round(scaleParams$prcp_seasonality_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$prcp_seasonality_s$`scaled:scale`,8), ") "))

tempNames <- str_replace_all(tempNames, pattern = "prcp_seasonality:", 
                         replacement = paste0("((prcp_seasonality - ", 
                                              round(scaleParams$prcp_seasonality_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$prcp_seasonality_s$`scaled:scale`,8), "):"))

tempNames <- str_replace_all(tempNames, pattern = "prcp_seasonality\\^", 
                         replacement = paste0("((prcp_seasonality - ", 
                                              round(scaleParams$prcp_seasonality_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$prcp_seasonality_s$`scaled:scale`,8), ")^"))

tempNames <- str_replace_all(tempNames, pattern = "tmean ", 
                         replacement = paste0("((tmean - ", 
                                              round(scaleParams$tmean_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$tmean_s$`scaled:scale`,8), ") "))

tempNames <- str_replace_all(tempNames, pattern = "tmean:", 
                         replacement = paste0("((tmean - ", 
                                              round(scaleParams$tmean_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$tmean_s$`scaled:scale`,8), "):"))

tempNames <- str_replace_all(tempNames, pattern = "tmean$", 
                         replacement = paste0("((tmean - ", 
                                              round(scaleParams$tmean_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$tmean_s$`scaled:scale`,8), ")"))

tempNames <- str_replace_all(tempNames, pattern = "tmean\\^", 
                         replacement = paste0("((tmean - ", 
                                              round(scaleParams$tmean_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$tmean_s$`scaled:scale`,8), ")^"))

tempNames <- str_replace_all(tempNames, pattern = "annWetDegDays ", 
                         replacement = paste0("((annWetDegDays - ", 
                                              round(scaleParams$annWetDegDays_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$annWetDegDays_s$`scaled:scale`,8), ") "))

tempNames <- str_replace_all(tempNames, pattern = "annWetDegDays:", 
                         replacement = paste0("((annWetDegDays - ", 
                                              round(scaleParams$annWetDegDays_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$annWetDegDays_s$`scaled:scale`,8), "):"))

tempNames <- str_replace_all(tempNames, pattern = "prcp_dry ", 
                         replacement = paste0("((prcp_dry - ", 
                                              round(scaleParams$prcp_dry_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$prcp_dry_s$`scaled:scale`,8), ") "))

tempNames <- str_replace_all(tempNames, pattern = "prcp_dry:", 
                         replacement = paste0("((prcp_dry - ", 
                                              round(scaleParams$prcp_dry_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$prcp_dry_s$`scaled:scale`,8), "):"))

tempNames <- str_replace_all(tempNames, pattern = "annWatDef_95", 
                         replacement = paste0("((annWatDef_95 - ", 
                                              round(scaleParams$annWatDef_95_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$annWatDef_95_s$`scaled:scale`,8), ")"))


## print scaled model statement

scaledModelName <- paste0(responseVar, "~ exp(", tempNames, ")")
  
return(list("scaledInputVars_ModelStatement" = unscaledModelName, 
            "unscaledInputVars_scaledModelStatement" = scaledModelName))
}


## bestLambda model equations
bestLambdaModStatement <- getModelStatement(coefficientTable = coefs, 
                  modelName = "coefficientValue_bestLambda", 
                  responseVar = "P(Forest)")
## second-best lambda model equations
secondBestLambdaModStatement <- getModelStatement(coefficientTable = coefs, 
                  modelName = ifelse(name_secondBestMod == "1/2 SE Model", 
                                      yes = "coefficientValue_halfseLambda", 
                                     no = "coefficientValue_1seLambda"), 
                  responseVar = "P(Forest)")
```

#### The equation for the bestLambda model if the inputs *are* scaled is: 
```{r}
print(bestLambdaModStatement$scaledInputVars_ModelStatement)
```

<!-- #### The equation for the bestLambda model if the inputs *are NOT* scaled is:  -->
<!-- ```{r} -->
<!-- print(bestLambdaModStatement$unscaledInputVars_scaledModelStatement) -->
<!-- ``` -->

#### The equation for the *second*-bestLambda model if the inputs *are* scaled is: 
```{r}
print(secondBestLambdaModStatement$scaledInputVars_ModelStatement)
```

<!-- #### The equation for the *second*-bestLambda model if the inputs *are NOT* scaled is:  -->
<!-- ```{r} -->
<!-- print(secondBestLambdaModStatement$unscaledInputVars_scaledModelStatement) -->
<!-- ``` -->

## Now, use the model to predict if a location is or is not trees, for all observations in the cover dataset (which we will use in the subsequent step of the cover-modeling process).
We use the 'modDat_1_s_temp' data.frame, which has all of the cover data, even for locations without observed tree cover, but has the climate and soils data that have been scaled in that same was as the data we used to fit the yes/no tree model
```{r}
# prep data
modDat_1_s_temp$rowID <- 1:nrow(modDat_1_s_temp)
predData <- modDat_1_s_temp %>% 
  select(rowID, Year, x, y, paste0(prednames, "_s")) %>% 
  rename_with(~str_remove(.x, "_s$"), 
              any_of(paste0(names(prednames), "_s")))
# for bestLambda model
predData$treeProb_bestLambda <- predict.glm(object = fit_glm_bestLambda, newdata = predData, type = "response")
# for scond-bestLambda model
predData$treeProb_secondBestLambda <- predict.glm(object = mod_secondBest, newdata = predData, type = "response")


# make continuous predictions binary according to the identified best threshold for each model 
predData <- predData %>% 
  mutate(treeProb_bestLambda_continuous = approxfun_bestLambda(treeProb_bestLambda)) %>% # use interpolation to shift the continuous values
  mutate(treeProb_bestLambda_continuous = replace(treeProb_bestLambda_continuous, list = treeProb_bestLambda < 0, 0)) %>% # replace NAs at the 'bottom' of the interpolation scale with 0s
  mutate(treeProb_bestLambda_binary = treeProb_bestLambda_continuous,
    treeProb_bestLambda_binary = replace(treeProb_bestLambda_binary, treeProb_bestLambda_binary > 0.1, 1), # now, make the interpolated continuous model predictions binary
         treeProb_bestLambda_binary = replace(treeProb_bestLambda_binary, treeProb_bestLambda_binary <= 0.1, 0)) 

# add predictions back to the original data frame and save 
modDat_treePreds <- modDat_1_s_temp %>% 
  left_join(predData %>% select(-c(tmean:AWHC)), by = c("rowID", "Year", "x", "y")) %>% 
  select(-c(tmin_s:AWHC_s))

saveRDS(modDat_treePreds,"../../../Data_processed/CoverData/CoverModelData_withTreeNoTreePreds.rds")

```


<!-- ```{r, eval = FALSE} -->
<!-- # get data -->
<!-- modDat_treePreds <- readRDS("../../../Data_processed/CoverData/CoverModelData_withTreeNoTreePreds.rds") -->

<!-- # make map -->
<!-- modDat_treePreds %>%  -->
<!--   filter(round(treeProb_bestLambda,2) == 0) %>%  -->
<!-- ggplot() +  -->
<!--   geom_sf(data = cropped_states_2, aes(NULL)) + -->
<!--   geom_point(aes(x = Long, y = Lat), size = 1) -->
<!-- #   geom_density(aes(treeProb_bestLambda)) +  -->
<!-- #   geom_density(aes(treeProb_bestLambda_binary)) -->
<!-- ``` -->


<!-- # Cross-validation -->
<!-- ## Using best lambda model -->
<!-- ### Use terms from global model to re-fit and predict on different held out regions -->
<!-- ### Figures show residuals for each of the models fit to held-out ecoregions -->
<!-- These models were fit to six ecoregions, and then predict on the indicated heldout ecoregion -->

<!-- ```{r cross-validation for best lambda model , fig.height = 7, fig.width = 9, warning=FALSE, message=FALSE } -->
<!-- if (length(prednames_fig) == 0) { -->
<!--   print("The model only contains one predictor (an intercept), so cross validation isn't practical") -->
<!-- } else { -->

<!-- ## code from Tredennick et al. 2020 -->
<!-- # try each separate level II ecoregion as a test set -->
<!-- # make a list to hold output data -->
<!-- outList <- vector(mode = "list", length = length(sort(unique(modDat_1_s$NA_L2NAME)))) -->
<!-- # obs_pred <- data.frame(ecoregion = character(),obs = numeric(), -->
<!-- #                        pred_opt = numeric(), pred_null = numeric()#, -->
<!-- #                        #pred_nopenalty = numeric() -->
<!-- #                        ) -->

<!-- ## get the model specification from the global model -->
<!-- mat <- as.matrix(coef(fit_glm_bestLambda, s = "lambda.min")) -->
<!-- mat2 <- mat[mat[,1] != 0,] -->

<!-- f_cv <- as.formula(paste0("TotalTreeCover_binom", " ~ ", paste0(names(mat2)[2:length(names(mat2))], collapse = " + "))) -->

<!-- X_cv <- model.matrix(object = f_cv, data = modDat_1_s) -->
<!-- # get response variable -->
<!-- y_cv <- as.matrix(modDat_1_s[,"TotalTreeCover_binom"]) -->


<!-- # now, loop through so with each iteration, a different ecoregion is held out -->
<!--  for(i_eco in sort(unique(modDat_1_s$NA_L2NAME))){ -->

<!--   # split into training and test sets -->
<!--   test_eco <- i_eco -->
<!--   print(test_eco) -->
<!--   # identify the rowID of observations to be in the training and test datasets -->
<!--   train <- which(modDat_1_s$NA_L2NAME!=test_eco) # data for all ecoregions that aren't 'i_eco' -->
<!--   test <- which(modDat_1_s$NA_L2NAME==test_eco) # data for the ecoregion that is 'i_eco' -->

<!--   trainDat_all <- modDat_1_s %>% -->
<!--     slice(train) %>% -->
<!--     dplyr::select(-newRegion) -->
<!--   testDat_all <- modDat_1_s %>% -->
<!--     slice(test) %>% -->
<!--     dplyr::select(-newRegion) -->

<!--   # get the model matrices for input and response variables for cross validation model specification -->
<!--   X_train <- as.matrix(X_cv[train,]) -->
<!--   X_test <- as.matrix(X_cv[test,]) -->

<!--   y_train <- modDat_1_s[train,"TotalTreeCover_binom"] -->
<!--   y_test <- modDat_1_s[test,"TotalTreeCover_binom"] -->

<!--   # get the model matrices for input and response variables for original model specification -->
<!--   X_train_glob <- as.matrix(X[train,]) -->
<!--   X_test_glob <- as.matrix(X[test,]) -->

<!--   y_train_glob <- modDat_1_s[train,"TotalTreeCover_binom"] -->
<!--   y_test_glob <- modDat_1_s[test,"TotalTreeCover_binom"] -->

<!--   train_eco <- modDat_1_s$NA_L2NAME[train] -->

<!--   ## just try a regular glm w/ the components from the global model -->
<!--   fit_i <- glm(data = trainDat_all, formula = f_cv, -->
<!--     , -->
<!--                family =  binomial -->
<!--     ) -->

<!--   # lasso model predictions with the optimal lambda (back transformed) -->
<!--   optimal_pred <- predict(fit_i, newdata = testDat_all, type = "response")  -->
<!--   # null model and predictions -->
<!--   # the "null" model in this case is the global model -->
<!--   # predict on the test data for this iteration w/ the global model (back transformed) -->
<!--   null_pred <- predict.glm(fit_glm_bestLambda, newdata = testDat_all, type = "response")  -->

<!--   # save data -->
<!--   tmp <- data.frame(ecoRegion_holdout = rep(test_eco,length(y_test)), -->
<!--                     obs=y_test, -->
<!--                     pred_opt=optimal_pred, -->
<!--                     pred_null=null_pred#, -->
<!--                     #pred_nopenalty=nopen_pred -->
<!--                     ) %>% -->
<!--     cbind(testDat_all) -->

<!--   # calculate RMSE, bias, etc. of -->
<!--   # RMSE of CV model -->
<!--   RMSE_optimal <- yardstick::rmse(data = data.frame(optimal_pred,"y_test" = (y_test)), truth = "y_test", estimate = "optimal_pred")[1,]$.estimate -->
<!--   # RMSE of global model -->
<!--   RMSE_null <- yardstick::rmse(data = data.frame(null_pred,"y_test" = (y_test)), truth = "y_test", estimate = "null_pred")[1,]$.estimate -->
<!--   # bias of CV model -->
<!--   bias_optimal <- mean((y_test) - optimal_pred) -->
<!--   # bias of global model -->
<!--   bias_null <-  mean((y_test) - null_pred ) -->

<!--   # put output into a list -->
<!--   tmpList <- list("testRegion" = i_eco, -->
<!--     "modelObject" = fit_i, -->
<!--        "modelPredictions" = tmp, -->
<!--     "performanceMetrics" = data.frame("RMSE_cvModel" = RMSE_optimal, -->
<!--                                       "RMSE_globalModel" = RMSE_null, -->
<!--                                       "bias_cvModel" = bias_optimal, -->
<!--                                       "bias_globalModel" = bias_null)) -->

<!--   # save model outputs -->
<!--   outList[[which(sort(unique(modDat_1_s$NA_L2NAME)) == i_eco)]] <- tmpList -->
<!--  } -->
<!-- } -->
<!-- ``` -->
<!-- ### Below are the RMSE and bias values for predictions made for each holdout level II ecoregion, compared to predictions from the global model for that same ecoregion -->
<!-- ```{r metrics output table for best lambda CV} -->
<!-- # table of model performance -->
<!-- purrr::map(outList, .f = function(x) { -->
<!--   cbind(data.frame("holdout region" = x$testRegion),  x$performanceMetrics) -->
<!-- } -->
<!-- ) %>% -->
<!--   purrr::list_rbind() %>% -->
<!--   kable(col.names = c("Held-out ecoregion", "RMSE of CV model", "RMSE of global model", -->
<!--                       "bias of CV model - mean(obs-pred.)", "bias of global model- mean(obs-pred.)"), -->
<!--         caption = "Performance of Cross Validation using 'best lambda' model specification") %>% -->
<!-- kable_styling(bootstrap_options = c("striped", "hover", "condensed")) -->
<!-- ``` -->

<!-- ```{r figures for best lambda CV} -->
<!-- # visualize model predictions -->
<!-- for (i in 1:length(unique(modDat_1_s$NA_L2NAME))) { -->
<!--   holdoutRegion <- outList[[i]]$testRegion -->
<!--   predictionData <- outList[[i]]$modelPredictions -->
<!--   modTerms <- as.matrix(coef(outList[[i]]$modelObject)) %>% -->
<!--     as.data.frame() %>% -->
<!--     filter(V1!=0) %>% -->
<!--     rownames() -->

<!--   # calculate residuals -->
<!--   predictionData <- predictionData %>% -->
<!--   mutate(resid = .[["obs"]] - .[["pred_opt"]] , -->
<!--          resid_globMod = .[["obs"]]  - .[["pred_null"]]) -->

<!-- # rasterize -->
<!-- # use 'test_rast' from earlier -->
<!--   # rasterize data -->
<!-- plotObs <- predictionData %>% -->
<!--          drop_na(paste("TotalTreeCover_binom")) %>% -->
<!--   #slice_sample(n = 5e4) %>% -->
<!--   terra::vect(geom = c("x", "y")) %>% -->
<!--   terra::set.crs(crs(test_rast)) %>% -->
<!--   terra::rasterize(y = test_rast, -->
<!--                    field = "resid", -->
<!--                    fun = mean) -->

<!-- tempExt <- crds(plotObs, na.rm = TRUE) -->

<!-- plotObs_2 <- plotObs %>% -->
<!--   terra::crop(ext(min(tempExt[,1]), max(tempExt[,1]), -->
<!--            min(tempExt[,2]), max(tempExt[,2])) -->
<!--        ) -->

<!-- # make figures -->
<!-- # make histogram -->
<!-- hist_i <- ggplot(predictionData) + -->
<!--   geom_histogram(aes(resid_globMod), col = "darkgrey", fill = "lightgrey") + -->
<!--   xlab(c("Residuals (obs. - pred.)")) -->
<!-- # make map -->
<!-- map_i <-  ggplot() + -->
<!-- geom_spatraster(data = plotObs_2) + -->
<!--   geom_sf(data=cropped_states_2,fill=NA ) + -->
<!--   geom_sf(data = mapRegions, fill = NA, col = "orchid", lwd = .5) + -->
<!-- labs(title = paste0("Residuals (obs. - pred.) for predictions of \n", holdoutRegion, " \n from a model fit to other ecoregions"), -->
<!--      subtitle = paste0("TotalTreeCover_binom", " ~ ", paste0( modTerms, collapse = " + "))) + -->
<!--   scale_fill_gradient2(low = "red", -->
<!--                        mid = "white" , -->
<!--                        high = "blue" , -->
<!--                        midpoint = 0,   na.value = "white", -->
<!--                        limits = c(-1, 1))   +  -->
<!--   xlim(st_bbox(plotObs_2)[1],st_bbox(plotObs_2)[3]) +  -->
<!--   ylim(st_bbox(plotObs_2)[2],st_bbox(plotObs_2)[4]) -->

<!--  assign(paste0("residPlot_",holdoutRegion), -->
<!--    value = ggarrange(map_i, hist_i, heights = c(3,1), ncol = 1) -->
<!-- ) -->

<!-- } -->

<!--   lapply(unique(modDat_1_s$NA_L2NAME), FUN = function(x) { -->
<!--     get(paste0("residPlot_", x)) -->
<!--   }) -->

<!-- ``` -->

<!-- ## Using second best lambda model (either 1se or 1/2se) -->
<!-- ### Use terms from global model to re-fit and predict on different held out regions -->
<!-- ### Figures show residuals for each of the models fit to held-out ecoregions -->
<!-- These models were fit to six ecoregions, and then predict on the indicated heldout ecoregion -->

<!-- ```{r cross-validation for 1 SE lambda model , fig.height = 7, fig.width = 9, warning=FALSE, message=FALSE } -->
<!-- if (length(prednames_secondBestMod) == 0) { -->
<!--   print("The model only contains one predictor (an intercept), so cross validation isn't practical") -->
<!-- } else { -->

<!-- ## code from Tredennick et al. 2020 -->
<!-- # try each separate level II ecoregion as a test set -->
<!-- # make a list to hold output data -->
<!-- outList <- vector(mode = "list", length = length(sort(unique(modDat_1_s$NA_L2NAME)))) -->

<!-- ## get the model specification from the global model -->
<!-- mat <- as.matrix(coef(mod_secondBest, s = "lambda.min")) -->
<!-- mat2 <- mat[mat[,1] != 0,] -->

<!-- f_cv <- as.formula(paste0("TotalTreeCover_binom", " ~ ", paste0(names(mat2)[2:length(names(mat2))], collapse = " + "))) -->

<!-- X_cv <- model.matrix(object = f_cv, data = modDat_1_s) -->
<!-- # get response variable -->
<!-- y_cv <- as.matrix(modDat_1_s[,"TotalTreeCover_binom"]) -->


<!-- # now, loop through so with each iteration, a different ecoregion is held out -->
<!--  for(i_eco in sort(unique(modDat_1_s$NA_L2NAME))){ -->

<!--   # split into training and test sets -->
<!--   test_eco <- i_eco -->
<!--   print(test_eco) -->
<!--   # identify the rowID of observations to be in the training and test datasets -->
<!--   train <- which(modDat_1_s$NA_L2NAME!=test_eco) # data for all ecoregions that aren't 'i_eco' -->
<!--   test <- which(modDat_1_s$NA_L2NAME==test_eco) # data for the ecoregion that is 'i_eco' -->

<!--   trainDat_all <- modDat_1_s %>% -->
<!--     slice(train) %>% -->
<!--     dplyr::select(-newRegion) -->
<!--   testDat_all <- modDat_1_s %>% -->
<!--     slice(test) %>% -->
<!--     dplyr::select(-newRegion) -->

<!--   # get the model matrices for input and response variables for cross validation model specification -->
<!--   X_train <- as.matrix(X_cv[train,]) -->
<!--   X_test <- as.matrix(X_cv[test,]) -->

<!--   y_train <- modDat_1_s[train,"TotalTreeCover_binom"] -->
<!--   y_test <- modDat_1_s[test,"TotalTreeCover_binom"] -->

<!--   # get the model matrices for input and response variables for original model specification -->
<!--   X_train_glob <- as.matrix(X[train,]) -->
<!--   X_test_glob <- as.matrix(X[test,]) -->

<!--   y_train_glob <- modDat_1_s[train,"TotalTreeCover_binom"] -->
<!--   y_test_glob <- modDat_1_s[test,"TotalTreeCover_binom"] -->

<!--   train_eco <- modDat_1_s$NA_L2NAME[train] -->

<!--   ## just try a regular glm w/ the components from the global model -->
<!--   fit_i <- glm(data = trainDat_all, formula = f_cv, -->
<!--     , -->
<!--                family =  binomial -->
<!--     ) -->

<!--     coef(fit_i) -->

<!--   # lasso model predictions with the optimal lambda -->
<!--   optimal_pred <- predict(fit_i, newdata= testDat_all, type = "response")  -->
<!--   # null model and predictions -->
<!--   # the "null" model in this case is the global model -->
<!--   # predict on the test data for this iteration w/ the global model -->
<!--   null_pred <- predict.glm(mod_secondBest, newdata = testDat_all, type = "response")  -->

<!--   # save data -->
<!--   tmp <- data.frame(ecoRegion_holdout = rep(test_eco,length(y_test)), -->
<!--                     obs=y_test, -->
<!--                     pred_opt=optimal_pred, -->
<!--                     pred_null=null_pred#, -->
<!--                     #pred_nopenalty=nopen_pred -->
<!--                     ) %>% -->
<!--     cbind(testDat_all) -->

<!--   # calculate RMSE, bias, etc. of -->
<!--   # RMSE of CV model -->
<!--   RMSE_optimal <- yardstick::rmse(data = data.frame(optimal_pred, "y_test" = (y_test)), truth = "y_test", estimate = "optimal_pred")[1,]$.estimate -->
<!--   # RMSE of global model -->
<!--   RMSE_null <- yardstick::rmse(data = data.frame(null_pred,  "y_test" = (y_test)), truth = "y_test", estimate = "null_pred")[1,]$.estimate -->
<!--   # bias of CV model -->
<!--   bias_optimal <- mean((y_test) - optimal_pred) -->
<!--   # bias of global model -->
<!--   bias_null <-  mean((y_test) - null_pred ) -->

<!--   # put output into a list -->
<!--   tmpList <- list("testRegion" = i_eco, -->
<!--     "modelObject" = fit_i, -->
<!--        "modelPredictions" = tmp, -->
<!--     "performanceMetrics" = data.frame("RMSE_cvModel" = RMSE_optimal, -->
<!--                                       "RMSE_globalModel" = RMSE_null, -->
<!--                                       "bias_cvModel" = bias_optimal, -->
<!--                                       "bias_globalModel" = bias_null)) -->

<!--   # save model outputs -->
<!--   outList[[which(sort(unique(modDat_1_s$NA_L2NAME)) == i_eco)]] <- tmpList -->
<!--  } -->
<!-- } -->
<!-- ``` -->

<!-- ### Below are the RMSE and bias values for predictions made for each holdout level II ecoregion, compared to predictions from the global model for that same ecoregion -->
<!-- ```{r metrics output table for 1SE lambda CV} -->

<!-- if (length(prednames_secondBestMod) == 0) { -->
<!--   print("The model only contains one predictor (an intercept), so cross validation isn't practical") -->
<!-- } else { -->
<!-- # table of model performance -->
<!-- purrr::map(outList, .f = function(x) { -->
<!--   cbind(data.frame("holdout region" = x$testRegion),  x$performanceMetrics) -->
<!-- } -->
<!-- ) %>% -->
<!--   purrr::list_rbind() %>% -->
<!--   kable(col.names = c("Held-out ecoregion", "RMSE of CV model", "RMSE of global model", -->
<!--                       "bias of CV model - mean(obs-pred.)", "bias of global model - mean(obs-pred.)"), -->
<!--         caption = "Performance of Cross Validation using 'second best lambda' model specification") %>% -->
<!-- kable_styling(bootstrap_options = c("striped", "hover", "condensed")) -->
<!-- } -->
<!-- ``` -->

<!-- ```{r figures of CV for 1 SE lambda model} -->
<!-- if (length(prednames_secondBestMod) == 0) { -->
<!--   print("The model only contains one predictor (an intercept), so cross validation isn't practical") -->
<!-- } else { -->
<!-- for (i in 1:length(unique(modDat_1_s$NA_L2NAME))) { -->
<!--   holdoutRegion <- outList[[i]]$testRegion -->
<!--   predictionData <- outList[[i]]$modelPredictions -->
<!--   modTerms <- as.matrix(coef(outList[[i]]$modelObject)) %>% -->
<!--     as.data.frame() %>% -->
<!--     filter(V1!=0) %>% -->
<!--     rownames() -->

<!--   # calculate residuals -->
<!--   predictionData <- predictionData %>% -->
<!--   mutate(resid = .[["obs"]] - .[["pred_opt"]] , -->
<!--          resid_globMod = .[["obs"]]  - .[["pred_null"]]) -->


<!-- # rasterize -->
<!-- # use 'test_rast' from earlier -->

<!--   # rasterize data -->
<!-- plotObs <- predictionData %>% -->
<!--          drop_na(paste("TotalTreeCover_binom")) %>% -->
<!--   #slice_sample(n = 5e4) %>% -->
<!--   terra::vect(geom = c("x", "y")) %>% -->
<!--   terra::set.crs(crs(test_rast)) %>% -->
<!--   terra::rasterize(y = test_rast, -->
<!--                    field = "resid", -->
<!--                    fun = mean)  -->

<!-- tempExt <- crds(plotObs, na.rm = TRUE) -->

<!-- plotObs_2 <- plotObs %>% -->
<!--   crop(ext(min(tempExt[,1]), max(tempExt[,1]), -->
<!--            min(tempExt[,2]), max(tempExt[,2])) -->
<!--        ) -->

<!-- # make figures -->
<!-- # make histogram -->
<!-- hist_i <- ggplot(predictionData) + -->
<!--   geom_histogram(aes(resid_globMod), col = "darkgrey", fill = "lightgrey") + -->
<!--   xlab(c("Residuals (obs. - pred.)")) -->
<!-- # make map -->
<!-- map_i <-  ggplot() + -->
<!-- geom_spatraster(data = plotObs_2) + -->
<!--   geom_sf(data=cropped_states_2,fill=NA ) + -->
<!--   geom_sf(data = mapRegions, fill = NA, col = "orchid", lwd = .5) + -->
<!-- labs(title = paste0("Residuals (obs. - pred.) for predictions of \n", holdoutRegion, " \n from a model fit to other ecoregions"), -->
<!--      subtitle = paste0("TotalTreeCover_binom", " ~ ", paste0( modTerms, collapse = " + "))) + -->
<!--   scale_fill_gradient2(low = "red", -->
<!--                        mid = "white" , -->
<!--                        high = "blue" , -->
<!--                        midpoint = 0,   na.value = "white", -->
<!--                        limits = c(-1, 1))  +  -->
<!--   xlim(st_bbox(plotObs_2)[1],st_bbox(plotObs_2)[3]) +  -->
<!--   ylim(st_bbox(plotObs_2)[2],st_bbox(plotObs_2)[4]) -->

<!--  assign(paste0("residPlot_",holdoutRegion), -->
<!--    value = ggarrange(map_i, hist_i, heights = c(3,1), ncol = 1) -->
<!-- ) -->

<!-- } -->

<!--   lapply(unique(modDat_1_s$NA_L2NAME), FUN = function(x) { -->
<!--     get(paste0("residPlot_", x)) -->
<!--   }) -->
<!-- } -->

<!-- ``` -->


<!-- # Save output -->

<!-- ```{r save_output} -->
<!-- ## save the coefficients for the models (best lambda, 1/2se lambda, 1se lambda) -->

<!-- saveRDS(coefs, file = paste0("./models/ForestMod_modelCoefficients.rds")) -->
<!-- saveRDS(uniqueCoeffs, file = paste0("./models/ForestMod_modelMetrics_trimAnom.rds")) -->

<!-- # make a table -->

<!-- ``` -->




<!-- ```{r pdp_glm, warning = FALSE, fig.width = 8, fig.height=8} -->
<!-- ## partial dependence plots -->
<!-- #vip::vip(mod_glmFinal, num_features = 15) -->

<!-- #pdp_all_vars(mod_glmFinal, mod_vars = pred_vars, ylab = 'probability',train = df_small) -->

<!-- #caret::varImp(fit) -->
<!-- ``` -->


<!-- # session info -->

<!-- Hash of current commit (i.e. to ID the version of the code used) -->

<!-- ```{r} -->
<!-- system("git rev-parse HEAD", intern=TRUE) -->
<!-- ``` -->

<!-- Packages etc. -->

<!-- ```{r} -->
<!-- sessionInfo() -->

<!-- ``` -->