---
title: "02_modelFitting_globalTreeModel"
author: "Alice Stears"
date: "`r lubridate::today()`"
output:
  html_document:
    toc: true
    df_print: paged
    code_folding: hide
    dev: "jpeg"
params:
  run: FALSE      
  save_figs: TRUE
  ecoregion: shrubGrass
  response: TotalTreeCover
  removeTexasLouisianaPlain: FALSE
  removeAllAnoms: TRUE
  trimAnomalies: FALSE
  autoKfold: FALSE
  whichSecondBestMod: "auto"# 1se OR halfse OR auto (chose the model w/ fewest # of predictors)
subtitle: "Based on code from Martin Holdrege's Sagebrush-Fire paper"
---
This workflow fits a model across all of CONUS that predicts whether a location does or does not have trees. 

The data consists of vegetation % cover by functional group from across CONUS (from AIM, FIA, LANDFIRE, and RAP), 
as well as climate variables from DayMet, which have been aggregated into mean interannual conditions accross multiple temporal windows. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE,
                      message = FALSE, dpi = 72)
```

# Dependencies 

Set user defined parameters

```{r}
print(params)
# set to true if want to run for a limited number of rows (i.e. for code testing)
test_run <- params$test_run
save_figs <- params$save_figs
response <- params$response
fit_sample <- TRUE # fit model to a sample of the data
n_train <- 5e4 # sample size of the training data
n_test <- 1e6 # sample size of the testing data (if this is too big the decile dotplot code throws memory errors)
trimAnom <- params$trimAnomalies
removeTLP <- params$removeTexasLouisianaPlain
run <- params$run
autoKfold <- params$autoKfold
removeAllAnoms <- params$removeAllAnoms
whichSecondBestMod <- params$whichSecondBestMod
```

Load packages and source functions
```{r warning=FALSE, message=FALSE}
# set option so resampled dataset created here reproduces earlier runs of this code with dplyr 1.0.10
source("../../../Functions/glmTransformsIterates.R")
source("../../../Functions/transformPreds.R")
source("../../../Functions/betaLASSO.R")

#source("../../../Functions/StepBeta_mine.R")
#source("src/fig_params.R")
#source("src/modeling_functions.R")
 
library(betareg)
library(ggspatial)
library(terra)
library(tidyterra)
library(sf)
library(caret)
library(tidyverse)
library(GGally) # for ggpairs()
library(pdp) # for partial dependence plots
library(gridExtra)
library(knitr)
library(patchwork) # for figure insets etc. 
library(ggtext)
library(StepBeta)
theme_set(theme_classic())
library(here)
library(rsample)
library(kableExtra)
library(glmnet)
library(USA.state.boundaries)
```

# Read in data

Data compiled in the `prepDataForModels.R` script

```{r}
here::i_am("Analysis/VegComposition/ModelFitting/02_ModelFitting_globalTreeModel.Rmd")
modDat <- readRDS( here("Data_processed", "CoverData", "DataForModels_spatiallyAveraged_withSoils_noSf_sampledLANDFIRE.rds")) %>% st_drop_geometry()
```

# Prep data

We will fit a binomial model that predicts whether or not there are trees at a location. Because the tree cover data we have is continuous between 0 and 100, we convert it to be binomial be forcing any values ≤ 10% to be 0, and any values > 10% to be 1. 

```{r}
modDat <- modDat %>% 
  mutate(TotalTreeCover_binom = replace(TotalTreeCover, TotalTreeCover <=10, 0)) %>% 
  mutate(TotalTreeCover_binom = replace(TotalTreeCover_binom, TotalTreeCover_binom > 10, 1))
```

```{r rename variables}
set.seed(1234)
# now, rename columns for brevity
modDat_1 <- modDat %>% 
  dplyr::select(-c(prcp_annTotal:annVPD_min)) %>% 
  # mutate(Lon = st_coordinates(.)[,1], 
  #        Lat = st_coordinates(.)[,2])  %>% 
  # st_drop_geometry() %>% 
  # filter(!is.na(newRegion))
  rename("tmin" = tmin_meanAnnAvg_CLIM, 
     "tmax" = tmax_meanAnnAvg_CLIM, #1
     "tmean" = tmean_meanAnnAvg_CLIM, 
     "prcp" = prcp_meanAnnTotal_CLIM, 
     "t_warm" = T_warmestMonth_meanAnnAvg_CLIM,
     "t_cold" = T_coldestMonth_meanAnnAvg_CLIM, 
     "prcp_wet" = precip_wettestMonth_meanAnnAvg_CLIM,
     "prcp_dry" = precip_driestMonth_meanAnnAvg_CLIM, 
     "prcp_seasonality" = precip_Seasonality_meanAnnAvg_CLIM, #2
     "prcpTempCorr" = PrecipTempCorr_meanAnnAvg_CLIM,  #3
     "abvFreezingMonth" = aboveFreezing_month_meanAnnAvg_CLIM, 
     "isothermality" = isothermality_meanAnnAvg_CLIM, #4
     "annWatDef" = annWaterDeficit_meanAnnAvg_CLIM, 
     "annWetDegDays" = annWetDegDays_meanAnnAvg_CLIM,
     "VPD_mean" = annVPD_mean_meanAnnAvg_CLIM, 
     "VPD_max" = annVPD_max_meanAnnAvg_CLIM, #5
     "VPD_min" = annVPD_min_meanAnnAvg_CLIM, #6
     "VPD_max_95" = annVPD_max_95percentile_CLIM, 
     "annWatDef_95" = annWaterDeficit_95percentile_CLIM, 
     "annWetDegDays_5" = annWetDegDays_5percentile_CLIM, 
     "frostFreeDays_5" = durationFrostFreeDays_5percentile_CLIM, 
     "frostFreeDays" = durationFrostFreeDays_meanAnnAvg_CLIM, 
     "soilDepth" = soilDepth, #7
     "clay" = surfaceClay_perc, 
     "sand" = avgSandPerc_acrossDepth, #8
     "coarse" = avgCoarsePerc_acrossDepth, #9
     "carbon" = avgOrganicCarbonPerc_0_3cm, #10
     "AWHC" = totalAvailableWaterHoldingCapacity,
     ## anomaly variables
     tmean_anom = tmean_meanAnnAvg_3yrAnom, #15
     tmin_anom = tmin_meanAnnAvg_3yrAnom, #16
     tmax_anom = tmax_meanAnnAvg_3yrAnom, #17
    prcp_anom = prcp_meanAnnTotal_3yrAnom, #18
      t_warm_anom = T_warmestMonth_meanAnnAvg_3yrAnom,  #19
     t_cold_anom = T_coldestMonth_meanAnnAvg_3yrAnom, #20
      prcp_wet_anom = precip_wettestMonth_meanAnnAvg_3yrAnom, #21
      precp_dry_anom = precip_driestMonth_meanAnnAvg_3yrAnom,  #22
    prcp_seasonality_anom = precip_Seasonality_meanAnnAvg_3yrAnom, #23 
     prcpTempCorr_anom = PrecipTempCorr_meanAnnAvg_3yrAnom, #24
      aboveFreezingMonth_anom = aboveFreezing_month_meanAnnAvg_3yrAnom, #25  
    isothermality_anom = isothermality_meanAnnAvg_3yrAnom, #26
       annWatDef_anom = annWaterDeficit_meanAnnAvg_3yrAnom, #27
     annWetDegDays_anom = annWetDegDays_meanAnnAvg_3yrAnom,  #28
      VPD_mean_anom = annVPD_mean_meanAnnAvg_3yrAnom, #29
      VPD_min_anom = annVPD_min_meanAnnAvg_3yrAnom,  #30
      VPD_max_anom = annVPD_max_meanAnnAvg_3yrAnom,  #31
     VPD_max_95_anom = annVPD_max_95percentile_3yrAnom, #32
      annWatDef_95_anom = annWaterDeficit_95percentile_3yrAnom, #33 
      annWetDegDays_5_anom = annWetDegDays_5percentile_3yrAnom ,  #34
    frostFreeDays_5_anom = durationFrostFreeDays_5percentile_3yrAnom, #35 
      frostFreeDays_anom = durationFrostFreeDays_meanAnnAvg_3yrAnom #36
  ) %>% 
  dplyr::select(-c(tmin_meanAnnAvg_3yr:durationFrostFreeDays_meanAnnAvg_3yr))

```

## Visualize the predictor variables

The following are the candidate predictor variables for this ecoregion: 
```{r get candidate predictor variables}
  prednames <- c(
    "tmean"               ,"prcp"               ,"prcp_seasonality", "prcpTempCorr"       ,  "isothermality"     ,     
"annWetDegDays"           ,"sand"               ,"coarse"         , "AWHC"                #, "tmin_anom"          ,    
#"tmax_anom"               ,"t_warm_anom", "prcp_wet_anom"      ,"precp_dry_anom" , "prcp_seasonality_anom", "prcpTempCorr_anom" ,     
#"aboveFreezingMonth_anom" ,"isothermality_anom" ,"annWatDef_anom" , "annWetDegDays_anom"  , "VPD_min_anom"     
  )
```

## Scale the predictor variables for the model-fitting process
```{r scaling predictors, fig.height=7, fig.width=10}
allPreds <- modDat_1 %>% 
  dplyr::select(tmin:frostFreeDays,tmean_anom:frostFreeDays_anom, soilDepth:AWHC) %>% 
  names()
modDat_1_s <- modDat_1 %>% 
  mutate(across(all_of(allPreds), base::scale, .names = "{.col}_s")) 
saveRDS(modDat_1_s, file = "./models/scaledModelInputData.rds")
```

## Remove the rows that have no observations for total tree cover
```{r subset data to that ecoregion}
# 
modDat_1_s <- modDat_1_s[!is.na(modDat_1_s[,"TotalTreeCover_binom"]),]
```


```{r}
# subset the data to only include these predictors, and remove any remaining NAs 
modDat_1_s <- modDat_1_s %>% 
  dplyr::select(prednames, paste0(prednames, "_s"), TotalTreeCover, TotalTreeCover_binom, newRegion, Year, x, y, NA_L1NAME, NA_L2NAME) %>% 
  drop_na()

names(prednames) <- prednames
df_pred <- modDat_1_s[, prednames]

response <- "TotalTreeCover"
```


## Visualize the response variable 
```{r visualize response variable}
ggplot(modDat_1_s) + 
  geom_histogram(aes(TotalTreeCover/100), fill = "darkgreen", col = "darkgreen", alpha = .5) + 
  xlab("Tree Cover") + 
  ggtitle("Untransformed, observed tree cover")

ggplot(modDat_1_s) + 
  geom_histogram(aes(TotalTreeCover_binom), fill = "purple", alpha = .5, col = "purple") +
  xlab("Tree Cover") + 
  ggtitle("Tree cover, converted to binomial with a 10% 'cutoff'")

```

```{r summary_table}
create_summary <- function(df) {
  df %>% 
    pivot_longer(cols = everything(),
                 names_to = 'variable') %>% 
    group_by(variable) %>% 
    summarise(across(value, .fns = list(mean = ~mean(.x, na.rm = TRUE), min = ~min(.x, na.rm = TRUE), 
                                        median = ~median(.x, na.rm = TRUE), max = ~max(.x, na.rm = TRUE)))) %>% 
    mutate(across(where(is.numeric), round, 4))
}

modDat_1_s[prednames] %>% 
  create_summary() %>% 
  knitr::kable(caption = 'summaries of possible predictor variables') %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed")) 

```

### Histograms of raw and scaled predictors
```{r}
  
scaleFigDat_1 <- modDat_1_s %>% 
  dplyr::select(c(x, y, Year, prednames)) %>% 
  pivot_longer(cols = all_of(names(prednames)), 
               names_to = "predNames", 
               values_to = "predValues_unScaled")
scaleFigDat_2 <- modDat_1_s %>% 
  dplyr::select(c(x, y, Year, paste0(prednames, "_s"))) %>% 
  pivot_longer(cols = all_of(paste0(prednames,"_s"
                                    )), 
               names_to = "predNames", 
               values_to = "predValues_scaled", 
               names_sep = ) %>% 
  mutate(predNames = str_replace(predNames, pattern = "_s$", replacement = ""))

scaleFigDat_3 <- scaleFigDat_1 %>% 
  left_join(scaleFigDat_2)

ggplot(scaleFigDat_3) + 
  facet_wrap(~predNames, scales = "free") +
  geom_histogram(aes(predValues_unScaled), fill = "lightgrey", col = "darkgrey") + 
  geom_histogram(aes(predValues_scaled), fill = "lightblue", col = "blue") +
  xlab ("predictor variable values") + 
  ggtitle("Comparing the distribution of unscaled (grey) to scaled (blue) predictor variables")
```

```{r}
modDat_1_s <- modDat_1_s %>% 
  rename_with(~paste0(.x, "_raw"), 
                any_of(names(prednames))) %>% 
  rename_with(~str_remove(.x, "_s$"), 
              any_of(paste0(names(prednames), "_s")))
  
```

## Predictor variables compared to binned response variables

```{r climVar_boxplots, fig.height=9, fig.width=8, eval=FALSE}
set.seed(12011993)
# vector of name of response variables
vars_response <- response
# longformat dataframes for making boxplots
df_sample_plots <-  modDat_1_s  %>% 
  slice_sample(n = 5e4) %>% 
   rename(response = all_of("TotalTreeCover_binom")) %>% 
  mutate(response = case_when(
    response == 0 ~ "0", 
    response > 0  ~ "1", 
  )) %>% 
  dplyr::select(c(response, prednames)) %>% 
  tidyr::pivot_longer(cols = unname(prednames), 
               names_to = "predictor", 
               values_to = "value"
               )  
 

  ggplot(df_sample_plots, aes_string(x= "response", y = 'value')) +
  geom_boxplot() +
  facet_wrap(~predictor , scales = 'free_y') + 
  ylab("Predictor Variable Values") + 
    xlab(response)

```


# Model Fitting

## Visualize the spatial blocks and how they differ across environmental space
First, if there are observations in Louisiana, sub-sample them so they're not so over-represented in the dataset
```{r}
## make data into spatial format
modDat_1_sf <- modDat_1_s %>% 
  st_as_sf(coords = c("x", "y"), crs = st_crs("EPSG:4326"))


# download map info for visualization
data(state_boundaries_wgs84) 

cropped_states <- suppressMessages(state_boundaries_wgs84 %>%
  dplyr::filter(NAME!="Hawaii") %>%
  dplyr::filter(NAME!="Alaska") %>%
  dplyr::filter(NAME!="Puerto Rico") %>%
  dplyr::filter(NAME!="American Samoa") %>%
  dplyr::filter(NAME!="Guam") %>%
  dplyr::filter(NAME!="Commonwealth of the Northern Mariana Islands") %>%
  dplyr::filter(NAME!="United States Virgin Islands") %>%
  sf::st_sf() %>%
  sf::st_transform(sf::st_crs(modDat_1_sf))) 

# if (ecoregion %in% c("Forest", "eastForest", "forest")){
# modDat_1_s$uniqueID <- 1:nrow(modDat_1_s)
# modDat_1_sf$uniqueID <- 1:nrow(modDat_1_sf)
# 
# # find which observations overlap with Louisiana
# obs_LA_temp <- st_intersects(modDat_1_sf, cropped_states[cropped_states$NAME == "Louisiana",], sparse = FALSE)
# if (sum(obs_LA_temp) > 0) {
#  obs_LA_1 <- modDat_1_sf[which(obs_LA_temp == TRUE, arr.ind = TRUE)[,1],]
# # now, find only those within the oversampled area (near the coast)
# dims <- c(xmin = 730439.1, xmax = 1042195.5, ymax = -1222745.2, ymin = -1390430.9)
# badBox <- st_bbox(dims) %>% 
#   st_as_sfc() %>% 
#   st_set_crs(value = st_crs(modDat_1_sf))
#   
# obs_LA_temp2 <- st_intersects(obs_LA_1, badBox, sparse = FALSE)
# obs_LA_2 <- obs_LA_1[which(obs_LA_temp2 == TRUE, arr.ind = TRUE)[,1],]
# 
# # subsample so there aren't so many 
# # get every 7th observation
# obs_LA_sampled <- obs_LA_2[seq(from = 1, to = nrow(obs_LA_2), by = 10),]
# # remove observations that aren't sampled from the larger dataset
# obsToRemove <- obs_LA_2[!(obs_LA_2$uniqueID %in% obs_LA_sampled$uniqueID),]
# 
# modDat_1_sf <- modDat_1_sf[!(modDat_1_sf$uniqueID %in% obsToRemove$uniqueID),]
# modDat_1_s <- modDat_1_s[!(modDat_1_s$uniqueID %in% obsToRemove$uniqueID),] 
# }
#}

```

```{r Visualize ecoregion distribution, fig.height=10, fig.width=14, message=FALSE, warning=FALSE}

## do a pca of climate across level 2 ecoregions
pca <- prcomp(modDat_1_s[,paste0(prednames)])
library(factoextra)
(fviz_pca_ind(pca, habillage = modDat_1_s$NA_L2NAME, label = "none", addEllipses = TRUE, ellipse.level = .95, ggtheme = theme_minimal(), alpha.ind = .1))

# if (ecoregion == "shrubGrass") {
#   print("We'll combine the 'Mediterranean California' and 'Western Sierra Madre Piedmont' ecoregions (into 'Mediterranean Piedmont'). We'll also combine `Tamaulipas-Texas semiarid plain,' 'Texas-Lousiana Coastal plain,' and 'South Central semiarid prairies' ecoregions (into (`Semiarid plain and prairies`)." )
# 
#   modDat_1_s[modDat_1_s$NA_L2NAME %in% c("MEDITERRANEAN CALIFORNIA", "WESTERN SIERRA MADRE PIEDMONT"), "NA_L2NAME"] <- "MEDITERRANEAN PIEDMONT"
# 
#   modDat_1_s[modDat_1_s$NA_L2NAME %in% c("TAMAULIPAS-TEXAS SEMIARID PLAIN", "TEXAS-LOUISIANA COASTAL PLAIN", "SOUTH CENTRAL SEMIARID PRAIRIES"), "NA_L2NAME"] <- "SEMIARID PLAIN AND PRAIRIES"
# 
#     modDat_1_s[modDat_1_s$NA_L2NAME %in% c("MARINE WEST COAST FOREST", "WESTERN CORDILLERA"), "NA_L2NAME"] <- "WESTERN CORDILLERA AND WEST COAST FOREST"
# 
#   modDat_1_s[modDat_1_s$NA_L2NAME %in% c("MEDITERRANEAN CALIFORNIA", "UPPER GILA MOUNTAINS"), "NA_L2NAME"] <- "MEDITERRANEAN CALIFORNIA AND UPPER GILA MOUNTAINS"
# 
# modDat_1_s[modDat_1_s$NA_L2NAME %in% c("MIXED WOOD PLAINS", "OZARK/OUACHITA-APPALACHIAN FORESTS"), "NA_L2NAME"] <- "OZARK/OUACHITA-APPALACHIAN FORESTS AND MIXED WOOD PLAINS"
# #////
# modDat_1_sf[modDat_1_sf$NA_L2NAME %in% c("MEDITERRANEAN CALIFORNIA", "WESTERN SIERRA MADRE PIEDMONT"), "NA_L2NAME"] <- "MEDITERRANEAN PIEDMONT"
# 
#   modDat_1_sf[modDat_1_sf$NA_L2NAME %in% c("TAMAULIPAS-TEXAS SEMIARID PLAIN", "TEXAS-LOUISIANA COASTAL PLAIN", "SOUTH CENTRAL SEMIARID PRAIRIES"), "NA_L2NAME"] <- "SEMIARID PLAIN AND PRAIRIES"
# 
#     modDat_1_sf[modDat_1_sf$NA_L2NAME %in% c("MARINE WEST COAST FOREST", "WESTERN CORDILLERA"), "NA_L2NAME"] <- "WESTERN CORDILLERA AND WEST COAST FOREST"
# 
#   modDat_1_sf[modDat_1_sf$NA_L2NAME %in% c("MEDITERRANEAN CALIFORNIA", "UPPER GILA MOUNTAINS"), "NA_L2NAME"] <- "MEDITERRANEAN CALIFORNIA AND UPPER GILA MOUNTAINS"
# 
# modDat_1_sf[modDat_1_sf$NA_L2NAME %in% c("MIXED WOOD PLAINS", "OZARK/OUACHITA-APPALACHIAN FORESTS"), "NA_L2NAME"] <- "OZARK/OUACHITA-APPALACHIAN FORESTS AND MIXED WOOD PLAINS"
#  if (response == "CAMCover") {
#    modDat_1_s[modDat_1_s$NA_L2NAME %in% c("MEDITERRANEAN PIEDMONT", "SEMIARID PLAIN AND PRAIRIES"), "NA_L2NAME"] <- "SEMIARID PLAIN AND PRAIRIES AND PIEDMONT"
#    modDat_1_sf[modDat_1_sf$NA_L2NAME %in% c("MEDITERRANEAN PIEDMONT", "SEMIARID PLAIN AND PRAIRIES"), "NA_L2NAME"] <- "SEMIARID PLAIN AND PRAIRIES AND PIEDMONT"
#  } else if (response %in% c("C4GramCover_prop", "C3GramCover_prop")) {
#      modDat_1_s[modDat_1_s$NA_L2NAME %in% c("SEMIARID PLAIN AND PRAIRIES", "TEMPERATE PRAIRIES"), "NA_L2NAME"] <- "SEMIARID AND TEMPERATE PRAIRIES"
#    modDat_1_sf[modDat_1_sf$NA_L2NAME %in% c("SEMIARID PLAIN AND PRAIRIES", "TEMPERATE PRAIRIES"), "NA_L2NAME"] <- "SEMIARID AND TEMPERATE PRAIRIES"
#  }
# 
# } else if (ecoregion == "CONUS") {
# 
#   modDat_1_s[modDat_1_s$NA_L2NAME %in% c("EVERGLADES", "MISSISSIPPI ALLUVIAL AND SOUTHEAST USA COASTAL PLAINS"), "NA_L2NAME"] <- "EVERGLADES MISSISSIPPI ALLUVIAL AND SOUTHEAST USA COASTAL PLAINS"
# 
#   modDat_1_s[modDat_1_s$NA_L2NAME %in% c("MARINE WEST COAST FOREST", "WESTERN CORDILLERA"), "NA_L2NAME"] <- "WESTERN CORDILLERA AND WEST COAST FOREST"
# 
#    modDat_1_s[modDat_1_s$NA_L2NAME %in% c("MIXED WOOD PLAINS", "OZARK/OUACHITA-APPALACHIAN FORESTS"), "NA_L2NAME"] <- "OZARK/OUACHITA-APPALACHIAN FORESTS AND MIXED WOOD PLAINS"
# 
#    modDat_1_s[modDat_1_s$NA_L2NAME %in% c("MEDITERRANEAN CALIFORNIA", "UPPER GILA MOUNTAINS"), "NA_L2NAME"] <- "MEDITERRANEAN CALIFORNIA AND UPPER GILA MOUNTAINS"
# 
#    modDat_1_s[modDat_1_s$NA_L2NAME %in% c("OZARK/OUACHITA-APPALACHIAN FORESTS AND MIXED WOOD PLAINS", "SOUTHEASTERN USA PLAINS",  "MISSISSIPPI ALLUVIAL AND SOUTHEAST USA COASTAL PLAINS", "EVERGLADES MISSISSIPPI ALLUVIAL AND SOUTHEAST USA COASTAL PLAINS"), "NA_L2NAME"] <- "SOUTHEASTERN AND MIXED WOOD PLAINS"
# 
#   modDat_1_s[modDat_1_s$NA_L2NAME %in% c("SOUTH CENTRAL SEMIARID PRAIRIES", "TEXAS-LOUISIANA COASTAL PLAIN"), "NA_L2NAME"] <- "SOUTH CENTRAL SEMIARID PRAIRIES"
#   #///
# 
#   modDat_1_sf[modDat_1_sf$NA_L2NAME %in% c("EVERGLADES", "MISSISSIPPI ALLUVIAL AND SOUTHEAST USA COASTAL PLAINS"), "NA_L2NAME"] <- "EVERGLADES MISSISSIPPI ALLUVIAL AND SOUTHEAST USA COASTAL PLAINS"
# 
#   modDat_1_sf[modDat_1_sf$NA_L2NAME %in% c("MARINE WEST COAST FOREST", "WESTERN CORDILLERA"), "NA_L2NAME"] <- "WESTERN CORDILLERA AND WEST COAST FOREST"
# 
#    modDat_1_sf[modDat_1_sf$NA_L2NAME %in% c("MIXED WOOD PLAINS", "OZARK/OUACHITA-APPALACHIAN FORESTS"), "NA_L2NAME"] <- "OZARK/OUACHITA-APPALACHIAN FORESTS AND MIXED WOOD PLAINS"
# 
#    modDat_1_sf[modDat_1_sf$NA_L2NAME %in% c("MEDITERRANEAN CALIFORNIA", "UPPER GILA MOUNTAINS"), "NA_L2NAME"] <- "MEDITERRANEAN CALIFORNIA AND UPPER GILA MOUNTAINS"
# 
#    modDat_1_sf[modDat_1_sf$NA_L2NAME %in% c("OZARK/OUACHITA-APPALACHIAN FORESTS AND MIXED WOOD PLAINS", "SOUTHEASTERN USA PLAINS",  "MISSISSIPPI ALLUVIAL AND SOUTHEAST USA COASTAL PLAINS", "EVERGLADES MISSISSIPPI ALLUVIAL AND SOUTHEAST USA COASTAL PLAINS"), "NA_L2NAME"] <- "SOUTHEASTERN AND MIXED WOOD PLAINS"
# 
#   modDat_1_sf[modDat_1_sf$NA_L2NAME %in% c("SOUTH CENTRAL SEMIARID PRAIRIES", "TEXAS-LOUISIANA COASTAL PLAIN"), "NA_L2NAME"] <- "SOUTH CENTRAL SEMIARID PRAIRIES"
# 
#   if (response %in% c("C4GramCover_prop")) {
#     modDat_1_s[modDat_1_s$NA_L2NAME %in% c("CENTRAL USA PLAINS", "TEMPERATE PRAIRIES", "SOUTHEASTERN AND MIXED WOOD PLAINS", "ATLANTIC HIGHLANDS", "MIXED WOOD SHIELD"), "NA_L2NAME"] <- "EASTERN AND MIXED WOOD PLAINS AND FOREST"
#   #///
# 
#   modDat_1_sf[modDat_1_sf$NA_L2NAME %in%c("CENTRAL USA PLAINS", "TEMPERATE PRAIRIES", "SOUTHEASTERN AND MIXED WOOD PLAINS", "ATLANTIC HIGHLANDS", "MIXED WOOD SHIELD"), "NA_L2NAME"] <- "EASTERN AND MIXED WOOD PLAINS AND FOREST"
#   }
# } else if (ecoregion == "forest"  & response != "CAMCover") {
#    modDat_1_s[modDat_1_s$NA_L2NAME %in% c("MISSISSIPPI ALLUVIAL AND SOUTHEAST USA COASTAL PLAINS", "EVERGLADES"), "NA_L2NAME"] <- "MISSISSIPPI ALLUVIAL AND SOUTHEAST USA COASTAL PLAINS	AND EVERGLADES"
# 
#    modDat_1_s[modDat_1_s$NA_L2NAME %in% c("UPPER GILA MOUNTAINS", "WESTERN CORDILLERA"), "NA_L2NAME"] <- "WESTERN CORDILLERA AND UPPER GILA MOUNTAINS"
# 
#    modDat_1_s[modDat_1_s$NA_L2NAME %in% c("MISSISSIPPI ALLUVIAL AND SOUTHEAST USA COASTAL PLAINS\tAND EVERGLADES", "SOUTHEASTERN USA PLAINS"), "NA_L2NAME"] <- "SOUTHEASTERN USA PLAINS"
# 
#    modDat_1_s[modDat_1_s$NA_L2NAME %in% c("ATLANTIC HIGHLANDS", "OZARK/OUACHITA-APPALACHIAN FORESTS"), "NA_L2NAME"] <- "HIGHLANDS AND APPALACHIAN FORESTS"
# 
#    modDat_1_s[modDat_1_s$NA_L2NAME %in% c("CENTRAL USA PLAINS", "MIXED WOOD PLAINS"), "NA_L2NAME"] <- "CENTRAL AND MIXED WOOD PLAINS"
# 
#    modDat_1_s[modDat_1_s$NA_L2NAME %in% c("MIXED WOOD SHIELD", "CENTRAL AND MIXED WOOD PLAINS"), "NA_L2NAME"] <- "CENTRAL AND MIXED WOOD PLAINS AND MIXED WOOD SHIELD"
# 
#        ## divide southeastern US plains into two regions, since it's by far the largest group
#   modDat_1_s[modDat_1_s$NA_L2NAME == "WESTERN CORDILLERA AND UPPER GILA MOUNTAINS" &
#                modDat_1_s$Long < -966595#-1773969
#                , "NA_L2NAME"] <- "WESTERN CORDILLERA AND UPPER GILA MOUNTAINS 1"
#   modDat_1_s[modDat_1_s$NA_L2NAME == "WESTERN CORDILLERA AND UPPER GILA MOUNTAINS", "NA_L2NAME"] <- "WESTERN CORDILLERA AND UPPER GILA MOUNTAINS 2"
#    #///
#    modDat_1_sf[modDat_1_sf$NA_L2NAME %in% c("MISSISSIPPI ALLUVIAL AND SOUTHEAST USA COASTAL PLAINS", "EVERGLADES"), "NA_L2NAME"] <- "MISSISSIPPI ALLUVIAL AND SOUTHEAST USA COASTAL PLAINS	AND EVERGLADES"
# 
#    modDat_1_sf[modDat_1_sf$NA_L2NAME %in% c("UPPER GILA MOUNTAINS", "WESTERN CORDILLERA"), "NA_L2NAME"] <- "WESTERN CORDILLERA AND UPPER GILA MOUNTAINS"
# 
#    modDat_1_sf[modDat_1_sf$NA_L2NAME %in% c("MISSISSIPPI ALLUVIAL AND SOUTHEAST USA COASTAL PLAINS\tAND EVERGLADES", "SOUTHEASTERN USA PLAINS"), "NA_L2NAME"] <- "SOUTHEASTERN USA PLAINS"
# 
#    modDat_1_sf[modDat_1_sf$NA_L2NAME %in% c("ATLANTIC HIGHLANDS", "OZARK/OUACHITA-APPALACHIAN FORESTS"), "NA_L2NAME"] <- "HIGHLANDS AND APPALACHIAN FORESTS"
# 
#    modDat_1_sf[modDat_1_sf$NA_L2NAME %in% c("CENTRAL USA PLAINS", "MIXED WOOD PLAINS"), "NA_L2NAME"] <- "CENTRAL AND MIXED WOOD PLAINS"
# 
#    modDat_1_sf[modDat_1_sf$NA_L2NAME %in% c("MIXED WOOD SHIELD", "CENTRAL AND MIXED WOOD PLAINS"), "NA_L2NAME"] <- "CENTRAL AND MIXED WOOD PLAINS AND MIXED WOOD SHIELD"
#           ## divide southeastern US plains into two regions, since it's by far the largest group
#   modDat_1_sf[modDat_1_sf$NA_L2NAME == "WESTERN CORDILLERA AND UPPER GILA MOUNTAINS" &
#               st_coordinates(modDat_1_sf)[,1] < -966595#-1773969
#                 , ]$NA_L2NAME <- "WESTERN CORDILLERA AND UPPER GILA MOUNTAINS 1"
#   modDat_1_sf[modDat_1_sf$NA_L2NAME == "WESTERN CORDILLERA AND UPPER GILA MOUNTAINS", "NA_L2NAME"] <- "WESTERN CORDILLERA AND UPPER GILA MOUNTAINS 2"
# 
#   if (response %in% c("C3GramCover_prop", "C4GramCover_prop") ) {
#      modDat_1_s[modDat_1_s$NA_L2NAME %in% c("MARINE WEST COAST FOREST", "WESTERN CORDILLERA AND UPPER GILA MOUNTAINS 1"), "NA_L2NAME"] <- "WESTERN CORDILLERA AND UPPER GILA MOUNTAINS 1"
#      modDat_1_s[modDat_1_s$NA_L2NAME %in% c("CENTRAL AND MIXED WOOD PLAINS AND MIXED WOOD SHIELD", "HIGHLANDS AND APPALACHIAN FORESTS"), "NA_L2NAME"] <- "CENTRAL AND MIXED WOODS AND HIGHLANDS FORESTS"
#      #//
#   modDat_1_sf[modDat_1_sf$NA_L2NAME %in% c("MARINE WEST COAST FOREST", "WESTERN CORDILLERA AND UPPER GILA MOUNTAINS 1"), "NA_L2NAME"] <- "WESTERN CORDILLERA AND UPPER GILA MOUNTAINS 1"
#   modDat_1_sf[modDat_1_sf$NA_L2NAME %in% c("CENTRAL AND MIXED WOOD PLAINS AND MIXED WOOD SHIELD", "HIGHLANDS AND APPALACHIAN FORESTS"), "NA_L2NAME"] <- "CENTRAL AND MIXED WOODS AND HIGHLANDS FORESTS"
# 
#   }
# 
# } else if (ecoregion == "forest" & response == "CAMCover") {
# 
#   modDat_1_s[modDat_1_s$NA_L2NAME %in% c("OZARK/OUACHITA-APPALACHIAN FORESTS", "SOUTHEASTERN USA PLAINS"), "NA_L2NAME"] <- "SOUTHEASTERN PLAINS AND APPALACHIAN FORESTS"
# 
#   modDat_1_s[modDat_1_s$NA_L2NAME %in% c("MARINE WEST COAST FOREST", "WESTERN CORDILLERA"), "NA_L2NAME"] <- "MARINE WEST COAST AND WESTERN CORDILLERA"
# 
#   modDat_1_s[modDat_1_s$NA_L2NAME %in% c("MIXED WOOD PLAINS", "SOUTHEASTERN PLAINS AND APPALACHIAN FORESTS"), "NA_L2NAME"] <- "SOUTHEASTERN PLAINS AND APPALACHIAN FORESTS"
# 
#   #///
#    modDat_1_sf[modDat_1_sf$NA_L2NAME %in% c("OZARK/OUACHITA-APPALACHIAN FORESTS", "SOUTHEASTERN USA PLAINS"), "NA_L2NAME"] <- "SOUTHEASTERN PLAINS AND APPALACHIAN FORESTS"
# 
#   modDat_1_sf[modDat_1_sf$NA_L2NAME %in% c("MARINE WEST COAST FOREST", "WESTERN CORDILLERA"), "NA_L2NAME"] <- "MARINE WEST COAST AND WESTERN CORDILLERA"
# 
#   modDat_1_sf[modDat_1_sf$NA_L2NAME %in% c("MIXED WOOD PLAINS", "SOUTHEASTERN PLAINS AND APPALACHIAN FORESTS"), "NA_L2NAME"] <- "SOUTHEASTERN PLAINS AND APPALACHIAN FORESTS"
# 
# } else if (ecoregion == "eastForest") {
#   modDat_1_s[modDat_1_s$NA_L2NAME %in% c("EVERGLADES", "MISSISSIPPI ALLUVIAL AND SOUTHEAST USA COASTAL PLAINS"), "NA_L2NAME"] <- "EVERGLADES MISSISSIPPI ALLUVIAL AND SOUTHEAST USA COASTAL PLAINS"
# 
#   modDat_1_s[modDat_1_s$NA_L2NAME %in% c("MIXED WOOD PLAINS","CENTRAL USA PLAINS"), "NA_L2NAME"] <- "SOUTHEASTERN AND CENTRAL USA PLAINS"
# 
#   modDat_1_s[modDat_1_s$NA_L2NAME %in% c("MIXED WOOD SHIELD", "ATLANTIC HIGHLANDS"), "NA_L2NAME"] <- "ATLANTIC HIGHLANDS AND MIXED WOOD SHIELD"
# 
#   modDat_1_s[modDat_1_s$NA_L2NAME %in% c("MIXED WOOD PLAINS", "ATLANTIC HIGHLANDS AND MIXED WOOD SHIELD"), "NA_L2NAME"] <- "ATLANTIC HIGHLANDS AND MIXED WOOD SHIELD AND PLAINS"
# 
#   modDat_1_s[modDat_1_s$NA_L2NAME %in% c("SOUTHEASTERN AND CENTRAL USA PLAINS", "ATLANTIC HIGHLANDS AND MIXED WOOD SHIELD AND PLAINS"), "NA_L2NAME"] <- "PLAINS AND HIGHLANDS AND SHIELD"
# 
#   modDat_1_s[modDat_1_s$NA_L2NAME %in% c("EVERGLADES MISSISSIPPI ALLUVIAL AND SOUTHEAST USA COASTAL PLAINS", "SOUTHEASTERN USA PLAINS"), "NA_L2NAME"] <- "SOUTHEASTERN PLAINS AND COAST"
# 
#   # #////
#    modDat_1_sf[modDat_1_sf$NA_L2NAME %in% c("EVERGLADES", "MISSISSIPPI ALLUVIAL AND SOUTHEAST USA COASTAL PLAINS"), "NA_L2NAME"] <- "EVERGLADES MISSISSIPPI ALLUVIAL AND SOUTHEAST USA COASTAL PLAINS"
# 
#    modDat_1_sf[modDat_1_sf$NA_L2NAME %in% c("MIXED WOOD PLAINS","CENTRAL USA PLAINS"), "NA_L2NAME"] <- "SOUTHEASTERN AND CENTRAL USA PLAINS"
# 
#    modDat_1_sf[modDat_1_sf$NA_L2NAME %in% c("MIXED WOOD SHIELD", "ATLANTIC HIGHLANDS"), "NA_L2NAME"] <- "ATLANTIC HIGHLANDS AND MIXED WOOD SHIELD"
# 
#    modDat_1_sf[modDat_1_sf$NA_L2NAME %in% c("MIXED WOOD PLAINS", "ATLANTIC HIGHLANDS AND MIXED WOOD SHIELD"), "NA_L2NAME"] <- "ATLANTIC HIGHLANDS AND MIXED WOOD SHIELD AND PLAINS"
# 
#    modDat_1_sf[modDat_1_sf$NA_L2NAME %in% c("SOUTHEASTERN AND CENTRAL USA PLAINS", "ATLANTIC HIGHLANDS AND MIXED WOOD SHIELD AND PLAINS"), "NA_L2NAME"] <- "PLAINS AND HIGHLANDS AND SHIELD"
# 
#    modDat_1_sf[modDat_1_sf$NA_L2NAME %in% c("EVERGLADES MISSISSIPPI ALLUVIAL AND SOUTHEAST USA COASTAL PLAINS", "SOUTHEASTERN USA PLAINS"), "NA_L2NAME"] <- "SOUTHEASTERN PLAINS AND COAST"
# 
#    ## divide southeastern US plains into two regions, since it's by far the largest group
#   # modDat_1_s[modDat_1_s$NA_L2NAME == "SOUTHEASTERN USA PLAINS" &
#   #              modDat_1_s$Lat < -590062, "NA_L2NAME"] <- "SOUTHEASTERN USA PLAINS 1"
#   # modDat_1_s[modDat_1_s$NA_L2NAME == "SOUTHEASTERN USA PLAINS", #&
#   #             # modDat_1_s$Lat < -590062,
#   #            "NA_L2NAME"] <- "SOUTHEASTERN USA PLAINS 2"
# 
#   #   ## divide southeastern US plains into two regions, since it's by far the largest group
#   # modDat_1_s[modDat_1_s$NA_L2NAME == "OZARK/OUACHITA-APPALACHIAN FORESTS" &
#   #              modDat_1_s$Long < 854862.2, "NA_L2NAME"] <- "OZARK/OUACHITA-APPALACHIAN FORESTS 1"
#   # modDat_1_s[modDat_1_s$NA_L2NAME == "OZARK/OUACHITA-APPALACHIAN FORESTS" &
#   #              modDat_1_s$Long  >=   854862.2, "NA_L2NAME"] <- "OZARK/OUACHITA-APPALACHIAN FORESTS 2"
# }
# make a table of n for each region

modDat_1_s %>% 
  group_by(NA_L2NAME) %>% 
  dplyr::summarize("Number_Of_Observations" = length(NA_L2NAME)) %>% 
  rename("Level_2_Ecoregion" = NA_L2NAME)%>% 
  kable() %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed")) 

```



## Then, look at the spatial distribution and environmental characteristics of the grouped ecoregions 
```{r fig.width = 14, fig.height=10, message=FALSE}

map1 <- ggplot() +
  geom_sf(data=cropped_states,fill='white') +
  geom_sf(data=modDat_1_sf#[modDat_1_sf$NA_L2NAME %in% c("MIXED WOOD PLAINS"),]
          ,
          aes(fill=as.factor(NA_L2NAME)),linewidth=0.5,alpha=0.5) +
  geom_point(data=modDat_1_s#[modDat_1_s$NA_L2NAME %in% c("MIXED WOOD PLAINS"),]
             ,
             alpha=0.5, 
             aes(x = x, y = y, color=as.factor(NA_L2NAME)), alpha = .1) +
  #scale_fill_okabeito() +
  #scale_color_okabeito() +
 # theme_default() +
  theme(legend.position = 'none') +
  labs(title = "Level 2 Ecoregions as spatial blocks")

hull <- modDat_1_sf %>%
  ungroup() %>%
  group_by(NA_L2NAME) %>%
  slice(chull(tmean, prcp))

plot1<-ggplot(data=modDat_1_sf,aes(x=tmean,y=prcp)) +
  geom_polygon(data = hull, alpha = 0.25,aes(fill=NA_L2NAME) )+
  geom_point(aes(group=NA_L2NAME,color=NA_L2NAME),alpha=0.25) +
  theme_minimal() + xlab("Annual Average T_mean - long-term average") +
  ylab("Annual Average Precip - long-term average") #+
  #scale_color_okabeito() +
  #scale_fill_okabeito()

plot2<-ggplot(data=modDat_1_sf %>%
                pivot_longer(cols=tmean:prcp),
              aes(x=value,group=name)) +
  # geom_polygon(data = hull, alpha = 0.25,aes(fill=fold) )+
  geom_density(aes(group=NA_L2NAME,fill=NA_L2NAME),alpha=0.25) +
  theme_minimal() +
  facet_wrap(~name,scales='free')# +
  #scale_color_okabeito() +
  #scale_fill_okabeito()
 
library(patchwork)
(combo <- (map1+plot1)/plot2) 
# 
# ggplot(data = modDat_1_s) +
#   geom_density(aes(ShrubCover_transformed, col = NA_L2NAME)) +
#   xlim(c(0,100))

```


## Fit a global model with all of the data 

### First, fit a LASSO regression model using the glmnet R package
- This regression is a beta glm with a logit link (using custom function from Daniel)
- Use cross validation across level 2 ecoregions to tune the lambda parameter in the LASSO model 
- this model is fit to using the scaled weather/climate/soils variables
- this list of possible predictors includes: 
  1. main effects
  2. interactions between all soils variables
  3. interactions between climate and weather variables
  4. transformed main effects (squared, log-transformed (add a uniform integer -- 20-- to all variables prior to log-transformation), square root-transformed (add a uniform integer -- 20-- to all variables prior to log-transformation))
```{r fit global model, echo = FALSE}
  # penalty values
  lambdas <- 10^seq(1,-5, length.out = 200)  # sequence of penalties to test
 
if (removeAllAnoms) { ## if the model parameters indicate that we don't want any anomalies... 
    # make model matrix (which includes transformations and interactions)
  ## only include interactions within and between climate and anomaly variables, and within and between soils variables

 ## break predictors into climate, weather, and soils 
  prednames_clim <- prednames[which(prednames %in% c("tmin" , "tmax",  "tmean",
     "prcp", "t_warm" , "t_cold" , "prcp_wet" , "prcp_dry" , "prcp_seasonality", "prcpTempCorr", "abvFreezingMonth" , "isothermality" , "annWatDef" , "annWetDegDays",  "VPD_mean",  "VPD_max",  "VPD_min" , "VPD_max_95", "annWatDef_95", "annWetDegDays_5",  "frostFreeDays_5", "frostFreeDays"))]

  prednames_soils <- prednames[which(prednames %in% c( "soilDepth" , "clay", 
     "sand" , "coarse" ,  "carbon", "AWHC"))]
   prednames_weath <- prednames[which(prednames %in% c("tmean_anom", "tmin_anom",  "tmax_anom", 
    "prcp_anom",      "t_warm_anom" , "t_cold_anom",  "prcp_wet_anom" , "precp_dry_anom" ,  "prcp_seasonality_anom" ,  "prcpTempCorr_anom" , "aboveFreezingMonth_anom",  "isothermality_anom" ,  "annWatDef_anom", "annWetDegDays_anom",  "VPD_mean_anom" , "VPD_min_anom" , "VPD_max_anom", "VPD_max_95_anom", "annWatDef_95_anom" , "annWetDegDays_5_anom",  "frostFreeDays_5_anom" , "frostFreeDays_anom"))]
  
  # get a text string of possible interactions between climate and weather variables
  combos <- data.frame(gtools::permutations(n = length(c(prednames_clim)),
                                            r = 2,
                                            v = c(prednames_clim)))
  
  combos$interactions <- paste0(combos$X1, "*", combos$X2)
  
  
  # get a text string of possible interactions between soils variables
  combos_2 <- data.frame(gtools::permutations(n = length(c(prednames_soils)),
                                            r = 2,
                                            v = c(prednames_soils)))
  
  combos_2$interactions <- paste0(combos_2$X1, "*", combos_2$X2)
  
  # make text string of all possible interactions
  int_string <-str_flatten(c(combos$interactions, combos_2$interactions), collapse = " + ")
  
  # make text string of transformations (log, squared)
  transform_string <- str_flatten(c(# add square root term 
                                    paste0("I(log(",prednames_clim,"+30))"), 
                                    paste0("I(log(",prednames_soils,"+30))"),
                                    
                                    paste0("I(sqrt(",prednames_clim,"+30))"), 
                                    paste0("I(sqrt(",prednames_soils,"+30))"),
                                    
                                    paste0("I(",prednames_clim,"^2)"), 
                                    paste0("I(",prednames_soils,"^2)"))
                                    , collapse = " + ")
              
  # get a text string of the possible untransformed model terms
  mod_string <- paste("TotalTreeCover_binom", "~", paste0(c(prednames_clim, prednames_soils), collapse = "+"))
  
  # make a text string of the model formula w/ interactions and transformations included
  modWithInteractions <- paste(mod_string, "+",int_string , "+", transform_string
                               )
   
} else { ## if the model parameters indicate that we DO want anomalies... 
    # make model matrix (which includes transformations and interactions)
  ## only include interactions within and between climate and anomaly variables, and within and between soils variables

 ## break predictors into climate, weather, and soils 
  prednames_clim <- prednames[which(prednames %in% c("tmin" , "tmax",  "tmean",
     "prcp", "t_warm" , "t_cold" , "prcp_wet" , "prcp_dry" , "prcp_seasonality", "prcpTempCorr", "abvFreezingMonth" , "isothermality" , "annWatDef" , "annWetDegDays",  "VPD_mean",  "VPD_max",  "VPD_min" , "VPD_max_95", "annWatDef_95", "annWetDegDays_5",  "frostFreeDays_5", "frostFreeDays"))]

  prednames_weath <- prednames[which(prednames %in% c("tmean_anom", "tmin_anom",  "tmax_anom", 
    "prcp_anom",      "t_warm_anom" , "t_cold_anom",  "prcp_wet_anom" , "precp_dry_anom" ,  "prcp_seasonality_anom" ,  "prcpTempCorr_anom" , "aboveFreezingMonth_anom",  "isothermality_anom" ,  "annWatDef_anom", "annWetDegDays_anom",  "VPD_mean_anom" , "VPD_min_anom" , "VPD_max_anom", "VPD_max_95_anom", "annWatDef_95_anom" , "annWetDegDays_5_anom",  "frostFreeDays_5_anom" , "frostFreeDays_anom"))]
  
  prednames_soils <- prednames[which(prednames %in% c( "soilDepth" , "clay", 
     "sand" , "coarse" ,  "carbon", "AWHC"))]
  
  
  # get a text string of possible interactions between climate and weather variables
  combos <- data.frame(gtools::permutations(n = length(c(prednames_clim, prednames_weath)),
                                            r = 2,
                                            v = c(prednames_clim, prednames_weath)))
  
  combos$interactions <- paste0(combos$X1, "*", combos$X2)
  
  
  # get a text string of possible interactions between soils variables
  combos_2 <- data.frame(gtools::permutations(n = length(c(prednames_soils)),
                                            r = 2,
                                            v = c(prednames_soils)))
  
  combos_2$interactions <- paste0(combos_2$X1, "*", combos_2$X2)
  
  # make text string of all possible interactions
  int_string <-str_flatten(c(combos$interactions, combos_2$interactions), collapse = " + ")
  
  # make text string of transformations (log, squared)
  transform_string <- str_flatten(c(# add square root term 
                                    paste0("I(log(",prednames_clim,"+30))"), 
                                    paste0("I(log(",prednames_weath,"+30))"),
                                    paste0("I(log(",prednames_soils,"+30))"),
                                    
                                    paste0("I(sqrt(",prednames_clim,"+30))"), 
                                    paste0("I(sqrt(",prednames_weath,"+30))"),
                                    paste0("I(sqrt(",prednames_soils,"+30))"),
                                    
                                    paste0("I(",prednames_clim,"^2)"), 
                                    paste0("I(",prednames_weath,"^2)"),
                                    paste0("I(",prednames_soils,"^2)"))
                                    , collapse = " + ")
              
  # get a text string of the possible untransformed model terms
  mod_string <- paste("TotalTreeCover_binom", "~", paste0(prednames, collapse = "+"))
  
  # make a text string of the model formula w/ interactions and transformations included
  modWithInteractions <- paste(mod_string, "+",int_string , "+", transform_string
                               )
   
}


  # now, make a model formula object
  f <- as.formula(modWithInteractions)
  
  # transform dataframe to model matrix
  X <- model.matrix(f, modDat_1_s)
  # get response variable
  y <- as.matrix(modDat_1_s[,"TotalTreeCover_binom"])

```
Get rid of transformed predictions and interactions that are correlated 
```{r}
# get first pass of names correlated variables
X_df <- X %>% 
  as.data.frame() %>% 
  dplyr::select(-'(Intercept)')  
corrNames_i <- X_df %>% 
  cor()  %>% 
   caret::findCorrelation(cutoff = .7, verbose = FALSE, names = TRUE, exact = TRUE)
# remove those names that are untransformed main effects 
  # vector of columns to remove 
badNames <- corrNames_i[!(corrNames_i %in% prednames)]
while (sum(!(corrNames_i %in% prednames))>0) {
 corrNames_i <-  X_df %>% 
    dplyr::select(-badNames) %>% 
     cor()  %>% 
   caret::findCorrelation(cutoff = .7, verbose = FALSE, names = TRUE, exact = TRUE)
 # update the vector of names to remove 
 badNames <- unique(c(badNames, corrNames_i[!(corrNames_i %in% prednames)]))
}

## see if there are any correlated variables left (would be all main effects at this point)
# if there are, step through and remove the variable that each is most correlated with 
if (length(corrNames_i)>1) {
  for (i in 1:length(corrNames_i)) {
    X_i <- X_df %>% 
      dplyr::select(-badNames)
    if (corrNames_i[i] %in% names(X_i)) {
    corMat_i <- cor(x = X_i[corrNames_i[i]], y = X_i %>% dplyr::select(-corrNames_i[i])) 
    badNames_i <- colnames(corMat_i)[abs(corMat_i)>=.7]
    # if there are any predictors in the 'badNames_i', remove them from this list
    if (length(badNames_i) > 0 & sum(c(badNames_i %in% prednames))>0) {
        badNames_i <- badNames_i[!(badNames_i %in% prednames)]
    }
    badNames <- unique(c(badNames, badNames_i))
    }
  }
}
## update the X matrix to exclude these correlated variables
X <- X[,!(colnames(X) %in% badNames)]
```

```{r}
if (run == TRUE) {
  # set up custom folds
    # get the ecoregions for training lambda
  train_eco <- modDat_1_s$NA_L2NAME#[train]
  
  # Fit model -----------------------------------------------
  # specify leave-one-year-out cross-validation
  my_folds <- as.numeric(as.factor(train_eco))

    # set up parallel processing
    library(doMC)
    # this computer has 16 cores (parallel::detectCores())
    registerDoMC(cores = 8)
    
    fit <- cv.glmnet(
      x = X[,2:ncol(X)], 
      y = y, 
      family = "binomial",
      keep = FALSE,
      alpha = 1,  # 0 == ridge regression, 1 == lasso, 0.5 ~~ elastic net
      lambda = lambdas,
      relax = ifelse(response == "ShrubCover", yes = TRUE, no = FALSE),
      #nlambda = 100,
      type.measure="mse",
      #penalty.factor = pen_facts,
      foldid = my_folds,
      #thresh = thresh,
      standardize = FALSE, ## scales variables prior to the model sequence... coefficients are always returned on the original scale
      parallel = TRUE
    )
    base::saveRDS(fit, paste0("../ModelFitting/models/yesOrNoTrees_globalLASSOmod_binomial.rds"))
    
  
     best_lambda <- fit$lambda.min
  # save the lambda for the most regularized model w/ an MSE that is still 1SE w/in the best lambda model
  lambda_1SE <- fit$lambda.1se
  # save the lambda for the most regularized model w/ an MSE that is still .5SE w/in the best lambda model
  lambda_halfSE <- best_lambda + ((lambda_1SE - best_lambda)/2)
 
## Now, we need to do stability selection to ensure the coefficients that are being chosen with each lambda are stable 

## stability selection for best lambda model 
# setup params
p <- ncol(X[,2:ncol(X)]) # of parameters
n <- length(y) # of observations
n_iter <- 100        # number of subsamples
sample_frac <- 0.75  # fraction of data to subsample
lambda_val <- best_lambda    # fixed lambda value (could be chosen via CV)

# Track selection
selection_counts_bestL <- matrix(0, nrow = p, ncol = 1)

for (i in 1:n_iter) {
  # Subsample rows
  sample_idx <- sample(1:n, size = floor(sample_frac * n), replace = FALSE)
  X_sub <- X[sample_idx, ]
  y_sub <- y[sample_idx]

  # Fit Lasso model
  fit_stab_i <- glmnet(x = X_sub[,2:ncol(X_sub)], y = y_sub, 
    family = "binomial",
    alpha = 1, lambda = lambda_val, standardize = FALSE)

  # Get non-zero coefficients (excluding intercept)
  select_bestL <- which(as.vector(coef(fit_stab_i))[-1] != 0)
  selection_counts_bestL[select_bestL] <- selection_counts_bestL[select_bestL] + 1
}

# Convert counts to selection probabilities (the probability that a variable is selected over 100 iterations)
selection_prob_bestL <- selection_counts_bestL / n_iter
selection_prob_bestL_df <- data.frame(
  VariableNumber = paste0("X", 1:p),
  VariableName = rownames(coef(fit_stab_i))[2:(p+1)],
  SelectionProb = as.vector(selection_prob_bestL)
)

# get those variables that are selected in ≥70% of subsets (Meinshausen and Bühlmann, 2010)
bestLambda_coef <- selection_prob_bestL_df[selection_prob_bestL_df$SelectionProb>=.7, c("VariableName", "SelectionProb")]

#//////
# stability selection for 1se lambda model
lambda_val <-  lambda_1SE    # fixed lambda value (could be chosen via CV)

# Track selection
selection_counts_1seL <- matrix(0, nrow = p, ncol = 1)

for (i in 1:n_iter) {
  # Subsample rows
  sample_idx <- sample(1:n, size = floor(sample_frac * n), replace = FALSE)
  X_sub <- X[sample_idx, ]
  y_sub <- y[sample_idx]

  # Fit Lasso model
  fit_stab_i <- glmnet(x = X_sub[,2:ncol(X_sub)], y = y_sub, 
    family = "binomial",
    alpha = 1, lambda = lambda_val, standardize = FALSE)

  # Get non-zero coefficients (excluding intercept)
  selected_1seL <- which(as.vector(coef(fit_stab_i))[-1] != 0)
  selection_counts_1seL[selected_1seL] <- selection_counts_1seL[selected_1seL] + 1
}

# Convert counts to selection probabilities (the probability that a variable is selected over 100 iterations)
selection_prob_1seL <- selection_counts_1seL / n_iter
selection_prob_1seL_df <- data.frame(
  VariableNumber = paste0("X", 1:p),
  VariableName = rownames(coef(fit_stab_i))[2:(p+1)],
  SelectionProb = as.vector(selection_prob_1seL)
)

# get those variables that are selected in ≥70% of subsets (Meinshausen and Bühlmann, 2010)
seLambda_coef <- selection_prob_1seL_df[selection_prob_1seL_df$SelectionProb>=.7, c("VariableName", "SelectionProb")]

# stability selection for half se lambda model
lambda_val <- lambda_halfSE    # fixed lambda value (could be chosen via CV)

# Track selection
selection_counts_halfseL <- matrix(0, nrow = p, ncol = 1)

for (i in 1:n_iter) {
  # Subsample rows
  sample_idx <- sample(1:n, size = floor(sample_frac * n), replace = FALSE)
  X_sub <- X[sample_idx, ]
  y_sub <- y[sample_idx]

  # Fit Lasso model
  fit_stab_i <- glmnet(x = X_sub[,2:ncol(X_sub)], y = y_sub, 
    family = "binomial",
    alpha = 1, lambda = lambda_val, standardize = FALSE)

  # Get non-zero coefficients (excluding intercept)
  selected_halfseL <- which(as.vector(coef(fit_stab_i))[-1] != 0)
  selection_counts_halfseL[selected_halfseL] <- selection_counts_halfseL[selected_halfseL] + 1
}

# Convert counts to selection probabilities (the probability that a variable is selected_halfseL over 100 iterations)
selection_prob_halfseL <- selection_counts_halfseL / n_iter
selection_prob_halfseL_df <- data.frame(
  VariableNumber = paste0("X", 1:p),
  VariableName = rownames(coef(fit_stab_i))[2:(p+1)],
  SelectionProb = as.vector(selection_prob_halfseL)
)

# get those variables that are selected_halfseL_halfseL in ≥70% of subsets (Meinshausen and Bühlmann, 2010)
halfseLambda_coef <- selection_prob_halfseL_df[selection_prob_halfseL_df$SelectionProb>=.7, c("VariableName", "SelectionProb")]

## fit w/ the identified coefficients from the 'best' lambda, but using the glm function
  mat_glmnet_best <- bestLambda_coef$VariableName 

  if (length(mat_glmnet_best) == 0) {
    f_glm_bestLambda <- as.formula(paste0("TotalTreeCover_binom", "~ 1"))
  } else {
  f_glm_bestLambda <- as.formula(paste0("TotalTreeCover_binom", " ~ ", paste0(mat_glmnet_best, collapse = " + ")))
  }
  
## fit using betareg
  fit_glm_bestLambda <- fit_glm_bestLambda_binomial <- glm(formula = f_glm_bestLambda, data = modDat_1_s, family = binomial)
    
   ## fit w/ the identified coefficients from the '1se' lambda, but using the glm function
  mat_glmnet_1se <- seLambda_coef$VariableName
    
  if (length(mat_glmnet_1se) == 0) {
    f_glm_1se <- as.formula(paste0("TotalTreeCover_binom", "~ 1"))
  } else {
  f_glm_1se <- as.formula(paste0("TotalTreeCover_binom", " ~ ", paste0(mat_glmnet_1se, collapse = " + ")))
  }


  fit_glm_se <- glm(formula = f_glm_1se, data = modDat_1_s, family = binomial)
  # glm(data = modDat_1_s, formula = f_glm_1se,
  #                   family =  stats::Gamma(link = "log"))
  
     ## fit w/ the identified coefficients from the '.5se' lambda, but using the glm function
  mat_glmnet_halfse <- halfseLambda_coef$VariableName
  
  if (length(mat_glmnet_halfse) == 0) {
    f_glm_halfse <- as.formula(paste0("TotalTreeCover_binom", "~ 1"))
  } else {
  f_glm_halfse <- as.formula(paste0("TotalTreeCover_binom", " ~ ", paste0(mat_glmnet_halfse, collapse = " + ")))
  }

  fit_glm_halfse <- glm(formula = f_glm_halfse, data = modDat_1_s, family = binomial )
  
  ## save models 
  saveRDS(fit_glm_bestLambda, paste0("./models/yesOrNoTrees_bestLambdaGLM.rds"))
  saveRDS(fit_glm_halfse, paste0("./models/yesOrNoTrees_halfSELambdaGLM.rds"))
  saveRDS(fit_glm_se, paste0("./models/yesOrNoTrees_oneSELambdaGLM.rds"))
    
  ## save the R environment after running the models 
  save(f_glm_halfse, mat_glmnet_halfse, halfseLambda_coef,
              f_glm_1se, mat_glmnet_1se, seLambda_coef,
              f_glm_bestLambda, mat_glmnet_best, bestLambda_coef,
              file = paste0("./models/interimModelFittingObjects_yesOrNoTrees_binomial.rds" ))
  } else {
    # read in LASSO object
    fit <- readRDS("../ModelFitting/models/yesOrNoTrees_globalLASSOmod_binomial.rds")
    
    # read in R objects having to do w/ model fitting 
    load(file = paste0("./models/interimModelFittingObjects_yesOrNoTrees_binomial.rds"))
      
  fit_glm_bestLambda <- readRDS("./models/yesOrNoTrees_bestLambdaGLM.rds")
  fit_glm_halfse <- readRDS("./models/yesOrNoTrees_halfSELambdaGLM.rds")
  fit_glm_se <- readRDS("./models/yesOrNoTrees_oneSELambdaGLM.rds")
  }


  # assess model fit
  # assess.glmnet(fit$fit.preval, #newx = X[,2:293], 
  #               newy = y, family = stats::Gamma(link = "log"))
  # save the minimum lambda
  best_lambda <- fit$lambda.min
  # save the lambda for the most regularized model w/ an MSE that is still 1SE w/in the best lambda model
  lambda_1SE <- fit$lambda.1se
  # save the lambda for the most regularized model w/ an MSE that is still .5SE w/in the best lambda model
  lambda_halfSE <- best_lambda + ((lambda_1SE - best_lambda)/2)
 
  print(fit)     
  
  plot(fit)

```


Then, we predict (on the training set) using both of these models (best lambda and 1se lambda) 
```{r}
  ## predict on the test data
  # lasso model predictions with the optimal lambda
  optimal_pred <- predict(fit_glm_bestLambda, newx=X[,2:ncol(X)], type = "response")
  optimal_pred_1se <-  predict(fit_glm_se, newx=X[,2:ncol(X)], type = "response")
  optimal_pred_halfse <- predict(fit_glm_halfse, newx = X[,2:ncol(X)], type = "response")
  
    null_fit <- glm(
      formula = y ~ 1, #data = modDat_1_s, 
      family = binomial
      )
  null_pred <- predict(null_fit, newdata = as.data.frame(X), type = "response"
                       )

  # save data
  fullModOut <- list(
    "modelObject" = fit,
    "nullModelObject" = null_fit,
    "modelPredictions" = data.frame(#ecoRegion_holdout = rep(test_eco,length(y)),
      obs=y,
                    pred_opt=optimal_pred, 
                    pred_opt_se = optimal_pred_1se,
                    pred_opt_halfse = optimal_pred_halfse,
                    pred_null=null_pred#,
                    #pred_nopenalty=nopen_pred
                    ))

ggplot() + 
  geom_point(aes(X[,2], fullModOut$modelPredictions$obs), col = "black", alpha = .1) + 
  geom_point(aes(X[,2], fullModOut$modelPredictions$pred_opt), col = "red", alpha = .1) + ## predictions w/ the CV model
  geom_point(aes(X[,2], fullModOut$modelPredictions$pred_opt_halfse), col = "orange", alpha = .1) + ## predictions w/ the CV model (.5se lambda)
  geom_point(aes(X[,2], fullModOut$modelPredictions$pred_opt_se), col = "green", alpha = .1) + ## predictions w/ the CV model (1se lambda)
  geom_point(aes(X[,2], fullModOut$modelPredictions$pred_null), col = "blue", alpha = .1) + 
  labs(title = "A rough comparison of observed and model-predicted values", 
       subtitle = "black = observed values \n red = predictions from 'best lambda' model \n orange = predictions for '1/2se' lambda model \n green = predictions from '1se' lambda model \n blue = predictions from null model") +
  xlab(colnames(X)[2])
  #ylim(c(0,200))
 
```

The internal cross-validation process to fit the global LASSO model identified an optimal lambda value (regularization parameter) of `r{print(best_lambda)}`. The lambda value such that the cross validation error is within 1 standard error of the minimum ("1se lambda") was `r{print(fit$lambda.1se)}`` . The following coefficients were kept in each model: 
```{r coefficients in the global model}
# the coefficient matrix from the 'best model' -- find and print those coefficients that aren't 0 in a table
coef_glm_bestLambda <- coef(fit_glm_bestLambda) %>% 
  data.frame() 
coef_glm_bestLambda$coefficientName <- rownames(coef_glm_bestLambda)
names(coef_glm_bestLambda)[1] <- "coefficientValue_bestLambda"
# coefficient matrix from the '1se' model 
coef_glm_1se <- coef(fit_glm_se) %>% 
  data.frame() 
coef_glm_1se$coefficientName <- rownames(coef_glm_1se)
names(coef_glm_1se)[1] <- "coefficientValue_1seLambda"
# coefficient matrix from the 'half se' model 
coef_glm_halfse <- coef(fit_glm_halfse) %>% 
  data.frame() 
coef_glm_halfse$coefficientName <- rownames(coef_glm_halfse)
names(coef_glm_halfse)[1] <- "coefficientValue_halfseLambda"
# add together
coefs <- full_join(coef_glm_bestLambda, coef_glm_halfse) %>% 
  full_join(coef_glm_1se) %>% 
  dplyr::select(coefficientName, coefficientValue_bestLambda,
                coefficientValue_halfseLambda, coefficientValue_1seLambda)

globModTerms <- coefs[!is.na(coefs$coefficientValue_bestLambda), "coefficientName"]

## also, get the number of unique variables in each model 
var_prop_pred <- paste0(response, "_pred")
response_vars <- c(response, var_prop_pred)
# for best lambda model
prednames_fig <- paste(str_split(globModTerms, ":", simplify = TRUE)) 
prednames_fig <- str_replace(prednames_fig, "I\\(", "")
prednames_fig <- str_replace(prednames_fig, "\\^2\\)", "")
prednames_fig <- unique(prednames_fig[prednames_fig>0])
prednames_fig <- prednames_fig
prednames_fig_num <- length(prednames_fig)
# for 1SE lambda model
globModTerms_1se <- coefs[!is.na(coefs$coefficientValue_1seLambda), "coefficientName"]
if (length(globModTerms_1se) == 1) {
prednames_fig_1se <- paste(str_split(globModTerms_1se, ":", simplify = TRUE)) 
prednames_fig_1se <- str_replace(prednames_fig_1se, "I\\(", "")
prednames_fig_1se <- str_replace(prednames_fig_1se, "\\^2\\)", "")
prednames_fig_1se <- unique(prednames_fig_1se[prednames_fig_1se>0])
prednames_fig_1se_num <- c(0)
} else {
prednames_fig_1se <- paste(str_split(globModTerms_1se, ":", simplify = TRUE)) 
prednames_fig_1se <- str_replace(prednames_fig_1se, "I\\(", "")
prednames_fig_1se <- str_replace(prednames_fig_1se, "\\^2\\)", "")
prednames_fig_1se <- unique(prednames_fig_1se[prednames_fig_1se>0])
prednames_fig_1se_num <- length(prednames_fig_1se)
}
# for 1/2SE lambda model
globModTerms_halfse <- coefs[!is.na(coefs$coefficientValue_halfseLambda), "coefficientName"]
if (length(globModTerms_halfse) == 1) {
prednames_fig_halfse <- paste(str_split(globModTerms_halfse, ":", simplify = TRUE)) 
prednames_fig_halfse <- str_replace(prednames_fig_halfse, "I\\(", "")
prednames_fig_halfse <- str_replace(prednames_fig_halfse, "\\^2\\)", "")
prednames_fig_halfse <- unique(prednames_fig_halfse[prednames_fig_halfse>0])
prednames_fig_halfse_num <- c(0)
} else {
prednames_fig_halfse <- paste(str_split(globModTerms_halfse, ":", simplify = TRUE)) 
prednames_fig_halfse <- str_replace(prednames_fig_halfse, "I\\(", "")
prednames_fig_halfse <- str_replace(prednames_fig_halfse, "\\^2\\)", "")
prednames_fig_halfse <- unique(prednames_fig_halfse[prednames_fig_halfse>0])
prednames_fig_halfse_num <- length(prednames_fig_halfse)
}
# make a table
kable(coefs, col.names = c("Coefficient Name", "Value from best lambda model", 
                           "Value from 1/2 se lambda", "Value from 1se lambda model")
      ) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed")) 

# calculate RMSE of all models 
RMSE_best <- yardstick::rmse(fullModOut$modelPredictions[,c("obs", "pred_opt")], truth = "obs", estimate = "pred_opt")$.estimate
RMSE_halfse <- yardstick::rmse(fullModOut$modelPredictions[,c("obs", "pred_opt_halfse")], truth = "obs", estimate = "pred_opt_halfse")$.estimate
RMSE_1se <- yardstick::rmse(fullModOut$modelPredictions[,c("obs", "pred_opt_se")], truth = "obs", estimate = "pred_opt_se")$.estimate
# calculate bias of all models
bias_best <-  mean((fullModOut$modelPredictions$obs) - fullModOut$modelPredictions$pred_opt)
bias_halfse <-  mean((fullModOut$modelPredictions$obs) - fullModOut$modelPredictions$pred_opt_halfse)
bias_1se <- mean((fullModOut$modelPredictions$obs) - fullModOut$modelPredictions$pred_opt_se)

uniqueCoeffs <- data.frame("Best lambda model" = c(signif(RMSE_best,3), as.character(signif(bias_best, 3)),
  as.integer(length(globModTerms)-1), as.integer(prednames_fig_num), 
                                                   as.integer(sum(prednames_fig %in% c(prednames_clim))),
                                                   as.integer(sum(prednames_fig %in% c(prednames_weath))),
                                                   as.integer(sum(prednames_fig %in% c(prednames_soils)))
                                                   ), 
                           "1/2 se lambda model" = c(signif(RMSE_halfse,3), as.character(signif(bias_halfse, 3)),
                             length(globModTerms_halfse)-1, prednames_fig_halfse_num,
                                                   sum(prednames_fig_halfse %in% c(prednames_clim)),
                                                   sum(prednames_fig_halfse %in% c(prednames_weath)),
                                                   sum(prednames_fig_halfse %in% c(prednames_soils))), 
                           "1se lambda model" = c(signif(RMSE_1se,3), as.character(signif(bias_1se, 3)),
                             length(globModTerms_1se)-1, prednames_fig_1se_num,
                                                   sum(prednames_fig_1se %in% c(prednames_clim)),
                                                   sum(prednames_fig_1se %in% c(prednames_weath)),
                                                   sum(prednames_fig_1se %in% c(prednames_soils))))
row.names(uniqueCoeffs) <- c("RMSE", "bias: mean(obs-pred.)", "Total number of coefficients", "Number of unique coefficients",
                             "Number of unique climate coefficients", 
                             "Number of unique weather coefficients",  
                             "Number of unique soils coefficients"
                             )

kable(uniqueCoeffs, 
      col.names = c("Best lambda model", "1/2 se lambda model", "1se lambda model"), row.names = TRUE) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed")) 
```

# Visualizations of Model Predictions and Residuals  -- using best lambda model
## observed vs. predicted values

As the alternative to the best lambda model, use the model (1se or 1/2se of best Lambda) that has the fewest number of unique predictors 
```{r}
if (whichSecondBestMod == "auto") {
  # name of model w/ fewest # of predictors (but more than 0)
uniqueCoeff_min <- min(as.numeric(uniqueCoeffs[4,2:3])[which(as.numeric(uniqueCoeffs[4,2:3]) > 0)])
alternativeModel <- names(uniqueCoeffs[4,2:3])[which(uniqueCoeffs[4,2:3] == uniqueCoeff_min)]

if (is.finite(uniqueCoeff_min)) {
  if (length(alternativeModel) == 1) {
  if (alternativeModel == "X1.2.se.lambda.model") {
  mod_secondBest <- fit_glm_halfse
  name_secondBestMod <- "1/2 SE Model"
  prednames_secondBestMod <- prednames_fig_halfse
} else if (alternativeModel == "X1se.lambda.model") {
  mod_secondBest <- fit_glm_se
  name_secondBestMod <- "1 SE Model"
  prednames_secondBestMod <- prednames_fig_1se
}
} else {
  # if both alternative models have the same number of unique coefficients, chose the model that has the fewest number of total predictors
  uniqueCoeff_min2 <- min(as.numeric(uniqueCoeffs[3,alternativeModel]))
alternativeModel2 <- names(uniqueCoeffs[3,alternativeModel])[which(uniqueCoeffs[3,alternativeModel] == uniqueCoeff_min2)]
if (alternativeModel2 == "X1.2.se.lambda.model") {
  mod_secondBest <- fit_glm_halfse
  name_secondBestMod <- "1/2 SE Model"
  prednames_secondBestMod <- prednames_fig_halfse
} else if (alternativeModel2 == "X1se.lambda.model") {
  mod_secondBest <- fit_glm_se
  name_secondBestMod <- "1 SE Model"
  prednames_secondBestMod <- prednames_fig_1se
}
}
  }else {
    mod_secondBest <- NULL
  name_secondBestMod <- "Intercept_Only"
  prednames_secondBestMod <- NULL
}
} else if (whichSecondBestMod == "1se") {
  mod_secondBest <- fit_glm_se
  name_secondBestMod <- "1 SE Model"
  prednames_secondBestMod <- prednames_fig_1se
} else if (whichSecondBestMod == "halfse") {
  mod_secondBest <- fit_glm_halfse
  name_secondBestMod <- "1/2 SE Model"
  prednames_secondBestMod <- prednames_fig_halfse
}
```

### Predicting on the data

```{r predict with best lambda on data used for fitting}
  # create prediction for each each model
# (i.e. for each fire proporation variable)
predict_by_response <- function(mod, df) {
  df_out <- df
  response_name <- paste0("TotalTreeCover_binom", "_pred")
  preds <- predict(mod, newx= df_out, #s="lambda.min", 
                                     type = "response")
  preds[preds<0] <- 0
  #preds[preds>100] <- 100
  df_out <- df_out %>% cbind(preds)
   colnames(df_out)[ncol(df_out)] <- response_name
  return(df_out)
}

pred_glm1 <- predict_by_response(fit_glm_bestLambda, X[,2:ncol(X)])

## back-transform the 
# add back in true y values
pred_glm1 <- pred_glm1 %>% 
  cbind(data.frame("TotalTreeCover_binom" = modDat_1_s$TotalTreeCover_binom))

# add back in lat/long data 
pred_glm1 <- pred_glm1 %>% 
  cbind(modDat_1_s[,c("x", "y", "Year")])

pred_glm1$resid <- pred_glm1[,"TotalTreeCover_binom"] - pred_glm1[,paste0("TotalTreeCover_binom", "_pred")]
pred_glm1$extremeResid <- NA
pred_glm1[pred_glm1$resid > .5 | pred_glm1$resid < -.5,"extremeResid"] <- 1

```

```{r predict with second best lambda model on data used for model fitting}
if ( name_secondBestMod == "Intercept_Only") {
  print("The next best lambda model only contains one predictor (an intercept)")
} else {

pred_glm1_1se <- predict_by_response(mod_secondBest, X[,2:ncol(X)])

# add back in true y values
pred_glm1_1se <- pred_glm1_1se %>% 
  cbind(data.frame("TotalTreeCover_binom" = modDat_1_s$TotalTreeCover_binom))

# add back in lat/x data 
pred_glm1_1se <- pred_glm1_1se %>% 
  cbind(modDat_1_s[,c("x", "y", "Year")])

pred_glm1_1se$resid <- pred_glm1_1se[,"TotalTreeCover_binom"] - pred_glm1_1se[,paste0("TotalTreeCover_binom", "_pred")]
pred_glm1_1se$extremeResid <- NA
pred_glm1_1se[pred_glm1_1se$resid > .5 | pred_glm1_1se$resid < -.5,"extremeResid"] <- 1
}
```

### Maps of Observations, Predictions, and Residuals=

Observations across the temporal range of the dataset
```{r Residual maps, fig.height = 7, fig.width = 9, }
# rasterize
# get reference raster
test_rast <-  rast("../../../Data_raw/dayMet/rawMonthlyData/orders/70e0da02b9d2d6e8faa8c97d211f3546/Daymet_Monthly_V4R1/data/daymet_v4_prcp_monttl_na_1980.tif") %>%
  terra::aggregate(fact = 3, fun = "mean") %>% 
  terra::project(crs("EPSG:4326"))
  # transform to match format of veg. data 
  
## add ecoregion boundaries (for our ecoregion level model)
regions <- sf::st_read(dsn = "../../../Data_raw/Level2Ecoregions/", layer = "NA_CEC_Eco_Level2") 
regions <- regions %>% 
  st_transform(crs = st_crs(test_rast)) %>% 
  st_make_valid() 

ecoregionLU <- data.frame("NA_L1NAME" = sort(unique(regions$NA_L1NAME)), 
                        "newRegion" = c(NA, "Forest", "dryShrubGrass", 
                                        "dryShrubGrass", "Forest", "dryShrubGrass",
                                       "dryShrubGrass", "Forest", "Forest", 
                                       "dryShrubGrass", "Forest", "Forest", 
                                       "Forest", "Forest", "dryShrubGrass", 
                                       NA
                                        ))
goodRegions <- regions %>% 
  left_join(ecoregionLU)
mapRegions <- goodRegions %>% 
  filter(!is.na(newRegion)) %>% 
  group_by(newRegion) %>% 
  summarise(geometry = sf::st_union(geometry)) %>% 
  ungroup() %>% 
  st_simplify(dTolerance = 1000) %>% 
  st_crop(ext(-130, -60, 20, 60))

# rasterize data
plotObs <- pred_glm1 %>% 
         drop_na("TotalTreeCover_binom") %>% 
  #slice_sample(n = 5e4) %>%
  terra::vect(geom = c("x", "y")) %>% 
  terra::set.crs(crs(test_rast)) %>% 
  terra::rasterize(y = test_rast, 
                   field = "TotalTreeCover_binom", 
                   fun = function(x) {
                     round(mean(x))
                   }) %>% 
  terra::crop(ext(-130, -60, 20, 60))

# make shapefile of cropped state boundaries in appropriate crs
cropped_states_2 <- cropped_states %>% 
  st_transform(crs = "EPSG:4326") %>% 
  st_make_valid() %>% 
  st_crop(ext(-130, -60, 20, 60))

# make figures
map_obs <- ggplot() +
geom_spatraster(data = plotObs) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
  geom_sf(data = mapRegions, fill = NA, col = "orchid", lwd = .5) +
labs(title = paste0("Observations of Binomial Total Tree Cover")) +
  scale_fill_gradient2(low = "brown",
                       mid = "wheat" ,
                       high = "darkgreen" , 
                       midpoint = 0,   na.value = "darkgrey") + 
  xlim(-125, -65) + 
  ylim(25, 50)

hist_obs <- ggplot(pred_glm1) + 
  geom_histogram(aes(TotalTreeCover_binom), fill = "lightgrey", col = "darkgrey")

library(ggpubr)
ggarrange(map_obs, hist_obs, heights = c(3,1), ncol = 1)
```

Predictions across the temporal range of the dataset
```{r predicted values map best lambda model, fig.height = 7, fig.width = 9, }

# rasterize data
plotPred <- pred_glm1 %>% 
         drop_na(paste0("TotalTreeCover_binom","_pred")) %>% 
  #slice_sample(n = 5e4) %>%
  terra::vect(geom = c("x", "y")) %>% 
  terra::set.crs(crs(test_rast)) %>% 
  terra::rasterize(y = test_rast, 
                   field = paste0("TotalTreeCover_binom","_pred"), 
                   fun = mean) %>% 
  terra::crop(ext(-130, -60, 20, 60))

# make figures
map_preds1 <- ggplot() +
geom_spatraster(data = plotPred) + 
  geom_sf(data=cropped_states_2,fill=NA )  + 
  geom_sf(data = mapRegions, fill = NA, col = "orchid", lwd = .5) +
labs(title = paste0("Predictions from the 'best lambda' fitted model of Yes/No Trees"),
     subtitle =  "bestLambda model")  +
  scale_fill_gradient2(low = "wheat",
                       mid = "darkgreen",
                       high = "red" , 
                       midpoint = 1,   na.value = "darkgrey",
                       limits = c(0,1)) + 
  xlim(-125, -65) + 
  ylim(25, 50)

hist_preds1 <- ggplot(pred_glm1) + 
  geom_histogram(aes(.data[[paste0("TotalTreeCover_binom", "_pred")]]), fill = "lightgrey", col = "darkgrey")+ 
  xlim(c(0,1))

ggarrange(map_preds1, hist_preds1, heights = c(3,1), ncol = 1)
```

```{r predicted values map 1se lambda model, fig.height = 7, fig.width = 9, }
if ( name_secondBestMod == "Intercept_Only") {
  print("The next best lambda model only contains one predictor (an intercept)")
} else {

# rasterize data
plotPred <- pred_glm1_1se %>% 
         drop_na(paste0("TotalTreeCover_binom","_pred")) %>% 
  #slice_sample(n = 5e4) %>%
  terra::vect(geom = c("x", "y")) %>% 
  terra::set.crs(crs(test_rast)) %>% 
  terra::rasterize(y = test_rast, 
                   field = paste0("TotalTreeCover_binom","_pred"), 
                   fun = mean) %>% 
     terra::crop(ext(-130, -60, 20, 60))

# make figures
map_preds2 <- ggplot() +
geom_spatraster(data = plotPred) + 
  geom_sf(data=cropped_states_2,fill=NA )  + 
  geom_sf(data = mapRegions, fill = NA, col = "orchid", lwd = .5) +
labs(title = paste0("Predictions from the ", name_secondBestMod, " of Yes/No Trees"),
     subtitle =  name_secondBestMod)  +
  scale_fill_gradient2(low = "wheat",
                       mid = "darkgreen",
                       high = "red" , 
                       midpoint = 1,   na.value = "darkgrey",
                       limits = c(0, 1))  + 
  xlim(-125, -65) + 
  ylim(25, 50)

hist_preds2 <- ggplot(pred_glm1_1se) + 
  geom_histogram(aes(.data[[paste0("TotalTreeCover_binom", "_pred")]]), fill = "lightgrey", col = "darkgrey")+ 
  xlim(c(0,1))

ggarrange(map_preds2, hist_preds2, heights = c(3,1), ncol = 1)
}
```

```{r Residual map from best lambda model, fig.height = 12, fig.width = 9, warning = FALSE }

# rasterize data
plotResid_rast <- pred_glm1 %>% 
         drop_na(resid) %>% 
  #slice_sample(n = 5e4) %>%
  terra::vect(geom = c("x", "y")) %>% 
  terra::set.crs(crs(test_rast)) %>% 
  terra::rasterize(y = test_rast, 
                   field = "resid", 
                   fun = mean) %>% 
     terra::crop(ext(-130, -60, 20, 60))

# identify locations where residuals are >.5 or < -.5
badResids_high <- pred_glm1 %>% 
  filter(resid > .5)  %>% 
  terra::vect(geom = c("x", "y")) %>% 
  terra::set.crs(crs(test_rast)) 
badResids_low <- pred_glm1 %>% 
  filter(resid < -.5)  %>% 
  terra::vect(geom = c("x", "y")) %>% 
  terra::set.crs(crs(test_rast)) 
# make figures
map <- ggplot() +
geom_spatraster(data =plotResid_rast) + 
  geom_sf(data=cropped_states_2,fill=NA )  + 
  geom_sf(data = mapRegions, fill = NA, col = "orchid", lwd = .5) +
  #geom_sf(data = badResids_high, col = "blue") +
  #geom_sf(data = badResids_low, col = "red") +
labs(title = paste0("Resids. (obs. - pred.) from the model of Yes/No Trees"),
     subtitle = "bestLambda model \n red points indicate locations that have residuals below -.5 \n blue points indicate locations that have residuals above .5") +
  scale_fill_gradient2(low = "red",
                       mid = "white" ,
                       high = "blue" , 
                       midpoint = 0,   na.value = "darkgrey",
                       limits = c(-1,1)
                       ) + 
  xlim(-125, -65) + 
  ylim(25, 50)

hist <- ggplot(pred_glm1) + 
  geom_histogram(aes(resid), fill = "lightgrey", col = "darkgrey") + 
  # geom_text(aes(x = min(resid)*.9, y = 1500, label = paste0("min = ", round(min(resid),2)))) +
  # geom_text(aes(x = max(resid)*.9, y = 1500, label = paste0("max = ", round(max(resid),2)))) + 
  geom_vline(aes(xintercept = mean(resid)))

ggarrange(map, hist, heights = c(3,1), ncol = 1)

```

```{r Residual map from 1 SE lambda model, fig.height = 12, fig.width = 9, warning = FALSE }
if ( name_secondBestMod == "Intercept_Only") {
  print("The next best lambda model only contains one predictor (an intercept)")
} else {

# rasterize data
plotResid_rast <- pred_glm1_1se %>% 
         drop_na(resid) %>% 
  #slice_sample(n = 5e4) %>%
  terra::vect(geom = c("x", "y")) %>% 
  terra::set.crs(crs(test_rast)) %>% 
  terra::rasterize(y = test_rast, 
                   field = "resid", 
                   fun = mean) %>% 
     terra::crop(ext(-130, -60, 20, 60))

# identify locations where residuals are >10 or < -10
badResids_high <- pred_glm1_1se %>% 
  filter(resid > .5)  %>% 
  terra::vect(geom = c("x", "y")) %>% 
  terra::set.crs(crs(test_rast)) 
badResids_low <- pred_glm1_1se %>% 
  filter(resid < -.5)  %>% 
  terra::vect(geom = c("x", "y")) %>% 
  terra::set.crs(crs(test_rast)) 
# make figures
map <- ggplot() +
geom_spatraster(data =plotResid_rast) + 
  geom_sf(data=cropped_states_2,fill=NA )  + 
  geom_sf(data = mapRegions, fill = NA, col = "orchid", lwd = .5) +
  # geom_sf(data = badResids_high, col = "blue") +
  # geom_sf(data = badResids_low, col = "red") +
labs(title = paste0("Resids. (obs. - pred.) from the model of Yes/No Trees"),
     subtitle = paste0(name_secondBestMod,"\n red points indicate locations that have residuals below -.5 \n blue points indicate locations that have residuals above .5")) +
  scale_fill_gradient2(low = "red",
                       mid = "white" ,
                       high = "blue" , 
                       midpoint = 0,   na.value = "darkgrey",
                       limits = c(-1,1)
                       )  + 
  xlim(-125, -65) + 
  ylim(25, 50)
hist <- ggplot(pred_glm1_1se) + 
  geom_histogram(aes(resid), fill = "lightgrey", col = "darkgrey") + 
 # geom_text(aes(x = min(resid)*.9, y = 1500, label = paste0("min = ", round(min(resid),2)))) +
  #geom_text(aes(x = max(resid)*.9, y = 1500, label = paste0("max = ", round(max(resid),2))))+ 
  geom_vline(aes(xintercept = mean(resid)))

ggarrange(map, hist, heights = c(3,1), ncol = 1)
}
```

### Are there biases of the model predictions across year/lat/long? 
```{r bias figures best lambda model, fig.align=8, fig.height=10}
if ( name_secondBestMod == "Intercept_Only") {
  print("The next best lambda model only contains one predictor (an intercept)")
  # plot residuals against Year
yearResidMod_bestLambda <- ggplot(pred_glm1) + 
  geom_point(aes(x = jitter(Year), y = resid), alpha = .1) + 
  geom_smooth(aes(x = Year, y = resid)) + 
  xlab("Year") + 
  ylab("Residuals") +
  ggtitle("from best lamba model")

# plot residuals against Lat
latResidMod_bestLambda <- ggplot(pred_glm1) + 
  geom_point(aes(x = y, y = resid), alpha = .1) + 
  geom_smooth(aes(x = y, y = resid)) + 
  xlab("Latitude") + 
  ylab("Residuals") +
  ggtitle("from best lamba model")

# plot residuals against Long
longResidMod_bestLambda <- ggplot(pred_glm1) + 
  geom_point(aes(x = x, y = resid), alpha = .1) + 
  geom_smooth(aes(x = x, y = resid)) + 
  xlab("Longitude") + 
  ylab("Residuals") +
  ggtitle("from best lamba model")
library(patchwork)
(yearResidMod_bestLambda ) / 
(  latResidMod_bestLambda ) /
(  longResidMod_bestLambda )
} else {

# plot residuals against Year
yearResidMod_bestLambda <- ggplot(pred_glm1) + 
  geom_point(aes(x = jitter(Year), y = resid), alpha = .1) + 
  geom_smooth(aes(x = Year, y = resid)) + 
  xlab("Year") + 
  ylab("Residuals") +
  ggtitle("from best lamba model")
yearResidMod_1seLambda <- ggplot(pred_glm1_1se) + 
  geom_point(aes(x = jitter(Year), y = resid), alpha = .1) + 
  geom_smooth(aes(x = Year, y = resid)) + 
  xlab("Year") + 
  ylab("Residuals") +
  ggtitle(paste0("from ", name_secondBestMod))

# plot residuals against Lat
latResidMod_bestLambda <- ggplot(pred_glm1) + 
  geom_point(aes(x = y, y = resid), alpha = .1) + 
  geom_smooth(aes(x = y, y = resid)) + 
  xlab("Latitude") + 
  ylab("Residuals") +
  ggtitle("from best lamba model")
latResidMod_1seLambda <- ggplot(pred_glm1_1se) + 
  geom_point(aes(x = y, y = resid), alpha = .1) + 
  geom_smooth(aes(x = y, y = resid)) + 
  xlab("Latitude") + 
  ylab("Residuals") +
  ggtitle(paste0("from ", name_secondBestMod))

# plot residuals against Long
longResidMod_bestLambda <- ggplot(pred_glm1) + 
  geom_point(aes(x = x, y = resid), alpha = .1) + 
  geom_smooth(aes(x = x, y = resid)) + 
  xlab("Longitude") + 
  ylab("Residuals") +
  ggtitle("from best lamba model")
longResidMod_1seLambda <- ggplot(pred_glm1_1se) + 
  geom_point(aes(x = x, y = resid), alpha = .1) + 
  geom_smooth(aes(x = x, y = resid)) + 
  xlab("Longitude") + 
  ylab("Residuals") +
  ggtitle(paste0("from ", name_secondBestMod))

library(patchwork)
(yearResidMod_bestLambda + yearResidMod_1seLambda) / 
(  latResidMod_bestLambda + latResidMod_1seLambda) /
(  longResidMod_bestLambda + longResidMod_1seLambda)
}
```

### Quantile plots

Binning predictor variables into "Quantiles"  and looking at the mean
predicted probability for each percentile.
```{r}
response_vars <- c("TotalTreeCover_binom", "TotalTreeCover_binom_pred")
# get deciles for best lambda model 
if (length(prednames_fig) == 0) {
  print("The best lambda model only contains one predictor (an intercept), so decile plots aren't possible to generate")
} else {
  pred_glm1_deciles <- predvars2deciles(pred_glm1,
                                      response_vars = response_vars,
                                        pred_vars = prednames_fig, 
                                       cut_points = seq(0, 1, 0.01))
}
# get deciles for 1 SE lambda model 
if ( name_secondBestMod == "Intercept_Only") {
  print("The next best lambda model only contains one predictor (an intercept)")} else {
  pred_glm1_deciles_1se <- predvars2deciles(pred_glm1_1se,
                                      response_vars = response_vars,
                                        pred_vars = prednames_secondBestMod, 
                                       cut_points = seq(0, 1, 0.01))
  }
```

Below are quantile plots for the *best lambda model* 
(note that the predictor variables are scaled) 
```{r Percentile_plot best lambda, fig.height= 16, fig.width = 12, warning=FALSE, message=FALSE}
if (length(prednames_fig) == 0) {
  print("The model only contains one predictor (an intercept), so decile plots aren't possible to generate")
} else {

# publication quality version
g3 <- decile_dotplot_pq(df = pred_glm1_deciles, response = c("TotalTreeCover_binom", paste0("TotalTreeCover_binom", "_pred")), IQR = FALSE,
                        CI = FALSE
                        ) + ggtitle("Decile Plot")

g4 <- add_dotplot_inset(g3, df = pred_glm1_deciles, response = c("TotalTreeCover_binom", paste0("TotalTreeCover_binom", "_pred")), dfRaw = pred_glm1, add_smooth = TRUE, deciles = FALSE)

  
if(save_figs) {
  # png(paste0("../../../Figures/CoverDatFigures/ figures/quantile_plots/quantile_plot_", response,  "_",ecoregion,".png"), 
  #    units = "in", res = 600, width = 5.5, height = 3.5 )
  #   print(g4)
  # dev.off()
}

g4
}

```

Below are percentile plots from the second best lambda model (`r print(name_secondBestMod)`)
```{r Percentile_plot 1 SE lambda, fig.height= 16, fig.width = 12, warning=FALSE, message=FALSE}

if ( name_secondBestMod == "Intercept_Only") {
  print("The next best lambda model only contains one predictor (an intercept)")
  } else {

# publication quality version
g3 <- decile_dotplot_pq(pred_glm1_deciles_1se, response = c("TotalTreeCover_binom", paste0("TotalTreeCover_binom", "_pred")), IQR = FALSE) + ggtitle("Decile Plot")

g4 <- add_dotplot_inset(g3, df = pred_glm1_deciles_1se, response = c("TotalTreeCover_binom", paste0("TotalTreeCover_binom", "_pred")), dfRaw = pred_glm1_1se, add_smooth = TRUE, deciles = FALSE)

  
if(save_figs) {
  # png(paste0("figures/quantile_plots/quantile_plot_", response,  "_",ecoregion,".png"), 
  #    units = "in", res = 600, width = 5.5, height = 3.5 )
  #   print(g4)
  # dev.off()
}

g4
}

```

### Filtered Quantiles
#### For the best lambda model

Filtered 'Quantile' plots of data. These plots show each vegetation variable,
but only based on data that falls into the upper and lower 20th percentiles of
each predictor variable. 

```{r glm_deciles_filtered best lambda model, fig.height = 24, fig.width = 24, message = FALSE}
if (length(prednames_fig) == 0) {
  print("The model only contains one predictor (an intercept), so decile plots aren't possible to generate")
} else {
pred_glm1_deciles_filt <- predvars2deciles( pred_glm1, 
                         response_vars = c("TotalTreeCover_binom", paste0("TotalTreeCover_binom", "_pred")),
                         pred_vars = prednames_fig,
                         filter_var = TRUE,
                         filter_vars = prednames_fig,
                         cut_points = seq(0, 1, 0.01)) 
g <- decile_dotplot_filtered_pq(df = pred_glm1_deciles_filt, response = "TotalTreeCover_binom", 
                                xvars = prednames_fig)

if(save_figs) {
# jpeg(paste0("figures/quantile_plots/quantile_plot_filtered_mid_v1", , ".jpeg"),
#      units = "in", res = 600, width = 5.5, height = 6 )
#   g 
# dev.off()
}
g
}
```

Filtered quantile figure with middle 2 deciles also shown

```{r filtered quantile plot for best lambda modle with middle 2 deciles, fig.height = 24, fig.width = 24, eval = TRUE, message = FALSE}
if ( name_secondBestMod == "Intercept_Only") {
  print("The next best lambda model only contains one predictor (an intercept)")
} else {
pred_glm1_deciles_filt_mid <- predvars2deciles(pred_glm1, 
                         response_vars = c("TotalTreeCover_binom", paste0("TotalTreeCover_binom", "_pred")),
                         pred_vars = prednames_fig,
                         filter_vars = prednames_fig,
                         filter_var = TRUE,
                         add_mid = TRUE,
                         cut_points = seq(0, 1, 0.01))

g <- decile_dotplot_filtered_pq(df = pred_glm1_deciles_filt_mid, response = "TotalTreeCover_binom", 
                                xvars = prednames_fig)

if(save_figs) {
# jpeg(paste0("figures/quantile_plots/quantile_plot_filtered_mid_v1", , ".jpeg"),
#      units = "in", res = 600, width = 5.5, height = 6)
#   g 
# dev.off()
}
g
}
```

#### For the second best lambda model (`r print(name_secondBestMod)`)

Filtered 'Quantile' plots of data. These plots show each vegetation variable,
but only based on data that falls into the upper and lower 20th percentiles of
each predictor variable. 

```{r glm_deciles_filtered second best lambda model, fig.height = 24, fig.width = 24, message = FALSE}
if (length(prednames_secondBestMod) == 0) {
  print("The model only contains one predictor (an intercept), so decile plots aren't possible to generate")
} else {
  
pred_glm1_deciles_filt <- predvars2deciles( pred_glm1_1se, 
                         response_vars = c("TotalTreeCover_binom", paste0("TotalTreeCover_binom", "_pred")),
                         pred_vars = prednames_secondBestMod,
                         filter_var = TRUE,
                         filter_vars = prednames_secondBestMod,
                         cut_points = seq(0, 1, 0.01)) 
g <- decile_dotplot_filtered_pq(df = pred_glm1_deciles_filt, response = "TotalTreeCover_binom", 
                                xvars = prednames_fig)

if(save_figs) {
# jpeg(paste0("figures/quantile_plots/quantile_plot_filtered_mid_v1", , ".jpeg"),
#      units = "in", res = 600, width = 5.5, height = 6 )
#   g 
# dev.off()
}
g
}
```

Filtered quantile figure with middle 2 deciles also shown

```{r filtered quantile plot for second best lambda modle with middle 2 deciles, fig.height = 24, fig.width = 24, eval = TRUE, message = FALSE}
if (length(prednames_secondBestMod) == 0) {
  print("The model only contains one predictor (an intercept), so decile plots aren't possible to generate")
} else {
pred_glm1_deciles_filt_mid <- predvars2deciles(pred_glm1, 
                         response_vars = c("TotalTreeCover_binom", paste0("TotalTreeCover_binom", "_pred")),
                         pred_vars = prednames_secondBestMod,
                         filter_vars = prednames_secondBestMod,
                         filter_var = TRUE,
                         add_mid = TRUE,
                         cut_points = seq(0, 1, 0.01))

g <- decile_dotplot_filtered_pq(df = pred_glm1_deciles_filt_mid, response = "TotalTreeCover_binom", 
                                xvars = prednames_fig)


if(save_figs) {
# jpeg(paste0("figures/quantile_plots/quantile_plot_filtered_mid_v1", , ".jpeg"),
#      units = "in", res = 600, width = 5.5, height = 6 )
#   g 
# dev.off()
}
g
}
```

### Show model RMSE w/in each quantile
```{r}
# get deciles for best lambda model 
if (length(prednames_fig) == 0) {
  print("The best lambda model only contains one predictor (an intercept), so decile plots aren't possible to generate")
} else {
  pred_glm1_deciles %>% 
    ggplot(aes(x = mean_value, y = RMSE)) +
    facet_wrap(~name, scales = "free_x") +
    geom_point(alpha = .2, size = .5) + 
    geom_smooth(lwd = .5) + 
    xlab("Scaled predictor value") + 
    ggtitle("RMSE by decile for bestLambda model")
}
# get deciles for 1 SE lambda model 
if (length(prednames_secondBestMod) == 0) {
  print("The 1SE (or 1/2 SE) lambda model only contains one predictor (an intercept), so decile plots aren't possible to generate")
} else {
  pred_glm1_deciles_1se %>% 
    ggplot(aes(x = mean_value, y = RMSE)) +
    facet_wrap(~name, scales = "free_x") +
    geom_point(alpha = .2, size = .5) + 
    geom_smooth(lwd = .5) + 
    xlab("Scaled predictor value") + 
    ggtitle(paste0("RMSE by decile for ", name_secondBestMod, "model"))
}
```

# Cross-validation
## Using best lambda model
### Use terms from global model to re-fit and predict on different held out regions
### Figures show residuals for each of the models fit to held-out ecoregions
These models were fit to six ecoregions, and then predict on the indicated heldout ecoregion

```{r cross-validation for best lambda model , fig.height = 7, fig.width = 9, warning=FALSE, message=FALSE }
if (length(prednames_fig) == 0) {
  print("The model only contains one predictor (an intercept), so cross validation isn't practical")
} else {

## code from Tredennick et al. 2020
# try each separate level II ecoregion as a test set
# make a list to hold output data
outList <- vector(mode = "list", length = length(sort(unique(modDat_1_s$NA_L2NAME))))
# obs_pred <- data.frame(ecoregion = character(),obs = numeric(),
#                        pred_opt = numeric(), pred_null = numeric()#,
#                        #pred_nopenalty = numeric()
#                        )

## get the model specification from the global model
mat <- as.matrix(coef(fit_glm_bestLambda, s = "lambda.min"))
mat2 <- mat[mat[,1] != 0,]

f_cv <- as.formula(paste0("TotalTreeCover_binom", " ~ ", paste0(names(mat2)[2:length(names(mat2))], collapse = " + ")))

X_cv <- model.matrix(object = f_cv, data = modDat_1_s)
# get response variable
y_cv <- as.matrix(modDat_1_s[,"TotalTreeCover_binom"])


# now, loop through so with each iteration, a different ecoregion is held out
 for(i_eco in sort(unique(modDat_1_s$NA_L2NAME))){

  # split into training and test sets
  test_eco <- i_eco
  print(test_eco)
  # identify the rowID of observations to be in the training and test datasets
  train <- which(modDat_1_s$NA_L2NAME!=test_eco) # data for all ecoregions that aren't 'i_eco'
  test <- which(modDat_1_s$NA_L2NAME==test_eco) # data for the ecoregion that is 'i_eco'

  trainDat_all <- modDat_1_s %>%
    slice(train) %>%
    dplyr::select(-newRegion)
  testDat_all <- modDat_1_s %>%
    slice(test) %>%
    dplyr::select(-newRegion)

  # get the model matrices for input and response variables for cross validation model specification
  X_train <- as.matrix(X_cv[train,])
  X_test <- as.matrix(X_cv[test,])

  y_train <- modDat_1_s[train,"TotalTreeCover_binom"]
  y_test <- modDat_1_s[test,"TotalTreeCover_binom"]

  # get the model matrices for input and response variables for original model specification
  X_train_glob <- as.matrix(X[train,])
  X_test_glob <- as.matrix(X[test,])

  y_train_glob <- modDat_1_s[train,"TotalTreeCover_binom"]
  y_test_glob <- modDat_1_s[test,"TotalTreeCover_binom"]

  train_eco <- modDat_1_s$NA_L2NAME[train]

  ## just try a regular glm w/ the components from the global model
  fit_i <- glm(data = trainDat_all, formula = f_cv,
    ,
               family =  binomial
    )
    
  # lasso model predictions with the optimal lambda (back transformed)
  optimal_pred <- predict(fit_i, newdata = testDat_all, type = "response") 
  # null model and predictions
  # the "null" model in this case is the global model
  # predict on the test data for this iteration w/ the global model (back transformed)
  null_pred <- predict.glm(fit_glm_bestLambda, newdata = testDat_all, type = "response") 

  # save data
  tmp <- data.frame(ecoRegion_holdout = rep(test_eco,length(y_test)),
                    obs=y_test,
                    pred_opt=optimal_pred,
                    pred_null=null_pred#,
                    #pred_nopenalty=nopen_pred
                    ) %>%
    cbind(testDat_all)

  # calculate RMSE, bias, etc. of
  # RMSE of CV model
  RMSE_optimal <- yardstick::rmse(data = data.frame(optimal_pred,"y_test" = (y_test)), truth = "y_test", estimate = "optimal_pred")[1,]$.estimate
  # RMSE of global model
  RMSE_null <- yardstick::rmse(data = data.frame(null_pred,"y_test" = (y_test)), truth = "y_test", estimate = "null_pred")[1,]$.estimate
  # bias of CV model
  bias_optimal <- mean((y_test) - optimal_pred)
  # bias of global model
  bias_null <-  mean((y_test) - null_pred )

  # put output into a list
  tmpList <- list("testRegion" = i_eco,
    "modelObject" = fit_i,
       "modelPredictions" = tmp,
    "performanceMetrics" = data.frame("RMSE_cvModel" = RMSE_optimal,
                                      "RMSE_globalModel" = RMSE_null,
                                      "bias_cvModel" = bias_optimal,
                                      "bias_globalModel" = bias_null))

  # save model outputs
  outList[[which(sort(unique(modDat_1_s$NA_L2NAME)) == i_eco)]] <- tmpList
 }
}
```
### Below are the RMSE and bias values for predictions made for each holdout level II ecoregion, compared to predictions from the global model for that same ecoregion
```{r metrics output table for best lambda CV}
# table of model performance
purrr::map(outList, .f = function(x) {
  cbind(data.frame("holdout region" = x$testRegion),  x$performanceMetrics)
}
) %>%
  purrr::list_rbind() %>%
  kable(col.names = c("Held-out ecoregion", "RMSE of CV model", "RMSE of global model",
                      "bias of CV model - mean(obs-pred.)", "bias of global model- mean(obs-pred.)"),
        caption = "Performance of Cross Validation using 'best lambda' model specification") %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

```{r figures for best lambda CV}
# visualize model predictions
for (i in 1:length(unique(modDat_1_s$NA_L2NAME))) {
  holdoutRegion <- outList[[i]]$testRegion
  predictionData <- outList[[i]]$modelPredictions
  modTerms <- as.matrix(coef(outList[[i]]$modelObject)) %>%
    as.data.frame() %>%
    filter(V1!=0) %>%
    rownames()

  # calculate residuals
  predictionData <- predictionData %>%
  mutate(resid = .[["obs"]] - .[["pred_opt"]] ,
         resid_globMod = .[["obs"]]  - .[["pred_null"]])

# rasterize
# use 'test_rast' from earlier
  # rasterize data
plotObs <- predictionData %>%
         drop_na(paste("TotalTreeCover_binom")) %>%
  #slice_sample(n = 5e4) %>%
  terra::vect(geom = c("x", "y")) %>%
  terra::set.crs(crs(test_rast)) %>%
  terra::rasterize(y = test_rast,
                   field = "resid",
                   fun = mean)

tempExt <- crds(plotObs, na.rm = TRUE)

plotObs_2 <- plotObs %>%
  terra::crop(ext(min(tempExt[,1]), max(tempExt[,1]),
           min(tempExt[,2]), max(tempExt[,2]))
       )

# make figures
# make histogram
hist_i <- ggplot(predictionData) +
  geom_histogram(aes(resid_globMod), col = "darkgrey", fill = "lightgrey") +
  xlab(c("Residuals (obs. - pred.)"))
# make map
map_i <-  ggplot() +
geom_spatraster(data = plotObs_2) +
  geom_sf(data=cropped_states_2,fill=NA ) +
  geom_sf(data = mapRegions, fill = NA, col = "orchid", lwd = .5) +
labs(title = paste0("Residuals (obs. - pred.) for predictions of \n", holdoutRegion, " \n from a model fit to other ecoregions"),
     subtitle = paste0("TotalTreeCover_binom", " ~ ", paste0( modTerms, collapse = " + "))) +
  scale_fill_gradient2(low = "red",
                       mid = "white" ,
                       high = "blue" ,
                       midpoint = 0,   na.value = "darkgrey",
                       limits = c(-1, 1))   + 
  xlim(st_bbox(plotObs_2)[1],st_bbox(plotObs_2)[3]) + 
  ylim(st_bbox(plotObs_2)[2],st_bbox(plotObs_2)[4])

 assign(paste0("residPlot_",holdoutRegion),
   value = ggarrange(map_i, hist_i, heights = c(3,1), ncol = 1)
)

}

  lapply(unique(modDat_1_s$NA_L2NAME), FUN = function(x) {
    get(paste0("residPlot_", x))
  })

```

## Using second best lambda model (either 1se or 1/2se)
### Use terms from global model to re-fit and predict on different held out regions
### Figures show residuals for each of the models fit to held-out ecoregions
These models were fit to six ecoregions, and then predict on the indicated heldout ecoregion

```{r cross-validation for 1 SE lambda model , fig.height = 7, fig.width = 9, warning=FALSE, message=FALSE }
if (length(prednames_secondBestMod) == 0) {
  print("The model only contains one predictor (an intercept), so cross validation isn't practical")
} else {

## code from Tredennick et al. 2020
# try each separate level II ecoregion as a test set
# make a list to hold output data
outList <- vector(mode = "list", length = length(sort(unique(modDat_1_s$NA_L2NAME))))

## get the model specification from the global model
mat <- as.matrix(coef(mod_secondBest, s = "lambda.min"))
mat2 <- mat[mat[,1] != 0,]

f_cv <- as.formula(paste0("TotalTreeCover_binom", " ~ ", paste0(names(mat2)[2:length(names(mat2))], collapse = " + ")))

X_cv <- model.matrix(object = f_cv, data = modDat_1_s)
# get response variable
y_cv <- as.matrix(modDat_1_s[,"TotalTreeCover_binom"])


# now, loop through so with each iteration, a different ecoregion is held out
 for(i_eco in sort(unique(modDat_1_s$NA_L2NAME))){

  # split into training and test sets
  test_eco <- i_eco
  print(test_eco)
  # identify the rowID of observations to be in the training and test datasets
  train <- which(modDat_1_s$NA_L2NAME!=test_eco) # data for all ecoregions that aren't 'i_eco'
  test <- which(modDat_1_s$NA_L2NAME==test_eco) # data for the ecoregion that is 'i_eco'

  trainDat_all <- modDat_1_s %>%
    slice(train) %>%
    dplyr::select(-newRegion)
  testDat_all <- modDat_1_s %>%
    slice(test) %>%
    dplyr::select(-newRegion)

  # get the model matrices for input and response variables for cross validation model specification
  X_train <- as.matrix(X_cv[train,])
  X_test <- as.matrix(X_cv[test,])

  y_train <- modDat_1_s[train,"TotalTreeCover_binom"]
  y_test <- modDat_1_s[test,"TotalTreeCover_binom"]

  # get the model matrices for input and response variables for original model specification
  X_train_glob <- as.matrix(X[train,])
  X_test_glob <- as.matrix(X[test,])

  y_train_glob <- modDat_1_s[train,"TotalTreeCover_binom"]
  y_test_glob <- modDat_1_s[test,"TotalTreeCover_binom"]

  train_eco <- modDat_1_s$NA_L2NAME[train]

  ## just try a regular glm w/ the components from the global model
  fit_i <- glm(data = trainDat_all, formula = f_cv,
    ,
               family =  binomial
    )

    coef(fit_i)

  # lasso model predictions with the optimal lambda
  optimal_pred <- predict(fit_i, newdata= testDat_all, type = "response") 
  # null model and predictions
  # the "null" model in this case is the global model
  # predict on the test data for this iteration w/ the global model
  null_pred <- predict.glm(mod_secondBest, newdata = testDat_all, type = "response") 

  # save data
  tmp <- data.frame(ecoRegion_holdout = rep(test_eco,length(y_test)),
                    obs=y_test,
                    pred_opt=optimal_pred,
                    pred_null=null_pred#,
                    #pred_nopenalty=nopen_pred
                    ) %>%
    cbind(testDat_all)

  # calculate RMSE, bias, etc. of
  # RMSE of CV model
  RMSE_optimal <- yardstick::rmse(data = data.frame(optimal_pred, "y_test" = (y_test)), truth = "y_test", estimate = "optimal_pred")[1,]$.estimate
  # RMSE of global model
  RMSE_null <- yardstick::rmse(data = data.frame(null_pred,  "y_test" = (y_test)), truth = "y_test", estimate = "null_pred")[1,]$.estimate
  # bias of CV model
  bias_optimal <- mean((y_test) - optimal_pred)
  # bias of global model
  bias_null <-  mean((y_test) - null_pred )

  # put output into a list
  tmpList <- list("testRegion" = i_eco,
    "modelObject" = fit_i,
       "modelPredictions" = tmp,
    "performanceMetrics" = data.frame("RMSE_cvModel" = RMSE_optimal,
                                      "RMSE_globalModel" = RMSE_null,
                                      "bias_cvModel" = bias_optimal,
                                      "bias_globalModel" = bias_null))

  # save model outputs
  outList[[which(sort(unique(modDat_1_s$NA_L2NAME)) == i_eco)]] <- tmpList
 }
}
```

### Below are the RMSE and bias values for predictions made for each holdout level II ecoregion, compared to predictions from the global model for that same ecoregion
```{r metrics output table for 1SE lambda CV}

if (length(prednames_secondBestMod) == 0) {
  print("The model only contains one predictor (an intercept), so cross validation isn't practical")
} else {
# table of model performance
purrr::map(outList, .f = function(x) {
  cbind(data.frame("holdout region" = x$testRegion),  x$performanceMetrics)
}
) %>%
  purrr::list_rbind() %>%
  kable(col.names = c("Held-out ecoregion", "RMSE of CV model", "RMSE of global model",
                      "bias of CV model - mean(obs-pred.)", "bias of global model - mean(obs-pred.)"),
        caption = "Performance of Cross Validation using 'second best lambda' model specification") %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
}
```

```{r figures of CV for 1 SE lambda model}
if (length(prednames_secondBestMod) == 0) {
  print("The model only contains one predictor (an intercept), so cross validation isn't practical")
} else {
for (i in 1:length(unique(modDat_1_s$NA_L2NAME))) {
  holdoutRegion <- outList[[i]]$testRegion
  predictionData <- outList[[i]]$modelPredictions
  modTerms <- as.matrix(coef(outList[[i]]$modelObject)) %>%
    as.data.frame() %>%
    filter(V1!=0) %>%
    rownames()

  # calculate residuals
  predictionData <- predictionData %>%
  mutate(resid = .[["obs"]] - .[["pred_opt"]] ,
         resid_globMod = .[["obs"]]  - .[["pred_null"]])


# rasterize
# use 'test_rast' from earlier

  # rasterize data
plotObs <- predictionData %>%
         drop_na(paste("TotalTreeCover_binom")) %>%
  #slice_sample(n = 5e4) %>%
  terra::vect(geom = c("x", "y")) %>%
  terra::set.crs(crs(test_rast)) %>%
  terra::rasterize(y = test_rast,
                   field = "resid",
                   fun = mean) 

tempExt <- crds(plotObs, na.rm = TRUE)

plotObs_2 <- plotObs %>%
  crop(ext(min(tempExt[,1]), max(tempExt[,1]),
           min(tempExt[,2]), max(tempExt[,2]))
       )

# make figures
# make histogram
hist_i <- ggplot(predictionData) +
  geom_histogram(aes(resid_globMod), col = "darkgrey", fill = "lightgrey") +
  xlab(c("Residuals (obs. - pred.)"))
# make map
map_i <-  ggplot() +
geom_spatraster(data = plotObs_2) +
  geom_sf(data=cropped_states_2,fill=NA ) +
  geom_sf(data = mapRegions, fill = NA, col = "orchid", lwd = .5) +
labs(title = paste0("Residuals (obs. - pred.) for predictions of \n", holdoutRegion, " \n from a model fit to other ecoregions"),
     subtitle = paste0("TotalTreeCover_binom", " ~ ", paste0( modTerms, collapse = " + "))) +
  scale_fill_gradient2(low = "red",
                       mid = "white" ,
                       high = "blue" ,
                       midpoint = 0,   na.value = "darkgrey",
                       limits = c(-1, 1))  + 
  xlim(st_bbox(plotObs_2)[1],st_bbox(plotObs_2)[3]) + 
  ylim(st_bbox(plotObs_2)[2],st_bbox(plotObs_2)[4])

 assign(paste0("residPlot_",holdoutRegion),
   value = ggarrange(map_i, hist_i, heights = c(3,1), ncol = 1)
)

}

  lapply(unique(modDat_1_s$NA_L2NAME), FUN = function(x) {
    get(paste0("residPlot_", x))
  })
}

```


# Save output

```{r save_output}
## save the coefficients for the models (best lambda, 1/2se lambda, 1se lambda)
if(trimAnom == TRUE) {
saveRDS(coefs, file = paste0("./models/yesOrNoTrees_modelCoefficients.rds"))
saveRDS(uniqueCoeffs, file = paste0("./models/yesOrNoTrees_modelMetrics_trimAnom.rds"))
} else {
saveRDS(coefs, file = paste0("./models/yesOrNoTrees_modelCoefficients.rds"))
saveRDS(uniqueCoeffs, file = paste0("./models/yesOrNoTrees_modelMetrics_trimAnom.rds"))
}
# make a table

```




```{r pdp_glm, warning = FALSE, fig.width = 8, fig.height=8}
## partial dependence plots
#vip::vip(mod_glmFinal, num_features = 15)

#pdp_all_vars(mod_glmFinal, mod_vars = pred_vars, ylab = 'probability',train = df_small)

#caret::varImp(fit)
```


# session info

Hash of current commit (i.e. to ID the version of the code used)

```{r}
system("git rev-parse HEAD", intern=TRUE)
```

Packages etc.

```{r}
sessionInfo()

```