---
title: "Evaluating spatial blocking of vegetation composition data for further modeling"
subtitle: "Based on code from Gregor Siegmund"
author: "Alice Stears"
date: "`r lubridate::today()`"
params: 
  test_run: FALSE
  save_figs: FALSE
  #inter: !r c('afgAGB:MAP' = "afgAGB:MAP")
  sample_group: 1
  byRegion: TRUE
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
---

The purpose of this script is to investigate whether spatial blocking (as recommended in (Roberts et al., 2016)) is a viable strategy for addressing structure in the vegetation composition and climate data that we plan to use in these analyses. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE,
                      message = FALSE)
```

# Prepare data and software pacakges
```{r, message= FALSE}
library(tidyverse)
library(blockCV)
library(khroma)
library(sf)
library(here)

# read in vegetation composition dataset (that includes climate and weather predictors) 
here::i_am("Analysis/VegComposition/ModelFitting/01_PredictorVarSelection.Rmd")
dat <- readRDS( here("Data_processed", "CoverData", "DataForModels_spatiallyAveraged_withSoils_noSf.rds"))
## there are some values of the annual wet degree days 5th percentile that have -Inf?? change to lowest value for now? 
dat[is.infinite(dat$annWetDegDays_5percentile_3yrAnom), "annWetDegDays_5percentile_3yrAnom"] <- -47.8
## same, but for annual water deficit 95th percentile 
dat[is.infinite(dat$annWaterDeficit_95percentile_3yrAnom), "annWaterDeficit_95percentile_3yrAnom"] <- -600

## break into grasslandshrub and forest datasets, and only include climate/soil/weather variables selected from the previous analysis
datForest <- dat %>% 
  filter(newRegion %in% c("westForest", "eastForest")) %>% 
  select("tmin_meanAnnAvg_CLIM"  , "prcp_meanAnnTotal_CLIM"  ,
         "T_warmestMonth_meanAnnAvg_CLIM" , "precip_driestMonth_meanAnnAvg_CLIM",
         "PrecipTempCorr_meanAnnAvg_CLIM" , "isothermality_meanAnnAvg_CLIM" ,
         "annWetDegDays_meanAnnAvg_CLIM" , "avgSandPerc_acrossDepth" ,
         "avgCoarsePerc_acrossDepth" , "avgOrganicCarbonPerc_0_3cm" , 
         "totalAvailableWaterHoldingCapacity" , "tmin_meanAnnAvg_3yrAnom", 
         "tmax_meanAnnAvg_3yrAnom"    , "prcp_meanAnnTotal_3yrAnom"          , "T_warmestMonth_meanAnnAvg_3yrAnom"           , "T_warmestMonth_meanAnnAvg_3yrAnom"            
,"precip_wettestMonth_meanAnnAvg_3yrAnom"           , "precip_driestMonth_meanAnnAvg_3yrAnom"     , 
"precip_Seasonality_meanAnnAvg_3yrAnom" , "PrecipTempCorr_meanAnnAvg_3yrAnom"      
,"aboveFreezing_month_meanAnnAvg_3yrAnom" , "isothermality_meanAnnAvg_3yrAnom" , "annWaterDeficit_meanAnnAvg_3yrAnom"        , "annWetDegDays_meanAnnAvg_3yrAnom"     
,"annVPD_min_meanAnnAvg_3yrAnom"            , "annVPD_max_95percentile_3yrAnom"    , "annWaterDeficit_95percentile_3yrAnom"     , "annWetDegDays_5percentile_3yrAnom"   
,"durationFrostFreeDays_5percentile_3yrAnom" , 
"TotalTreeCover", "Year.x", "CAMCover", "TotalHerbaceousCover"        ,  "BareGroundCover" ,                          "BroadleavedTreeCover_prop"    ,  "NeedleLeavedTreeCover_prop",  "C4Cover_prop"                 ,  "C3Cover_prop",                              "ForbCover_prop"                 ,  "ShrubCover"                     ,  "Long" ,        "Lat"  )

datGrassShrub <- dat %>% 
  filter(newRegion == "dryShrubGrass") %>% 
  select("tmin_meanAnnAvg_CLIM", "tmean_meanAnnAvg_CLIM", "prcp_meanAnnTotal_CLIM", 
         "precip_Seasonality_meanAnnAvg_CLIM", "PrecipTempCorr_meanAnnAvg_CLIM", "avgSandPerc_acrossDepth",                   "avgCoarsePerc_acrossDepth" ,   "totalAvailableWaterHoldingCapacity"  , "tmin_meanAnnAvg_3yrAnom" , "tmax_meanAnnAvg_3yrAnom" ,  "prcp_meanAnnTotal_3yrAnom", "T_warmestMonth_meanAnnAvg_3yrAnom"       , "T_coldestMonth_meanAnnAvg_3yrAnom" ,"precip_driestMonth_meanAnnAvg_3yrAnom"   , "precip_Seasonality_meanAnnAvg_3yrAnom" ,  "PrecipTempCorr_meanAnnAvg_3yrAnom" , "aboveFreezing_month_meanAnnAvg_3yrAnom", "isothermality_meanAnnAvg_3yrAnom" ,     "annWaterDeficit_meanAnnAvg_3yrAnom" ,   "annWetDegDays_meanAnnAvg_3yrAnom"  , "annVPD_min_meanAnnAvg_3yrAnom"    ,        "annVPD_max_95percentile_3yrAnom" ,        "annWaterDeficit_95percentile_3yrAnom"   ,    "annWetDegDays_5percentile_3yrAnom", "durationFrostFreeDays_5percentile_3yrAnom" , 
"TotalTreeCover", "Year.x", "CAMCover", "TotalHerbaceousCover"        ,  "BareGroundCover" ,                          "BroadleavedTreeCover_prop"    ,  "NeedleLeavedTreeCover_prop",  "C4Cover_prop"                 ,  "C3Cover_prop",                              "ForbCover_prop"                 ,  "ShrubCover"                     ,  "Long" ,        "Lat"    
         )
```

# Determining which type of blocking as best suited to our approach
## Forest

Prepare data for spatial blocking
```{r}

datForest_sf <- 
  datForest %>% 
  st_as_sf(coords = c("Long", "Lat"), 
           crs = "PROJCRS[\"unnamed\",\n    BASEGEOGCRS[\"unknown\",\n        DATUM[\"unknown\",\n            ELLIPSOID[\"Spheroid\",6378137,298.257223563,\n                LENGTHUNIT[\"metre\",1,\n                    ID[\"EPSG\",9001]]]],\n        PRIMEM[\"Greenwich\",0,\n            ANGLEUNIT[\"degree\",0.0174532925199433,\n                ID[\"EPSG\",9122]]]],\n    CONVERSION[\"Lambert Conic Conformal (2SP)\",\n        METHOD[\"Lambert Conic Conformal (2SP)\",\n            ID[\"EPSG\",9802]],\n        PARAMETER[\"Latitude of false origin\",42.5,\n            ANGLEUNIT[\"degree\",0.0174532925199433],\n            ID[\"EPSG\",8821]],\n        PARAMETER[\"Longitude of false origin\",-100,\n            ANGLEUNIT[\"degree\",0.0174532925199433],\n            ID[\"EPSG\",8822]],\n        PARAMETER[\"Latitude of 1st standard parallel\",25,\n            ANGLEUNIT[\"degree\",0.0174532925199433],\n            ID[\"EPSG\",8823]],\n        PARAMETER[\"Latitude of 2nd standard parallel\",60,\n            ANGLEUNIT[\"degree\",0.0174532925199433],\n            ID[\"EPSG\",8824]],\n        PARAMETER[\"Easting at false origin\",0,\n            LENGTHUNIT[\"metre\",1],\n            ID[\"EPSG\",8826]],\n        PARAMETER[\"Northing at false origin\",0,\n            LENGTHUNIT[\"metre\",1],\n            ID[\"EPSG\",8827]]],\n    CS[Cartesian,2],\n        AXIS[\"easting\",east,\n            ORDER[1],\n            LENGTHUNIT[\"metre\",1,\n                ID[\"EPSG\",9001]]],\n        AXIS[\"northing\",north,\n            ORDER[2],\n            LENGTHUNIT[\"metre\",1,\n                ID[\"EPSG\",9001]]]]")

# - + Map: all survey locations  ----

us_states <- tigris::states()

cropped_states <- us_states %>%
  dplyr::filter(STUSPS!="HI") %>%
  dplyr::filter(STUSPS!="AK") %>%
  dplyr::filter(STUSPS!="Puerto Rico") %>%
  dplyr::filter(STUSPS!="American Samoa") %>%
  dplyr::filter(STUSPS!="Guam") %>%
  dplyr::filter(STUSPS!="Commonwealth of the Northern Mariana Islands") %>%
  dplyr::filter(STUSPS!="United States Virgin Islands") %>%

  sf::st_sf() %>%
  sf::st_transform(sf::st_crs(datForest_sf)) %>%
  sf::st_crop(sf::st_bbox(datForest_sf)+c(-1,-1,1,1))
plot(cropped_states$geometry)
```

```{r}
# - + Spatial block ---
datForest_sf %>% 
  st_drop_geometry() %>% 
  select(TotalTreeCover, CAMCover:ShrubCover) %>% 
  summarize_all(.funs = function(x) {sum(!is.na(x))})
## shrub cover has the most observations, so use this 

datForest_sf.sp <- datForest_sf %>%
  select(tmin_meanAnnAvg_CLIM:durationFrostFreeDays_5percentile_3yrAnom, ShrubCover, geometry) %>% 
  filter(!is.na(ShrubCover)) %>% 
  drop_na()

# create the folds (k = 6)
block.sp <- cv_spatial(x = datForest_sf.sp, 
                       hexagon = FALSE, 
                       k = 6, report = FALSE, plot = FALSE, progress = FALSE)

# add the fold information to the raw data
datForest_sf.fold <- datForest_sf.sp %>% 
  dplyr::bind_cols(fold = block.sp$folds_ids) %>%
  dplyr::mutate(fold = as.factor(fold)) %>% 
  left_join(datForest %>% select(names(datForest_sf)[1:27], "Long", "Lat"))
```
Explore spatial autocorrelation
```{r}
autocor <- cv_spatial_autocor(x = datForest_sf.sp, 
                   column = "ShrubCover")
autocor$plots
```


Determine regular spatial blocking using the most common cover variable as the response 

k = 6


```{r make figures of spatial blocking}
#plot(datForest_sf.fold$geometry, col = as.factor(datForest_sf.fold$fold))

map1 <- ggplot() +
  geom_sf(data=cropped_states,fill='white') +
  geom_sf(data=block.sp$blocks,aes(fill=as.factor(folds)),linewidth=0.5,alpha=0.5) +
  geom_point(data=datForest_sf.fold,alpha=0.8, 
             aes(x = Long, y = Lat, color=as.factor(datForest_sf.fold$fold))) +
  scale_fill_okabeito() +
  scale_color_okabeito() +
 # theme_default() +
  theme(legend.position = 'none') +
  labs(title = "Spatial blocks (k=5)")

# how are tmin_clim and MAP correlated w/in a fold? 
# (datForest_sf.fold %>%
#   ggplot() +
#   geom_point(aes(x=tmin_meanAnnAvg_CLIM,y=prcp_meanAnnTotal_CLIM,color=fold), alpha = .1)  + 
#   geom_smooth(aes(x=tmin_meanAnnAvg_CLIM,y=prcp_meanAnnTotal_CLIM,color=fold), method = "lm") + 
#   xlab("Annual Average T_min - long-term average") +
#   ylab("Annual Average Precip - long-term average") + 
#     theme_minimal())

hull <- datForest_sf.fold %>%
  ungroup() %>%
  group_by(fold) %>%
  slice(chull(tmin_meanAnnAvg_CLIM, prcp_meanAnnTotal_CLIM))

plot1<-ggplot(data=datForest_sf.fold,aes(x=tmin_meanAnnAvg_CLIM,y=prcp_meanAnnTotal_CLIM)) +
  geom_polygon(data = hull, alpha = 0.25,aes(fill=fold) )+
  geom_point(aes(group=fold,color=fold),alpha=0.25) +
  theme_minimal() + xlab("Annual Average T_min - long-term average") +
  ylab("Annual Average Precip - long-term average") +
  scale_color_okabeito() +
  scale_fill_okabeito()

plot2<-ggplot(data=datForest_sf.fold %>%
                pivot_longer(cols=tmin_meanAnnAvg_CLIM:prcp_meanAnnTotal_CLIM),
              aes(x=value,group=name)) +
  # geom_polygon(data = hull, alpha = 0.25,aes(fill=fold) )+
  geom_density(aes(group=fold,fill=fold),alpha=0.25) +
  theme_minimal() +
  facet_wrap(~name,scales='free') +
  scale_color_okabeito() +
  scale_fill_okabeito()
 
library(patchwork)
(combo <- (map1+plot1)/plot2) 

```

Try different k values: 
k = 5 
```{r}

# create the folds (k = 6)
block.sp <- cv_spatial(x = datForest_sf.sp, 
                       hexagon = FALSE, 
                       k = 5, report = FALSE, plot = FALSE, progress = FALSE)

# add the fold information to the raw data
datForest_sf.fold <- datForest_sf.sp %>% 
  dplyr::bind_cols(fold = block.sp$folds_ids) %>%
  dplyr::mutate(fold = as.factor(fold)) %>% 
  left_join(datForest %>% select(names(datForest_sf)[1:27], "Long", "Lat"))

#plot(datForest_sf.fold$geometry, col = as.factor(datForest_sf.fold$fold))

map1 <- ggplot() +
  geom_sf(data=cropped_states,fill='white') +
  geom_sf(data=block.sp$blocks,aes(fill=as.factor(folds)),linewidth=0.5,alpha=0.5) +
  geom_point(data=datForest_sf.fold,alpha=0.8, 
             aes(x = Long, y = Lat, color=as.factor(datForest_sf.fold$fold))) +
  scale_fill_okabeito() +
  scale_color_okabeito() +
 # theme_default() +
  theme(legend.position = 'none') +
  labs(title = "Spatial blocks (k=5)")

# how are tmin_clim and MAP correlated w/in a fold? 
# (datForest_sf.fold %>%
#   ggplot() +
#   geom_point(aes(x=tmin_meanAnnAvg_CLIM,y=prcp_meanAnnTotal_CLIM,color=fold), alpha = .1)  + 
#   geom_smooth(aes(x=tmin_meanAnnAvg_CLIM,y=prcp_meanAnnTotal_CLIM,color=fold), method = "lm") + 
#   xlab("Annual Average T_min - long-term average") +
#   ylab("Annual Average Precip - long-term average") + 
#     theme_minimal())

hull <- datForest_sf.fold %>%
  ungroup() %>%
  group_by(fold) %>%
  slice(chull(tmin_meanAnnAvg_CLIM, prcp_meanAnnTotal_CLIM))

plot1<-ggplot(data=datForest_sf.fold,aes(x=tmin_meanAnnAvg_CLIM,y=prcp_meanAnnTotal_CLIM)) +
  geom_polygon(data = hull, alpha = 0.25,aes(fill=fold) )+
  geom_point(aes(group=fold,color=fold),alpha=0.25) +
  theme_minimal() + xlab("Annual Average T_min - long-term average") +
  ylab("Annual Average Precip - long-term average") +
  scale_color_okabeito() +
  scale_fill_okabeito()

plot2<-ggplot(data=datForest_sf.fold %>%
                pivot_longer(cols=tmin_meanAnnAvg_CLIM:prcp_meanAnnTotal_CLIM),
              aes(x=value,group=name)) +
  # geom_polygon(data = hull, alpha = 0.25,aes(fill=fold) )+
  geom_density(aes(group=fold,fill=fold),alpha=0.25) +
  theme_minimal() +
  facet_wrap(~name,scales='free') +
  scale_color_okabeito() +
  scale_fill_okabeito()

(combo <- (map1+plot1)/plot2) 
```

k = 3
```{r}

# create the folds (k = 6)
block.sp <- cv_spatial(x = datForest_sf.sp, 
                       hexagon = FALSE, 
                       k = 3, report = FALSE, plot = FALSE, progress = FALSE)

# add the fold information to the raw data
datForest_sf.fold <- datForest_sf.sp %>% 
  dplyr::bind_cols(fold = block.sp$folds_ids) %>%
  dplyr::mutate(fold = as.factor(fold)) %>% 
  left_join(datForest %>% select(names(datForest_sf)[1:27], "Long", "Lat"))

#plot(datForest_sf.fold$geometry, col = as.factor(datForest_sf.fold$fold))

map1 <- ggplot() +
  geom_sf(data=cropped_states,fill='white') +
  geom_sf(data=block.sp$blocks,aes(fill=as.factor(folds)),linewidth=0.5,alpha=0.5) +
  geom_point(data=datForest_sf.fold,alpha=0.8, 
             aes(x = Long, y = Lat, color=as.factor(datForest_sf.fold$fold))) +
  scale_fill_okabeito() +
  scale_color_okabeito() +
 # theme_default() +
  theme(legend.position = 'none') +
  labs(title = "Spatial blocks (k=5)")

# how are tmin_clim and MAP correlated w/in a fold? 
# (datForest_sf.fold %>%
#   ggplot() +
#   geom_point(aes(x=tmin_meanAnnAvg_CLIM,y=prcp_meanAnnTotal_CLIM,color=fold), alpha = .1)  + 
#   geom_smooth(aes(x=tmin_meanAnnAvg_CLIM,y=prcp_meanAnnTotal_CLIM,color=fold), method = "lm") + 
#   xlab("Annual Average T_min - long-term average") +
#   ylab("Annual Average Precip - long-term average") + 
#     theme_minimal())

hull <- datForest_sf.fold %>%
  ungroup() %>%
  group_by(fold) %>%
  slice(chull(tmin_meanAnnAvg_CLIM, prcp_meanAnnTotal_CLIM))

plot1<-ggplot(data=datForest_sf.fold,aes(x=tmin_meanAnnAvg_CLIM,y=prcp_meanAnnTotal_CLIM)) +
  geom_polygon(data = hull, alpha = 0.25,aes(fill=fold) )+
  geom_point(aes(group=fold,color=fold),alpha=0.25) +
  theme_minimal() + xlab("Annual Average T_min - long-term average") +
  ylab("Annual Average Precip - long-term average") +
  scale_color_okabeito() +
  scale_fill_okabeito()

plot2<-ggplot(data=datForest_sf.fold %>%
                pivot_longer(cols=tmin_meanAnnAvg_CLIM:prcp_meanAnnTotal_CLIM),
              aes(x=value,group=name)) +
  # geom_polygon(data = hull, alpha = 0.25,aes(fill=fold) )+
  geom_density(aes(group=fold,fill=fold),alpha=0.25) +
  theme_minimal() +
  facet_wrap(~name,scales='free') +
  scale_color_okabeito() +
  scale_fill_okabeito()

(combo <- (map1+plot1)/plot2) 
```

Look at potential block sizes
```{r}
cv_block_size(x = 
datForest_sf.sp)
```

```{r}
# 
# library(patchwork)
# 
# combo <- (map1+plot1)/plot2 
# 
# combo %>%   
#   ggsave(filename="~/Downloads/spatial-block.png",width=10,height=8,  bg = "transparent")
# 
# datForest_sf.fold %>%
#   sf::st_drop_geometry() %>%
#   dplyr::group_by(fold,yearSeeding) %>%
#   dplyr::filter(!is.na(fold)) %>%
#   tally() %>%
#   ggplot(aes(x=yearSeeding,y=n)) +
#   geom_col(aes(fill=fold)) +
# #  facet_wrap(~fold) +
#   theme_default() +
#   scale_fill_okabeito()
# 
# df.split <- datForest_sf.fold %>%
#   sf::st_drop_geometry() %>%
#   dplyr::group_by(fold,yearSeeding) %>%
#   dplyr::filter(!is.na(fold)) %>%
#   tally() %>%
#   group_by(fold) %>%
#   dplyr::mutate(p=cumsum(n)/sum(n),
#                 split = case_when(p>=0.8~'test',.default='train')) 
# 
# spaceTimeBlock <- datForest_sf.fold %>%
#   sf::st_drop_geometry() %>%
#   dplyr::group_by(fold,yearSeeding) %>%
#   dplyr::filter(!is.na(fold)) %>%
#   tally() %>%
#   left_join(df.split) %>%
#   mutate(split.alpha = case_when(split=='train'~1,.default=0.25)) %>%
#   ggplot(aes(x=yearSeeding,y=n)) +
#   geom_col(aes(fill=fold,alpha=split.alpha)) +
#   #  facet_wrap(~fold) +
#   theme_default() +
#   scale_fill_okabeito() +
#   theme(legend.position='none')
# 
# spaceTimeBlock %>%   
#   ggsave(filename="~/Downloads/spatial-time-split.png",width=8,height=4,  bg = "transparent")
# 
# # - + Environmental block ---
# 
# sampr <- datForest_sf.sp %>% select(PPT,Tmean) %>% sf::st_drop_geometry()
# 
# kms <- stats::kmeans(sampr, centers = 6, iter.max = 500, 
#                      nstart = 25)
# datForest_sf.sp$fold = kms$cluster
# 
# datForest_sf.fold <- datForest_sf.sp %>% 
#   dplyr::mutate(fold = as.factor(fold))
# 
# map1 <- ggplot() +
#   geom_sf(data=cropped_states,fill='white') +
#  # geom_sf(data=block.sp$blocks,aes(fill=as.factor(folds)),linewidth=0.5,alpha=0.5) +
#   geom_sf(data=datForest_sf.sp,aes(col=as.factor(fold)),alpha=0.25) +
#   scale_fill_okabeito() +
#   theme_default() +
#   theme(legend.position = 'none') +
#   labs(title = "Environmental blocks (k=6)")
# 
# datForest_sf.fold %>%
#   ggplot() +
#   geom_point(aes(x=Tmean,y=PPT,color=fold)) +
#   theme_default() 
# 
# hull <- datForest_sf.fold %>%
#   ungroup() %>%
#   group_by(fold) %>%
#   slice(chull(Tmean, PPT))
# 
# plot1<-ggplot(data=datForest_sf.fold,aes(x=Tmean,y=PPT)) +
#   geom_polygon(data = hull, alpha = 0.25,aes(fill=fold) )+
#   geom_point(aes(group=fold,color=fold),alpha=0.25) +
#   theme_default() +
#   scale_color_okabeito() +
#   scale_fill_okabeito()
# 
# plot2<-ggplot(data=datForest_sf.fold %>%
#                 pivot_longer(cols=Tmean:PPT),
#               aes(x=value,group=name)) +
#   # geom_polygon(data = hull, alpha = 0.25,aes(fill=fold) )+
#   geom_density(aes(group=fold,fill=fold),alpha=0.25) +
#   theme_default() +
#   facet_wrap(~name,scales='free') +
#   scale_color_okabeito() +
#   scale_fill_okabeito()
# 
# library(patchwork)
# 
# combo <- (map1+plot1)/plot2 
# 
# combo %>%   
#   ggsave(filename="~/Downloads/environmental-block.png",width=10,height=8,  bg = "transparent")
# 
# datForest_sf.fold %>%
#   sf::st_drop_geometry() %>%
#   dplyr::group_by(fold,yearSeeding) %>%
#   dplyr::filter(!is.na(fold)) %>%
#   tally() %>%
#   ggplot(aes(x=yearSeeding,y=n)) +
#   geom_col(aes(fill=fold)) +
#   #  facet_wrap(~fold) +
#   theme_default() +
#   scale_fill_okabeito()
# 
# df.split <- datForest_sf.fold %>%
#   sf::st_drop_geometry() %>%
#   dplyr::group_by(fold,yearSeeding) %>%
#   dplyr::filter(!is.na(fold)) %>%
#   tally() %>%
#   group_by(fold) %>%
#   dplyr::mutate(p=cumsum(n)/sum(n),
#                 split = case_when(p>=0.8~'test',.default='train')) 
# 
# spaceTimeBlock <- datForest_sf.fold %>%
#   sf::st_drop_geometry() %>%
#   dplyr::group_by(fold,yearSeeding) %>%
#   dplyr::filter(!is.na(fold)) %>%
#   tally() %>%
#   left_join(df.split) %>%
#   mutate(split.alpha = case_when(split=='train'~1,.default=0.25)) %>%
#   ggplot(aes(x=yearSeeding,y=n)) +
#   geom_col(aes(fill=fold,alpha=split.alpha)) +
#   #  facet_wrap(~fold) +
#   theme_default() +
#   scale_fill_okabeito() +
#   theme(legend.position='none')
# 
# spaceTimeBlock %>%   
#   ggsave(filename="~/Downloads/spatial-time-split-env.png",width=8,height=4,  bg = "transparent")
```



