---
title: "02_modelFitting_zeroTreeModel"
author: "Alice Stears"
date: "`r lubridate::today()`"
output:
  html_document:
    toc: true
    df_print: paged
    code_folding: hide
    dev: "jpeg"
params:
  run: FALSE      
  save_figs: TRUE
  ecoregion: CONUS
  response: fracNotForest
  treeThreshold: 10
  whichSecondBestMod: "auto"# 1se OR halfse OR auto (chose the model w/ fewest # of predictors)
  thresholdMethod: "PredPrev=Obs" # approach for identifying best threshold
subtitle: "Based on code from Martin Holdrege's Sagebrush-Fire paper"
---
This workflow fits a model across all of CONUS that predicts whether a location has trees or no trees. 

The training data consists of classificaiton of treed or un-treed based on RAP, 
as well as climate variables from DayMet, which have been aggregated into mean interannual conditions accross multiple temporal windows. 

```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = TRUE, warning=FALSE,
#                       message = FALSE, dpi = 144)
```

# Dependencies 

Set user defined parameters

```{r}
print(params)
# set to true if want to run for a limited number of rows (i.e. for code testing)
test_run <- params$test_run
save_figs <- params$save_figs
response <- params$response
fit_sample <- TRUE # fit model to a sample of the data
n_train <- 5e4 # sample size of the training data
n_test <- 1e6 # sample size of the testing data (if this is too big the decile dotplot code throws memory errors)
run <- params$run
whichSecondBestMod <- params$whichSecondBestMod
treeThreshold <- params$treeThreshold
thresholdMethod <- params$thresholdMethod
```

Load packages and source functions
```{r warning=FALSE, message=FALSE}
# set option so resampled dataset created here reproduces earlier runs of this code with dplyr 1.0.10
source("../../../Functions/glmTransformsIterates.R")
source("../../../Functions/transformPreds.R")
source("../../../Functions/betaLASSO.R")

#source("../../../Functions/StepBeta_mine.R")
#source("src/fig_params.R")
#source("src/modeling_functions.R")
 
library(betareg)
library(ggspatial)
library(terra)
library(tidyterra)
library(sf)
library(caret)
library(tidyverse)
library(GGally) # for ggpairs()
library(pdp) # for partial dependence plots
library(gridExtra)
library(knitr)
library(patchwork) # for figure insets etc. 
library(ggtext)
library(StepBeta)
theme_set(theme_classic())
library(here)
library(rsample)
library(kableExtra)
library(glmnet)
library(USAboundaries)
library(cvms)
library(rsvg)
library(ggimage)
library(doMC)
library(parallel)
library(pROC)

## functions
           get_inset <- function(df, densPlotName) {
              p <- ggplot(data = df) +
             #facet_wrap(~lyr, ncol = 1, scales = "free_y") +
                geom_density(aes(densPlotName)) +
                theme_bw(base_size=9) +  ## makes everything smaller
                theme(panel.background = element_rect(fill="transparent"),  ## white plot background
                      panel.border = element_blank(),
          axis.title.y = element_blank(),
          axis.title.x = element_blank(),
          axis.text.x = element_text(size=rel(2)), ## tiny axis text
          axis.text.y = element_text(size=rel(2)), ## tiny axis text
          axis.line.x = element_line(size = .5),
          axis.line.y = element_line(size = .5),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          plot.background = element_blank(),
          strip.background = element_blank(),
          strip.text = element_blank()) 
  return(p)
           }
           # custom function to add inset to main plot
           annotation_custom2 <- function (grob, xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, data)
{
  layer(data = data, stat = StatIdentity, position = PositionIdentity,
        geom = ggplot2:::GeomCustomAnn,
        inherit.aes = TRUE, params = list(grob = grob,
                                          xmin = xmin, xmax = xmax,
                                          ymin = ymin, ymax = ymax))
           }
```

Source the script that calculates the proportion of CONUS by area that is un-forested, which we will use to calibrate our model
```{r message=FALSE}
source("../DataPrep/09_non-forest_area.R")
# object w/ the raster of the forest locations is called notForest
# object w/ the proportion of the unmasked raster area that is not forested (<1% trees in a grid cell) is called area_not_forest_km2
# object w/ the mean proportion non-forest (area-weighted) is called mean_frac
```

# Read in data

Data compiled in the `prepDataForModels.R` script

```{r}
here::i_am("Analysis/VegComposition/ModelFitting/02B_modelFitting_zeroTreeModel.Rmd")
modDat <- readRDS( here("Data_processed", "CoverData", "DataForModels_spatiallyAveraged_withSoils_noSf_sampledLANDFIRE.rds")) %>% st_drop_geometry()
# get scaled data
modDat_1_s <- readRDS("./models/scaledModelInputData.rds")
```

# Prep data

We will fit a beta glm that predicts the continuous probability that a location is un-forested 
```{r}
## get cover data, which we will use to get climate and soils data
modDat <- modDat %>% 
  mutate(TotalTreeCover_binom = replace(TotalTreeCover, TotalTreeCover <=treeThreshold, 0)) %>% 
  mutate(TotalTreeCover_binom = replace(TotalTreeCover_binom, TotalTreeCover_binom > treeThreshold, 1))
## get dayMet raster
test_rast_dayMet_TEMP <- rast("../../../Data_raw/dayMet/rawMonthlyData/orders/70e0da02b9d2d6e8faa8c97d211f3546/Daymet_Monthly_V4R1/data/daymet_v4_prcp_monttl_na_1980.tif") 
```
```{r}
## get tree/no-tree data

###Get data that shows un-forested locations, based on RAP. This dataset is created by first identifying RAP 30-m pixels that have 3% or less of tree cover. Then that information is aggregated up to the dayMet scale, and the value provided is the percentage of RAP pixels within a given dayMet grid cell that have <3% tree cover. 

## note that this dataset uses the the same LCMAP land-use mask that the cover data used, and uses an even more conservative burned-area raster. 
## get RAP unforested cover dataset from Martin
  #
nonForest_rast <- terra::rast("../../../Data_processed/RAP_v3_fracNotForest_lt3_mask-lcmap50_2019-2023_1000m.tif") %>% 
  # make sure the crs is the same
  terra::project(crs(test_rast_dayMet_TEMP))
# ## make true zeros 0.000001 and 1 0.999999 (so the beta regression will run)
# nonForest_rast_FULL <- nonForest_rastTemp
# nonForest_rast_FULL[nonForest_rast_FULL$fracNotForest==0] <- 0.000001
# nonForest_rast_FULL[nonForest_rast_FULL$fracNotForest==1] <- 0.99999
# 
# ## get RAP forest cover data. We trim the nonForest_rast layer to locations with 50% or lower tree cover, so that the model has less data than CONUS-wide to fit the non-forest model, but also has enough variation in tree/no-tree to fit the model
# RAP_50PercentForestCover <- terra::rast("../../../Data_processed/RAP_v3_cover_2019-2023_1000m.tif")
# RAP_50PercentForestCover <- RAP_50PercentForestCover$totalTreeCov
# # filter to only show data from locations w/ 50% or less of tree cover
# RAP_50PercentForestCover[RAP_50PercentForestCover>50] <- NA
# # apply the mask from the nonForest_rast raster to the RAP_50PercentForestCover raster
# RAP_50PercentForestCover <- RAP_50PercentForestCover %>% mask(nonForest_rast_FULL, maskvalue = NA)
# 
# ## now, trim the nonForest data such that we are only training the model on data from locations where RAP total tree cover is <=50%
# nonForest_rast <- nonForest_rast_FULL %>% mask(RAP_50PercentForestCover, maskvalue = NA)
```

This tree/un-treed data gives the proportion of 30-m RAP pixels within a 1km dayMet grid cell that have no trees (which we define as <3% tree cover). The resulting raster has pixels with values on a continuous range between 0 and 1 (0 is completely un-treed, 1 is completely treed). We want this model to reliably predict locations that are completely un-treed, so we convert this dataset into "binary" values, such that dayMet grid cells with no trees (50% or fewer of the RAP cells in the dayMet grid cell have <3% tree cover)(value = 1) receive a value of 0, and dayMet grid cells with trees (value <0.5) receive a value of 0. 

```{r, warning=FALSE}
## reclassify the raster according to our purposes 
nonForest_rast <- 1 - nonForest_rast
##  make values binary
nonForest_rast[nonForest_rast>.5] <- 1
nonForest_rast[nonForest_rast<=.5] <- 0
## download wall-to-wall climate data, which we'll use to fit the model
## contemporary climate data (wall-to-wall, no anomalies)
climDat_Temp <- readRDS("../../../Data_processed/WallToWallClimateData/DayMetData_allCONUS_2023ClimateValues_withSoils.rds")

# rename
climDat <- climDat_Temp %>% 
  #filter(year == 2016) %>% 
  dplyr::select(tmin_meanAnnAvg_CLIM:durationFrostFreeDays_meanAnnAvg_CLIM,
                #tmean_meanAnnAvg_3yrAnom:durationFrostFreeDays_meanAnnAvg_3yrAnom,
               soilDepth:totalAvailableWaterHoldingCapacity) %>%
rename("tmin" = tmin_meanAnnAvg_CLIM, 
     "tmax" = tmax_meanAnnAvg_CLIM, #1 
     "tmean" = tmean_meanAnnAvg_CLIM, 
     "prcp" = prcp_meanAnnTotal_CLIM, 
     "t_warm" = T_warmestMonth_meanAnnAvg_CLIM,
     "t_cold" = T_coldestMonth_meanAnnAvg_CLIM, 
     "prcp_wet" = precip_wettestMonth_meanAnnAvg_CLIM,
     "prcp_dry" = precip_driestMonth_meanAnnAvg_CLIM, 
     "prcp_seasonality" = precip_Seasonality_meanAnnAvg_CLIM, #2
     "prcpTempCorr" = PrecipTempCorr_meanAnnAvg_CLIM,  #3
     "abvFreezingMonth" = aboveFreezing_month_meanAnnAvg_CLIM, 
     "isothermality" = isothermality_meanAnnAvg_CLIM, #4
     "annWatDef" = annWaterDeficit_meanAnnAvg_CLIM, 
     "annWetDegDays" = annWetDegDays_meanAnnAvg_CLIM,
     "VPD_mean" = annVPD_mean_meanAnnAvg_CLIM, 
     "VPD_max" = annVPD_max_meanAnnAvg_CLIM, #5
     "VPD_min" = annVPD_min_meanAnnAvg_CLIM, #6
     "VPD_max_95" = annVPD_max_95percentile_CLIM, 
     "annWatDef_95" = annWaterDeficit_95percentile_CLIM, 
     "annWetDegDays_5" = annWetDegDays_5percentile_CLIM, 
     "frostFreeDays_5" = durationFrostFreeDays_5percentile_CLIM, 
     "frostFreeDays" = durationFrostFreeDays_meanAnnAvg_CLIM, 
     "soilDepth" = soilDepth, #7
     "clay" = surfaceClay_perc, 
     "sand" = avgSandPerc_acrossDepth, #8
     "coarse" = avgCoarsePerc_acrossDepth, #9
     "carbon" = avgOrganicCarbonPerc_0_3cm, #10
     "AWHC" = totalAvailableWaterHoldingCapacity,
  )
#rm(climDat_Temp) 
#gc()

## now, scale the contemporary climate data for use in cover models 
# get the scaling factors 
scaleParams <- modDat_1_s %>% 
  sf::st_drop_geometry() %>% 
  #filter(Year == 2016) %>% 
  dplyr::select(tmin_s:AWHC_s) %>% 
  reframe(across(all_of(names(.)), attributes)) 

# apply the scaling factors to the contemporary climate data 
climDat <- climDat %>% 
  mutate("Lon" = st_coordinates(.)[,1],
         "Lat" = st_coordinates(.)[,2])

climDat_noSF <- climDat %>% 
  st_drop_geometry()

namesToScale <- climDat_noSF %>% 
  dplyr::select(tmin:frostFreeDays, soilDepth:carbon, AWHC) %>% 
  names()

climDat_scaled <- map(namesToScale, .f = function(x) {
  x_new <- (climDat_noSF[,x] - scaleParams[,paste0(x, "_s")]$`scaled:center`)/scaleParams[,paste0(x, "_s")]$`scaled:scale`
  return(data.frame(x_new))
}) %>% 
  purrr::list_cbind()
names(climDat_scaled) <- paste0(namesToScale, "_s")

climDatPred <- climDat %>% 
  dplyr::select(Lon, Lat, geometry) %>% 
  cbind(climDat_scaled)
names(climDatPred)[3:31] <- str_remove(names(climDatPred)[3:31], pattern = "_s$")

## use the climate data points to transform the no tree data into a data.frame
nonForest_points <- terra::extract(x = nonForest_rast, 
                                   y = terra::vect(climDatPred), bind = TRUE)
  #terra::as.points(nonForest_rast)
nonForest_points <- nonForest_points %>%
 sf::st_as_sf()

## remove the climate observations for locations that don't have no tree information
nonForest_points <- nonForest_points %>% 
  filter(!is.na(fracNotForest))
# ggplot(nonForest_points) + 
#   geom_point(aes(x = Lon, y = Lat, col = fracNotForest), size = .1)
```

```{r rename variables}  
set.seed(1234)
# now, rename columns for brevity
modDat_1 <- nonForest_points 
rm(nonForest_points)
# get necessary raster data for background data on maps
test_rastTemp <-  rast("/Users/astears/Documents/Dropbox_static/Work/NAU_USGS_postdoc/cleanPED/PED_vegClimModels/Data_raw/dayMet/rawMonthlyData/orders/70e0da02b9d2d6e8faa8c97d211f3546/Daymet_Monthly_V4R1/data/daymet_v4_prcp_monttl_na_1980.tif") %>%
  terra::crop(ext(-2000000, 2600000, -1900000, 1500000))

## want to use x_vegDat_sf and y_vegDat_sf, so rename to x and y
modDat_1 <- modDat_1 %>% 
  rename(x = Lon, 
         y = Lat)
```


#### Predictors
The following are the candidate predictor variables for this ecoregion: 
```{r get candidate predictor variables}
  prednames <- c(
      "tmean","prcp","prcpTempCorr","isothermality", "annWetDegDays", "clay","coarse", "AWHC"
 
 #"tmean", "prcp", "prcp_seasonality", "prcpTempCorr", "isothermality", "annWetDegDays", "sand", "coarse", "AWHC"
  )

print(prednames)
```

The variables in the training data are already scaled
```{r scaling predictors}

# Remove the rows that have no observations for total tree cover
modDat_1_s <- modDat_1 
# subset the data to only include these predictors, and remove any remaining NAs 
modDat_1_s <- modDat_1_s %>% 
  dplyr::select(all_of(prednames), fracNotForest, x, y) %>% 
  drop_na()

names(prednames) <- prednames
df_pred <- modDat_1_s[, prednames]
```

Need to add in information about ecoregion, which we use for internal cross validation within the LASSO
```{r}
# get the ecoregion data
regions <- sf::st_read(dsn = "../../../Data_raw/Level2Ecoregions/", layer = "NA_CEC_Eco_Level2") 
# transform it to match the dayMet projection/crs
regions_2 <- st_transform(regions, crs = st_crs(test_rastTemp)) %>% 
  st_make_valid()

# add to the notree data
modDat_1_s <- sf::st_join(modDat_1_s, regions_2)

# remove spatial information
modDat_1_s <- modDat_1_s %>% 
  st_drop_geometry()

## get rid of any nas
modDat_1_s <- modDat_1_s %>% 
  drop_na()
```


Because the dataset is so large (on the dayMet grid), subset to 300,000 observations to make model fitting viable
```{r}
## subset data set for testing purposes
# save full version
modDat_1_s_FULL <- modDat_1_s
# now subset
modDat_1_s <- modDat_1_s_FULL %>% 
  slice_sample(n = 500000)
```

#### Visualize the response variable 
```{r visualize response variable , fig.height=4, fig.width=5}
ggplot(modDat_1_s) + 
  geom_histogram(aes(.data[[response]]), fill = "forestgreen", col = "forestgreen", alpha = .5) + 
  xlab("P(Non-tree)") + 
  ggtitle("P(Non-tree) in the training data")


```

```{r summary_table}
create_summary <- function(df) {
  df %>% 
    pivot_longer(cols = everything(),
                 names_to = 'variable') %>% 
    group_by(variable) %>% 
    summarise(across(value, .fns = list(mean = ~mean(.x, na.rm = TRUE), min = ~min(.x, na.rm = TRUE), 
                                        median = ~median(.x, na.rm = TRUE), max = ~max(.x, na.rm = TRUE)))) %>% 
    mutate(across(where(is.numeric), round, 4))
}

modDat_1_s[prednames] %>% 
  create_summary() %>% 
  knitr::kable(caption = 'summaries of possible predictor variables') %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed")) 

```

#### Predictor variables compared to binned response variables

```{r climVar_boxplots, fig.height=9, fig.width=8, eval=FALSE}
set.seed(12011993)
# vector of name of response variables
vars_response <- response
# longformat dataframes for making boxplots
df_sample_plots <-  modDat_1_s  %>% 
  slice_sample(n = 5e4) %>% 
   rename(response = all_of(response)) %>% 
  mutate(response = case_when(
    response < 0.33 ~ "low", 
    response >= 0.33  & response < 0.66 ~ "med", 
    response >= 0.66   ~ "high", 
  )) %>% 
  dplyr::select(c(response, prednames)) %>% 
  tidyr::pivot_longer(cols = unname(prednames), 
               names_to = "predictor", 
               values_to = "value"
               )  
 

  ggplot(df_sample_plots, aes_string(x= "response", y = 'value')) +
  geom_boxplot() +
  facet_wrap(~predictor , scales = 'free_y') + 
  ylab("Predictor Variable Values") + 
    xlab(response)

```

# Model Fitting

## Visualize the spatial blocks and how they differ across environmental space
First, if there are observations in Louisiana, sub-sample them so they're not so over-represented in the dataset
```{r}
## make data into spatial format
modDat_1_sf <- modDat_1_s %>% 
  st_as_sf(coords = c("x", "y"), crs = st_crs("EPSG:4326"))


# get state boundaries for plotting
cropped_states <- us_states(resolution = "low")
cropped_states <- suppressMessages(cropped_states %>%
  dplyr::filter(name!="Hawaii") %>%
  dplyr::filter(name!="Alaska") %>%
  dplyr::filter(name!="Puerto Rico") %>%
  dplyr::filter(name!="American Samoa") %>%
  dplyr::filter(name!="Guam") %>%
  dplyr::filter(name!="Commonwealth of the Northern Mariana Islands") %>%
  dplyr::filter(name!="United States Virgin Islands")  %>%
  sf::st_transform(sf::st_crs(modDat_1_sf)))

```

```{r Visualize ecoregion distribution, fig.height=10, fig.width=14, message=FALSE, warning=FALSE}
## do a pca of climate across level 2 ecoregions
pca <- prcomp(modDat_1_s[,paste0(prednames)])
library(factoextra)
(fviz_pca_ind(pca, habillage = modDat_1_s$NA_L2NAME, label = "none", addEllipses = TRUE, ellipse.level = .95, ggtheme = theme_minimal(), alpha.ind = .1))

# make a table of n for each region
modDat_1_s %>% 
  group_by(NA_L2NAME) %>% 
  dplyr::summarize("Number_Of_Observations" = length(NA_L2NAME)) %>% 
  rename("Level_2_Ecoregion" = NA_L2NAME)%>% 
  kable() %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed")) 

```
## Then, look at the spatial distribution and environmental characteristics of the grouped ecoregions 
```{r fig.width = 14, fig.height=10, message=FALSE}

map1 <- ggplot() +
  geom_sf(data=cropped_states,fill='white') +
  geom_sf(data=modDat_1_sf#[modDat_1_sf$NA_L2NAME %in% c("MIXED WOOD PLAINS"),]
          ,
          aes(fill=as.factor(NA_L2NAME)),linewidth=0.5,alpha=0.5) +
  geom_point(data=modDat_1_s#[modDat_1_s$NA_L2NAME %in% c("MIXED WOOD PLAINS"),]
             ,
             alpha=0.5, 
             aes(x = x, y = y, color=as.factor(NA_L2NAME)), alpha = .1) +
  #scale_fill_okabeito() +
  #scale_color_okabeito() +
 # theme_default() +
  theme(legend.position = 'none') +
  labs(title = "Level 2 Ecoregions as spatial blocks")

hull <- modDat_1_sf %>%
  ungroup() %>%
  group_by(NA_L2NAME) %>%
  slice(chull(tmean, prcp))

plot1<-ggplot(data=modDat_1_sf,aes(x=tmean,y=prcp)) +
  geom_polygon(data = hull, alpha = 0.25,aes(fill=NA_L2NAME) )+
  geom_point(aes(group=NA_L2NAME,color=NA_L2NAME),alpha=0.25) +
  theme_minimal() + xlab("Annual Average T_mean - long-term average") +
  ylab("Annual Average Precip - long-term average") #+
  #scale_color_okabeito() +
  #scale_fill_okabeito()

plot2<-ggplot(data=modDat_1_sf %>%
                pivot_longer(cols=tmean:prcp),
              aes(x=value,group=name)) +
  # geom_polygon(data = hull, alpha = 0.25,aes(fill=fold) )+
  geom_density(aes(group=NA_L2NAME,fill=NA_L2NAME),alpha=0.25) +
  theme_minimal() +
  facet_wrap(~name,scales='free')
 
library(patchwork)
(combo <- (map1+plot1)/plot2) 
rm(modDat_1_sf)
gc()
```


## Fit a global model with all of the data 

### First, fit a LASSO regression model using the glmnet R package
- This regression is a beta glm with a logit link (using custom function from Daniel)
- Use cross validation across level 2 ecoregions to tune the lambda parameter in the LASSO model 
- this model is fit to using the scaled weather/climate/soils variables
- this list of possible predictors includes: 
  1. main effects
  2. interactions between all soils variables
  3. interactions between climate and weather variables
  4. transformed main effects (squared, log-transformed (add a uniform integer -- 20-- to all variables prior to log-transformation), square root-transformed (add a uniform integer -- 20-- to all variables prior to log-transformation))
```{r fit global model, echo = FALSE}
# penalty values
  lambdas <- 10^seq(1,-5, length.out = 200)  # sequence of penalties to test

 ## if the model parameters indicate that we DO want anomalies... 
    # make model matrix (which includes transformations and interactions)
  ## only include interactions within and between climate and anomaly variables, and within and between soils variables

 ## break predictors into climate, weather, and soils 
  prednames_clim <- prednames[which(prednames %in% c("tmin" , "tmax",  "tmean",
     "prcp", "t_warm" , "t_cold" , "prcp_wet" , "prcp_dry" , "prcp_seasonality", "prcpTempCorr", "abvFreezingMonth" , "isothermality" , "annWatDef" , "annWetDegDays",  "VPD_mean",  "VPD_max",  "VPD_min" , "VPD_max_95", "annWatDef_95", "annWetDegDays_5",  "frostFreeDays_5", "frostFreeDays"))]

  prednames_weath <- prednames[which(prednames %in% c("tmean_anom", "tmin_anom",  "tmax_anom", 
    "prcp_anom",      "t_warm_anom" , "t_cold_anom",  "prcp_wet_anom" , "precp_dry_anom" ,  "prcp_seasonality_anom" ,  "prcpTempCorr_anom" , "aboveFreezingMonth_anom",  "isothermality_anom" ,  "annWatDef_anom", "annWetDegDays_anom",  "VPD_mean_anom" , "VPD_min_anom" , "VPD_max_anom", "VPD_max_95_anom", "annWatDef_95_anom" , "annWetDegDays_5_anom",  "frostFreeDays_5_anom" , "frostFreeDays_anom"))]
  
  prednames_soils <- prednames[which(prednames %in% c( "soilDepth" , "clay", 
     "sand" , "coarse" ,  "carbon", "AWHC"))]
  
  
  # get a text string of possible interactions between climate and weather variables
  combos <- data.frame(gtools::permutations(n = length(c(prednames_clim, prednames_weath)),
                                            r = 2,
                                            v = c(prednames_clim, prednames_weath)))
  
  combos$interactions <- paste0(combos$X1, "*", combos$X2)
  
  
  # get a text string of possible interactions between soils variables
  combos_2 <- data.frame(gtools::permutations(n = length(c(prednames_soils)),
                                            r = 2,
                                            v = c(prednames_soils)))
  
  combos_2$interactions <- paste0(combos_2$X1, "*", combos_2$X2)
  
  # make text string of all possible interactions
  int_string <-str_flatten(c(combos$interactions, combos_2$interactions), collapse = " + ")
  
  # make text string of transformations (log, squared)
  transform_string <- str_flatten(c(# add square root term 
                                    paste0("I(log(",prednames_clim,"+30))"), 
                                    paste0("I(log(",prednames_soils,"+30))"),
                                    # 
                                    paste0("I(sqrt(",prednames_clim,"+30))"),
                                    paste0("I(sqrt(",prednames_soils,"+30))"),

                                    paste0("I(",prednames_clim,"^2)"),
                                    paste0("I(",prednames_soils,"^2)")),
                                    collapse = " + ")
              
  # get a text string of the possible untransformed model terms
  mod_string <- paste(response, "~", paste0(prednames, collapse = "+"))
  
  # make a text string of the model formula w/ interactions and transformations included
  modWithInteractions <- paste(mod_string, "+",int_string , "+", transform_string
                               )

  # now, make a model formula object
  f <- as.formula(modWithInteractions)
  
  # transform dataframe to model matrix
  X <- model.matrix(f, modDat_1_s)
  # get response variable
  y <- as.matrix(modDat_1_s[,response])

```

#### Get rid of transformed predictions and interactions that are correlated 
```{r}
# get first pass of names correlated variables
X_df <- X %>% 
  as.data.frame() %>% 
  dplyr::select(-'(Intercept)')  
corrNames_i <- X_df %>% 
  cor()  %>% 
   caret::findCorrelation(cutoff = .7, verbose = FALSE, names = TRUE, exact = TRUE)
# remove those names that are untransformed main effects 
  # vector of columns to remove 
badNames <- corrNames_i[!(corrNames_i %in% prednames)]
while (sum(!(corrNames_i %in% prednames))>0) {
 corrNames_i <-  X_df %>% 
    dplyr::select(-badNames) %>% 
     cor()  %>% 
   caret::findCorrelation(cutoff = .7, verbose = FALSE, names = TRUE, exact = TRUE)
 # update the vector of names to remove 
 badNames <- unique(c(badNames, corrNames_i[!(corrNames_i %in% prednames)]))
}

## see if there are any correlated variables left (would be all main effects at this point)
# if there are, step through and remove the variable that each is most correlated with 
if (length(corrNames_i)>1) {
  for (i in 1:length(corrNames_i)) {
    X_i <- X_df %>% 
      dplyr::select(-badNames)
    if (corrNames_i[i] %in% names(X_i)) {
    corMat_i <- cor(x = X_i[corrNames_i[i]], y = X_i %>% dplyr::select(-corrNames_i[i])) 
    badNames_i <- colnames(corMat_i)[abs(corMat_i)>=.7]
    # if there are any predictors in the 'badNames_i', remove them from this list
    if (length(badNames_i) > 0 & sum(c(badNames_i %in% prednames))>0) {
        badNames_i <- badNames_i[!(badNames_i %in% prednames)]
    }
    badNames <- unique(c(badNames, badNames_i))
    }
  }
}
## update the X matrix to exclude these correlated variables
X <- X[,!(colnames(X) %in% badNames)]

rm(X_df)
```

#### Run the global LASSO model
```{r}
if (run == TRUE) {
  # set up custom folds
    # get the ecoregions for training lambda
  train_eco <- modDat_1_s$NA_L2NAME#[train]
 
  # Fit model -----------------------------------------------
  # specify leave-one-year-out cross-validation
  my_folds <- as.numeric(as.factor(train_eco))

    # set up parallel processing
    # this computer has 16 cores (parallel::detectCores())
    registerDoMC(cores = 8)
 
    fit <- cv.glmnet(
      x = X[,2:ncol(X)], 
      y = y, 
      family = "binomial", #quasibeta_family(),
      keep = FALSE,
      alpha = 1,  # 0 == ridge regression, 1 == lasso, 0.5 ~~ elastic net
      lambda = lambdas,
      relax = ifelse(response == "ShrubCover", yes = TRUE, no = FALSE),
      #nlambda = 100,
      type.measure="mse",
      #penalty.factor = pen_facts,
      foldid = my_folds,
      #thresh = thresh,
      standardize = FALSE, ## scales variables prior to the model sequence... coefficients are always returned on the original scale
      parallel = TRUE
    )
    base::saveRDS(fit, paste0("../ModelFitting/models/yesOrNoTrees_globalLASSOmod_binomial_treeCutoff_",treeThreshold,".rds"))
    
  
     best_lambda <- fit$lambda.min
  # save the lambda for the most regularized model w/ an MSE that is still 1SE w/in the best lambda model
  lambda_1SE <- fit$lambda.1se
  # save the lambda for the most regularized model w/ an MSE that is still .5SE w/in the best lambda model
  lambda_halfSE <- best_lambda + ((lambda_1SE - best_lambda)/2)
 
## Now, we need to do stability selection to ensure the coefficients that are being chosen with each lambda are stable 

## stability selection for best lambda model 
# setup params
p <- ncol(X[,2:ncol(X)]) # of parameters
n <- length(y) # of observations
n_iter <- 100        # number of subsamples
sample_frac <- 0.75  # fraction of data to subsample
lambda_val <- best_lambda    # fixed lambda value (could be chosen via CV)

# Track selection
selection_counts_bestL <- matrix(0, nrow = p, ncol = 1)

for (i in 1:n_iter) {
  # Subsample rows
  sample_idx <- sample(1:n, size = floor(sample_frac * n), replace = FALSE)
  X_sub <- X[sample_idx, ]
  y_sub <- y[sample_idx]

  # Fit Lasso model
  fit_stab_i <- glmnet(x = X_sub[,2:ncol(X_sub)], y = y_sub, 
    family = "binomial",
    alpha = 1, lambda = lambda_val, standardize = FALSE)

  # Get non-zero coefficients (excluding intercept)
  select_bestL <- which(as.vector(coef(fit_stab_i))[-1] != 0)
  selection_counts_bestL[select_bestL] <- selection_counts_bestL[select_bestL] + 1
}

# Convert counts to selection probabilities (the probability that a variable is selected over 100 iterations)
selection_prob_bestL <- selection_counts_bestL / n_iter
selection_prob_bestL_df <- data.frame(
  VariableNumber = paste0("X", 1:p),
  VariableName = rownames(coef(fit_stab_i))[2:(p+1)],
  SelectionProb = as.vector(selection_prob_bestL)
)

# get those variables that are selected in ≥70% of subsets (Meinshausen and Bühlmann, 2010)
bestLambda_coef <- selection_prob_bestL_df[selection_prob_bestL_df$SelectionProb>=.7, c("VariableName", "SelectionProb")]

#//////
# stability selection for 1se lambda model
lambda_val <-  lambda_1SE    # fixed lambda value (could be chosen via CV)

# Track selection
selection_counts_1seL <- matrix(0, nrow = p, ncol = 1)

for (i in 1:n_iter) {
  # Subsample rows
  sample_idx <- sample(1:n, size = floor(sample_frac * n), replace = FALSE)
  X_sub <- X[sample_idx, ]
  y_sub <- y[sample_idx]

  # Fit Lasso model
  fit_stab_i <- glmnet(x = X_sub[,2:ncol(X_sub)], y = y_sub, 
    family = "binomial",
    alpha = 1, lambda = lambda_val, standardize = FALSE)

  # Get non-zero coefficients (excluding intercept)
  selected_1seL <- which(as.vector(coef(fit_stab_i))[-1] != 0)
  selection_counts_1seL[selected_1seL] <- selection_counts_1seL[selected_1seL] + 1
}

# Convert counts to selection probabilities (the probability that a variable is selected over 100 iterations)
selection_prob_1seL <- selection_counts_1seL / n_iter
selection_prob_1seL_df <- data.frame(
  VariableNumber = paste0("X", 1:p),
  VariableName = rownames(coef(fit_stab_i))[2:(p+1)],
  SelectionProb = as.vector(selection_prob_1seL)
)

# get those variables that are selected in ≥70% of subsets (Meinshausen and Bühlmann, 2010)
seLambda_coef <- selection_prob_1seL_df[selection_prob_1seL_df$SelectionProb>=.7, c("VariableName", "SelectionProb")]

# stability selection for half se lambda model
lambda_val <- lambda_halfSE    # fixed lambda value (could be chosen via CV)

# Track selection
selection_counts_halfseL <- matrix(0, nrow = p, ncol = 1)

for (i in 1:n_iter) {
  # Subsample rows
  sample_idx <- sample(1:n, size = floor(sample_frac * n), replace = FALSE)
  X_sub <- X[sample_idx, ]
  y_sub <- y[sample_idx]

  # Fit Lasso model
  fit_stab_i <- glmnet(x = X_sub[,2:ncol(X_sub)], y = y_sub, 
    family = "binomial",
    alpha = 1, lambda = lambda_val, standardize = FALSE)

  # Get non-zero coefficients (excluding intercept)
  selected_halfseL <- which(as.vector(coef(fit_stab_i))[-1] != 0)
  selection_counts_halfseL[selected_halfseL] <- selection_counts_halfseL[selected_halfseL] + 1
}

# Convert counts to selection probabilities (the probability that a variable is selected_halfseL over 100 iterations)
selection_prob_halfseL <- selection_counts_halfseL / n_iter
selection_prob_halfseL_df <- data.frame(
  VariableNumber = paste0("X", 1:p),
  VariableName = rownames(coef(fit_stab_i))[2:(p+1)],
  SelectionProb = as.vector(selection_prob_halfseL)
)

# get those variables that are selected_halfseL_halfseL in ≥70% of subsets (Meinshausen and Bühlmann, 2010)
halfseLambda_coef <- selection_prob_halfseL_df[selection_prob_halfseL_df$SelectionProb>=.7, c("VariableName", "SelectionProb")]

## fit w/ the identified coefficients from the 'best' lambda, but using the glm function
  mat_glmnet_best <- bestLambda_coef$VariableName 

  if (length(mat_glmnet_best) == 0) {
    f_glm_bestLambda <- as.formula(paste0(response, "~ 1"))
  } else {
  f_glm_bestLambda <- as.formula(paste0(response, " ~ ", paste0(mat_glmnet_best, collapse = " + ")))
  }
  
## fit
  fit_glm_bestLambda <- fit_glm_bestLambda_binomial <- glm(formula = f_glm_bestLambda, data = modDat_1_s, family = "binomial")
    
   ## fit w/ the identified coefficients from the '1se' lambda, but using the glm function
  mat_glmnet_1se <- seLambda_coef$VariableName
    
  if (length(mat_glmnet_1se) == 0) {
    f_glm_1se <- as.formula(paste0(response, "~ 1"))
  } else {
  f_glm_1se <- as.formula(paste0(response, " ~ ", paste0(mat_glmnet_1se, collapse = " + ")))
  }


  fit_glm_se <- glm(formula = f_glm_1se, data = modDat_1_s, family = "binomial")
  # glm(data = modDat_1_s, formula = f_glm_1se,
  #                   family =  stats::Gamma(link = "log"))
  
     ## fit w/ the identified coefficients from the '.5se' lambda, but using the glm function
  mat_glmnet_halfse <- halfseLambda_coef$VariableName
  
  if (length(mat_glmnet_halfse) == 0) {
    f_glm_halfse <- as.formula(paste0(response, "~ 1"))
  } else {
  f_glm_halfse <- as.formula(paste0(response, " ~ ", paste0(mat_glmnet_halfse, collapse = " + ")))
  }

  fit_glm_halfse <- glm(formula = f_glm_halfse, data = modDat_1_s, family = "binomial")
  
  ## save models 
  saveRDS(fit_glm_bestLambda, paste0("./models/yesOrNoTrees_bestLambdaGLM_",treeThreshold,".rds"))
  saveRDS(fit_glm_halfse, paste0("./models/yesOrNoTrees_halfSELambdaGLM_",treeThreshold,".rds"))
  saveRDS(fit_glm_se, paste0("./models/yesOrNoTrees_oneSELambdaGLM_",treeThreshold,".rds"))
    
  ## save the R environment after running the models 
  save(f_glm_halfse, mat_glmnet_halfse, halfseLambda_coef,
              f_glm_1se, mat_glmnet_1se, seLambda_coef,
              f_glm_bestLambda, mat_glmnet_best, bestLambda_coef,
              file = paste0("./models/interimModelFittingObjects_yesOrNoTrees_binomial_",treeThreshold,".rds"))
  } else {
    # read in LASSO object
    fit <- readRDS(paste0("../ModelFitting/models/yesOrNoTrees_globalLASSOmod_binomial_treeCutoff_",treeThreshold,".rds"))
    
    # read in R objects having to do w/ model fitting 
    load(file = paste0("./models/interimModelFittingObjects_yesOrNoTrees_binomial_",treeThreshold,".rds"))
      
  fit_glm_bestLambda <- readRDS(paste0("./models/yesOrNoTrees_bestLambdaGLM_",treeThreshold,".rds"))
  fit_glm_halfse <- readRDS(paste0("./models/yesOrNoTrees_halfSELambdaGLM_",treeThreshold,".rds"))
  fit_glm_se <- readRDS(paste0("./models/yesOrNoTrees_oneSELambdaGLM_",treeThreshold,".rds"))
  }


  # assess model fit
  # assess.glmnet(fit$fit.preval, #newx = X[,2:293], 
  #               newy = y, family = stats::Gamma(link = "log"))
  # save the minimum lambda
  best_lambda <- fit$lambda.min
  # save the lambda for the most regularized model w/ an MSE that is still 1SE w/in the best lambda model
  lambda_1SE <- fit$lambda.1se
  # save the lambda for the most regularized model w/ an MSE that is still .5SE w/in the best lambda model
  lambda_halfSE <- best_lambda + ((lambda_1SE - best_lambda)/2)
 
  print(fit)     
  
  plot(fit)

```

#### Then, we predict (on the training set) using both of these models (best lambda and 1se lambda) as an itial test
```{r}
  ## predict on the test data
  # lasso model predictions with the optimal lambda
  optimal_pred <- stats::predict(fit_glm_bestLambda, #newx=X[,2:ncol(X)], 
                                 type = "response")
  optimal_pred_1se <-  stats::predict(fit_glm_se, newx=X[,2:ncol(X)], type = "response")
  optimal_pred_halfse <- stats::predict(fit_glm_halfse, newx = X[,2:ncol(X)], type = "response")
  
    null_fit <- glm(
      formula = y ~ 1, #data = modDat_1_s, 
      family = binomial
      )
  null_pred <- predict(null_fit, newdata = as.data.frame(X), type = "response"
                       )

  # save data
  fullModOut <- list(
    "modelObject" = fit,
    "nullModelObject" = null_fit,
    "modelPredictions" = data.frame(#ecoRegion_holdout = rep(test_eco,length(y)),
      obs=y,
                    pred_opt=optimal_pred, 
                    pred_opt_se = optimal_pred_1se,
                    pred_opt_halfse = optimal_pred_halfse,
                    pred_null=null_pred#,
                    #pred_nopenalty=nopen_pred
                    ))

ggplot() + 
  geom_point(aes(X[,2], fullModOut$modelPredictions$obs), col = "black", alpha = .1) + 
  geom_point(aes(X[,2], fullModOut$modelPredictions$pred_opt), col = "red", alpha = .1) + ## predictions w/ the CV model (best lambda)
  geom_point(aes(X[,2], fullModOut$modelPredictions$pred_opt_halfse), col = "orange", alpha = .1) + ## predictions w/ the CV model (.5se lambda)
  geom_point(aes(X[,2], fullModOut$modelPredictions$pred_opt_se), col = "green", alpha = .1) + ## predictions w/ the CV model (1se lambda)
  geom_point(aes(X[,2], fullModOut$modelPredictions$pred_null), col = "blue", alpha = .1) + 
  labs(title = "A rough comparison of observed and model-predicted values", 
       subtitle = "black = observed values \n red = predictions from 'best lambda' model \n orange = predictions for '1/2se' lambda model \n green = predictions from '1se' lambda model \n blue = predictions from null model") +
  xlab(colnames(X)[2])
  #ylim(c(0,200))
 
```

#### The internal cross-validation process to fit the global LASSO model identified an optimal lambda value (regularization parameter) of `r{print(best_lambda)}`. The lambda value such that the cross validation error is within 1 standard error of the minimum ("1se lambda") was `r{print(fit$lambda.1se)}``. The following coefficients were kept in each model: 
```{r coefficients in the global model}
# the coefficient matrix from the 'best model' -- find and print those coefficients that aren't 0 in a table
coef_glm_bestLambda <- coef(fit_glm_bestLambda) %>% 
  data.frame() 
coef_glm_bestLambda$coefficientName <- rownames(coef_glm_bestLambda)
names(coef_glm_bestLambda)[1] <- "coefficientValue_bestLambda"
# coefficient matrix from the '1se' model 
coef_glm_1se <- coef(fit_glm_se) %>% 
  data.frame() 
coef_glm_1se$coefficientName <- rownames(coef_glm_1se)
names(coef_glm_1se)[1] <- "coefficientValue_1seLambda"
# coefficient matrix from the 'half se' model 
coef_glm_halfse <- coef(fit_glm_halfse) %>% 
  data.frame() 
coef_glm_halfse$coefficientName <- rownames(coef_glm_halfse)
names(coef_glm_halfse)[1] <- "coefficientValue_halfseLambda"
# add together
coefs <- full_join(coef_glm_bestLambda, coef_glm_halfse) %>% 
  full_join(coef_glm_1se) %>% 
  dplyr::select(coefficientName, coefficientValue_bestLambda,
                coefficientValue_halfseLambda, coefficientValue_1seLambda)

globModTerms <- coefs[!is.na(coefs$coefficientValue_bestLambda), "coefficientName"]

## also, get the number of unique variables in each model 
var_prop_pred <- paste0(response, "_pred")
response_vars <- c(response, var_prop_pred)
# for best lambda model
prednames_fig <- paste(str_split(globModTerms, ":", simplify = TRUE)) 
prednames_fig <- str_replace(prednames_fig, "I\\(", "")
prednames_fig <- str_replace(prednames_fig, "\\^2\\)", "")
prednames_fig <- unique(prednames_fig[prednames_fig>0])
prednames_fig <- prednames_fig
prednames_fig_num <- length(prednames_fig)
# for 1SE lambda model
globModTerms_1se <- coefs[!is.na(coefs$coefficientValue_1seLambda), "coefficientName"]
if (length(globModTerms_1se) == 1) {
prednames_fig_1se <- paste(str_split(globModTerms_1se, ":", simplify = TRUE)) 
prednames_fig_1se <- str_replace(prednames_fig_1se, "I\\(", "")
prednames_fig_1se <- str_replace(prednames_fig_1se, "\\^2\\)", "")
prednames_fig_1se <- unique(prednames_fig_1se[prednames_fig_1se>0])
prednames_fig_1se_num <- c(0)
} else {
prednames_fig_1se <- paste(str_split(globModTerms_1se, ":", simplify = TRUE)) 
prednames_fig_1se <- str_replace(prednames_fig_1se, "I\\(", "")
prednames_fig_1se <- str_replace(prednames_fig_1se, "\\^2\\)", "")
prednames_fig_1se <- unique(prednames_fig_1se[prednames_fig_1se>0])
prednames_fig_1se_num <- length(prednames_fig_1se)
}
# for 1/2SE lambda model
globModTerms_halfse <- coefs[!is.na(coefs$coefficientValue_halfseLambda), "coefficientName"]
if (length(globModTerms_halfse) == 1) {
prednames_fig_halfse <- paste(str_split(globModTerms_halfse, ":", simplify = TRUE)) 
prednames_fig_halfse <- str_replace(prednames_fig_halfse, "I\\(", "")
prednames_fig_halfse <- str_replace(prednames_fig_halfse, "\\^2\\)", "")
prednames_fig_halfse <- unique(prednames_fig_halfse[prednames_fig_halfse>0])
prednames_fig_halfse_num <- c(0)
} else {
prednames_fig_halfse <- paste(str_split(globModTerms_halfse, ":", simplify = TRUE)) 
prednames_fig_halfse <- str_replace(prednames_fig_halfse, "I\\(", "")
prednames_fig_halfse <- str_replace(prednames_fig_halfse, "\\^2\\)", "")
prednames_fig_halfse <- unique(prednames_fig_halfse[prednames_fig_halfse>0])
prednames_fig_halfse_num <- length(prednames_fig_halfse)
}
# make a table
kable(coefs, col.names = c("Coefficient Name", "Value from best lambda model", 
                           "Value from 1/2 se lambda", "Value from 1se lambda model")
      ) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed")) 

# calculate RMSE of all models 
RMSE_best <- yardstick::rmse(fullModOut$modelPredictions[,c("obs", "pred_opt")], truth = "obs", estimate = "pred_opt")$.estimate
RMSE_halfse <- yardstick::rmse(fullModOut$modelPredictions[,c("obs", "pred_opt_halfse")], truth = "obs", estimate = "pred_opt_halfse")$.estimate
RMSE_1se <- yardstick::rmse(fullModOut$modelPredictions[,c("obs", "pred_opt_se")], truth = "obs", estimate = "pred_opt_se")$.estimate
# calculate bias of all models
bias_best <-  mean((fullModOut$modelPredictions$obs) - fullModOut$modelPredictions$pred_opt)
bias_halfse <-  mean((fullModOut$modelPredictions$obs) - fullModOut$modelPredictions$pred_opt_halfse)
bias_1se <- mean((fullModOut$modelPredictions$obs) - fullModOut$modelPredictions$pred_opt_se)

uniqueCoeffs <- data.frame("Best lambda model" = c(signif(RMSE_best,3), as.character(signif(bias_best, 3)),
  as.integer(length(globModTerms)-1), as.integer(prednames_fig_num), 
                                                   as.integer(sum(prednames_fig %in% c(prednames_clim))),
                                                   as.integer(sum(prednames_fig %in% c(prednames_weath))),
                                                   as.integer(sum(prednames_fig %in% c(prednames_soils)))
                                                   ), 
                           "1/2 se lambda model" = c(signif(RMSE_halfse,3), as.character(signif(bias_halfse, 3)),
                             length(globModTerms_halfse)-1, prednames_fig_halfse_num,
                                                   sum(prednames_fig_halfse %in% c(prednames_clim)),
                                                   sum(prednames_fig_halfse %in% c(prednames_weath)),
                                                   sum(prednames_fig_halfse %in% c(prednames_soils))), 
                           "1se lambda model" = c(signif(RMSE_1se,3), as.character(signif(bias_1se, 3)),
                             length(globModTerms_1se)-1, prednames_fig_1se_num,
                                                   sum(prednames_fig_1se %in% c(prednames_clim)),
                                                   sum(prednames_fig_1se %in% c(prednames_weath)),
                                                   sum(prednames_fig_1se %in% c(prednames_soils))))
row.names(uniqueCoeffs) <- c("RMSE", "bias: mean(obs-pred.)", "Total number of coefficients", "Number of unique coefficients",
                             "Number of unique climate coefficients", 
                             "Number of unique weather coefficients",  
                             "Number of unique soils coefficients"
                             )

kable(uniqueCoeffs, 
      col.names = c("Best lambda model", "1/2 se lambda model", "1se lambda model"), row.names = TRUE) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed")) 
```

# Calibration and Visualizations of Model Predictions and Residuals 

In the following figures, we show model predictions made using the best lambda model, as well as an alternative "second-best" lambda model. As the alternative to the best lambda model, we use the model (1se or 1/2se of best Lambda) that has the fewest number of unique predictors, but is not a null model. 
```{r}
if (whichSecondBestMod == "auto") {
  # name of model w/ fewest # of predictors (but more than 0)
uniqueCoeff_min <- min(as.numeric(uniqueCoeffs[4,2:3])[which(as.numeric(uniqueCoeffs[4,2:3]) > 0)])
alternativeModel <- names(uniqueCoeffs[4,2:3])[which(uniqueCoeffs[4,2:3] == uniqueCoeff_min)]

if (is.finite(uniqueCoeff_min)) {
  if (length(alternativeModel) == 1) {
  if (alternativeModel == "X1.2.se.lambda.model") {
  mod_secondBest <- fit_glm_halfse
  name_secondBestMod <- "1/2 SE Model"
  prednames_secondBestMod <- prednames_fig_halfse
  # get the name of the column that contains the prediction data for only the training data, for this model
trainDatPredName <- "pred_opt_halfse"
} else if (alternativeModel == "X1se.lambda.model") {
  mod_secondBest <- fit_glm_se
  name_secondBestMod <- "1 SE Model"
  prednames_secondBestMod <- prednames_fig_1se
  # get the name of the column that contains the prediction data for only the training data, for this model
trainDatPredName <- "pred_opt_se"
}
} else {
  # if both alternative models have the same number of unique coefficients, chose the model that has the fewest number of total predictors
  uniqueCoeff_min2 <- min(as.numeric(uniqueCoeffs[3,alternativeModel]))
alternativeModel2 <- names(uniqueCoeffs[3,alternativeModel])[which(uniqueCoeffs[3,alternativeModel] == uniqueCoeff_min2)]
if (length(alternativeModel2) == 1) {
  if (alternativeModel2 == "X1.2.se.lambda.model") {
  mod_secondBest <- fit_glm_halfse
  name_secondBestMod <- "1/2 SE Model"
  prednames_secondBestMod <- prednames_fig_halfse
   # get the name of the column that contains the prediction data for only the training data, for this model
trainDatPredName <- "pred_opt_halfse"
} else if (alternativeModel2 == "X1se.lambda.model") {
  mod_secondBest <- fit_glm_se
  name_secondBestMod <- "1 SE Model"
  prednames_secondBestMod <- prednames_fig_1se
  # get the name of the column that contains the prediction data for only the training data, for this model
trainDatPredName <- "pred_opt_se"
}
} else {
  mod_secondBest <- fit_glm_halfse
  name_secondBestMod <- "1/2 SE Model"
  prednames_secondBestMod <- prednames_fig_halfse
  # get the name of the column that contains the prediction data for only the training data, for this model
trainDatPredName <- "pred_opt_halfse"
}

}
  }else {
    mod_secondBest <- NULL
  name_secondBestMod <- "Intercept_Only"
  prednames_secondBestMod <- NULL
}
} else if (whichSecondBestMod == "1se") {
  mod_secondBest <- fit_glm_se
  name_secondBestMod <- "1 SE Model"
  prednames_secondBestMod <- prednames_fig_1se
# get the name of the column that contains the prediction data for only the training data, for this model
trainDatPredName <- "pred_opt_se"
} else if (whichSecondBestMod == "halfse") {
  mod_secondBest <- fit_glm_halfse
  name_secondBestMod <- "1/2 SE Model"
  prednames_secondBestMod <- prednames_fig_halfse
  # get the name of the column that contains the prediction data for only the training data, for this model
trainDatPredName <- "pred_opt_halfse"
}
```


```{r}
## Get wall-to-wall climate data calculated from dayMet, which we will use for calibration (get version w/ predictors that are already scaled)
# read in unscaled climate data
climDatPred <- readRDS("../../../Data_processed/WallToWallClimateData/DayMetData_allCONUS_2023_SCALEDcliamteAndSoilsPredictors.rds")
```

<!-- Make sure that the locations we're using for the model-fitting are the same as the locations used to determine the proportion of CONUS that is un-forested (i.e. masks out the same locations due to fire and/or land use) -->
```{r}
# rasterize the model-fitting data on the dayMet grid

## compare to the the RAP forest data from Martin (is the 'notForest' object)
# reproject notForest to match data I'm using
notForest2 <- notForest %>%
  terra::project(crs(climDatPred))

# notForestTrim <- notForest2 %>%
#   terra::crop(terra::ext(climDatPred[1:1000,]))
# make sure crs matches
#crs(notForest2) == crs(climDatPred)

climDatPred_vect <- climDatPred %>%
  terra::vect()

# crop climDatPred to match the extent of the notForest data
climDatPred_extract <- notForest2 %>%
  terra::extract(y = climDatPred_vect) %>%
  cbind(climDatPred) %>%
  st_as_sf()

## remove climate data from the locations that have been masked in Martin's tree data
climDatPred_new <- climDatPred_extract %>%
  filter(!is.na(fracNotForest))
```

## Predict on the complete dataset, and then the masked wall-to-wall data

```{r predict with best lambda on data used for fitting}
  # create prediction for each each model
# (i.e. for each fire proporation variable)
predict_by_response <- function(mod, df, scale = FALSE) {
  df_out <- df
    if(scale == TRUE) {
  df_out <- df_out %>% 
  mutate(across(all_of(prednames), base::scale,scale = FALSE, center = FALSE)) 
  }
  response_name <- paste0(response, "_pred")
  preds <- predict(mod, newdata = df_out, #s="lambda.min", 
                                     type = "response")
  #preds[preds<0] <- 0
  #preds[preds>100] <- 100
  df_out <- df_out %>% cbind(preds)
   colnames(df_out)[ncol(df_out)] <- response_name
  return(df_out)
}
## predict on entire dataset
pred_glm1_full <- predict_by_response(fit_glm_bestLambda, modDat_1_s_FULL)#X[,2:ncol(X)])
pred_glm1_full$resid <- pred_glm1_full[,response] - pred_glm1_full[,paste0(response, "_pred")]
pred_glm1_full$extremeResid <- NA
pred_glm1_full[pred_glm1_full$resid > .5 | pred_glm1_full$resid < -.5,"extremeResid"] <- 1

## predict on the wall-to-wall dataset
pred_glm1_w2w <- predict_by_response(fit_glm_bestLambda, climDatPred %>% st_drop_geometry(), scale = FALSE)

```

```{r predict with second best lambda model on data used for model fitting}
if ( name_secondBestMod == "Intercept_Only") {
  print("The next best lambda model only contains one predictor (an intercept)")
} else {

## for full dataset
pred_glm1_1se_full <- predict_by_response(mod_secondBest, modDat_1_s_FULL)#X[,2:ncol(X)])
pred_glm1_1se_full$resid <- pred_glm1_1se_full[,response] - pred_glm1_1se_full[,paste0(response, "_pred")]
pred_glm1_1se_full$extremeResid <- NA
pred_glm1_1se_full[pred_glm1_1se_full$resid > .5 | pred_glm1_1se_full$resid < -.5,"extremeResid"] <- 1

## predict on the wall-to-wall dataset
pred_glm1_1se_w2w <- predict_by_response(mod_secondBest, climDatPred %>% st_drop_geometry(), scale = FALSE)
}
```



## Making predictions binary
Identifying the optimal threshold to make the continuous predictions binomial based on the prevalence in the training data. 

For the bestLambda model: 
```{r threshold for best lambda model, warning=FALSE}
library(PresenceAbsence)
DATA <- fullModOut$modelPredictions[,c("obs", "pred_opt")] # only training data (full dataset)
DATA$ID <- 1:nrow(DATA)
## make the 'observations' binary (i.e. locations > 50% are 'treed' and locations < 50% are 'un-treed')
# in model-prediction data
DATA$obs_binary <- DATA$obs
DATA[DATA$obs_binary <= .5, "obs_binary"] <- 0
DATA[DATA$obs_binary > .5, "obs_binary"] <- 1
# in full training data
modDat_1_s_FULL$fracNotForest_binary <- modDat_1_s_FULL$fracNotForest
modDat_1_s_FULL[modDat_1_s_FULL$fracNotForest_binary <= .5, "fracNotForest_binary"] <- 0
modDat_1_s_FULL[modDat_1_s_FULL$fracNotForest_binary > .5, "fracNotForest_binary"] <- 1

## calculate the observed prevalence from the FULL dataset (not the training subset) (the average of the observed )
obs.prev_MINE <- mean(modDat_1_s_FULL$fracNotForest)

## now, identify the optimal threshold
optimalThresholds <- optimal.thresholds(DATA = DATA[,c("ID", "obs_binary", "pred_opt")],
threshold = 200, obs.prev = obs.prev_MINE#mean(DATA[,"obs"])
) # tests 100 thresholds)

## find the cutoff for making the predictions binary such that the prevalence of treed locations is the same in the model predictions as in the prediction area 
rm(DATA)
best_threshold_bestLambda <- optimalThresholds[optimalThresholds$Method == thresholdMethod, 2]
cat(paste0("Optimal threshold for best lambda model based on ROC curve, based on the ",thresholdMethod, " method: ", best_threshold_bestLambda, "\n"))

## save this prevalence for later 
saveRDS(best_threshold_bestLambda, "./models/YesNoTreeModelPrevalence.rds")

```

For the second-bestLambda model: 
```{r threshold for next best lambda model}
# best_threshold_secondBest <- coords$threshold
DATA <- fullModOut$modelPredictions[,c("obs", trainDatPredName)]
DATA$ID <- 1:nrow(DATA)

## make the 'observations' binary (i.e. locations > 50% are 'treed' and locations < 50% are 'un-treed')
# in model-prediction data
DATA$obs_binary <- DATA$obs
DATA[DATA$obs_binary <= .5, "obs_binary"] <- 0
DATA[DATA$obs_binary > .5, "obs_binary"] <- 1

optimalThresholds <- optimal.thresholds(DATA = DATA[,c("ID", "obs_binary", trainDatPredName)], 
                                        threshold = 200, obs.prev = obs.prev_MINE) # tests 100 thresholds)

rm(DATA)
best_threshold_secondBest <- optimalThresholds[optimalThresholds$Method == thresholdMethod, 2]
cat(paste0("Optimal threshold for second-best lambda model based on ROC curve, based on the ",thresholdMethod, " method: ", best_threshold_secondBest, "\n"))


## save this prevalence for later 
saveRDS(best_threshold_secondBest, "./models/YesNoTreeModelPrevalence_secondBest.rds")

```

Below are figures of the empirical cumulative distribution functions for model-predicted probability of no trees from the best and second-best lambda models *before* we make the values binary. The vertical dashed line shows the threshold identified above to make the predicted values binary. 

For best lambda model
```{r}
## make an empirical cumulative probability function plot of the model predictions on wall-to-wall data
pred_glm1_w2w_cdf <- stats::ecdf(pred_glm1_w2w[,"fracNotForest_pred"])
# ## find the quantile of observations that corresponds to the desired proportion of unforested locations
thresholdPercentile_bestLambda <- pred_glm1_w2w_cdf(best_threshold_bestLambda)
  #quantile(pred_glm1_w2w$TotalTreeCover_binom_pred %>% na.omit(), probs = best_threshold_bestLambda) 
# get P(forest) for percentiles +/- 0.05 from the percentile of the optimal threshold

# plot the ecdf
plot(pred_glm1_w2w_cdf, xlab = c("Proportion of RAP cells in a dayMet \n cell that have <3% tree cover- continuous"), ylab = "Percentile",
     main = "Empirical Cumulative Distribution Function for \nUn-transformed model predictions of \nproportion of treed cells from the best lambda model")

# lines for optimal threshold
abline(v = best_threshold_bestLambda, lty = 2)
abline(h = thresholdPercentile_bestLambda, lty = 2)
```

For second-best lambda model
```{r}
## make an empirical cumulative probability function plot of the model predictions on wall-to-wall data
pred_glm1_1se_w2w_cdf <- stats::ecdf(pred_glm1_1se_w2w[,"fracNotForest_pred"])
# ## find the quantile of observations that corresponds to the desired proportion of unforested locations
thresholdPercentile_secondBestLambda <- pred_glm1_1se_w2w_cdf(best_threshold_secondBest)
  #quantile(pred_glm1_w2w$TotalTreeCover_binom_pred %>% na.omit(), probs = best_threshold_bestLambda) 
# get P(forest) for percentiles +/- 0.05 from the percentile of the optimal threshold

# plot the ecdf
plot(pred_glm1_1se_w2w_cdf, xlab = c("Proportion of RAP cells in a dayMet \n cell that have <3% tree cover- continuous"), ylab = "Percentile",
     main = "Empirical Cumulative Distribution Function for \nUn-transformed model predictions of \nproportion of treed cells from the second-best lambda model")

# lines for optimal threshold
abline(v = best_threshold_secondBest, lty = 2)
abline(h = thresholdPercentile_secondBestLambda, lty = 2)
```

We use the thresholds identified above to make continuous predictions of proportion of trees binary
```{r}
## for full dataset
# best lambda model
# "binomialize" the continuous predictions
pred_glm1_full <- pred_glm1_full %>% 
  mutate(fracNotForest_pred_binary = fracNotForest_pred,
    fracNotForest_pred_binary = replace(fracNotForest_pred_binary, fracNotForest_pred_binary > best_threshold_bestLambda, 1),
         fracNotForest_pred_binary = replace(fracNotForest_pred_binary, fracNotForest_pred_binary <= best_threshold_bestLambda, 0)) 

# second-best lambda model
# "binomialize" the continuouse predictions
pred_glm1_1se_full <- pred_glm1_1se_full %>% 
  mutate(fracNotForest_pred_binary = fracNotForest_pred,
    fracNotForest_pred_binary = replace(fracNotForest_pred_binary, fracNotForest_pred_binary > best_threshold_bestLambda, 1),
         fracNotForest_pred_binary = replace(fracNotForest_pred_binary, fracNotForest_pred_binary <= best_threshold_bestLambda, 0)) 

## for wall-to-wall data
# best lambda model
# "binomialize" the continuous predictions
pred_glm1_w2w <- pred_glm1_w2w %>% 
  mutate(fracNotForest_pred_binary = fracNotForest_pred,
    fracNotForest_pred_binary = replace(fracNotForest_pred_binary, fracNotForest_pred_binary > best_threshold_bestLambda, 1),
         fracNotForest_pred_binary = replace(fracNotForest_pred_binary, fracNotForest_pred_binary <= best_threshold_bestLambda, 0)) 

# second-best lambda model
# "binomialize" the continuouse predictions
pred_glm1_1se_w2w <- pred_glm1_1se_w2w %>% 
 mutate(fracNotForest_pred_binary = fracNotForest_pred,
    fracNotForest_pred_binary = replace(fracNotForest_pred_binary, fracNotForest_pred_binary > best_threshold_bestLambda, 1),
         fracNotForest_pred_binary = replace(fracNotForest_pred_binary, fracNotForest_pred_binary <= best_threshold_bestLambda, 0)) 
```


### Identify Misclassifications
```{r}

## first, need to bin prediction data 
# best lambda model
pred_glm1_full$fracNotForest_binary <- pred_glm1_full$fracNotForest
pred_glm1_full[pred_glm1_full$fracNotForest_binary <= .5, "fracNotForest_binary"] <- 0
pred_glm1_full[pred_glm1_full$fracNotForest_binary > .5, "fracNotForest_binary"] <- 1
# second-best lambda model
pred_glm1_1se_full$fracNotForest_binary <- pred_glm1_1se_full$fracNotForest
pred_glm1_1se_full[pred_glm1_1se_full$fracNotForest_binary <= .5, "fracNotForest_binary"] <- 0
pred_glm1_1se_full[pred_glm1_1se_full$fracNotForest_binary > .5, "fracNotForest_binary"] <- 1

# true positive = tree (1)
# false positive = model says tree, data says no-tree (2) 
# true negative = no tree (3) 
# false negative = model says no tree, data says tree (4)

# best-lambda model -- full dataset
pred_glm1_full$missClass <- NA
pred_glm1_full[pred_glm1_full$fracNotForest_binary == 1 & 
                 pred_glm1_full$fracNotForest_pred_binary == 1, "missClass"] <- 1
pred_glm1_full[pred_glm1_full$fracNotForest_binary == 0 & 
                 pred_glm1_full$fracNotForest_pred_binary == 0, "missClass"] <- 3
pred_glm1_full[pred_glm1_full$fracNotForest_binary == 0 & 
                 pred_glm1_full$fracNotForest_pred_binary == 1, "missClass"] <- 2
pred_glm1_full[pred_glm1_full$fracNotForest_binary == 1 & 
                 pred_glm1_full$fracNotForest_pred_binary == 0, "missClass"] <- 4

# second-best lambda model - full dataset
pred_glm1_1se_full$missClass <- NA
pred_glm1_1se_full[pred_glm1_1se_full$fracNotForest_binary == 1 & 
                 pred_glm1_1se_full$fracNotForest_pred_binary == 1, "missClass"] <- 1
pred_glm1_1se_full[pred_glm1_1se_full$fracNotForest_binary == 0 & 
                 pred_glm1_1se_full$fracNotForest_pred_binary == 0, "missClass"] <- 3
pred_glm1_1se_full[pred_glm1_1se_full$fracNotForest_binary == 0 & 
                 pred_glm1_1se_full$fracNotForest_pred_binary == 1, "missClass"] <- 2
pred_glm1_1se_full[pred_glm1_1se_full$fracNotForest_binary == 1 & 
                 pred_glm1_1se_full$fracNotForest_pred_binary == 0, "missClass"] <- 4

```

```{r}
# read in .tif w/ predictions of forest classification from the forest model, which we'll use later oneon dayMet grid
forestPreds_dmGrid <- terra::rast(x= "../../../Data_processed/CoverData/finalCoverData/predictionsOfForestProbOnDaymetScale.tif")

```


## Maps of Observations, Predictions, and Residuals
### Observations - continuous and binary total tree cover for the full dataset 
```{r Residual maps, fig.height = 6.5, fig.width = 9, warning = FALSE}

## clean up the environment
rm(modDat_1_s, modDat_1, nonForest_rast, notForest, notForest2, test_rast_dayMet_TEMP, X_i, X_sub, climDat, climDat_noSF, climDat_scaled, climDat_Temp, climDatPred, climDatPred_extract, climDatPred_new)
gc()

## get information for plotting
# aggregate to a grid of 8km x 8km cells
test_rast <- test_rastTemp #%>%
 #terra::aggregate(fact = 8, fun = "mean")#%>% 

# function for getting the statistical mode
Mode <- function(x, ...) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

# string w/ crs of veg data
crs_veg <- "PROJCRS[\"unnamed\",\n    BASEGEOGCRS[\"unknown\",\n        DATUM[\"unknown\",\n            ELLIPSOID[\"Spheroid\",6378137,298.257223563,\n                LENGTHUNIT[\"metre\",1,\n                    ID[\"EPSG\",9001]]]],\n        PRIMEM[\"Greenwich\",0,\n            ANGLEUNIT[\"degree\",0.0174532925199433,\n                ID[\"EPSG\",9122]]]],\n    CONVERSION[\"Lambert Conic Conformal (2SP)\",\n        METHOD[\"Lambert Conic Conformal (2SP)\",\n            ID[\"EPSG\",9802]],\n        PARAMETER[\"Latitude of false origin\",42.5,\n            ANGLEUNIT[\"degree\",0.0174532925199433],\n            ID[\"EPSG\",8821]],\n        PARAMETER[\"Longitude of false origin\",-100,\n            ANGLEUNIT[\"degree\",0.0174532925199433],\n            ID[\"EPSG\",8822]],\n        PARAMETER[\"Latitude of 1st standard parallel\",25,\n            ANGLEUNIT[\"degree\",0.0174532925199433],\n            ID[\"EPSG\",8823]],\n        PARAMETER[\"Latitude of 2nd standard parallel\",60,\n            ANGLEUNIT[\"degree\",0.0174532925199433],\n            ID[\"EPSG\",8824]],\n        PARAMETER[\"Easting at false origin\",0,\n            LENGTHUNIT[\"metre\",1],\n            ID[\"EPSG\",8826]],\n        PARAMETER[\"Northing at false origin\",0,\n            LENGTHUNIT[\"metre\",1],\n            ID[\"EPSG\",8827]]],\n    CS[Cartesian,2],\n        AXIS[\"easting\",east,\n            ORDER[1],\n            LENGTHUNIT[\"metre\",1,\n                ID[\"EPSG\",9001]]],\n        AXIS[\"northing\",north,\n            ORDER[2],\n            LENGTHUNIT[\"metre\",1,\n                ID[\"EPSG\",9001]]]]"
# make a multi-layer raster w/ layers for each variable of interest
pred_glm1_full_sf <- pred_glm1_full %>% 
  st_as_sf(coords = c("x", "y"), crs = crs(crs_veg)) %>% 
  st_transform(crs(test_rast)) %>% 
  terra::vect()
testRast2 <- lapply(c("fracNotForest_binary", "fracNotForest_pred", "fracNotForest_pred_binary", "resid", "missClass"), FUN = function(x) {
  # S4 method for class 'data.frame'
  if (x %in% c("fracNotForest_binary", "fracNotForest_pred", "fracNotForest_pred_binary", "resid")) {
    temp <- terra::rasterize(x = pred_glm1_full_sf, y = test_rast, 
                         field = x, fun = mean, na.rm = TRUE)
    names(temp) <- x
  } else {
    temp <- terra::rasterize(x = pred_glm1_full_sf, y = test_rast, 
                         field = x, fun = Mode, na.rm = TRUE)
    # get the % aggreement between values in the dayMetx8 cell 
    temp2 <- terra::rasterize(x = pred_glm1_full_sf, y = test_rast$daymet_v4_prcp_monttl_na_1980_1, 
                         field = x, fun = function(y) {
                           yTemp <- y[!is.na(y)]
                           yMode <- Mode(yTemp)
                           (sum(yTemp == yMode)/length(yTemp))*100
                         })
    temp <- c(as.factor(temp$missClass), temp2)
    names(temp) <- c("missClass", "PercAgreement")
  }
    
 return(temp)
  }
)

pred_glm1_full_RAST <- c(testRast2[[1]], testRast2[[2]], testRast2[[3]], testRast2[[4]], testRast2[[5]])

# reproject state boundaries
cropped_states_2 <- cropped_states  %>%
  sf::st_transform(sf::st_crs(test_rast))

# make figure of treed/un-treed - binary
map_obs_bin <- ggplot() +
#geom_spatraster(data = pred_glm1_full_RAST, aes(fill = fracNotForest), maxcell = 50000000) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
  geom_raster(data = pred_glm1_full_RAST, aes(x = x, 
                   y = y, 
                   fill = as.character(fracNotForest_binary)), maxcell = 5e+10) + 
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) +
  coord_sf(datum = st_crs(test_rastTemp)) +
             scale_fill_viridis_c(na.value = "white", "magma", direction = -1, limits = c(0,212)) +
 #geom_sf(data = pred_glm1_full_RAST, aes(col = TotalTreeCover)) + 
  # stat_summary_2d(data = plotObs, aes(x = x, y = y, z = TotalTreeCover), fun = mean, binwidth = .05) + 
labs(title = paste0("Observations of tree presence - binary"), x = "", y = ""
     ) +
   scale_fill_manual(name = "", 
                     values = c("wheat", "darkgreen"), 
                     labels = c("No trees", "Trees"),
                    na.translate = FALSE) +
  # scale_fill_gradient2(low = "brown",
  #                      mid = "wheat" ,
  #                      high = "forestgreen" , 
  #                      midpoint = 0,   na.value = "white") +
  theme(axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) 

# # make figure of treed/un-treed - continuous
# map_obs_cont <- ggplot() +
# geom_spatraster(data = pred_glm1_full_RAST, aes(fill = fracNotForest), maxcell = 50000000) + 
#   geom_sf(data=cropped_states_2 ,fill=NA ) +
#   #geom_raster(data = pred_glm1_full_RAST, aes(x = x, 
#    #                y = y, 
#     #               fill = as.character(fracNotForest)), maxcell = 5e+10) + 
#  xlim(-2150000, 2510000) +
#  ylim(-2005000, 1000000) +
#   coord_sf(datum = st_crs(test_rastTemp)) +
#              scale_fill_viridis_c(na.value = "white", "magma", direction = -1, limits = c(0,212)) +
#  #geom_sf(data = pred_glm1_full_RAST, aes(col = TotalTreeCover)) + 
#   # stat_summary_2d(data = plotObs, aes(x = x, y = y, z = TotalTreeCover), fun = mean, binwidth = .05) + 
# labs(title = paste0("Observations of tree presence - continuous"), x = "", y = ""
#      ) +
#    # scale_fill_manual(name = "", 
#    #                   values = c("wheat", "darkgreen"), 
#    #                   labels = c("No trees", "Trees"),
#    #                  na.translate = FALSE) +
#   scale_fill_gradient2(low = "brown",
#                        mid = "wheat" ,
#                        high = "forestgreen" ,
#                        midpoint = 0,   na.value = "white") +
#   theme(axis.text = element_blank(),
#                    strip.text = element_text(size = 15),
#                    title = element_text(size = 18),
#                    axis.ticks = element_blank()) 
# 
# #add inset
#  inset_plot <-  ggplot(pred_glm1_full) +
#    geom_density(aes(fracNotForest)) +
#       theme_bw(base_size=9) +  ## makes everything smaller
#       theme(panel.background = element_rect(fill="transparent"),  ## white plot background
#             panel.border = element_blank(),
# axis.title.y = element_blank(),
# axis.title.x = element_blank(),
# axis.text.x = element_text(size=rel(2)), ## tiny axis text
# axis.text.y = element_text(size=rel(2)), ## tiny axis text
# axis.line.x = element_line(size = .5),
# axis.line.y = element_line(size = .5),
# panel.grid.major = element_blank(),
# panel.grid.minor = element_blank(),
# plot.background = element_blank(),
# strip.background = element_blank(),
# strip.text = element_blank())
#      
#  map_obs_cont <-    map_obs_cont +   annotation_custom(grob = ggplotGrob(inset_plot), ymin = -2028000, ymax=-1205000, xmin=-2280000, xmax=-180000 ) 

(map_obs_bin)
 # (map_obs_cont)
```

### Predictions on the full dataset using the *best lambda* model
```{r predicted values map best lambda model, fig.height = 6.5, fig.width = 9, warning=FALSE, message=FALSE}

# make figure of binomialized tree/no-tree
map_preds1 <- ggplot() +
   geom_raster(data = pred_glm1_full_RAST, aes(x = x, 
                   y = y, 
                   fill = as.character(fracNotForest_pred_binary)), maxcell = 5e+10) + 
#geom_spatraster(data = pred_glm1_full_RAST, aes(fill = fracNotForest_pred_binary), maxcell = 20000000) + 
    geom_sf(data=cropped_states_2 ,fill=NA ) +
labs(title = paste0("Predictions from the bestaLambda model of tree presence \nbinary probabilities"), x = "", y = "") +
   scale_fill_manual(name = "", 
                     values = c("wheat", "darkgreen"), 
                     labels = c("No trees", "Trees"),
                    na.translate = FALSE) + 
  coord_sf(datum = st_crs(test_rastTemp)) +
    theme(axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) 

#  make figure of continuous tree/no-tree
map_preds1_cont <- ggplot() +
geom_spatraster(data = pred_glm1_full_RAST, aes(fill = fracNotForest_pred), maxcell = 50000000) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
  #geom_raster(data = pred_glm1_full_RAST, aes(x = x, 
   #                y = y, 
    #               fill = as.character(fracNotForest)), maxcell = 5e+10) + 
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) +
  coord_sf(datum = st_crs(test_rastTemp)) +
             scale_fill_viridis_c(na.value = "white", "magma", direction = -1, limits = c(0,212)) +
 #geom_sf(data = pred_glm1_full_RAST, aes(col = TotalTreeCover)) + 
  # stat_summary_2d(data = plotObs, aes(x = x, y = y, z = TotalTreeCover), fun = mean, binwidth = .05) + 
labs(title = paste0("Predictions from the bestaLambda model of tree presence \ncontinous predictions"), x = "", y = ""
     ) +
   # scale_fill_manual(name = "", 
   #                   values = c("wheat", "darkgreen"), 
   #                   labels = c("No trees", "Trees"),
   #                  na.translate = FALSE) +
  scale_fill_gradient2(low = "brown",
                       mid = "wheat" ,
                       high = "forestgreen" ,
                       midpoint = 0,   na.value = "white") +
  theme(axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) 

#add inset
 inset_plot <-  ggplot(pred_glm1_full) +
   geom_density(aes(fracNotForest_pred)) +
      theme_bw(base_size=9) +  ## makes everything smaller
      theme(panel.background = element_rect(fill="transparent"),  ## white plot background
            panel.border = element_blank(),
axis.title.y = element_blank(),
axis.title.x = element_blank(),
axis.text.x = element_text(size=rel(2)), ## tiny axis text
axis.text.y = element_text(size=rel(2)), ## tiny axis text
axis.line.x = element_line(size = .5),
axis.line.y = element_line(size = .5),
panel.grid.major = element_blank(),
panel.grid.minor = element_blank(),
plot.background = element_blank(),
strip.background = element_blank(),
strip.text = element_blank())
     
 map_preds1_cont <-    map_preds1_cont +   annotation_custom(grob = ggplotGrob(inset_plot), ymin = -2028000, ymax=-1205000, xmin=-2280000, xmax=-180000 ) 
 

(map_preds1)
(map_preds1_cont)

```

### Predictions on the full dataset using the *second best lambda* model
```{r predicted values map 1se lambda model, fig.height = 6.5, fig.width = 9, warning=FALSE, message=FALSE}
if ( name_secondBestMod == "Intercept_Only") {
  print("The next best lambda model only contains one predictor (an intercept)")
} else {
  pred_glm1_1se_full_sf <- pred_glm1_1se_full %>% 
  st_as_sf(coords = c("x", "y"), crs = crs(crs_veg)) %>% 
  st_transform(crs(test_rast)) %>% 
  terra::vect()
  # make a raster of the full dataset
# make a multi-layer raster w/ layers for each variable of interest
testRast3 <- lapply(c( "fracNotForest_binary", "fracNotForest_pred", "fracNotForest_pred_binary", "resid", "missClass"), FUN = function(x) {
  # S4 method for class 'data.frame'
 if (x %in% c("fracNotForest_binary", "fracNotForest_pred", "fracNotForest_pred_binary", "resid")) {
    temp <- terra::rasterize(x = pred_glm1_1se_full_sf, y = test_rast, 
                         field = x, fun = mean, na.rm = TRUE)
    names(temp) <- x
  } else {
    temp <- terra::rasterize(x = pred_glm1_1se_full_sf, y = test_rast, 
                         field = x, fun = Mode, na.rm = TRUE)
    # get the % aggreement between values in the dayMetx8 cell 
    temp2 <- terra::rasterize(x = pred_glm1_1se_full_sf, y = test_rast$daymet_v4_prcp_monttl_na_1980_1, 
                         field = x, fun = function(y) {
                           yTemp <- y[!is.na(y)]
                           yMode <- Mode(yTemp)
                           (sum(yTemp == yMode)/length(yTemp))*100
                         })
    temp <- c(as.factor(temp$missClass), temp2)
    names(temp) <- c("missClass", "PercAgreement")
  }
    
 return(temp)
  }
)
pred_glm1_1se_full_RAST <- c(testRast3[[1]], testRast3[[2]], testRast3[[3]], testRast3[[4]], testRast3[[5]])

# make figures of predictions w/ the full dataset
map_preds2_binary <- ggplot() +
   geom_raster(data = pred_glm1_1se_full_RAST, aes(x = x, 
                   y = y, 
                   fill = as.character(fracNotForest_pred_binary)), maxcell = 5e+10) + 
#geom_spatraster(data = pred_glm1_full_RAST, aes(fill = fracNotForest_pred_binary), maxcell = 20000000) + 
    geom_sf(data=cropped_states_2 ,fill=NA ) +
labs(title = paste0("Predictions from the ", name_secondBestMod, " of tree presence \nbinary probabilities"), x = "", y = "") +
  scale_fill_manual(name = "", 
                     values = c("wheat", "darkgreen"), 
                     labels = c("No trees", "Trees"),
                    na.translate = FALSE) + 
  coord_sf(datum = st_crs(test_rastTemp)) +
    theme(axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) 


#  make figure of continuous tree/no-tree
map_preds2_cont <- ggplot() +
geom_spatraster(data = pred_glm1_1se_full_RAST, aes(fill = fracNotForest_pred), maxcell = 50000000) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
  #geom_raster(data = pred_glm1_full_RAST, aes(x = x, 
   #                y = y, 
    #               fill = as.character(fracNotForest)), maxcell = 5e+10) + 
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) +
  coord_sf(datum = st_crs(test_rastTemp)) +
             scale_fill_viridis_c(na.value = "white", "magma", direction = -1, limits = c(0,212)) +
 #geom_sf(data = pred_glm1_full_RAST, aes(col = TotalTreeCover)) + 
  # stat_summary_2d(data = plotObs, aes(x = x, y = y, z = TotalTreeCover), fun = mean, binwidth = .05) + 
labs(title = paste0("Predictions from the ", name_secondBestMod, " of tree presence \ncontinuous probabilities"), x = "", y = ""
     ) +
   # scale_fill_manual(name = "", 
   #                   values = c("wheat", "darkgreen"), 
   #                   labels = c("No trees", "Trees"),
   #                  na.translate = FALSE) +
  scale_fill_gradient2(low = "brown",
                       mid = "wheat" ,
                       high = "forestgreen" ,
                       midpoint = 0,   na.value = "white") +
  theme(axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) 

#add inset
 inset_plot <-  ggplot(pred_glm1_1se_full) +
   geom_density(aes(fracNotForest_pred)) +
      theme_bw(base_size=9) +  ## makes everything smaller
      theme(panel.background = element_rect(fill="transparent"),  ## white plot background
            panel.border = element_blank(),
axis.title.y = element_blank(),
axis.title.x = element_blank(),
axis.text.x = element_text(size=rel(2)), ## tiny axis text
axis.text.y = element_text(size=rel(2)), ## tiny axis text
axis.line.x = element_line(size = .5),
axis.line.y = element_line(size = .5),
panel.grid.major = element_blank(),
panel.grid.minor = element_blank(),
plot.background = element_blank(),
strip.background = element_blank(),
strip.text = element_blank())
     
 map_preds2_cont <-    map_preds2_cont +   annotation_custom(grob = ggplotGrob(inset_plot), ymin = -2028000, ymax=-1205000, xmin=-2280000, xmax=-180000 ) 
 

(map_preds2_binary)
(map_preds2_cont)
 
}
```

### A map of classification error of predictions on the full dataset made using the *best lambda model*
```{r classification error map for best lambda model, fig.height = 6.5, fig.width = 9, warning = FALSE }
# true positive = tree (1) (forest green)
# true negative = no tree (3) (wheat) 
# false positive = model says tree, data says no-tree (2) (red)
# false negative = model says no tree, data says tree (4) (blue)
# make figures
map_missClass <- ggplot() +
geom_spatraster(data =pred_glm1_full_RAST
                , aes(fill= missClass), maxcell = 2000000000000) + 
  geom_sf(data=cropped_states_2,fill=NA )  + 
labs(title = paste0("Misclassification and correct classifications \nfrom the bestLambda model of Tree Presence")#,
     #subtitle = "bestLambda model \nvalues show are the most common classification in each 8km x 8km grid cell"
     ) +
    scale_fill_discrete(palette = 
                        c("forestgreen", "red", "wheat", "blue"),
                        na.value = "white",
                        labels = c("trees", "Data = no trees; Model = trees", "no trees", "Data = trees; Model = no trees", "")) +
  coord_sf(datum = st_crs(test_rastTemp)) +
    theme(axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) 

map_missClass
```

```{r fig.width = 7, fig.height=7}
# make a confusion matrix 
# prepare data
matData <- pred_glm1_full %>% 
  mutate(predClass = fracNotForest_pred_binary, 
         obsClass = fracNotForest_binary) %>% 
  mutate(predClass = replace(predClass, predClass == 0, "no trees"),
         predClass = replace(predClass, predClass == 1, "trees"),
         obsClass = replace(obsClass, obsClass == 0, "no trees"),
         obsClass = replace(obsClass, obsClass == 1, "trees")) %>% 
mutate(predClass = as.factor(predClass) ,
       obsClass = as.factor(obsClass))

# make matrix as a data.frame
# confMat <-  confusionMatrix(data = matData$predClass,
#                  reference = matData$obsClass)
ConfusionTableR::binary_visualiseR(
  train_labels = as.factor(matData$predClass),
  truth_labels = as.factor(matData$obsClass),
   class_label1 = "No trees",
   class_label2 = "Trees",
   quadrant_col1 = "wheat",
   quadrant_col2 = "forestgreen",
  text_col = "white", custom_title = "Confusion Matrix: classification of full dataset \nwith the best lambda model  using the full dataset"
)
```

### A map of classification error of predictions on the full datasets made using the *second best lambda model*
```{r classification error map for second best lambda model, fig.height = 6.5, fig.width = 9, warning = FALSE }
### full dataset
# make figures
map_missClass <- ggplot() +
geom_spatraster(data =pred_glm1_1se_full_RAST, aes(fill= missClass), maxcell = 200000000000) + 
  geom_sf(data=cropped_states_2,fill=NA )  + 
labs(title = paste0("Misclassification and correct classifications \nfrom the second-best lambda model of Yes/No Trees \nfull dataset"),
     subtitle = "second-best lambda model \nvalues shown are the most common classification in each 1km x 1km grid cell") + 
  scale_fill_discrete(palette = 
                        c("forestgreen", "red", "wheat", "blue"),
                        na.value = "white",
                       labels = c("trees", "Data = no trees; Model = trees", "no trees", "Data = trees; Model = no trees", ""))  +
  coord_sf(datum = st_crs(test_rastTemp)) +
    theme(axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) 
  
map_missClass
```

```{r fig.width = 7, fig.height=6.5}  
# make a confusion matrix 
# prepare data
matData <- pred_glm1_1se_full %>% 
  mutate(predClass = fracNotForest_pred_binary, 
         obsClass = fracNotForest_binary) %>% 
  mutate(predClass = replace(predClass, predClass == 0, "no trees"),
         predClass = replace(predClass, predClass == 1, "trees"),
         obsClass = replace(obsClass, obsClass == 0, "no trees"),
         obsClass = replace(obsClass, obsClass == 1, "trees")) %>% 
mutate(predClass = as.factor(predClass) ,
       obsClass = as.factor(obsClass))

# make matrix as a data.frame
# confMat <-  confusionMatrix(data = matData$predClass,
#                  reference = matData$obsClass)
ConfusionTableR::binary_visualiseR(
  train_labels = as.factor(matData$predClass),
  truth_labels = as.factor(matData$obsClass),
   class_label1 = "No trees",
   class_label2 = "Trees",
   quadrant_col1 = "wheat",
   quadrant_col2 = "forestgreen",
  text_col = "white", custom_title = "Confusion Matrix: classification of full dataset \nwith the second-best lambda model using the full dataset"
)

```

## Map of classification using best lambda model, *only in locations that the forest model predicts to be un-forested*
```{r classification error map for best lambda model Only in Unforested Locations, fig.height = 6.5, fig.width = 9, warning = FALSE }
## mask data based on forest classification
missClassRast_bestLambda_forestOnly <- pred_glm1_full_RAST$missClass %>% 
  terra::crop(ext(forestPreds_dmGrid)) %>% 
  terra::mask(mask = forestPreds_dmGrid, maskvalues = 1)
# make figures
map_missClass_nonForestOnly <- ggplot() +
geom_spatraster(data =missClassRast_bestLambda_forestOnly, aes(fill= missClass), maxcell = 2000000000) + 
  geom_sf(data=cropped_states_2,fill=NA )  + 
labs(title = paste0("Misclassification and correct classifications \nfrom the best lambda model of Yes/No Trees \nOnly in model-predicted unforested locations"),
     subtitle = "best lambda model") + 
  scale_fill_discrete(palette = 
                        c("forestgreen", "red", "wheat", "blue"),
                        na.value = "white",
                       labels = c("trees", "Data = no trees; Model = trees", "no trees", "Data = trees; Model = no trees", ""))  +
  coord_sf(datum = st_crs(test_rastTemp)) +
    theme(axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) 
  
map_missClass_nonForestOnly
```

```{r fig.width = 7, fig.height=6.5}  
# make a confusion matrix 
# prepare data
matData <- terra::values(missClassRast_bestLambda_forestOnly$missClass) %>% 
  as.data.frame() %>% 
  drop_na()
matData$observation <- NA
matData$prediction <- NA
matData[matData$missClass == 1, c("observation", "prediction")] <- "trees"

matData[matData$missClass == 2, c("observation")] <- "no trees"
matData[matData$missClass == 2, c("prediction")] <- "trees"

matData[matData$missClass == 3, c("observation", "prediction")] <- "no trees"

matData[matData$missClass == 4, c("observation")] <- "trees"
matData[matData$missClass == 4, c("prediction")] <- "no trees"

# make matrix as a data.frame
# confMat <-  confusionMatrix(data = matData$predClass,
#                  reference = matData$obsClass)
ConfusionTableR::binary_visualiseR(
  train_labels = as.factor(matData$prediction),
  truth_labels = as.factor(matData$observation),
   class_label1 = "No trees",
   class_label2 = "Trees",
   quadrant_col1 = "wheat",
   quadrant_col2 = "forestgreen",
  text_col = "white", custom_title = "Confusion Matrix: classification of full dataset \nwith the best lambda model--only in NonForest"
)

```

## Map of classification using second-best lambda model, *only in locations that the forest model predicts to be un-forested*
```{r classification error map for second-best lambda model Only in Unforested Locations, fig.height = 6.5, fig.width = 9, warning = FALSE }
## mask data based on forest classification
missClassRast_1SE_Lambda_forestOnly <- pred_glm1_1se_full_RAST$missClass %>% 
  terra::crop(ext(forestPreds_dmGrid)) %>% 
  terra::mask(mask = forestPreds_dmGrid, maskvalues = 1)
# make figures
map_missClass_nonForestOnly_secondBestLambda <- ggplot() +
geom_spatraster(data =missClassRast_1SE_Lambda_forestOnly, aes(fill= missClass), maxcell = 20000000) + 
  geom_sf(data=cropped_states_2,fill=NA )  + 
labs(title = paste0("Misclassification and correct classifications \nfrom the second-best lambda model of Yes/No Trees \nOnly in model-predicted unforested locations"),
     subtitle = "second-best lambda model") + 
  scale_fill_discrete(palette = 
                        c("forestgreen", "red", "wheat", "blue"),
                        na.value = "white",
                       labels = c("trees", "Data = no trees; Model = trees", "no trees", "Data = trees; Model = no trees", ""))  +
  coord_sf(datum = st_crs(test_rastTemp)) +
    theme(axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) 
  
map_missClass_nonForestOnly_secondBestLambda
```

```{r fig.width = 7, fig.height=6.5}  
# make a confusion matrix 
# prepare data
matData <- terra::values(missClassRast_1SE_Lambda_forestOnly$missClass) %>% 
  as.data.frame() %>% 
  drop_na()
matData$observation <- NA
matData$prediction <- NA
matData[matData$missClass == 1, c("observation", "prediction")] <- "trees"

matData[matData$missClass == 2, c("observation")] <- "no trees"
matData[matData$missClass == 2, c("prediction")] <- "trees"

matData[matData$missClass == 3, c("observation", "prediction")] <- "no trees"

matData[matData$missClass == 4, c("observation")] <- "trees"
matData[matData$missClass == 4, c("prediction")] <- "no trees"

# make matrix as a data.frame
# confMat <-  confusionMatrix(data = matData$predClass,
#                  reference = matData$obsClass)
ConfusionTableR::binary_visualiseR(
  train_labels = as.factor(matData$prediction),
  truth_labels = as.factor(matData$observation),
   class_label1 = "No trees",
   class_label2 = "Trees",
   quadrant_col1 = "wheat",
   quadrant_col2 = "forestgreen",
  text_col = "white", custom_title = "Confusion Matrix: classification of full dataset \nwith the second-best lambda model--only in NonForest"
)

```

## Maps of wall-to-wall model predictions, *including* all locations that are developed, agriculture, or water, or have been burned
```{r , fig.height = 6.5, fig.width = 9, warning=FALSE}
### best lambda model
## rasterize
 pred_glm1_w2w_sf <- pred_glm1_w2w %>%
  st_as_sf(coords = c("Lon", "Lat"), crs = crs(crs_veg)) %>%
  st_transform(crs(test_rast)) %>%
  terra::vect()
  # make a raster of the full dataset
# make a multi-layer raster w/ layers for each variable of interest
testRast6 <- lapply(c("fracNotForest_pred_binary"), FUN = function(x) {
  # S4 method for class 'data.frame'
    temp <- terra::rasterize(x = pred_glm1_w2w_sf, y = test_rast,
                         field = x, fun = mean, na.rm = TRUE)
    names(temp) <- x
 return(temp)
  }
)
pred_glm1_w2w_RAST <- c(testRast6[[1]])

## map of binary predictions
pred_glm1_w2w_bin <- ggplot() +
geom_raster(data =  pred_glm1_w2w_RAST #%>% terra::crop(ext(-90000, 10000, -1000, 90000))#pred_glm1_w2w_RAST
                , aes(x =x, y = y,fill = as.character(fracNotForest_pred_binary)), maxcell = 20000000000) +
  geom_sf(data=cropped_states_2 ,fill=NA ) +
labs(title = paste0("Predictions from the best lambda model of Tree Presence \nbinary predictions on the wall-to-wall dayMet data"), x = "", y = "") + 
  scale_fill_manual(name = "",
                     values = c("wheat", "darkgreen"),
                     labels = c("No trees", "Trees"),
                    na.translate = FALSE) +
  coord_sf(datum = st_crs(test_rastTemp)) +
    theme(axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000)

   
### second best lambda model
## rasterize
 pred_glm1_1se_w2w_sf <- pred_glm1_1se_w2w %>%
  st_as_sf(coords = c("Lon", "Lat"), crs = crs(crs_veg)) %>%
  st_transform(crs(test_rast)) %>%
  terra::vect()
  # make a raster of the full dataset
# make a multi-layer raster w/ layers for each variable of interest
testRast7 <- lapply(c( "fracNotForest_pred_binary"), FUN = function(x) {
  # S4 method for class 'data.frame'
    temp <- terra::rasterize(x = pred_glm1_1se_w2w_sf, y = test_rast,
                         field = x, fun = mean, na.rm = TRUE)
    names(temp) <- x
 return(temp)
  }
)
pred_glm1_1se_w2w_RAST <- c(testRast7[[1]])

## map of binary predictions
pred_glm1_1se_w2w_bin <- ggplot() +
geom_raster(data = pred_glm1_1se_w2w_RAST, aes(x = x, y = y, fill = as.character(fracNotForest_pred_binary)), maxcell = 20000000) +
  geom_sf(data=cropped_states_2 ,fill=NA ) +
labs(title = paste0("Predictions from the second-best lambda model of Tree Presence \nbinary probabilities on the wall-to-wall dayMet data"), x = "", y = "") +
  scale_fill_manual(name = "",
                     values = c("wheat", "darkgreen"),
                     labels = c("No trees", "Trees"),
                    na.translate = FALSE) +
  coord_sf(datum = st_crs(test_rastTemp)) +
    theme(axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000)

pred_glm1_w2w_bin
pred_glm1_1se_w2w_bin
```

<!-- ### Are there biases of the model predictions across year/lat/long? (for the full dataset) -->
```{r Misclassification figures best lambda model, fig.align=8, fig.height=10, eval = FALSE}
if ( name_secondBestMod == "Intercept_Only") {
  print("The next best lambda model only contains one predictor (an intercept)")
  
  ## calculate Misclassification in the input pred_glm1 or pred_glm1se datasets
  pred_glm1_full$missClass2 <- pred_glm1_full$fracNotForest - pred_glm1_full$fracNotForest_pred_binary
  
  # plot misclassifications against Year
# yearResidMod_bestLambda <- ggplot(pred_glm1_full) + 
#   geom_point(aes(x = jitter(Year), y = jitter(missClass2)), alpha = .1) + 
#   geom_hline(aes(yintercept = 0), lty = 2, col = "red") + 
#   geom_smooth(aes(x = Year, y = missClass2), method = "lm") + 
#   xlab("Year") + 
#   ylab("misclassifications") +
#   ggtitle("from best lamba model")

# plot misclassifications against Lat
latResidMod_bestLambda <- ggplot(pred_glm1_full) + 
  geom_point(aes(x = y, y = jitter(missClass2)), alpha = .1) + 
  geom_hline(aes(yintercept = 0), lty = 2, col = "red") + 
  geom_smooth(aes(x = y, y = missClass2)) + 
  xlab("Latitude") + 
  ylab("misclassifications") +
  ggtitle("from best lamba model")

# plot misclassifications against Long
longResidMod_bestLambda <- ggplot(pred_glm1_full) + 
  geom_point(aes(x = x, y = jitter(missClass2)), alpha = .1) + 
  geom_hline(aes(yintercept = 0), lty = 2, col = "red") + 
  geom_smooth(aes(x = x, y = missClass2)) + 
  xlab("Longitude") + 
  ylab("misclassifications") +
  ggtitle("from best lamba model")
library(patchwork)
#(yearResidMod_bestLambda ) / 
(  latResidMod_bestLambda ) /
(  longResidMod_bestLambda )
} else {
  
  ## calculate missClassification in the input pred_glm1 or pred_glm1se datasets
  pred_glm1_full$missClass2 <- pred_glm1_full$fracNotForest - pred_glm1_full$fracNotForest_pred_binary
  pred_glm1_1se_full$missClass2 <- pred_glm1_1se_full$fracNotForest - pred_glm1_1se_full$fracNotForest_pred_binary
  
# plot misclassifications against Year
# yearResidMod_bestLambda <- ggplot(pred_glm1_full) + 
#   geom_point(aes(x = jitter(Year), y = jitter(missClass2)), alpha = .1) + 
#   geom_hline(aes(yintercept = 0), lty = 2, col = "red") + 
#   geom_smooth(aes(x = Year, y = missClass2), method = "lm") + 
#   xlab("Year") + 
#   ylab("misclassifications") +
#   ggtitle("from best lamba model")
# yearResidMod_1seLambda <- ggplot(pred_glm1_1se_full) + 
#   geom_point(aes(x = jitter(Year), y = jitter(missClass2)), alpha = .1) + 
#   geom_hline(aes(yintercept = 0), lty = 2, col = "red") + 
#   geom_smooth(aes(x = Year, y = missClass2), method = "lm") + 
#   xlab("Year") + 
#   ylab("misclassifications") +
#   ggtitle(paste0("from ", name_secondBestMod))

# plot misclassifications against Lat
latResidMod_bestLambda <- ggplot(pred_glm1_full) + 
  geom_point(aes(x = y, y = jitter(missClass2)), alpha = .1) + 
  geom_hline(aes(yintercept = 0), lty = 2, col = "red") + 
  geom_smooth(aes(x = y, y = missClass2)) + 
  xlab("Latitude") + 
  ylab("misclassifications") +
  ggtitle("from best lamba model")
latResidMod_1seLambda <- ggplot(pred_glm1_1se_full) + 
  geom_point(aes(x = y, y = jitter(missClass2)), alpha = .1) + 
  geom_hline(aes(yintercept = 0), lty = 2, col = "red") + 
  geom_smooth(aes(x = y, y = missClass2)) + 
  xlab("Latitude") + 
  ylab("misclassifications") +
  ggtitle(paste0("from ", name_secondBestMod))

# plot misclassifications against Long
longResidMod_bestLambda <- ggplot(pred_glm1_full) + 
  geom_point(aes(x = x, y = jitter(missClass2)), alpha = .1) + 
  geom_hline(aes(yintercept = 0), lty = 2, col = "red") + 
  geom_smooth(aes(x = x, y = missClass2)) + 
  xlab("Longitude") + 
  ylab("misclassifications") +
  ggtitle("from best lamba model")
longResidMod_1seLambda <- ggplot(pred_glm1_1se_full %>% sample) + 
  geom_point(aes(x = x, y = jitter(missClass2)), alpha = .1) + 
  geom_hline(aes(yintercept = 0), lty = 2, col = "red") + 
  geom_smooth(aes(x = x, y = missClass2)) + 
  xlab("Longitude") + 
  ylab("misclassifications") +
  ggtitle(paste0("from ", name_secondBestMod))

library(patchwork)
#(yearResidMod_bestLambda + yearResidMod_1seLambda) / 
(  latResidMod_bestLambda + latResidMod_1seLambda) /
(  longResidMod_bestLambda + longResidMod_1seLambda)
}
```


### Quantile plots - comparing "binary" training data to "binary" model-predictions (i.e. converting the continuous model predictions to discrete values (1 or 0) based on the previously-identified optimal threshold)

```{r}
response_vars <- c("fracNotForest_binary", "fracNotForest_pred_binary")
# get deciles for best lambda model 
if (length(prednames_fig) == 0) {
  print("The best lambda model only contains one predictor (an intercept), so decile plots aren't possible to generate")
} else {
  pred_glm1_deciles <- predvars2deciles(pred_glm1_full,
                                      response_vars = response_vars,
                                        pred_vars = prednames_fig, 
                                       cut_points = seq(0, 1, 0.01))
}
# get deciles for 1 SE lambda model 
if ( name_secondBestMod == "Intercept_Only") {
  print("The next best lambda model only contains one predictor (an intercept)")} else {
  pred_glm1_deciles_1se <- predvars2deciles(pred_glm1_1se_full,
                                      response_vars = response_vars,
                                        pred_vars = prednames_secondBestMod, 
                                       cut_points = seq(0, 1, 0.01))
  }
```

Below are quantile plots for the *best lambda model* 
(note that the predictor variables are scaled) 
```{r quantile_plot best lambda binary response, fig.height= 16, fig.width = 12, warning=FALSE, message=FALSE}
if (length(prednames_fig) == 0) {
  print("The model only contains one predictor (an intercept), so decile plots aren't possible to generate")
} else {

  #####
# publication quality version
g3 <- decile_dotplot_pq(df = pred_glm1_deciles, response = c("fracNotForest_binary", paste0("fracNotForest", "_pred_binary")), IQR = FALSE,
                        CI = FALSE
                        ) + ggtitle("Quantile Plot") + 
  ggplot2::labs(y = "Treed (1) or Un-treed (0) - binary")

g4 <- add_dotplot_inset(g3, df = pred_glm1_deciles, response = c("fracNotForest_binary", paste0("fracNotForest", "_pred_binary")), dfRaw = pred_glm1_full, add_smooth = TRUE, deciles = FALSE)

  
if(save_figs) {
  # png(paste0("../../../Figures/CoverDatFigures/ figures/quantile_plots/quantile_plot_", response,  "_",ecoregion,".png"), 
  #    units = "in", res = 600, width = 5.5, height = 3.5 )
  #   print(g4)
  # dev.off()
}

# make a density plot 
 g4 <- ggarrange(g4, ggplot(pred_glm1_full) + 
  geom_density(aes(.data[["fracNotForest_binary"]])) + 
  geom_density(aes(.data[["fracNotForest_pred_binary"]]), col = "lightblue") + 
    ggtitle("Density of observed values (black) and predicted values (blue)"),
  ncol = 1, heights = c(.75, .25)
 )
 
 g4
 
}
```

Below are quantile plots from the second best lambda model (not that the predictor variables are scaled)
```{r quantile_plot 1 SE lambda binary response, fig.height= 16, fig.width = 12, warning=FALSE, message=FALSE}

if ( name_secondBestMod == "Intercept_Only") {
  print("The next best lambda model only contains one predictor (an intercept)")
  } else {

# publication quality version
g3 <- decile_dotplot_pq(pred_glm1_deciles_1se, response = c("fracNotForest_binary", paste0("fracNotForest", "_pred_binary")), IQR = FALSE) + ggtitle("Quantile Plot") + 
  ggplot2::labs(y = "Treed (1) or Un-treed (0) - binary")

g4 <- add_dotplot_inset(g3, df = pred_glm1_deciles_1se, response = c("fracNotForest_binary", paste0("fracNotForest", "_pred_binary")), dfRaw = pred_glm1_1se_full, add_smooth = TRUE, deciles = FALSE)

  
if(save_figs) {
  # png(paste0("figures/quantile_plots/quantile_plot_", response,  "_",ecoregion,".png"), 
  #    units = "in", res = 600, width = 5.5, height = 3.5 )
  #   print(g4)
  # dev.off()
}

 g4 <- ggarrange(g4, ggplot(pred_glm1_1se_full) + 
  geom_density(aes(.data[["fracNotForest_binary"]])) + 
  geom_density(aes(.data[["fracNotForest_pred_binary"]]), col = "lightblue") + 
    ggtitle("Density of observed values (black) and predicted values (blue)"),
  ncol = 1, heights = c(.75, .25)
 )
 
 g4

}

```

### Show model RMSE w/in each quantile (RMSE calculated by comparing binary predictions to binary observations)
```{r}
# get deciles for best lambda model 
if (length(prednames_fig) == 0) {
  print("The best lambda model only contains one predictor (an intercept), so decile plots aren't possible to generate")
} else {
  pred_glm1_deciles %>% 
    ggplot(aes(x = mean_value, y = RMSE)) +
    facet_wrap(~name, scales = "free_x") +
    geom_point(alpha = .2, size = .5) + 
    geom_smooth(lwd = .5) + 
    xlab("Scaled predictor value") + 
    ggtitle("RMSE by decile for bestLambda model")
}
# get deciles for 1 SE lambda model 
if (length(prednames_secondBestMod) == 0) {
  print("The 1SE (or 1/2 SE) lambda model only contains one predictor (an intercept), so decile plots aren't possible to generate")
} else {
  pred_glm1_deciles_1se %>% 
    ggplot(aes(x = mean_value, y = RMSE)) +
    facet_wrap(~name, scales = "free_x") +
    geom_point(alpha = .2, size = .5) + 
    geom_smooth(lwd = .5) + 
    xlab("Scaled predictor value") + 
    ggtitle(paste0("RMSE by decile for ", name_secondBestMod, "model"))
}
```


### Partial dependence plots (continuous model predictions)
For best lambda 
```{r, fig.width = 10, fig.height = 8}
# get names of coefficients in models
predNames_pdp_bestLambda <- names(coefficients(fit_glm_bestLambda)[-1])

## make a 'model.matrix' for the full dataset
X_pdp_bestLambda <- model.matrix(fit_glm_bestLambda$formula, modDat_1_s_FULL) %>% 
  as.data.frame()
X_pdp_bestLambda <- X_pdp_bestLambda %>% 
  left_join(modDat_1_s_FULL[prednames_fig])

for (i in 1:length(predNames_pdp_bestLambda)) {
  pd_bestLambda_i <- pdp::partial(object = fit_glm_bestLambda, pred.var = predNames_pdp_bestLambda[i], train = X_pdp_bestLambda, type = "regression", )
  names(pd_bestLambda_i)[1] <- "predictorVariable_value"
  pd_bestLambda_i$predictorVariable_Name <- predNames_pdp_bestLambda[i]
  if(i == 1) {
    pd_bestLambda <- pd_bestLambda_i
  } else {
    pd_bestLambda <- rbind(pd_bestLambda, pd_bestLambda_i)
  }
}
 
## update predictor names to be more interpretable
lookup_md <- c(
    "AWHC" = "Available Soil \nWater Holding Capacity",
    "VPD_mean_anom" = "VPD mean \n- 3 yr. anomaly", 
    "VPD_max" = "VPD max - 30yr. avg.",
    "VPD_max_95_anom" = "VPD max 95th percentile \n- 30yr. avg.",
    "VPD_min_anom" = "VPD min - 3 yr. anomaly",
    "aboveFreezingMonth_anom" = "First Above Freezing Month \n- 3 yr. anomaly",
    "annWatDef" = "Ann. Water Deficit \n- 30yr. avg.", 
    "annWatDef_anom" = "Ann. Water Deficit \n- 3 yr. anomaly", 
    "annWetDegDays_anom" = "Annual Wet Degree Days \n- 3 yr. anomaly", 
    "annWetDegDays" = "Annual Wet Degree Days \n- 30yr. avg.",
    "coarse" = "% coarse", 
    "clay" = "% clay", 
    "frostFreeDays_5_anom" = "No. Frost-Free Days 5th \nPercentile - 3 yr. anomaly",
    "isothermality" = "Isothermality - 30yr. avg.",
    "isothermality_anom" = "Isothermality - 3 yr. anomaly", 
    "prcp" = "MAP - 30yr. avg.",
    "prcp_dry" = "Precip. of Driest Month \n- 30yr. avg.",
    "prcpTempCorr" = "Precip. Temp. Correlation \n- 30yr. avg.",
    "prcpTempCorr_anom" = "Precip. Temp. Correlation \n- 3 yr. anomaly",
    "prcp_seasonality" = "Precip. Seasonality \n- 30yr. avg.", 
    "prcp_seasonality_anom" = "Precip. Seasonality\n - 3 yr. anomaly",
    "prcp_wet_anom" = "Precip. of Wettest Month \n- 3 yr. anomaly", 
    "precp_dry_anom" = "Precip. of Driest Month \n- 3 yr. anomaly", 
    "prcp_anom" = "MAP - 3 yr. anomaly",
    "sand" = "% sand", 
    "carbon" = "% carbon in soil surface", 
    "t_cold_anom" = "Temp. of Coldest Month \n- 3 yr. anomaly",
    "t_warm_anom" = "Temp. of Warmest Month \n- 3 yr. anomaly",
    "tmax_anom" = "Ann. Maximum Temp. \n- 3 yr. anomaly", 
    "tmean" = "MAT - 30yr. avg.",
    "tmin_anom" = "Ann. Minimum Temp. \n- 3 yr. anomaly",
    "prcp_wet" = "Precip. of Wettest Month \n- 30yr. avg.",
    "t_warm" = "Temp. of Warmest Month \n- 30yr. avg.",
    "soilDepth" = "Soil Depth", 
    "t_cold" = "Temp. of Coldest Month \n- 30yr. avg.",
    "annWatDef_95" = "Ann. Water Deficit 95th \nPercentile - 30yr. avg.",
    # squared terms
    "I(prcp^2)" = "(MAP - 30yr. avg.)^2", 
    "I(prcpTempCorr^2)" = "(Precip. Temp. Correlation \n- 30yr. avg.)^2", 
    "I(isothermality^2)" = "(Isothermality - 30yr. avg.)^2",
    "I(clay^2)" = "(% clay)^2",                 
    "I(AWHC^2)" = "(Available Soil \nWater Holding Capacity)^2",
    # interactions
    "prcp:annWetDegDays" = "(MAP - 30yr. avg.) * \nAnnual Wet Degree Days \n- 30yr. avg.",
    "prcp:isothermality" = "(MAP - 30yr. avg.) * \nIsothermality - 30yr. avg.",
    "prcpTempCorr:isothermality" = "Precip. Temp. Correlation \n- 30yr. avg. * \nIsothermality - 30yr. avg.",
    "isothermality:tmean" = " Isothermality - 30yr. avg. * \nMAT - 30yr. avg.",
    "prcpTempCorr:tmean" = "Precip. Temp. Correlation \n- 30yr. avg. * \nMAT - 30yr. avg.",
    "AWHC:clay" = "Available Soil \nWater Holding Capacity * \n% clay",
    "coarse:AWHC" = "% coarse * \nAvailable Soil \nWater Holding Capacity",
    "coarse:clay" =  "% coarse * \nAvailable Soil \n% clay"
  ) %>% 
  data.frame()
names(lookup_md) <- c("longName")
lookup_md$predName <- rownames(lookup_md)

pd_bestLambda <- pd_bestLambda %>% 
  left_join(lookup_md, by = c("predictorVariable_Name" = "predName"))

ggplot(pd_bestLambda) + 
  facet_wrap(~longName,scales = "free_x") + 
  geom_line(aes(x = predictorVariable_value, y = yhat)) + 
  labs(title = "Partial Dependence Plot - predictions from best lambda model",
          x = "Predictor Variable Value", 
          y = paste0("Marginal effect of predictor on tree/no-tree"))
```

For Second-best lambda plot
```{r, fig.width = 10, fig.height = 8}
if ( name_secondBestMod == "Intercept_Only") {
  print("The next best lambda model only contains one predictor (an intercept), so partial dependence plots are impractical")
  } else {# get names of coefficients in models
predNames_pdp_secondBestLambda <- names(coefficients(mod_secondBest)[-1])

## make a 'model.matrix' for the full dataset
X_pdp_secondBestLambda <- model.matrix(mod_secondBest$formula, modDat_1_s_FULL) %>% 
  as.data.frame()
X_pdp_secondBestLambda <- X_pdp_secondBestLambda %>% 
  left_join(modDat_1_s_FULL[prednames_secondBestMod])

for (i in 1:length(predNames_pdp_secondBestLambda)) {
  pd_secondBestLambda_i <- pdp::partial(object = mod_secondBest, pred.var = predNames_pdp_secondBestLambda[i], train = X_pdp_secondBestLambda, type = "regression", )
  names(pd_secondBestLambda_i)[1] <- "predictorVariable_value"
  pd_secondBestLambda_i$predictorVariable_Name <- predNames_pdp_secondBestLambda[i]
  if(i == 1) {
    pd_secondBestLambda <- pd_secondBestLambda_i
  } else {
    pd_secondBestLambda <- rbind(pd_secondBestLambda, pd_secondBestLambda_i)
  }
}
 
## update predictor names to be more interpretable
lookup_md <- c(
    "AWHC" = "Available Soil \nWater Holding Capacity",
    "VPD_mean_anom" = "VPD mean \n- 3 yr. anomaly", 
    "VPD_max" = "VPD max - 30yr. avg.",
    "VPD_max_95_anom" = "VPD max 95th percentile \n- 30yr. avg.",
    "VPD_min_anom" = "VPD min - 3 yr. anomaly",
    "aboveFreezingMonth_anom" = "First Above Freezing Month \n- 3 yr. anomaly",
    "annWatDef" = "Ann. Water Deficit \n- 30yr. avg.", 
    "annWatDef_anom" = "Ann. Water Deficit \n- 3 yr. anomaly", 
    "annWetDegDays_anom" = "Annual Wet Degree Days \n- 3 yr. anomaly", 
    "annWetDegDays" = "Annual Wet Degree Days \n- 30yr. avg.",
    "coarse" = "% coarse", 
    "clay" = "% clay", 
    "frostFreeDays_5_anom" = "No. Frost-Free Days 5th \nPercentile - 3 yr. anomaly",
    "isothermality" = "Isothermality - 30yr. avg.",
    "isothermality_anom" = "Isothermality - 3 yr. anomaly", 
    "prcp" = "MAP - 30yr. avg.",
    "prcp_dry" = "Precip. of Driest Month \n- 30yr. avg.",
    "prcpTempCorr" = "Precip. Temp. Correlation \n- 30yr. avg.",
    "prcpTempCorr_anom" = "Precip. Temp. Correlation \n- 3 yr. anomaly",
    "prcp_seasonality" = "Precip. Seasonality \n- 30yr. avg.", 
    "prcp_seasonality_anom" = "Precip. Seasonality\n - 3 yr. anomaly",
    "prcp_wet_anom" = "Precip. of Wettest Month \n- 3 yr. anomaly", 
    "precp_dry_anom" = "Precip. of Driest Month \n- 3 yr. anomaly", 
    "prcp_anom" = "MAP - 3 yr. anomaly",
    "sand" = "% sand", 
    "carbon" = "% carbon in soil surface", 
    "t_cold_anom" = "Temp. of Coldest Month \n- 3 yr. anomaly",
    "t_warm_anom" = "Temp. of Warmest Month \n- 3 yr. anomaly",
    "tmax_anom" = "Ann. Maximum Temp. \n- 3 yr. anomaly", 
    "tmean" = "MAT - 30yr. avg.",
    "tmin_anom" = "Ann. Minimum Temp. \n- 3 yr. anomaly",
    "prcp_wet" = "Precip. of Wettest Month \n- 30yr. avg.",
    "t_warm" = "Temp. of Warmest Month \n- 30yr. avg.",
    "soilDepth" = "Soil Depth", 
    "t_cold" = "Temp. of Coldest Month \n- 30yr. avg.",
    "annWatDef_95" = "Ann. Water Deficit 95th \nPercentile - 30yr. avg.",
    # squared terms
    "I(prcp^2)" = "(MAP - 30yr. avg.)^2", 
    "I(prcpTempCorr^2)" = "(Precip. Temp. Correlation \n- 30yr. avg.)^2", 
    "I(isothermality^2)" = "(Isothermality - 30yr. avg.)^2",
    "I(clay^2)" = "(% clay)^2",                 
    "I(AWHC^2)" = "(Available Soil \nWater Holding Capacity)^2",
    # interactions
    "prcp:annWetDegDays" = "(MAP - 30yr. avg.) * \nAnnual Wet Degree Days \n- 30yr. avg.",
    "prcp:isothermality" = "(MAP - 30yr. avg.) * \nIsothermality - 30yr. avg.",
    "prcpTempCorr:isothermality" = "Precip. Temp. Correlation \n- 30yr. avg. * \nIsothermality - 30yr. avg.",
    "isothermality:tmean" = " Isothermality - 30yr. avg. * \nMAT - 30yr. avg.",
    "prcpTempCorr:tmean" = "Precip. Temp. Correlation \n- 30yr. avg. * \nMAT - 30yr. avg.",
    "AWHC:clay" = "Available Soil \nWater Holding Capacity * \n% clay",
    "coarse:AWHC" = "% coarse * \nAvailable Soil \nWater Holding Capacity",
    "coarse:clay" =  "% coarse * \nAvailable Soil \n% clay"
  ) %>% 
  data.frame()
names(lookup_md) <- c("longName")
lookup_md$predName <- rownames(lookup_md)

pd_secondBestLambda <- pd_secondBestLambda %>% 
  left_join(lookup_md, by = c("predictorVariable_Name" = "predName"))

ggplot(pd_secondBestLambda) + 
  facet_wrap(~longName,scales = "free_x") + 
  geom_line(aes(x = predictorVariable_value, y = yhat)) + 
  labs(title = "Partial Dependence Plot - predictions from best lambda model",
          x = "Predictor Variable Value", 
          y = paste0("Marginal effect of predictor on tree/no-tree"))
}
```


```{r}
# function for printing model statement
## function to get model statements
getModelStatement <- function(coefficientTable, # name of the d.f that has model coefficients
                  modelName, # name of the column in the coefficient table that has the parameters of interest
                  responseVar # name of the response variable
                  ) {
  ##
  # coefficientTable <- grassShrub_totalHerb_trimAnoms
  # modelName <- "coefficientValue_bestLambda"
  # responseVar <- "TotalHerbaceousCover"
  ##
  temp <- coefficientTable[,c("coefficientName", modelName)] %>% 
    drop_na()
  rownames(temp) <- temp$coefficientName
  temp[,modelName] <- round(temp[,modelName], 8)
  
  # print out coefficients w/ coefficient names
tempNames <- paste0(
  apply(temp, MARGIN = 1, FUN = function(x) {
    if (x["coefficientName"] == "(Intercept)") {
      paste0(x[modelName])
    } else {  
      paste0(x[modelName], "*", x["coefficientName"])
    }
    }
  ),
  collapse = " + "
)

# print the unscaled model statement
  unscaledModelName <- paste0(responseVar, "~ 1/(1 + exp(-(", tempNames, ")))")

## print scaled model statement

return(list("scaledInputVars_ModelStatement" = unscaledModelName))
}


## bestLambda model equations
bestLambdaModStatement <- getModelStatement(coefficientTable = coefs, 
                  modelName = "coefficientValue_bestLambda", 
                  responseVar = "P(Tree)")
## second-best lambda model equations
secondBestLambdaModStatement <- getModelStatement(coefficientTable = coefs, 
                  modelName = ifelse(name_secondBestMod == "1/2 SE Model", 
                                      yes = "coefficientValue_halfseLambda", 
                                     no = "coefficientValue_1seLambda"), 
                  responseVar = "P(Tree)")
```

#### The equation for the bestLambda model if the inputs *are* scaled is: 
```{r}
print(bestLambdaModStatement$scaledInputVars_ModelStatement)
```

#### The equation for the *second*-bestLambda model if the inputs *are* scaled is: 
```{r}
print(secondBestLambdaModStatement$scaledInputVars_ModelStatement)
```


