---
title: "03_modelFitting_testingBetaLASSO"
author: "Alice Stears"
date: "`r lubridate::today()`"
output:
  html_document:
    toc: true
    df_print: paged
    code_folding: hide
    dev: "jpeg"
params:
  run: FALSE     
  save_figs: TRUE
  ecoregion: CONUS # trees or noTrees
  response: BareGroundCover
  allowAllTreeData: TRUE
  removeTexasLouisianaPlain: FALSE
  removeAllAnoms: FALSE
  trimAnomalies: TRUE
  autoKfold: FALSE
  whichSecondBestMod: "auto"# 1se OR halfse OR auto (chose the model w/ fewest # of predictors)
subtitle: "Based on code from Martin Holdrege's Sagebrush-Fire paper"
---
The data consists of vegetation % cover by functional group from across CONUS (from AIM, FIA, LANDFIRE, and RAP), 
as well as climate variables from DayMet, which have been aggregated into mean interannual conditions accross multiple temporal windows. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE,
                      message = FALSE)
```

# Dependencies 

Set user defined parameters

```{r}
print(params)
# set to true if want to run for a limited number of rows (i.e. for code testing)
test_run <- params$test_run
save_figs <- params$save_figs
response <- params$response
fit_sample <- TRUE # fit model to a sample of the data
n_train <- 5e4 # sample size of the training data
n_test <- 1e6 # sample size of the testing data (if this is too big the decile dotplot code throws memory errors)
trimAnom <- params$trimAnomalies
removeTLP <- params$removeTexasLouisianaPlain
run <- params$run
autoKfold <- params$autoKfold
removeAllAnoms <- params$removeAllAnoms
whichSecondBestMod <- params$whichSecondBestMod
allowAllTreeData <- params$allowAllTreeData # should the total tree cover data be curtailed to only include locations that also have NL and BL tree cover information? 
## crs of veg data
crs_veg <- "PROJCRS[\"unnamed\",\n    BASEGEOGCRS[\"unknown\",\n        DATUM[\"unknown\",\n            ELLIPSOID[\"Spheroid\",6378137,298.257223563,\n                LENGTHUNIT[\"metre\",1,\n                    ID[\"EPSG\",9001]]]],\n        PRIMEM[\"Greenwich\",0,\n            ANGLEUNIT[\"degree\",0.0174532925199433,\n                ID[\"EPSG\",9122]]]],\n    CONVERSION[\"Lambert Conic Conformal (2SP)\",\n        METHOD[\"Lambert Conic Conformal (2SP)\",\n            ID[\"EPSG\",9802]],\n        PARAMETER[\"Latitude of false origin\",42.5,\n            ANGLEUNIT[\"degree\",0.0174532925199433],\n            ID[\"EPSG\",8821]],\n        PARAMETER[\"Longitude of false origin\",-100,\n            ANGLEUNIT[\"degree\",0.0174532925199433],\n            ID[\"EPSG\",8822]],\n        PARAMETER[\"Latitude of 1st standard parallel\",25,\n            ANGLEUNIT[\"degree\",0.0174532925199433],\n            ID[\"EPSG\",8823]],\n        PARAMETER[\"Latitude of 2nd standard parallel\",60,\n            ANGLEUNIT[\"degree\",0.0174532925199433],\n            ID[\"EPSG\",8824]],\n        PARAMETER[\"Easting at false origin\",0,\n            LENGTHUNIT[\"metre\",1],\n            ID[\"EPSG\",8826]],\n        PARAMETER[\"Northing at false origin\",0,\n            LENGTHUNIT[\"metre\",1],\n            ID[\"EPSG\",8827]]],\n    CS[Cartesian,2],\n        AXIS[\"easting\",east,\n            ORDER[1],\n            LENGTHUNIT[\"metre\",1,\n                ID[\"EPSG\",9001]]],\n        AXIS[\"northing\",north,\n            ORDER[2],\n            LENGTHUNIT[\"metre\",1,\n                ID[\"EPSG\",9001]]]]"
```

Load packages and source functions
```{r warning=FALSE, message=FALSE}
# set option so resampled dataset created here reproduces earlier runs of this code with dplyr 1.0.10
source("../../../Functions/glmTransformsIterates.R")
source("../../../Functions/transformPreds.R")
source("../../../Functions/betaLASSO.R")

#source("../../../Functions/StepBeta_mine.R")
#source("src/fig_params.R")
#source("src/modeling_functions.R")
 
library(betareg)
library(ggspatial)
library(terra)
library(tidyterra)
library(sf)
library(caret)
library(tidyverse)
library(GGally) # for ggpairs()
library(pdp) # for partial dependence plots
library(gridExtra)
library(patchwork) # for figure insets etc. 
library(ggtext)
library(StepBeta)
theme_set(theme_classic())
library(here)
library(rsample)
library(kableExtra)
library(glmnet)
library(USA.state.boundaries)
```

# Read in data

Data compiled in the `prepDataForModels.R` script

```{r}
here::i_am("Analysis/VegComposition/ModelFitting/03_modelFitting_testingBetaLASSO.Rmd")

# get data that already have forest classification from previous model 
modDat <- readRDS( here("Data_processed", "CoverData", "CoverModelData_withTreeNoTreePreds.rds")) %>% st_drop_geometry()

# change Level 1 cover variables to be between 0 and 1 rather than 0 and 100
modDat[,"TotalTreeCover"] <- modDat[,"TotalTreeCover"]/100
modDat[,"TotalHerbaceousCover"] <- modDat[,"TotalHerbaceousCover"]/100
modDat[,"BareGroundCover"] <- modDat[,"BareGroundCover"]/100
modDat[,"ShrubCover"] <- modDat[,"ShrubCover"]/100

# For all response variables, make sure there are no 0s add  .0001 from each, since the Beta model framework can't handle that
modDat[modDat$TotalTreeCover == 0 & !is.na(modDat$TotalTreeCover), "TotalTreeCover"] <- 0.0001
modDat[modDat$TotalHerbaceousCover == 0  & !is.na(modDat$TotalHerbaceousCover), "TotalHerbaceousCover"] <- 0.0001
modDat[modDat$BareGroundCover == 0 & !is.na(modDat$BareGroundCover), "BareGroundCover"] <- 0.0001
modDat[modDat$ShrubCover == 0 & !is.na(modDat$ShrubCover), "ShrubCover"] <- 0.0001
modDat[modDat$AngioTreeCover_prop == 0 & !is.na(modDat$AngioTreeCover_prop), "AngioTreeCover_prop"] <- 0.0001
modDat[modDat$ConifTreeCover_prop == 0 & !is.na(modDat$ConifTreeCover_prop), "ConifTreeCover_prop"] <- 0.0001
modDat[modDat$C4GramCover_prop == 0 & !is.na(modDat$C4GramCover_prop), "C4GramCover_prop"] <- 0.0001
modDat[modDat$C3GramCover_prop == 0 & !is.na(modDat$C3GramCover_prop), "C3GramCover_prop"] <- 0.0001
modDat[modDat$ForbCover_prop == 0 & !is.na(modDat$ForbCover_prop), "ForbCover_prop"] <- 0.0001

# For all response variables, make sure there are no values of 1 -- subtract .0001 from each, since the Beta model framework can't handle that
modDat[modDat$TotalTreeCover == 1 & !is.na(modDat$TotalTreeCover), "TotalTreeCover"] <- 0.9999
modDat[modDat$TotalHerbaceousCover == 1  & !is.na(modDat$TotalHerbaceousCover), "TotalHerbaceousCover"] <- 0.9999
modDat[modDat$BareGroundCover == 1 & !is.na(modDat$BareGroundCover), "BareGroundCover"] <- 0.9999
modDat[modDat$ShrubCover == 1 & !is.na(modDat$ShrubCover), "ShrubCover"] <- 0.9999
modDat[modDat$AngioTreeCover_prop == 1 & !is.na(modDat$AngioTreeCover_prop), "AngioTreeCover_prop"] <- 0.9999
modDat[modDat$ConifTreeCover_prop == 1 & !is.na(modDat$ConifTreeCover_prop), "ConifTreeCover_prop"] <- 0.9999
modDat[modDat$C4GramCover_prop == 1 & !is.na(modDat$C4GramCover_prop), "C4GramCover_prop"] <- 0.9999
modDat[modDat$C3GramCover_prop == 1 & !is.na(modDat$C3GramCover_prop), "C3GramCover_prop"] <- 0.9999
modDat[modDat$ForbCover_prop == 1 & !is.na(modDat$ForbCover_prop), "ForbCover_prop"] <- 0.9999
```

# Prep data

## Identify the ecoregion and response variable type to use in this model run (i.e. trees or no trees)
```{r get ecoregion variable}
ecoregion <- params$ecoregion
response <- params$response
print(paste0("In this model run, the ecoregion is ", ecoregion," and the response variable is ",response))
```

## Visualize the predictor variables

The following are the candidate predictor variables for this ecoregion: 
```{r get candidate predictor variables}
if (ecoregion == "noTrees") {
  # select potential predictor variables for the ecoregion of interest
        prednames <-
          c(
"tmean"                   ,"prcp_dry"           , "prcpTempCorr"  , "isothermality"           ,"clay"             ,
"coarse"                  ,"AWHC"               , "tmin_anom"     , "tmax_anom"               ,"t_warm_anom"      ,
"t_cold_anom"             ,"prcp_wet_anom"      , "precp_dry_anom", "prcp_seasonality_anom"   ,"prcpTempCorr_anom",
"aboveFreezingMonth_anom" ,"isothermality_anom" , "annWatDef_anom", "annWetDegDays_anom"    
)
} else if (ecoregion == "trees") {
  # select potential predictor variables for the ecoregion of interest
  prednames <- 
    c(
 "tmean"            , "prcp"                    , "prcp_dry"          , "prcpTempCorr"  , "isothermality"        ,  
 "annWatDef"        , "clay"                    , "sand"              , "coarse"        , "AWHC"                 ,  
 "tmax_anom"        , "t_warm_anom"             , "prcp_wet_anom"     , "precp_dry_anom", "prcp_seasonality_anom",  
 "prcpTempCorr_anom", "aboveFreezingMonth_anom" , "isothermality_anom", "annWatDef_anom", "annWetDegDays_anom"   ,  
 "VPD_min_anom" 
)
} else if (ecoregion == "CONUS") {
  prednames <- c("tmean"             , "prcp"          , "prcpTempCorr"        , "isothermality"     ,"annWetDegDays" ,         
"clay"              , "coarse"        , "AWHC"                , "tmax_anom"         ,"t_warm_anom" ,     
"prcp_wet_anom"     , "precp_dry_anom", "prcp_seasonality_anom", "prcpTempCorr_anom", "aboveFreezingMonth_anom",
"isothermality_anom", "annWatDef_anom", "annWetDegDays_anom"  , "VPD_min_anom"            
  )
  
  
 
} 

```

### Scale the predictor variables for the model-fitting process
```{r scaling predictors, fig.height=7, fig.width=10}
modDat_1 <- modDat 
allPreds <- modDat_1 %>% 
  dplyr::select(tmin:frostFreeDays,tmean_anom:frostFreeDays_anom, soilDepth:AWHC) %>% 
  names()
modDat_1_s <- modDat_1 %>% 
  mutate(across(all_of(allPreds), base::scale, .names = "{.col}_s")) 
saveRDS(modDat_1_s, file = "./models/scaledModelInputData.rds")
```

## If predicting total tree cover *and* if desired, restrict the data only to the locations where we have observations for total tree cover and the proportion of that total tree cover that is either needle-leaved or broad-leaved
```{r}
if (allowAllTreeData == FALSE) {
  if(response == "TotalTreeCover") {
  modDat_1_s <- modDat_1_s %>% 
    filter(!is.na(response) & !is.na(AngioTreeCover_prop) & !is.na(ConifTreeCover_prop))
}
}

```

### If the response is tree cover-- Subset the data to only include data for the ecoregion of interest based on whether or not the location is forested (i.e. total tree cover is >10%). Otherwise, determine ecoregion (trees or no trees) based on teh model-predicted classification using the forest model. 
```{r subset data to that ecoregion}
if (response %in% c("TotalTreeCover", "AngioTreeCover_prop", "ConifTreeCover_prop")) {
  if (ecoregion == "noTrees") {
  # select data for the ecoregion of interest
  modDat_1_s <- modDat_1_s %>%
    filter(TotalTreeCover <=0.10)
} else if (ecoregion == "trees") {
  # select data for the ecoregion of interest
  modDat_1_s <- modDat_1_s %>%
    filter(TotalTreeCover > 0.10)
} else if (ecoregion == "CONUS") {
  modDat_1_s <- modDat_1_s
}
} else {
    if (ecoregion == "noTrees") {
  # select data for the ecoregion of interest
  modDat_1_s <- modDat_1_s %>%
    filter(treeProb_bestLambda_binary == 0)
} else if (ecoregion == "trees") {
  # select data for the ecoregion of interest
  modDat_1_s <- modDat_1_s %>%
    filter(treeProb_bestLambda_binary == 1)
} else if (ecoregion == "CONUS") {
  modDat_1_s <- modDat_1_s
}
}
# remove the rows that have no observations for the response variable of interest
modDat_1_s <- modDat_1_s[!is.na(modDat_1_s[,response]),]
```


```{r}
# subset the data to only include these predictors, and remove any remaining NAs 
modDat_1_s <- modDat_1_s %>% 
  dplyr::select(prednames, paste0(prednames, "_s"), response, newRegion, Year, x, y, NA_L1NAME, NA_L2NAME) %>% 
  drop_na()

names(prednames) <- prednames
df_pred <- modDat_1_s[, prednames]
```

## Visualize the response variable 
```{r visualize response variable}
hist(modDat_1_s[,response], main = paste0("Histogram of ",response),
     xlab = paste0(response), breaks = 50)
# plot(density(modDat_1_s[,response]), col = "brown", main = "TotalTreeCover values: observed dataset \nblack = complete dataset; brown = un-forested locations; \ngreen = forested locations", xlab = "Total Tree Cover")
# lines(density(modDat[!is.na(modDat$TotalTreeCover),response]), col = "black")
# lines(density(modDat[(modDat$treeProb_bestLambda_binary == 1) & (!is.na(modDat$TotalTreeCover)) & !is.na(modDat$treeProb_bestLambda_binary),
#                      response]), col = "green")
```


```{r summary_table}
create_summary <- function(df) {
  df %>% 
    pivot_longer(cols = everything(),
                 names_to = 'variable') %>% 
    group_by(variable) %>% 
    summarise(across(value, .fns = list(mean = ~mean(.x, na.rm = TRUE), min = ~min(.x, na.rm = TRUE), 
                                        median = ~median(.x, na.rm = TRUE), max = ~max(.x, na.rm = TRUE)))) %>% 
    mutate(across(where(is.numeric), round, 4))
}

modDat_1_s[prednames] %>% 
  create_summary() %>% 
  knitr::kable(caption = 'summaries of possible predictor variables') %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed")) 

```

### Histograms of raw and scaled predictors
```{r}
  
scaleFigDat_1 <- modDat_1_s %>% 
  dplyr::select(c(x, y, Year, prednames)) %>% 
  pivot_longer(cols = all_of(names(prednames)), 
               names_to = "predNames", 
               values_to = "predValues_unScaled")
scaleFigDat_2 <- modDat_1_s %>% 
  dplyr::select(c(x, y, Year, paste0(prednames, "_s"))) %>% 
  pivot_longer(cols = all_of(paste0(prednames,"_s"
                                    )), 
               names_to = "predNames", 
               values_to = "predValues_scaled", 
               names_sep = ) %>% 
  mutate(predNames = str_replace(predNames, pattern = "_s$", replacement = ""))

scaleFigDat_3 <- scaleFigDat_1 %>% 
  left_join(scaleFigDat_2)

ggplot(scaleFigDat_3) + 
  facet_wrap(~predNames, scales = "free") +
  geom_histogram(aes(predValues_unScaled), fill = "lightgrey", col = "darkgrey") + 
  geom_histogram(aes(predValues_scaled), fill = "lightblue", col = "blue") +
  xlab ("predictor variable values") + 
  ggtitle("Comparing the distribution of unscaled (grey) to scaled (blue) predictor variables")
```

```{r}
modDat_1_s <- modDat_1_s %>% 
  rename_with(~paste0(.x, "_raw"), 
                any_of(names(prednames))) %>% 
  rename_with(~str_remove(.x, "_s$"), 
              any_of(paste0(names(prednames), "_s")))
  
```


## Predictor variables compared to binned response variables

```{r climVar_boxplots, fig.height=9, fig.width=8, eval=TRUE}
set.seed(12011993)
# vector of name of response variables
vars_response <- response
# longformat dataframes for making boxplots
df_sample_plots <-  modDat_1_s  %>% 
  slice_sample(n = 5e4) %>% 
   rename(response = all_of(response)) %>% 
  mutate(response = case_when(
    response <= .25 ~ ".25", 
    response > .25 & response <=.50 ~ ".50", 
    response > .50 & response <=.75 ~ ".75", 
    response >= .75  ~ "1", 
  )) %>% 
  dplyr::select(c(response, prednames)) %>% 
  tidyr::pivot_longer(cols = unname(prednames), 
               names_to = "predictor", 
               values_to = "value"
               )  
 

  ggplot(df_sample_plots, aes_string(x= "response", y = 'value')) +
  geom_boxplot() +
  facet_wrap(~predictor , scales = 'free_y') + 
  ylab("Predictor Variable Values") + 
    xlab(response)

```


# Model Fitting


For the model of total tree cover in non-forested regions, we transform the response variable because all observations, by definition, are between 0 and 0.1. We transform the response variable to be between 0 and 1 by multiplying it by 10, and then will back-transform the model predictions
```{r}
if (response == "TotalTreeCover" & ecoregion == "noTrees") {
 modDat_1_s <- modDat_1_s %>% 
  mutate(TotalTreeCover_transformed = TotalTreeCover * 10)
# because the beta glm does not allow values of 1, make the highest values very slightly less than 1
modDat_1_s[modDat_1_s$TotalTreeCover_transformed == 1, "TotalTreeCover_transformed"] <- 0.9999
response <- "TotalTreeCover_transformed"
}
```


## Visualize the spatial blocks and how they differ across environmental space
First, if there are observations in Louisiana, sub-sample them so they're not so over-represented in the dataset
```{r}
## make data into spatial format
modDat_1_sf <- modDat_1_s %>% 
  st_as_sf(coords = c("x", "y"), crs = st_crs(crs_veg))


# download map info for visualization
data(state_boundaries_wgs84) 

cropped_states <- suppressMessages(state_boundaries_wgs84 %>%
  dplyr::filter(NAME!="Hawaii") %>%
  dplyr::filter(NAME!="Alaska") %>%
  dplyr::filter(NAME!="Puerto Rico") %>%
  dplyr::filter(NAME!="American Samoa") %>%
  dplyr::filter(NAME!="Guam") %>%
  dplyr::filter(NAME!="Commonwealth of the Northern Mariana Islands") %>%
  dplyr::filter(NAME!="United States Virgin Islands") %>%
  sf::st_sf() %>%
  sf::st_transform(sf::st_crs(modDat_1_sf))) #%>%
  #sf::st_crop(sf::st_bbox(modDat_1_sf)+c(-1,-1,1,1))
if (ecoregion %in% c("noTrees", "trees", "CONUS")){
modDat_1_s$uniqueID <- 1:nrow(modDat_1_s)
modDat_1_sf$uniqueID <- 1:nrow(modDat_1_sf)

# find which observations overlap with Louisiana
obs_LA_temp <- st_intersects(modDat_1_sf, cropped_states[cropped_states$NAME == "Louisiana",], sparse = FALSE)
if (sum(obs_LA_temp) > 0) {
 obs_LA_1 <- modDat_1_sf[which(obs_LA_temp == TRUE, arr.ind = TRUE)[,1],]
# now, find only those within the oversampled area (near the coast)
dims <- c(xmin = 730439.1, xmax = 1042195.5, ymax = -1222745.2, ymin = -1390430.9)
badBox <- st_bbox(dims) %>% 
  st_as_sfc() %>% 
  st_set_crs(value = st_crs(modDat_1_sf))
  
obs_LA_temp2 <- st_intersects(obs_LA_1, badBox, sparse = FALSE)
obs_LA_2 <- obs_LA_1[which(obs_LA_temp2 == TRUE, arr.ind = TRUE)[,1],]

# subsample so there aren't so many 
# get every 7th observation
obs_LA_sampled <- obs_LA_2[seq(from = 1, to = nrow(obs_LA_2), by = 10),]
# remove observations that aren't sampled from the larger dataset
obsToRemove <- obs_LA_2[!(obs_LA_2$uniqueID %in% obs_LA_sampled$uniqueID),]

modDat_1_sf <- modDat_1_sf[!(modDat_1_sf$uniqueID %in% obsToRemove$uniqueID),]
modDat_1_s <- modDat_1_s[!(modDat_1_s$uniqueID %in% obsToRemove$uniqueID),] 
}
}

```

```{r Visualize ecoregion distribution, fig.height=10, fig.width=14, message=FALSE, warning=FALSE}

## do a pca of climate across level 2 ecoregions
pca <- prcomp(modDat_1_s[,paste0(prednames)])
library(factoextra)
(fviz_pca_ind(pca, habillage = modDat_1_s$NA_L2NAME, label = "none", addEllipses = TRUE, ellipse.level = .95, ggtheme = theme_minimal(), alpha.ind = .1))

# for ecoregions that are very tiny/extremely different from the rest of the other recoregions climatically, aggregate them together
if(ecoregion == "noTrees" & response != "TotalTreeCover_transformed") {
   modDat_1_s[modDat_1_s$NA_L2NAME %in% c("EVERGLADES", "SOUTHEASTERN USA PLAINS", "TEXAS-LOUISIANA COASTAL PLAIN", "MISSISSIPPI ALLUVIAL AND SOUTHEAST USA COASTAL PLAINS", "TAMAULIPAS-TEXAS SEMIARID PLAIN"), "NA_L2NAME"] <- "SOUTHEASTERN USA PLAINS AND EVERGLADES AND LOUISIANA PLAIN"
   modDat_1_sf[modDat_1_sf$NA_L2NAME %in% c("EVERGLADES", "SOUTHEASTERN USA PLAINS", "TEXAS-LOUISIANA COASTAL PLAIN", "MISSISSIPPI ALLUVIAL AND SOUTHEAST USA COASTAL PLAINS", "TAMAULIPAS-TEXAS SEMIARID PLAIN"), "NA_L2NAME"] <- "SOUTHEASTERN USA PLAINS AND EVERGLADES AND LOUISIANA PLAIN"
} else if(ecoregion == "noTrees" & response == "TotalTreeCover_transformed") {
   # 
   # modDat_1_s[modDat_1_s$NA_L2NAME %in% c("EVERGLADES", "SOUTHEASTERN USA PLAINS", "TEXAS-LOUISIANA COASTAL PLAIN", "MISSISSIPPI ALLUVIAL AND SOUTHEAST USA COASTAL PLAINS", "TAMAULIPAS-TEXAS SEMIARID PLAIN"), "NA_L2NAME"] <- "SOUTHEASTERN USA PLAINS AND EVERGLADES AND LOUISIANA PLAIN"
   # modDat_1_sf[modDat_1_sf$NA_L2NAME %in% c("EVERGLADES", "SOUTHEASTERN USA PLAINS", "TEXAS-LOUISIANA COASTAL PLAIN", "MISSISSIPPI ALLUVIAL AND SOUTHEAST USA COASTAL PLAINS", "TAMAULIPAS-TEXAS SEMIARID PLAIN"), "NA_L2NAME"] <- "SOUTHEASTERN USA PLAINS AND EVERGLADES AND LOUISIANA PLAIN"
   # 
   #  modDat_1_s[modDat_1_s$NA_L2NAME %in% c("ATLANTIC HIGHLANDS", "MIXED WOOD PLAINS"), "NA_L2NAME"] <- "MIXED WOOD PLAINS AND ATLANTIC HIGHLANDS"
   # modDat_1_sf[modDat_1_sf$NA_L2NAME %in% c("ATLANTIC HIGHLANDS", "MIXED WOOD PLAINS"), "NA_L2NAME"] <- "MIXED WOOD PLAINS AND ATLANTIC HIGHLANDS"
   # 
   # 
   #    modDat_1_s[modDat_1_s$NA_L2NAME %in% c("SOUTHEASTERN USA PLAINS AND EVERGLADES AND LOUISIANA PLAIN", "OZARK/OUACHITA-APPALACHIAN FORESTS"), "NA_L2NAME"] <- "SOUTHEASTERN USA PLAINS AND EVERGLADES AND LOUISIANA PLAIN AND APPALACHIAN FORESTS"
   # modDat_1_sf[modDat_1_sf$NA_L2NAME %in% c("SOUTHEASTERN USA PLAINS AND EVERGLADES AND LOUISIANA PLAIN", "OZARK/OUACHITA-APPALACHIAN FORESTS"), "NA_L2NAME"] <- "SOUTHEASTERN USA PLAINS AND EVERGLADES AND LOUISIANA PLAIN AND APPALACHIAN FORESTS"

  #    
  #    modDat_1_s[modDat_1_s$NA_L2NAME %in% c("MIXED WOOD PLAINS", "MIXED WOOD SHIELD"), "NA_L2NAME"] <- "MIXED WOOD PLAINS AND SHIELD"
  #  modDat_1_sf[modDat_1_sf$NA_L2NAME %in% c("MIXED WOOD PLAINS", "MIXED WOOD SHIELD"), "NA_L2NAME"] <- "MIXED WOOD PLAINS AND SHIELD"
  # 
  #  modDat_1_s[modDat_1_s$NA_L2NAME %in% c("MARINE WEST COAST FOREST", "WESTERN CORDILLERA"), "NA_L2NAME"] <- "MARINE WEST COAST FOREST AND WESTERN CORDILLERA"
  #  modDat_1_sf[modDat_1_sf$NA_L2NAME %in% c("MARINE WEST COAST FOREST", "WESTERN CORDILLERA"), "NA_L2NAME"] <- "MARINE WEST COAST FOREST AND WESTERN CORDILLERA"
  #  
  #  
  #   modDat_1_s[modDat_1_s$NA_L2NAME %in% c("SOUTH CENTRAL SEMIARID PRAIRIES", "SOUTHEASTERN USA PLAINS AND EVERGLADES AND LOUISIANA PLAIN"), "NA_L2NAME"] <- "SOUTH CENTRAL SEMIARID PRAIRIES AND COASTAL PLAINS"
  #  modDat_1_sf[modDat_1_sf$NA_L2NAME %in% c("SOUTH CENTRAL SEMIARID PRAIRIES", "SOUTHEASTERN USA PLAINS AND EVERGLADES AND LOUISIANA PLAIN"), "NA_L2NAME"] <- "SOUTH CENTRAL SEMIARID PRAIRIES AND COASTAL PLAINS"
  #  
  #  modDat_1_s[modDat_1_s$NA_L2NAME %in% c("MARINE WEST COAST FOREST AND WESTERN CORDILLERA", "TEMPERATE PRAIRIES"), "NA_L2NAME"] <- "TEMPERATE PRAIRIES"
  #  modDat_1_sf[modDat_1_sf$NA_L2NAME %in% c("MARINE WEST COAST FOREST AND WESTERN CORDILLERA", "TEMPERATE PRAIRIES"), "NA_L2NAME"] <- "TEMPERATE PRAIRIES"
  #  
  #  modDat_1_s[modDat_1_s$NA_L2NAME %in% c("MIXED WOOD PLAINS AND SHIELD", "TEMPERATE PRAIRIES"), "NA_L2NAME"] <- "MIXED WOOD AND TEMPERATE PRAIRIES"
  #  modDat_1_sf[modDat_1_sf$NA_L2NAME %in% c("MIXED WOOD PLAINS AND SHIELD", "TEMPERATE PRAIRIES"), "NA_L2NAME"] <- "MIXED WOOD AND TEMPERATE PRAIRIES"
  # 
  }


# make a table of n for each region
modDat_1_s %>% 
  group_by(NA_L2NAME) %>% 
  dplyr::summarize("Number_Of_Observations" = length(NA_L2NAME)) %>% 
  rename("Level_2_Ecoregion" = NA_L2NAME)%>% 
  kable() %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed")) 

```

## Then, look at the spatial distribution and environmental characteristics of the grouped ecoregions 
```{r fig.width = 14, fig.height=10, message=FALSE}
if (ecoregion == "noTrees") {
  map1 <- ggplot() +
  geom_sf(data=cropped_states,fill='white') +
  # geom_sf(data=modDat_1_sf#[modDat_1_sf$NA_L2NAME %in% c("MIXED WOOD PLAINS"),]
  #         ,
  #         aes(fill=as.factor(NA_L2NAME)),linewidth=0.5,alpha=0.5, size = 1) +
  geom_point(data=modDat_1_s#[modDat_1_s$NA_L2NAME %in% c("MIXED WOOD PLAINS"),]
             ,
             alpha=0.5, 
             aes(x = x, y = y, color=as.factor(NA_L2NAME)), alpha = .1, size = 1) +
  #scale_fill_okabeito() +
  #scale_color_okabeito() +
 # theme_default() +
  theme(legend.position = 'none') +
  labs(title = "Level 2 Ecoregions as spatial blocks")

hull <- modDat_1_sf %>%
  ungroup() %>%
  group_by(NA_L2NAME) %>%
  slice(chull(tmean, prcp_dry))

plot1<-ggplot(data=modDat_1_sf,aes(x=tmean,y=prcp_dry)) +
  geom_polygon(data = hull, alpha = 0.25,aes(fill=NA_L2NAME) )+
  geom_point(aes(group=NA_L2NAME,color=NA_L2NAME),alpha=0.25) +
  theme_minimal() + xlab("30-yr average tmean (scaled)") +
  ylab("30-yr avergae prcp of the driest month (scaled)") #+
  #scale_color_okabeito() +
  #scale_fill_okabeito()

plot2<-ggplot(data=modDat_1_sf %>%
                pivot_longer(cols=c(tmean,prcp_dry)),
              aes(x=value,group=name)) +
  # geom_polygon(data = hull, alpha = 0.25,aes(fill=fold) )+
  geom_density(aes(group=NA_L2NAME,fill=NA_L2NAME),alpha=0.25) +
  theme_minimal() +
  facet_wrap(~name,scales='free')# +
  #scale_color_okabeito() +
  #scale_fill_okabeito()
} else {
  map1 <- ggplot() +
  geom_sf(data=cropped_states,fill='white') +
  geom_sf(data=modDat_1_sf#[modDat_1_sf$NA_L2NAME %in% c("MIXED WOOD PLAINS"),]
          ,
          aes(fill=as.factor(NA_L2NAME)),linewidth=0.5,alpha=0.5, size = 1) +
  geom_point(data=modDat_1_s#[modDat_1_s$NA_L2NAME %in% c("MIXED WOOD PLAINS"),]
             ,
             alpha=0.5, 
             aes(x = x, y = y, color=as.factor(NA_L2NAME)), alpha = .1) +
  #scale_fill_okabeito() +
  #scale_color_okabeito() +
 # theme_default() +
  theme(legend.position = 'none') +
  labs(title = "Level 2 Ecoregions as spatial blocks")

hull <- modDat_1_sf %>%
  ungroup() %>%
  group_by(NA_L2NAME) %>%
  slice(chull(tmean, prcp))

plot1<-ggplot(data=modDat_1_sf,aes(x=tmean,y=prcp)) +
  geom_polygon(data = hull, alpha = 0.25,aes(fill=NA_L2NAME) )+
  geom_point(aes(group=NA_L2NAME,color=NA_L2NAME),alpha=0.25) +
  theme_minimal() + xlab("30-yr avergae tmean (scaled)") +
  ylab("30-yr avergae prcp (scaled)") #+
  #scale_color_okabeito() +
  #scale_fill_okabeito()

plot2<-ggplot(data=modDat_1_sf %>%
                pivot_longer(cols=tmean:prcp),
              aes(x=value,group=name)) +
  # geom_polygon(data = hull, alpha = 0.25,aes(fill=fold) )+
  geom_density(aes(group=NA_L2NAME,fill=NA_L2NAME),alpha=0.25) +
  theme_minimal() +
  facet_wrap(~name,scales='free')# +
  #scale_color_okabeito() +
  #scale_fill_okabeito()
}

 
library(patchwork)
(combo <- (map1+plot1)/plot2) 

```

## Fit a global model with all of the data 

### First, fit a LASSO regression model using the glmnet R package
- This regression is a beta glm with a logit link (using custom function from Daniel)
- Use cross validation across level 2 ecoregions to tune the lambda parameter in the LASSO model 
- this model is fit to using the scaled weather/climate/soils variables
- this list of possible predictors includes: 
  1. main effects
  2. interactions between all soils variables
  3. interactions between climate and weather variables
  4. transformed main effects (squared, log-transformed (add a uniform integer -- 30-- to all variables prior to log-transformation), square root-transformed (add a uniform integer -- 30-- to all variables prior to log-transformation))
```{r fit global model, echo = FALSE}
if (removeAllAnoms) { ## if the model parameters indicate that we don't want any anomalies... 
    # make model matrix (which includes transformations and interactions)
  ## only include interactions within and between climate and anomaly variables, and within and between soils variables

 ## break predictors into climate, weather, and soils 
  prednames_clim <- prednames[which(prednames %in% c("tmin" , "tmax",  "tmean",
     "prcp", "t_warm" , "t_cold" , "prcp_wet" , "prcp_dry" , "prcp_seasonality", "prcpTempCorr", "abvFreezingMonth" , "isothermality" , "annWatDef" , "annWetDegDays",  "VPD_mean",  "VPD_max",  "VPD_min" , "VPD_max_95", "annWatDef_95", "annWetDegDays_5",  "frostFreeDays_5", "frostFreeDays"))]

  prednames_soils <- prednames[which(prednames %in% c( "soilDepth" , "clay", 
     "sand" , "coarse" ,  "carbon", "AWHC"))]
   prednames_weath <- prednames[which(prednames %in% c("tmean_anom", "tmin_anom",  "tmax_anom", 
    "prcp_anom",      "t_warm_anom" , "t_cold_anom",  "prcp_wet_anom" , "precp_dry_anom" ,  "prcp_seasonality_anom" ,  "prcpTempCorr_anom" , "aboveFreezingMonth_anom",  "isothermality_anom" ,  "annWatDef_anom", "annWetDegDays_anom",  "VPD_mean_anom" , "VPD_min_anom" , "VPD_max_anom", "VPD_max_95_anom", "annWatDef_95_anom" , "annWetDegDays_5_anom",  "frostFreeDays_5_anom" , "frostFreeDays_anom"))]
  
  # get a text string of possible interactions between climate and weather variables
  combos <- data.frame(gtools::permutations(n = length(c(prednames_clim)),
                                            r = 2,
                                            v = c(prednames_clim)))
  
  combos$interactions <- paste0(combos$X1, "*", combos$X2)
  
  
  # get a text string of possible interactions between soils variables
  combos_2 <- data.frame(gtools::permutations(n = length(c(prednames_soils)),
                                            r = 2,
                                            v = c(prednames_soils)))
  
  combos_2$interactions <- paste0(combos_2$X1, "*", combos_2$X2)
  
  # make text string of all possible interactions
  int_string <-str_flatten(c(combos$interactions, combos_2$interactions), collapse = " + ")
  
  # make text string of transformations (log, squared)
  transform_string <- str_flatten(c(# add square root term 
                                    paste0("I(log(",prednames_clim,"+30))"), 
                                    paste0("I(log(",prednames_soils,"+30))"),
                                    
                                    paste0("I(sqrt(",prednames_clim,"+30))"), 
                                    paste0("I(sqrt(",prednames_soils,"+30))"),
                                    
                                    paste0("I(",prednames_clim,"^2)"), 
                                    paste0("I(",prednames_soils,"^2)"))
                                    , collapse = " + ")
              
  # get a text string of the possible untransformed model terms
  mod_string <- paste(response, "~", paste0(c(prednames_clim, prednames_soils), collapse = "+"))
  
  # make a text string of the model formula w/ interactions and transformations included
  modWithInteractions <- paste(mod_string, "+",int_string , "+", transform_string
                               )
   
} else { ## if the model parameters indicate that we DO want anomalies... 
    # make model matrix (which includes transformations and interactions)
  ## only include interactions within and between climate and anomaly variables, and within and between soils variables

 ## break predictors into climate, weather, and soils 
  prednames_clim <- prednames[which(prednames %in% c("tmin" , "tmax",  "tmean",
     "prcp", "t_warm" , "t_cold" , "prcp_wet" , "prcp_dry" , "prcp_seasonality", "prcpTempCorr", "abvFreezingMonth" , "isothermality" , "annWatDef" , "annWetDegDays",  "VPD_mean",  "VPD_max",  "VPD_min" , "VPD_max_95", "annWatDef_95", "annWetDegDays_5",  "frostFreeDays_5", "frostFreeDays"))]

  prednames_weath <- prednames[which(prednames %in% c("tmean_anom", "tmin_anom",  "tmax_anom", 
    "prcp_anom",      "t_warm_anom" , "t_cold_anom",  "prcp_wet_anom" , "precp_dry_anom" ,  "prcp_seasonality_anom" ,  "prcpTempCorr_anom" , "aboveFreezingMonth_anom",  "isothermality_anom" ,  "annWatDef_anom", "annWetDegDays_anom",  "VPD_mean_anom" , "VPD_min_anom" , "VPD_max_anom", "VPD_max_95_anom", "annWatDef_95_anom" , "annWetDegDays_5_anom",  "frostFreeDays_5_anom" , "frostFreeDays_anom"))]
  
  prednames_soils <- prednames[which(prednames %in% c( "soilDepth" , "clay", 
     "sand" , "coarse" ,  "carbon", "AWHC"))]
  
  
  # get a text string of possible interactions between climate and weather variables
  combos <- data.frame(gtools::permutations(n = length(c(prednames_clim, prednames_weath)),
                                            r = 2,
                                            v = c(prednames_clim, prednames_weath)))
  
  combos$interactions <- paste0(combos$X1, "*", combos$X2)
  
  
  # get a text string of possible interactions between soils variables
  combos_2 <- data.frame(gtools::permutations(n = length(c(prednames_soils)),
                                            r = 2,
                                            v = c(prednames_soils)))
  
  combos_2$interactions <- paste0(combos_2$X1, "*", combos_2$X2)
  
  # make text string of all possible interactions
  int_string <-str_flatten(c(combos$interactions, combos_2$interactions), collapse = " + ")
  
  # make text string of transformations (log, squared)
  transform_string <- str_flatten(c(# add square root term 
                                    paste0("I(log(",prednames_clim,"+30))"), 
                                    paste0("I(log(",prednames_weath,"+30))"),
                                    paste0("I(log(",prednames_soils,"+30))"),
                                    
                                    paste0("I(sqrt(",prednames_clim,"+30))"), 
                                    paste0("I(sqrt(",prednames_weath,"+30))"),
                                    paste0("I(sqrt(",prednames_soils,"+30))"),
                                    
                                    paste0("I(",prednames_clim,"^2)"), 
                                    paste0("I(",prednames_weath,"^2)"),
                                    paste0("I(",prednames_soils,"^2)"))
                                    , collapse = " + ")
              
  # get a text string of the possible untransformed model terms
  mod_string <- paste(response, "~", paste0(prednames, collapse = "+"))
  
  # make a text string of the model formula w/ interactions and transformations included
  modWithInteractions <- paste(mod_string, "+",int_string , "+", transform_string
                               )
}


  # now, make a model formula object
  f <- as.formula(modWithInteractions)
  
  # transform dataframe to model matrix
  X <- model.matrix(f, modDat_1_s)
  # get response variable
  y <- as.matrix(modDat_1_s[,response])

```
Get rid of transformed predictions and interactions that are correlated 
```{r}
# get first pass of names correlated variables
X_df <- X %>% 
  as.data.frame() %>% 
  dplyr::select(-'(Intercept)')  
corrNames_i <- X_df %>% 
  cor()  %>% 
   caret::findCorrelation(cutoff = .7, verbose = FALSE, names = TRUE, exact = TRUE)
# remove those names that are untransformed main effects 
  # vector of columns to remove 
badNames <- corrNames_i[!(corrNames_i %in% prednames)]
while (sum(!(corrNames_i %in% prednames))>0) {
 corrNames_i <-  X_df %>% 
    dplyr::select(-badNames) %>% 
     cor()  %>% 
   caret::findCorrelation(cutoff = .7, verbose = FALSE, names = TRUE, exact = TRUE)
 # update the vector of names to remove 
 badNames <- unique(c(badNames, corrNames_i[!(corrNames_i %in% prednames)]))
}

## see if there are any correlated variables left (would be all main effects at this point)
# if there are, step through and remove the variable that each is most correlated with 
if (length(corrNames_i)>1) {
  for (i in 1:length(corrNames_i)) {
    X_i <- X_df %>% 
      dplyr::select(-badNames)
    if (corrNames_i[i] %in% names(X_i)) {
    corMat_i <- cor(x = X_i[corrNames_i[i]], y = X_i %>% dplyr::select(-corrNames_i[i])) 
    badNames_i <- colnames(corMat_i)[abs(corMat_i)>=.7]
    # if there are any predictors in the 'badNames_i', remove them from this list
    if (length(badNames_i) > 0 & sum(c(badNames_i %in% prednames))>0) {
        badNames_i <- badNames_i[!(badNames_i %in% prednames)]
    }
    badNames <- unique(c(badNames, badNames_i))
    }
  }
}
## update the X matrix to exclude these correlated variables
X <- X[,!(colnames(X) %in% badNames)]
```

```{r}
if (run == TRUE) {
  # set up custom folds
    # get the ecoregions for training lambda
  train_eco <- modDat_1_s$NA_L2NAME#[train]
  
  # parameter for the 'relax' argument in the cv.glmnet function
  relaxParam <- ifelse(response == "ShrubCover", yes = TRUE, no = 
                         ifelse(c(response == "TotalTreeCover_transformed" & ecoregion == "noTrees"), yes = TRUE, no = FALSE))
  # Fit model -----------------------------------------------
  # specify leave-one-year-out cross-validation
  my_folds <- as.numeric(as.factor(train_eco))

    # set up parallel processing
    library(doMC)
    # this computer has 16 cores (parallel::detectCores())
    registerDoMC(cores = 8)
    
    fit <- cv.glmnet(
      x = X[,2:ncol(X)], 
      y = y, 
      family = quasibeta_family(),
      keep = FALSE,
      alpha = 1,  # 0 == ridge regression, 1 == lasso, 0.5 ~~ elastic net
      #lambda = lambdas, 
      relax = relaxParam,
      #nlambda = 100,
      type.measure="mse",
      #penalty.factor = pen_facts,
      foldid = my_folds,
      #thresh = thresh,
      standardize = FALSE, ## scaled variables prior to the model sequence... coefficients are always returned on the original scale
      parallel = TRUE
    )
     base::saveRDS(fit, paste0("../ModelFitting/models/betaLASSO/", response, "_globalLASSOmod_betaLogitLink_",ecoregion,"_removeAnomalies", removeAllAnoms,"observeDatForestClassification.rds"))
    
     best_lambda <- fit$lambda.min
  # save the lambda for the most regularized model w/ an MSE that is still 1SE w/in the best lambda model
  lambda_1SE <- fit$lambda.1se
  # save the lambda for the most regularized model w/ an MSE that is still .5SE w/in the best lambda model
  lambda_halfSE <- best_lambda + ((lambda_1SE - best_lambda)/2)
 
## Now, we need to do stability selection to ensure the coefficients that are being chosen with each lambda are stable 

## stability selection for best lambda model 
# setup params
p <- ncol(X[,2:ncol(X)]) # of parameters
n <- length(y) # of observations
n_iter <- 100        # number of subsamples
sample_frac <- 0.75  # fraction of data to subsample
lambda_val <- best_lambda    # fixed lambda value (could be chosen via CV)

# Track selection
selection_counts_bestL <- matrix(0, nrow = p, ncol = 1)

for (i in 1:n_iter) {
  # Subsample rows
  sample_idx <- sample(1:n, size = floor(sample_frac * n), replace = FALSE)
  X_sub <- X[sample_idx, ]
  y_sub <- y[sample_idx]

  # Fit Lasso model
  fit_stab_i <- glmnet(x = X_sub[,2:ncol(X_sub)], y = y_sub, 
    family = quasibeta_family(),
    alpha = 1, lambda = lambda_val, standardize = FALSE)

  # Get non-zero coefficients (excluding intercept)
  select_bestL <- which(as.vector(coef(fit_stab_i))[-1] != 0)
  selection_counts_bestL[select_bestL] <- selection_counts_bestL[select_bestL] + 1
}

# Convert counts to selection probabilities (the probability that a variable is selected over 100 iterations)
selection_prob_bestL <- selection_counts_bestL / n_iter
selection_prob_bestL_df <- data.frame(
  VariableNumber = paste0("X", 1:p),
  VariableName = rownames(coef(fit_stab_i))[2:(p+1)],
  SelectionProb = as.vector(selection_prob_bestL)
)

# get those variables that are selected in ≥70% of subsets (Meinshausen and Bühlmann, 2010)
bestLambda_coef <- selection_prob_bestL_df[selection_prob_bestL_df$SelectionProb>=.7, c("VariableName", "SelectionProb")]

#//////
# stability selection for 1se lambda model
lambda_val <-  lambda_1SE    # fixed lambda value (could be chosen via CV)

# Track selection
selection_counts_1seL <- matrix(0, nrow = p, ncol = 1)

for (i in 1:n_iter) {
  # Subsample rows
  sample_idx <- sample(1:n, size = floor(sample_frac * n), replace = FALSE)
  X_sub <- X[sample_idx, ]
  y_sub <- y[sample_idx]

  # Fit Lasso model
  fit_stab_i <- glmnet(x = X_sub[,2:ncol(X_sub)], y = y_sub, 
    family = quasibeta_family(),
    alpha = 1, lambda = lambda_val, standardize = FALSE)

  # Get non-zero coefficients (excluding intercept)
  selected_1seL <- which(as.vector(coef(fit_stab_i))[-1] != 0)
  selection_counts_1seL[selected_1seL] <- selection_counts_1seL[selected_1seL] + 1
}

# Convert counts to selection probabilities (the probability that a variable is selected over 100 iterations)
selection_prob_1seL <- selection_counts_1seL / n_iter
selection_prob_1seL_df <- data.frame(
  VariableNumber = paste0("X", 1:p),
  VariableName = rownames(coef(fit_stab_i))[2:(p+1)],
  SelectionProb = as.vector(selection_prob_1seL)
)

# get those variables that are selected in ≥70% of subsets (Meinshausen and Bühlmann, 2010)
seLambda_coef <- selection_prob_1seL_df[selection_prob_1seL_df$SelectionProb>=.7, c("VariableName", "SelectionProb")]

# stability selection for half se lambda model
lambda_val <- lambda_halfSE    # fixed lambda value (could be chosen via CV)

# Track selection
selection_counts_halfseL <- matrix(0, nrow = p, ncol = 1)

for (i in 1:n_iter) {
  # Subsample rows
  sample_idx <- sample(1:n, size = floor(sample_frac * n), replace = FALSE)
  X_sub <- X[sample_idx, ]
  y_sub <- y[sample_idx]

  # Fit Lasso model
  fit_stab_i <- glmnet(x = X_sub[,2:ncol(X_sub)], y = y_sub, 
    family = quasibeta_family(),
    alpha = 1, lambda = lambda_val, standardize = FALSE)

  # Get non-zero coefficients (excluding intercept)
  selected_halfseL <- which(as.vector(coef(fit_stab_i))[-1] != 0)
  selection_counts_halfseL[selected_halfseL] <- selection_counts_halfseL[selected_halfseL] + 1
}

# Convert counts to selection probabilities (the probability that a variable is selected_halfseL over 100 iterations)
selection_prob_halfseL <- selection_counts_halfseL / n_iter
selection_prob_halfseL_df <- data.frame(
  VariableNumber = paste0("X", 1:p),
  VariableName = rownames(coef(fit_stab_i))[2:(p+1)],
  SelectionProb = as.vector(selection_prob_halfseL)
)

# get those variables that are selected_halfseL_halfseL in ≥70% of subsets (Meinshausen and Bühlmann, 2010)
halfseLambda_coef <- selection_prob_halfseL_df[selection_prob_halfseL_df$SelectionProb>=.7, c("VariableName", "SelectionProb")]

## If prompted to do so by the input arguments, remove any predictors that are for weather anomalies whose corresponding climate predictor is not included in the model 
if (trimAnom == TRUE) {
  # get unique predictors
  bestLamb_all <- bestLambda_coef$VariableName %>% #[str_detect(bestLambda_coef$VariableName, pattern = "_anom_s")] %>% 
    str_remove(pattern = "I\\(") %>% 
    str_remove(pattern = "\\^2\\)") %>% 
    str_remove(pattern = "\\^2\\)") %>% 
    str_split(pattern = ":", simplify = TRUE) 
  if (ncol(bestLamb_all) == 1) {
    bestLamb_all <- unique(bestLamb_all)
  } else {
    bestLamb_all <-  c(bestLamb_all[bestLamb_all[,1] !="",1], bestLamb_all[bestLamb_all[,2] !="",2]) %>% 
      unique()
  }
  # get anomalies
   bestLamb_anom <- bestLamb_all[bestLamb_all %in% prednames_weath]
  # get climate
   bestLamb_clim <- bestLamb_all[bestLamb_all %in% prednames_clim]
  # which anomalies are present in the predictors, but their corresponding climate variable isn't
   bestLamb_anomsWithMissingClim <- bestLamb_anom[!(str_remove(bestLamb_anom, "_anom") %in% bestLamb_clim)]
   # remove anomalies (and all interaction terms w/ those anomalies) from the predictor list

   if (length(bestLamb_anomsWithMissingClim) != 0) {
   
        bestLambda_coef_NEW <- bestLambda_coef
      for (i in 1:length(bestLamb_anomsWithMissingClim)) {
     bestLambda_coef_NEW <- bestLambda_coef_NEW[!str_detect(bestLambda_coef_NEW$VariableName, pattern = bestLamb_anomsWithMissingClim[i]),]
      }
     
   bestLambda_coef <- bestLambda_coef_NEW
   }
   
   
  #//// 1 se lambda model
      if (nrow(seLambda_coef) !=0) {
        # get unique predictors
  seLamb_all <- seLambda_coef$VariableName %>% #[str_detect(seLambda_coef$VariableName, pattern = "_anom_s")] %>% 
    str_remove(pattern = "I\\(") %>% 
    str_remove(pattern = "_s\\^2\\)") %>% 
    str_remove(pattern = "\\^2\\)") %>% 
    str_split(pattern = ":", simplify = TRUE) 
  if (ncol(seLamb_all) == 1) {
    seLamb_all <- unique(seLamb_all)
  } else {
    seLamb_all <-  c(seLamb_all[seLamb_all[,1] !="",1], seLamb_all[seLamb_all[,2] !="",2]) %>% 
      unique()
  }
  # get anomalies
   seLamb_anom <- seLamb_all[seLamb_all %in% prednames_weath]
  # get climate
   seLamb_clim <- seLamb_all[seLamb_all %in% prednames_clim]
  # which anomalies are present in the predictors, but their corresponding climate variable isn't
   seLamb_anomsWithMissingClim <- seLamb_anom[!(str_remove(seLamb_anom, "_anom") %in%  seLamb_clim)]
   # remove anomalies (and all interaction terms w/ those anomalies) from the predictor list
      if (length(seLamb_anomsWithMissingClim) != 0) {
    seLambda_coef_NEW <- seLambda_coef
   for (i in 1:length(seLamb_anomsWithMissingClim)) {
     seLambda_coef_NEW <- seLambda_coef_NEW[!str_detect(seLambda_coef_NEW$VariableName, pattern = seLamb_anomsWithMissingClim[i]),]
   }
   seLambda_coef <- seLambda_coef_NEW
      }
   }
  
  #//// 1 se lambda model
      if (nrow(halfseLambda_coef) !=0) {
        # get unique predictors
  halfseLamball<- halfseLambda_coef$VariableName %>% #[str_detect(halfseLambda_coef$VariableName, pattern = "_anom_s")] %>% 
    str_remove(pattern = "I\\(") %>% 
    str_remove(pattern = "_s\\^2\\)") %>% 
    str_remove(pattern = "\\^2\\)") %>% 
    str_split(pattern = ":", simplify = TRUE) 
  
    if (ncol(halfseLamball) == 1) {
    halfseLamball <- unique(halfseLamball)
  } else {
    halfseLamball <-  c(halfseLamball[halfseLamball[,1] !="",1], halfseLamball[halfseLamball[,2] !="",2]) %>% 
      unique()
  }
  
  # get anomalies
   halfseLambanom <- halfseLamball[halfseLamball %in% prednames_weath]
  # get climate
   halfseLambclim <- halfseLamball[halfseLamball %in% prednames_clim]
  # which anomalies are present in the predictors, but their corresponding climate variable isn't
   halfseLambanomsWithMissingClim <- halfseLambanom[!(str_remove(halfseLambanom, "_anom_s") %in%  halfseLambclim)]
   # remove anomalies (and all interaction terms w/ those anomalies) from the predictor list

   
   ##
        if (length(halfseLambanomsWithMissingClim) != 0) {
   halfseLambda_coef_NEW <- halfseLambda_coef
   for (i in 1:length(halfseLambanomsWithMissingClim)) {
     halfseLambda_coef_NEW <- halfseLambda_coef_NEW[!str_detect(halfseLambda_coef_NEW$VariableName, pattern = halfseLambanomsWithMissingClim[i]),]
   }
   halfseLambda_coef <- halfseLambda_coef_NEW
      }
   }
    ##
  
      }


## 
## fit w/ the identified coefficients from the 'best' lambda, but using the glm function
  # make the modDat_1_s matrix into a data.frame for model fitting
  X_dataFrame <- X[,2:ncol(X)] %>% 
    cbind(y) %>% as.data.frame() 
  names(X_dataFrame)[ncol(X_dataFrame)] <- response
  
mat_glmnet_best <- bestLambda_coef$VariableName 

  if (length(mat_glmnet_best) == 0) {
    f_glm_bestLambda <- as.formula(paste0(response, "~ 1"))
  } else {
  f_glm_bestLambda <- as.formula(paste0(response, " ~ ", paste0(mat_glmnet_best, collapse = " + ")))
  }
  
## fit using betareg
  fit_glm_bestLambda <- fit_glm_bestLambda_quasiBETA <- glm(formula = f_glm_bestLambda,data = X_dataFrame#modDat_1_s
                                                            , family =quasibeta_family(link = "logit") )
    
  # fit_glm_bestLambda_BETAREG <- betareg(formula =  f_glm_bestLambda, data = X_dataFrame#modDat_1_s
  #                                       , link = "logit")
  ## fit using quasi-beta
  # ## compare coefficients
  # modCoeffsTest <- data.frame("coefficientName" = names(coefficients(fit_glm_bestLambda_BETAREG)),
  #                             "coefficientValue" = coefficients(fit_glm_bestLambda_BETAREG), 
  #                             "confint_low" = confint(fit_glm_bestLambda_BETAREG)[,1],
  #                             "confint_high" = confint(fit_glm_bestLambda_BETAREG)[,2],
  #                             "type" = "betareg") %>% 
  #   rbind(data.frame("coefficientName" = names(coefficients(fit_glm_bestLambda_quasiBETA)),
  #                             "coefficientValue" = coefficients(fit_glm_bestLambda_quasiBETA), 
  #                             "confint_low" = coefficients(fit_glm_bestLambda_quasiBETA) - ( summary(fit_glm_bestLambda_quasiBETA)$coefficients[,2] * 1.96) ,
  #                             "confint_high" = coefficients(fit_glm_bestLambda_quasiBETA) + ( summary(fit_glm_bestLambda_quasiBETA)$coefficients[,2] * 1.96),
  #                             "type" = "quasiBeta")) %>% 
  #   filter(coefficientName != "(phi)")
  # ggplot(data = modCoeffsTest) + 
  #   geom_point(aes(x = coefficientValue, y = coefficientName, col = type)) + 
  #   geom_errorbar(aes(xmin = confint_low, xmax = confint_high, y = coefficientName, col = type)) + 
  #   geom_vline(aes(xintercept = 0), lty = 3)
  # 
   ## fit w/ the identified coefficients from the '1se' lambda, but using the glm function
  mat_glmnet_1se <- seLambda_coef$VariableName
    
  if (length(mat_glmnet_1se) == 0) {
    f_glm_1se <- as.formula(paste0(response, "~ 1"))
  } else {
  f_glm_1se <- as.formula(paste0(response, " ~ ", paste0(mat_glmnet_1se, collapse = " + ")))
  }

  fit_glm_se <- glm(formula = f_glm_1se, data = X_dataFrame#modDat_1_s
                    , family =quasibeta_family(link = "logit") )
  # glm(data = modDat_1_s, formula = f_glm_1se,
  #                   family =  stats::Gamma(link = "log"))
  
     ## fit w/ the identified coefficients from the '.5se' lambda, but using the glm function
  mat_glmnet_halfse <- halfseLambda_coef$VariableName
  
  if (length(mat_glmnet_halfse) == 0) {
    f_glm_halfse <- as.formula(paste0(response, "~ 1"))
  } else {
  f_glm_halfse <- as.formula(paste0(response, " ~ ", paste0(mat_glmnet_halfse, collapse = " + ")))
  }

  fit_glm_halfse <- glm(formula = f_glm_halfse, data = X_dataFrame#modDat_1_s
                        , family =quasibeta_family(link = "logit") )
  
  ## save models 
  if (trimAnom == TRUE) {
    saveRDS(fit_glm_bestLambda, paste0("./models/betaLASSO/",response,"_",ecoregion, "_noTLP_",removeTLP,"_removeAnomalies", removeAllAnoms,"_trimAnom_bestLambdaGLM_observeDatForestClassification.rds"))
  saveRDS(fit_glm_halfse, paste0("./models/betaLASSO/",response,"_",ecoregion, "_noTLP_",removeTLP,"_removeAnomalies", removeAllAnoms,"_trimAnom_halfSELambdaGLM_observeDatForestClassification.rds"))
  saveRDS(fit_glm_se, paste0("./models/betaLASSO/",response,"_",ecoregion, "_noTLP_",removeTLP,"_removeAnomalies", removeAllAnoms,"_trimAnom_oneSELambdaGLM_observeDatForestClassification.rds"))
  } else {
    saveRDS(fit_glm_bestLambda, paste0("./models/betaLASSO/",response,"_",ecoregion, "_noTLP_",removeTLP,"_removeAnomalies", removeAllAnoms,"_bestLambdaGLM_observeDatForestClassification.rds"))
  saveRDS(fit_glm_halfse, paste0("./models/betaLASSO/",response,"_",ecoregion, "_noTLP_",removeTLP,"_removeAnomalies", removeAllAnoms,"_halfSELambdaGLM_observeDatForestClassification.rds"))
  saveRDS(fit_glm_se, paste0("./models/betaLASSO/",response,"_",ecoregion, "_noTLP_",removeTLP,"_removeAnomalies", removeAllAnoms,"_oneSELambdaGLM_observeDatForestClassification.rds"))
    
  }
  
  ## save the R environment after running the models 
  save(f_glm_halfse, mat_glmnet_halfse, halfseLambda_coef,
              f_glm_1se, mat_glmnet_1se, seLambda_coef,
              f_glm_bestLambda, mat_glmnet_best, bestLambda_coef,
              file = paste0("./models/interimModelFittingObjects_", response, "_", ecoregion, "_removeAnomalies", removeAllAnoms,"_betaReg_observeDatForestClassification.rds" ))
  } else {
    if (response == "TotalTreeCover_transformed" & ecoregion == "noTrees") {
      # read in LASSO object
    fit <- readRDS(paste0("../ModelFitting/models/betaLASSO/", response, "_globalLASSOmod_betaLogitLink_",ecoregion,"_removeAnomalies", removeAllAnoms,"observeDatForestClassification.rds"))
    
    # read in R objects having to do w/ model fitting 
    load(file = paste0("./models/interimModelFittingObjects_", response, "_", ecoregion, "_removeAnomalies", removeAllAnoms,"_betaReg_observeDatForestClassification.rds" ))
    
     if (trimAnom == TRUE) {
    fit_glm_bestLambda <- readRDS( paste0("./models/betaLASSO/",response,"_",ecoregion, "_noTLP_",removeTLP,"_removeAnomalies", removeAllAnoms,"_trimAnom_bestLambdaGLM_observeDatForestClassification.rds"))
  fit_glm_halfse <- readRDS( paste0("./models/betaLASSO/",response,"_",ecoregion, "_noTLP_",removeTLP,"_removeAnomalies", removeAllAnoms,"_trimAnom_halfSELambdaGLM_observeDatForestClassification.rds"))
 fit_glm_se <- readRDS( paste0("./models/betaLASSO/",response,"_",ecoregion, "_noTLP_",removeTLP,"_removeAnomalies", removeAllAnoms,"_trimAnom_oneSELambdaGLM_observeDatForestClassification.rds"))
  } else {
    fit_glm_bestLambda <- readRDS( paste0("./models/betaLASSO/",response,"_",ecoregion, "_noTLP_",removeTLP,"_removeAnomalies", removeAllAnoms,"_bestLambdaGLM_observeDatForestClassification.rds"))
  fit_glm_halfse <- readRDS( paste0("./models/betaLASSO/",response,"_",ecoregion, "_noTLP_",removeTLP,"_removeAnomalies", removeAllAnoms,"_halfSELambdaGLM_observeDatForestClassification.rds"))
  fit_glm_se <- readRDS( paste0("./models/betaLASSO/",response,"_",ecoregion, "_noTLP_",removeTLP,"_removeAnomalies", removeAllAnoms,"_oneSELambdaGLM_observeDatForestClassification.rds"))
    
  }
    
    } else {
      # read in LASSO object
    fit <- readRDS(paste0("../ModelFitting/models/betaLASSO/", response, "_globalLASSOmod_betaLogitLink_",ecoregion,"_removeAnomalies", removeAllAnoms, ".rds"))
    
    # read in R objects having to do w/ model fitting 
    load(file = paste0("./models/interimModelFittingObjects_", response, "_", ecoregion, "_removeAnomalies", removeAllAnoms,"_betaReg.rds" ))
    
      
    if (trimAnom == TRUE) {
    fit_glm_bestLambda <- readRDS( paste0("./models/betaLASSO/",response,"_",ecoregion, "_noTLP_",removeTLP,"_removeAnomalies", removeAllAnoms,"_trimAnom_bestLambdaGLM.rds"))
  fit_glm_halfse <- readRDS( paste0("./models/betaLASSO/",response,"_",ecoregion, "_noTLP_",removeTLP,"_removeAnomalies", removeAllAnoms,"_trimAnom_halfSELambdaGLM.rds"))
 fit_glm_se <- readRDS( paste0("./models/betaLASSO/",response,"_",ecoregion, "_noTLP_",removeTLP,"_removeAnomalies", removeAllAnoms,"_trimAnom_oneSELambdaGLM.rds"))
  } else {
    fit_glm_bestLambda <- readRDS( paste0("./models/betaLASSO/",response,"_",ecoregion, "_noTLP_",removeTLP,"_removeAnomalies", removeAllAnoms,"_bestLambdaGLM.rds"))
  fit_glm_halfse <- readRDS( paste0("./models/betaLASSO/",response,"_",ecoregion, "_noTLP_",removeTLP,"_removeAnomalies", removeAllAnoms,"_halfSELambdaGLM.rds"))
  fit_glm_se <- readRDS( paste0("./models/betaLASSO/",response,"_",ecoregion, "_noTLP_",removeTLP,"_removeAnomalies", removeAllAnoms,"_oneSELambdaGLM.rds"))
    
  }
    
    }
  }


  # assess model fit
  # assess.glmnet(fit$fit.preval, #newx = X[,2:293], 
  #               newy = y, family = stats::Gamma(link = "log"))
  # save the minimum lambda
  best_lambda <- fit$lambda.min
  # save the lambda for the most regularized model w/ an MSE that is still 1SE w/in the best lambda model
  lambda_1SE <- fit$lambda.1se
  # save the lambda for the most regularized model w/ an MSE that is still .5SE w/in the best lambda model
  lambda_halfSE <- best_lambda + ((lambda_1SE - best_lambda)/2)
 
  print(fit)     
  
  plot(fit)

```

Then, we predict (on the training set) using both of these models (best lambda and 1se lambda) 
```{r Prediction}
## predict on the test data
## newdata for predictions
newdata_DF <- data.frame(X[,2:ncol(X)])
names(newdata_DF) <- colnames(X[,2:ncol(X)])  

# predictions w/ the 'best lambda' glm model   
optimal_pred <- predict.glm(fit_glm_bestLambda, newdata = newdata_DF, type = "response")
# predictions w/ the 'best lambda' model fit with betareg()
#optimal_pred_betareg <- betareg:::predict.betareg(object = fit_glm_bestLambda_BETAREG, newdata = newdata_DF, type = "response")
  # plot(density(optimal_pred_OLD)) 
  # lines(density(optimal_pred_NEW), col = "red")
  
 optimal_pred_ByHand <- exp(predict.glm(fit_glm_bestLambda, newdata = newdata_DF))/
   (1 +  exp(predict.glm(fit_glm_bestLambda, newdata = newdata_DF)))
  optimal_pred_1se <-  predict.glm(fit_glm_se, newdata = newdata_DF, type = "response")
  optimal_pred_halfse <- predict.glm(fit_glm_halfse, newdata = newdata_DF, type = "response")
  
    null_fit <- glm(#data = data.frame("y" = y, X[,paste0(prednames, "_s")]), 
      formula = y ~ 1, #data = modDat_1_s, 
      family = quasibeta_family(link = "logit")
      #formula = y ~ 1, family = stats::Gamma(link = "log")
      )
  null_pred <- predict.glm(null_fit, newdata = as.data.frame(X), type = "response"
                       )

  # ## snap values above 100 and below 0 to be 100 or z
  # #optimal_pred[optimal_pred>100] <- 100
  # optimal_pred[optimal_pred<0] <- 0
  # #optimal_pred_1se[optimal_pred_1se>100] <- 100
  # optimal_pred_1se[optimal_pred_1se<0] <- 0
  # #optimal_pred_halfse[optimal_pred_halfse>100] <- 100
  # optimal_pred_halfse[optimal_pred_halfse<0] <- 0
  # 
  
  if(response == "TotalTreeCover_transformed") {
       # save data
  fullModOut <- list(
    "modelObject" = fit,
    "nullModelObject" = null_fit,
    "modelPredictions" = data.frame(#ecoRegion_holdout = rep(test_eco,length(y)),
      obs=y/10,
                    pred_opt=optimal_pred/10, 
                    pred_opt_se = optimal_pred_1se/10,
                    pred_opt_halfse = optimal_pred_halfse/10,
                    pred_null=null_pred/10#,
                    #pred_nopenalty=nopen_pred
                    ))
  } else {
    # save data
  fullModOut <- list(
    "modelObject" = fit,
    "nullModelObject" = null_fit,
    "modelPredictions" = data.frame(#ecoRegion_holdout = rep(test_eco,length(y)),
      obs=y,
                    pred_opt=optimal_pred, 
                    pred_opt_se = optimal_pred_1se,
                    pred_opt_halfse = optimal_pred_halfse,
                    pred_null=null_pred#,
                    #pred_nopenalty=nopen_pred
                    ))
  }
  

ggplot() + 
  geom_point(aes(X[,2], fullModOut$modelPredictions$obs), col = "black", alpha = .1) + 
  geom_point(aes(X[,2], fullModOut$modelPredictions$pred_opt), col = "red", alpha = .1) + ## predictions w/ the CV model
  geom_point(aes(X[,2], fullModOut$modelPredictions$pred_opt_halfse), col = "orange", alpha = .1) + ## predictions w/ the CV model (.5se lambda)
  geom_point(aes(X[,2], fullModOut$modelPredictions$pred_opt_se), col = "green", alpha = .1) + ## predictions w/ the CV model (1se lambda)
  geom_point(aes(X[,2], fullModOut$modelPredictions$pred_null), col = "blue", alpha = .1) + 
  labs(title = "A rough comparison of observed and model-predicted values", 
       subtitle = "black = observed values \n red = predictions from 'best lambda' model \n orange = predictions for '1/2se' lambda model \n green = predictions from '1se' lambda model \n blue = predictions from null model") +
  xlab(colnames(X)[2])
  #ylim(c(0,200))
 
```

The internal cross-validation process to fit the global LASSO model identified an optimal lambda value (regularization parameter) of `r{print(best_lambda)}`. The lambda value such that the cross validation error is within 1 standard error of the minimum ("1se lambda") was `r{print(fit$lambda.1se)}`` . The following coefficients were kept in each model: 
```{r coefficients in the global model}
# the coefficient matrix from the 'best model' -- find and print those coefficients that aren't 0 in a table
# the coefficient   matrix from the 'best model' -- find and print those coefficients that aren't 0 in a table
coef_glm_bestLambda <- coef(fit_glm_bestLambda) %>% 
  data.frame() 
coef_glm_bestLambda$coefficientName <- rownames(coef_glm_bestLambda)
names(coef_glm_bestLambda)[1] <- "coefficientValue_bestLambda"
# coefficient matrix from the '1se' model 
coef_glm_1se <- coef(fit_glm_se) %>% 
  data.frame() 
coef_glm_1se$coefficientName <- rownames(coef_glm_1se)
names(coef_glm_1se)[1] <- "coefficientValue_1seLambda"
# coefficient matrix from the 'half se' model 
coef_glm_halfse <- coef(fit_glm_halfse) %>% 
  data.frame() 
coef_glm_halfse$coefficientName <- rownames(coef_glm_halfse)
names(coef_glm_halfse)[1] <- "coefficientValue_halfseLambda"
# add together
coefs <- full_join(coef_glm_bestLambda, coef_glm_halfse) %>% 
  full_join(coef_glm_1se) %>% 
  dplyr::select(coefficientName, coefficientValue_bestLambda,
                coefficientValue_halfseLambda, coefficientValue_1seLambda)

globModTerms <- coefs[!is.na(coefs$coefficientValue_bestLambda), "coefficientName"]

## also, get the number of unique variables in each model 
var_prop_pred <- paste0(response, "_pred")
response_vars <- c(response, var_prop_pred)
# for best lambda model
prednames_fig <- paste(str_split(globModTerms, ":", simplify = TRUE)) 
prednames_fig <- str_replace(prednames_fig, "I\\(", "")
prednames_fig <- str_replace(prednames_fig, "\\^2\\)", "")
prednames_fig <- unique(prednames_fig[prednames_fig>0])
prednames_fig <- prednames_fig
prednames_fig_num <- length(prednames_fig)
# for 1SE lambda model
globModTerms_1se <- coefs[!is.na(coefs$coefficientValue_1seLambda), "coefficientName"]
if (length(globModTerms_1se) == 1) {
prednames_fig_1se <- paste(str_split(globModTerms_1se, ":", simplify = TRUE)) 
prednames_fig_1se <- str_replace(prednames_fig_1se, "I\\(", "")
prednames_fig_1se <- str_replace(prednames_fig_1se, "\\^2\\)", "")
prednames_fig_1se <- unique(prednames_fig_1se[prednames_fig_1se>0])
prednames_fig_1se_num <- c(0)
} else {
prednames_fig_1se <- paste(str_split(globModTerms_1se, ":", simplify = TRUE)) 
prednames_fig_1se <- str_replace(prednames_fig_1se, "I\\(", "")
prednames_fig_1se <- str_replace(prednames_fig_1se, "\\^2\\)", "")
prednames_fig_1se <- unique(prednames_fig_1se[prednames_fig_1se>0])
prednames_fig_1se_num <- length(prednames_fig_1se)
}
# for 1/2SE lambda model
globModTerms_halfse <- coefs[!is.na(coefs$coefficientValue_halfseLambda), "coefficientName"]
if (length(globModTerms_halfse) == 1) {
prednames_fig_halfse <- paste(str_split(globModTerms_halfse, ":", simplify = TRUE)) 
prednames_fig_halfse <- str_replace(prednames_fig_halfse, "I\\(", "")
prednames_fig_halfse <- str_replace(prednames_fig_halfse, "\\^2\\)", "")
prednames_fig_halfse <- unique(prednames_fig_halfse[prednames_fig_halfse>0])
prednames_fig_halfse_num <- c(0)
} else {
prednames_fig_halfse <- paste(str_split(globModTerms_halfse, ":", simplify = TRUE)) 
prednames_fig_halfse <- str_replace(prednames_fig_halfse, "I\\(", "")
prednames_fig_halfse <- str_replace(prednames_fig_halfse, "\\^2\\)", "")
prednames_fig_halfse <- unique(prednames_fig_halfse[prednames_fig_halfse>0])
prednames_fig_halfse_num <- length(prednames_fig_halfse)
}
# make a table
kable(coefs, col.names = c("Coefficient Name", "Value from best lambda model", 
                           "Value from 1/2 se lambda", "Value from 1se lambda model")
      ) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed")) 

# calculate RMSE of all models 
RMSE_best <- yardstick::rmse(fullModOut$modelPredictions[,c("obs", "pred_opt")], truth = "obs", estimate = "pred_opt")$.estimate
RMSE_halfse <- yardstick::rmse(fullModOut$modelPredictions[,c("obs", "pred_opt_halfse")], truth = "obs", estimate = "pred_opt_halfse")$.estimate
RMSE_1se <- yardstick::rmse(fullModOut$modelPredictions[,c("obs", "pred_opt_se")], truth = "obs", estimate = "pred_opt_se")$.estimate
# calculate bias of all models
bias_best <-  mean((fullModOut$modelPredictions$obs) - fullModOut$modelPredictions$pred_opt)
bias_halfse <-  mean((fullModOut$modelPredictions$obs) - fullModOut$modelPredictions$pred_opt_halfse)
bias_1se <- mean((fullModOut$modelPredictions$obs) - fullModOut$modelPredictions$pred_opt_se)


uniqueCoeffs <- data.frame("Best lambda model" = c(signif(RMSE_best,3), as.character(signif(bias_best, 3)),
  as.integer(length(globModTerms)-1), as.integer(prednames_fig_num), 
                                                   as.integer(sum(prednames_fig %in% c(prednames_clim))),
                                                   as.integer(sum(prednames_fig %in% c(prednames_weath))),
                                                   as.integer(sum(prednames_fig %in% c(prednames_soils)))
                                                   ), 
                           "1/2 se lambda model" = c(signif(RMSE_halfse,3), as.character(signif(bias_halfse, 3)),
                             length(globModTerms_halfse)-1, prednames_fig_halfse_num,
                                                   sum(prednames_fig_halfse %in% c(prednames_clim)),
                                                   sum(prednames_fig_halfse %in% c(prednames_weath)),
                                                   sum(prednames_fig_halfse %in% c(prednames_soils))), 
                           "1se lambda model" = c(signif(RMSE_1se,3), as.character(signif(bias_1se, 3)),
                             length(globModTerms_1se)-1, prednames_fig_1se_num,
                                                   sum(prednames_fig_1se %in% c(prednames_clim)),
                                                   sum(prednames_fig_1se %in% c(prednames_weath)),
                                                   sum(prednames_fig_1se %in% c(prednames_soils))))
row.names(uniqueCoeffs) <- c("RMSE", "bias: mean(obs-pred.)", "Total number of coefficients", "Number of unique coefficients",
                             "Number of unique climate coefficients", 
                             "Number of unique weather coefficients",  
                             "Number of unique soils coefficients"
                             )

kable(uniqueCoeffs, 
      col.names = c("Best lambda model", "1/2 se lambda model", "1se lambda model"), row.names = TRUE) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed")) 
```

# Visualizations of Model Predictions and Residuals  -- using best lambda model
## observed vs. predicted values

As the alternative to the best lambda model, use the model (1se or 1/2se of best Lambda) that has the fewest number of unique predictors 
```{r}
if (whichSecondBestMod == "auto") {
  # name of model w/ fewest # of predictors (but more than 0)
uniqueCoeff_min <- min(as.numeric(uniqueCoeffs[4,2:3])[which(as.numeric(uniqueCoeffs[4,2:3]) > 0)])
alternativeModel <- names(uniqueCoeffs[4,2:3])[which(uniqueCoeffs[4,2:3] == uniqueCoeff_min)]

if (is.finite(uniqueCoeff_min)) {
  if (length(alternativeModel) == 1) {
  if (alternativeModel == "X1.2.se.lambda.model") {
  mod_secondBest <- fit_glm_halfse
  name_secondBestMod <- "1/2 SE Model"
  prednames_secondBestMod <- prednames_fig_halfse
} else if (alternativeModel == "X1se.lambda.model") {
  mod_secondBest <- fit_glm_se
  name_secondBestMod <- "1 SE Model"
  prednames_secondBestMod <- prednames_fig_1se
}
} else {
  # if both alternative models have the same number of unique coefficients, chose the model that has the fewest number of total predictors
  uniqueCoeff_min2 <- min(as.numeric(uniqueCoeffs[3,alternativeModel]))
alternativeModel2 <- names(uniqueCoeffs[3,alternativeModel])[which(uniqueCoeffs[3,alternativeModel] == uniqueCoeff_min2)]
# if both models still have the same number of unique coefficeints, chose the 1se lambda model
if (length(alternativeModel2 == 2)) {
  alternativeModel2 <- "X1se.lambda.model"
}
if (alternativeModel2 == "X1.2.se.lambda.model") {
  mod_secondBest <- fit_glm_halfse
  name_secondBestMod <- "1/2 SE Model"
  prednames_secondBestMod <- prednames_fig_halfse
} else if (alternativeModel2 == "X1se.lambda.model") {
  mod_secondBest <- fit_glm_se
  name_secondBestMod <- "1 SE Model"
  prednames_secondBestMod <- prednames_fig_1se
}
}
  }else {
    mod_secondBest <- NULL
  name_secondBestMod <- "Intercept_Only"
  prednames_secondBestMod <- NULL
}
} else if (whichSecondBestMod == "1se") {
  mod_secondBest <- fit_glm_se
  name_secondBestMod <- "1 SE Model"
  prednames_secondBestMod <- prednames_fig_1se
} else if (whichSecondBestMod == "halfse") {
  mod_secondBest <- fit_glm_halfse
  name_secondBestMod <- "1/2 SE Model"
  prednames_secondBestMod <- prednames_fig_halfse
}
```

### Predicting on the data

```{r predict with best lambda on data used for fitting}
  # create prediction for each each model
# (i.e. for each fire proporation variable)
if (response == "TotalTreeCover_transformed") {
  predict_by_response <- function(mod, df) {
  df_out <- df
  response_name <- paste0(response, "_pred")
  preds <- predict(mod, newdata = df_out, #s="lambda.min", 
                                     type = "response")/10
  #preds[preds<0] <- 0
  #preds[preds>100] <- 100
  df_out <- df_out %>% cbind(preds)
   colnames(df_out)[ncol(df_out)] <- response_name
  return(df_out)
}
} else {
  predict_by_response <- function(mod, df) {
  df_out <- df
  response_name <- paste0(response, "_pred")
  preds <- predict(mod, newdata = df_out, #s="lambda.min", 
                                     type = "response")
  #preds[preds<0] <- 0
  #preds[preds>100] <- 100
  df_out <- df_out %>% cbind(preds)
   colnames(df_out)[ncol(df_out)] <- response_name
  return(df_out)
}
}

pred_glm1 <- predict_by_response(fit_glm_bestLambda, newdata_DF#X[,2:ncol(X)]
                                 )
# add back in true y values
pred_glm1 <- pred_glm1 %>% 
  cbind(data.frame(response = modDat_1_s[,c(response)])) 
names(pred_glm1)[names(pred_glm1) == "response"] <- response

# add back in lat/long data 
pred_glm1 <- pred_glm1 %>% 
  cbind(modDat_1_s[,c("x", "y", "Year")])

```

```{r predict with second best lambda model on data used for model fitting}
if ( name_secondBestMod == "Intercept_Only") {
  print("The next best lambda model only contains one predictor (an intercept)")
} else {

pred_glm1_1se <- predict_by_response(mod_secondBest, newdata_DF)

# add back in true y values
pred_glm1_1se <- pred_glm1_1se %>% 
  cbind(data.frame(response = modDat_1_s[,c(response)])) 
names(pred_glm1_1se)[names(pred_glm1_1se) == "response"] <- response

# add back in lat/long data 
pred_glm1_1se <- pred_glm1_1se %>% 
  cbind(modDat_1_s[,c("x", "y", "Year")])

}
```

```{r}

## get climate data for mapping
## Get wall-to-wall climate data calculated from dayMet, which we will use for calibration
# read in unscaled climate data
climDat_Temp <- readRDS("../../../Data_processed/WallToWallClimateData/DayMetData_allCONUS_2023ClimateValues_withSoils.rds")
#climDat_Temp <- readRDS("../../../Data_processed/EcoRegion_climSoilData.rds")

# rename
climDat <- climDat_Temp %>% 
  #filter(year == 2016) %>% 
  dplyr::select(year, tmin_meanAnnAvg_CLIM:geometry) %>%
rename("tmin" = tmin_meanAnnAvg_CLIM, 
     "tmax" = tmax_meanAnnAvg_CLIM, #1 
     "tmean" = tmean_meanAnnAvg_CLIM, 
     "prcp" = prcp_meanAnnTotal_CLIM, 
     "t_warm" = T_warmestMonth_meanAnnAvg_CLIM,
     "t_cold" = T_coldestMonth_meanAnnAvg_CLIM, 
     "prcp_wet" = precip_wettestMonth_meanAnnAvg_CLIM,
     "prcp_dry" = precip_driestMonth_meanAnnAvg_CLIM, 
     "prcp_seasonality" = precip_Seasonality_meanAnnAvg_CLIM, #2
     "prcpTempCorr" = PrecipTempCorr_meanAnnAvg_CLIM,  #3
     "abvFreezingMonth" = aboveFreezing_month_meanAnnAvg_CLIM, 
     "isothermality" = isothermality_meanAnnAvg_CLIM, #4
     "annWatDef" = annWaterDeficit_meanAnnAvg_CLIM, 
     "annWetDegDays" = annWetDegDays_meanAnnAvg_CLIM,
     "VPD_mean" = annVPD_mean_meanAnnAvg_CLIM, 
     "VPD_max" = annVPD_max_meanAnnAvg_CLIM, #5
     "VPD_min" = annVPD_min_meanAnnAvg_CLIM, #6
     "VPD_max_95" = annVPD_max_95percentile_CLIM, 
     "annWatDef_95" = annWaterDeficit_95percentile_CLIM, 
     "annWetDegDays_5" = annWetDegDays_5percentile_CLIM, 
     "frostFreeDays_5" = durationFrostFreeDays_5percentile_CLIM, 
     "frostFreeDays" = durationFrostFreeDays_meanAnnAvg_CLIM, 
     "soilDepth" = soilDepth, #7
     "clay" = surfaceClay_perc, 
     "sand" = avgSandPerc_acrossDepth, #8
     "coarse" = avgCoarsePerc_acrossDepth, #9
     "carbon" = avgOrganicCarbonPerc_0_3cm, #10
     "AWHC" = totalAvailableWaterHoldingCapacity,
  ) %>% 
  # since the dayMet wall-to-wall data doesn't have anomalies, add 0s for each anomaly (so predictions are more temporaly stable)
  mutate(
    "tmean_anom" = 0,                           
 "tmin_anom"   = 0            ,
 "tmax_anom"   = 0            , 
 "prcp_anom"   = 0       ,                   
 "t_warm_anom"     = 0        ,
 "t_cold_anom"     = 0        , 
 "prcp_wet_anom"   = 0   ,                   
 "precp_dry_anom"  = 0        ,
 "prcp_seasonality_anom"  = 0 , 
 "prcpTempCorr_anom" = 0 ,                   
 "aboveFreezingMonth_anom" = 0,
 "isothermality_anom" = 0     , 
 "annWatDef_anom"  = 0   ,                   
 "annWetDegDays_anom" = 0     ,
 "VPD_mean_anom"      = 0     , 
 "VPD_min_anom"       = 0,                   
 "VPD_max_anom"       = 0     ,
 "VPD_max_95_anom"    = 0     , 
 "annWatDef_95_anom"  = 0,                   
 "annWetDegDays_5_anom"  = 0  ,
 "frostFreeDays_5_anom"  = 0  , 
 "frostFreeDays_anom" = 0,
  )

## read in scaled data (w/ all predictirs) 
modDat_1_s_temp <- readRDS("./models/scaledModelInputData.rds")

#. now, scale the model predictions according to the scale parameters used in the fitting data 
scaleParams <- modDat_1_s_temp %>% 
  #filter(Year == 2016) %>% 
  dplyr::select(tmin_s:AWHC_s) %>% 
  reframe(across(all_of(names(.)), attributes)) 

# apply the scaling factors to the contemporary climate data 
namesToScale <- climDat %>% 
  st_drop_geometry() %>% 
  dplyr::select(tmin:frostFreeDays, soilDepth:carbon, AWHC, tmean_anom:frostFreeDays_anom) %>% 
  names()
climDat_noSF <- climDat %>% 
  st_drop_geometry()
climDat_scaled <- map(namesToScale, .f = function(x) {
  x_new <- (climDat_noSF[,x] - scaleParams[,paste0(x, "_s")]$`scaled:center`)/scaleParams[,paste0(x, "_s")]$`scaled:scale`
  return(data.frame(x_new))
}) %>% 
  purrr::list_cbind()
names(climDat_scaled) <- paste0(namesToScale, "_s")

# get the Lon and Lat from the climDat geometry (crs is the same as dayMet)
climDatPred <- climDat %>% 
  mutate("Lon" = st_coordinates(.)[,1],
         "Lat" = st_coordinates(.)[,2]) %>% 
  #st_drop_geometry() %>% 
  dplyr::select(Lon, Lat) %>% 
  cbind(climDat_scaled)
names(climDatPred)[3:52] <- str_remove(names(climDatPred)[3:52], pattern = "_s$")
rm(climDat_Temp, climDat_noSF, climDatPred_extract, climDat_scaled, climDat_noSF, climDat)
gc()

# for ecoregion-level models, trim the climDat data frame just to the ecoregion of focus
# predict before scaling/interpolation
climDatPred <- climDatPred %>% 
  mutate(ForestPred_raw =  as.numeric(1/ (1 + 
                exp( - (3.38565522 + -0.79822221*tmean +  4.74536062*prcp + -0.49798680*prcpTempCorr +  0.22745494*isothermality + -0.39051271*clay +  0.53857137*coarse + -0.40059849*AWHC + -0.83406990*I(prcp^2) + -0.94850342*I(prcpTempCorr^2) +  0.19450621*I(isothermality^2) + -0.04772718*I(clay^2) + -0.26519959*I(coarse^2) +  0.07438930*I(AWHC^2) + -0.29049829*prcp*isothermality + -0.16800569*tmean*isothermality +  0.07625545*tmean*prcp + -0.55536190*tmean*prcpTempCorr + -0.37720980*clay*AWHC)    
                )
    )))

# now, make predictions binary based on the threshold determined in the forest model fitting process
# get threshold
best_threshold_bestLambda <- readRDS( "./models/ForestModelPrevalence.rds")

climDatPred <- climDatPred %>% 
  mutate(ForestPred_binary = ForestPred_raw) %>% 
  mutate(ForestPred_binary = replace(ForestPred_binary, ForestPred_raw > best_threshold_bestLambda, 1), 
         ForestPred_binary = replace(ForestPred_binary, ForestPred_raw <= best_threshold_bestLambda, 0)) 

if (ecoregion != "CONUS") {
  ##Classify locations as forested or not with our forest model, using the wall-to-wall climate data we acquired above

### Now, get a raster of locations that the model classifies as unforested

if (ecoregion == "noTrees") {
  # filter to only get locations that are model-predicted to be un-forested
climDatPred_ecoregion <- climDatPred %>% 
  filter(ForestPred_binary == 0)
} else {
  # filter to only get locations that are model-predicted to be forested
  climDatPred_ecoregion <- climDatPred %>% 
  filter(ForestPred_binary == 1)
}

} else {
climDatPred_ecoregion <- climDatPred
  
}

# get dayMet raster
test_rast_dayMet_TEMP <- rast("../../../Data_raw/dayMet/rawMonthlyData/orders/70e0da02b9d2d6e8faa8c97d211f3546/Daymet_Monthly_V4R1/data/daymet_v4_prcp_monttl_na_1980.tif") 



```

Interpolate the model predictions and back-transform the observations, if necessary (only for total tree cover in non-forested locations)
```{r}
if(response == "TotalTreeCover_transformed") {
  # back-transform the response variable
  pred_glm1$TotalTreeCover_transformed <- pred_glm1$TotalTreeCover_transformed/10
  
   # back-transform the response variable
  pred_glm1_1se$TotalTreeCover_transformed <- pred_glm1_1se$TotalTreeCover_transformed/10
}

```

Calculate residuals
```{r}
# best lambda model
pred_glm1$resid  <- pred_glm1[,response] - pred_glm1[,paste0(response, "_pred")]
pred_glm1$extremeResid <- NA
pred_glm1[pred_glm1$resid > 1 | pred_glm1$resid < 1,"extremeResid"] <- 1

# second-best lambda model
if ( name_secondBestMod == "Intercept_Only") {
  print("The next best lambda model only contains one predictor (an intercept)")
} else {

pred_glm1_1se$resid <- pred_glm1_1se[,response] - pred_glm1_1se[,paste0(response, "_pred")]
pred_glm1_1se$extremeResid <- NA
pred_glm1_1se[pred_glm1_1se$resid > 1 | pred_glm1_1se$resid < -1,"extremeResid"] <- 1
}
```

### Maps of Observations, Predictions, and Residuals=

Observations across the temporal range of the dataset
```{r Residual maps5, fig.height = 7, fig.width = 9}
# rasterize
# get reference raster
# use the gridMet raster 
test_rast_dayMet <- rast("../../../Data_raw/dayMet/rawMonthlyData/orders/70e0da02b9d2d6e8faa8c97d211f3546/Daymet_Monthly_V4R1/data/daymet_v4_prcp_monttl_na_1980.tif") %>%
  #terra::project(terra::crs(crs_veg)) %>%
  terra::crop(ext(-2000000, 2600000, -1900000, 1500000))

## aggregate to a 4x4km^2 grid
test_rast <- test_rast_dayMet %>% 
  terra::aggregate(fact = 8, fun = "mean")
# test_rast_Untranfsormed <-  rast("../../../Data_raw/gridMet/pr_2020.nc") #%>% ## mean cell size of unaggregated grid size (16709380m^2)
# test_rast <- test_rast_Untranfsormed %>% 
#    terra::aggregate(fact = 2, fun = "mean") #%>% 
regions <- sf::st_read(dsn = "../../../Data_raw/Level2Ecoregions/", layer = "NA_CEC_Eco_Level2") 
regions <- regions %>% 
  st_transform(crs = st_crs(crs(test_rast))) %>% 
  st_make_valid() 

# make shapefile of cropped state boundaries in appropriate crs
cropped_states_2 <- cropped_states %>% 
  st_transform(crs = crs(test_rast)) %>% 
  st_make_valid()  %>% 
  sf::st_crop(c("xmin" = -2000000, "xmax" = 2600000, "ymin" = -1900000, "ymax" = 1500000))

# make a multi-layer raster w/ layers for each variable of interest
pred_glm1_sf <- pred_glm1 %>% 
  st_as_sf(coords = c("x", "y"), crs = crs(crs_veg)) %>% 
  st_transform(crs(test_rast)) %>% 
  terra::vect()

# rasterize the observations and predictions
testRast2 <- lapply(c(response, paste0(response, "_pred"), "resid"), FUN = function(x) {
  # S4 method for class 'data.frame'
temp <- terra::rasterize(x = pred_glm1_sf, y = test_rast, 
                         field = x, fun = mean, na.rm = TRUE)
names(temp) <- x
  # rast(pred_glm1_full[,c("x", "y", x)] %>% drop_na(), type="xyz", crs=terra::crs(test_rast$`precipitation_amount_day=43829`), digits=6, extent=test_rast$`precipitation_amount_day=43829`)
 return(temp)
  }
)
pred_glm1_RAST <- c(testRast2[[1]], testRast2[[2]], testRast2[[3]])


if(response == "TotalTreeCover_transformed") {
  limits <- c(0,0.1)
} else {
  limits <- c(0,1)
}

# make figure of observations
map_obs <- ggplot() +
geom_spatraster(data = pred_glm1_RAST, aes(fill = .data[[response]]), maxcell = 50000000) + 
 #geom_sf(data = pred_glm1_RAST, aes(col = TotalTreeCover)) + 
  # stat_summary_2d(data = plotObs, aes(x = x, y = y, z = TotalTreeCover), fun = mean, binwidth = .05) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
labs(title = paste0("Observations of ", response, " in the ",ecoregion, " ecoregion")) +
  scale_fill_gradient2(low = "brown",
                       mid = "tan" ,
                       high = "forestgreen" , 
                       midpoint = 0,   na.value = "white", 
                       limits = limits) + 
  coord_sf(datum = st_crs(pred_glm1_RAST)) +
   # xlim(-104, -100) + 
   # ylim(30, 33)
  xlim(-2000000, 2500000) + 
  ylim(-1800000, 1200000)
# histogram of observations
hist_obs <- pred_glm1%>% 
  ggplot() + 
  geom_histogram(aes(.data[[response]]), fill = "lightgrey", col = "darkgrey", bins = 50)

library(ggpubr)
ggarrange( map_obs, hist_obs, heights = c(3, 1), ncol = 1)

```

Predictions across the temporal range of the dataset
```{r predicted values map best lambda model, fig.height = 7, fig.width = 9, }

map_preds1 <- ggplot() +
geom_spatraster(data = pred_glm1_RAST, aes(fill = .data[[paste0(response,"_pred")]]), maxcell = 50000000) +
  #geom_sf(data = highPred_points, col = "red") +
geom_sf(data=cropped_states_2 ,fill=NA ) +
  labs(title = paste0("Predictions from the 'best lambda' fitted model of ", response, " in the ",ecoregion, " ecoregion"),
     subtitle =  "bestLambda model")  +
  # scale_fill_gradient2(low = "brown",
  #                      mid = "wheat" ,
  #                      high = "forestgreen" , 
  #                      midpoint = 0,   na.value = "white", 
  #                      limits = limits) + 
  scale_fill_gradientn(
    colors = c("wheat", "tan", "forestgreen"),
     values = c(0, 1e-10, 1), 
    limits = limits,
    na.value = "white"
  ) +
  xlim(-2000000, 2500000) + 
     ylim(-1800000, 1200000)



hist_preds1 <- ggplot(pred_glm1) + 
  geom_histogram(aes(.data[[paste0(response, "_pred")]]), fill = "lightgrey", col = "darkgrey")#+ 
  #xlim(limits[1]-, limits[2])

ggarrange(map_preds1, hist_preds1, heights = c(3,1), ncol = 1)
```

```{r predicted values map 1se lambda model, fig.height = 7, fig.width = 9, }
if ( name_secondBestMod == "Intercept_Only") {
  print("The next best lambda model only contains one predictor (an intercept)")
} else {

# rasterize data
# make a multi-layer raster w/ layers for each variable of interest
pred_glm1_1se_sf <- pred_glm1_1se %>% 
  st_as_sf(coords = c("x", "y"), crs = crs(crs_veg)) %>% 
  st_transform(crs(test_rast)) %>% 
  terra::vect()

# rasterize the observations and predictions
testRast3 <- lapply(c(response, paste0(response, "_pred"), "resid"), FUN = function(x) {
  # S4 method for class 'data.frame'
temp <- terra::rasterize(x = pred_glm1_1se_sf, y = test_rast, 
                         field = x, fun = mean, na.rm = TRUE)
names(temp) <- x
  # rast(pred_glm1_full[,c("x", "y", x)] %>% drop_na(), type="xyz", crs=terra::crs(test_rast$`precipitation_amount_day=43829`), digits=6, extent=test_rast$`precipitation_amount_day=43829`)
 return(temp)
  }
)
pred_glm1_1se_RAST <- c(testRast3[[1]], testRast3[[2]], testRast3[[3]])

# make figures
map_preds2 <- ggplot() +
geom_spatraster(data = pred_glm1_RAST, aes(fill = .data[[paste0(response,"_pred")]]), maxcell = 50000000) + 
 geom_sf(data=cropped_states_2 ,fill=NA ) +
labs(title = paste0("Predictions from the ", name_secondBestMod, "fitted model of ", response, " in the ",ecoregion, " ecoregion"),
     subtitle =  name_secondBestMod)  +
  # scale_fill_gradient2(low = "wheat",
  #                      mid = "darkgreen",
  #                      high = "red" , 
  #                      midpoint = 1,   na.value = "white",
  #                      limits = c(0, 1)) +
  scale_fill_gradientn(
    colors = c("wheat", "tan", "forestgreen"),
     values = c(0, 1e-10, 1), 
    limits = limits,
    na.value = "white"
  ) +
  xlim(-2000000, 2500000) + 
     ylim(-1800000, 1200000)


hist_preds2 <- ggplot(pred_glm1_1se) + 
  geom_histogram(aes(.data[[paste0(response, "_pred")]]), fill = "lightgrey", col = "darkgrey")#+ 
  #xlim(c(0,1))

ggarrange(map_preds2, hist_preds2, heights = c(3,1), ncol = 1)
}
```
```{r Absolute Residual map from best lambda model, fig.height = 12, fig.width = 9, warning = FALSE }
if (response == "TotalTreeCover_transformed") {
  
limitsResids <- c(-0.1, 0.1)
} else {
  
limitsResids <- c(-1,1)
}
# make figures
map <- ggplot() +
geom_spatraster(data = pred_glm1_RAST, aes(fill = resid), maxcell = 50000000) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
  #geom_sf(data = badResids_high, col = "blue") +
  #geom_sf(data = badResids_low, col = "red") +
labs(title = paste0("Absolute Resids. (obs. - pred.) from the ", ecoregion," ecoregion-wide model of ", response),
     subtitle = "bestLambda model") +
  scale_fill_gradientn(colors = c("red", "orange", "white", "green", "blue"),
                                                                #values = c(-1, -.5, 0, 0.5, 1), 
                                                                breaks = c(-1, -.5, 0, 0.5, 1), 
                                                               limits = c(-1,1),
                                                                na.value = "lightgrey") + 
  xlim(-2000000, 2500000) + 
     ylim(-1800000, 1200000)

hist <- ggplot(pred_glm1) + 
  geom_histogram(aes(resid), fill = "lightgrey", col = "darkgrey") + 
  geom_text(aes(x = min(resid)*.9, y = 1500, label = paste0("min = ", round(min(resid),2)))) +
  geom_text(aes(x = max(resid)*.9, y = 1500, label = paste0("max = ", round(max(resid),2)))) + 
  geom_vline(aes(xintercept = mean(resid)))

ggarrange(map, hist, heights = c(3,1), ncol = 1)

```
```{r Relative Residual map from best lambda model, fig.height = 10, fig.width = 9, warning = FALSE }
# calculate actual residuals
relativeResid <- (pred_glm1_RAST[[response]] -  pred_glm1_RAST[[paste0(response, "_pred")]])/pred_glm1_RAST[[paste0(response, "_pred")]]
# calculate relative residuals
names(relativeResid) <- "relResid"
# make figures
map <- ggplot() +
geom_spatraster(data = relativeResid, aes(fill = relResid), maxcell = 50000000) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
  #geom_sf(data = badResids_high, col = "blue") +
  #geom_sf(data = badResids_low, col = "red") +
labs(title = paste0("Relative Resids. (obs. - pred.)/(obs.) from the ", ecoregion," ecoregion-wide model of ", response),
     subtitle = "bestLambda model") +
  scale_fill_gradient2(low = "red",
                       mid = "white" ,
                       high = "blue" ,
                       midpoint = 0,   na.value = "grey30"#,
                       #limits = limitsResids
                       ) +
  xlim(-2000000, 2500000) + 
     ylim(-1800000, 1200000)

hist <- ggplot(data.frame(values(relativeResid))) + 
  geom_histogram(aes(relResid), fill = "lightgrey", col = "darkgrey") + 
  geom_text(aes(x = min(relResid)*.9, y = 1500, label = paste0("min = ", round(min(relResid),2)))) +
  geom_text(aes(x = max(relResid)*.9, y = 1500, label = paste0("max = ", round(max(relResid),2)))) + 
  geom_vline(aes(xintercept = mean(relResid)))

ggarrange(map, hist, heights = c(3,1), ncol = 1)

```

```{r Absolute Residual map from 1 SE lambda model, fig.height = 10, fig.width = 9, warning = FALSE }
if ( name_secondBestMod == "Intercept_Only") {
  print("The next best lambda model only contains one predictor (an intercept)")
} else {

# make figures
map <- ggplot() +
geom_spatraster(data = pred_glm1_1se_RAST, aes(fill = resid), maxcell = 50000000) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
labs(title = paste0("Resids. (obs. - pred.) from the ", ecoregion," ecoregion-wide model of ", response),
     subtitle = "second-bestLambda model") +
  scale_fill_gradientn(colors = c("red", "orange", "white", "green", "blue"),
                                                                breaks = c(-1, -.5, 0, 0.5, 1), 
                                                               limits = c(-1,1),
                                                                na.value = "lightgrey") + 
  xlim(-2000000, 2500000) + 
     ylim(-1800000, 1200000)

hist <- ggplot(pred_glm1_1se) + 
  geom_histogram(aes(resid), fill = "lightgrey", col = "darkgrey") + 
  geom_text(aes(x = min(resid)*.9, y = 1500, label = paste0("min = ", round(min(resid),2)))) +
  geom_text(aes(x = max(resid)*.9, y = 1500, label = paste0("max = ", round(max(resid),2)))) + 
  geom_vline(aes(xintercept = mean(resid, na.rm = TRUE)))

ggarrange(map, hist, heights = c(3,1), ncol = 1)

}
```

```{r Relative Residual map from second-best lambda model, fig.height = 12, fig.width = 9, warning = FALSE }
# calculate actual residuals
relativeResid <- (pred_glm1_1se_RAST[[response]] -  pred_glm1_1se_RAST[[paste0(response, "_pred")]])/pred_glm1_1se_RAST[[paste0(response, "_pred")]]
names(relativeResid) <- "relResid"
# calculate relative residuals

# make figures
map <- ggplot() +
geom_spatraster(data = relativeResid, aes(fill = relResid), maxcell = 50000000) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
  #geom_sf(data = badResids_high, col = "blue") +
  #geom_sf(data = badResids_low, col = "red") +
labs(title = paste0("Relative Resids. (obs. - pred.)/(obs.) from the ", ecoregion," ecoregion-wide model of ", response),
     subtitle = "second-bestLambda model") +
  scale_fill_gradient2(low = "red",
                       mid = "white" ,
                       high = "blue" , 
                       midpoint = 0,   na.value = "grey30"#,
                       #limits = limitsResids
                       ) +
  xlim(-2000000, 2500000) + 
     ylim(-1800000, 1200000)

hist <- ggplot(data.frame(values(relativeResid))) + 
  geom_histogram(aes(relResid), fill = "lightgrey", col = "darkgrey") + 
  geom_text(aes(x = min(relResid)*.9, y = 1500, label = paste0("min = ", round(min(relResid),2)))) +
  geom_text(aes(x = max(relResid)*.9, y = 1500, label = paste0("max = ", round(max(relResid),2)))) + 
  geom_vline(aes(xintercept = mean(relResid)))

ggarrange(map, hist, heights = c(3,1), ncol = 1)
```

### Now we show model predictions using data on the dayMet scale (1km-1km grid) across the focal ecoregion
```{r w2w preds with the bestLambda model}
# the data.frame to use for these predictions is "climDatPred_ecoregion"
climDatPred_ecoregion_glm1 <- predict_by_response(mod = fit_glm_bestLambda, df = climDatPred_ecoregion %>% st_drop_geometry())


climDatPred_ecoregion_glm1_rast <- climDatPred_ecoregion_glm1 %>%
  st_as_sf(coords = c("Lon", "Lat"), crs = crs("PROJCRS[\"unnamed\",\n    BASEGEOGCRS[\"unknown\",\n        DATUM[\"unknown\",\n            ELLIPSOID[\"Spheroid\",6378137,298.257223563,\n                LENGTHUNIT[\"metre\",1,\n                    ID[\"EPSG\",9001]]]],\n        PRIMEM[\"Greenwich\",0,\n            ANGLEUNIT[\"degree\",0.0174532925199433,\n                ID[\"EPSG\",9122]]]],\n    CONVERSION[\"Lambert Conic Conformal (2SP)\",\n        METHOD[\"Lambert Conic Conformal (2SP)\",\n            ID[\"EPSG\",9802]],\n        PARAMETER[\"Latitude of false origin\",42.5,\n            ANGLEUNIT[\"degree\",0.0174532925199433],\n            ID[\"EPSG\",8821]],\n        PARAMETER[\"Longitude of false origin\",-100,\n            ANGLEUNIT[\"degree\",0.0174532925199433],\n            ID[\"EPSG\",8822]],\n        PARAMETER[\"Latitude of 1st standard parallel\",25,\n            ANGLEUNIT[\"degree\",0.0174532925199433],\n            ID[\"EPSG\",8823]],\n        PARAMETER[\"Latitude of 2nd standard parallel\",60,\n            ANGLEUNIT[\"degree\",0.0174532925199433],\n            ID[\"EPSG\",8824]],\n        PARAMETER[\"Easting at false origin\",0,\n            LENGTHUNIT[\"metre\",1],\n            ID[\"EPSG\",8826]],\n        PARAMETER[\"Northing at false origin\",0,\n            LENGTHUNIT[\"metre\",1],\n            ID[\"EPSG\",8827]]],\n    CS[Cartesian,2],\n        AXIS[\"easting\",east,\n            ORDER[1],\n            LENGTHUNIT[\"metre\",1,\n                ID[\"EPSG\",9001]]],\n        AXIS[\"northing\",north,\n            ORDER[2],\n            LENGTHUNIT[\"metre\",1,\n                ID[\"EPSG\",9001]]]]")) %>%
  terra::vect() %>%
  terra::rasterize(y = test_rast_dayMet_TEMP, field = paste0(response, "_pred"))

# make a map
map_obs <- ggplot() +
geom_spatraster(data = climDatPred_ecoregion_glm1_rast, aes(fill = last), maxcell = 50000000) + 
 #geom_sf(data = pred_glm1_RAST, aes(col = TotalTreeCover)) + 
  # stat_summary_2d(data = plotObs, aes(x = x, y = y, z = TotalTreeCover), fun = mean, binwidth = .05) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
labs(title = paste0("Model-predictions of ", response, " in the ",ecoregion, " ecoregion, on the dayMet scale"),
     subtitle = "Best-lambda model") +
  scale_fill_gradientn(
    colors = c("wheat", "tan", "forestgreen"),
     values = c(0, 1e-10, 1), 
    limits = limits,
    na.value = "white"
  ) +
   # xlim(-104, -100) + 
   # ylim(30, 33)
  xlim(-2000000, 2500000) + 
  ylim(-1800000, 1200000)

# histogram of observations
hist_obs <- climDatPred_ecoregion_glm1%>% 
  ggplot() + 
  geom_histogram(aes(.data[[ paste0(response, "_pred")]]), fill = "lightgrey", col = "darkgrey", bins = 50)

ggarrange( map_obs, hist_obs, heights = c(3, 1), ncol = 1)
```

```{r w2w preds with the second-bestLambda model}
# the data.frame to use for these predictions is "climDatPred_ecoregion"
climDatPred_ecoregion_glm1_1se <- predict_by_response(mod = mod_secondBest, df = climDatPred_ecoregion %>% st_drop_geometry())


# now, rasterize the predictions
climDatPred_ecoregion_glm1_1se_rast <- climDatPred_ecoregion_glm1_1se %>%
  st_as_sf(coords = c("Lon", "Lat"), crs = crs("PROJCRS[\"unnamed\",\n    BASEGEOGCRS[\"unknown\",\n        DATUM[\"unknown\",\n            ELLIPSOID[\"Spheroid\",6378137,298.257223563,\n                LENGTHUNIT[\"metre\",1,\n                    ID[\"EPSG\",9001]]]],\n        PRIMEM[\"Greenwich\",0,\n            ANGLEUNIT[\"degree\",0.0174532925199433,\n                ID[\"EPSG\",9122]]]],\n    CONVERSION[\"Lambert Conic Conformal (2SP)\",\n        METHOD[\"Lambert Conic Conformal (2SP)\",\n            ID[\"EPSG\",9802]],\n        PARAMETER[\"Latitude of false origin\",42.5,\n            ANGLEUNIT[\"degree\",0.0174532925199433],\n            ID[\"EPSG\",8821]],\n        PARAMETER[\"Longitude of false origin\",-100,\n            ANGLEUNIT[\"degree\",0.0174532925199433],\n            ID[\"EPSG\",8822]],\n        PARAMETER[\"Latitude of 1st standard parallel\",25,\n            ANGLEUNIT[\"degree\",0.0174532925199433],\n            ID[\"EPSG\",8823]],\n        PARAMETER[\"Latitude of 2nd standard parallel\",60,\n            ANGLEUNIT[\"degree\",0.0174532925199433],\n            ID[\"EPSG\",8824]],\n        PARAMETER[\"Easting at false origin\",0,\n            LENGTHUNIT[\"metre\",1],\n            ID[\"EPSG\",8826]],\n        PARAMETER[\"Northing at false origin\",0,\n            LENGTHUNIT[\"metre\",1],\n            ID[\"EPSG\",8827]]],\n    CS[Cartesian,2],\n        AXIS[\"easting\",east,\n            ORDER[1],\n            LENGTHUNIT[\"metre\",1,\n                ID[\"EPSG\",9001]]],\n        AXIS[\"northing\",north,\n            ORDER[2],\n            LENGTHUNIT[\"metre\",1,\n                ID[\"EPSG\",9001]]]]")) %>%
  terra::vect() %>%
  terra::rasterize(y = test_rast_dayMet_TEMP, field = paste0(response, "_pred"))

# make a map
map_obs <- ggplot() +
geom_spatraster(data = climDatPred_ecoregion_glm1_1se_rast, aes(fill = last), maxcell = 50000000) + 
 #geom_sf(data = pred_glm1_RAST, aes(col = TotalTreeCover)) + 
  # stat_summary_2d(data = plotObs, aes(x = x, y = y, z = TotalTreeCover), fun = mean, binwidth = .05) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
labs(title = paste0("Model-predictions of ", response, " in the ",ecoregion, " ecoregion, on the dayMet scale"),
     subtitle = "Best-lambda model") +
  scale_fill_gradientn(
    colors = c("wheat", "tan", "forestgreen"),
     values = c(0, 1e-10, 1), 
    limits = limits,
    na.value = "white"
  ) +
   # xlim(-104, -100) + 
   # ylim(30, 33)
  xlim(-2000000, 2500000) + 
  ylim(-1800000, 1200000)

# histogram of observations
hist_obs <- climDatPred_ecoregion_glm1_1se%>% 
  ggplot() + 
  geom_histogram(aes(.data[[ paste0(response, "_pred")]]), fill = "lightgrey", col = "darkgrey", bins = 50)

ggarrange( map_obs, hist_obs, heights = c(3, 1), ncol = 1)
```


### Are there biases of the model predictions across year/lat/long? 
```{r bias figures best lambda model, fig.align=8, fig.height=10, warning=FALSE}
rm(map_obs, hist_obs, climDatPred_ecoregion_glm1, climDatPred_ecoregion_glm1_1se, climDatPred_ecoregion_glm1_1se_rast, climDatPred_ecoregion_glm1_1se_rast, climDatPred)
gc()

if ( name_secondBestMod == "Intercept_Only") {
  print("The next best lambda model only contains one predictor (an intercept)")
  # plot residuals against Year
yearResidMod_bestLambda <- ggplot(pred_glm1) + 
  geom_point(aes(x = jitter(Year), y = resid), alpha = .1) + 
  geom_smooth(aes(x = Year, y = resid)) + 
  xlab("Year") + 
  ylab("Residuals") +
  ggtitle("from best lamba model")

# plot residuals against Lat
latResidMod_bestLambda <- ggplot(pred_glm1) + 
  geom_point(aes(x = y, y = resid), alpha = .1) + 
  geom_smooth(aes(x = y, y = resid)) + 
  xlab("Latitude") + 
  ylab("Residuals") +
  ggtitle("from best lamba model")

# plot residuals against Long
longResidMod_bestLambda <- ggplot(pred_glm1) + 
  geom_point(aes(x = x, y = resid), alpha = .1) + 
  geom_smooth(aes(x = x, y = resid)) + 
  xlab("Longitude") + 
  ylab("Residuals") +
  ggtitle("from best lamba model")
library(patchwork)
(yearResidMod_bestLambda ) / 
(  latResidMod_bestLambda ) /
(  longResidMod_bestLambda )
} else {

# plot residuals against Year
yearResidMod_bestLambda <- ggplot(pred_glm1) + 
  geom_point(aes(x = jitter(Year), y = resid), alpha = .1) + 
  geom_smooth(aes(x = Year, y = resid)) + 
  xlab("Year") + 
  ylab("Residuals") +
  ggtitle("from best lamba model")
yearResidMod_1seLambda <- ggplot(pred_glm1_1se) + 
  geom_point(aes(x = jitter(Year), y = resid), alpha = .1) + 
  geom_smooth(aes(x = Year, y = resid)) + 
  xlab("Year") + 
  ylab("Residuals") +
  ggtitle(paste0("from ", name_secondBestMod))

# plot residuals against Lat
latResidMod_bestLambda <- ggplot(pred_glm1) + 
  geom_point(aes(x = y, y = resid), alpha = .1) + 
  geom_smooth(aes(x = y, y = resid)) + 
  xlab("Latitude") + 
  ylab("Residuals") +
  ggtitle("from best lamba model")
latResidMod_1seLambda <- ggplot(pred_glm1_1se) + 
  geom_point(aes(x = y, y = resid), alpha = .1) + 
  geom_smooth(aes(x = y, y = resid)) + 
  xlab("Latitude") + 
  ylab("Residuals") +
  ggtitle(paste0("from ", name_secondBestMod))

# plot residuals against Long
longResidMod_bestLambda <- ggplot(pred_glm1) + 
  geom_point(aes(x = x, y = resid), alpha = .1) + 
  geom_smooth(aes(x = x, y = resid)) + 
  xlab("Longitude") + 
  ylab("Residuals") +
  ggtitle("from best lamba model")
longResidMod_1seLambda <- ggplot(pred_glm1_1se) + 
  geom_point(aes(x = x, y = resid), alpha = .1) + 
  geom_smooth(aes(x = x, y = resid)) + 
  xlab("Longitude") + 
  ylab("Residuals") +
  ggtitle(paste0("from ", name_secondBestMod))

library(patchwork)
(yearResidMod_bestLambda + yearResidMod_1seLambda) / 
(  latResidMod_bestLambda + latResidMod_1seLambda) /
(  longResidMod_bestLambda + longResidMod_1seLambda)
}
```

### Partial Dependence Plots
For best lambda plot
```{r}
## make data frame for training dataset
# figure out which of the columns has an interaction with a name in the incorrect order and fix
badInteractionNames <- (names(coefficients(fit_glm_bestLambda)[-1]))[!(names(coefficients(fit_glm_bestLambda)[-1]) %in% names(X_df))]
if (length(badInteractionNames) > 0) {
  for (i in 1:length(badInteractionNames)) {
    intTerms <- badInteractionNames[i] %>% 
      str_split(":", simplify = TRUE)
    newInteraction <- paste0(intTerms[,2], ":", intTerms[,1])
    names(X_df)[names(X_df)==newInteraction] <- badInteractionNames[i]
  }
}
for (i in 1:length(names(coefficients(fit_glm_bestLambda)[-1]))) {
  pd_bestLambda_i <- pdp::partial(object = fit_glm_bestLambda, pred.var = names(coefficients(fit_glm_bestLambda)[-1])[i], train = X_df, type = "regression")
  names(pd_bestLambda_i)[1] <- "predictorVariable_value"
  pd_bestLambda_i$predictorVariable_Name <- prednames_fig[i]
  if(i == 1) {
    pd_bestLambda <- pd_bestLambda_i
  } else {
    pd_bestLambda <- rbind(pd_bestLambda, pd_bestLambda_i)
  }
}
 
## update predictor names to be more interpretable
lookup_md <- c(
    "AWHC" = "Available Soil \nWater Holding Capacity",
    "VPD_mean_anom" = "VPD mean \n- 3 yr. anomaly", 
    "VPD_max" = "VPD max - 30yr. avg.",
    "VPD_max_95_anom" = "VPD max 95th percentile \n- 30yr. avg.",
    "VPD_min_anom" = "VPD min - 3 yr. anomaly",
    "aboveFreezingMonth_anom" = "First Above Freezing Month \n- 3 yr. anomaly",
    "annWatDef" = "Ann. Water Deficit \n- 30yr. avg.", 
    "annWatDef_anom" = "Ann. Water Deficit \n- 3 yr. anomaly", 
    "annWetDegDays_anom" = "Annual Wet Degree Days \n- 3 yr. anomaly", 
    "annWetDegDays" = "Annual Wet Degree Days \n- 30yr. avg.",
    "coarse" = "% coarse", 
    "clay" = "% clay", 
    "frostFreeDays_5_anom" = "No. Frost-Free Days 5th \nPercentile - 3 yr. anomaly",
    "isothermality" = "Isothermality - 30yr. avg.",
    "isothermality_anom" = "Isothermality - 3 yr. anomaly", 
    "prcp" = "MAP - 30yr. avg.",
    "prcp_dry" = "Precip. of Driest Month \n- 30yr. avg.",
    "prcpTempCorr" = "Precip. Temp. Correlation \n- 30yr. avg.",
    "prcpTempCorr_anom" = "Precip. Temp. Correlation \n- 3 yr. anomaly",
    "prcp_seasonality" = "Precip. Seasonality \n- 30yr. avg.", 
    "prcp_seasonality_anom" = "Precip. Seasonality\n - 3 yr. anomaly",
    "prcp_wet_anom" = "Precip. of Wettest Month \n- 3 yr. anomaly", 
    "precp_dry_anom" = "Precip. of Driest Month \n- 3 yr. anomaly", 
    "prcp_anom" = "MAP - 3 yr. anomaly",
    "sand" = "% sand", 
    "carbon" = "% carbon in soil surface", 
    "t_cold_anom" = "Temp. of Coldest Month \n- 3 yr. anomaly",
    "t_warm_anom" = "Temp. of Warmest Month \n- 3 yr. anomaly",
    "tmax_anom" = "Ann. Maximum Temp. \n- 3 yr. anomaly", 
    "tmean" = "MAT - 30yr. avg.",
    "tmin_anom" = "Ann. Minimum Temp. \n- 3 yr. anomaly",
    "prcp_wet" = "Precip. of Wettest Month \n- 30yr. avg.",
    "t_warm" = "Temp. of Warmest Month \n- 30yr. avg.",
    "soilDepth" = "Soil Depth", 
    "t_cold" = "Temp. of Coldest Month \n- 30yr. avg.",
    "annWatDef_95" = "Ann. Water Deficit 95th \nPercentile - 30yr. avg."
  ) %>% 
  data.frame()
names(lookup_md) <- c("longName")
lookup_md$predName <- rownames(lookup_md)

pd_bestLambda <- pd_bestLambda %>% 
  left_join(lookup_md, by = c("predictorVariable_Name" = "predName"))

ggplot(pd_bestLambda) + 
  facet_wrap(~longName,scales = "free_x") + 
  geom_line(aes(x = predictorVariable_value, y = yhat)) + 
  labs(title = "Partial Dependence Plot - predictions from best lambda model",
          x = "Predictor Variable Value", 
          y = paste0("Marginal effect of predictor on ", response,""))
```

For Second-best lambda plot
```{r}
if ( name_secondBestMod == "Intercept_Only") {
  print("The next best lambda model only contains one predictor (an intercept), so partial dependence plots are impractical")
  } else {
      
    # figure out which of the columns has an interaction with a name in the incorrect order and fix
badInteractionNames <- (names(coefficients(mod_secondBest)[-1]))[!(names(coefficients(mod_secondBest)[-1]) %in% names(X_df))]
if (length(badInteractionNames) > 0) {
  for (i in 1:length(badInteractionNames)) {
    intTerms <- badInteractionNames[i] %>% 
      str_split(":", simplify = TRUE)
    newInteraction <- paste0(intTerms[,2], ":", intTerms[,1])
    names(X_df)[names(X_df)==newInteraction] <- badInteractionNames[i]
  }
}
for (i in 1:length(prednames_secondBestMod)) {
  pd_secondBestLambda_i <- pdp::partial(object = mod_secondBest, pred.var = prednames_secondBestMod[i], train = X_df, type = "regression")
  names(pd_secondBestLambda_i)[1] <- "predictorVariable_value"
  pd_secondBestLambda_i$predictorVariable_Name <- prednames_secondBestMod[i]
  if(i == 1) {
    pd_secondBestLambda <- pd_secondBestLambda_i
  } else {
    pd_secondBestLambda <- rbind(pd_secondBestLambda, pd_secondBestLambda_i)
  }
}
 
## update predictor names to be more interpretable
lookup_md <- c(
    "AWHC" = "Available Soil \nWater Holding Capacity",
    "VPD_mean_anom" = "VPD mean \n- 3 yr. anomaly", 
    "VPD_max" = "VPD max - 30yr. avg.",
    "VPD_max_95_anom" = "VPD max 95th percentile \n- 30yr. avg.",
    "VPD_min_anom" = "VPD min - 3 yr. anomaly",
    "aboveFreezingMonth_anom" = "First Above Freezing Month \n- 3 yr. anomaly",
    "annWatDef" = "Ann. Water Deficit \n- 30yr. avg.", 
    "annWatDef_anom" = "Ann. Water Deficit \n- 3 yr. anomaly", 
    "annWetDegDays_anom" = "Annual Wet Degree Days \n- 3 yr. anomaly", 
    "annWetDegDays" = "Annual Wet Degree Days \n- 30yr. avg.",
    "coarse" = "% coarse", 
    "clay" = "% clay", 
    "frostFreeDays_5_anom" = "No. Frost-Free Days 5th \nPercentile - 3 yr. anomaly",
    "isothermality" = "Isothermality - 30yr. avg.",
    "isothermality_anom" = "Isothermality - 3 yr. anomaly", 
    "prcp" = "MAP - 30yr. avg.",
    "prcp_dry" = "Precip. of Driest Month \n- 30yr. avg.",
    "prcpTempCorr" = "Precip. Temp. Correlation \n- 30yr. avg.",
    "prcpTempCorr_anom" = "Precip. Temp. Correlation \n- 3 yr. anomaly",
    "prcp_seasonality" = "Precip. Seasonality \n- 30yr. avg.", 
    "prcp_seasonality_anom" = "Precip. Seasonality\n - 3 yr. anomaly",
    "prcp_wet_anom" = "Precip. of Wettest Month \n- 3 yr. anomaly", 
    "precp_dry_anom" = "Precip. of Driest Month \n- 3 yr. anomaly", 
    "prcp_anom" = "MAP - 3 yr. anomaly",
    "sand" = "% sand", 
    "carbon" = "% carbon in soil surface", 
    "t_cold_anom" = "Temp. of Coldest Month \n- 3 yr. anomaly",
    "t_warm_anom" = "Temp. of Warmest Month \n- 3 yr. anomaly",
    "tmax_anom" = "Ann. Maximum Temp. \n- 3 yr. anomaly", 
    "tmean" = "MAT - 30yr. avg.",
    "tmin_anom" = "Ann. Minimum Temp. \n- 3 yr. anomaly",
    "prcp_wet" = "Precip. of Wettest Month \n- 30yr. avg.",
    "t_warm" = "Temp. of Warmest Month \n- 30yr. avg.",
    "soilDepth" = "Soil Depth", 
    "t_cold" = "Temp. of Coldest Month \n- 30yr. avg.",
    "annWatDef_95" = "Ann. Water Deficit 95th \nPercentile - 30yr. avg."
  ) %>% 
  data.frame()
names(lookup_md) <- c("longName")
lookup_md$predName <- rownames(lookup_md)

pd_secondBestLambda <- pd_secondBestLambda %>% 
  left_join(lookup_md, by = c("predictorVariable_Name" = "predName"))

ggplot(pd_secondBestLambda) + 
  facet_wrap(~longName,scales = "free_x") + 
  geom_line(aes(x = predictorVariable_value, y = yhat)) + 
  labs(title = paste0("Partial Dependence Plot - predictions from", name_secondBestMod),
          x = "Predictor Variable Value", 
          y = paste0("P(", response,")"))
}
```


### Quantile plots

Binning predictor variables into "Quantiles"  and looking at the mean
predicted probability for each percentile.
```{r}
# get deciles for best lambda model 
if (length(prednames_fig) == 0) {
  print("The best lambda model only contains one predictor (an intercept), so decile plots aren't possible to generate")
} else {
  pred_glm1_deciles <- predvars2deciles(df = pred_glm1,
                                      response_vars = response_vars,
                                        pred_vars = prednames_fig, 
                                       cut_points = seq(0, 1, 0.01))
}
# get deciles for 1 SE lambda model 
if ( name_secondBestMod == "Intercept_Only") {
  print("The next best lambda model only contains one predictor (an intercept)")} else {
  pred_glm1_deciles_1se <- predvars2deciles(pred_glm1_1se,
                                      response_vars = response_vars,
                                        pred_vars = prednames_secondBestMod, 
                                       cut_points = seq(0, 1, 0.01))
  }
```

Below are quantile plots for the *best lambda model* 
(note that the predictor variables are scaled) 
```{r Percentile_plot best lambda, fig.height= 16, fig.width = 12, warning=FALSE, message=FALSE}
if (length(prednames_fig) == 0) {
  print("The model only contains one predictor (an intercept), so decile plots aren't possible to generate")
} else {

# publication quality version
g3 <- decile_dotplot_pq(df = pred_glm1_deciles, response = c(response, paste0(response, "_pred")), IQR = FALSE,
                        CI = TRUE
                        ) + ggtitle("Quantile Plot")

g4 <- add_dotplot_inset(g3, df = pred_glm1_deciles, response = c(response, paste0(response, "_pred")), dfRaw = pred_glm1, add_smooth = TRUE, deciles = FALSE)

  # make a density plot 
 g4 <- ggarrange(g4, ggplot(pred_glm1) + 
  geom_density(aes(.data[[response]])) + 
  geom_density(aes(.data[[paste0(response,"_pred")]]), col = "lightblue") + 
    ggtitle("Density of observed values (black) and predicted values (blue)"),
  ncol = 1, heights = c(.75, .25)
 )

# ggplot(pred_glm1) + 
#   geom_point(aes(y =.data[[response]], 
#                   x = .data[[paste0(response,"_pred")]])) +
# geom_abline(aes(slope = 1, intercept = 0), col = "red") + 
#   geom_smooth(aes(y =.data[[response]], 
#                   x = .data[[paste0(response,"_pred")]]), col = "grey")
  
if(save_figs) {
  # png(paste0("../../../Figures/CoverDatFigures/ figures/quantile_plots/quantile_plot_", response,  "_",ecoregion,".png"), 
  #    units = "in", res = 600, width = 5.5, height = 3.5 )
  #   print(g4)
  # dev.off()
}

g4
}

```

Below are percentile plots from the second best lambda model (`r print(name_secondBestMod)`)
```{r Percentile_plot 1 SE lambda, fig.height= 16, fig.width = 12, warning=FALSE, message=FALSE}

if ( name_secondBestMod == "Intercept_Only") {
  print("The next best lambda model only contains one predictor (an intercept)")
  } else {

# publication quality version
g3 <- decile_dotplot_pq(pred_glm1_deciles_1se, response = c(response, paste0(response, "_pred")), IQR = TRUE) + ggtitle("Quantile Plot")

g4 <- add_dotplot_inset(g3, df = pred_glm1_deciles_1se, response = c(response, paste0(response, "_pred")), dfRaw = pred_glm1_1se, add_smooth = TRUE, deciles = FALSE)

    # make a density plot 
 g4 <- ggarrange(g4, ggplot(pred_glm1_1se) + 
  geom_density(aes(.data[[response]])) + 
  geom_density(aes(.data[[paste0(response,"_pred")]]), col = "lightblue") + 
    ggtitle("Density of observed values (black) and predicted values (blue)"),
  ncol = 1, heights = c(.75, .25)
 )
if(save_figs) {
  # png(paste0("figures/quantile_plots/quantile_plot_", response,  "_",ecoregion,".png"), 
  #    units = "in", res = 600, width = 5.5, height = 3.5 )
  #   print(g4)
  # dev.off()
}

g4
}

```

### Filtered Quantiles
#### For the best lambda model

Filtered 'Quantile' plots of data. These plots show each vegetation variable,
but only based on data that falls into the upper and lower 20th percentiles of
each predictor variable. 

```{r glm_deciles_filtered best lambda model, fig.height = 24, fig.width = 24, message = FALSE}
if (length(prednames_fig) == 0) {
  print("The model only contains one predictor (an intercept), so quantile plots aren't possible to generate")
} else {
pred_glm1_deciles_filt <- predvars2deciles( pred_glm1, 
                         response_vars = c(response, paste0(response, "_pred")),
                         pred_vars = prednames_fig,
                         filter_var = TRUE,
                         filter_vars = prednames_fig,
                         cut_points = seq(0, 1, 0.01)) 
g <- decile_dotplot_filtered_pq(df = pred_glm1_deciles_filt, response = response, 
                                xvars = prednames_fig)

if(save_figs) {
# jpeg(paste0("figures/quantile_plots/quantile_plot_filtered_mid_v1", , ".jpeg"),
#      units = "in", res = 600, width = 5.5, height = 6 )
#   g 
# dev.off()
}
g
}
```

Filtered quantile figure with middle 2 deciles also shown

```{r filtered quantile plot for best lambda modle with middle 2 deciles, fig.height = 24, fig.width = 24, eval = TRUE, message = FALSE}
if ( name_secondBestMod == "Intercept_Only") {
  print("The next best lambda model only contains one predictor (an intercept)")
} else {
pred_glm1_deciles_filt_mid <- predvars2deciles(pred_glm1, 
                         response_vars = c(response, paste0(response, "_pred")),
                         pred_vars = prednames_fig,
                         filter_vars = prednames_fig,
                         filter_var = TRUE,
                         add_mid = TRUE,
                         cut_points = seq(0, 1, 0.01))

g <- decile_dotplot_filtered_pq(df = pred_glm1_deciles_filt_mid, response = response, 
                                xvars = prednames_fig)

if(save_figs) {
# jpeg(paste0("figures/quantile_plots/quantile_plot_filtered_mid_v1", , ".jpeg"),
#      units = "in", res = 600, width = 5.5, height = 6)
#   g 
# dev.off()
}
g
}
```

#### For the second best lambda model (`r print(name_secondBestMod)`)

Filtered 'Quantile' plots of data. These plots show each vegetation variable,
but only based on data that falls into the upper and lower 20th percentiles of
each predictor variable. 

```{r glm_deciles_filtered second best lambda model, fig.height = 24, fig.width = 24, message = FALSE}
if (length(prednames_secondBestMod) == 0) {
  print("The model only contains one predictor (an intercept), so quantile plots aren't possible to generate")
} else {
  
pred_glm1_deciles_filt <- predvars2deciles( pred_glm1_1se, 
                         response_vars = c(response, paste0(response, "_pred")),
                         pred_vars = prednames_secondBestMod,
                         filter_var = TRUE,
                         filter_vars = prednames_secondBestMod,
                         cut_points = seq(0, 1, 0.01)) 
g <- decile_dotplot_filtered_pq(df = pred_glm1_deciles_filt, response = response, 
                                xvars = prednames_fig)

if(save_figs) {
# jpeg(paste0("figures/quantile_plots/quantile_plot_filtered_mid_v1", , ".jpeg"),
#      units = "in", res = 600, width = 5.5, height = 6 )
#   g 
# dev.off()
}
g
}
```

Filtered quantile figure with middle 2 deciles also shown

```{r filtered quantile plot for second best lambda modle with middle 2 deciles, fig.height = 24, fig.width = 24, eval = TRUE, message = FALSE}
if (length(prednames_secondBestMod) == 0) {
  print("The model only contains one predictor (an intercept), so quantile plots aren't possible to generate")
} else {
pred_glm1_deciles_filt_mid <- predvars2deciles(pred_glm1, 
                         response_vars = c(response, paste0(response, "_pred")),
                         pred_vars = prednames_secondBestMod,
                         filter_vars = prednames_secondBestMod,
                         filter_var = TRUE,
                         add_mid = TRUE,
                         cut_points = seq(0, 1, 0.01))

g <- decile_dotplot_filtered_pq(df = pred_glm1_deciles_filt_mid, response = response, 
                                xvars = prednames_fig)


if(save_figs) {
# jpeg(paste0("figures/quantile_plots/quantile_plot_filtered_mid_v1", , ".jpeg"),
#      units = "in", res = 600, width = 5.5, height = 6 )
#   g 
# dev.off()
}
g
}
```

### Show model RMSE w/in each quantile
```{r}
# get deciles for best lambda model 
if (length(prednames_fig) == 0) {
  print("The best lambda model only contains one predictor (an intercept), so quantile plots aren't possible to generate")
} else {
  pred_glm1_deciles %>% 
    ggplot(aes(x = mean_value, y = RMSE)) +
    facet_wrap(~name, scales = "free_x") +
    geom_point(alpha = .2, size = .5) + 
    geom_smooth(lwd = .5) + 
    xlab("Scaled predictor value") + 
    ggtitle("RMSE by decile for bestLambda model")
}
# get deciles for 1 SE lambda model 
if (length(prednames_secondBestMod) == 0) {
  print("The 1SE (or 1/2 SE) lambda model only contains one predictor (an intercept), so quantile plots aren't possible to generate")
} else {
  pred_glm1_deciles_1se %>% 
    ggplot(aes(x = mean_value, y = RMSE)) +
    facet_wrap(~name, scales = "free_x") +
    geom_point(alpha = .2, size = .5) + 
    geom_smooth(lwd = .5) + 
    xlab("Scaled predictor value") + 
    ggtitle(paste0("RMSE by decile for ", name_secondBestMod, "model"))
}
```

# Model Equations

Below are printed the model equations that predict the response using *scaled* climate, soil, and weather predictors. The scaling factors are provided in the file "./Documentation/CoverModelPredictorInformation.csv"

```{r}
# get the parameters that were used to scale the data
scaleParams <- modDat_1_s_temp%>% 
  #filter(Year == 2016) %>% 
  dplyr::select(tmin_s:AWHC_s) %>% 
  reframe(across(all_of(names(.)), attributes)) 

# function for printing model statement
## function to get model statements
getModelStatement <- function(coefficientTable, # name of the d.f that has model coefficients
                  modelName, # name of the column in the coefficient table that has the parameters of interest
                  responseVar # name of the response variable
                  ) {
  ##
  # coefficientTable <- grassShrub_totalHerb_trimAnoms
  # modelName <- "coefficientValue_bestLambda"
  # responseVar <- "TotalHerbaceousCover"
  ##
  temp <- coefficientTable[,c("coefficientName", modelName)] %>% 
    drop_na()
  rownames(temp) <- temp$coefficientName
  temp[,modelName] <- round(temp[,modelName], 8)
  
  # print out coefficients w/ coefficient names
tempNames <- paste0(
  apply(temp, MARGIN = 1, FUN = function(x) {
    if (x["coefficientName"] == "(Intercept)") {
      paste0(x[modelName])
    } else {  
      paste0(x[modelName], "*", x["coefficientName"])
    }
    }
  ),
  collapse = " + "
)

# print the unscaled model statement
  unscaledModelName <- paste0(responseVar, "~ 1/(1 + exp(-(", tempNames, ")))")
  
  # now add in the scale parameters
tempNames <- str_replace_all(tempNames, pattern = "annWetDegDays_anom", 
                         replacement = paste0("((annWetDegDays_anom - ", 
                                              round(scaleParams$annWetDegDays_anom_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$annWetDegDays_anom_s$`scaled:scale`,8), ")"))

tempNames <- str_replace_all(tempNames, pattern = "prcpTempCorr_anom", 
                         replacement = paste0("((prcpTempCorr_anom - ", 
                                              round(scaleParams$prcpTempCorr_anom_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$prcpTempCorr_anom_s$`scaled:scale`,8), ")"))


tempNames <- str_replace_all(tempNames, pattern = "prcp_seasonality_anom", 
                         replacement = paste0("((prcp_seasonality_anom - ", 
                                              round(scaleParams$prcp_seasonality_anom_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$prcp_seasonality_anom_s$`scaled:scale`,8), ") "))

tempNames <- str_replace_all(tempNames, pattern = "annWatDef_anom", 
                         replacement = paste0("((annWatDef_anom - ", 
                                              round(scaleParams$annWatDef_anom_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$annWatDef_anom_s$`scaled:scale`,8), ") "))

tempNames <- str_replace_all(tempNames, pattern = "isothermality_anom", 
                         replacement = paste0("((isothermality_anom - ", 
                                              round(scaleParams$isothermality_anom_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$isothermality_anom_s$`scaled:scale`,8), ") "))

tempNames <- str_replace_all(tempNames, pattern = "prcp_anom", 
                         replacement = paste0("((prcp_anom - ", 
                                              round(scaleParams$prcp_anom_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$prcp_anom_s$`scaled:scale`,8), ") "))

tempNames <- str_replace_all(tempNames, pattern = "prcp ", 
                         replacement = paste0("((prcp - ", 
                                              round(scaleParams$prcp_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$prcp_s$`scaled:scale`,8), ") "))

tempNames <- str_replace_all(tempNames, pattern = "prcp\\^", 
                         replacement = paste0("((prcp - ", 
                                              round(scaleParams$prcp_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$prcp_s$`scaled:scale`,8), ")^"))

tempNames <- str_replace_all(tempNames, pattern = "prcp:", 
                         replacement = paste0("((prcp - ", 
                                              round(scaleParams$prcp_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$prcp_s$`scaled:scale`,8), "):"))

tempNames <- str_replace_all(tempNames, pattern = "prcpTempCorr ", 
                         replacement = paste0("((prcpTempCorr - ", 
                                              round(scaleParams$prcpTempCorr_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$prcpTempCorr_s$`scaled:scale`,8), ") "))

tempNames <- str_replace_all(tempNames, pattern = "prcpTempCorr\\^", 
                         replacement = paste0("((prcpTempCorr - ", 
                                              round(scaleParams$prcpTempCorr_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$prcpTempCorr_s$`scaled:scale`,8), ")^"))

tempNames <- str_replace_all(tempNames, pattern = "prcpTempCorr:", 
                         replacement = paste0("((prcpTempCorr - ", 
                                              round(scaleParams$prcpTempCorr_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$prcpTempCorr_s$`scaled:scale`,8), "):"))

tempNames <- str_replace_all(tempNames, pattern = "prcpTempCorr\\)", 
                         replacement = paste0("((prcpTempCorr - ", 
                                              round(scaleParams$prcpTempCorr_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$prcpTempCorr_s$`scaled:scale`,8), "))"))

tempNames <- str_replace_all(tempNames, pattern = "prcpTempCorr$", 
                         replacement = paste0("((prcpTempCorr - ", 
                                              round(scaleParams$prcpTempCorr_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$prcpTempCorr_s$`scaled:scale`,8), ")"))

tempNames <- str_replace_all(tempNames, pattern = "isothermality ", 
                         replacement = paste0("((isothermality - ", 
                                              round(scaleParams$isothermality_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$isothermality_s$`scaled:scale`,8), ") "))

tempNames <- str_replace_all(tempNames, pattern = "isothermality\\^", 
                         replacement = paste0("((isothermality - ", 
                                              round(scaleParams$isothermality_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$isothermality_s$`scaled:scale`,8), ")^"))

tempNames <- str_replace_all(tempNames, pattern = "isothermality:", 
                         replacement = paste0("((isothermality - ", 
                                              round(scaleParams$isothermality_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$isothermality_s$`scaled:scale`,8), "):"))

tempNames <- str_replace_all(tempNames, pattern = "sand", 
                         replacement = paste0("((sand - ", 
                                              round(scaleParams$sand_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$sand_s$`scaled:scale`,8), ")"))

tempNames <- str_replace_all(tempNames, pattern = "coarse", 
                         replacement = paste0("((coarse - ", 
                                              round(scaleParams$coarse_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$coarse_s$`scaled:scale`,8), ")"))

tempNames <- str_replace_all(tempNames, pattern = "AWHC", 
                         replacement = paste0("((AWHC - ", 
                                              round(scaleParams$AWHC_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$AWHC_s$`scaled:scale`,8), ")"))

tempNames <- str_replace_all(tempNames, pattern = "carbon", 
                         replacement = paste0("((carbon - ", 
                                              round(scaleParams$carbon_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$carbon_s$`scaled:scale`,8), ")"))

tempNames <- str_replace_all(tempNames, pattern = "clay", 
                         replacement = paste0("((clay - ", 
                                              round(scaleParams$clay_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$clay_s$`scaled:scale`,8), ")"))

tempNames <- str_replace_all(tempNames, pattern = "annWatDef ", 
                         replacement = paste0("((annWatDef - ", 
                                              round(scaleParams$annWatDef_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$annWatDef_s$`scaled:scale`,8), ") "))

tempNames <- str_replace_all(tempNames, pattern = "annWatDef:", 
                         replacement = paste0("((annWatDef - ", 
                                              round(scaleParams$annWatDef_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$annWatDef_s$`scaled:scale`,8), "):"))

tempNames <- str_replace_all(tempNames, pattern = "prcp_seasonality ", 
                         replacement = paste0("((prcp_seasonality - ", 
                                              round(scaleParams$prcp_seasonality_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$prcp_seasonality_s$`scaled:scale`,8), ") "))

tempNames <- str_replace_all(tempNames, pattern = "prcp_seasonality:", 
                         replacement = paste0("((prcp_seasonality - ", 
                                              round(scaleParams$prcp_seasonality_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$prcp_seasonality_s$`scaled:scale`,8), "):"))

tempNames <- str_replace_all(tempNames, pattern = "prcp_seasonality\\^", 
                         replacement = paste0("((prcp_seasonality - ", 
                                              round(scaleParams$prcp_seasonality_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$prcp_seasonality_s$`scaled:scale`,8), ")^"))

tempNames <- str_replace_all(tempNames, pattern = "tmean ", 
                         replacement = paste0("((tmean - ", 
                                              round(scaleParams$tmean_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$tmean_s$`scaled:scale`,8), ") "))

tempNames <- str_replace_all(tempNames, pattern = "tmean:", 
                         replacement = paste0("((tmean - ", 
                                              round(scaleParams$tmean_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$tmean_s$`scaled:scale`,8), "):"))

tempNames <- str_replace_all(tempNames, pattern = "tmean$", 
                         replacement = paste0("((tmean - ", 
                                              round(scaleParams$tmean_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$tmean_s$`scaled:scale`,8), ")"))

tempNames <- str_replace_all(tempNames, pattern = "tmean\\^", 
                         replacement = paste0("((tmean - ", 
                                              round(scaleParams$tmean_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$tmean_s$`scaled:scale`,8), ")^"))

tempNames <- str_replace_all(tempNames, pattern = "annWetDegDays ", 
                         replacement = paste0("((annWetDegDays - ", 
                                              round(scaleParams$annWetDegDays_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$annWetDegDays_s$`scaled:scale`,8), ") "))

tempNames <- str_replace_all(tempNames, pattern = "annWetDegDays:", 
                         replacement = paste0("((annWetDegDays - ", 
                                              round(scaleParams$annWetDegDays_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$annWetDegDays_s$`scaled:scale`,8), "):"))

tempNames <- str_replace_all(tempNames, pattern = "prcp_dry ", 
                         replacement = paste0("((prcp_dry - ", 
                                              round(scaleParams$prcp_dry_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$prcp_dry_s$`scaled:scale`,8), ") "))

tempNames <- str_replace_all(tempNames, pattern = "prcp_dry:", 
                         replacement = paste0("((prcp_dry - ", 
                                              round(scaleParams$prcp_dry_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$prcp_dry_s$`scaled:scale`,8), "):"))

tempNames <- str_replace_all(tempNames, pattern = "annWatDef_95", 
                         replacement = paste0("((annWatDef_95 - ", 
                                              round(scaleParams$annWatDef_95_s$`scaled:center`,8), ") / ", 
                                              round(scaleParams$annWatDef_95_s$`scaled:scale`,8), ")"))


## print scaled model statement

scaledModelName <- paste0(responseVar, "~ exp(", tempNames, ")")
  
return(list("scaledInputVars_ModelStatement" = unscaledModelName, 
            "unscaledInputVars_scaledModelStatement" = scaledModelName))
}


## bestLambda model equations
bestLambdaModStatement <- getModelStatement(coefficientTable = coefs, 
                  modelName = "coefficientValue_bestLambda", 
                  responseVar = response)
## second-best lambda model equations
secondBestLambdaModStatement <- getModelStatement(coefficientTable = coefs, 
                  modelName = ifelse(name_secondBestMod == "1/2 SE Model", 
                                      yes = "coefficientValue_halfseLambda", 
                                     no = "coefficientValue_1seLambda"), 
                  responseVar = response)
```

#### The equation for the bestLambda model if the inputs *are* scaled is: 
```{r}
print(bestLambdaModStatement$scaledInputVars_ModelStatement)
```
#### The equation for the second-bestLambda model if the inputs *are* scaled is: 
```{r}
print(secondBestLambdaModStatement$scaledInputVars_ModelStatement)
```


# Cross-validation
## Using best lambda model
### Use terms from global model to re-fit and predict on different held out regions
### Figures show residuals for each of the models fit to held-out ecoregions
These models were fit to six ecoregions, and then predict on the indicated heldout ecoregion

```{r cross-validation for best lambda model , fig.height = 7, fig.width = 9, warning=FALSE, message=FALSE }
if (length(prednames_fig) == 0) {
  print("The model only contains one predictor (an intercept), so cross validation isn't practical")
} else {

## code from Tredennick et al. 2020
# try each separate level II ecoregion as a test set
# make a list to hold output data
outList <- vector(mode = "list", length = length(sort(unique(modDat_1_s$NA_L2NAME))))
# obs_pred <- data.frame(ecoregion = character(),obs = numeric(),
#                        pred_opt = numeric(), pred_null = numeric()#,
#                        #pred_nopenalty = numeric()
#                        )

## get the model specification from the global model
mat <- as.matrix(coef(fit_glm_bestLambda, s = "lambda.min"))
mat2 <- mat[mat[,1] != 0,]

f_cv <- as.formula(paste0(response, " ~ ", paste0(names(mat2)[2:length(names(mat2))], collapse = " + ")))

X_cv <- model.matrix(object = f_cv, data = modDat_1_s)
# get response variable
y_cv <- as.matrix(modDat_1_s[,response])


# now, loop through so with each iteration, a different ecoregion is held out
 for(i_eco in sort(unique(modDat_1_s$NA_L2NAME))){

  # split into training and test sets
  test_eco <- i_eco
  print(test_eco)
  # identify the rowID of observations to be in the training and test datasets
  train <- which(modDat_1_s$NA_L2NAME!=test_eco) # data for all ecoregions that aren't 'i_eco'
  test <- which(modDat_1_s$NA_L2NAME==test_eco) # data for the ecoregion that is 'i_eco'

  trainDat_all <- modDat_1_s %>%
    slice(train) %>%
    dplyr::select(-newRegion)
  testDat_all <- modDat_1_s %>%
    slice(test) %>%
    dplyr::select(-newRegion)
## make testDat_all a data.frame that matches data used for fit_glm_bestLambda model (fit to matrix)
testDat_DF <- data.frame(matrix(X[test,2:ncol(X)], nrow = length(test))) #data.frame(X[test,2:ncol(X)])
names(testDat_DF) <- colnames(X)[2:ncol(X)]
  
# get the model matrices for input and response variables for cross validation model specification
  X_train <- as.matrix(X_cv[train,])
  X_test <- as.matrix(X_cv[test,])

  y_train <- modDat_1_s[train,response]
  y_test <- modDat_1_s[test,response]

  # get the model matrices for input and response variables for original model specification
  X_train_glob <- as.matrix(X[train,])
  X_test_glob <- as.matrix(X[test,])

  y_train_glob <- modDat_1_s[train,response]
  y_test_glob <- modDat_1_s[test,response]

  train_eco <- modDat_1_s$NA_L2NAME[train]

  ## just try a regular glm w/ the components from the global model
  fit_i <- glm(data = trainDat_all, formula = f_cv,
    ,
               family =  quasibeta_family(link = "logit")
    )
    
  # lasso model predictions with the optimal lambda (back transformed)
  optimal_pred <- predict.glm(fit_i, newdata= testDat_all, type = "response") 
  # null model and predictions
  # the "null" model in this case is the global model
  # predict on the test data for this iteration w/ the global model (back transformed)

  null_pred <- predict.glm(fit_glm_bestLambda, newdata = testDat_DF, type = "response") 


  # save data
  tmp <- data.frame(ecoRegion_holdout = rep(test_eco,length(y_test)),
                    obs=y_test,
                    pred_opt=optimal_pred,
                    pred_null=null_pred#,
                    #pred_nopenalty=nopen_pred
                    ) %>%
    cbind(testDat_all)

  # calculate RMSE, bias, etc. of
  # RMSE of CV model
  RMSE_optimal <- yardstick::rmse(data = data.frame(optimal_pred,"y_test" = (y_test)), truth = "y_test", estimate = "optimal_pred")[1,]$.estimate
  # RMSE of global model
  RMSE_null <- yardstick::rmse(data = data.frame(null_pred,"y_test" = (y_test)), truth = "y_test", estimate = "null_pred")[1,]$.estimate
  # bias of CV model
  bias_optimal <- mean((y_test) - optimal_pred)
  # bias of global model
  bias_null <-  mean((y_test) - null_pred )

  # put output into a list
  tmpList <- list("testRegion" = i_eco,
    "modelObject" = fit_i,
       "modelPredictions" = tmp,
    "performanceMetrics" = data.frame("RMSE_cvModel" = RMSE_optimal,
                                      "RMSE_globalModel" = RMSE_null,
                                      "bias_cvModel" = bias_optimal,
                                      "bias_globalModel" = bias_null))

  # save model outputs
  outList[[which(sort(unique(modDat_1_s$NA_L2NAME)) == i_eco)]] <- tmpList
 }
}
```
### Below are the RMSE and bias values for predictions made for each holdout level II ecoregion, compared to predictions from the global model for that same ecoregion
```{r metrics output table for best lambda CV}
# table of model performance
purrr::map(outList, .f = function(x) {
  cbind(data.frame("holdout region" = x$testRegion),  x$performanceMetrics)
}
) %>%
  purrr::list_rbind() %>%
  kable(col.names = c("Held-out ecoregion", "RMSE of CV model", "RMSE of global model",
                      "bias of CV model - mean(obs-pred.)", "bias of global model- mean(obs-pred.)"),
        caption = "Performance of Cross Validation using 'best lambda' model specification") %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

```{r figures for best lambda CV}
# visualize model predictions
for (i in 1:length(unique(modDat_1_s$NA_L2NAME))) {
  holdoutRegion <- outList[[i]]$testRegion
  predictionData <- outList[[i]]$modelPredictions
  modTerms <- as.matrix(coef(outList[[i]]$modelObject)) %>%
    as.data.frame() %>%
    filter(V1!=0) %>%
    rownames()

  # calculate residuals
  predictionData <- predictionData %>%
  mutate(resid = .[["obs"]] - .[["pred_opt"]] ,
         resid_globMod = .[["obs"]]  - .[["pred_null"]])

# rasterize
# use 'test_rast' from earlier

  # rasterize data
plotObs <- predictionData %>%
         drop_na(paste(response)) %>%
  #slice_sample(n = 5e4) %>%
  terra::vect(geom = c("x", "y")) %>%
  terra::set.crs(crs(crs_veg)) %>%
  terra::project(crs(test_rast)) %>%
  terra::rasterize(y = test_rast,
                   field = "resid",
                   fun = mean) #%>%
  #terra::aggregate(fact = 2, fun = mean, na.rm = TRUE) %>%
  #terra::crop(ext(-1950000, 1000000, -1800000, 1000000))

tempExt <- crds(plotObs, na.rm = TRUE)
plotObs_2 <- plotObs %>%  
  crop(ext(
    min(tempExt[,1]), max(tempExt[,1])+1,
           min(tempExt[,2]), max(tempExt[,2])+1
    )
       )

# identify locations where residuals are >100 or < -100
badResids_high <- predictionData %>%
  filter(resid > .5)  %>%
  terra::vect(geom = c("x", "y")) %>%
  terra::set.crs(crs(crs_veg)) %>%
  terra::project(crs(test_rast))
badResids_low <- predictionData %>%
  filter(resid < -.5)  %>%
  terra::vect(geom = c("x", "y")) %>%
  terra::set.crs(crs(crs_veg)) %>%
  terra::project(crs(test_rast)) 

# make figures
# make histogram
hist_i <- ggplot(predictionData) +
  geom_histogram(aes(resid_globMod), col = "darkgrey", fill = "lightgrey") +
  xlab(c("Residuals (obs. - pred.)"))
# make map
map_i <-  ggplot() +
geom_spatraster(data = plotObs_2) +
  geom_sf(data=cropped_states, fill=NA ) +
  geom_sf(data = badResids_high, col = "blue") +
  geom_sf(data = badResids_low, col = "red") +
labs(title = paste0("Residuals (obs. - pred.) for predictions of \n", holdoutRegion, " \n from a model fit to other ecoregions"),
     subtitle = paste0(response, " ~ ", paste0( modTerms, collapse = " + "))) +
  scale_fill_gradient2(low = "red",
                       mid = "white" ,
                       high = "blue" ,
                       midpoint = 0,   na.value = "grey20",
                       limits = c(-1, 1)) +
  xlim(c(min(tempExt[,1]), max(tempExt[,1])+1)) + 
  ylim(c(min(tempExt[,2]), max(tempExt[,2])+1))

 assign(paste0("residPlot_",holdoutRegion),
   value = ggarrange(map_i, hist_i, heights = c(3,1), ncol = 1)
)

}

  lapply(unique(modDat_1_s$NA_L2NAME), FUN = function(x) {
    get(paste0("residPlot_", x))
  })

```

## Using second best lambda model (either 1se or 1/2se)
### Use terms from global model to re-fit and predict on different held out regions
### Figures show residuals for each of the models fit to held-out ecoregions
These models were fit to six ecoregions, and then predict on the indicated heldout ecoregion

```{r cross-validation for 1 SE lambda model , fig.height = 7, fig.width = 9, warning=FALSE, message=FALSE }
if (length(prednames_secondBestMod) == 0) {
  print("The model only contains one predictor (an intercept), so cross validation isn't practical")
} else {

## code from Tredennick et al. 2020
# try each separate level II ecoregion as a test set
# make a list to hold output data
outList <- vector(mode = "list", length = length(sort(unique(modDat_1_s$NA_L2NAME))))
# obs_pred <- data.frame(ecoregion = character(),obs = numeric(),
#                        pred_opt = numeric(), pred_null = numeric()#,
#                        #pred_nopenalty = numeric()
#                        )

## get the model specification from the global model
mat <- as.matrix(coef(mod_secondBest, s = "lambda.min"))
mat2 <- mat[mat[,1] != 0,]

f_cv <- as.formula(paste0(response, " ~ ", paste0(names(mat2)[2:length(names(mat2))], collapse = " + ")))

X_cv <- model.matrix(object = f_cv, data = modDat_1_s)
# get response variable
y_cv <- as.matrix(modDat_1_s[,response])


# now, loop through so with each iteration, a different ecoregion is held out
 for(i_eco in sort(unique(modDat_1_s$NA_L2NAME))){

  # split into training and test sets
  test_eco <- i_eco
  print(test_eco)
  # identify the rowID of observations to be in the training and test datasets
  train <- which(modDat_1_s$NA_L2NAME!=test_eco) # data for all ecoregions that aren't 'i_eco'
  test <- which(modDat_1_s$NA_L2NAME==test_eco) # data for the ecoregion that is 'i_eco'

  trainDat_all <- modDat_1_s %>%
    slice(train) %>%
    dplyr::select(-newRegion)
  testDat_all <- modDat_1_s %>%
    slice(test) %>%
    dplyr::select(-newRegion)
## make testDat_all a data.frame that matches data used for fit_glm_bestLambda model (fit to matrix)
testDat_DF <- data.frame(matrix(X[test,2:ncol(X)], nrow = length(test)))#testDat_DF <- data.frame(X[test,2:ncol(X)])
names(testDat_DF) <- colnames(X)[2:ncol(X)]
  # get the model matrices for input and response variables for cross validation model specification
  X_train <- as.matrix(X_cv[train,])
  X_test <- as.matrix(X_cv[test,])

  y_train <- modDat_1_s[train,response]
  y_test <- modDat_1_s[test,response]

  # get the model matrices for input and response variables for original model specification
  X_train_glob <- as.matrix(X[train,])
  X_test_glob <- as.matrix(X[test,])

  y_train_glob <- modDat_1_s[train,response]
  y_test_glob <- modDat_1_s[test,response]

  train_eco <- modDat_1_s$NA_L2NAME[train]

  ## just try a regular glm w/ the components from the global model
  fit_i <- glm(data = trainDat_all, formula = f_cv,
    ,
               family =  quasibeta_family(link = "logit")
    )

    coef(fit_i)

  # lasso model predictions with the optimal lambda
  optimal_pred <- predict(fit_i, newdata= testDat_all, type = "response") 
  # null model and predictions
  # the "null" model in this case is the global model
  # predict on the test data for this iteration w/ the global model
  null_pred <- predict.glm(mod_secondBest, newdata = testDat_DF, type = "response") 

  # save data
  tmp <- data.frame(ecoRegion_holdout = rep(test_eco,length(y_test)),
                    obs=y_test,
                    pred_opt=optimal_pred,
                    pred_null=null_pred#,
                    #pred_nopenalty=nopen_pred
                    ) %>%
    cbind(testDat_all)

  # calculate RMSE, bias, etc. of
  # RMSE of CV model
  RMSE_optimal <- yardstick::rmse(data = data.frame(optimal_pred, "y_test" = (y_test)), truth = "y_test", estimate = "optimal_pred")[1,]$.estimate
  # RMSE of global model
  RMSE_null <- yardstick::rmse(data = data.frame(null_pred,  "y_test" = (y_test)), truth = "y_test", estimate = "null_pred")[1,]$.estimate
  # bias of CV model
  bias_optimal <- mean((y_test) - optimal_pred)
  # bias of global model
  bias_null <-  mean((y_test) - null_pred )

  # put output into a list
  tmpList <- list("testRegion" = i_eco,
    "modelObject" = fit_i,
       "modelPredictions" = tmp,
    "performanceMetrics" = data.frame("RMSE_cvModel" = RMSE_optimal,
                                      "RMSE_globalModel" = RMSE_null,
                                      "bias_cvModel" = bias_optimal,
                                      "bias_globalModel" = bias_null))

  # save model outputs
  outList[[which(sort(unique(modDat_1_s$NA_L2NAME)) == i_eco)]] <- tmpList
 }
}
```

### Below are the RMSE and bias values for predictions made for each holdout level II ecoregion, compared to predictions from the global model for that same ecoregion
```{r metrics output table for 1SE lambda CV}

if (length(prednames_secondBestMod) == 0) {
  print("The model only contains one predictor (an intercept), so cross validation isn't practical")
} else {
# table of model performance
purrr::map(outList, .f = function(x) {
  cbind(data.frame("holdout region" = x$testRegion),  x$performanceMetrics)
}
) %>%
  purrr::list_rbind() %>%
  kable(col.names = c("Held-out ecoregion", "RMSE of CV model", "RMSE of global model",
                      "bias of CV model - mean(obs-pred.)", "bias of global model - mean(obs-pred.)"),
        caption = "Performance of Cross Validation using 'second best lambda' model specification") %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
}
```

```{r figures of CV for 1 SE lambda model}
if (length(prednames_secondBestMod) == 0) {
  print("The model only contains one predictor (an intercept), so cross validation isn't practical")
} else {
for (i in 1:length(unique(modDat_1_s$NA_L2NAME))) {
  holdoutRegion <- outList[[i]]$testRegion
  predictionData <- outList[[i]]$modelPredictions
  modTerms <- as.matrix(coef(outList[[i]]$modelObject)) %>%
    as.data.frame() %>%
    filter(V1!=0) %>%
    rownames()

  # calculate residuals
  predictionData <- predictionData %>%
  mutate(resid = .[["obs"]] - .[["pred_opt"]] ,
         resid_globMod = .[["obs"]]  - .[["pred_null"]])

  # rasterize data
plotObs <- predictionData %>%
         drop_na(paste(response)) %>%
  #slice_sample(n = 5e4) %>%
  terra::vect(geom = c("x", "y")) %>%
  terra::set.crs(crs(crs_veg)) %>%
  terra::project(crs(test_rast)) %>% 
  terra::rasterize(y = test_rast,
                   field = "resid",
                   fun = mean) #%>%
  #terra::aggregate(fact = 2, fun = mean, na.rm = TRUE) %>%
  #terra::crop(ext(-1950000, 1000000, -1800000, 1000000))

tempExt <- crds(plotObs, na.rm = TRUE)

plotObs_2 <- plotObs %>%
  crop(ext(min(tempExt[,1]), max(tempExt[,1])+1,
           min(tempExt[,2]), max(tempExt[,2])+1)
       )

# identify locations where residuals are >100 or < -100
badResids_high <- predictionData %>%
  filter(resid > .5)  %>%
  terra::vect(geom = c("x", "y")) %>%
  terra::set.crs(crs(crs_veg)) %>%
  terra::project(crs(test_rast)) 
badResids_low <- predictionData %>%
  filter(resid < -.5)  %>%
  terra::vect(geom = c("x", "y")) %>% 
  terra::set.crs(crs(crs_veg)) %>%
  terra::project(crs(test_rast)) 


# make figures
# make histogram
hist_i <- ggplot(predictionData) +
  geom_histogram(aes(resid_globMod), col = "darkgrey", fill = "lightgrey") +
  xlab(c("Residuals (obs. - pred.)"))
# make map
map_i <-  ggplot() +
geom_spatraster(data = plotObs_2) +
  geom_sf(data=cropped_states,fill=NA ) +
  geom_sf(data = badResids_high, col = "blue") +
  geom_sf(data = badResids_low, col = "red") +
labs(title = paste0("Residuals (obs. - pred.) for predictions of \n", holdoutRegion, " \n from a model fit to other ecoregions"),
     subtitle = paste0(response, " ~ ", paste0( modTerms, collapse = " + "))) +
  scale_fill_gradient2(low = "red",
                       mid = "white" ,
                       high = "blue" ,
                       midpoint = 0,   na.value = "grey20",
                       limits = c(-1, 1))  +
  xlim(st_bbox(plotObs_2)[c(1,3)]) +
  ylim(st_bbox(plotObs_2)[c(2,4)])

 assign(paste0("residPlot_",holdoutRegion),
   value = ggarrange(map_i, hist_i, heights = c(3,1), ncol = 1)
)

}

  lapply(unique(modDat_1_s$NA_L2NAME), FUN = function(x) {
    get(paste0("residPlot_", x))
  })
}

```


# Save output

```{r save_output}
## save the coefficients for the models (best lambda, 1/2se lambda, 1se lambda)
if(trimAnom == TRUE) {
saveRDS(coefs, file = paste0("./models/modelCoefficients_trimAnom_", ecoregion, "_", response,"_removeAnomalies", removeAllAnoms,"_observeDatForestClassification.rds"))
saveRDS(uniqueCoeffs, file = paste0("./models/modelMetrics_trimAnom_", ecoregion, "_", response,"_removeAnomalies", removeAllAnoms,"_observeDatForestClassification.rds"))
} else {
saveRDS(coefs, file = paste0("./models/modelCoefficients_", ecoregion, "_", response,"_removeAnomalies", removeAllAnoms,"_observeDatForestClassification.rds"))
saveRDS(uniqueCoeffs, file = paste0("./models/modelMetrics_", ecoregion, "_", response,"_removeAnomalies", removeAllAnoms,"_observeDatForestClassification.rds"))
}
# make a table

```




```{r pdp_glm, warning = FALSE, fig.width = 8, fig.height=8}
## partial dependence plots
#vip::vip(mod_glmFinal, num_features = 15)

#pdp_all_vars(mod_glmFinal, mod_vars = pred_vars, ylab = 'probability',train = df_small)

#caret::varImp(fit)
```


# session info

Hash of current commit (i.e. to ID the version of the code used)

```{r}
system("git rev-parse HEAD", intern=TRUE)
```

Packages etc.

```{r}
sessionInfo()

```