---
title: "Visualizing cover data at each step of the data preparation workflow"
output:
  html_document:
    code_folding: hide
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE,
                      message = FALSE)
```

```{r}
library(tidyverse) 
library(sf) 
library(terra)
library(ggpubr)
library(tidyterra)
library(USAboundaries)
ggplot2::theme_set(theme_classic())
# custom functions for plotting inset figures
            # make function to add inset
           get_inset <- function(df) {
              p <- ggplot(data = df) +
             facet_wrap(~lyr, ncol = 1, scales = "free_y") +
                geom_density(aes(Cover)) +
                theme_bw(base_size=9) +  ## makes everything smaller
                theme(panel.background = element_rect(fill="transparent"),  ## white plot background
                      panel.border = element_blank(),
          axis.title.y = element_blank(),
          axis.title.x = element_blank(),
          axis.text.x = element_text(size=rel(0.8)), ## tiny axis text
          axis.text.y = element_text(size=rel(0.8)), ## tiny axis text
          axis.line.x = element_line(size = .5),
          axis.line.y = element_line(size = .5),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          plot.background = element_blank(),
          strip.background = element_blank(),
          strip.text = element_blank()) 
  return(p)
           }
           # custom function to add inset to main plot
           annotation_custom2 <- function (grob, xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, data)
{
  layer(data = data, stat = StatIdentity, position = PositionIdentity,
        geom = ggplot2:::GeomCustomAnn,
        inherit.aes = TRUE, params = list(grob = grob,
                                          xmin = xmin, xmax = xmax,
                                          ymin = ymin, ymax = ymax))
}
```

This document visualizes the data of vegetation cover by functional type that we use in subsequent analysis at each step of the data-processing workflow. These maps serve both as a reference and a check that the workflow is performing as intended. 

For the sake of simplified comparison, I show data from 2016, 2020, and 2022, which are years for which we have easy to access rasters of FIA data to compare to. 
```{r}
# get necessary raster data for background data on maps
test_rastTemp <-  rast("../../../Data_raw/dayMet/rawMonthlyData/orders/70e0da02b9d2d6e8faa8c97d211f3546/Daymet_Monthly_V4R1/data/daymet_v4_prcp_monttl_na_1980.tif") %>%
  terra::crop(ext(-2000000, 2600000, -1900000, 1500000))

## aggregate test_rastTemp by 13, the scale that the dayMet data are aggregated to 
test_rast_RAP <- test_rastTemp %>% 
  terra::aggregate(fact = 13, fun = "mean") #%>% 
#terra::project(crs("EPSG:4326")) %>% 
  #terra::crop(ext(-150, -50, 20, 60))
  
# aggregate to regular plotting density for other data types
test_rast <- test_rastTemp %>%
 terra::aggregate(fact = 8, fun = "mean")#%>% 
#terra::project(crs("EPSG:4326")) %>% 
 # terra::crop(ext(-150, -50, 20, 60))
  # transform to match format of veg. data 

cropped_states_2 <- us_states(resolution = "low")
cropped_states_2 <- suppressMessages(cropped_states_2 %>%
  dplyr::filter(name!="Hawaii") %>%
  dplyr::filter(name!="Alaska") %>%
  dplyr::filter(name!="Puerto Rico") %>%
  dplyr::filter(name!="American Samoa") %>%
  dplyr::filter(name!="Guam") %>%
  dplyr::filter(name!="Commonwealth of the Northern Mariana Islands") %>%
  dplyr::filter(name!="United States Virgin Islands") %>%
  sf::st_transform(sf::st_crs(test_rast))) 
```

# RAP 
## Initial dataset
```{r, echo = FALSE, fig.width= 20, fig.height = 14, warning=FALSE}
# Read in data 
dat_1temp <- st_read(dsn = "../../../Data_processed/CoverData/DataForAnalysisPoints", layer = "vegCompPoints") %>% #, driver = "ESRI Shapefile")
  st_transform(crs(test_rast))

# trim to be only later than 2000, which is what we want to use for further analysis
dat_1 <- dat_1temp %>% 
  filter(Year >= 2000) %>% 
  mutate(Lon = st_coordinates(.)[,1],
         Lat = st_coordinates(.)[,2]) %>% 
  rename(ShrubCover = ShrbCvr, 
         TotalTreeCover = TtlTrCv, 
         TotalHerbaceousCover = TtlHrbC,
         BareGroundCover = BrGrndC,
         ForbCover = ForbCvr,
         C3GramCover = C3GrmCv,
         C4GramCover = C4GrmCv, 
         BroadLeavedTreeCover = AngTrCv,
         NeedleLeavedTreeCover = CnfTrCv)
 # st_drop_geometry()

###### rasterized version
# make a multi-layer raster w/ layers for each variable of interest
dat_1_RAST <- 
  lapply(c("FIA", "LANDFIRE", "LDC","RAP"), 
         FUN = function(y) {
           tempDat <- dat_1 %>% 
             filter(Source == y )
           tempDat1 <- tempDat %>% 
             st_drop_geometry()
           # get vector of years for the specific data type 
           if (y %in% c("FIA", "LDC", "RAP")) {
             datYears <- c("2016", "2020", "2022")
             if (y %in% "RAP") {
               # remove data from before 2010
               tempDat1 <- tempDat1 %>% 
                 filter(Year >=2010)
             }
           } else {
             datYears <- c("2003", "2007", "2015")
           }
           if (y == c( "RAP")) {
             # get data for each cover type by year
           tempOut <- apply(expand.grid(c("ShrubCover", "TotalTreeCover", "TotalHerbaceousCover", "BareGroundCover", "ForbCover", "C3GramCover",  "C4GramCover",  "BroadLeavedTreeCover", "NeedleLeavedTreeCover"), datYears, stringsAsFactors = FALSE), 
                            MARGIN = 1,
                            FUN = function(x) {
                              tempDat2 <- tempDat %>% 
                                filter(Year == x["Var2"])
                              
                              temp <- terra::rasterize(x = tempDat2,#tempDat2[,c("Lon", "Lat")], 
                                               y = test_rast_RAP,
                                               field = x["Var1"],
                                               #values = tempDat2[,names(tempDat2) == x["Var1"]],
                                               fun = mean, 
                                               na.rm = TRUE
                                               ) 
                      names(temp) <-  paste0(x["Var1"], "_", x["Var2"])
                      return(temp)
                    }
                    )
           # get dat across all years
           tempOut2 <- lapply(c("ShrubCover", "TotalTreeCover", "TotalHerbaceousCover", "BareGroundCover", "ForbCover", "C3GramCover",  "C4GramCover",  "BroadLeavedTreeCover", "NeedleLeavedTreeCover"), 
                            FUN = function(z) {
                              temp <- terra::rasterize(x = tempDat1[,c("Lon", "Lat")], 
                                               y = test_rast_RAP,
                                               values = tempDat1[,z],
                                               fun = mean
                                               ) 
                      
                      names(temp) <-  paste0(z, "_all")
                      return(temp)
                    }
                    )
           } else {
             # get data for each cover type by year
           tempOut <- apply(expand.grid(c("ShrubCover", "TotalTreeCover", "TotalHerbaceousCover", "BareGroundCover", "ForbCover", "C3GramCover",  "C4GramCover",  "BroadLeavedTreeCover", "NeedleLeavedTreeCover"), datYears, stringsAsFactors = FALSE), 
                            MARGIN = 1,
                            FUN = function(x) {
                              tempDat2 <- tempDat %>% 
                                filter(Year == x["Var2"])
                              
                              temp <- terra::rasterize(x = tempDat2,#tempDat2[,c("Lon", "Lat")], 
                                               y = test_rast,
                                               field = x["Var1"],
                                               #values = tempDat2[,names(tempDat2) == x["Var1"]],
                                               fun = mean, 
                                               na.rm = TRUE
                                               ) 
                      names(temp) <-  paste0(x["Var1"], "_", x["Var2"])
                      return(temp)
                    }
                    )
           # get dat across all years
           tempOut2 <- lapply(c("ShrubCover", "TotalTreeCover", "TotalHerbaceousCover", "BareGroundCover", "ForbCover", "C3GramCover",  "C4GramCover",  "BroadLeavedTreeCover", "NeedleLeavedTreeCover"), 
                            FUN = function(z) {
                              temp <- terra::rasterize(x = tempDat1[,c("Lon", "Lat")], 
                                               y = test_rast,
                                               values = tempDat1[,z],
                                               fun = mean
                                               ) 
                      
                      names(temp) <-  paste0(z, "_all")
                      return(temp)
                    }
                    )
           }
           
           # for each data type, put data for each year in one raster in the list, and the data across all years in another raster
           outDat <- list(c(tempOut[[1]], tempOut[[2]], tempOut[[3]], tempOut[[4]], tempOut[[5]], tempOut[[6]], tempOut[[7]], tempOut[[8]], tempOut[[9]]), 
                          c(tempOut[[10]], tempOut[[11]], tempOut[[12]], tempOut[[13]], tempOut[[14]], tempOut[[15]], tempOut[[16]], tempOut[[17]], tempOut[[18]]),
                          c(tempOut[[19]], tempOut[[20]], tempOut[[21]], tempOut[[22]], tempOut[[23]], tempOut[[24]], tempOut[[25]], tempOut[[26]], tempOut[[27]]), 
                          c(tempOut2[[1]], tempOut2[[2]], tempOut2[[3]], tempOut2[[4]], tempOut2[[5]], tempOut2[[6]], tempOut2[[7]], tempOut2[[8]], tempOut2[[9]]))
           names(outDat) <- c(datYears, "allYears")
           return(outDat)
         }
)
names(dat_1_RAST) <- c("FIA", "LANDFIRE", "LDC","RAP")

# ##testing 
# mapview(temp2 %>% terra::crop(ext(c(-1400000, -500000, -600000, 0))), maxpixels = 540000) + mapview(tempDat)
# plot(temp %>% terra::crop(ext(c(-1200000, -800000, -400000, -200000))))
# points(tempDat$geometry)
## 
## make figure for RAP 
testFigs <- lapply(names(dat_1_RAST$RAP)[[4]], FUN = function(y) {
  # make a map for this year
           tempFig <- 
  ggplot() +
             facet_wrap(~lyr, ncol = 1) +
geom_spatraster(data = dat_1_RAST$RAP[[y]][[1:4]], maxcell = 5e+10) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) +
  coord_sf(datum = st_crs(test_rastTemp)) +
             scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"), 
                          limits = c(0,100), na.value = "white") 
  # add an inset showing a density plot  
           # get data
           data_inset <- terra::values(dat_1_RAST$RAP[[y]][[1:4]]) %>% 
               as_tibble() %>% 
               pivot_longer(1:4, names_to = "lyr", values_to = "Cover")
 
           # make insets
           inset_plots <- data_inset %>%
             split(f = .$lyr) %>%
             purrr::map(~annotation_custom2(
    grob = ggplotGrob(get_inset(.) #+
                        #scale_y_continuous(limits=c(0,105), breaks = c(0, 50, 100))
                      ),
    data = data.frame(lyr=unique(.$lyr)),
    ymin = -2025000, ymax=-1205000, xmin=-2280000, xmax=-180000)
  )

         tempFig2 <- tempFig + inset_plots

           return(tempFig2)
})

jpeg(filename = "/Users/astears/Documents/Dropbox_static/Work/NAU_USGS_postdoc/cleanPED/PED_vegClimModels/Figures/CoverDatFigures/RAPInputs_step1.jpeg",
    width = 800, height = 2000)
ggarrange(#testFigs[[1]], 
          #testFigs[[2]], 
          ##testFigs[[3]], 
          testFigs[[1]], 
          nrow = 1, common.legend = TRUE, legend = "bottom"
          ) %>% 
  annotate_figure(top = "Map of RAP data on a dayMet x 13 grid")
dev.off()
```
```{r}
# barplots of # of observations per data type per year
dat_1 %>% 
  st_drop_geometry() %>% 
  filter(Source == "RAP" & Year >=2010) %>%
  group_by(Year) %>% 
  summarize(ShrubCover = sum(!is.na(ShrubCover)), 
            TotalTreeCover = sum(!is.na(TotalTreeCover)),
            TotalHerbaceousCover = sum(!is.na(TotalHerbaceousCover)),
            BareGroundCover = sum(!is.na(BareGroundCover))) %>% 
  pivot_longer(2:5, values_to = "No_Observations", names_to = "functionalGroup") %>% 
  ggplot() + 
  geom_col(aes(x = Year, y = No_Observations, fill = functionalGroup), position = "dodge", col = "grey20")  + 
   scale_fill_discrete(type = c("#000000", #"#E69F00" ,"#56B4E9", "#009E73", "#F0E442" ,"#0072B2" ,
                                "#D55E00", "#CC79A7", "#999999"),
                       labels = c("BareGroundCover", #"BroadLeavedTreeCover", "C3GramCover","C4GramCover","ForbCover", "NeedleLeavedTreeCover",
                                  "ShrubCover",  "TotalHerbaceousCover", "TotalTreeCover")
   )
```

<!-- ## After removing burned area -->
<!-- ```{r , fig.width= 20, fig.height = 14} -->
<!-- # Read in data -->
<!-- dat_2temp <- readRDS(file = "../../../Data_processed/CoverData/dataForAnalysis_fireRemoved.rds") %>%  -->
<!--   st_transform(crs(test_rast)) -->
<!-- # trim to be only later than 2000, which is what we want to use for further analysis -->
<!-- dat_2 <- dat_2temp %>%  -->
<!--   filter(Year >= 2000) %>%  -->
<!--   mutate(Lon = st_coordinates(.)[,1], -->
<!--          Lat = st_coordinates(.)[,2]) %>%  -->
<!--   rename(ShrubCover = ShrbCvr,  -->
<!--          TotalTreeCover = TtlTrCv,  -->
<!--          TotalHerbaceousCover = TtlHrbC, -->
<!--          BareGroundCover = BrGrndC, -->
<!--          ForbCover = ForbCvr, -->
<!--          C3GramCover = C3GrmCv, -->
<!--          C4GramCover = C4GrmCv,  -->
<!--          BroadLeavedTreeCover = AngTrCv, -->
<!--          NeedleLeavedTreeCover = CnfTrCv) -->
<!--  # st_drop_geometry() -->

<!-- ###### rasterized version -->
<!-- # make a multi-layer raster w/ layers for each variable of interest -->
<!-- dat_2_RAST <-  -->
<!--   lapply(c("FIA", "LANDFIRE", "LDC","RAP"),  -->
<!--          FUN = function(y) { -->
<!--            tempDat <- dat_2 %>%  -->
<!--              filter(Source == y)  -->
<!--            tempDat1 <- tempDat %>%  -->
<!--              st_drop_geometry() -->
<!--            # get vector of years for the specific data type  -->
<!--            if (y %in% c("FIA", "LDC", "RAP")) { -->
<!--              datYears <- c("2016", "2020", "2022") -->
<!--              if (y %in% "RAP") { -->
<!--                # remove data from before 2010 -->
<!--                tempDat1 <- tempDat1 %>%  -->
<!--                  filter(Year >=2010) -->
<!--              } -->
<!--            } else { -->
<!--              datYears <- c("2003", "2007", "2015") -->
<!--            } -->
<!--            # get data for each cover type by year -->
<!--            tempOut <- apply(expand.grid(c("ShrubCover", "TotalTreeCover", "TotalHerbaceousCover", "BareGroundCover", "ForbCover", "C3GramCover",  "C4GramCover",  "BroadLeavedTreeCover", "NeedleLeavedTreeCover"), datYears, stringsAsFactors = FALSE),  -->
<!--                             MARGIN = 1, -->
<!--                             FUN = function(x) { -->
<!--                               tempDat2 <- tempDat1 %>%  -->
<!--                                 filter(Year == x["Var2"]) -->
<!--                               temp <- terra::rasterize(x = tempDat2[,c("Lon", "Lat")],  -->
<!--                                                y = test_rast, -->
<!--                                                values = tempDat2[,names(tempDat2) == x["Var1"]], -->
<!--                                                fun = mean,  -->
<!--                                                na.rm = TRUE -->
<!--                                                ) -->
<!--                       names(temp) <-  paste0(x["Var1"], "_", x["Var2"]) -->
<!--                       return(temp) -->
<!--                     } -->
<!--                     ) -->
<!--            # get dat across all years -->
<!--            tempOut2 <- lapply(c("ShrubCover", "TotalTreeCover", "TotalHerbaceousCover", "BareGroundCover", "ForbCover", "C3GramCover",  "C4GramCover",  "BroadLeavedTreeCover", "NeedleLeavedTreeCover"),  -->
<!--                                FUN = function(z) { -->
<!--                               temp <- terra::rasterize(x = tempDat1[,c("Lon", "Lat")],  -->
<!--                                                y = test_rast, -->
<!--                                                values = tempDat1[,z], -->
<!--                                                fun = mean -->
<!--                                                ) -->

<!--                       names(temp) <-  paste0(z, "_all") -->
<!--                       return(temp) -->
<!--                     } -->
<!--                     ) -->
<!--            # for each data type, put data for each year in one raster in the list, and the data across all years in another raster -->
<!--                   outDat <- list(c(tempOut[[1]], tempOut[[2]], tempOut[[3]], tempOut[[4]], tempOut[[5]], tempOut[[6]], tempOut[[7]], tempOut[[8]], tempOut[[9]]),  -->
<!--                           c(tempOut[[10]], tempOut[[11]], tempOut[[12]], tempOut[[13]], tempOut[[14]], tempOut[[15]], tempOut[[16]], tempOut[[17]], tempOut[[18]]), -->
<!--                           c(tempOut[[19]], tempOut[[20]], tempOut[[21]], tempOut[[22]], tempOut[[23]], tempOut[[24]], tempOut[[25]], tempOut[[26]], tempOut[[27]]),  -->
<!--                           c(tempOut2[[1]], tempOut2[[2]], tempOut2[[3]], tempOut2[[4]], tempOut2[[5]], tempOut2[[6]], tempOut2[[7]], tempOut2[[8]], tempOut2[[9]])) -->
<!--            names(outDat) <- c(datYears, "allYears") -->
<!--            return(outDat) -->
<!--          } -->
<!-- ) -->

<!-- names(dat_2_RAST) <- c("FIA", "LANDFIRE", "LDC","RAP") -->

<!-- ## make figure for RAP  -->
<!-- testFigs <- lapply(names(dat_2_RAST$RAP), FUN = function(y) { -->
<!--   # make a map for this year -->
<!--            tempFig <- ggplot() + -->
<!--              facet_wrap(~lyr, ncol = 1) + -->
<!-- geom_spatraster(data = dat_2_RAST$RAP[[y]][[1:4]], maxcell = 5e+10) +  -->
<!--   geom_sf(data=cropped_states_2 ,fill=NA ) + -->
<!--  xlim(-140, -65) + -->
<!--  ylim(16, 50) + -->
<!--   coord_sf(datum = st_crs(test_rastTemp)) + -->
<!--              scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"),  -->
<!--                           limits = c(0,100), na.value = "white")  -->
<!--   # add an inset showing a density plot   -->
<!--            # get data -->
<!--            data_inset <- terra::values(dat_2_RAST$RAP[[y]][[1:4]]) %>%  -->
<!--                as_tibble() %>%  -->
<!--                pivot_longer(1:4, names_to = "lyr", values_to = "Cover") -->

<!--            # make insets -->
<!--            inset_plots <- data_inset %>% -->
<!--              split(f = .$lyr) %>% -->
<!--              purrr::map(~annotation_custom2( -->
<!--     grob = ggplotGrob(get_inset(.) #+ -->
<!--                         #scale_y_continuous(limits=c(0,105), breaks = c(0, 50, 100)) -->
<!--                       ), -->
<!--     data = data.frame(lyr=unique(.$lyr)), -->
<!--     ymin = 15, ymax=30, xmin=-140, xmax=-105) -->
<!--   ) -->

<!--          tempFig2 <- tempFig + inset_plots -->

<!--            return(tempFig2) -->
<!-- }) -->

<!-- ggarrange(testFigs[[1]],  -->
<!--           testFigs[[2]],  -->
<!--           ##testFigs[[3]],  -->
<!--           testFigs[[4]],  -->
<!--           nrow = 1, common.legend = TRUE, legend = "bottom" -->
<!--           ) %>%  -->
<!--   annotate_figure(top = "Map of RAP data on a dayMet x 8 grid") -->
<!-- # testFigs <- lapply(names(dat_2_RAST$RAP), FUN = function(y) { -->
<!-- #            tempFig <- ggplot() + -->
<!-- #              facet_wrap(~lyr, ncol = 1) + -->
<!-- # geom_spatraster(data = dat_2_RAST$RAP[[y]]) +  -->
<!-- #   geom_sf(data=cropped_states_2 ,fill=NA ) + -->
<!-- #              scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"),  -->
<!-- #                           limits = c(0,100)) + -->
<!-- #   xlim(-125, -65) +  -->
<!-- #   ylim(25, 50) -->
<!-- #            return(tempFig) -->
<!-- # }) -->
<!-- # ggarrange(testFigs[[1]],  -->
<!-- #           testFigs[[2]],  -->
<!-- #           testFigs[[3]],  -->
<!-- #           testFigs[[4]],  -->
<!-- #           nrow = 1, common.legend = TRUE) -->
<!-- ``` -->

<!-- ## After 'coarsifying' LANDFIRE data -->
<!-- ```{r , fig.width= 20, fig.height = 14} -->
<!-- # Read in data -->
<!-- dat_3_temp <- readRDS(file = "../../../Data_processed/CoverData/data_beforeSpatialAveraging_sampledLANDFIRE.rds") -->
<!-- # trim to be only later than 2000, which is what we want to use for further analysis -->
<!-- dat_3 <- dat_3_temp %>% -->
<!--   filter(Year >= 2000) %>%  -->
<!--   st_drop_geometry() -->

<!-- # make a multi-layer raster w/ layers for each variable of interest -->
<!-- dat_3_RAST <-  -->
<!--   lapply(c("FIA", "LANDFIRE", "LDC","RAP"),  -->
<!--          FUN = function(y) { -->
<!--            tempDat <- dat_3 %>%  -->
<!--              filter(Source == y) -->
<!--            # get vector of years for the specific data type  -->
<!--            if (y %in% c("FIA", "LDC", "RAP")) { -->
<!--              datYears <- c("2016", "2020", "2022") -->
<!--            } else { -->
<!--              datYears <- c("2003", "2007", "2015") -->
<!--            } -->
<!--            # get data for each cover type by year -->
<!--            tempOut <- apply(expand.grid(c("ShrubCover", "TotalTreeCover", "TotalHerbaceousCover", "BareGroundCover"), datYears, stringsAsFactors = FALSE),  -->
<!--                             MARGIN = 1, -->
<!--                             FUN = function(x) { -->
<!--                               tempDat2 <- tempDat %>%  -->
<!--                                 filter(Year == x["Var2"]) -->
<!--                               temp <- terra::rasterize(x = tempDat2[,c("Lon", "Lat")],  -->
<!--                                                y = test_rast, -->
<!--                                                values = tempDat2[,names(tempDat2) == x["Var1"]], -->
<!--                                                fun = mean,  -->
<!--                                                na.rm = TRUE,  -->
<!--                                                by  -->
<!--                                                ) -->
<!--                       names(temp) <-  paste0(x["Var1"], "_", x["Var2"]) -->
<!--                       return(temp) -->
<!--                     } -->
<!--                     ) -->
<!--            # get dat across all years -->
<!--            tempOut2 <- lapply(c("ShrubCover", "TotalTreeCover", "TotalHerbaceousCover", "BareGroundCover"),  -->
<!--                             FUN = function(x) { -->
<!--                               temp <- terra::rasterize(x = tempDat[,c("Lon", "Lat")],  -->
<!--                                                y = test_rast, -->
<!--                                                values = tempDat[,x], -->
<!--                                                fun = mean,  -->
<!--                                                na.rm = TRUE,  -->
<!--                                                by  -->
<!--                                                ) -->
<!--                       names(temp) <-  paste0(x, "_all") -->
<!--                       return(temp) -->
<!--                     } -->
<!--                     ) -->
<!--            # for each data type, put data for each year in one raster in the list, and the data across all years in another raster -->
<!--            outDat <- list(c(tempOut[[1]], tempOut[[2]], tempOut[[3]], tempOut[[4]]),  -->
<!--                           c(tempOut[[5]], tempOut[[6]], tempOut[[7]], tempOut[[8]]), -->
<!--                           c(tempOut[[9]], tempOut[[10]], tempOut[[11]], tempOut[[12]]),  -->
<!--                           c(tempOut2[[1]], tempOut2[[2]], tempOut2[[3]], tempOut2[[4]])) -->
<!--            names(outDat) <- c(datYears, "allYears") -->
<!--            return(outDat) -->
<!--          } -->
<!-- ) -->
<!-- names(dat_3_RAST) <- c("FIA", "LANDFIRE", "LDC","RAP") -->

<!-- ## make figure for RAP  -->
<!-- testFigs <- lapply(names(dat_3_RAST$RAP), FUN = function(y) { -->
<!--            tempFig <- ggplot() + -->
<!--              facet_wrap(~lyr, ncol = 1) + -->
<!-- geom_spatraster(data = dat_3_RAST$RAP[[y]]) +  -->
<!--   geom_sf(data=cropped_states_2 ,fill=NA ) + -->
<!--              scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"),  -->
<!--                           limits = c(0,100)) + -->
<!--   xlim(-125, -65) +  -->
<!--   ylim(25, 50) -->
<!--            return(tempFig) -->
<!-- }) -->

<!-- ggarrange(testFigs[[1]],  -->
<!--           testFigs[[2]],  -->
<!--           testFigs[[3]],  -->
<!--           testFigs[[4]],  -->
<!--           nrow = 1, common.legend = TRUE) -->
<!-- ``` -->

# AIM (BLM Assessment, Inventory and Monitoring data)
## Initial Dataset
```{r fig.width= 8, fig.height =25}
## make figure for LDC
testFigs <- lapply(names(dat_1_RAST$LDC)[4], FUN = function(y) {
  # make a map for this year
           tempFig <- 
  ggplot() +
             facet_wrap(~lyr, ncol = 1) +
geom_spatraster(data = dat_1_RAST$LDC[[y]], maxcell = 5e+10) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) +
  coord_sf(datum = st_crs(test_rastTemp)) +
             scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"), 
                          limits = c(0,100), na.value = "white") 
  # add an inset showing a density plot  
           # get data
           data_inset <- terra::values(dat_1_RAST$LDC[[y]]) %>% 
               as_tibble() %>% 
               pivot_longer(1:9, names_to = "lyr", values_to = "Cover")
           

           # make insets
           inset_plots <- data_inset %>%
             split(f = .$lyr) %>%
             purrr::map(~annotation_custom2(
    grob = ggplotGrob(get_inset(.) #+
                        #scale_y_continuous(limits=c(0,105), breaks = c(0, 50, 100))
                      ),
    data = data.frame(lyr=unique(.$lyr)),
    ymin = -2025000, ymax=-1205000, xmin=-2280000, xmax=-180000)
  )

         tempFig2 <- tempFig + inset_plots

           return(tempFig2)
})

testFigs_splitForPlotting <- lapply(names(dat_1_RAST$LDC)[4], FUN = function(y) {
  # make a map for this year
           tempFig_a <- 
  ggplot() +
             facet_wrap(~lyr, ncol = 1) +
geom_spatraster(data = dat_1_RAST$LDC[[y]][[1:4]], maxcell = 5e+10) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) +
  coord_sf(datum = st_crs(test_rastTemp)) +
             scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"), 
                          limits = c(0,100), na.value = "white") 
  # add an inset showing a density plot  
           # get data
           data_inset_a <- terra::values(dat_1_RAST$LDC[[y]][[1:4]]) %>% 
               as_tibble() %>% 
               pivot_longer(1:4, names_to = "lyr", values_to = "Cover")
           

           # make insets
           inset_plots_a <- data_inset_a %>%
             split(f = .$lyr) %>%
             purrr::map(~annotation_custom2(
    grob = ggplotGrob(get_inset(.) #+
                        #scale_y_continuous(limits=c(0,105), breaks = c(0, 50, 100))
                      ),
    data = data.frame(lyr=unique(.$lyr)),
    ymin = -2025000, ymax=-1205000, xmin=-2280000, xmax=-180000)
  )

         tempFig2_a <- tempFig_a + inset_plots_a
         
         ## second subset
                    tempFig_b <- 
  ggplot() +
             facet_wrap(~lyr, ncol = 1) +
geom_spatraster(data = dat_1_RAST$LDC[[y]][[5:7]], maxcell = 5e+10) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) +
  coord_sf(datum = st_crs(test_rastTemp)) +
             scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"), 
                          limits = c(0,100), na.value = "white") 
  # add an inset showing a density plot  
           # get data
           data_inset_b <- terra::values(dat_1_RAST$LDC[[y]][[5:7]]) %>% 
               as_tibble() %>% 
               pivot_longer(1:3, names_to = "lyr", values_to = "Cover")
           

           # make insets
           inset_plots_b <- data_inset_b %>%
             split(f = .$lyr) %>%
             purrr::map(~annotation_custom2(
    grob = ggplotGrob(get_inset(.) #+
                        #scale_y_continuous(limits=c(0,105), breaks = c(0, 50, 100))
                      ),
    data = data.frame(lyr=unique(.$lyr)),
    ymin = -2025000, ymax=-1205000, xmin=-2280000, xmax=-180000)
  )

         tempFig2_b <- tempFig_b + inset_plots_b
         
                  ## third subset
                    tempFig_c <- 
  ggplot() +
             facet_wrap(~lyr, ncol = 1) +
geom_spatraster(data = dat_1_RAST$LDC[[y]][[8:9]], maxcell = 5e+10) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) +
  coord_sf(datum = st_crs(test_rastTemp)) +
             scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"), 
                          limits = c(0,100), na.value = "white") 
  # add an inset showing a density plot  
           # get data
           data_inset_c <- terra::values(dat_1_RAST$LDC[[y]][[8:9]]) %>% 
               as_tibble() %>% 
               pivot_longer(1:2, names_to = "lyr", values_to = "Cover")
           

           # make insets
           inset_plots_c <- data_inset_c %>%
             split(f = .$lyr) %>%
             purrr::map(~annotation_custom2(
    grob = ggplotGrob(get_inset(.) #+
                        #scale_y_continuous(limits=c(0,105), breaks = c(0, 50, 100))
                      ),
    data = data.frame(lyr=unique(.$lyr)),
    ymin = -2025000, ymax=-1205000, xmin=-2280000, xmax=-180000)
  )

         tempFig2_c <- tempFig_c + inset_plots_c

           return(list(tempFig2_a, tempFig2_b, tempFig2_c))
})
jpeg(filename = "/Users/astears/Documents/Dropbox_static/Work/NAU_USGS_postdoc/cleanPED/PED_vegClimModels/Figures/CoverDatFigures/AIMInputs_step1_partA.jpeg",
    width = 800, height = 2000)
ggarrange(#testFigs[[1]], 
          #testFigs[[2]], 
          ##testFigs[[3]], 
          #testFigs[[1]], 
  testFigs_splitForPlotting[[1]][[1]],
          nrow = 1, common.legend = TRUE, legend = "bottom"
          ) %>% 
  annotate_figure(top = "Map of AIM data on a dayMet x 8 grid")
dev.off()


jpeg(filename = "/Users/astears/Documents/Dropbox_static/Work/NAU_USGS_postdoc/cleanPED/PED_vegClimModels/Figures/CoverDatFigures/AIMInputs_step1_partB.jpeg",
    width = 800, height = 1500)
  testFigs_splitForPlotting[[1]][[2]]
dev.off()


jpeg(filename = "/Users/astears/Documents/Dropbox_static/Work/NAU_USGS_postdoc/cleanPED/PED_vegClimModels/Figures/CoverDatFigures/AIMInputs_step1_partC.jpeg",
    width = 800, height = 1000)
  testFigs_splitForPlotting[[1]][[3]]
dev.off()
```

```{r}
# barplots of # of observations per data type per year
dat_1 %>% 
  st_drop_geometry() %>% 
  filter(Source == "LDC" & Year >=2010) %>%
  group_by(Year) %>% 
  summarize(ShrubCover = sum(!is.na(ShrubCover)), 
            TotalTreeCover = sum(!is.na(TotalTreeCover)),
            TotalHerbaceousCover = sum(!is.na(TotalHerbaceousCover)),
           BareGroundCover = sum(!is.na(BareGroundCover)),
            BroadLeavedTreeCover = sum(!is.na(BroadLeavedTreeCover)),
            NeedleLeavedTreeCover = sum(!is.na(NeedleLeavedTreeCover)),
            ForbCover = sum(!is.na(ForbCover)),
            C3GramCover = sum(!is.na(C3GramCover)),
            C4GramCover = sum(!is.na(C4GramCover))
            ) %>% 
  pivot_longer(2:10, values_to = "No_Observations", names_to = "functionalGroup") %>% 
  ggplot() + 
  geom_col(aes(x = Year, y = No_Observations, fill = functionalGroup), position = "dodge", col = "grey20")  + 
   scale_fill_discrete(type = c("#000000", "#E69F00" ,"#56B4E9", "#009E73", "#F0E442" ,"#0072B2" ,
                                "#D55E00", "#CC79A7", "#999999"),
                       labels = c("BareGroundCover", "BroadLeavedTreeCover", "C3GramCover","C4GramCover","ForbCover", "NeedleLeavedTreeCover",
                                  "ShrubCover",  "TotalHerbaceousCover", "TotalTreeCover")
   )
```

<!-- ## After removing burned area -->
<!-- ```{r , fig.width= 20, fig.height = 14} -->
<!-- ## make figure for LDC -->
<!-- testFigs <- lapply(names(dat_2_RAST$LDC), FUN = function(y) { -->
<!--            tempFig <- ggplot() + -->
<!--              facet_wrap(~lyr, ncol = 1) + -->
<!-- geom_spatraster(data = dat_2_RAST$LDC[[y]]) +  -->
<!--   coord_sf(datum = st_crs(test_rastTemp)) + -->
<!--   geom_sf(data=cropped_states_2 ,fill=NA ) + -->
<!--              scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"),  -->
<!--                           limits = c(0,100)) + -->
<!--   xlim(-125, -65) +  -->
<!--   ylim(25, 50) -->
<!--            return(tempFig) -->
<!-- }) -->
<!-- ggarrange(testFigs[[1]],  -->
<!--           testFigs[[2]],  -->
<!--           testFigs[[3]],  -->
<!--           testFigs[[4]],  -->
<!--           nrow = 1, common.legend = TRUE) -->

<!-- ``` -->


<!-- ## After 'coarsifying' LANDFIRE data -->
<!-- ```{r , fig.width= 20, fig.height = 14} -->
<!-- ## make figure for AIM -->
<!-- # get data just for AIM -->
<!-- AIM_3 <- dat_3 %>% -->
<!--   filter(Source == "LDC") %>% -->
<!--   pivot_longer(cols = c(ShrubCover, TotalTreeCover, TotalHerbaceousCover, BareGroundCover#, AngTrC_, CnfTrC_, C3GrmC_, C4GrmC_, FrbCvr_ -->
<!--                         ), -->
<!--                names_to = "coverType", -->
<!--                values_to = "coverValues" -->
<!--                 ) %>% -->
<!--   filter(!is.na(coverValues)) -->

<!-- ggarrange(ggplot(AIM_3 %>% -->
<!--   filter(Year %in% c(2016, 2020, 2022))) + -->
<!--   facet_grid(rows = vars(coverType), cols = vars(Year)) + -->
<!--   stat_summary_2d(aes(x = Lon, y = Lat, z = coverValues), fun = mean, binwidth = .1) + -->
<!--      scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"), -->
<!--                           limits = c(0,100)) + -->
<!--     #geom_raster(aes(x = Lon, y = Lat, fill = coverValues)) + -->
<!--     #geom_point(aes(x = Lon, y = Lat, col = coverValues)) + -->
<!--   xlim(c(-130,-65)) + -->
<!--     ylim(c(25, 50)) + -->
<!--   ggtitle("After 'coarsifying' LANDFIRE data - AIM"), -->

<!--   ggplot(AIM_3) + -->
<!--   facet_grid(rows = vars(coverType)) + -->
<!--   stat_summary_2d(aes(x = Lon, y = Lat, z = coverValues), fun = mean, binwidth = .1) + -->
<!--      scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"), -->
<!--                           limits = c(0,100)) + -->
<!--     #geom_raster(aes(x = Lon, y = Lat, fill = coverValues)) + -->
<!--     #geom_point(aes(x = Lon, y = Lat, col = coverValues)) + -->
<!--   xlim(c(-130,-65)) + -->
<!--     ylim(c(25, 50)) + -->
<!--   ggtitle("After 'coarsifying' LANDFIRE data, \nall years - AIM "), -->

<!--   ncol = 2, align = "h", -->
<!--   common.legend = TRUE, -->
<!--   widths = c(.7, .3) -->
<!-- ) -->
<!-- ``` -->


# FIA (USFS Forest Inventory and Analysis data)
## Initial Dataset
```{r fig.width= 20, fig.height = 26}
## make figure for FIA
testFigs <- lapply(names(dat_1_RAST$FIA)[4], FUN = function(y) {
  # make a map for this year
           tempFig <- 
  ggplot() +
             facet_wrap(~lyr, ncol = 1) +
geom_spatraster(data = dat_1_RAST$FIA[[y]][[c(1:5,8:9)]], maxcell = 5e+10) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) +
  coord_sf(datum = st_crs(test_rastTemp)) +
             scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"), 
                          limits = c(0,100), na.value = "white") 
  # add an inset showing a density plot  
           # get data
           data_inset <- terra::values(dat_1_RAST$FIA[[y]][[c(1:5,8:9)]]) %>% 
               as_tibble() %>% 
               pivot_longer(1:7, names_to = "lyr", values_to = "Cover")
           
           # make insets
           inset_plots <- data_inset %>%
             split(f = .$lyr) %>%
             purrr::map(~annotation_custom2(
    grob = ggplotGrob(get_inset(.) #+
                        #scale_y_continuous(limits=c(0,105), breaks = c(0, 50, 100))
                      ),
    data = data.frame(lyr=unique(.$lyr)),
    ymin = -2025000, ymax=-1205000, xmin=-2280000, xmax=-180000)
  )

         tempFig2 <- tempFig + inset_plots

           return(tempFig2)
})

testFigs_splitForPlotting <- lapply(names(dat_1_RAST$FIA)[4], FUN = function(y) {
  # make a map for this year
           tempFig_a <- 
  ggplot() +
             facet_wrap(~lyr, ncol = 1) +
geom_spatraster(data = dat_1_RAST$FIA[[y]][[c(1:4)]], maxcell = 5e+10) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) +
  coord_sf(datum = st_crs(test_rastTemp)) +
             scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"), 
                          limits = c(0,100), na.value = "white") 
  # add an inset showing a density plot  
           # get data
           data_inset_a <- terra::values(dat_1_RAST$FIA[[y]][[c(1:4)]]) %>% 
               as_tibble() %>% 
               pivot_longer(1:4, names_to = "lyr", values_to = "Cover")
           

           # make insets
           inset_plots_a <- data_inset_a %>%
             split(f = .$lyr) %>%
             purrr::map(~annotation_custom2(
    grob = ggplotGrob(get_inset(.) #+
                        #scale_y_continuous(limits=c(0,105), breaks = c(0, 50, 100))
                      ),
    data = data.frame(lyr=unique(.$lyr)),
    ymin = -2025000, ymax=-1205000, xmin=-2280000, xmax=-180000)
  )

         tempFig2_a <- tempFig_a + inset_plots_a
         
         ## second subset
                    tempFig_b <- 
  ggplot() +
             facet_wrap(~lyr, ncol = 1) +
geom_spatraster(data = dat_1_RAST$FIA[[y]][[c(5,8:9)]], maxcell = 5e+10) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) +
  coord_sf(datum = st_crs(test_rastTemp)) +
             scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"), 
                          limits = c(0,100), na.value = "white") 
  # add an inset showing a density plot  
           # get data
           data_inset_b <- terra::values(dat_1_RAST$FIA[[y]][[c(5,8:9)]]) %>% 
               as_tibble() %>% 
               pivot_longer(1:3, names_to = "lyr", values_to = "Cover")
           

           # make insets
           inset_plots_b <- data_inset_b %>%
             split(f = .$lyr) %>%
             purrr::map(~annotation_custom2(
    grob = ggplotGrob(get_inset(.) #+
                        #scale_y_continuous(limits=c(0,105), breaks = c(0, 50, 100))
                      ),
    data = data.frame(lyr=unique(.$lyr)),
    ymin = -2025000, ymax=-1205000, xmin=-2280000, xmax=-180000)
  )

         tempFig2_b <- tempFig_b + inset_plots_b
         
                
           return(list(tempFig2_a, tempFig2_b))
})
jpeg(filename = "/Users/astears/Documents/Dropbox_static/Work/NAU_USGS_postdoc/cleanPED/PED_vegClimModels/Figures/CoverDatFigures/FIAInputs_step1_partA.jpeg",
    width = 800, height = 2000)
ggarrange(#testFigs[[1]], 
          #testFigs[[2]], 
          ##testFigs[[3]], 
          #testFigs[[1]], 
  testFigs_splitForPlotting[[1]][[1]],
          nrow = 1, common.legend = TRUE, legend = "bottom"
          ) %>% 
  annotate_figure(top = "Map of FIA data on a dayMet x 8 grid")
dev.off()


jpeg(filename = "/Users/astears/Documents/Dropbox_static/Work/NAU_USGS_postdoc/cleanPED/PED_vegClimModels/Figures/CoverDatFigures/FIAInputs_step1_partB.jpeg",
    width = 800, height = 1500)
  testFigs_splitForPlotting[[1]][[2]]
dev.off()
```


```{r fig.width=9, fig.height=6}
# barplots of # of observations per data type per year
dat_1 %>% 
  st_drop_geometry() %>% 
  filter(Source == "FIA") %>%
  group_by(Year) %>% 
  summarize(ShrubCover = sum(!is.na(ShrubCover)), 
            TotalTreeCover = sum(!is.na(TotalTreeCover)),
            TotalHerbaceousCover = sum(!is.na(TotalHerbaceousCover)),
           BareGroundCover = sum(!is.na(BareGroundCover)),
            BroadLeavedTreeCover = sum(!is.na(BroadLeavedTreeCover)),
            NeedleLeavedTreeCover = sum(!is.na(NeedleLeavedTreeCover)),
            ForbCover = sum(!is.na(ForbCover))#,
            #C3GramCover = sum(!is.na(C3GramCover)),
            #C4GramCover = sum(!is.na(C4GramCover))
            ) %>% 
  pivot_longer(2:8, values_to = "No_Observations", names_to = "functionalGroup") %>% 
  ggplot() + 
  geom_col(aes(x = Year, y = No_Observations, fill = functionalGroup), position = "dodge", col = "grey20") + 
   scale_fill_discrete(type = c("#000000",
     "#E69F00" ,#"#56B4E9", "#009E73", 
     "#F0E442" ,"#0072B2" ,"#D55E00", "#CC79A7", "#999999"),
                       labels = c("BareGroundCover",
                         "BroadLeavedTreeCover", #"C3GramCover","C4GramCover",
                         "ForbCover", "NeedleLeavedTreeCover",
                                  "ShrubCover",  "TotalHerbaceousCover", "TotalTreeCover")
 )
```

<!-- ## After removing burned area -->
<!-- ```{r , fig.width= 20, fig.height = 14} -->
<!-- ## make figure for FIA -->
<!-- testFigs <- lapply(names(dat_2_RAST$FIA), FUN = function(y) { -->
<!--            tempFig <- ggplot() + -->
<!--              facet_wrap(~lyr, ncol = 1) + -->
<!-- geom_spatraster(data = dat_2_RAST$FIA[[y]]) +  -->
<!--   coord_sf(datum = st_crs(test_rastTemp)) + -->
<!--   geom_sf(data=cropped_states_2 ,fill=NA ) + -->
<!--              scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"),  -->
<!--                           limits = c(0,100)) + -->
<!--   xlim(-125, -65) +  -->
<!--   ylim(25, 50) -->
<!--            return(tempFig) -->
<!-- }) -->
<!-- ggarrange(testFigs[[1]],  -->
<!--           testFigs[[2]],  -->
<!--           testFigs[[3]],  -->
<!--           testFigs[[4]],  -->
<!--           nrow = 1, common.legend = TRUE) -->
<!-- ``` -->


<!-- ## After 'coarsifying' LANDFIRE data -->
<!-- ```{r , fig.width= 20, fig.height = 14} -->
<!-- ## make figure for FIA -->
<!-- # get data just for FIA -->
<!-- FIA_3 <- dat_3 %>% -->
<!--   filter(Source == "FIA") %>% -->
<!--   pivot_longer(cols = c(ShrubCover, TotalTreeCover, TotalHerbaceousCover, BareGroundCover#, AngTrC_, CnfTrC_, C3GrmC_, C4GrmC_, FrbCvr_ -->
<!--                         ), -->
<!--                names_to = "coverType", -->
<!--                values_to = "coverValues" -->
<!--                 ) %>% -->
<!--   filter(!is.na(coverValues)) -->

<!-- ggarrange(ggplot(FIA_3 %>% -->
<!--   filter(Year %in% c(2016, 2020, 2022))) + -->
<!--   facet_grid(rows = vars(coverType), cols = vars(Year)) + -->
<!--   stat_summary_2d(aes(x = Lon, y = Lat, z = coverValues), fun = mean, binwidth = .1) + -->
<!--      scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"), -->
<!--                           limits = c(0,100)) + -->
<!--     #geom_raster(aes(x = Lon, y = Lat, fill = coverValues)) + -->
<!--     #geom_point(aes(x = Lon, y = Lat, col = coverValues)) + -->
<!--   xlim(c(-130,-65)) + -->
<!--     ylim(c(25, 50)) + -->
<!--   ggtitle("After 'coarsifying' LANDFIRE data - FIA"), -->

<!--   ggplot(FIA_3) + -->
<!--   facet_grid(rows = vars(coverType)) + -->
<!--   stat_summary_2d(aes(x = Lon, y = Lat, z = coverValues), fun = mean, binwidth = .1) + -->
<!--      scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"), -->
<!--                           limits = c(0,100)) + -->
<!--     #geom_raster(aes(x = Lon, y = Lat, fill = coverValues)) + -->
<!--     #geom_point(aes(x = Lon, y = Lat, col = coverValues)) + -->
<!--   xlim(c(-130,-65)) + -->
<!--     ylim(c(25, 50)) + -->
<!--   ggtitle("After 'coarsifying' LANDFIRE data, \nall years - FIA "), -->

<!--   ncol = 2, align = "h", -->
<!--   common.legend = TRUE, -->
<!--   widths = c(.7, .3) -->
<!-- ) -->
<!-- ``` -->


# LANDFIRE Reference database  (use different years for LANDFIRE due to data availability)
## Initial Dataset
```{r fig.width=16, fig.height = 26, warning = FALSE}
## make figure for LANDFIRE
testFigs <- lapply(names(dat_1_RAST$LANDFIRE)[4], FUN = function(y) {
  # make a map for this year
           tempFig <- 
  ggplot() +
             facet_wrap(~lyr, ncol = 1) +
geom_spatraster(data = dat_1_RAST$LANDFIRE[[y]][[c(1:3,5:9)]], maxcell = 5e+10) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) +
  coord_sf(datum = st_crs(test_rastTemp)) +
             scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"), 
                          limits = c(0,100), na.value = "white") 
  # add an inset showing a density plot  
           # get data
           data_inset <- terra::values(dat_1_RAST$LANDFIRE[[y]][[c(1:3,5:9)]]) %>% 
               as_tibble() %>% 
               pivot_longer(1:8, names_to = "lyr", values_to = "Cover")
           
           # make insets
           inset_plots <- data_inset %>%
             split(f = .$lyr) %>%
             purrr::map(~annotation_custom2(
    grob = ggplotGrob(get_inset(.) #+
                        #scale_y_continuous(limits=c(0,105), breaks = c(0, 50, 100))
                      ),
    data = data.frame(lyr=unique(.$lyr)),
    ymin = -2025000, ymax=-1205000, xmin=-2280000, xmax=-180000)
  )

         tempFig2 <- tempFig + inset_plots

           return(tempFig2)
})

testFigs_splitForPlotting <- lapply(names(dat_1_RAST$LANDFIRE)[4], FUN = function(y) {
  # make a map for this year
           tempFig_a <- 
  ggplot() +
             facet_wrap(~lyr, ncol = 1) +
geom_spatraster(data = dat_1_RAST$LANDFIRE[[y]][[c(1:3,5)]], maxcell = 5e+10) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) +
  coord_sf(datum = st_crs(test_rastTemp)) +
             scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"), 
                          limits = c(0,100), na.value = "white") 
  # add an inset showing a density plot  
           # get data
           data_inset_a <- terra::values(dat_1_RAST$LANDFIRE[[y]][[c(1:3,5)]]) %>% 
               as_tibble() %>% 
               pivot_longer(1:4, names_to = "lyr", values_to = "Cover")
           

           # make insets
           inset_plots_a <- data_inset_a %>%
             split(f = .$lyr) %>%
             purrr::map(~annotation_custom2(
    grob = ggplotGrob(get_inset(.) #+
                        #scale_y_continuous(limits=c(0,105), breaks = c(0, 50, 100))
                      ),
    data = data.frame(lyr=unique(.$lyr)),
    ymin = -2025000, ymax=-1205000, xmin=-2280000, xmax=-180000)
  )

         tempFig2_a <- tempFig_a + inset_plots_a
         
         ## second subset
                    tempFig_b <- 
  ggplot() +
             facet_wrap(~lyr, ncol = 1) +
geom_spatraster(data = dat_1_RAST$LANDFIRE[[y]][[c(6:9)]], maxcell = 5e+10) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) +
  coord_sf(datum = st_crs(test_rastTemp)) +
             scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"), 
                          limits = c(0,100), na.value = "white") 
  # add an inset showing a density plot  
           # get data
           data_inset_b <- terra::values(dat_1_RAST$LANDFIRE[[y]][[c(6:9)]]) %>% 
               as_tibble() %>% 
               pivot_longer(1:4, names_to = "lyr", values_to = "Cover")
           

           # make insets
           inset_plots_b <- data_inset_b %>%
             split(f = .$lyr) %>%
             purrr::map(~annotation_custom2(
    grob = ggplotGrob(get_inset(.) #+
                        #scale_y_continuous(limits=c(0,105), breaks = c(0, 50, 100))
                      ),
    data = data.frame(lyr=unique(.$lyr)),
    ymin = -2025000, ymax=-1205000, xmin=-2280000, xmax=-180000)
  )

         tempFig2_b <- tempFig_b + inset_plots_b
         
                
           return(list(tempFig2_a, tempFig2_b))
})
jpeg(filename = "/Users/astears/Documents/Dropbox_static/Work/NAU_USGS_postdoc/cleanPED/PED_vegClimModels/Figures/CoverDatFigures/LANDFIREInputs_step1_partA.jpeg",
    width = 800, height = 2000)
ggarrange(#testFigs[[1]], 
          #testFigs[[2]], 
          ##testFigs[[3]], 
          #testFigs[[1]], 
  testFigs_splitForPlotting[[1]][[1]],
          nrow = 1, common.legend = TRUE, legend = "bottom"
          ) %>% 
  annotate_figure(top = "Map of LANDFIRE data on a dayMet x 8 grid")
dev.off()


jpeg(filename = "/Users/astears/Documents/Dropbox_static/Work/NAU_USGS_postdoc/cleanPED/PED_vegClimModels/Figures/CoverDatFigures/LANDFIREInputs_step1_partB.jpeg",
    width = 800, height = 2000)
  testFigs_splitForPlotting[[1]][[2]]
dev.off()
```

```{r fig.width=9, fig.height=6}
# barplots of # of observations per data type per year
dat_1 %>% 
  st_drop_geometry() %>% 
  filter(Source == "LANDFIRE") %>%
  group_by(Year) %>% 
  summarize(ShrubCover = sum(!is.na(ShrubCover)), 
            TotalTreeCover = sum(!is.na(TotalTreeCover)),
            TotalHerbaceousCover = sum(!is.na(TotalHerbaceousCover)),
           # BareGroundCover = sum(!is.na(BareGroundCover)),
            BroadLeavedTreeCover = sum(!is.na(BroadLeavedTreeCover)),
            NeedleLeavedTreeCover = sum(!is.na(NeedleLeavedTreeCover)),
            ForbCover = sum(!is.na(ForbCover)),
            C3GramCover = sum(!is.na(C3GramCover)),
            C4GramCover = sum(!is.na(C4GramCover))
            ) %>% 
  pivot_longer(2:9, values_to = "No_Observations", names_to = "functionalGroup") %>% 
  ggplot() + 
  geom_col(aes(x = Year, y = No_Observations, fill = functionalGroup), position = "dodge", col = "grey20") + 
   scale_fill_discrete(type = c(#"#000000", 
     "#E69F00" ,"#56B4E9", "#009E73", "#F0E442" ,"#0072B2" ,"#D55E00", "#CC79A7", "#999999"),
                       labels = c(#"BareGroundCover", 
                         "BroadLeavedTreeCover", "C3GramCover","C4GramCover","ForbCover", "NeedleLeavedTreeCover",
                                  "ShrubCover",  "TotalHerbaceousCover", "TotalTreeCover")
 )
```

<!-- ## After removing burned area -->
<!-- ```{r , fig.width= 20, fig.height = 14} -->
<!-- ## make figure for LANDFIRE -->
<!-- testFigs <- lapply(names(dat_2_RAST$LANDFIRE), FUN = function(y) { -->
<!--            tempFig <- ggplot() + -->
<!--              facet_wrap(~lyr, ncol = 1) + -->
<!-- geom_spatraster(data = dat_2_RAST$LANDFIRE[[y]]) +  -->
<!--   coord_sf(datum = st_crs(test_rastTemp)) + -->
<!--   geom_sf(data=cropped_states_2 ,fill=NA ) + -->
<!--              scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"),  -->
<!--                           limits = c(0,100)) + -->
<!--   xlim(-125, -65) +  -->
<!--   ylim(25, 50) -->
<!--            return(tempFig) -->
<!-- }) -->

<!-- ggarrange(testFigs[[1]],  -->
<!--           testFigs[[2]],  -->
<!--           testFigs[[3]],  -->
<!--           testFigs[[4]],  -->
<!--           nrow = 1, common.legend = TRUE) -->
<!-- ``` -->

<!-- ## After 'coarsifying' LANDFIRE data -->
<!-- ```{r , fig.width= 20, fig.height = 14} -->
<!-- ## make figure for LANDFIRE -->
<!-- LANDFIRE_3 <- dat_3 %>% -->
<!--   filter(Source == "LANDFIRE") %>% -->
<!--   pivot_longer(cols = c(ShrubCover, TotalTreeCover, TotalHerbaceousCover, BareGroundCover#, AngTrC_, CnfTrC_, C3GrmC_, C4GrmC_, FrbCvr_ -->
<!--                         ), -->
<!--                names_to = "coverType", -->
<!--                values_to = "coverValues" -->
<!--                 ) %>% -->
<!--   filter(!is.na(coverValues)) -->

<!-- ggarrange(ggplot(LANDFIRE_3 %>% -->
<!--   filter(Year %in% c(2003, 2007, 2015))) + -->
<!--   facet_grid(rows = vars(coverType), cols = vars(Year)) + -->
<!--   stat_summary_2d(aes(x = Lon, y = Lat, z = coverValues), fun = mean, binwidth = .1) + -->
<!--      scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"), -->
<!--                           limits = c(0,100)) + -->
<!--     #geom_raster(aes(x = Lon, y = Lat, fill = coverValues)) + -->
<!--     #geom_point(aes(x = Lon, y = Lat, col = coverValues)) + -->
<!--   xlim(c(-130,-65)) + -->
<!--     ylim(c(25, 50)) + -->
<!--   ggtitle("After 'coarsifying' LANDFIRE data - LANDFIRE"), -->

<!--   ggplot(LANDFIRE_3) + -->
<!--   facet_grid(rows = vars(coverType)) + -->
<!--   stat_summary_2d(aes(x = Lon, y = Lat, z = coverValues), fun = mean, binwidth = .1) + -->
<!--      scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"), -->
<!--                           limits = c(0,100)) + -->
<!--     #geom_raster(aes(x = Lon, y = Lat, fill = coverValues)) + -->
<!--     #geom_point(aes(x = Lon, y = Lat, col = coverValues)) + -->
<!--   xlim(c(-130,-65)) + -->
<!--     ylim(c(25, 50)) + -->
<!--   ggtitle("After 'coarsifying' LANDFIRE data, \nall years - LANDFIRE "), -->

<!--   ncol = 2, align = "h", -->
<!--   common.legend = TRUE, -->
<!--   widths = c(.7, .3) -->
<!-- ) -->
<!-- ``` -->

# Map of locations w/ fire history
```{r}
mtbs <- st_read(dsn = "../../../Data_raw/mtbs_perimeter_data/", layer = "mtbs_perims_DD")

#combined wildland fire 
cwf_layers <- st_layers(dsn = "../../../Data_raw/combinedWildlandFireDB/Fire_Feature_Data_Pro2_8_Geodatabase/Fire_Feature_Data.gdb/")
cwf <- st_read(dsn = "../../../Data_raw/combinedWildlandFireDB/Fire_Feature_Data_Pro2_8_Geodatabase/Fire_Feature_Data.gdb/", layer = "USGS_Wildland_Fire_Combined_Dataset")
 # transform crs to match mtbs
cwf <- cwf %>% 
  st_transform(st_crs(mtbs)) %>% 
  st_sf() %>% 
  st_cast()

## fix those geometries that are "multisurface"
cwf_multi <- cwf[st_geometry_type(cwf) == "MULTISURFACE",]
cwf_nonMulti <- cwf[st_geometry_type(cwf) != "MULTISURFACE",]

# cwf_multi_fixed <- cwf_multi %>% 
# 
#   #st_make_valid() %>% 
#   surf_to_poly()

cwf_new <- #cwf_multi_fixed %>% 
 # rbind(cwf_nonMulti) %>% 
  cwf_nonMulti

str(cwf)
library(mapview)
mapview(cwf_new)
```

# Map of locations we exclude because of land use 
```{r}

```

# After data has been spatially averaged (can no longer look at data by data source)
## After initial spatial averaging
```{r, echo = FALSE, , fig.width= 20, fig.height = 14}
# Read in data
dat_4 <- readRDS("../../../Data_processed/CoverData/spatiallyAverageData_intermediate_test5_sampledLANDFIRE.rds")

## make figure for RAP
# get data just for RAP
dat_4 <- dat_4 %>%
  pivot_longer(cols = c(ShrubCover, TotalTreeCover, TotalHerbaceousCover, BareGroundCover#, AngTrC_, CnfTrC_, C3GrmC_, C4GrmC_, FrbCvr_
                        ),
               names_to = "coverType",
               values_to = "coverValues"
                ) %>%
  filter(!is.na(coverValues))

ggarrange(ggplot(dat_4 %>%
  filter(Year %in% c(2016, 2020, 2022))) +
  facet_grid(rows = vars(coverType), cols = vars(Year)) +
  stat_summary_2d(aes(x = x, y = y, z = coverValues), fun = mean, binwidth = .1) +
     scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"),
                          limits = c(0,100)) +
    #geom_raster(aes(x = Lon, y = Lat, fill = coverValues)) +
    #geom_point(aes(x = Lon, y = Lat, col = coverValues)) +
  xlim(c(-130,-65)) +
    ylim(c(25, 50)) +
  ggtitle("Spatially Averaged Data - step 1"),

  ggplot(dat_4) +
  facet_grid(rows = vars(coverType)) +
  stat_summary_2d(aes(x = x, y = y, z = coverValues), fun = mean, binwidth = .1) +
     scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"),
                          limits = c(0,100)) +
    #geom_raster(aes(x = Lon, y = Lat, fill = coverValues)) +
    #geom_point(aes(x = Lon, y = Lat, col = coverValues)) +
  xlim(c(-130,-65)) +
    ylim(c(25, 50)) +
  ggtitle("Spatially Averaged Data - step 1 - n\all Years"),

  ncol = 2, align = "h",
  common.legend = TRUE,
  widths = c(.7, .3)
)
```
## After adding climate/weather data
```{r, echo = FALSE,  fig.width= 20, fig.height = 14}
# Read in data
dat_5_temp <- readRDS("../../../Data_processed/CoverData/DataForModels_spatiallyAveraged_sf_sampledLANDFIRE.rds")

## make figures
dat_5 <- dat_5_temp %>%
  pivot_longer(cols = c(ShrubCover, TotalTreeCover, TotalHerbaceousCover, BareGroundCover#, AngTrC_, CnfTrC_, C3GrmC_, C4GrmC_, FrbCvr_
                        ),
               names_to = "coverType",
               values_to = "coverValues"
                ) %>%
  filter(!is.na(coverValues))

ggarrange(ggplot(dat_5 %>%
  filter(Year %in% c(2016, 2020, 2022))) +
  facet_grid(rows = vars(coverType), cols = vars(Year)) +
  stat_summary_2d(aes(x = x, y = y, z = coverValues), fun = mean, binwidth = .1) +
     scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"),
                          limits = c(0,100)) +
    #geom_raster(aes(x = Lon, y = Lat, fill = coverValues)) +
    #geom_point(aes(x = Lon, y = Lat, col = coverValues)) +
  xlim(c(-130,-65)) +
    ylim(c(25, 50)) +
  ggtitle("Spatially Averaged Data - after adding climate/weather"),

  ggplot(dat_5) +
  facet_grid(rows = vars(coverType)) +
  stat_summary_2d(aes(x = x, y = y, z = coverValues), fun = mean, binwidth = .1) +
     scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"),
                          limits = c(0,100)) +
    #geom_raster(aes(x = Lon, y = Lat, fill = coverValues)) +
    #geom_point(aes(x = Lon, y = Lat, col = coverValues)) +
  xlim(c(-130,-65)) +
    ylim(c(25, 50)) +
  ggtitle("Spatially Averaged Data - \nafter adding clim/weath - n\all Years"),

  ncol = 2, align = "h",
  common.legend = TRUE,
  widths = c(.7, .3)
)
```
## After adding ecoregion data
```{r, echo = FALSE, , fig.width= 20, fig.height = 14}
# Read in data
dat_6 <- readRDS("../../../Data_processed/CoverData/DataForModels_withEcoregion_sampledLANDFIRE.rds")

## make figures
dat_6 <- dat_6 %>%
  pivot_longer(cols = c(ShrubCover, TotalTreeCover, TotalHerbaceousCover, BareGroundCover#, AngTrC_, CnfTrC_, C3GrmC_, C4GrmC_, FrbCvr_
                        ),
               names_to = "coverType",
               values_to = "coverValues"
                ) %>%
  filter(!is.na(coverValues))

ggarrange(ggplot(dat_6 %>%
  filter(Year %in% c(2016, 2020, 2022))) +
  facet_grid(rows = vars(coverType), cols = vars(Year)) +
  stat_summary_2d(aes(x = x, y = y, z = coverValues), fun = mean, binwidth = .1) +
     scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"),
                          limits = c(0,100)) +
    #geom_raster(aes(x = Lon, y = Lat, fill = coverValues)) +
    #geom_point(aes(x = Lon, y = Lat, col = coverValues)) +
  xlim(c(-130,-65)) +
    ylim(c(25, 50)) +
  ggtitle("Spatially Averaged Data - after adding ecoregion"),

  ggplot(dat_6) +
  facet_grid(rows = vars(coverType)) +
  stat_summary_2d(aes(x = x, y = y, z = coverValues), fun = mean, binwidth = .1) +
     scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"),
                          limits = c(0,100)) +
    #geom_raster(aes(x = Lon, y = Lat, fill = coverValues)) +
    #geom_point(aes(x = Lon, y = Lat, col = coverValues)) +
  xlim(c(-130,-65)) +
    ylim(c(25, 50)) +
  ggtitle("Spatially Averaged Data - \nafter adding ecoregion - n\all Years"),

  ncol = 2, align = "h",
  common.legend = TRUE,
  widths = c(.7, .3)
)
```


## After adding soils data and removing RAP points that are in ag/developed areas
```{r, echo = FALSE, , fig.width= 20, fig.height = 14}
# Read in data
dat_7_temp <- readRDS("../../../Data_processed/CoverData/DataForModels_spatiallyAveraged_withSoils_noSf_sampledLANDFIRE.rds")

## make figures
dat_7 <- dat_7_temp %>%
  pivot_longer(cols = c(ShrubCover, TotalTreeCover, TotalHerbaceousCover, BareGroundCover#, AngTrC_, CnfTrC_, C3GrmC_, C4GrmC_, FrbCvr_
                        ),
               names_to = "coverType",
               values_to = "coverValues"
                ) %>%
  filter(!is.na(coverValues))

ggarrange(ggplot(dat_7 %>%
  filter(Year %in% c(2016, 2020, 2022))) +
  facet_grid(rows = vars(coverType), cols = vars(Year)) +
  stat_summary_2d(aes(x = x, y = y, z = coverValues), fun = mean, binwidth = .1) +
     scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"),
                          limits = c(0,100)) +
    #geom_raster(aes(x = Lon, y = Lat, fill = coverValues)) +
    #geom_point(aes(x = Lon, y = Lat, col = coverValues)) +
  xlim(c(-130,-65)) +
    ylim(c(25, 50)) +
  ggtitle("Spatially Averaged Data - after adding soils"),

  ggplot(dat_7) +
  facet_grid(rows = vars(coverType)) +
  stat_summary_2d(aes(x = x, y = y, z = coverValues), fun = mean, binwidth = .1) +
     scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"),
                          limits = c(0,100)) +
    #geom_raster(aes(x = Lon, y = Lat, fill = coverValues)) +
    #geom_point(aes(x = Lon, y = Lat, col = coverValues)) +
  xlim(c(-130,-65)) +
    ylim(c(25, 50)) +
  ggtitle("Spatially Averaged Data - \nafter adding soils - n\all Years"),

  ncol = 2, align = "h",
  common.legend = TRUE,
  widths = c(.7, .3)
)
```

### Making density plots of all of the cover type by data source (after removing burned area)
First, predict forest classification using the model we later fit to the data to determine forest or non-forest
```{r}
#dat_5_temp has the climate data, so I think I should be able to add the climate data from there to the data in dat_3_temp? 
dat_7_clim <- dat_7_temp %>% 
  st_as_sf(coords = c("x", "y"), crs = st_crs(dat_5_temp), remove = FALSE) %>% 
  #cbind(dat_7_temp %>% select("x", "y")) %>% 
          st_transform(crs = st_crs(dat_3_temp), remove = FALSE) %>% 
  mutate(locationID = paste0(round(.data[["x"]], 5),"_", round(.data[["y"]], 5)))

dat_3_tempClim <- dat_3_temp %>% 
  #slice(1) %>% 
 sf::st_join(
    dat_7_clim %>% 
      select(locationID),
      join = st_nearest_feature,#st_nearest_feature,
    left = TRUE
  ) 
dat_3_tempClim <- dat_3_tempClim %>% 
  st_drop_geometry() %>% 
  left_join(dat_7_clim %>% 
              st_drop_geometry() %>% 
              select(Year, x:locationID), 
            by = c("Year" = "Year", "locationID" = "locationID")
             )

dat_3_tempClim <- 
  dat_3_tempClim %>% 
  mutate(ForestPred =  as.numeric(1/ (1 + exp(-(-0.67478165 +  
                                            1.71345065*((prcp_meanAnnTotal_CLIM - 483.34337211) / 321.52170566) + 
                                            -0.01633932*((precip_Seasonality_meanAnnAvg_CLIM - 0.97338094) / 0.23365057) + 
                                            -0.91285243*((PrecipTempCorr_meanAnnAvg_CLIM - 0.01681097) / 0.39882063) +  
                                            0.53377439*((annWetDegDays_meanAnnAvg_CLIM - 1808.61043539) / 1083.25298131) + 
                                            -0.60523936*((annWaterDeficit_95percentile_CLIM - 164.68611646) / 104.14892564) +  
                                            0.20578879*((avgCoarsePerc_acrossDepth - 9.66120017) / 10.16883531) + 
                                            -0.61568228*((totalAvailableWaterHoldingCapacity - 14.61190409) / 5.65765542) + 
                                            -0.31818631*I(((tmean_meanAnnAvg_CLIM - 11.0291275) / 4.89048541)^2) + 
                                            0.20104126*I(((precip_Seasonality_meanAnnAvg_CLIM - 0.97338094) / 0.23365057)^2) + 
                                            -0.89055646*I(((PrecipTempCorr_meanAnnAvg_CLIM - 0.01681097) / 0.39882063)^2) +  
                                            0.14226917*I(((isothermality_meanAnnAvg_CLIM - 37.76731082) / 5.2256935)^2) + 
                                            -0.11810211*I(((surfaceClay_perc - 19.59171509) / 8.51944876)^2) +  
                                            0.11278028*I(((totalAvailableWaterHoldingCapacity - 14.61190409) / 5.65765542)^2) +  
                                            0.91656456*((annWetDegDays_meanAnnAvg_CLIM - 1808.61043539) / 1083.25298131)*((annWaterDeficit_95percentile_CLIM - 164.68611646) / 104.14892564) + 
                                            -0.29650844*((precip_driestMonth_meanAnnAvg_CLIM - 3.18622575) / 5.36400552)*((isothermality_meanAnnAvg_CLIM - 37.76731082) / 5.2256935) + 
                                            -0.02061048*((PrecipTempCorr_meanAnnAvg_CLIM - 0.01681097) / 0.39882063)*((isothermality_meanAnnAvg_CLIM - 37.76731082) / 5.2256935) +
                                            -0.28504255*((isothermality_meanAnnAvg_CLIM - 37.76731082) / 5.2256935)*((tmean_meanAnnAvg_CLIM - 11.0291275) / 4.89048541) + 
                                            0.50549093*((PrecipTempCorr_meanAnnAvg_CLIM - 0.01681097) / 0.39882063)*((precip_driestMonth_meanAnnAvg_CLIM - 3.18622575) / 5.36400552) +
                                            -0.25285633*((PrecipTempCorr_meanAnnAvg_CLIM - 0.01681097) / 0.39882063)*((tmean_meanAnnAvg_CLIM - 11.0291275) / 4.89048541) + 
                                            -0.16290284*((totalAvailableWaterHoldingCapacity - 14.61190409) / 5.65765542)*((surfaceClay_perc - 19.59171509) / 8.51944876) +  
                                            0.09163249*((avgCoarsePerc_acrossDepth - 9.66120017) / 10.16883531)*((surfaceClay_perc - 19.59171509) / 8.51944876)))))) %>% 
  drop_na(ForestPred)
# ggplot(dat_3_tempClim) + 
#   geom_point(aes(x = Lon, y = Lat, col = ForestPred))

# convert the continuous probability predictions to binary using previously-identified threshold 
dat_3_tempClim <- dat_3_tempClim %>% 
  mutate(pred_binary = ForestPred,
    pred_binary = replace(pred_binary, pred_binary > 0.3819095, 1),
         pred_binary = replace(pred_binary, pred_binary <= 0.3819095, 0))

# interpolate the continuous predictions
dat_3_tempClim$preds_interp <- NA
dat_3_tempClim[!is.na(dat_3_tempClim$ForestPred), "preds_interp"] <- approxfun_pred(dat_3_tempClim[ !is.na(dat_3_tempClim$ForestPred), "ForestPred"])

```

```{r}
# data are in the dat_3_temp data.frame
# are there any locations where we have NL and BL tree but not total tree? 
#dat_3_temp[is.finite(dat_3_temp$AngioTreeCover_prop) & is.na(dat_3_temp$TotalTreeCover),]

  ggplot() + 
    facet_wrap(vars(Source, pred_binary), ncol = 2, scales = "free") + 
    geom_density(data = dat_3_tempClim, aes(TotalTreeCover, col = Source)) + 
    geom_density(data = dat_3_tempClim[is.finite(dat_3_tempClim$AngioTreeCover_prop),], aes(TotalTreeCover, col = Source),lty = 2) + 
    ggtitle("Density plots of total tree cover data (prior to spatial averaging) \nsolid lines = all available total tree cover data \ndashed line = total tree cover data in locations where we also have \nNL and BL tree cover data")
  
kable(x = data.frame("FIA" = c(table(dat_3_temp[is.finite(dat_3_temp$TotalTreeCover), "Source"] %>% st_drop_geometry())["FIA"], 
                                table(dat_3_temp[is.finite(dat_3_temp$AngioTreeCover_prop) ,"Source"] %>% st_drop_geometry())["FIA"]), 
                     "LANDFIRE" = c(table(dat_3_temp[is.finite(dat_3_temp$TotalTreeCover), "Source"] %>% st_drop_geometry() )["LANDFIRE"],
                                table(dat_3_temp[is.finite(dat_3_temp$AngioTreeCover_prop) ,"Source"] %>% st_drop_geometry())["LANDFIRE"]),
                     "AIM" = c(table(dat_3_temp[is.finite(dat_3_temp$TotalTreeCover), "Source"] %>% st_drop_geometry() )["LDC"],
                                table(dat_3_temp[is.finite(dat_3_temp$AngioTreeCover_prop) ,"Source"] %>% st_drop_geometry())["LDC"]),
                     "RAP" = c(table(dat_3_temp[is.finite(dat_3_temp$TotalTreeCover), "Source"] %>% st_drop_geometry() )["RAP"], NA),
                     row.names = c("allTotalTreeCoverData", "completeTreeData")
                     ), 
      row.names = TRUE,
      format = "html"
) %>% 
kable_styling(bootstrap_options = c("striped", "hover", "condensed")) 
  
# maps 
## locations where we have observations for total tree cover vs. locations where we have observations for all tree types 
dat_3_temp$totalTreeDataAvail <- NA
dat_3_temp$completeTreeDataAvail <- NA
dat_3_temp[!is.na(dat_3_temp$TotalTreeCover), "totalTreeDataAvail"] <- TRUE
dat_3_temp[!is.na(dat_3_temp$TotalTreeCover) & !is.na(dat_3_temp$ConifTreeCover_prop) & !is.na(dat_3_temp$AngioTreeCover_prop), "completeTreeDataAvail"] <- TRUE
ggarrange(
  ggplot(data = dat_3_temp %>% filter(!is.na(totalTreeDataAvail))) + 
    facet_grid(row = vars(Source)) + 
    geom_point(aes(x = Lon, y = Lat,  col = Source), pch = 20) + 
    ggtitle("Locations with total tree cover data"),
 ggarrange( ggplot(data = dat_3_temp  %>% filter(!is.na(completeTreeDataAvail))) +
    facet_grid(row = vars(Source)) +  
    geom_point(aes(x = Lon, y = Lat,  col = Source), pch = 20) + 
    ggtitle("Locations with total tree cover data + NL and BL tree cover data"),
    ggplot() + geom_point(), ncol = 1, heights = c(.75, .25)), 
  align = "h"
)
```

<!-- #Initial data, across all years and all data types -->
<!-- ```{r fig.width= 9, fig.height = 6} -->
<!-- # Inital FIA data, all years by year -->
<!-- ggplot(dat_1temp %>% -->
<!--          filter(Source == "FIA")) + -->
<!--   facet_wrap(vars(Year)) + -->
<!--   stat_summary_2d(aes(x = Lon, y = Lat, z = TtlTrCv), fun = mean, binwidth = .1) + -->
<!--      scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"), -->
<!--                           limits = c(0,100)) + -->
<!--     #geom_raster(aes(x = Lon, y = Lat, fill = coverValues)) + -->
<!--     #geom_point(aes(x = Lon, y = Lat, col = coverValues)) + -->
<!--   xlim(c(-130,-65)) + -->
<!--     ylim(c(25, 50)) + -->
<!--   ggtitle("Initial Data: Total Tree Cover from FIA") -->
<!-- ``` -->