---
title: "Visualizations of Model Parameters in Veg.-Climate Relationship Models"
author: "Alice Stears"
date: "`r lubridate::today()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE,
                      message = FALSE)
```

# Dependencies 

```{r warning=FALSE, message=FALSE}
library(tidyverse)
library(GGally) # for ggpairs()
library(pdp) # for partial dependence plots
library(gridExtra)
library(knitr)
library(patchwork) # for figure insets etc. 
library(ggtext)
library(betareg)
library(ggplot2)
theme_set(theme_classic())
```

# read in data

Data compiled in the `prepDataForModels.R` script

```{r}
modDat <- readRDS("../../../data/DataForModels.rds")
```

Here are the climate variables we could potentially use in the models 
```{r}
kable(names(modDat)[c(24:41,100:138)])
```

# Correlation of potential predictors 
This is using a subset of the data (50,000 rows), just for runtime purposes 

Below is the correlation between only climate predictors that are averaged across 30 years
 * Dropped tmin, tmax, t_warmest month, and t_coldest month and replaced w/ MAT
 * Drop prcp wettest month - use MAP only
 * dropped precip of driest month--was highly correlated w/ precip and precip seasonality
 * drop prcpSeasonality -- highly correlated w/ water deficit and wet degree days
 * Replace VPDmean with VPDmax
 * Probably drop FreezeMon
** also dropped vp (which we didn't talk about, but is highly correlated w/ VPD and pretty highly correlated w/ precip)


```{r}
(corrPlot_30yrMean <- 
   modDat %>% 
  rename("swe" = swe_meanAnnAvg_30yr, "tmean" = tmean_meanAnnAvg_30yr, tmax = tmax_meanAnnAvg_30yr, "vp" = vp_meanAnnAvg_30yr, "prcp" = prcp_meanAnnTotal_30yr, T_warmest = T_warmestMonth_meanAnnAvg_30yr, T_coldest = T_coldestMonth_meanAnnAvg_30yr, prcpWettest = precip_wettestMonth_meanAnnAvg_30yr, prcpDriest = precip_driestMonth_meanAnnAvg_30yr, prcpSeasonality = precip_Seasonality_meanAnnAvg_30yr, prcpTempCorr = PrecipTempCorr_meanAnnAvg_30yr, abvFreezeMonth = aboveFreezing_month_meanAnnAvg_30yr, isothermality = isothermality_meanAnnAvg_30yr, WatDeficit = annWaterDeficit_meanAnnAvg_30yr, WetDegDays = annWetDegDays_meanAnnAvg_30yr, VPDmean = annVPD_mean_meanAnnAvg_30yr, VPD_max = annVPD_max_meanAnnAvg_30yr, VPD_min = annVPD_min_meanAnnAvg_30yr) %>% 
select(swe, tmean, prcp,  prcpTempCorr, isothermality, WatDeficit, 
       WetDegDays, VPD_max ) %>% 
  slice_sample(n = 5e4) %>% 
  #select(-matches("_")) %>% 
ggpairs(lower = list(continuous = GGally::wrap("points", alpha = 0.1, size=0.2))))
```
spatial distribution of VPD_max
```{r}
# raster to get the dayMet resolution we want for visualization
test_rast <-  rast("../../../data/dayMet/rawMonthlyData/orders/70e0da02b9d2d6e8faa8c97d211f3546/Daymet_Monthly_V4R1/data/daymet_v4_prcp_monttl_na_1980.tif") %>% 
  terra::aggregate(fact = 16, fun = "mean")

# rasterize data
plotRast_annVPD <- modDat %>% 
         drop_na(annVPD_max_meanAnnAvg_30yr) %>% 
  #slice_sample(n = 5e4) %>%
  terra::vect(geom = c("Lon", "Lat")) %>% 
  terra::set.crs(crs(test_rast)) %>% 
  terra::rasterize(y = test_rast, 
                   field = "annVPD_max_meanAnnAvg_30yr", 
                   fun = mean) %>% 
  terra::crop(ext(-2000000, 2500000, -2000000, 1200000))
terra::plot(plotRast_annVPD, main = "VPD Max (30 yr avg.) at 16km2 scale", clip = TRUE, 
            plg = list(title = "VPD Max (KPa)"))

```

spatial distribution of Wet degree days
```{r}
# rasterize data
plotRast_wetDD <- modDat %>% 
         drop_na(annWetDegDays_meanAnnAvg_30yr) %>% 
  #slice_sample(n = 5e4) %>%
  terra::vect(geom = c("Lon", "Lat")) %>% 
  terra::set.crs(crs(test_rast)) %>% 
  terra::rasterize(y = test_rast, 
                   field = "annWetDegDays_meanAnnAvg_30yr", 
                   fun = mean) %>% 
  terra::crop(ext(-2000000, 2500000, -2000000, 1200000))
terra::plot(plotRast_wetDD, main = "Wet Deg. Days (30 yr avg.) at 16km2 scale", clip = TRUE, 
            plg = list(title = "Deg. Days"), 
            col = map.pal("ryb"))

```
spatial distribution of water deficit
```{r}
# rasterize data
plotRast_watDef <- modDat %>% 
         drop_na(annWaterDeficit_meanAnnAvg_30yr) %>% 
  #slice_sample(n = 5e4) %>%
  terra::vect(geom = c("Lon", "Lat")) %>% 
  terra::set.crs(crs(test_rast)) %>% 
  terra::rasterize(y = test_rast, 
                   field = "annWaterDeficit_meanAnnAvg_30yr", 
                   fun = mean) %>% 
  terra::crop(ext(-2000000, 2500000, -2000000, 1200000))
terra::plot(plotRast_watDef, main = "Water Deficit (30 yr avg.) at 16km2 scale", clip = TRUE, 
            plg = list(title = "Deg. Days"), 
            col = map.pal("plasma"))

```

visualize the residuals from the 'best mod' for total tree cover
```{r}
treeMod <- betareg(TotalTreeCover_dec ~ stats::poly(I(log10(I(swe_meanAnnAvg_5yr+1))),2,raw=TRUE) + stats::poly(tmin_meanAnnAvg_5yr,2,raw=TRUE) + stats::poly(prcp_meanAnnTotal_5yr,2,raw=TRUE) + stats::poly(I(sqrt(precip_Seasonality_meanAnnAvg_5yr)),2,raw=TRUE) + stats::poly(PrecipTempCorr_meanAnnAvg_5yr,2, raw = TRUE), data = modDat %>% slice_sample(n = 50000), link = c("logit"), link.phi = NULL,
                              type = c("ML"))

# predict responses
response <- "TotalTreeCover_dec"
predict_by_response <- function(mod, df) {
  df_out <- df

  response_name <- paste0(response, "_pred")
  df_out[[response_name]] <- predict(mod, df, type = 'response')
  df_out
}
testDat <- modDat %>% slice_sample(n = 50000)
pred_glm1 <- predict_by_response(treeMod, testDat)
pred_glm1$resid <- pred_glm1$TotalTreeCover_dec - pred_glm1$TotalTreeCover_dec_pred


# rasterize
test_rast <-  rast("../../../data/dayMet/rawMonthlyData/orders/70e0da02b9d2d6e8faa8c97d211f3546/Daymet_Monthly_V4R1/data/daymet_v4_prcp_monttl_na_1980.tif") %>% 
  terra::aggregate(fact = 32, fun = "mean")

# rasterize data
plotResid_rast <- pred_glm1 %>% 
         drop_na(resid) %>% 
  #slice_sample(n = 5e4) %>%
  terra::vect(geom = c("Lon", "Lat")) %>% 
  terra::set.crs(crs(test_rast)) %>% 
  terra::rasterize(y = test_rast, 
                   field = "resid", 
                   fun = mean) %>% 
  terra::crop(ext(-2000000, 2500000, -2000000, 1200000))
terra::plot(plotResid_rast, main = "Residuals from total tree cover model at 32km2 scale", clip = TRUE, 
            plg = list(title = "residual"), 
            col = map.pal("curvature"))
```

