---
title: "Visualizations of Model Parameters in Veg.-Climate Relationship Models"
author: "Alice Stears"
date: "`r lubridate::today()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE,
                      message = FALSE)
```

# Dependencies 

```{r warning=FALSE, message=FALSE}
library(tidyverse)
library(GGally) # for ggpairs()
library(pdp) # for partial dependence plots
library(gridExtra)
library(knitr)
library(patchwork) # for figure insets etc. 
library(ggtext)
library(betareg)
library(ggplot2)
library(terra)
library(sf)
theme_set(theme_classic())
```

# read in data

Data compiled in the `prepDataForModels.R` script

```{r}
modDat <- readRDS("../../../Data_processed/CoverData/DataForModels_spatiallyAveraged_withSoils_noSf.rds")
```

Here are the climate variables (30-year lag) we could potentially use in the models -- however I think we should remove the 5th and 95th percentile variables here, since we're interested in long-term averages rather than extremes 
```{r, echo = FALSE}
kable(modDat %>% 
        select(swe_meanAnnAvg_30yr:durationFrostFreeDays_meanAnnAvg_30yr) %>% 
        names()
)
```

Here are the weather anomaly variables (5-year lag) we could potentially use in the models 
```{r, echo = FALSE}
kable(modDat %>% 
        select(swe_meanAnn_5yrAnom:durationFrostFreeDays_meanAnnAvg_5yrAnom) %>% 
        names()
)
```

And here are the soil variables we could potentially use in the models 
```{r, echo = FALSE}
kable(modDat %>% 
        select(soilDepth:totalAvailableWaterHoldingCapacity) %>% 
        names()
)
```

# Correlation of potential weather predictors 
This is using a subset of the data (50,000 rows), just for runtime purposes 

Below is the correlation between only climate predictors that are averaged across 30 years

 * excluded 95th and 5th percentile variables 
 * Dropped tmin, tmax, t_warmest month, and t_coldest month and replaced w/ MAT
 * Dropped month above freezing (highly correlated with frost-free days in all ecoregions)

<!--  * Drop prcp wettest month - use MAP only -->
<!--  * dropped precip of driest month--was highly correlated w/ precip and precip seasonality -->
<!--  * drop prcpSeasonality -- highly correlated w/ water deficit and wet degree days -->
<!--  * Replace VPDmean with VPDmax -->
<!--  * Probably drop FreezeMon -->
<!-- ** also dropped vp (which we didn't talk about, but is highly correlated w/ VPD and pretty highly correlated w/ precip) -->


```{r, fig.dim = c(10,10)}

# function to add colors to correlation plot
 my_fn <- function(data, mapping, method="p", use="pairwise", ...){

              # grab data
              x <- eval_data_col(data, mapping$x)
              y <- eval_data_col(data, mapping$y)

              # calculate correlation
              corr <- cor(x, y, method=method, use=use)

              # calculate colour based on correlation value
              # Here I have set a correlation of minus one to blue, 
              # zero to white, and one to red 
              # Change this to suit: possibly extend to add as an argument of `my_fn`
              colFn <- colorRampPalette(c("red", "white", "blue"), interpolate ='spline')
              fill <- colFn(100)[findInterval(corr, seq(-1, 1, length=100))]

              ggally_cor(data = data, mapping = mapping, stars = FALSE, 
                         digits = 2, colour = I("black"),...) + 
                theme_void() +
                theme(panel.background = element_rect(fill=fill))
              
 }

(corrPlot_30yrMean <- 
   modDat %>% 
   select(c(swe_meanAnnAvg_30yr, tmean_meanAnnAvg_30yr:prcp_meanAnnTotal_30yr,
            precip_wettestMonth_meanAnnAvg_30yr:PrecipTempCorr_meanAnnAvg_30yr, 
            isothermality_meanAnnAvg_30yr:annVPD_min_meanAnnAvg_30yr, 
            durationFrostFreeDays_meanAnnAvg_30yr)) %>% 
  rename("swe" = swe_meanAnnAvg_30yr,  "tmean" = tmean_meanAnnAvg_30yr,
         "vp" = vp_meanAnnAvg_30yr, "prcp" = prcp_meanAnnTotal_30yr, 
         "prcp \n Wettest" = precip_wettestMonth_meanAnnAvg_30yr, "prcpDriest" = precip_driestMonth_meanAnnAvg_30yr, 
         "prcp \n Seasonality" = precip_Seasonality_meanAnnAvg_30yr, "prcp \n TempCorr" = PrecipTempCorr_meanAnnAvg_30yr,  "isothermality" = isothermality_meanAnnAvg_30yr,
         "Wat \n Deficit" = annWaterDeficit_meanAnnAvg_30yr, "Wet \n DegDays" = annWetDegDays_meanAnnAvg_30yr, 
         "VPD \n mean" = annVPD_mean_meanAnnAvg_30yr, "VPD \n max" = annVPD_max_meanAnnAvg_30yr, "VPD \n min" = annVPD_min_meanAnnAvg_30yr,
         "frost \n FreeDays" = durationFrostFreeDays_meanAnnAvg_30yr) %>% 
  slice_sample(n = 5e4) %>% 
  #select(-matches("_")) %>% 
ggpairs( upper = list(continuous = my_fn), lower = list(continuous = GGally::wrap("points", alpha = 0.1, size=0.2)), progress = FALSE))

```

# Correlations at Ecoregion scales

### Here are the ecoregions, which are grouped Level 1 Ecoregions 

```{r, echo = FALSE}
# get data for model fitting
modRegions <- readRDS("../../../Data_processed/CoverData/DataForModels_withSubEcoreg.rds")
# also get data for ecoregions
ecoRegions <- sf::st_read("../../../Data_raw/Level1Ecoregions/", layer = "NA_CEC_Eco_Level1")
```

```{r, echo = FALSE}
# filter the ecoregion data by CONUS boundary
# add the ecsoregions together as we've done in the model data
CONUS <- st_read("../../../Data_raw/CONUS_extent", "CONUS_boundary") %>% 
  st_transform(st_crs(ecoRegions)) %>% 
  st_make_valid()
# trim the ecoregions by CONUS extent
ecoRegCrop <- ecoRegions %>% 
  st_make_valid() %>% 
  st_intersection( CONUS)

unique(st_drop_geometry(modRegions)[,c("NA_L1NAME", "newRegion")])
ecoRegCrop$newRegion <- NA
ecoRegCrop[ecoRegCrop$NA_L1NAME %in% c("NORTHWESTERN FORESTED MOUNTAINS", "TEMPERATE SIERRAS", "MARINE WEST COAST FOREST"), "newRegion"] <- "Western Forests"

ecoRegCrop[ecoRegCrop$NA_L1NAME %in% c("EASTERN TEMPERATE FORESTS", "NORTHERN FORESTS", "TROPICAL WET FORESTS"), "newRegion"] <- "Eastern Forests"

ecoRegCrop[ecoRegCrop$NA_L1NAME %in% c("NORTH AMERICAN DESERTS", "SOUTHERN SEMIARID HIGHLANDS", "GREAT PLAINS", "MEDITERRANEAN CALIFORNIA"), "newRegion"] <- "Grass and Shrubland"

ecoReg <- ecoRegCrop %>% 
  aggregate(by = list(ecoRegCrop$newRegion),
            FUN = unique,
            do_union = TRUE)
saveRDS(ecoReg, "../../../Data_processed/CoverData/ecoRegionExtents.RDS")

ggplot(data = ecoReg) + 
  geom_sf(aes(fill = newRegion)) + 
  scale_fill_discrete(guide = 
  guide_legend(title = c("Ecoregion")))

# pdf(file = "../../../figures/EcoregionMap.pdf")
# ggplot(data = ecoReg) + 
#   geom_sf(aes(fill = newRegion)) + 
#   scale_fill_discrete(guide = 
#   guide_legend(title = c("Ecoregion")))
# dev.off()
```

#### Dry shrublands and grasslands

 * excluded 95th and 5th percentile variables 
 * Dropped tmin, tmax, t_warmest month, and t_coldest month and replaced w/ MAT
 * Dropped month above freezing (highly correlated with frost-free days in all ecoregions)
 * removed precip of the wettest month (correlated w/ precip and precip of the driest month)
 * remove all of the VPD values (highly correlated w/ tmean)
 
```{r, fig.dim = c(10,10), echo = FALSE}
#modDat_ecoregions <- readRDS("../../../data/DataForModels_withEcoregion.rds")
# just use modDat data
# get data 
 plotDat <- 
   modDat %>% 
   filter(newRegion == "dryShrubGrass") %>% 
   #st_drop_geometry() %>% 
  select(c(swe_meanAnnAvg_30yr, tmean_meanAnnAvg_30yr,                     
           vp_meanAnnAvg_30yr,prcp_meanAnnTotal_30yr, precip_driestMonth_meanAnnAvg_30yr,        
           precip_Seasonality_meanAnnAvg_30yr,  PrecipTempCorr_meanAnnAvg_30yr, 
           isothermality_meanAnnAvg_30yr, annWaterDeficit_meanAnnAvg_30yr,  annWetDegDays_meanAnnAvg_30yr,             
           #annVPD_mean_meanAnnAvg_30yr, 
           #annVPD_max_meanAnnAvg_30yr,                
           #annVPD_min_meanAnnAvg_30yr,  
           durationFrostFreeDays_meanAnnAvg_30yr)) %>% 
  rename("swe" = swe_meanAnnAvg_30yr,  "tmean" = tmean_meanAnnAvg_30yr,
         "vp" = vp_meanAnnAvg_30yr, "prcp" = prcp_meanAnnTotal_30yr, 
         "prcpDriest" = precip_driestMonth_meanAnnAvg_30yr, 
         "prcp \n Seasonality" = precip_Seasonality_meanAnnAvg_30yr, "prcp \n TempCorr" = PrecipTempCorr_meanAnnAvg_30yr,  "isothermality" = isothermality_meanAnnAvg_30yr,
         "Wat \n Deficit" = annWaterDeficit_meanAnnAvg_30yr, "Wet \n DegDays" = annWetDegDays_meanAnnAvg_30yr, 
         #"VPD \n mean" = annVPD_mean_meanAnnAvg_30yr, 
         #"VPD \n max" = annVPD_max_meanAnnAvg_30yr, 
         #"VPD \n min" = annVPD_min_meanAnnAvg_30yr,
         "frost \n FreeDays" = durationFrostFreeDays_meanAnnAvg_30yr) %>% 
     drop_na(swe, tmean, prcp,  "prcp \n TempCorr", isothermality, "Wat \n Deficit", 
       "Wet \n DegDays", #"VPD \n max" 
       ) %>% 
  slice_sample(n = 5e5) 
  #select(-matches("_")) %>% 
 
  p1 <- ggpairs(plotDat, upper = list(continuous = my_fn),
                lower = list(continuous = GGally::wrap("points", alpha = 0.1, size=0.2)), title = "Dry shrubland and grassland", progress = FALSE)
print(p1)
```

#### Eastern forests
 
 * excluded 95th and 5th percentile variables 
 * Dropped tmin, tmax, t_warmest month, and t_coldest month and replaced w/ MAT
 * Dropped month above freezing (highly correlated with frost-free days in all ecoregions)
 * removed precip of the wettest month (correlated w/ precip and precip of the driest month and vp)
  * remove all of the VPD values (highly correlated w/ tmean)
  
```{r, fig.dim = c(10,10), echo = FALSE}
#modDat_ecoregions <- readRDS("../../../data/DataForModels_withEcoregion.rds")
# just use modDat data
# get data 
 plotDat <- 
   modDat %>% 
   filter(newRegion == "eastForest") %>% 
   #st_drop_geometry() %>% 
  select(c(swe_meanAnnAvg_30yr, tmean_meanAnnAvg_30yr,                     
           vp_meanAnnAvg_30yr,prcp_meanAnnTotal_30yr, precip_driestMonth_meanAnnAvg_30yr,        
           precip_Seasonality_meanAnnAvg_30yr,  PrecipTempCorr_meanAnnAvg_30yr, 
           isothermality_meanAnnAvg_30yr, annWaterDeficit_meanAnnAvg_30yr,  annWetDegDays_meanAnnAvg_30yr,             
           #annVPD_mean_meanAnnAvg_30yr, annVPD_max_meanAnnAvg_30yr,                
           #annVPD_min_meanAnnAvg_30yr,  
           durationFrostFreeDays_meanAnnAvg_30yr)) %>% 
  rename("swe" = swe_meanAnnAvg_30yr,  "tmean" = tmean_meanAnnAvg_30yr,
         "vp" = vp_meanAnnAvg_30yr, "prcp" = prcp_meanAnnTotal_30yr, 
          "prcpDriest" = precip_driestMonth_meanAnnAvg_30yr, 
         "prcp \n Seasonality" = precip_Seasonality_meanAnnAvg_30yr, "prcp \n TempCorr" = PrecipTempCorr_meanAnnAvg_30yr,  "isothermality" = isothermality_meanAnnAvg_30yr,
         "Wat \n Deficit" = annWaterDeficit_meanAnnAvg_30yr, "Wet \n DegDays" = annWetDegDays_meanAnnAvg_30yr, 
         #"VPD \n mean" = annVPD_mean_meanAnnAvg_30yr, "VPD \n max" = annVPD_max_meanAnnAvg_30yr, "VPD \n min" = annVPD_min_meanAnnAvg_30yr,
         "frost \n FreeDays" = durationFrostFreeDays_meanAnnAvg_30yr) %>% 
     drop_na(swe, tmean, prcp,  "prcp \n TempCorr", isothermality, "Wat \n Deficit", 
       "Wet \n DegDays"#, "VPD \n max" 
       ) %>% 
  slice_sample(n = 5e5) 
  #select(-matches("_")) %>% 
 
  p1 <- ggpairs(plotDat, upper = list(continuous = my_fn),
                lower = list(continuous = GGally::wrap("points", alpha = 0.1, size=0.2)), 
                title = "Eastern Forests", progress = FALSE)
print(p1)
```

#### Western forests

 * excluded 95th and 5th percentile variables 
 * Dropped tmin, tmax, t_warmest month, and t_coldest month and replaced w/ MAT
 * Dropped month above freezing (highly correlated with frost-free days in all ecoregions)
 * removed precip of the wettest month (correlated w/ precip and precip of the driest month and vp)
 * remove all of the VPD values (highly correlated w/ tmean)
 
```{r, fig.dim = c(10,10)}
#modDat_ecoregions <- readRDS("../../../data/DataForModels_withEcoregion.rds")
# just use modDat data
# get data 
 plotDat <- 
   modDat %>% 
   filter(newRegion == "westForest") %>% 
   #st_drop_geometry() %>% 
   select(c(swe_meanAnnAvg_30yr, tmean_meanAnnAvg_30yr,                     
           vp_meanAnnAvg_30yr,prcp_meanAnnTotal_30yr, precip_driestMonth_meanAnnAvg_30yr,        
           precip_Seasonality_meanAnnAvg_30yr,  PrecipTempCorr_meanAnnAvg_30yr, 
           isothermality_meanAnnAvg_30yr, annWaterDeficit_meanAnnAvg_30yr,  annWetDegDays_meanAnnAvg_30yr,             
           #annVPD_mean_meanAnnAvg_30yr, annVPD_max_meanAnnAvg_30yr,                
           #annVPD_min_meanAnnAvg_30yr,  
           durationFrostFreeDays_meanAnnAvg_30yr)) %>% 
  rename("swe" = swe_meanAnnAvg_30yr,  "tmean" = tmean_meanAnnAvg_30yr,
         "vp" = vp_meanAnnAvg_30yr, "prcp" = prcp_meanAnnTotal_30yr, 
          "prcpDriest" = precip_driestMonth_meanAnnAvg_30yr, 
         "prcp \n Seasonality" = precip_Seasonality_meanAnnAvg_30yr, "prcp \n TempCorr" = PrecipTempCorr_meanAnnAvg_30yr,  "isothermality" = isothermality_meanAnnAvg_30yr,
         "Wat \n Deficit" = annWaterDeficit_meanAnnAvg_30yr, "Wet \n DegDays" = annWetDegDays_meanAnnAvg_30yr, 
         #"VPD \n mean" = annVPD_mean_meanAnnAvg_30yr, "VPD \n max" = annVPD_max_meanAnnAvg_30yr, "VPD \n min" = annVPD_min_meanAnnAvg_30yr,
         "frost \n FreeDays" = durationFrostFreeDays_meanAnnAvg_30yr) %>% 
     drop_na(swe, tmean, prcp,  "prcp \n TempCorr", isothermality, "Wat \n Deficit", 
       "Wet \n DegDays"#, "VPD \n max" 
       ) %>% 
  slice_sample(n = 5e5) 
  #select(-matches("_")) %>% 
 
  p1 <- ggpairs(plotDat, upper = list(continuous = my_fn),
                lower = list(continuous = GGally::wrap("points", alpha = 0.1, size=0.2)), 
                title = "Western Forests", progress = FALSE)
print(p1)
```

<!-- # Visualization of some predictors -->
<!-- spatial distribution of VPD_max -->
<!-- ```{r} -->
<!-- # raster to get the dayMet resolution we want for visualization -->
<!-- test_rast <-  rast("../../../data/dayMet/rawMonthlyData/orders/70e0da02b9d2d6e8faa8c97d211f3546/Daymet_Monthly_V4R1/data/daymet_v4_prcp_monttl_na_1980.tif") %>%  -->
<!--   terra::aggregate(fact = 16, fun = "mean") -->

<!-- # rasterize data -->
<!-- plotRast_annVPD <- modDat %>%  -->
<!--          drop_na(annVPD_max_meanAnnAvg_30yr) %>%  -->
<!--   #slice_sample(n = 5e4) %>% -->
<!--   terra::vect(geom = c("Lon", "Lat")) %>%  -->
<!--   terra::set.crs(crs(test_rast)) %>%  -->
<!--   terra::rasterize(y = test_rast,  -->
<!--                    field = "annVPD_max_meanAnnAvg_30yr",  -->
<!--                    fun = mean) %>%  -->
<!--   terra::crop(ext(-2000000, 2500000, -2000000, 1200000)) -->
<!-- terra::plot(plotRast_annVPD, main = "VPD Max (30 yr avg.) at 16km2 scale", clip = TRUE,  -->
<!--             plg = list(title = "VPD Max (KPa)")) -->

<!-- ``` -->

<!-- spatial distribution of Wet degree days -->
<!-- ```{r} -->
<!-- # rasterize data -->
<!-- plotRast_wetDD <- modDat %>%  -->
<!--          drop_na(annWetDegDays_meanAnnAvg_30yr) %>%  -->
<!--   #slice_sample(n = 5e4) %>% -->
<!--   terra::vect(geom = c("Lon", "Lat")) %>%  -->
<!--   terra::set.crs(crs(test_rast)) %>%  -->
<!--   terra::rasterize(y = test_rast,  -->
<!--                    field = "annWetDegDays_meanAnnAvg_30yr",  -->
<!--                    fun = mean) %>%  -->
<!--   terra::crop(ext(-2000000, 2500000, -2000000, 1200000)) -->
<!-- terra::plot(plotRast_wetDD, main = "Wet Deg. Days (30 yr avg.) at 16km2 scale", clip = TRUE,  -->
<!--             plg = list(title = "Deg. Days"),  -->
<!--             col = map.pal("ryb")) -->

<!-- ``` -->
<!-- spatial distribution of water deficit -->
<!-- ```{r} -->
<!-- # rasterize data -->
<!-- plotRast_watDef <- modDat %>%  -->
<!--          drop_na(annWaterDeficit_meanAnnAvg_30yr) %>%  -->
<!--   #slice_sample(n = 5e4) %>% -->
<!--   terra::vect(geom = c("Lon", "Lat")) %>%  -->
<!--   terra::set.crs(crs(test_rast)) %>%  -->
<!--   terra::rasterize(y = test_rast,  -->
<!--                    field = "annWaterDeficit_meanAnnAvg_30yr",  -->
<!--                    fun = mean) %>%  -->
<!--   terra::crop(ext(-2000000, 2500000, -2000000, 1200000)) -->
<!-- terra::plot(plotRast_watDef, main = "Water Deficit (30 yr avg.) at 16km2 scale", clip = TRUE,  -->
<!--             plg = list(title = "Deg. Days"),  -->
<!--             col = map.pal("plasma")) -->

<!-- ``` -->

<!-- # visualize the residuals from the 'best mod' for total tree cover -->
<!-- ```{r} -->
<!-- treeMod <- betareg(TotalTreeCover_dec ~ stats::poly(I(log10(I(swe_meanAnnAvg_5yr+1))),2,raw=TRUE) + stats::poly(tmin_meanAnnAvg_5yr,2,raw=TRUE) + stats::poly(prcp_meanAnnTotal_5yr,2,raw=TRUE) + stats::poly(I(sqrt(precip_Seasonality_meanAnnAvg_5yr)),2,raw=TRUE) + stats::poly(PrecipTempCorr_meanAnnAvg_5yr,2, raw = TRUE), data = modDat %>% slice_sample(n = 50000), link = c("logit"), link.phi = NULL, -->
<!--                               type = c("ML")) -->

<!-- # predict responses -->
<!-- response <- "TotalTreeCover_dec" -->
<!-- predict_by_response <- function(mod, df) { -->
<!--   df_out <- df -->

<!--   response_name <- paste0(response, "_pred") -->
<!--   df_out[[response_name]] <- predict(mod, df, type = 'response') -->
<!--   df_out -->
<!-- } -->
<!-- testDat <- modDat %>% slice_sample(n = 50000) -->
<!-- pred_glm1 <- predict_by_response(treeMod, testDat) -->
<!-- pred_glm1$resid <- pred_glm1$TotalTreeCover_dec - pred_glm1$TotalTreeCover_dec_pred -->


<!-- # rasterize -->
<!-- test_rast <-  rast("../../../data/dayMet/rawMonthlyData/orders/70e0da02b9d2d6e8faa8c97d211f3546/Daymet_Monthly_V4R1/data/daymet_v4_prcp_monttl_na_1980.tif") %>%  -->
<!--   terra::aggregate(fact = 32, fun = "mean") -->

<!-- # rasterize data -->
<!-- plotResid_rast <- pred_glm1 %>%  -->
<!--          drop_na(resid) %>%  -->
<!--   #slice_sample(n = 5e4) %>% -->
<!--   terra::vect(geom = c("Lon", "Lat")) %>%  -->
<!--   terra::set.crs(crs(test_rast)) %>%  -->
<!--   terra::rasterize(y = test_rast,  -->
<!--                    field = "resid",  -->
<!--                    fun = mean) %>%  -->
<!--   terra::crop(ext(-2000000, 2500000, -2000000, 1200000)) -->
<!-- terra::plot(plotResid_rast, main = "Residuals from total tree cover model at 32km2 scale", clip = TRUE,  -->
<!--             plg = list(title = "residual"),  -->
<!--             col = map.pal("curvature")) -->
<!-- ``` -->

