---
title: "Cover-Climate Relationships Methods Overview"
output:
  html_document:
    code_folding: hide
    df_print: paged
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE,
                      message = FALSE)
```

```{r}
library(tidyverse) 
library(sf) 
library(terra)
library(ggpubr)
library(tidyterra)
library(USAboundaries)
ggplot2::theme_set(theme_classic())
# custom functions for plotting inset figures
            # make function to add inset
           get_inset <- function(df) {
              p <- ggplot(data = df) +
             facet_wrap(~lyr, ncol = 1, scales = "free_y") +
                geom_density(aes(Cover)) +
                theme_bw(base_size=9) +  ## makes everything smaller
                theme(panel.background = element_rect(fill="transparent"),  ## white plot background
                      panel.border = element_blank(),
          axis.title.y = element_blank(),
          axis.title.x = element_blank(),
          axis.text.x = element_text(size=rel(2)), ## tiny axis text
          axis.text.y = element_text(size=rel(2)), ## tiny axis text
          axis.line.x = element_line(size = .5),
          axis.line.y = element_line(size = .5),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          plot.background = element_blank(),
          strip.background = element_blank(),
          strip.text = element_blank()) 
  return(p)
           }
           # custom function to add inset to main plot
           annotation_custom2 <- function (grob, xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, data)
{
  layer(data = data, stat = StatIdentity, position = PositionIdentity,
        geom = ggplot2:::GeomCustomAnn,
        inherit.aes = TRUE, params = list(grob = grob,
                                          xmin = xmin, xmax = xmax,
                                          ymin = ymin, ymax = ymax))
           }
           ## custom inset plot for density plots w/ two lines
                      get_inset_2 <- function(df) {
              p <- ggplot(data = df) +
             facet_wrap(~lyr, ncol = 1, scales = "free_y") +
                geom_density(aes(Cover, col = state)) +
                theme_bw(base_size=9) +  ## makes everything smaller
                theme(panel.background = element_rect(fill="transparent"),  ## white plot background
                      panel.border = element_blank(),
          axis.title.y = element_blank(),
          axis.title.x = element_blank(),
          axis.text.x = element_text(size=rel(2)), ## tiny axis text
          axis.text.y = element_text(size=rel(2)), ## tiny axis text
          axis.line.x = element_line(size = .5),
          axis.line.y = element_line(size = .5),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          plot.background = element_blank(),
          strip.background = element_blank(),
          strip.text = element_blank()) 
  return(p)
           }
```

# Project Overview

We aim to generate a set of closed-form equations that determine coarse vegetation functional composition and phenology and quantity of biomass based on climate and soil inputs.

# Methods - Overview
1) Acquire cover data from FIA, LANDFIRE, AIM and RAP
    * CONUS-wide data for FIA and AIM (all available locations)
    * Subset LANDFIRE to be more spatially coarse across CONUS (more on par with the spatial sampling density of FIA), since the sampling density is very inconsistent across space in this dataset
    * For RAP, use CONUS-wide values for total tree cover (satellite-based estimates of tree cover are likely to be reliable in forest and non-forest locations). Because satellite-based estimates of cover for understory functional types are less reliable in forested areas, exclude cover for non-forest types in locations where tree cover is > 10%
2) Remove cover observations from locations that have been recently burned (according to perimeters from mtbs and cwf) 
    * For locations with burn history within the 20 years prior to collection of cover data, remove all cover data
    * For locations with burn history more than 20 years prior to collection of cover data, retain cover observations for non-tree functional groups, remove cover observations for trees 
3) Remove cover observations from locations that are developed (agriculture or cities, according to LCMAP land-cover use) 
4) "Coarsify" data from LANDFIRE to make the sampling density more spatially consistent across CONUS and on-par with the sampling density of FIA. 
5) Average cover values together that occur in the same dayMet grid cell in the same year
    * Two versions for tree cover:
        i) All available observations of total tree cover (even those where we don’t have information about the breakdown of that tree cover into broad-leaved vs. needle-leaved)
        ii) Only observations of total tree cover where we have information about the breakdown of tree cover into broad-leaved vs. needle-leaved
    * Also two versions for total herbaceous cover
        i) All available observations of total herbaceous cover (even those where we don’t have information about the breakdown of that herbaceous cover into C3 grass vs. C4 grass vs. Forb)
        ii) Only observations of total herbaceous cover where we have information about the breakdown of herbaceous cover into C3 grass vs. C4 grass vs. Forb
6) Acquire weather and climate data for each spatially-averaged location, derived from dayMet
    * Climate: average of values over the 30 years prior to observation
    * Weather: the anomalies of average of values over the three years prior to observation relative to the climate values
7) Acquire soil information for each spatially averaged location, derived from SOLUS100
    * Is not dynamic over time
8) Use this harmonized cover dataset to fit a model that classifies a location as forest or non-forest
    * Use the complete set of total tree cover data, plus climate and soils predictors
    * Identify a starting list of potential climate and soils predictors that are uncorrelated (r ≤ 0.7)
    * Convert the continuous tree cover data into a binary variable, where locations with >10% total tree cover are “forest,” while locations with ≤10% total tree cover are “unforested” (similar to the cutoff that FIA uses to identify forested areas for further sampling)
    * With a training subset of the data (data for all years except 2003, 2012, and 2019), use a binomial LASSO to identify a regularized set of predictors to include in the final model. The LASSO uses internal cross-validation to tune the lambda parameter, with data held out for the cross-validation by EPA level 2 ecoregion.
        i) Identify three possible predictor sets, based on three levels of lambda identified in the LASSO (“best” lambda, “1SE” lambda, and a value halfway between each of these that we call “1/2SE” lambda).
        ii) Fit binomial glms to the training data using each of these predictor sets and estimate coefficients.
        iii) For “best lambda” glm *and* either the “1SE” or “1/2SE” lambda glms (use whichever one has the fewest unique predictors but is not a null model). 
              * predict on the test dataset (data from 2003, 2012 and 2019), as well as to predict on the complete dataset
              * identify the optimal threshold for turning the continuous predictions of the binomial model back to binary classifications of forest vs. non forest. We tested various methods, and determined that using the prevalence of forest in the observed data as the threshold optimizes the precision and accuracy of the model
              * use the optimal threshold to translate the continuous model predictions of probability of forest into classifications of “forest” and “non-forest”
              * For both the test dataset as well as the complete dataset, compare the model-predictions to the test data and generate maps of classifications (including miss-classifications), confusion matrices, and confusion matrix statistics.
              * Make quantile plots that compare average model predictions of forest classification to average observations of forest classification within each quantile of each predictor variable.
        iv) Based on visualizations and model performance, identify a model that best represents the data (i.e. the “best,” “1SE,” or “1/2SE” lambda glm) and use that model to predict, for each location in the complete, harmonized cover dataset, whether that location is forested or unforested
9) Fit a Model that predicts cover for each functional type
    * Identify a set of predictors (climate, weather and soils variables) that are uncorrelated in locations that the above model identifies as forested. Do the same for unforested locations. 
    * For each cover type, fit models to three subsets of data: (1) all available data CONUS-wide, (2) data in locations that the forest model predicts as forest, (3) and data in locations the model predicts as unforested. For each subset of data, do the following: 

# Methods - Specifics

## (1) Acquire cover data from FIA, LANDFIRE, AIM and RAP

```{r, echo = FALSE, message=FALSE}
# get necessary raster data for background data on maps
test_rastTemp <-  rast("/Users/astears/Documents/Dropbox_static/Work/NAU_USGS_postdoc/cleanPED/PED_vegClimModels/Data_raw/dayMet/rawMonthlyData/orders/70e0da02b9d2d6e8faa8c97d211f3546/Daymet_Monthly_V4R1/data/daymet_v4_prcp_monttl_na_1980.tif") %>%
  terra::crop(ext(-2000000, 2600000, -1900000, 1500000))

## aggregate test_rastTemp by 13, the scale that the dayMet data are aggregated to 
test_rast_RAP <- test_rastTemp %>% 
  terra::aggregate(fact = 13, fun = "mean") #%>% 
#terra::project(crs("EPSG:4326")) %>% 
  #terra::crop(ext(-150, -50, 20, 60))
  
# aggregate to regular plotting density for other data types
test_rast <- test_rastTemp %>%
 terra::aggregate(fact = 8, fun = "mean")#%>% 
#terra::project(crs("EPSG:4326")) %>% 
 # terra::crop(ext(-150, -50, 20, 60))
  # transform to match format of veg. data 

cropped_states_2 <- us_states(resolution = "low")
cropped_states_2 <- suppressMessages(cropped_states_2 %>%
  dplyr::filter(name!="Hawaii") %>%
  dplyr::filter(name!="Alaska") %>%
  dplyr::filter(name!="Puerto Rico") %>%
  dplyr::filter(name!="American Samoa") %>%
  dplyr::filter(name!="Guam") %>%
  dplyr::filter(name!="Commonwealth of the Northern Mariana Islands") %>%
  dplyr::filter(name!="United States Virgin Islands") %>%
  sf::st_transform(sf::st_crs(test_rast))) 
```

### RAP 
* Manuscript detailing the Rangeland Analysis Platform: Allred, B.W., B.T. Bestelmeyer, C.S. Boyd, C. Brown, K.W. Davies, M.C. Duniway, L.M. Ellsworth, T.A. Erickson, S.D. Fuhlendorf, T.V. Griffiths, V. Jansen, M.O. Jones, J. Karl, A. Knight, J.D. Maestas, J.J. Maynard, S.E. McCord, D.E. Naugle, H.D. Starns, D. Twidwell, and D.R. Uden. 2021. Improving Landsat predictions of rangeland fractional cover with multitask learning and uncertainty. Methods in Ecology and Evolution. http://dx.doi.org/10.1111/2041-210x.13564

* Data were acquired from Google Earth Engine (GEE)
    * asset name: projects/rangeland-analysis-platform/vegetation-cover-v3
    * date accessed: Data from forested regions accessed March, 2026; data from un-forested regions accessed December, 2025

RAP data contain information about fractional cover on a remotely-sensed 30-m grid across CONUS. RAP provides fractional cover for **total trees**, **shrubs**, **bare ground**, and annual and perennial forb and grass cover. We combined values for annual and perennial forb and grass cover into one value for **total herbaceous cover**. While RAP does not have information about cover of needle leaved or broad leaved tree cover or forb, C3 grass, or C4 grass cover, when total tree cover from RAP is zero, we assume that needle leaved and broad leaved tree cover is also zero. Also, when total herbaceous cover from RAP is zero, we assume that forb, C3 grass and C4 grass cover is also zero. We acquired data from the same locations for each year from **2010 - 2024**. 

To sub-sample the RAP data, we used a process so that the sampled data was more similar in spatial sampling density to the field-sampled datasets: 
1) Start with generating 1 random point location within each 1km^2 dayMet grid cell within the sample region, which is across CONUS excluding locations that are developed, cropland, or water according to LCMAP land use data.
    - LCMAP data: https://eros.usgs.gov/lcmap/apps/data-downloads We used data from 2021, the most recent year data are available 
2) Sample the four closest RAP pixels around those random points and average the values together - this single point represents the ‘observation’ within that dayMet grid cell
3) Then, make the data coarser by averaging values that fall within the same ‘dayMet x 13’ grid cell, which gives a spatial sampling density that is qualitatively similar to that of FIA 

Below are maps of RAP data across all years after they’ve been sampled according to the process outlined above: (pending update w/ data from forested regions)
```{r, echo = FALSE, fig.width = 12, fig.height = 30, warning=FALSE, message = FALSE}
# Read in data 
dat_1temp <- st_read(dsn = "/Users/astears/Documents/Dropbox_static/Work/NAU_USGS_postdoc/cleanPED/PED_vegClimModels/Data_processed/CoverData/DataForAnalysisPoints", layer = "vegCompPoints") %>% #, driver = "ESRI Shapefile")
  st_transform(crs(test_rast))

# trim to be only later than 2000 or before 2024 (last year for which we have climate data was 2023) which is what we want to use for further analysis
dat_1 <- dat_1temp %>% 
  filter(Year >= 2000 & Year < 2024) %>% 
  mutate(Lon = st_coordinates(.)[,1],
         Lat = st_coordinates(.)[,2]) %>% 
  rename(ShrubCover = ShrbCvr, 
         TotalTreeCover = TtlTrCv, 
         TotalHerbaceousCover = TtlHrbC,
         BareGroundCover = BrGrndC,
         ForbCover = ForbCvr,
         C3GramCover = C3GrmCv,
         C4GramCover = C4GrmCv, 
         BroadLeavedTreeCover = AngTrCv,
         NeedleLeavedTreeCover = CnfTrCv)
 # st_drop_geometry()

###### rasterized version
# make a multi-layer raster w/ layers for each variable of interest
dat_1_RAST <- 
  lapply(c("FIA", "LANDFIRE", "LDC","RAP"), 
         FUN = function(y) {
           tempDat <- dat_1 %>% 
             filter(Source == y )
           tempDat1 <- tempDat %>% 
             st_drop_geometry()
           # get vector of years for the specific data type 
           if (y %in% c("FIA", "LDC", "RAP")) {
             datYears <- c("2016", "2020", "2022")
             if (y %in% "RAP") {
               # remove data from before 2010
               tempDat1 <- tempDat1 %>% 
                 filter(Year >=2010)
             }
           } else {
             datYears <- c("2003", "2007", "2015")
           }
           if (y == c( "RAP")) {
             # get data for each cover type by year
           tempOut <- apply(expand.grid(c("ShrubCover", "TotalTreeCover", "TotalHerbaceousCover", "BareGroundCover", "ForbCover", "C3GramCover",  "C4GramCover",  "BroadLeavedTreeCover", "NeedleLeavedTreeCover"), datYears, stringsAsFactors = FALSE), 
                            MARGIN = 1,
                            FUN = function(x) {
                              tempDat2 <- tempDat %>% 
                                filter(Year == x["Var2"])
                              
                              temp <- terra::rasterize(x = tempDat2,#tempDat2[,c("Lon", "Lat")], 
                                               y = test_rast_RAP,
                                               field = x["Var1"],
                                               #values = tempDat2[,names(tempDat2) == x["Var1"]],
                                               fun = mean, 
                                               na.rm = TRUE
                                               ) 
                      names(temp) <-  paste0(x["Var1"], "_", x["Var2"])
                      return(temp)
                    }
                    )
           # get dat across all years
           tempOut2 <- lapply(c("ShrubCover", "TotalTreeCover", "TotalHerbaceousCover", "BareGroundCover", "ForbCover", "C3GramCover",  "C4GramCover",  "BroadLeavedTreeCover", "NeedleLeavedTreeCover"), 
                            FUN = function(z) {
                              temp <- terra::rasterize(x = tempDat1[,c("Lon", "Lat")], 
                                               y = test_rast_RAP,
                                               values = tempDat1[,z],
                                               fun = mean
                                               ) 
                      
                      names(temp) <-  paste0(z, "_all")
                      return(temp)
                    }
                    )
           } else {
             # get data for each cover type by year
           tempOut <- apply(expand.grid(c("ShrubCover", "TotalTreeCover", "TotalHerbaceousCover", "BareGroundCover", "ForbCover", "C3GramCover",  "C4GramCover",  "BroadLeavedTreeCover", "NeedleLeavedTreeCover"), datYears, stringsAsFactors = FALSE), 
                            MARGIN = 1,
                            FUN = function(x) {
                              tempDat2 <- tempDat %>% 
                                filter(Year == x["Var2"])
                              
                              temp <- terra::rasterize(x = tempDat2,#tempDat2[,c("Lon", "Lat")], 
                                               y = test_rast,
                                               field = x["Var1"],
                                               #values = tempDat2[,names(tempDat2) == x["Var1"]],
                                               fun = mean, 
                                               na.rm = TRUE
                                               ) 
                      names(temp) <-  paste0(x["Var1"], "_", x["Var2"])
                      return(temp)
                    }
                    )
           # get dat across all years
           tempOut2 <- lapply(c("ShrubCover", "TotalTreeCover", "TotalHerbaceousCover", "BareGroundCover", "ForbCover", "C3GramCover",  "C4GramCover",  "BroadLeavedTreeCover", "NeedleLeavedTreeCover"), 
                            FUN = function(z) {
                              temp <- terra::rasterize(x = tempDat1[,c("Lon", "Lat")], 
                                               y = test_rast,
                                               values = tempDat1[,z],
                                               fun = mean
                                               ) 
                      
                      names(temp) <-  paste0(z, "_all")
                      return(temp)
                    }
                    )
           }
           
           # for each data type, put data for each year in one raster in the list, and the data across all years in another raster
           outDat <- list(c(tempOut[[1]], tempOut[[2]], tempOut[[3]], tempOut[[4]], tempOut[[5]], tempOut[[6]], tempOut[[7]], tempOut[[8]], tempOut[[9]]), 
                          c(tempOut[[10]], tempOut[[11]], tempOut[[12]], tempOut[[13]], tempOut[[14]], tempOut[[15]], tempOut[[16]], tempOut[[17]], tempOut[[18]]),
                          c(tempOut[[19]], tempOut[[20]], tempOut[[21]], tempOut[[22]], tempOut[[23]], tempOut[[24]], tempOut[[25]], tempOut[[26]], tempOut[[27]]), 
                          c(tempOut2[[1]], tempOut2[[2]], tempOut2[[3]], tempOut2[[4]], tempOut2[[5]], tempOut2[[6]], tempOut2[[7]], tempOut2[[8]], tempOut2[[9]]))
           names(outDat) <- c(datYears, "allYears")
           return(outDat)
         }
) %>% invisible()
names(dat_1_RAST) <- c("FIA", "LANDFIRE", "LDC","RAP")

## make figures
## make figure for RAP 
testFigs <- lapply(names(dat_1_RAST$RAP)[[4]], FUN = function(y) {
  # make a map for this year
           tempFig <- 
  ggplot() +
             facet_wrap(~lyr, ncol = 1) +
geom_spatraster(data = dat_1_RAST$RAP[[y]][[1:4]], maxcell = 5e+10) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) +
  coord_sf(datum = st_crs(test_rastTemp)) +
             paletteer::scale_fill_paletteer_c("scico::roma", guide = guide_colorbar(title = "% cover"),
                                               limits = c(0,100), na.value = "white") + 
             theme(
                   axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) + 
             ggtitle("Map of RAP data on a dayMet x 13 grid")
             
  # add an inset showing a density plot  
           # get data
           data_inset <- #terra::values(dat_1_RAST$RAP[[y]][[1:4]]) %>% 
             dat_1 %>% 
             st_drop_geometry() %>% 
             filter(Source == "RAP") %>% 
               #as_tibble() %>% 
             rename(ShrubCover_all = ShrubCover, 
                    TotalTreeCover_all = TotalTreeCover, 
                    TotalHerbaceousCover_all = TotalHerbaceousCover, 
                    BareGroundCover_all = BareGroundCover) %>% 
               pivot_longer(c(ShrubCover_all, TotalTreeCover_all, TotalHerbaceousCover_all, BareGroundCover_all), names_to = "lyr", values_to = "Cover")
 
           # make insets
           inset_plots <- data_inset %>%
             split(f = .$lyr) %>%
             purrr::map(~annotation_custom2(
    grob = ggplotGrob(get_inset(.) #+
                        #scale_y_continuous(limits=c(0,105), breaks = c(0, 50, 100))
                      ),
    data = data.frame(lyr=unique(.$lyr)),
    ymin = -2025000, ymax=-1205000, xmin=-2280000, xmax=-180000)
  )

         tempFig2 <- tempFig + inset_plots

           return(tempFig2)
})

RAPfigs <- testFigs[[1]]
RAPfigs
# 
# # jpeg(filename = "/Users/astears/Documents/Dropbox_static/Work/NAU_USGS_postdoc/cleanPED/PED_vegClimModels/Figures/CoverDatFigures/RAPInputs_step1.jpeg",
# #     width = 800, height = 2000)
# ggarrange(#testFigs[[1]], 
#           #testFigs[[2]], 
#           ##testFigs[[3]], 
#           testFigs[[1]], 
#           nrow = 1, common.legend = TRUE, legend = "bottom"
#           ) %>% 
#   annotate_figure(top = "Map of RAP data on a dayMet x 13 grid", fig.lab.size = 4)
# #dev.off()
```

This barplot shows the sample size for each functional group across years in the subsampled RAP data we’re using (pending update w/ data from forested regions) 
```{r, fig.width=7, fig.height = 5.5}
# barplots of # of observations per data type per year
dat_1 %>% 
  st_drop_geometry() %>% 
  filter(Source == "RAP" & Year >=2010) %>%
  group_by(Year) %>% 
  summarize(ShrubCover = sum(!is.na(ShrubCover)), 
            TotalTreeCover = sum(!is.na(TotalTreeCover)),
            TotalHerbaceousCover = sum(!is.na(TotalHerbaceousCover)),
            BareGroundCover = sum(!is.na(BareGroundCover))) %>% 
  pivot_longer(2:5, values_to = "No_Observations", names_to = "functionalGroup") %>% 
  ggplot() + 
  geom_col(aes(x = Year, y = No_Observations, fill = functionalGroup), position = "dodge", col = "grey20")  + 
   scale_fill_discrete(type = c("#000000", #"#E69F00" ,"#56B4E9", "#009E73", "#F0E442" ,"#0072B2" ,
                                "#D55E00", "#CC79A7", "#999999"),
                      
                        labels = c("BareGround",
                         #"BroadLeavedTree", "C3Grass","C4Grass","Forb", "NeedleLeavedTree",
                                  "Shrub",  "TotalHerbaceous", "TotalTree")
   ) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  ylim(c(0,NA)) +
  ylab(c("Number of Observations"))
```

### AIM
* Manuscript detailing the Landscape Data Commons: McCord, Sarah E., Nicholas P. Webb, Brandon T. Bestelmeyer, Kristopher Bonefont, Joseph R. Brehm, Joel Brown, Ericha M. Courtright, Chris Dietrich, Michael C. Duniway, Brandon Edwards, Christopher Fraser, Jeffrey E. Herrick, Anna C. Knight, Loretta Metz, Justin W. Van Zee, and Craig Tweedie. 2023. The Landscape Data Commons: A system for standardizing, accessing, and applying large environmental datasets for agroecosystem research and management. Agricultural & Environmental Letters. https://doi.org/10.1002/ael2.20120
* Citation for the AIM dataset itself: Taylor, J.J., E.J. Kachergis, G.R. Toevs, J.W. Karl, M.R. Bobo, M. Karl, S. Miller, and C.S. Spurrier. 2014. AIM-Monitoring: A Component of the BLM Assessment, Inventory, and Monitoring Strategy. Technical Note 445. U.S. Department of the Interior, Bureau of Land Management, National Operations Center, Denver, CO. 
* Information about the trex R package: https://github.com/Landscape-Data-Commons/trex

* AIM Data were acquired via the Landscape Data Commons (LDC), which harmonizes and aggregates data from AIM and other sources. We used the 'trex' R package to access AIM data stored within the LDC.  
    * We used AIM data from all plots in the LDC, which we accessed on 3/19/2024. 

The BLM- AIM database provides information about vegetation cover from a standardized survey protocol. It includes information about **total tree cover**, **shrub cover**, **total herbaceous cover**, **needle-leaved tree cover**, **broad-leaved tree cover**, **C3 grass cover**, **C4 grass cover**, and **forb cover**, and **bare ground cover**. We use AIM data from **2000 to 2017**. 

Below are maps of AIM data aggregated across all years:
```{r fig.width= 12, fig.height =67.5}
## make figure for LDC
testFigs <- lapply(names(dat_1_RAST$LDC)[4], FUN = function(y) {
  # make a map for this year
           tempFig <-
  ggplot() +
             facet_wrap(~lyr, ncol = 1) +
geom_spatraster(data = dat_1_RAST$LDC[[y]], maxcell = 5e+10) +
  geom_sf(data=cropped_states_2 ,fill=NA ) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) +
  coord_sf(datum = st_crs(test_rastTemp)) +
             paletteer::scale_fill_paletteer_c("scico::roma", guide = guide_colorbar(title = "% cover"),
                                               limits = c(0,100), na.value = "white") + 
             theme(
                   axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) + 
             ggtitle("Map of AIM data on a dayMet x 8 grid")
  # add an inset showing a density plot
           # get data
           data_inset <- dat_1 %>% 
             st_drop_geometry() %>% 
             filter(Source == "LDC") %>% 
             rename(ShrubCover_all = ShrubCover, 
                    TotalTreeCover_all = TotalTreeCover, 
                    TotalHerbaceousCover_all = TotalHerbaceousCover, 
                    BareGroundCover_all = BareGroundCover,
                    ForbCover_all = ForbCover, 
                    C3GramCover_all = C3GramCover, 
                    C4GramCover_all = C4GramCover, 
                    BroadLeavedTreeCover_all = BroadLeavedTreeCover,
                    NeedleLeavedTreeCover_all = NeedleLeavedTreeCover) %>% 
               pivot_longer(c(ShrubCover_all, TotalTreeCover_all, TotalHerbaceousCover_all, BareGroundCover_all,
                              ForbCover_all,  C3GramCover_all,  C4GramCover_all,  BroadLeavedTreeCover_all, NeedleLeavedTreeCover_all), 
                            names_to = "lyr", values_to = "Cover")
 
           
           # data_inset <- terra::values(dat_1_RAST$LDC[[y]]) %>%
           #     as_tibble() %>%
           #     pivot_longer(1:9, names_to = "lyr", values_to = "Cover")


           # make insets
           inset_plots <- data_inset %>%
             split(f = .$lyr) %>%
             purrr::map(~annotation_custom2(
    grob = ggplotGrob(get_inset(.) #+
                        #scale_y_continuous(limits=c(0,105), breaks = c(0, 50, 100))
                      ),
    data = data.frame(lyr=unique(.$lyr)),
    ymin = -2025000, ymax=-1205000, xmin=-2280000, xmax=-180000)
  )

         tempFig2 <- tempFig + inset_plots

           return(tempFig2)
})  %>% invisible()

AIMfigs <- testFigs[[1]]
AIMfigs
```
This barplot shows the sample size for each functional group across years in the AIM data we’re using: 
```{r fig.width=8, fig.height = 5.5}
# barplots of # of observations per data type per year
dat_1 %>%
  st_drop_geometry() %>%
  filter(Source == "LDC" #& Year >=2010
         ) %>%
  group_by(Year) %>%
  summarize(ShrubCover = sum(!is.na(ShrubCover)),
            TotalTreeCover = sum(!is.na(TotalTreeCover)),
            TotalHerbaceousCover = sum(!is.na(TotalHerbaceousCover)),
           BareGroundCover = sum(!is.na(BareGroundCover)),
            BroadLeavedTreeCover = sum(!is.na(BroadLeavedTreeCover)),
            NeedleLeavedTreeCover = sum(!is.na(NeedleLeavedTreeCover)),
            ForbCover = sum(!is.na(ForbCover)),
            C3GramCover = sum(!is.na(C3GramCover)),
            C4GramCover = sum(!is.na(C4GramCover))
            ) %>%
  pivot_longer(2:10, values_to = "No_Observations", names_to = "functionalGroup") %>%
  ggplot() +
  geom_col(aes(x = Year, y = No_Observations, fill = functionalGroup), position = "dodge", col = "grey20")  +
   scale_fill_discrete(type = c("#000000", "#E69F00" ,"#56B4E9", "#009E73", "#F0E442" ,"#0072B2" ,
                                "#D55E00", "#CC79A7", "#999999"),
                        labels = c("BareGround",
                         "BroadLeavedTree", "C3Grass","C4Grass","Forb", "NeedleLeavedTree",
                                  "Shrub",  "TotalHerbaceous", "TotalTree")
   )+ 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  ylim(c(0,NA)) +
  ylab(c("Number of Observations"))
```

### FIA
FIA data The BLM- AIM database provides information about vegetation cover from a standardized survey protocol. It includes information about **total tree cover**, **shrub cover**, **total herbaceous cover**, **needle-leaved tree cover**, **broad-leaved tree cover**, **forb cover**, and **bare ground cover**. While FIA does not have information about cover of C3 or C4 grass cover, when total herbaceous cover from FIA is zero, we assume that forb, C3 grass and C4 grass cover is also zero. We use FIA data from **2011 to 2022** (FIA data prior to 2011 don’t measure total tree cover in a standardized way across CONUS). Below are maps of data from FIA across all years:
```{r fig.width= 12, fig.height = 52.5}
## make figure for FIA
testFigs <- lapply(names(dat_1_RAST$FIA)[4], FUN = function(y) {
  # make a map for this year
           tempFig <-
  ggplot() +
             facet_wrap(~lyr, ncol = 1) +
geom_spatraster(data = dat_1_RAST$FIA[[y]][[c(1:5,8:9)]], maxcell = 5e+10) +
  geom_sf(data=cropped_states_2 ,fill=NA ) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) + 
  coord_sf(datum = st_crs(test_rastTemp)) +
             paletteer::scale_fill_paletteer_c("scico::roma", guide = guide_colorbar(title = "% cover"),
                                               limits = c(0,100), na.value = "white") + 
             theme(
                   axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) + 
             ggtitle("Map of FIA data on a dayMet x 8 grid")
  # add an inset showing a density plot
           # get data
                      data_inset <- dat_1 %>% 
             st_drop_geometry() %>% 
             filter(Source == "FIA") %>% 
             rename(ShrubCover_all = ShrubCover, 
                    TotalTreeCover_all = TotalTreeCover, 
                    TotalHerbaceousCover_all = TotalHerbaceousCover, 
                    BareGroundCover_all = BareGroundCover,
                    ForbCover_all = ForbCover, 
                    C3GramCover_all = C3GramCover, 
                    C4GramCover_all = C4GramCover, 
                    BroadLeavedTreeCover_all = BroadLeavedTreeCover,
                    NeedleLeavedTreeCover_all = NeedleLeavedTreeCover) %>% 
               pivot_longer(c(ShrubCover_all, TotalTreeCover_all, TotalHerbaceousCover_all, BareGroundCover_all,
                              ForbCover_all, BroadLeavedTreeCover_all, NeedleLeavedTreeCover_all), 
                            names_to = "lyr", values_to = "Cover")
           # data_inset <- terra::values(dat_1_RAST$FIA[[y]][[c(1:5,8:9)]]) %>%
           #     as_tibble() %>%
           #     pivot_longer(1:7, names_to = "lyr", values_to = "Cover")

           # make insets
           inset_plots <- data_inset %>%
             split(f = .$lyr) %>%
             purrr::map(~annotation_custom2(
    grob = ggplotGrob(get_inset(.) #+
                        #scale_y_continuous(limits=c(0,105), breaks = c(0, 50, 100))
                      ),
    data = data.frame(lyr=unique(.$lyr)),
    ymin = -2025000, ymax=-1205000, xmin=-2280000, xmax=-180000)
  )

         tempFig2 <- tempFig + inset_plots

           return(tempFig2)
})  %>% invisible()
FIAfigs <- testFigs[[1]]
FIAfigs
```

This barplot shows the sample size for each functional group across years in the AIM data we’re using:
```{r fig.width=8, fig.height = 5.5}
# barplots of # of observations per data type per year
dat_1 %>%
  st_drop_geometry() %>%
  filter(Source == "FIA") %>%
  group_by(Year) %>%
  summarize(ShrubCover = sum(!is.na(ShrubCover)),
            TotalTreeCover = sum(!is.na(TotalTreeCover)),
            TotalHerbaceousCover = sum(!is.na(TotalHerbaceousCover)),
           BareGroundCover = sum(!is.na(BareGroundCover)),
            BroadLeavedTreeCover = sum(!is.na(BroadLeavedTreeCover)),
            NeedleLeavedTreeCover = sum(!is.na(NeedleLeavedTreeCover)),
            ForbCover = sum(!is.na(ForbCover))#,
            #C3GramCover = sum(!is.na(C3GramCover)),
            #C4GramCover = sum(!is.na(C4GramCover))
            ) %>%
  pivot_longer(2:8, values_to = "No_Observations", names_to = "functionalGroup") %>%
  ggplot() +
  geom_col(aes(x = Year, y = No_Observations, fill = functionalGroup), position = "dodge", col = "grey20") +
   scale_fill_discrete(type = c("#000000",
     "#E69F00" ,#"#56B4E9", "#009E73",
     "#F0E442" ,"#0072B2" ,"#D55E00", "#CC79A7", "#999999"),
              
      labels = c("BareGround",
                         "BroadLeavedTree", #"C3Grass","C4Grass",
                 "Forb", "NeedleLeavedTree",
                                  "Shrub",  "TotalHerbaceous", "TotalTree")
 )+ 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  ylim(c(0,NA)) +
  ylab(c("Number of Observations"))
```

### LANDFIRE
The LANDFIRE reference database combines information about vegetation cover from many different sources (state surveys, regional surveys, etc.) It includes information about **total tree cover**, **shrub cover**, **total herbaceous cover**, **needle-leaved tree cover**, **broad-leaved tree cover**, **C3 grass cover**, **C4 grass cover**, and **forb cover**, although not every observation includes information about all functional groups. We use LANDFIRE data from **2000 to 2017**. 
Below are maps of LANDFIRE data summarized across all years:
```{r fig.width=12, fig.height = 60, warning = FALSE}
## make figure for LANDFIRE
testFigs <- lapply(names(dat_1_RAST$LANDFIRE)[4], FUN = function(y) {
  # make a map for this year
           tempFig <-
  ggplot() +
             facet_wrap(~lyr, ncol = 1) +
geom_spatraster(data = dat_1_RAST$LANDFIRE[[y]][[c(1:3,5:9)]], maxcell = 5e+10) +
  geom_sf(data=cropped_states_2 ,fill=NA ) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) +
  coord_sf(datum = st_crs(test_rastTemp)) +
             paletteer::scale_fill_paletteer_c("scico::roma", guide = guide_colorbar(title = "% cover"),
                                               limits = c(0,100), na.value = "white") + 
             theme(
                   axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) + 
             ggtitle("Map of LANDFIRE data on a dayMet x 13 grid")
  # add an inset showing a density plot
           # get data
           # data_inset <- terra::values(dat_1_RAST$LANDFIRE[[y]][[c(1:3,5:9)]]) %>%
           #     as_tibble() %>%
           #     pivot_longer(1:8, names_to = "lyr", values_to = "Cover")
        data_inset <- dat_1 %>% 
             st_drop_geometry() %>% 
             filter(Source == "LANDFIRE") %>% 
             rename(ShrubCover_all = ShrubCover, 
                    TotalTreeCover_all = TotalTreeCover, 
                    TotalHerbaceousCover_all = TotalHerbaceousCover, 
                    BareGroundCover_all = BareGroundCover,
                    ForbCover_all = ForbCover, 
                    C3GramCover_all = C3GramCover, 
                    C4GramCover_all = C4GramCover, 
                    BroadLeavedTreeCover_all = BroadLeavedTreeCover,
                    NeedleLeavedTreeCover_all = NeedleLeavedTreeCover) %>% 
               pivot_longer(c(ShrubCover_all, TotalTreeCover_all, TotalHerbaceousCover_all, 
ForbCover_all, C3GramCover_all, C4GramCover_all, BroadLeavedTreeCover_all, NeedleLeavedTreeCover_all), 
                            names_to = "lyr", values_to = "Cover")
           # make insets
           inset_plots <- data_inset %>%
             split(f = .$lyr) %>%
             purrr::map(~annotation_custom2(
    grob = ggplotGrob(get_inset(.) #+
                        #scale_y_continuous(limits=c(0,105), breaks = c(0, 50, 100))
                      ),
    data = data.frame(lyr=unique(.$lyr)),
    ymin = -2025000, ymax=-1205000, xmin=-2280000, xmax=-180000)
  )

         tempFig2 <- tempFig + inset_plots

           return(tempFig2)
}) %>% invisible()

LANDFIREfigs <- testFigs[[1]]
LANDFIREfigs
```
This barplot shows the sample size for each functional group across years in the LANDFIRE data we’re using: 
```{r fig.width=8, fig.height=5.5}
# barplots of # of observations per data type per year
dat_1 %>%
  st_drop_geometry() %>%
  filter(Source == "LANDFIRE") %>%
  group_by(Year) %>%
  summarize(ShrubCover = sum(!is.na(ShrubCover)),
            TotalTreeCover = sum(!is.na(TotalTreeCover)),
            TotalHerbaceousCover = sum(!is.na(TotalHerbaceousCover)),
           # BareGroundCover = sum(!is.na(BareGroundCover)),
            BroadLeavedTreeCover = sum(!is.na(BroadLeavedTreeCover)),
            NeedleLeavedTreeCover = sum(!is.na(NeedleLeavedTreeCover)),
            ForbCover = sum(!is.na(ForbCover)),
            C3GramCover = sum(!is.na(C3GramCover)),
            C4GramCover = sum(!is.na(C4GramCover))
            ) %>%
  pivot_longer(2:9, values_to = "No_Observations", names_to = "functionalGroup") %>%
  ggplot() +
  geom_col(aes(x = Year, y = No_Observations, fill = functionalGroup), position = "dodge", col = "grey20") +
   scale_fill_discrete(type = c(#"#000000",
     "#E69F00" ,"#56B4E9", "#009E73", "#F0E442" ,"#0072B2" ,"#D55E00", "#CC79A7", "#999999"),
                       labels = c(#"BareGround",
                         "BroadLeavedTree", "C3Grass","C4Grass","Forb", "NeedleLeavedTree",
                                  "Shrub",  "TotalHerbaceous", "TotalTree")
 )+ 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  ylim(c(0,NA)) +
  ylab(c("Number of Observations"))
```

<!-- ### side-by-side maps of all data types across sources -->
<!-- ```{r fig.width = 20, fig.} -->
<!-- ggarrange(RAPfigs, AIMfigs, FIAfigs, LANDFIREfigs, ncol = 4) -->
<!-- ``` -->


## (2) Remove cover observations from locations that have been recently burned 

We combine fire perimeter information from two data sources: 
  1) mtbs
  2) cwf
If the location where cover data was collected was burned *within* 20 years prior to the year when cover data were collected, we do not use that cover information. 
If the location was burned *more* than 20 years prior to the year when cover data were collected, we retain cover information for non-tree cover classes, but exclude information about total tree cover, needle-leaved tree cover, and broad-leaved tree cover. 

```{r, fig.width=12, fig.height=8}
# get version of the data w/ fire info added
dat_fireRemoved <- readRDS(file = "/Users/astears/Documents/Dropbox_static/Work/NAU_USGS_postdoc/cleanPED/PED_vegClimModels/Data_processed/CoverData/dataForAnalysis_FireDataPRE_fireRemoved.rds") %>% 
  st_transform(crs(dat_1)) %>% 
  mutate(geom_x = st_coordinates(.)[,1],
         geom_y = st_coordinates(.)[,2]) %>% 
  filter(Year >=2000)

# locations where we removed all observations (present in pre-fire, but not on post-fire)
dat_fireCompletelyRemoved <- dat_fireRemoved %>% 
  filter(burnedMoreThan20YearsAgo == "RECENTLY_BURNED_REMOVE")

## rasterize this dataset
 dat_fireCompletelyRemoved_RAST <- terra::rasterize(x = dat_fireCompletelyRemoved,#tempDat2[,c("Lon", "Lat")], 
                                               y = test_rast,
                                               field ="Year",
                                               #values = tempDat2[,names(tempDat2) == x["Var1"]],
                                               fun = mean, 
                                               na.rm = TRUE
                                               ) 

## data excluded completely b/c of fire <20 years prior to sampling
## make a map of all burned locations in CONUS
burnPlot_20yr <- ggplot() +
  #facet_wrap(~Year) +
  geom_spatraster(data = dat_fireCompletelyRemoved_RAST) +
  #geom_sf(data = dat_fireCompletelyRemoved# %>% filter(Year == 2012)
   #       , aes(col = Year), pch = 20) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) +
              paletteer::scale_fill_paletteer_c("scico::tofino", guide = guide_colorbar(title = "Year"),
                                                limits = c(2000,2023), na.value = "white") + 
             theme(
                   axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) +
             ggtitle("Locations where we excluded all cover data \ndue to fire within 20 years prior to sampling",
                     subtitle = "Color corresponds to the year when excluded cover data were collected")

burnPlot_20yr
```


```{r, fig.width=12, fig.height=8}
# locations where we removed tree \
dat_firePartlyRemoved <- dat_fireRemoved %>% 
  filter(burnedMoreThan20YearsAgo == TRUE)

## rasterize this dataset
 dat_firePartlyRemoved_RAST <- terra::rasterize(x = dat_firePartlyRemoved,#tempDat2[,c("Lon", "Lat")], 
                                               y = test_rast,
                                               field ="Year",
                                               #values = tempDat2[,names(tempDat2) == x["Var1"]],
                                               fun = mean, 
                                               na.rm = TRUE
                                               ) 

## data excluded completely b/c of fire <20 years prior to sampling
## make a map of all burned locations in CONUS
burnPlot_all <- ggplot() +
  #facet_wrap(~Year) +
  geom_spatraster(data = dat_firePartlyRemoved_RAST) +
  #geom_sf(data = dat_firePartlyRemoved# %>% filter(Year == 2012)
   #       , aes(col = Year), pch = 20) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) +
              paletteer::scale_fill_paletteer_c("scico::tofino", guide = guide_colorbar(title = "Year"),
                                                limits = c(2000,2023), na.value = "white") + 
             theme(
                   axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) +
             ggtitle("Locations where we excluded tree cover data \ndue to any fire prior to sampling",
                     subtitle = "Color corresponds to the year when excluded cover data were collected")

burnPlot_all
```


```{r fig.width=10}
## add plots showing density of observations by functional group both before 
dat_postBurnRemoval <- dat_fireRemoved %>%  
  rename(ShrubCover = ShrbCvr, 
         TotalTreeCover = TtlTrCv, 
         TotalHerbaceousCover = TtlHrbC,
         BareGroundCover = BrGrndC,
         ForbCover = ForbCvr,
         C3GramCover = C3GrmCv,
         C4GramCover = C4GrmCv, 
         BroadLeavedTreeCover = AngTrCv,
         NeedleLeavedTreeCover = CnfTrCv) %>% 
  dplyr::select(-c(SttUntC, Plot, PltCndt, Month, Day, TtlGrmC, AnnlHGC, PrnnHGC, TrBsA_2, LttrCvr, LttrDpt, AngTrC_, CnfTrC_, C4GrmC_, C3GrmC_, FrbCvr_, group)) %>% 
  pivot_longer(cols = c(TotalTreeCover, TotalHerbaceousCover, ShrubCover, BareGroundCover, C3GramCover, C4GramCover, ForbCover, BroadLeavedTreeCover, NeedleLeavedTreeCover), 
               names_to = "CoverType", values_to = "CoverValue") %>% 
  filter(burnedMoreThan20YearsAgo != "RECENTLY_BURNED_REMOVE") %>%  # remove all observations that were burned recently
  mutate(CoverValue = replace(x = CoverValue, 
                              list = (CoverType %in% c("TotalTreeCover", "BroadLeavedTreeCover", "NeedleLeavedTreeCover")) & burnedMoreThan20YearsAgo == TRUE, 
                              values = NA),  # for plots that were burned at any point, remove observations for tree data 
  CoverValue = replace(x = CoverValue, 
                       list = (CoverType %in% c("BroadLeavedTreeCover", "NeedleLeavedTreeCover", "C3GramCover", "C4GramCover", "ForbCover")) & Source == "RAP", 
                       values = NA))# for RAP data, remove values for BroadLeavedTreeCover, C3GramCover, C4GramCOver, ForbCover, and NeedleLeavedTreeCover (put zeros when the corresponding level-1 value is 0, but those are the only values)
  # make pre-fire removal data 'long' 
  dat_1_long <- dat_1 %>% 
    pivot_longer(cols = c(TotalTreeCover, TotalHerbaceousCover, ShrubCover, BareGroundCover, C3GramCover, C4GramCover, ForbCover, BroadLeavedTreeCover, NeedleLeavedTreeCover), 
               names_to = "CoverType", values_to = "CoverValue") %>% 
    mutate(CoverValue = replace(x = CoverValue, 
                       list = (CoverType %in% c("BroadLeavedTreeCover", "NeedleLeavedTreeCover", "C3GramCover", "C4GramCover", "ForbCover")) & Source == "RAP", 
                       values = NA))# for RAP data, remove values for BroadLeavedTreeCover, C3GramCover, C4GramCOver, ForbCover, and NeedleLeavedTreeCover (put zeros when the corresponding level-1 value is 0, but those are the only values)
  ggplot() +
    facet_wrap(~CoverType, scales = "free") + 
    geom_density(data = dat_1_long, aes(CoverValue#, col= Source
                                                 ), col = "grey") + 
    geom_density(data = dat_postBurnRemoval , aes(CoverValue#, col = Source
                                        ),  col = "black") + 
    ggtitle("Density plots of cover by functional group across all datasets", 
            subtitle = "Grey = Data prior to excluding burned locations; \nBlack = Data after excluding burned locations")
```

## (3) Remove cover observations from locations that are developed 
We excluded vegetation cover ata from locations that the LCMAP dataset designated as developed (LCMAP class 1), cropland (LCMAP class 2) or water (LCMAP class 5)

If a point location (or the RAP data associated with a randomly-selected point location in a dayMet grid cell) was within a 30-m^2 LCMAP grid cell that was classified as developed, cropland, or water, we excluded cover data from that point location. 
We used data from LCMAP version 1.3 from 2021, which was the most recent year for which data were available, as it was likely the most "conservative" (i.e. classified the most area as developed or cropland).  
Dataset source: https://eros.usgs.gov/lcmap/apps/data-downloads
Access date: 05 June, 2024

Because the LCMAP data is so fine-resolution, it isn't practical to display locations where we excluded cover data due to LCMAP cover type on the native LCMAP raster grid. In this figure below, we show 4km x 4km grid cells (the dayMet grid scaled up by a factor of four) within in which we excluded at least one data point. The color of the grid cell corresponds to the land cover type that was most commonly associated with excluded cover data observations in that grid cell. 

```{r fig.width=13, fig.height=10}
# data after removing developed/ag. locations
dat_LCMAPremoved <- readRDS("../Data_processed/CoverData/DataForModels.RDS")

# # # raster layer of locations excluded based on LCMAP classification
# LCMAP_mask <- terra::rast("../Data_raw/LCMAP/LCMAP_CU_2021_V13_LCPRI.tif")
# # change the NA values to 0, for plotting purposes
# LCMAP_mask2 <- LCMAP_mask %>%
#   terra::classify(rcl = matrix(c(1,2,3,4,5,6,7,8, 1,2,NA,NA,5,NA,NA,NA), ncol = 2))
# # put on the dayMet grid
# LCMAP_mask3 <- LCMAP_mask2 %>%
#   terra::project(terra::crs(test_rastTemp))
# 
# # mask the LCMAP data so there aren't any values outside of the state boundary shapefile
# LCMAP_mask4 <- LCMAP_mask3 %>%
#   terra::mask(mask =  vect(cropped_states_2)) %>%
#   round() %>%
#   terra::classify(rcl = matrix(c(1,2,3,4,5, as.integer(1), as.integer(2), NA, NA, as.integer(5)), ncol = 2)) %>%
#   as.factor()
# lvls <- data.frame(id=c(1,2,5), cover=c("developed", "cropland", "water"))
# levels(LCMAP_mask4) <- lvls
# 
# # save the values to use later
# terra::writeRaster(LCMAP_mask4, "../Data_processed/LCMAP/LCMAP_forPlotting_dayMetGrid.tif", overwrite = TRUE)
# read in raster for plotting
LCMAP_mask4 <- terra::rast("../Data_processed/LCMAP/LCMAP_forPlotting_dayMetGrid.tif")

## now, get locations that overlap with the cover data (get locations where we DID exclude cover information based on LCMAP)
# data, pre-LCMAP filtering
# find where this data overlaps w/ the LCMAP mask
# dat_LCMAPfiltered_temp <- terra::extract(x = LCMAP_mask4, y = vect(dat_postBurnRemoval))
# dat_LCMAPfiltered <- dat_postBurnRemoval %>% 
#   cbind(dat_LCMAPfiltered_temp) %>% 
#   select(-ID)
#`dat_LCMAPfiltered` contains information about the locations that were filtered out as well as those that were retained
#saveRDS(dat_LCMAPfiltered, file = "../Data_processed/LCMAP/CoverDataWithLCMAPExclusionInfoForPlotting.rds")
# get point locations that have been filtered out because they overlap w/ a LCMAP land cover type we don't want
# dat_LCMAPfiltered <- readRDS("../Data_processed/LCMAP/CoverDataWithLCMAPExclusionInfoForPlotting.rds")
# dat_LCMAPfiltered <- dat_LCMAPfiltered %>% 
#   mutate(coverType_code = as.integer(cover))
# # 
# # # crop the LCMAP rast extent for testing
# # LCMAP_mask4_test <- LCMAP_mask4 %>% 
# #   terra::crop(terra::ext(c(500000, 520000, 
# #                          -20000, 0)))
# 
# # rasterize the data that were excluded, and color-code by land cover type (rasterize on the dayMet grid (on dayMet x 4 grid, so 4km x 4km), since the LCMAP grid is too fine to visualize accurately at CONUS-scale)--so the resulting map shows 4km x 4km grid cells where we excluded at least one location because of land cover type we didn't want 
# rast_LCMAPfiltered <- dat_LCMAPfiltered %>%
#   filter(!is.na(coverType_code)) %>%
#   #slice(30000:40000) %>%
#   vect() %>%
#   terra::rasterize(y = test_rastTemp %>% terra::aggregate(fact = 4), field = "coverType_code", fun =
#                      function(x) {as.integer(round(mean(x, na.rm = TRUE)))} #as.integer(mean)#DescTools::Mode
#                )
# # round values so that they're integers
#  rast_LCMAPfiltered <- rast_LCMAPfiltered %>%
# #   round() %>%
#    terra::as.int()
# 
# # save raster for plotting 
#  terra::writeRaster(rast_LCMAPfiltered, "../Data_processed/LCMAP/LCMAP_DataRemoved_forPlotting_dayMetGrid.tif", overwrite = TRUE, datatype="INTIU")
rast_LCMAPfiltered <- terra::rast( "../Data_processed/LCMAP/LCMAP_DataRemoved_forPlotting_dayMetGrid.tif")

# now find the gridcells that we removed ()
 ggplot(rast_LCMAPfiltered) +
   geom_raster(aes(x = x, 
                   y = y, 
                   fill = as.character(coverType_code)), maxcell = 1000000 ) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
   scale_fill_manual(name = "Land cover", 
                     values = c("firebrick3", "darkgoldenrod2", "royalblue", NA), 
                     labels = c("Developed", "Cropland", "Water", ""),
                    na.translate = FALSE) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) +
             theme(
                   axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) +
             ggtitle("Locations where we excluded data due to land cover type",
                     subtitle = "Land cover type information from LCMAP")

```


```{r fig.width=10}
## add plots showing density of observations by functional group both before and after excluding burned areas 
## for pre-LCMAP filtering, the data is in the d.f: 
#dat_postBurnRemoval 

  # make  post-LCMAP removal data 'long' 
  dat_LCMAPremoved_long  <- dat_LCMAPremoved %>% 
  st_drop_geometry() %>% 
  rename(ShrubCover = ShrbCvr, 
         TotalTreeCover = TtlTrCv, 
         TotalHerbaceousCover = TtlHrbC,
         BareGroundCover = BrGrndC,
         ForbCover = ForbCvr,
         C3GramCover = C3GrmCv,
         C4GramCover = C4GrmCv, 
         BroadLeavedTreeCover = AngTrCv,
         NeedleLeavedTreeCover = CnfTrCv) %>% 
    pivot_longer(cols = c(TotalTreeCover, TotalHerbaceousCover, ShrubCover, BareGroundCover, C3GramCover, C4GramCover, ForbCover, BroadLeavedTreeCover, NeedleLeavedTreeCover), 
               names_to = "CoverType", values_to = "CoverValue") %>% 
    mutate(CoverValue = replace(x = CoverValue, 
                       list = (CoverType %in% c("BroadLeavedTreeCover", "NeedleLeavedTreeCover", "C3GramCover", "C4GramCover", "ForbCover")) & Source == "RAP", 
                       values = NA))# for RAP data, remove values for BroadLeavedTreeCover, C3GramCover, C4GramCOver, ForbCover, and NeedleLeavedTreeCover (put zeros when the corresponding level-1 value is 0, but those are the only values)
  ggplot() +
    facet_wrap(~CoverType, scales = "free") + 
    geom_density(data = dat_postBurnRemoval, aes(CoverValue#, col= Source
                                                 ), col = "grey") + 
    geom_density(data = dat_LCMAPremoved_long , aes(CoverValue#, col = Source
                                        ),  col = "black") + 
    ggtitle("Density plots of cover by functional group across all datasets \n before and after removing observations due to unsuitable land cover type", 
            subtitle = "Grey = Data prior to excluding locations with unsuitable land cover type ;  \nBlack = Data after excluding locations with unsuitable land cover type")
```

## 4) "Coarsify" data from LANDFIRE 
Data from the LANDFIRE reference database is a conglomeration of plot data from various sources across the United States, and as a result, some geographic locations have a very high number of cover observations, while other locations are completely unrepresented in the dataset. This spatial heterogeneity is potentially problematic, particularly since it results in over-representation of some vegetation types. To attempt to ameliorate this issue, we've "coarsified" the cover data from th LANDFIRE reference database to result in a sampling density (in locations where data are available) that is similar to the  density of FIA plots. 
The median distance between FIA plots is 11,879 m, whereas the median distance between plots in the LANDFIRE reference database prior to this "coarsifying" step was 1,319 m. To make the spatial density of LANDFIRE plots coarser, we used a ~3km x ~3km across CONUS (based on the dayMet grid, which we scaled up by a factor of three). If more than one LANDFIRE plot that was measured in a given year was within the same 3km^2 grid cell, the cover observations from those plots were added together. We chose a 3km^2 grid because the final median distance between LANDFIRE plots after using the coarsifying process with that grid cell size (which was 11,637 m) was closest to the median distance between FIA plots. Below we show density figures of the LANDFIRE data both before and after this coarsifying process.

```{r}
# file is "data_beforeSpatialAveraging_sampledLANDFIRE.rds" 
dat_LFfilt <- readRDS("../Data_processed/CoverData/data_beforeSpatialAveraging_sampledLANDFIRE.rds") %>% 
  st_transform(crs(dat_1)) %>% 
  filter(Source == "LANDFIRE") #%>% 
  #select(Year:TotalHerbaceousCover, BareGroundCover, geometry)
# make the new LF filtered data into a 'long' format
dat_LFfilt_long <- dat_LFfilt %>% 
  st_drop_geometry() %>% 
  mutate(C3GramCover = C3GramCover_prop * TotalHerbaceousCover,
         C4GramCover = C4GramCover_prop * TotalHerbaceousCover, 
         ForbCover = ForbCover_prop * TotalHerbaceousCover, 
         BroadLeavedTreeCover = AngioTreeCover_prop * TotalTreeCover ,
         NeedleLeavedTreeCover = ConifTreeCover_prop * TotalTreeCover) %>% 
  pivot_longer(cols = c(TotalTreeCover, TotalHerbaceousCover, ShrubCover, C3GramCover, C4GramCover, ForbCover, BroadLeavedTreeCover, NeedleLeavedTreeCover), names_to = "CoverType", values_to = "CoverValue")

# make pre-LF filtered data long 
dat_LCMAPremoved_long_LF <-  dat_LCMAPremoved %>% 
  st_drop_geometry() %>% 
  filter(Source == "LANDFIRE") %>% 
  rename(ShrubCover = ShrbCvr, 
         TotalTreeCover = TtlTrCv, 
         TotalHerbaceousCover = TtlHrbC,
         BareGroundCover = BrGrndC,
         ForbCover = ForbCvr,
         C3GramCover = C3GrmCv,
         C4GramCover = C4GrmCv, 
         BroadLeavedTreeCover = AngTrCv,
         NeedleLeavedTreeCover = CnfTrCv) %>% 
    pivot_longer(cols = c(TotalTreeCover, TotalHerbaceousCover, ShrubCover, C3GramCover, C4GramCover, ForbCover, BroadLeavedTreeCover, NeedleLeavedTreeCover), 
               names_to = "CoverType", values_to = "CoverValue")

# make a density plot 
  ggplot() +
    facet_wrap(~CoverType, scales = "free") + 
    geom_density(data = dat_LCMAPremoved_long_LF, aes(CoverValue # before filtering
                                                 ), col = "grey") + 
    geom_density(data = dat_LFfilt_long , aes(CoverValue # after filtering
                                        ),  col = "black") + 
    ggtitle("Density plots of cover by functional group in the LANDFIRe data \nbefore and after `coarsifying` the LANDFIRE data to have only one \nobservation per 3 km x 3 km grid cell", 
            subtitle = "Grey = Data prior to `coarsifying`;  \nBlack = Data after `coarsifying`")
```

## 5) Average cover values together that occur in the same dayMet grid cell in the same year
To harmonize the four data sets we used for this project, we performed a "spatial averaging" process, where we average together cover measurements that were measured within the same dayMet grid cell in the same year. 
### Below are the vegetation cover data for each functional type and each data set, *prior* to the spatial averaging step (after removing burned area, locations with unsuitable land use, and "coarsified" LANDFIRE data). 

### RAP - data after filtering but before spatial averaging
```{r, fig.width = 12, fig.height = 30}
# Read in data 
dat_preAvg <- readRDS("../Data_processed/CoverData/data_beforeSpatialAveraging_sampledLANDFIRE.rds") %>% 
  st_transform(crs(dat_1)) 

# trim to be only later than 2000 or before 2024 (last year for which we have climate data was 2023) which is what we want to use for further analysis
dat_preAvg <- dat_preAvg %>% 
  filter(Year >= 2000 & Year < 2024) %>% 
  mutate(Lon = st_coordinates(.)[,1],
         Lat = st_coordinates(.)[,2]) %>% 
  mutate(
         ForbCover = ForbCover_prop * TotalHerbaceousCover,
         C3GramCover = C3GramCover_prop * TotalHerbaceousCover,
         C4GramCover = C4GramCover_prop * TotalHerbaceousCover, 
         BroadLeavedTreeCover = AngioTreeCover_prop * TotalTreeCover,
         NeedleLeavedTreeCover = ConifTreeCover_prop * TotalTreeCover)
 # st_drop_geometry()

###### rasterized version
# make a multi-layer raster w/ layers for each variable of interest
dat_preAvg_RAST <- 
  lapply(c("FIA", "LANDFIRE", "LDC","RAP"), 
         FUN = function(y) {
           tempDat <- dat_preAvg %>% 
             filter(Source == y )
           tempDat1 <- tempDat %>% 
             st_drop_geometry()
           if (y == c( "RAP")) {
           # get dat across all years
           tempOut2 <- lapply(c("ShrubCover", "TotalTreeCover", "TotalHerbaceousCover", "BareGroundCover", "ForbCover", "C3GramCover",  "C4GramCover",  "BroadLeavedTreeCover", "NeedleLeavedTreeCover"), 
                            FUN = function(z) {
                              temp <- terra::rasterize(x = tempDat1[,c("Lon", "Lat")], 
                                               y = test_rast_RAP,
                                               values = tempDat1[,z],
                                               fun = mean
                                               ) 
                      
                      names(temp) <-  paste0(z, "_all")
                      return(temp)
                    }
                    )
           } else {
           # get dat across all years
           tempOut2 <- lapply(c("ShrubCover", "TotalTreeCover", "TotalHerbaceousCover", "BareGroundCover", "ForbCover", "C3GramCover",  "C4GramCover",  "BroadLeavedTreeCover", "NeedleLeavedTreeCover"), 
                            FUN = function(z) {
                              temp <- terra::rasterize(x = tempDat1[,c("Lon", "Lat")], 
                                               y = test_rast,
                                               values = tempDat1[,z],
                                               fun = mean
                                               ) 
                      
                      names(temp) <-  paste0(z, "_all")
                      return(temp)
                    }
                    )
           }
           
           # for each data type, put data for each year in one raster in the list, and the data across all years in another raster
           outDat <- list( 
                          c(tempOut2[[1]], tempOut2[[2]], tempOut2[[3]], tempOut2[[4]], tempOut2[[5]], tempOut2[[6]], tempOut2[[7]], tempOut2[[8]], tempOut2[[9]]))
           names(outDat) <- c( "allYears")
           return(outDat)
         }
) %>% invisible()
names(dat_preAvg_RAST) <- c("FIA", "LANDFIRE", "LDC","RAP")

## make figures
## make figure for RAP 
testFigs_preAvg <- lapply(names(dat_preAvg_RAST$RAP), FUN = function(y) {
  # make a map for this year
           tempFig <- 
  ggplot() +
             facet_wrap(~lyr, ncol = 1) +
geom_spatraster(data = dat_preAvg_RAST$RAP[[y]][[1:4]], maxcell = 5e+10) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) +
  coord_sf(datum = st_crs(test_rastTemp)) +
             paletteer::scale_fill_paletteer_c("scico::roma", guide = guide_colorbar(title = "% cover"),
                                               limits = c(0,100), na.value = "white") + 
             theme(
                   axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) + 
             ggtitle("Map of RAP data on a dayMet x 13 grid - \nAfter filtering out burned area & developed and Ag. locations")
             
  # add an inset showing a density plot  
           # w/ data *prior* to any filtering for burned area, LCMAP, etc. 
           # get data
           dat_inset_pre <- #terra::values(dat_preAvg_RAST$RAP[[y]][[1:4]]) %>% 
             dat_1 %>% 
             st_drop_geometry() %>% 
             filter(Source == "RAP") %>% 
               #as_tibble() %>% 
             rename(ShrubCover_all = ShrubCover, 
                    TotalTreeCover_all = TotalTreeCover, 
                    TotalHerbaceousCover_all = TotalHerbaceousCover, 
                    BareGroundCover_all = BareGroundCover) %>% 
               pivot_longer(c(ShrubCover_all, TotalTreeCover_all, TotalHerbaceousCover_all, BareGroundCover_all), names_to = "lyr", values_to = "Cover") %>% 
             mutate(state = "pre-filtering") %>% 
             select(lyr, Cover, state)
 
           dat_inset_post <- 
             dat_preAvg %>% 
             st_drop_geometry %>% 
             filter(Source == "RAP") %>% 
             rename(ShrubCover_all = ShrubCover, 
                    TotalTreeCover_all = TotalTreeCover, 
                    TotalHerbaceousCover_all = TotalHerbaceousCover, 
                    BareGroundCover_all = BareGroundCover) %>% 
               pivot_longer(c(ShrubCover_all, TotalTreeCover_all, TotalHerbaceousCover_all, BareGroundCover_all), names_to = "lyr", values_to = "Cover")%>% 
             mutate(state = "post-filtering") %>% 
             select(lyr, Cover, state)
             
           dat_inset <- rbind(dat_inset_pre, dat_inset_post)
           
            # make insets
           inset_plots <- dat_inset %>%
             split(f = .$lyr) %>%
             purrr::map(~annotation_custom2(
      grob = ggplotGrob(get_inset_2(.)  + 
                          scale_color_discrete(palette = c("black", "grey")) +
                          guides(color = guide_legend(title = NULL)) + 
  theme(legend.position = c(.85,.5), legend.just = c(1,0), legend.text = element_text(size = 10))
                        #scale_y_continuous(limits=c(0,105), breaks = c(0, 50, 100))
                      ),
    data = data.frame(lyr=unique(.$lyr)),
    ymin = -2025000, ymax=-1205000, xmin=-2280000, xmax=-180000)
  )
         tempFig2 <- tempFig + inset_plots

           return(tempFig2)
  
  
}) %>% invisible()

RAPfigs_preAvg <- testFigs_preAvg[[1]]
RAPfigs_preAvg
```

### AIM - data after filtering but before spatial averaging
```{r, fig.width = 12, fig.height = 67.5, warning = FALSE}

## make figures
## make figure for RAP 
testFigs_preAvg <- lapply(names(dat_preAvg_RAST$LDC), FUN = function(y) {
  # make a map for this year
           tempFig <- 
  ggplot() +
             facet_wrap(~lyr, ncol = 1) +
geom_spatraster(data = dat_preAvg_RAST$LDC[[y]], maxcell = 5e+10) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) +
  coord_sf(datum = st_crs(test_rastTemp)) +
             paletteer::scale_fill_paletteer_c("scico::roma", guide = guide_colorbar(title = "% cover"),
                                               limits = c(0,100), na.value = "white") + 
             theme(
                   axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) + 
             ggtitle("Map of AIM data on a dayMet x 8 grid - \nAfter filtering out burned area & developed and Ag. locations")
             
  # add an inset showing a density plot  
           # w/ data *prior* to any filtering for burned area, LCMAP, etc. 
           # get data
           dat_inset_pre <- #terra::values(dat_preAvg_RAST$RAP[[y]][[1:4]]) %>% 
             dat_1 %>% 
             st_drop_geometry() %>% 
             filter(Source == "LDC") %>% 
               #as_tibble() %>% 
             rename(ShrubCover_all = ShrubCover, 
                    TotalTreeCover_all = TotalTreeCover, 
                    TotalHerbaceousCover_all = TotalHerbaceousCover, 
                    BareGroundCover_all = BareGroundCover,
                    ForbCover_all = ForbCover, 
                    C3GramCover_all = C3GramCover, 
                    C4GramCover_all = C4GramCover, 
                    BroadLeavedTreeCover_all = BroadLeavedTreeCover,
                    NeedleLeavedTreeCover_all = NeedleLeavedTreeCover) %>% 
               pivot_longer(c(ShrubCover_all, TotalTreeCover_all, TotalHerbaceousCover_all, BareGroundCover_all,
                              ForbCover_all,  C3GramCover_all,  C4GramCover_all,  BroadLeavedTreeCover_all, NeedleLeavedTreeCover_all), names_to = "lyr", values_to = "Cover") %>% 
             mutate(state = "pre-filtering") %>% 
             select(lyr, Cover, state)
 
                      
           dat_inset_post <- 
             dat_preAvg %>% 
             st_drop_geometry %>% 
             filter(Source == "LDC") %>% 
              rename(ShrubCover_all = ShrubCover, 
                    TotalTreeCover_all = TotalTreeCover, 
                    TotalHerbaceousCover_all = TotalHerbaceousCover, 
                    BareGroundCover_all = BareGroundCover,
                    ForbCover_all = ForbCover, 
                    C3GramCover_all = C3GramCover, 
                    C4GramCover_all = C4GramCover, 
                    BroadLeavedTreeCover_all = BroadLeavedTreeCover,
                    NeedleLeavedTreeCover_all = NeedleLeavedTreeCover) %>% 
               pivot_longer(c(ShrubCover_all, TotalTreeCover_all, TotalHerbaceousCover_all, BareGroundCover_all,
                              ForbCover_all,  C3GramCover_all,  C4GramCover_all,  BroadLeavedTreeCover_all, NeedleLeavedTreeCover_all), names_to = "lyr", values_to = "Cover")%>% 
             mutate(state = "post-filtering") %>% 
             select(lyr, Cover, state)
             
           dat_inset <- rbind(dat_inset_pre, dat_inset_post)
           
            # make insets
           inset_plots <- dat_inset %>%
             split(f = .$lyr) %>%
             purrr::map(~annotation_custom2(
      grob = ggplotGrob(get_inset_2(.)  + 
                          scale_color_discrete(palette = c("black", "grey")) +
                          guides(color = guide_legend(title = NULL)) + 
  theme(legend.position = c(.85,.5), legend.just = c(1,0), legend.text = element_text(size = 10))
                        #scale_y_continuous(limits=c(0,105), breaks = c(0, 50, 100))
                      ),
    data = data.frame(lyr=unique(.$lyr)),
    ymin = -2025000, ymax=-1205000, xmin=-2280000, xmax=-180000)
  )

         tempFig2 <- tempFig + inset_plots

           return(tempFig2)
  
  
}) %>% invisible()

AIMfigs_preAvg <- testFigs_preAvg[[1]]
AIMfigs_preAvg
```

### FIA - data after filtering but before spatial averaging
```{r, fig.width = 12, fig.height = 52.5, warning = FALSE}
## make figures
## make figure for FIA 
testFigs_preAvg <- lapply(names(dat_preAvg_RAST$FIA), FUN = function(y) {
  # make a map for this year
           tempFig <- 
  ggplot() +
             facet_wrap(~lyr, ncol = 1) +
geom_spatraster(data = dat_preAvg_RAST$FIA[[y]][[c(1:5,8:9)]], maxcell = 5e+10) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) +
  coord_sf(datum = st_crs(test_rastTemp)) +
             paletteer::scale_fill_paletteer_c("scico::roma", guide = guide_colorbar(title = "% cover"),
                                               limits = c(0,100), na.value = "white") + 
             theme(
                   axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) + 
             ggtitle("Map of FIA data on a dayMet x 8 grid - \nAfter filtering out burned area & developed and Ag. locations")
             
  # add an inset showing a density plot  
           # w/ data *prior* to any filtering for burned area, LCMAP, etc. 
           # get data
           dat_inset_pre <- #terra::values(dat_preAvg_RAST$RAP[[y]][[1:4]]) %>% 
             dat_1 %>% 
             st_drop_geometry() %>% 
             filter(Source == "FIA") %>% 
               #as_tibble() %>% 
                  rename(ShrubCover_all = ShrubCover, 
                    TotalTreeCover_all = TotalTreeCover, 
                    TotalHerbaceousCover_all = TotalHerbaceousCover, 
                    BareGroundCover_all = BareGroundCover,
                    ForbCover_all = ForbCover, 
                    C3GramCover_all = C3GramCover, 
                    C4GramCover_all = C4GramCover, 
                    BroadLeavedTreeCover_all = BroadLeavedTreeCover,
                    NeedleLeavedTreeCover_all = NeedleLeavedTreeCover) %>% 
               pivot_longer(c(ShrubCover_all, TotalTreeCover_all, TotalHerbaceousCover_all, BareGroundCover_all,
                              ForbCover_all, BroadLeavedTreeCover_all, NeedleLeavedTreeCover_all), 
                            names_to = "lyr", values_to = "Cover") %>% 
             mutate(state = "pre-filtering") %>% 
             select(lyr, Cover, state)
 
                      
           dat_inset_post <- 
             dat_preAvg %>% 
             st_drop_geometry %>% 
             filter(Source == "FIA") %>% 
                    rename(ShrubCover_all = ShrubCover, 
                    TotalTreeCover_all = TotalTreeCover, 
                    TotalHerbaceousCover_all = TotalHerbaceousCover, 
                    BareGroundCover_all = BareGroundCover,
                    ForbCover_all = ForbCover, 
                    C3GramCover_all = C3GramCover, 
                    C4GramCover_all = C4GramCover, 
                    BroadLeavedTreeCover_all = BroadLeavedTreeCover,
                    NeedleLeavedTreeCover_all = NeedleLeavedTreeCover) %>% 
               pivot_longer(c(ShrubCover_all, TotalTreeCover_all, TotalHerbaceousCover_all, BareGroundCover_all,
                              ForbCover_all, BroadLeavedTreeCover_all, NeedleLeavedTreeCover_all), 
                            names_to = "lyr", values_to = "Cover")%>% 
             mutate(state = "post-filtering") %>% 
             select(lyr, Cover, state)
             
           dat_inset <- rbind(dat_inset_pre, dat_inset_post)
           
            # make insets
           inset_plots <- dat_inset %>%
             split(f = .$lyr) %>%
             purrr::map(~annotation_custom2(
      grob = ggplotGrob(get_inset_2(.)  + 
                          scale_color_discrete(palette = c("black", "grey")) +
                          guides(color = guide_legend(title = NULL)) + 
  theme(legend.position = c(.85,.5), legend.just = c(1,0), legend.text = element_text(size = 10))
                        #scale_y_continuous(limits=c(0,105), breaks = c(0, 50, 100))
                      ),
    data = data.frame(lyr=unique(.$lyr)),
    ymin = -2025000, ymax=-1205000, xmin=-2280000, xmax=-180000)
  )

         tempFig2 <- tempFig + inset_plots

           return(tempFig2)
  
  
}) %>% invisible()

FIAfigs_preAvg <- testFigs_preAvg[[1]]
FIAfigs_preAvg
```

### LANDFIRE - data after filtering but before spatial averaging
```{r, fig.width = 12, fig.height = 60, warning = FALSE}

## make figures
## make figure for LANDFIRE 
testFigs_preAvg <- lapply(names(dat_preAvg_RAST$LANDFIRE), FUN = function(y) {
  # make a map for this year
           tempFig <- 
  ggplot() +
             facet_wrap(~lyr, ncol = 1) +
geom_spatraster(data = dat_preAvg_RAST$LANDFIRE[[y]][[c(1:3,5:9)]], maxcell = 5e+10) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) +
  coord_sf(datum = st_crs(test_rastTemp)) +
             paletteer::scale_fill_paletteer_c("scico::roma", guide = guide_colorbar(title = "% cover"),
                                               limits = c(0,100), na.value = "white") + 
             theme(
                   axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) + 
             ggtitle("Map of LANDFIRE data on a dayMet x 8 grid - \nAfter filtering out burned area & developed and Ag. locations")
             
  # add an inset showing a density plot  
           # w/ data *prior* to any filtering for burned area, LCMAP, etc. 
           # get data
           dat_inset_pre <- #terra::values(dat_preAvg_RAST$RAP[[y]][[1:4]]) %>% 
             dat_1 %>% 
             st_drop_geometry() %>% 
             filter(Source == "LANDFIRE") %>% 
               #as_tibble() %>% 
                   rename(ShrubCover_all = ShrubCover, 
                    TotalTreeCover_all = TotalTreeCover, 
                    TotalHerbaceousCover_all = TotalHerbaceousCover, 
                    BareGroundCover_all = BareGroundCover,
                    ForbCover_all = ForbCover, 
                    C3GramCover_all = C3GramCover, 
                    C4GramCover_all = C4GramCover, 
                    BroadLeavedTreeCover_all = BroadLeavedTreeCover,
                    NeedleLeavedTreeCover_all = NeedleLeavedTreeCover) %>% 
               pivot_longer(c(ShrubCover_all, TotalTreeCover_all, TotalHerbaceousCover_all, 
ForbCover_all, C3GramCover_all, C4GramCover_all, BroadLeavedTreeCover_all, NeedleLeavedTreeCover_all), 
                            names_to = "lyr", values_to = "Cover") %>% 
             mutate(state = "pre-filtering") %>% 
             select(lyr, Cover, state)
 
                      
           dat_inset_post <- 
             dat_preAvg %>% 
             st_drop_geometry %>% 
             filter(Source == "LANDFIRE") %>% 
                     rename(ShrubCover_all = ShrubCover, 
                    TotalTreeCover_all = TotalTreeCover, 
                    TotalHerbaceousCover_all = TotalHerbaceousCover, 
                    BareGroundCover_all = BareGroundCover,
                    ForbCover_all = ForbCover, 
                    C3GramCover_all = C3GramCover, 
                    C4GramCover_all = C4GramCover, 
                    BroadLeavedTreeCover_all = BroadLeavedTreeCover,
                    NeedleLeavedTreeCover_all = NeedleLeavedTreeCover) %>% 
               pivot_longer(c(ShrubCover_all, TotalTreeCover_all, TotalHerbaceousCover_all, 
ForbCover_all, C3GramCover_all, C4GramCover_all, BroadLeavedTreeCover_all, NeedleLeavedTreeCover_all), 
                            names_to = "lyr", values_to = "Cover")%>% 
             mutate(state = "post-filtering") %>% 
             select(lyr, Cover, state)
             
           dat_inset <- rbind(dat_inset_pre, dat_inset_post)
           
            # make insets
           inset_plots <- dat_inset %>%
             split(f = .$lyr) %>%
             purrr::map(~annotation_custom2(
      grob = ggplotGrob(get_inset_2(.)  + 
                          scale_color_discrete(palette = c("black", "grey")) +
                          guides(color = guide_legend(title = NULL)) + 
  theme(legend.position = c(.85,.5), legend.just = c(1,0), legend.text = element_text(size = 10))
                        #scale_y_continuous(limits=c(0,105), breaks = c(0, 50, 100))
                      ),
    data = data.frame(lyr=unique(.$lyr)),
    ymin = -2025000, ymax=-1205000, xmin=-2280000, xmax=-180000)
  )

         tempFig2 <- tempFig + inset_plots

           return(tempFig2)
  
  
}) %>% invisible()

LANDFIREfigs_preAvg <- testFigs_preAvg[[1]]
LANDFIREfigs_preAvg
```


### Below are tables showing sample size in the cover dataset, *prior* to the spatial averaging step
```{r}
dat_preAvg %>% 
  st_drop_geometry() %>% 
  select(Year, Source, ShrubCover, TotalTreeCover, TotalHerbaceousCover, BareGroundCover, ForbCover, C3GramCover, C4GramCover, BroadLeavedTreeCover, NeedleLeavedTreeCover) %>% 
  mutate(Source = replace(Source, Source == "LDC", "AIM"),
         Year = as.factor(Year)) %>% 
  group_by(Year, Source) %>% 
  summarize(Shrubs = sum(!is.na(ShrubCover)), 
            "Total\nTrees" = sum(!is.na(TotalTreeCover)),
            "Total\nHerbaceous" = sum(!is.na(TotalHerbaceousCover)),
            "Bare\nGround" = sum(!is.na(BareGroundCover)),
            Forbs = sum(!is.na(ForbCover)),
            "C3\nGrams" = sum(!is.na(C3GramCover)),
            "C4\nGrams" = sum(!is.na(C4GramCover)),
            "Broad\nLeaved\nTrees" = sum(!is.na(BroadLeavedTreeCover)),
            "Needle\nLeaved\nTrees" = sum(!is.na(NeedleLeavedTreeCover))) %>% 
  flextable::flextable() %>% 
  flextable::theme_vanilla() %>% 
  flextable::vline(j = c("Year", "Source"), border = flextable::fp_border_default()) %>% 
  flextable::set_caption(caption = flextable::as_paragraph(flextable::colorize("Sample Size for Each Functional Group For Each Year and Data Source", color = "black")), align_with_table = TRUE)   
```

```{r}
dat_preAvg %>% 
  st_drop_geometry() %>% 
  select(Year, Source, ShrubCover, TotalTreeCover, TotalHerbaceousCover, BareGroundCover, ForbCover, C3GramCover, C4GramCover, BroadLeavedTreeCover, NeedleLeavedTreeCover) %>% 
  mutate(Source = replace(Source, Source == "LDC", "AIM"),
         Year = as.factor(Year)) %>% 
  group_by(Year) %>% 
  summarize(Shrubs = sum(!is.na(ShrubCover)), 
            "Total\nTrees" = sum(!is.na(TotalTreeCover)),
            "Total\nHerbaceous" = sum(!is.na(TotalHerbaceousCover)),
            "Bare\nGround" = sum(!is.na(BareGroundCover)),
            Forbs = sum(!is.na(ForbCover)),
            "C3\nGrams" = sum(!is.na(C3GramCover)),
            "C4\nGrams" = sum(!is.na(C4GramCover)),
            "Broad\nLeaved\nTrees" = sum(!is.na(BroadLeavedTreeCover)),
            "Needle\nLeaved\nTrees" = sum(!is.na(NeedleLeavedTreeCover))) %>% 
  flextable::flextable() %>% 
  flextable::theme_vanilla() %>% 
  flextable::vline(j = c("Year"), border = flextable::fp_border_default()) %>% 
  flextable::set_caption(caption = flextable::as_paragraph(flextable::colorize("Sample Size for Each Functional Group in Each Year, Summarized Across All Data Sources", color = "black")), align_with_table = TRUE)   
```

### Below are maps of the vegetation cover data *after* the spatial averaging step. 
#### First we show the number of observations that we averaged together within each dayMet grid cell
```{r fig.width = 14, fig.height =40, warning = FALSE}
# read in data
dat_avgsBySource <- readRDS( "../Data_processed/CoverData/spatiallyAverageData_ByDataSource_intermediate_test5_sampledLANDFIRE.rds") %>% 
  filter(Year < 2024)
# make spatial and update crs to match the dayMet crs
dat_avgsBySource <- dat_avgsBySource %>% 
  dplyr::ungroup() %>% 
  st_as_sf(coords = c("x", "y"), crs = "GEOGCRS[\"NAD83\",\n    DATUM[\"North American Datum 1983\",\n        ELLIPSOID[\"GRS 1980\",6378137,298.257222101,\n            LENGTHUNIT[\"metre\",1]]],\n    PRIMEM[\"Greenwich\",0,\n        ANGLEUNIT[\"degree\",0.0174532925199433]],\n    CS[ellipsoidal,2],\n        AXIS[\"geodetic latitude (Lat)\",north,\n            ORDER[1],\n            ANGLEUNIT[\"degree\",0.0174532925199433]],\n        AXIS[\"geodetic longitude (Lon)\",east,\n            ORDER[2],\n            ANGLEUNIT[\"degree\",0.0174532925199433]],\n    ID[\"EPSG\",4269]]") %>% 
  st_transform(crs(dat_preAvg))

# rasterize for each data source and each cover type (across all years; values in the density plot are not averaged across years)
dat_avgsBySource <- dat_avgsBySource %>% 
  mutate(Lon = st_coordinates(.)[,1],
         Lat = st_coordinates(.)[,2]) %>% 
  mutate(
         ForbCover = ForbCover_prop * TotalHerbaceousCover,
         C3GramCover = C3GramCover_prop * TotalHerbaceousCover,
         C4GramCover = C4GramCover_prop * TotalHerbaceousCover, 
         BroadLeavedTreeCover = AngioTreeCover_prop * TotalTreeCover,
         NeedleLeavedTreeCover = ConifTreeCover_prop * TotalTreeCover,
         ForbCover_n = ForbCover_prop_n,
         C3GramCover_n = C3GramCover_prop_n,
         C4GramCover_n = C4GramCover_prop_n, 
         BroadLeavedTreeCover_n = AngioTreeCover_prop_n,
         NeedleLeavedTreeCover_n = ConifTreeCover_prop_n,
         ForbCover_sd = ForbCover_prop_sd,
         C3GramCover_sd = C3GramCover_prop_sd,
         C4GramCover_sd = C4GramCover_prop_sd, 
         BroadLeavedTreeCover_sd = AngioTreeCover_prop_sd,
         NeedleLeavedTreeCover_sd = ConifTreeCover_prop_sd,
         Source = replace(Source, Source == "LDC", "AIM")
         )
 # st_drop_geometry()
# for plotting purposes, change 0s to NAs
dat_avgsBySource[dat_avgsBySource$ShrubCover_n == 0 & !is.na(dat_avgsBySource$ShrubCover_n),]$ShrubCover_n <- NA
dat_avgsBySource[dat_avgsBySource$TotalTreeCover_n == 0 & !is.na(dat_avgsBySource$TotalTreeCover_n),]$TotalTreeCover_n <- NA
dat_avgsBySource[dat_avgsBySource$TotalHerbaceousCover_n == 0 & !is.na(dat_avgsBySource$TotalHerbaceousCover_n),]$TotalHerbaceousCover_n <- NA
dat_avgsBySource[dat_avgsBySource$TotalTreeCover_compObs_n == 0 & !is.na(dat_avgsBySource$TotalTreeCover_compObs_n),]$TotalTreeCover_compObs_n <- NA
dat_avgsBySource[dat_avgsBySource$BareGroundCover_n == 0 & !is.na(dat_avgsBySource$BareGroundCover_n),]$BareGroundCover_n <- NA
dat_avgsBySource[dat_avgsBySource$ForbCover_n == 0 & !is.na(dat_avgsBySource$ForbCover_n),]$ForbCover_n <- NA
dat_avgsBySource[dat_avgsBySource$C3GramCover_n == 0 & !is.na(dat_avgsBySource$C3GramCover_n),]$C3GramCover_n <- NA
dat_avgsBySource[dat_avgsBySource$C4GramCover_n == 0 & !is.na(dat_avgsBySource$C4GramCover_n),]$C4GramCover_n <- NA
dat_avgsBySource[dat_avgsBySource$BroadLeavedTreeCover_n == 0 & !is.na(dat_avgsBySource$BroadLeavedTreeCover_n),]$BroadLeavedTreeCover_n <- NA
dat_avgsBySource[dat_avgsBySource$NeedleLeavedTreeCover_n == 0 & !is.na(dat_avgsBySource$NeedleLeavedTreeCover_n),]$NeedleLeavedTreeCover_n <- NA

###### rasterized version
# make a multi-layer raster w/ layers for each variable of interest
dat_spatAvg_size_RAST <- 
  lapply(c("FIA", "LANDFIRE", "AIM","RAP"), 
         FUN = function(y) {
           tempDat <- dat_avgsBySource %>% 
             filter(Source == y)
           tempDat1 <- tempDat %>% 
             st_drop_geometry()
           if (y == c( "RAP")) {
           # get dat across all years
           tempOut2 <- lapply(c("ShrubCover", "TotalTreeCover", "TotalHerbaceousCover", "TotalTreeCover_compObs", "BareGroundCover", "ForbCover", "C3GramCover",  "C4GramCover",  "BroadLeavedTreeCover", "NeedleLeavedTreeCover",
                                "ShrubCover_n", "TotalTreeCover_n", "TotalHerbaceousCover_n", "TotalTreeCover_compObs_n", "BareGroundCover_n", "ForbCover_n", "C3GramCover_n",  "C4GramCover_n",  "BroadLeavedTreeCover_n", "NeedleLeavedTreeCover_n",
                                "ShrubCover_sd", "TotalTreeCover_sd", "TotalHerbaceousCover_sd", "TotalTreeCover_compObs_sd", "BareGroundCover_sd", "ForbCover_sd", "C3GramCover_sd",  "C4GramCover_sd",  "BroadLeavedTreeCover_sd", "NeedleLeavedTreeCover_sd"), 
                            FUN = function(z) {
                              temp <- terra::rasterize(x = tempDat1[,c("Lon", "Lat")], 
                                               y = test_rast_RAP,
                                               values = tempDat1[,z],
                                               fun = mean
                                               ) 
                      
                      names(temp) <-  paste0(z)
                      return(temp)
                    }
                    )
           } else {
           # get dat across all years
           tempOut2 <- lapply(c("ShrubCover", "TotalTreeCover", "TotalHerbaceousCover", "TotalTreeCover_compObs", "BareGroundCover", "ForbCover", "C3GramCover",  "C4GramCover",  "BroadLeavedTreeCover", "NeedleLeavedTreeCover",
                                "ShrubCover_n", "TotalTreeCover_n", "TotalHerbaceousCover_n", "TotalTreeCover_compObs_n", "BareGroundCover_n", "ForbCover_n", "C3GramCover_n",  "C4GramCover_n",  "BroadLeavedTreeCover_n", "NeedleLeavedTreeCover_n",
                                "ShrubCover_sd", "TotalTreeCover_sd", "TotalHerbaceousCover_sd", "TotalTreeCover_compObs_sd", "BareGroundCover_sd", "ForbCover_sd", "C3GramCover_sd",  "C4GramCover_sd",  "BroadLeavedTreeCover_sd", "NeedleLeavedTreeCover_sd"), 
                            FUN = function(z) {
                              temp <- terra::rasterize(x = tempDat1[,c("Lon", "Lat")], 
                                               y = test_rast,
                                               values = tempDat1[,z],
                                               fun = mean
                                               ) 
                      
                      names(temp) <-  paste0(z)
                      return(temp)
                    }
                    )
           }
           
           # for each data type, put data for each year in one raster in the list, and the data across all years in another raster
           outDat <- list( 
                          c(tempOut2[[1]], tempOut2[[2]], tempOut2[[3]], tempOut2[[4]], tempOut2[[5]], tempOut2[[6]], tempOut2[[7]], tempOut2[[8]], tempOut2[[9]],
                          tempOut2[[10]], tempOut2[[11]],tempOut2[[12]], tempOut2[[13]], tempOut2[[14]], tempOut2[[15]], tempOut2[[16]], tempOut2[[17]], tempOut2[[18]], tempOut2[[19]],
                          tempOut2[[20]], tempOut2[[21]],tempOut2[[22]], tempOut2[[23]], tempOut2[[24]], tempOut2[[25]], tempOut2[[26]], tempOut2[[27]], tempOut2[[28]], tempOut2[[29]], tempOut2[[30]]))
           names(outDat) <- c( "allYears")
           return(outDat)
         }
 ) %>% invisible()
 names(dat_spatAvg_size_RAST) <- c("FIA", "LANDFIRE", "AIM","RAP")

## make figures of number of observations per location
## make figure for RAP 
testFigs_spatAvg_n <- lapply(c("FIA", "LANDFIRE", "AIM","RAP"), FUN = function(y) {
  # make a map for the number of observations for each cover type for this dataset
           tempFig <- 
  ggplot() +
             facet_wrap(~lyr, ncol = 1) +
geom_spatraster(data = dat_spatAvg_size_RAST[[y]]$allYears[[11:20
                                                            ]], maxcell = 5e+10) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) +
  coord_sf(datum = st_crs(test_rastTemp)) +
             scale_fill_viridis_c(na.value = "white", "magma", direction = -1, limits = c(0,212)) +
             #scale_fill_distiller(type = "seq", palette = "PuBuGn", na.value = "grey70", direction = -1 ) +
             # paletteer::scale_fill_paletteer_c("wesanderson::Zissou1Continuous", guide = guide_colorbar(title = "N"),
             #                                   #limits = c(0,100)
             #                                   , na.value = "white") +
             theme(
                   axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) + 
             ggtitle(paste0("Number of observations from ",y," \nin a dayMet grid cell"),
                     subtitle = ifelse(y == "RAP",  yes = "Maps are shown on a dayMet x 13 grid, \nso values shown are the \nnumber of observations in a dayMet grid cell, \naveraged across all grid cells that occur \nin the 13km^2 mapping grid", no = "Maps are shown on a dayMet x 8 grid, \nso values shown are the \nnumber of observations in a dayMet grid cell, \naveraged across all grid cells that occur \nin the 8km^2 mapping grid"))
          
           
  # add an inset showing a density plot  
           # get data
              data_inset <- #terra::values(dat_1_RAST$RAP[[y]][[1:4]]) %>% 
             dat_avgsBySource %>% 
             st_drop_geometry() %>% 
             filter(Source == y) %>% 
               pivot_longer(c("ShrubCover_n"  ,"TotalTreeCover_n","TotalHerbaceousCover_n",   "TotalTreeCover_compObs_n", "BareGroundCover_n","ForbCover_n",  "C3GramCover_n" ,"C4GramCover_n", "BroadLeavedTreeCover_n" ,  "NeedleLeavedTreeCover_n" ), names_to = "lyr", values_to = "Cover")
            # make insets
           inset_plots <- data_inset %>%
             split(f = .$lyr) %>%
             purrr::map(~annotation_custom2(
      grob = ggplotGrob(get_inset(.)
                      ),
    data = data.frame(lyr=unique(.$lyr)),
    ymin = -2025000, ymax=-1205000, xmin=-2280000, xmax=-180000)
  )

         tempFig2 <- tempFig + inset_plots

           return(tempFig2)
  
}) %>% invisible()
names(testFigs_spatAvg_n) <- c("FIA", "LANDFIRE", "AIM","RAP")
ggarrange(testFigs_spatAvg_n$RAP, testFigs_spatAvg_n$AIM, testFigs_spatAvg_n$FIA, testFigs_spatAvg_n$LANDFIRE, ncol = 2)
```
 
#### Then, we show the spatially averaged cover values themselves
```{r, fig.width = 12, fig.height = 70, warning = FALSE, echo=FALSE}
# read in data
dat_avgsAll <- readRDS( "../Data_processed/CoverData/spatiallyAverageData_intermediate_test5_sampledLANDFIRE.rds") %>% 
  filter(Year < 2024)
# make spatial and update crs to match the dayMet crs
dat_avgsAll <- dat_avgsAll %>% 
  dplyr::ungroup() %>% 
  st_as_sf(coords = c("x", "y"), crs = "GEOGCRS[\"NAD83\",\n    DATUM[\"North American Datum 1983\",\n        ELLIPSOID[\"GRS 1980\",6378137,298.257222101,\n            LENGTHUNIT[\"metre\",1]]],\n    PRIMEM[\"Greenwich\",0,\n        ANGLEUNIT[\"degree\",0.0174532925199433]],\n    CS[ellipsoidal,2],\n        AXIS[\"geodetic latitude (Lat)\",north,\n            ORDER[1],\n            ANGLEUNIT[\"degree\",0.0174532925199433]],\n        AXIS[\"geodetic longitude (Lon)\",east,\n            ORDER[2],\n            ANGLEUNIT[\"degree\",0.0174532925199433]],\n    ID[\"EPSG\",4269]]") %>% 
  st_transform(crs(dat_preAvg))

# rasterize for each data source and each cover type (across all years; values in the density plot are not averaged across years)
dat_avgsAll <- dat_avgsAll %>% 
  mutate(Lon = st_coordinates(.)[,1],
         Lat = st_coordinates(.)[,2]) %>% 
  mutate(
         ForbCover = ForbCover_prop * TotalHerbaceousCover,
         C3GramCover = C3GramCover_prop * TotalHerbaceousCover,
         C4GramCover = C4GramCover_prop * TotalHerbaceousCover, 
         BroadLeavedTreeCover = AngioTreeCover_prop * TotalTreeCover,
         NeedleLeavedTreeCover = ConifTreeCover_prop * TotalTreeCover,
         ForbCover_n = ForbCover_prop_n,
         C3GramCover_n = C3GramCover_prop_n,
         C4GramCover_n = C4GramCover_prop_n, 
         BroadLeavedTreeCover_n = AngioTreeCover_prop_n,
         NeedleLeavedTreeCover_n = ConifTreeCover_prop_n,
         ForbCover_sd = ForbCover_prop_sd,
         C3GramCover_sd = C3GramCover_prop_sd,
         C4GramCover_sd = C4GramCover_prop_sd, 
         BroadLeavedTreeCover_sd = AngioTreeCover_prop_sd,
         NeedleLeavedTreeCover_sd = ConifTreeCover_prop_sd)
 # st_drop_geometry()
# for plotting purposes, change 0s to NAs
dat_avgsAll[dat_avgsAll$ShrubCover_n == 0 & !is.na(dat_avgsAll$ShrubCover_n),]$ShrubCover_n <- NA
dat_avgsAll[dat_avgsAll$TotalTreeCover_n == 0 & !is.na(dat_avgsAll$TotalTreeCover_n),]$TotalTreeCover_n <- NA
dat_avgsAll[dat_avgsAll$TotalHerbaceousCover_n == 0 & !is.na(dat_avgsAll$TotalHerbaceousCover_n),]$TotalHerbaceousCover_n <- NA
dat_avgsAll[dat_avgsAll$TotalTreeCover_compObs_n == 0 & !is.na(dat_avgsAll$TotalTreeCover_compObs_n),]$TotalTreeCover_compObs_n <- NA
dat_avgsAll[dat_avgsAll$BareGroundCover_n == 0 & !is.na(dat_avgsAll$BareGroundCover_n),]$BareGroundCover_n <- NA
dat_avgsAll[dat_avgsAll$ForbCover_n == 0 & !is.na(dat_avgsAll$ForbCover_n),]$ForbCover_n <- NA
dat_avgsAll[dat_avgsAll$C3GramCover_n == 0 & !is.na(dat_avgsAll$C3GramCover_n),]$C3GramCover_n <- NA
dat_avgsAll[dat_avgsAll$C4GramCover_n == 0 & !is.na(dat_avgsAll$C4GramCover_n),]$C4GramCover_n <- NA
dat_avgsAll[dat_avgsAll$BroadLeavedTreeCover_n == 0 & !is.na(dat_avgsAll$BroadLeavedTreeCover_n),]$BroadLeavedTreeCover_n <- NA
dat_avgsAll[dat_avgsAll$NeedleLeavedTreeCover_n == 0 & !is.na(dat_avgsAll$NeedleLeavedTreeCover_n),]$NeedleLeavedTreeCover_n <- NA

###### rasterized version
# make a multi-layer raster w/ layers for each variable of interest
dat_spatAvg_all_RAST <- 
  lapply(c("all"), 
         FUN = function(y) {
           tempDat <- dat_avgsAll 
           tempDat1 <- tempDat %>% 
             st_drop_geometry()
          
           # get dat across all years
           tempOut2 <- lapply(c("ShrubCover", "TotalTreeCover", "TotalHerbaceousCover", "TotalTreeCover_compObs", "BareGroundCover", "ForbCover", "C3GramCover",  "C4GramCover",  "BroadLeavedTreeCover", "NeedleLeavedTreeCover",
                                "ShrubCover_n", "TotalTreeCover_n", "TotalHerbaceousCover_n", "TotalTreeCover_compObs_n", "BareGroundCover_n", "ForbCover_n", "C3GramCover_n",  "C4GramCover_n",  "BroadLeavedTreeCover_n", "NeedleLeavedTreeCover_n",
                                "ShrubCover_sd", "TotalTreeCover_sd", "TotalHerbaceousCover_sd", "TotalTreeCover_compObs_sd", "BareGroundCover_sd", "ForbCover_sd", "C3GramCover_sd",  "C4GramCover_sd",  "BroadLeavedTreeCover_sd", "NeedleLeavedTreeCover_sd"), 
                            FUN = function(z) {
                              temp <- terra::rasterize(x = tempDat1[,c("Lon", "Lat")], 
                                               y = test_rast,
                                               values = tempDat1[,z],
                                               fun = mean
                                               ) 
                      
                      names(temp) <-  paste0(z)
                      return(temp)
                    }
                    )
           
           
           # for each data type, put data for each year in one raster in the list, and the data across all years in another raster
           outDat <- list( 
                          c(tempOut2[[1]], tempOut2[[2]], tempOut2[[3]], tempOut2[[4]], tempOut2[[5]], tempOut2[[6]], tempOut2[[7]], tempOut2[[8]], tempOut2[[9]],
                          tempOut2[[10]], tempOut2[[11]],tempOut2[[12]], tempOut2[[13]], tempOut2[[14]], tempOut2[[15]], tempOut2[[16]], tempOut2[[17]], tempOut2[[18]], tempOut2[[19]],
                          tempOut2[[20]], tempOut2[[21]],tempOut2[[22]], tempOut2[[23]], tempOut2[[24]], tempOut2[[25]], tempOut2[[26]], tempOut2[[27]], tempOut2[[28]], tempOut2[[29]], tempOut2[[30]]))
           names(outDat) <- c( "allYears")
           return(outDat)
         }
) %>% invisible()
names(dat_spatAvg_all_RAST) <- c("allData")

## make figures of number of observations per location
## make figure for RAP 
testFigs_spatAvg_n <- lapply(c("allData"), FUN = function(y) {
  # make a map for the number of observations for each cover type for this dataset
           tempFig <- 
  ggplot() +
             facet_wrap(~lyr, ncol = 1) +
geom_spatraster(data = dat_spatAvg_all_RAST[[y]]$allYears[[1:10
                                                            ]], maxcell = 5e+10) + 
  geom_sf(data=cropped_states_2 ,fill=NA ) +
 xlim(-2150000, 2510000) +
 ylim(-2005000, 1000000) +
  coord_sf(datum = st_crs(test_rastTemp)) +
             paletteer::scale_fill_paletteer_c("scico::roma", guide = guide_colorbar(title = "% Cover"),
                                               limits = c(0,100)
                                               , na.value = "white") +
             theme(
                   axis.text = element_blank(),
                   strip.text = element_text(size = 15),
                   title = element_text(size = 18),
                   axis.ticks = element_blank()) + 
             ggtitle(paste0("Spatially averaged cover data"),
                     subtitle = "Maps are shown on a dayMet x 8 grid, so values shown are the \n cover value in a in a dayMet grid cell, averaged across all grid cells that occur in the 8km^2 mapping grid")
          
           
  # add an inset showing a density plot  
           # get data
              data_inset <- #terra::values(dat_1_RAST$RAP[[y]][[1:4]]) %>% 
             dat_avgsAll %>% 
             st_drop_geometry() %>% 
               pivot_longer(c("ShrubCover"  ,"TotalTreeCover","TotalHerbaceousCover",   "TotalTreeCover_compObs", "BareGroundCover","ForbCover",  "C3GramCover" ,"C4GramCover", "BroadLeavedTreeCover" ,  "NeedleLeavedTreeCover" ), names_to = "lyr", values_to = "Cover")
            # make insets
           inset_plots <- data_inset %>%
             split(f = .$lyr) %>%
             purrr::map(~annotation_custom2(
      grob = ggplotGrob(get_inset(.)
                      ),
    data = data.frame(lyr=unique(.$lyr)),
    ymin = -2025000, ymax=-1205000, xmin=-2280000, xmax=-180000)
  )

         tempFig2 <- tempFig + inset_plots

           return(tempFig2)
  
})%>% invisible()
names(testFigs_spatAvg_n) <- c("allData")
testFigs_spatAvg_n
```
 

### Below are tables showing sample size in the cover dataset, *after* to the spatial averaging step
```{r}
dat_avgsAll %>% 
  st_drop_geometry() %>% 
  select(Year, ShrubCover, TotalTreeCover, TotalHerbaceousCover, BareGroundCover, ForbCover, C3GramCover, C4GramCover, BroadLeavedTreeCover, NeedleLeavedTreeCover) %>% 
  mutate(
         Year = as.factor(Year)) %>% 
  group_by(Year) %>% 
  summarize(Shrubs = sum(!is.na(ShrubCover)), 
            "Total\nTrees" = sum(!is.na(TotalTreeCover)),
            "Total\nHerbaceous" = sum(!is.na(TotalHerbaceousCover)),
            "Bare\nGround" = sum(!is.na(BareGroundCover)),
            Forbs = sum(!is.na(ForbCover)),
            "C3\nGrams" = sum(!is.na(C3GramCover)),
            "C4\nGrams" = sum(!is.na(C4GramCover)),
            "Broad\nLeaved\nTrees" = sum(!is.na(BroadLeavedTreeCover)),
            "Needle\nLeaved\nTrees" = sum(!is.na(NeedleLeavedTreeCover))) %>% 
  flextable::flextable() %>% 
  flextable::theme_vanilla() %>% 
  flextable::vline(j = c("Year"), border = flextable::fp_border_default()) %>% 
  flextable::set_caption(caption = flextable::as_paragraph(flextable::colorize("Sample Size for Each Functional Group For Each Year in the Spatially Averaged Dataset", color = "black")), align_with_table = TRUE)   
```


<!-- # After data has been spatially averaged (can no longer look at data by data source) -->
<!-- ## After initial spatial averaging -->
<!-- ```{r, echo = FALSE, , fig.width= 20, fig.height = 14} -->
<!-- # Read in data -->
<!-- dat_4 <- readRDS("../../../Data_processed/CoverData/spatiallyAverageData_intermediate_test5_sampledLANDFIRE.rds") -->

<!-- ## make figure for RAP -->
<!-- # get data just for RAP -->
<!-- dat_4 <- dat_4 %>% -->
<!--   pivot_longer(cols = c(ShrubCover, TotalTreeCover, TotalHerbaceousCover, BareGroundCover#, AngTrC_, CnfTrC_, C3GrmC_, C4GrmC_, FrbCvr_ -->
<!--                         ), -->
<!--                names_to = "coverType", -->
<!--                values_to = "coverValues" -->
<!--                 ) %>% -->
<!--   filter(!is.na(coverValues)) -->

<!-- ggarrange(ggplot(dat_4 %>% -->
<!--   filter(Year %in% c(2016, 2020, 2022))) + -->
<!--   facet_grid(rows = vars(coverType), cols = vars(Year)) + -->
<!--   stat_summary_2d(aes(x = x, y = y, z = coverValues), fun = mean, binwidth = .1) + -->
<!--      scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"), -->
<!--                           limits = c(0,100)) + -->
<!--     #geom_raster(aes(x = Lon, y = Lat, fill = coverValues)) + -->
<!--     #geom_point(aes(x = Lon, y = Lat, col = coverValues)) + -->
<!--   xlim(c(-130,-65)) + -->
<!--     ylim(c(25, 50)) + -->
<!--   ggtitle("Spatially Averaged Data - step 1"), -->

<!--   ggplot(dat_4) + -->
<!--   facet_grid(rows = vars(coverType)) + -->
<!--   stat_summary_2d(aes(x = x, y = y, z = coverValues), fun = mean, binwidth = .1) + -->
<!--      scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"), -->
<!--                           limits = c(0,100)) + -->
<!--     #geom_raster(aes(x = Lon, y = Lat, fill = coverValues)) + -->
<!--     #geom_point(aes(x = Lon, y = Lat, col = coverValues)) + -->
<!--   xlim(c(-130,-65)) + -->
<!--     ylim(c(25, 50)) + -->
<!--   ggtitle("Spatially Averaged Data - step 1 - n\all Years"), -->

<!--   ncol = 2, align = "h", -->
<!--   common.legend = TRUE, -->
<!--   widths = c(.7, .3) -->
<!-- ) -->
<!-- ``` -->
<!-- ## After adding climate/weather data -->
<!-- ```{r, echo = FALSE,  fig.width= 20, fig.height = 14} -->
<!-- # Read in data -->
<!-- dat_5_temp <- readRDS("../../../Data_processed/CoverData/DataForModels_spatiallyAveraged_sf_sampledLANDFIRE.rds") -->

<!-- ## make figures -->
<!-- dat_5 <- dat_5_temp %>% -->
<!--   pivot_longer(cols = c(ShrubCover, TotalTreeCover, TotalHerbaceousCover, BareGroundCover#, AngTrC_, CnfTrC_, C3GrmC_, C4GrmC_, FrbCvr_ -->
<!--                         ), -->
<!--                names_to = "coverType", -->
<!--                values_to = "coverValues" -->
<!--                 ) %>% -->
<!--   filter(!is.na(coverValues)) -->

<!-- ggarrange(ggplot(dat_5 %>% -->
<!--   filter(Year %in% c(2016, 2020, 2022))) + -->
<!--   facet_grid(rows = vars(coverType), cols = vars(Year)) + -->
<!--   stat_summary_2d(aes(x = x, y = y, z = coverValues), fun = mean, binwidth = .1) + -->
<!--      scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"), -->
<!--                           limits = c(0,100)) + -->
<!--     #geom_raster(aes(x = Lon, y = Lat, fill = coverValues)) + -->
<!--     #geom_point(aes(x = Lon, y = Lat, col = coverValues)) + -->
<!--   xlim(c(-130,-65)) + -->
<!--     ylim(c(25, 50)) + -->
<!--   ggtitle("Spatially Averaged Data - after adding climate/weather"), -->

<!--   ggplot(dat_5) + -->
<!--   facet_grid(rows = vars(coverType)) + -->
<!--   stat_summary_2d(aes(x = x, y = y, z = coverValues), fun = mean, binwidth = .1) + -->
<!--      scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"), -->
<!--                           limits = c(0,100)) + -->
<!--     #geom_raster(aes(x = Lon, y = Lat, fill = coverValues)) + -->
<!--     #geom_point(aes(x = Lon, y = Lat, col = coverValues)) + -->
<!--   xlim(c(-130,-65)) + -->
<!--     ylim(c(25, 50)) + -->
<!--   ggtitle("Spatially Averaged Data - \nafter adding clim/weath - n\all Years"), -->

<!--   ncol = 2, align = "h", -->
<!--   common.legend = TRUE, -->
<!--   widths = c(.7, .3) -->
<!-- ) -->
<!-- ``` -->
<!-- ## After adding ecoregion data -->
<!-- ```{r, echo = FALSE, , fig.width= 20, fig.height = 14} -->
<!-- # Read in data -->
<!-- dat_6 <- readRDS("../../../Data_processed/CoverData/DataForModels_withEcoregion_sampledLANDFIRE.rds") -->

<!-- ## make figures -->
<!-- dat_6 <- dat_6 %>% -->
<!--   pivot_longer(cols = c(ShrubCover, TotalTreeCover, TotalHerbaceousCover, BareGroundCover#, AngTrC_, CnfTrC_, C3GrmC_, C4GrmC_, FrbCvr_ -->
<!--                         ), -->
<!--                names_to = "coverType", -->
<!--                values_to = "coverValues" -->
<!--                 ) %>% -->
<!--   filter(!is.na(coverValues)) -->

<!-- ggarrange(ggplot(dat_6 %>% -->
<!--   filter(Year %in% c(2016, 2020, 2022))) + -->
<!--   facet_grid(rows = vars(coverType), cols = vars(Year)) + -->
<!--   stat_summary_2d(aes(x = x, y = y, z = coverValues), fun = mean, binwidth = .1) + -->
<!--      scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"), -->
<!--                           limits = c(0,100)) + -->
<!--     #geom_raster(aes(x = Lon, y = Lat, fill = coverValues)) + -->
<!--     #geom_point(aes(x = Lon, y = Lat, col = coverValues)) + -->
<!--   xlim(c(-130,-65)) + -->
<!--     ylim(c(25, 50)) + -->
<!--   ggtitle("Spatially Averaged Data - after adding ecoregion"), -->

<!--   ggplot(dat_6) + -->
<!--   facet_grid(rows = vars(coverType)) + -->
<!--   stat_summary_2d(aes(x = x, y = y, z = coverValues), fun = mean, binwidth = .1) + -->
<!--      scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"), -->
<!--                           limits = c(0,100)) + -->
<!--     #geom_raster(aes(x = Lon, y = Lat, fill = coverValues)) + -->
<!--     #geom_point(aes(x = Lon, y = Lat, col = coverValues)) + -->
<!--   xlim(c(-130,-65)) + -->
<!--     ylim(c(25, 50)) + -->
<!--   ggtitle("Spatially Averaged Data - \nafter adding ecoregion - n\all Years"), -->

<!--   ncol = 2, align = "h", -->
<!--   common.legend = TRUE, -->
<!--   widths = c(.7, .3) -->
<!-- ) -->
<!-- ``` -->


<!-- ## After adding soils data and removing RAP points that are in ag/developed areas -->
<!-- ```{r, echo = FALSE, , fig.width= 20, fig.height = 14} -->
<!-- # Read in data -->
<!-- dat_7_temp <- readRDS("../../../Data_processed/CoverData/DataForModels_spatiallyAveraged_withSoils_noSf_sampledLANDFIRE.rds") -->

<!-- ## make figures -->
<!-- dat_7 <- dat_7_temp %>% -->
<!--   pivot_longer(cols = c(ShrubCover, TotalTreeCover, TotalHerbaceousCover, BareGroundCover#, AngTrC_, CnfTrC_, C3GrmC_, C4GrmC_, FrbCvr_ -->
<!--                         ), -->
<!--                names_to = "coverType", -->
<!--                values_to = "coverValues" -->
<!--                 ) %>% -->
<!--   filter(!is.na(coverValues)) -->

<!-- ggarrange(ggplot(dat_7 %>% -->
<!--   filter(Year %in% c(2016, 2020, 2022))) + -->
<!--   facet_grid(rows = vars(coverType), cols = vars(Year)) + -->
<!--   stat_summary_2d(aes(x = x, y = y, z = coverValues), fun = mean, binwidth = .1) + -->
<!--      scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"), -->
<!--                           limits = c(0,100)) + -->
<!--     #geom_raster(aes(x = Lon, y = Lat, fill = coverValues)) + -->
<!--     #geom_point(aes(x = Lon, y = Lat, col = coverValues)) + -->
<!--   xlim(c(-130,-65)) + -->
<!--     ylim(c(25, 50)) + -->
<!--   ggtitle("Spatially Averaged Data - after adding soils"), -->

<!--   ggplot(dat_7) + -->
<!--   facet_grid(rows = vars(coverType)) + -->
<!--   stat_summary_2d(aes(x = x, y = y, z = coverValues), fun = mean, binwidth = .1) + -->
<!--      scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"), -->
<!--                           limits = c(0,100)) + -->
<!--     #geom_raster(aes(x = Lon, y = Lat, fill = coverValues)) + -->
<!--     #geom_point(aes(x = Lon, y = Lat, col = coverValues)) + -->
<!--   xlim(c(-130,-65)) + -->
<!--     ylim(c(25, 50)) + -->
<!--   ggtitle("Spatially Averaged Data - \nafter adding soils - n\all Years"), -->

<!--   ncol = 2, align = "h", -->
<!--   common.legend = TRUE, -->
<!--   widths = c(.7, .3) -->
<!-- ) -->
<!-- ``` -->

<!-- ### Making density plots of all of the cover type by data source (after removing burned area) -->
<!-- First, predict forest classification using the model we later fit to the data to determine forest or non-forest -->
<!-- ```{r} -->
<!-- #dat_5_temp has the climate data, so I think I should be able to add the climate data from there to the data in dat_3_temp?  -->
<!-- dat_7_clim <- dat_7_temp %>%  -->
<!--   st_as_sf(coords = c("x", "y"), crs = st_crs(dat_5_temp), remove = FALSE) %>%  -->
<!--   #cbind(dat_7_temp %>% select("x", "y")) %>%  -->
<!--           st_transform(crs = st_crs(dat_3_temp), remove = FALSE) %>%  -->
<!--   mutate(locationID = paste0(round(.data[["x"]], 5),"_", round(.data[["y"]], 5))) -->

<!-- dat_3_tempClim <- dat_3_temp %>%  -->
<!--   #slice(1) %>%  -->
<!--  sf::st_join( -->
<!--     dat_7_clim %>%  -->
<!--       select(locationID), -->
<!--       join = st_nearest_feature,#st_nearest_feature, -->
<!--     left = TRUE -->
<!--   )  -->
<!-- dat_3_tempClim <- dat_3_tempClim %>%  -->
<!--   st_drop_geometry() %>%  -->
<!--   left_join(dat_7_clim %>%  -->
<!--               st_drop_geometry() %>%  -->
<!--               select(Year, x:locationID),  -->
<!--             by = c("Year" = "Year", "locationID" = "locationID") -->
<!--              ) -->

<!-- dat_3_tempClim <-  -->
<!--   dat_3_tempClim %>%  -->
<!--   mutate(ForestPred =  as.numeric(1/ (1 + exp(-(-0.67478165 +   -->
<!--                                             1.71345065*((prcp_meanAnnTotal_CLIM - 483.34337211) / 321.52170566) +  -->
<!--                                             -0.01633932*((precip_Seasonality_meanAnnAvg_CLIM - 0.97338094) / 0.23365057) +  -->
<!--                                             -0.91285243*((PrecipTempCorr_meanAnnAvg_CLIM - 0.01681097) / 0.39882063) +   -->
<!--                                             0.53377439*((annWetDegDays_meanAnnAvg_CLIM - 1808.61043539) / 1083.25298131) +  -->
<!--                                             -0.60523936*((annWaterDeficit_95percentile_CLIM - 164.68611646) / 104.14892564) +   -->
<!--                                             0.20578879*((avgCoarsePerc_acrossDepth - 9.66120017) / 10.16883531) +  -->
<!--                                             -0.61568228*((totalAvailableWaterHoldingCapacity - 14.61190409) / 5.65765542) +  -->
<!--                                             -0.31818631*I(((tmean_meanAnnAvg_CLIM - 11.0291275) / 4.89048541)^2) +  -->
<!--                                             0.20104126*I(((precip_Seasonality_meanAnnAvg_CLIM - 0.97338094) / 0.23365057)^2) +  -->
<!--                                             -0.89055646*I(((PrecipTempCorr_meanAnnAvg_CLIM - 0.01681097) / 0.39882063)^2) +   -->
<!--                                             0.14226917*I(((isothermality_meanAnnAvg_CLIM - 37.76731082) / 5.2256935)^2) +  -->
<!--                                             -0.11810211*I(((surfaceClay_perc - 19.59171509) / 8.51944876)^2) +   -->
<!--                                             0.11278028*I(((totalAvailableWaterHoldingCapacity - 14.61190409) / 5.65765542)^2) +   -->
<!--                                             0.91656456*((annWetDegDays_meanAnnAvg_CLIM - 1808.61043539) / 1083.25298131)*((annWaterDeficit_95percentile_CLIM - 164.68611646) / 104.14892564) +  -->
<!--                                             -0.29650844*((precip_driestMonth_meanAnnAvg_CLIM - 3.18622575) / 5.36400552)*((isothermality_meanAnnAvg_CLIM - 37.76731082) / 5.2256935) +  -->
<!--                                             -0.02061048*((PrecipTempCorr_meanAnnAvg_CLIM - 0.01681097) / 0.39882063)*((isothermality_meanAnnAvg_CLIM - 37.76731082) / 5.2256935) + -->
<!--                                             -0.28504255*((isothermality_meanAnnAvg_CLIM - 37.76731082) / 5.2256935)*((tmean_meanAnnAvg_CLIM - 11.0291275) / 4.89048541) +  -->
<!--                                             0.50549093*((PrecipTempCorr_meanAnnAvg_CLIM - 0.01681097) / 0.39882063)*((precip_driestMonth_meanAnnAvg_CLIM - 3.18622575) / 5.36400552) + -->
<!--                                             -0.25285633*((PrecipTempCorr_meanAnnAvg_CLIM - 0.01681097) / 0.39882063)*((tmean_meanAnnAvg_CLIM - 11.0291275) / 4.89048541) +  -->
<!--                                             -0.16290284*((totalAvailableWaterHoldingCapacity - 14.61190409) / 5.65765542)*((surfaceClay_perc - 19.59171509) / 8.51944876) +   -->
<!--                                             0.09163249*((avgCoarsePerc_acrossDepth - 9.66120017) / 10.16883531)*((surfaceClay_perc - 19.59171509) / 8.51944876)))))) %>%  -->
<!--   drop_na(ForestPred) -->
<!-- # ggplot(dat_3_tempClim) +  -->
<!-- #   geom_point(aes(x = Lon, y = Lat, col = ForestPred)) -->

<!-- # convert the continuous probability predictions to binary using previously-identified threshold  -->
<!-- dat_3_tempClim <- dat_3_tempClim %>%  -->
<!--   mutate(pred_binary = ForestPred, -->
<!--     pred_binary = replace(pred_binary, pred_binary > 0.3819095, 1), -->
<!--          pred_binary = replace(pred_binary, pred_binary <= 0.3819095, 0)) -->

<!-- # interpolate the continuous predictions -->
<!-- dat_3_tempClim$preds_interp <- NA -->
<!-- dat_3_tempClim[!is.na(dat_3_tempClim$ForestPred), "preds_interp"] <- approxfun_pred(dat_3_tempClim[ !is.na(dat_3_tempClim$ForestPred), "ForestPred"]) -->

<!-- ``` -->

<!-- ```{r} -->
<!-- # data are in the dat_3_temp data.frame -->
<!-- # are there any locations where we have NL and BL tree but not total tree?  -->
<!-- #dat_3_temp[is.finite(dat_3_temp$AngioTreeCover_prop) & is.na(dat_3_temp$TotalTreeCover),] -->

<!--   ggplot() +  -->
<!--     facet_wrap(vars(Source, pred_binary), ncol = 2, scales = "free") +  -->
<!--     geom_density(data = dat_3_tempClim, aes(TotalTreeCover, col = Source)) +  -->
<!--     geom_density(data = dat_3_tempClim[is.finite(dat_3_tempClim$AngioTreeCover_prop),], aes(TotalTreeCover, col = Source),lty = 2) +  -->
<!--     ggtitle("Density plots of total tree cover data (prior to spatial averaging) \nsolid lines = all available total tree cover data \ndashed line = total tree cover data in locations where we also have \nNL and BL tree cover data") -->

<!-- kable(x = data.frame("FIA" = c(table(dat_3_temp[is.finite(dat_3_temp$TotalTreeCover), "Source"] %>% st_drop_geometry())["FIA"],  -->
<!--                                 table(dat_3_temp[is.finite(dat_3_temp$AngioTreeCover_prop) ,"Source"] %>% st_drop_geometry())["FIA"]),  -->
<!--                      "LANDFIRE" = c(table(dat_3_temp[is.finite(dat_3_temp$TotalTreeCover), "Source"] %>% st_drop_geometry() )["LANDFIRE"], -->
<!--                                 table(dat_3_temp[is.finite(dat_3_temp$AngioTreeCover_prop) ,"Source"] %>% st_drop_geometry())["LANDFIRE"]), -->
<!--                      "AIM" = c(table(dat_3_temp[is.finite(dat_3_temp$TotalTreeCover), "Source"] %>% st_drop_geometry() )["LDC"], -->
<!--                                 table(dat_3_temp[is.finite(dat_3_temp$AngioTreeCover_prop) ,"Source"] %>% st_drop_geometry())["LDC"]), -->
<!--                      "RAP" = c(table(dat_3_temp[is.finite(dat_3_temp$TotalTreeCover), "Source"] %>% st_drop_geometry() )["RAP"], NA), -->
<!--                      row.names = c("allTotalTreeCoverData", "completeTreeData") -->
<!--                      ),  -->
<!--       row.names = TRUE, -->
<!--       format = "html" -->
<!-- ) %>%  -->
<!-- kable_styling(bootstrap_options = c("striped", "hover", "condensed"))  -->

<!-- # maps  -->
<!-- ## locations where we have observations for total tree cover vs. locations where we have observations for all tree types  -->
<!-- dat_3_temp$totalTreeDataAvail <- NA -->
<!-- dat_3_temp$completeTreeDataAvail <- NA -->
<!-- dat_3_temp[!is.na(dat_3_temp$TotalTreeCover), "totalTreeDataAvail"] <- TRUE -->
<!-- dat_3_temp[!is.na(dat_3_temp$TotalTreeCover) & !is.na(dat_3_temp$ConifTreeCover_prop) & !is.na(dat_3_temp$AngioTreeCover_prop), "completeTreeDataAvail"] <- TRUE -->
<!-- ggarrange( -->
<!--   ggplot(data = dat_3_temp %>% filter(!is.na(totalTreeDataAvail))) +  -->
<!--     facet_grid(row = vars(Source)) +  -->
<!--     geom_point(aes(x = Lon, y = Lat,  col = Source), pch = 20) +  -->
<!--     ggtitle("Locations with total tree cover data"), -->
<!--  ggarrange( ggplot(data = dat_3_temp  %>% filter(!is.na(completeTreeDataAvail))) + -->
<!--     facet_grid(row = vars(Source)) +   -->
<!--     geom_point(aes(x = Lon, y = Lat,  col = Source), pch = 20) +  -->
<!--     ggtitle("Locations with total tree cover data + NL and BL tree cover data"), -->
<!--     ggplot() + geom_point(), ncol = 1, heights = c(.75, .25)),  -->
<!--   align = "h" -->
<!-- ) -->
<!-- ``` -->

<!-- <!-- #Initial data, across all years and all data types --> -->
<!-- <!-- ```{r fig.width= 9, fig.height = 6} --> -->
<!-- <!-- # Inital FIA data, all years by year --> -->
<!-- <!-- ggplot(dat_1temp %>% --> -->
<!-- <!--          filter(Source == "FIA")) + --> -->
<!-- <!--   facet_wrap(vars(Year)) + --> -->
<!-- <!--   stat_summary_2d(aes(x = Lon, y = Lat, z = TtlTrCv), fun = mean, binwidth = .1) + --> -->
<!-- <!--      scale_fill_viridis_c(option = "A", guide = guide_colorbar(title = "% cover"), --> -->
<!-- <!--                           limits = c(0,100)) + --> -->
<!-- <!--     #geom_raster(aes(x = Lon, y = Lat, fill = coverValues)) + --> -->
<!-- <!--     #geom_point(aes(x = Lon, y = Lat, col = coverValues)) + --> -->
<!-- <!--   xlim(c(-130,-65)) + --> -->
<!-- <!--     ylim(c(25, 50)) + --> -->
<!-- <!--   ggtitle("Initial Data: Total Tree Cover from FIA") --> -->
<!-- <!-- ``` --> -->